### 2022-03-09 appendix.Rmd
first commit

\renewcommand{\appendixname}{\bf{Supplementary Material}}
\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\tablename}{Supplementary Table}
\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
\renewcommand{\theequation}{S\arabic{table}} \setcounter{equation}{0}

#

## Supplementary Methods

### Calculating global geographic and linguistic proximity matrices

Geographic distance between two nations was calculated as the logged geodesic distance between country capital cities (data from the R package _maps_[@Brownrigg2018]) using the R package _geosphere_[@Hijmans2019]. The geographic proximity matrix was computed as one minus the log geographic distance matrix scaled between 0 and 1.

Linguistic proximity between two nations was calculated as the cultural proximity between all languages spoken within those nations, weighted by speaker percentages. We acquired cultural proximity data by combining the language family trees provided by Glottolog v3.0[@Glottolog] into one global language tree (undated and unresolved). We calculated cultural proximity $s$ between two languages $j$ and $k$ as the distance (in number of nodes traversed) of their most recent common ancestor $i$ to the root of the tree, through the formula:

$$
s_{jk} = \frac{n_{r}-n_{i}}{n_{r}}
$$

where $n_{r}$ is the maximum path length (in number of nodes traversed) leading to the pan-human root $r$, and $n_{i}$ is the maximum path length leading to node $i$. We then combined these proximities with speaker data from Ethnologue 21[@Ethnologue] and compared every language spoken within those nations by at least 1 permille of the population, weighted by speaker percentages, through the formula:

$$
w_{lm} = {\Sigma}{\Sigma}p_{lj}p_{mk}s_{jk}
$$

where $p_{lj}$ is the percentage of the population in nation $l$ speaking language $j$, $p_{mk}$ is the percentage of the population in nation $m$ speaking language $k$, and $s_{jk}$ is the proximity measure between languages $j$ and $k$[@Eff2008]. The resulting linguistic proximity matrix was also scaled between 0 and 1 before analysis.

### Bayesian models for reanalysis

We provide model formulae for our reanalyses of cross-national correlations, for a general bivariate case with standardised outcome $Y$ and predictor $X$ variables. In the naive regression model without controls for non-independence:

$$
\begin{aligned}
Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
\mu_{i} &= \alpha + \beta X_{i} \\
\alpha &\sim \text{Normal}(0, 0.4) \\
\beta &\sim \text{Normal}(0, 0.4) \\
\sigma &\sim \text{Exponential}(5)
\end{aligned}
$$
The priors in this model were arrived at by prior predictive checks, with wider priors making predictions beyond the scale of standardised outcome variables and narrower priors being too informative.

\newpage

To control for spatial non-independence, we add a Gaussian process to this model and feed it a scaled geographic distance matrix $D$ based on Euclidean distances between latitude and longitude coordinates. This distance matrix is computed internally by the R package _brms_[@Burkner2017]. The Gaussian process uses an exponentiated quadratic covariance kernel, the only covariance kernel currently supported by _brms_. The model formula is:

$$
\begin{aligned}
Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
\mu_{i} &= \alpha + \kappa_{\text{NATION}[i]} + \beta X_{i} \\
\begin{pmatrix}
\kappa_{1} \\ \kappa_{2} \\ \kappa_{3} \\ ... \\ \kappa_{n}
\end{pmatrix} &\sim \text{MVNormal}
\begin{pmatrix}
\begin{pmatrix}
0 \\ 0 \\ 0 \\ ... \\ 0
\end{pmatrix},\textbf{K}
\end{pmatrix}\\
\textbf{K}_{ij} &= sdgp^2 \text{exp} \big (-D_{ij}^2 / (2 lscale^2) \big )\\
\alpha &\sim \text{Normal}(0, 0.4) \\
\beta &\sim \text{Normal}(0, 0.4) \\
\sigma &\sim \text{Exponential}(5) \\
sdgp &\sim \text{Exponential}(5) \\
lscale &\sim \text{InverseGamma}(?,?)
\end{aligned}
$$
where $n$ is the number of nations, and $D^2_{ij}$ reflects the squared Euclidean distances between latitude and longitude coordinates for the $i$-th and $j$-th nations. Notice that the inverse gamma prior on $lscale$ is left undetermined. This is because the _brms_ package intelligently tunes the prior for this parameter based on the covariates of the Gaussian process (see https://betanalpha.github.io/assets/case_studies/gp_part3/part3.html).

\newpage

To control for cultural phylogenetic non-independence, we manually specify the covariance structure for nation random intercepts using a pre-computed linguistic proximity matrix $L$ (see previous section). The covariance between two nations is assumed to be linearly proportional to the linguistic proximity between those nations. This assumption is justified if we assume that cultural traits evolve via Brownian motion along a language phylogeny. The non-centered parameterisation of this model is:

$$
\begin{aligned}
Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
\mu_{i} &= \alpha + z_{\text{NATION}[i]}\sigma_{\alpha}L + \beta X_{i} \\
\alpha &\sim \text{Normal}(0, 0.4) \\
\beta &\sim \text{Normal}(0, 0.4) \\
z_{j} &\sim \text{Normal}(0, 1)\\
\sigma_{\alpha} &\sim \text{Exponential}(5) \\
\sigma &\sim \text{Exponential}(5)
\end{aligned}
$$
\newpage

Finally, we can control for spatial and cultural phylogenetic non-independence simultaneously by including both a Gaussian process over latitude and longitude coordinates _and_ nation random intercepts that covary according to linguistic proximity. The resulting model is as follows:

$$
\begin{aligned}
Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
\mu_{i} &= \alpha + \kappa_{\text{NATION}[i]} + z_{\text{NATION}[i]}\sigma_{\alpha}L + \beta X_{i} \\
\begin{pmatrix}
\kappa_{1} \\ \kappa_{2} \\ \kappa_{3} \\ ... \\ \kappa_{n}
\end{pmatrix} &\sim \text{MVNormal}
\begin{pmatrix}
\begin{pmatrix}
0 \\ 0 \\ 0 \\ ... \\ 0
\end{pmatrix},\textbf{K}
\end{pmatrix}\\
\textbf{K}_{ij} &= sdgp^2 \text{exp} \big (-D_{ij}^2 / (2 lscale^2) \big )\\
\alpha &\sim \text{Normal}(0, 0.4) \\
\beta &\sim \text{Normal}(0, 0.4) \\
z_{j} &\sim \text{Normal}(0, 1)\\
\sigma_{\alpha} &\sim \text{Exponential}(5) \\
\sigma &\sim \text{Exponential}(5) \\
sdgp &\sim \text{Exponential}(5) \\
lscale &\sim \text{InverseGamma}(?,?)
\end{aligned}
$$

\newpage

## Supplementary Results

### Geographic and cultural phylogenetic signal estimates for Human Development Index and Inglehart's value dimensions

```{r load signal}
tar_load(c(hdiGeo, hdiLin, tradGeo, tradLin, survGeo, survLin))
```

Controlling for shared cultural ancestry, the proportion of national-level variance explained by spatial proximity was `r printnum(hdiGeo$hypothesis$Estimate, digits = 2)` for the Human Development Index (95% credible interval [`r printnum(hdiGeo$hypothesis$CI.Lower, digits = 2)` `r printnum(hdiGeo$hypothesis$CI.Upper, digits = 2)`]; $\text{BF}_{\neq0}$ > 100), `r printnum(tradGeo$hypothesis$Estimate, digits = 2)` for traditional values (95% CI [`r printnum(tradGeo$hypothesis$CI.Lower, digits = 2)` `r printnum(tradGeo$hypothesis$CI.Upper, digits = 2)`]; $\text{BF}_{\neq0}$ > 100), and `r printnum(survGeo$hypothesis$Estimate, digits = 2)` for survival values (95% CI [`r printnum(survGeo$hypothesis$CI.Lower, digits = 2)` `r printnum(survGeo$hypothesis$CI.Upper, digits = 2)`]; $\text{BF}_{\neq0}$ = `r printnum(1 / survGeo$hypothesis$Evid.Ratio, digits = 2)`). Controlling for spatial proximity, the proportion of variance explained by shared cultural ancestry was `r printnum(hdiLin$hypothesis$Estimate, digits = 2)` for the Human Development Index (95% CI [`r printnum(hdiLin$hypothesis$CI.Lower, digits = 2)` `r printnum(hdiLin$hypothesis$CI.Upper, digits = 2)`]; $\text{BF}_{\neq0}$ > 100), `r printnum(tradLin$hypothesis$Estimate, digits = 2)` for traditional values (95% CI [`r printnum(tradLin$hypothesis$CI.Lower, digits = 2)` `r printnum(tradLin$hypothesis$CI.Upper, digits = 2)`]; $\text{BF}_{\neq0}$ > 100), and `r printnum(survLin$hypothesis$Estimate, digits = 2)` for survival values (95% CI [`r printnum(survLin$hypothesis$CI.Lower, digits = 2)` `r printnum(survLin$hypothesis$CI.Upper, digits = 2)`]; $\text{BF}_{\neq0}$ > 100).

\newpage

## Supplementary Figures

(ref:dagCaption) _A causal directed acyclic graph of spatial and cultural phylogenetic non-independence in cross-national studies._ We are interested in estimating the direct effect of national-level exposure $X$ on national-level outcome $Y$. But these variables are confounded by their common unobserved cause $U$. $U$ is a stand-in for shared environmental, ecological, and geographic causes (e.g. climate, biodiversity, physical topography) and cultural and institutional causes (e.g. cultural norms, political systems). In this causal model, we cannot condition on $U$ directly, as it is unobserved. However, geographic $G$ and linguistic $L$ relationships between societies influence $U$, since changing a nation's spatial distance to or shared cultural ancestry with other nations will change its environmental and cultural traits. We can thus use $G$ and $L$ to model the covariation between $X$ and $Y$ induced by $U$. Failing to do this and simply estimating the bivariate correlation between $X$ and $Y$ will produce spurious relationships and residuals that are spatially and culturally non-independent around the world (see Figure \@ref(fig:spatialExample)).

```{r dag, fig.width=6, fig.height=3.5, fig.cap="(ref:dagCaption)"}
tar_read(dag)
```

\newpage

(ref:plotSim1Caption) _Distribution of cross-national correlations from simulation study under moderate spatial autocorrelation._ In these simulations, the strength of spatial autocorrelation is set to 0.5 for both outcome and predictor variables. For frequentist regression models, point ranges represent correlation estimates and 95% confidence intervals. For Bayesian Gaussian process regression models, point ranges represent posterior means and 95% credible intervals. Correlations are ordered by effect size independently in each panel. Red point ranges indicate that the slope is "significant" (i.e. the 95% confidence / credible interval excludes zero). Black point ranges indicate that the slope is "not significant".

```{r plotSim1, fig.width=7.5, fig.height=5, fig.cap="(ref:plotSim1Caption)"}
tar_read(plotSim1)
```

\newpage

(ref:plotReplications2Caption) _Reanalysis models fitted to raw data, for economic development (a) and cultural values (b) studies._ Data points are labelled using ISO 3166-1 alpha-2 letter country codes. In all reanalyses, outcome and predictor variables are standardised, making regression slopes comparable to Pearson's correlation coefficients. Lines and shaded areas represent posterior median regression lines and 95% credible intervals. For models with covariates (Adamzyck and Pitt 2009; Gelfand et al. 2011), marginal effects are presented holding all covariates at zero or their reference categories.

```{r plotReplications2, fig.width=8, fig.height=8, fig.cap="(ref:plotReplications2Caption)"}
tar_read(plotReplications2)
```

\newpage

(ref:plotReplications3Caption) _Posterior estimates of Gaussian process functions mapping spatial autocorrelation onto geographic distance from our reanalyses of economic development (a) and cultural values (b) studies._ Estimates are from models additionally controlling for cultural phylogenetic non-independence. The y-axis represents the amount of spatial autocorrelation between data points with increasing distance between those points on the x-axis (logged distance in kilometres). Lines and shaded areas represent median posterior spatial autocorrelation functions and 50% and 95% credible intervals.

```{r plotReplications3, fig.width=8.2, fig.height=6, fig.cap="(ref:plotReplications3Caption)"}
tar_read(plotReplications3)
```

\newpage

(ref:plotReplications4Caption) _Posterior estimates of cultural phylogenetic signal from our reanalyses._ Estimates are from models additionally controlling for spatial non-independence. Cultural phylogenetic signal is operationalised as the proportion of national-level variance explained by linguistic proximity between nations. Ridges are full posterior distributions, and point ranges are posterior medians and 95% credible intervals.

```{r plotReplications4, fig.width=5, fig.height=5, fig.cap="(ref:plotReplications4Caption)"}
tar_read(plotReplications4)
```

\newpage

## Supplementary References

### 2022-03-09 appendix.tex
first commit

\clearpage
\makeatletter
\efloat@restorefloats
\makeatother

\begin{appendix}
\renewcommand{\appendixname}{\bf{Supplementary Material}}
\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\tablename}{Supplementary Table}
\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
\renewcommand{\theequation}{S\arabic{table}} \setcounter{equation}{0}

\hypertarget{section}{%
\section{}\label{section}}

\hypertarget{supplementary-methods}{%
\subsection{Supplementary Methods}\label{supplementary-methods}}

\hypertarget{calculating-global-geographic-and-linguistic-proximity-matrices}{%
\subsubsection{Calculating global geographic and linguistic proximity
matrices}\label{calculating-global-geographic-and-linguistic-proximity-matrices}}

Geographic distance between two nations was calculated as the logged
geodesic distance between country capital cities (data from the R
package \emph{maps}\textsuperscript{1}) using the R package
\emph{geosphere}\textsuperscript{2}. The geographic proximity matrix was
computed as one minus the log geographic distance matrix scaled between
0 and 1.

Linguistic proximity between two nations was calculated as the cultural
proximity between all languages spoken within those nations, weighted by
speaker percentages. We acquired cultural proximity data by combining
the language family trees provided by Glottolog v3.0\textsuperscript{3}
into one global language tree (undated and unresolved). We calculated
cultural proximity \(s\) between two languages \(j\) and \(k\) as the
distance (in number of nodes traversed) of their most recent common
ancestor \(i\) to the root of the tree, through the formula:

\[
s_{jk} = \frac{n_{r}-n_{i}}{n_{r}}
\]

where \(n_{r}\) is the maximum path length (in number of nodes
traversed) leading to the pan-human root \(r\), and \(n_{i}\) is the
maximum path length leading to node \(i\). We then combined these
proximities with speaker data from Ethnologue 21\textsuperscript{4} and
compared every language spoken within those nations by at least 1
permille of the population, weighted by speaker percentages, through the
formula:

\[
w_{lm} = {\Sigma}{\Sigma}p_{lj}p_{mk}s_{jk}
\]

where \(p_{lj}\) is the percentage of the population in nation \(l\)
speaking language \(j\), \(p_{mk}\) is the percentage of the population
in nation \(m\) speaking language \(k\), and \(s_{jk}\) is the proximity
measure between languages \(j\) and \(k\)\textsuperscript{5}. The
resulting linguistic proximity matrix was also scaled between 0 and 1
before analysis.

\hypertarget{bayesian-models-for-reanalysis}{%
\subsubsection{Bayesian models for
reanalysis}\label{bayesian-models-for-reanalysis}}

We provide model formulae for our reanalyses of cross-national
correlations, for a general bivariate case with standardised outcome
\(Y\) and predictor \(X\) variables. In the naive regression model
without controls for non-independence:

\[
\begin{aligned}
Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
\mu_{i} &= \alpha + \beta X_{i} \\
\alpha &\sim \text{Normal}(0, 0.4) \\
\beta &\sim \text{Normal}(0, 0.4) \\
\sigma &\sim \text{Exponential}(5)
\end{aligned}
\] The priors in this model were arrived at by prior predictive checks,
with wider priors making predictions beyond the scale of standardised
outcome variables and narrower priors being too informative.

\newpage

To control for spatial non-independence, we add a Gaussian process to
this model and feed it a scaled geographic distance matrix \(D\) based
on Euclidean distances between latitude and longitude coordinates. This
distance matrix is computed internally by the R package
\emph{brms}\textsuperscript{6}. The Gaussian process uses an
exponentiated quadratic covariance kernel, the only covariance kernel
currently supported by \emph{brms}. The model formula is:

\[
\begin{aligned}
Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
\mu_{i} &= \alpha + \kappa_{\text{NATION}[i]} + \beta X_{i} \\
\begin{pmatrix}
\kappa_{1} \\ \kappa_{2} \\ \kappa_{3} \\ ... \\ \kappa_{n}
\end{pmatrix} &\sim \text{MVNormal}
\begin{pmatrix}
\begin{pmatrix}
0 \\ 0 \\ 0 \\ ... \\ 0
\end{pmatrix},\textbf{K}
\end{pmatrix}\\
\textbf{K}_{ij} &= sdgp^2 \text{exp} \big (-D_{ij}^2 / (2 lscale^2) \big )\\
\alpha &\sim \text{Normal}(0, 0.4) \\
\beta &\sim \text{Normal}(0, 0.4) \\
\sigma &\sim \text{Exponential}(5) \\
sdgp &\sim \text{Exponential}(5) \\
lscale &\sim \text{InverseGamma}(?,?)
\end{aligned}
\] where \(n\) is the number of nations, and \(D^2_{ij}\) reflects the
squared Euclidean distances between latitude and longitude coordinates
for the \(i\)-th and \(j\)-th nations. Notice that the inverse gamma
prior on \(lscale\) is left undetermined. This is because the
\emph{brms} package intelligently tunes the prior for this parameter
based on the covariates of the Gaussian process (see
https://betanalpha.github.io/assets/case\_studies/gp\_part3/part3.html).

\newpage

To control for cultural phylogenetic non-independence, we manually
specify the covariance structure for nation random intercepts using a
pre-computed linguistic proximity matrix \(L\) (see previous section).
The covariance between two nations is assumed to be linearly
proportional to the linguistic proximity between those nations. This
assumption is justified if we assume that cultural traits evolve via
Brownian motion along a language phylogeny. The non-centered
parameterisation of this model is:

\[
\begin{aligned}
Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
\mu_{i} &= \alpha + z_{\text{NATION}[i]}\sigma_{\alpha}L + \beta X_{i} \\
\alpha &\sim \text{Normal}(0, 0.4) \\
\beta &\sim \text{Normal}(0, 0.4) \\
z_{j} &\sim \text{Normal}(0, 1)\\
\sigma_{\alpha} &\sim \text{Exponential}(5) \\
\sigma &\sim \text{Exponential}(5)
\end{aligned}
\] \newpage

Finally, we can control for spatial and cultural phylogenetic
non-independence simultaneously by including both a Gaussian process
over latitude and longitude coordinates \emph{and} nation random
intercepts that covary according to linguistic proximity. The resulting
model is as follows:

\[
\begin{aligned}
Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
\mu_{i} &= \alpha + \kappa_{\text{NATION}[i]} + z_{\text{NATION}[i]}\sigma_{\alpha}L + \beta X_{i} \\
\begin{pmatrix}
\kappa_{1} \\ \kappa_{2} \\ \kappa_{3} \\ ... \\ \kappa_{n}
\end{pmatrix} &\sim \text{MVNormal}
\begin{pmatrix}
\begin{pmatrix}
0 \\ 0 \\ 0 \\ ... \\ 0
\end{pmatrix},\textbf{K}
\end{pmatrix}\\
\textbf{K}_{ij} &= sdgp^2 \text{exp} \big (-D_{ij}^2 / (2 lscale^2) \big )\\
\alpha &\sim \text{Normal}(0, 0.4) \\
\beta &\sim \text{Normal}(0, 0.4) \\
z_{j} &\sim \text{Normal}(0, 1)\\
\sigma_{\alpha} &\sim \text{Exponential}(5) \\
\sigma &\sim \text{Exponential}(5) \\
sdgp &\sim \text{Exponential}(5) \\
lscale &\sim \text{InverseGamma}(?,?)
\end{aligned}
\]

\newpage

\hypertarget{supplementary-results}{%
\subsection{Supplementary Results}\label{supplementary-results}}

\hypertarget{geographic-and-cultural-phylogenetic-signal-estimates-for-human-development-index-and-ingleharts-value-dimensions}{%
\subsubsection{Geographic and cultural phylogenetic signal estimates for
Human Development Index and Inglehart's value
dimensions}\label{geographic-and-cultural-phylogenetic-signal-estimates-for-human-development-index-and-ingleharts-value-dimensions}}

Controlling for shared cultural ancestry, the proportion of
national-level variance explained by spatial proximity was 0.38 for the
Human Development Index (95\% credible interval {[}0.20 0.60{]};
\(\text{BF}_{\neq0}\) \textgreater{} 100), 0.43 for traditional values
(95\% CI {[}0.18 0.72{]}; \(\text{BF}_{\neq0}\) \textgreater{} 100), and
0.18 for survival values (95\% CI {[}0.00 0.44{]}; \(\text{BF}_{\neq0}\)
= 1.45). Controlling for spatial proximity, the proportion of variance
explained by shared cultural ancestry was 0.60 for the Human Development
Index (95\% CI {[}0.39 0.79{]}; \(\text{BF}_{\neq0}\) \textgreater{}
100), 0.55 for traditional values (95\% CI {[}0.26 0.79{]};
\(\text{BF}_{\neq0}\) \textgreater{} 100), and 0.79 for survival values
(95\% CI {[}0.53 0.97{]}; \(\text{BF}_{\neq0}\) \textgreater{} 100).

\newpage

\hypertarget{supplementary-figures}{%
\subsection{Supplementary Figures}\label{supplementary-figures}}

(ref:dagCaption) \emph{A causal directed acyclic graph of spatial and
cultural phylogenetic non-independence in cross-national studies.} We
are interested in estimating the direct effect of national-level
exposure \(X\) on national-level outcome \(Y\). But these variables are
confounded by their common unobserved cause \(U\). \(U\) is a stand-in
for shared environmental, ecological, and geographic causes
(e.g.~climate, biodiversity, physical topography) and cultural and
institutional causes (e.g.~cultural norms, political systems). In this
causal model, we cannot condition on \(U\) directly, as it is
unobserved. However, geographic \(G\) and linguistic \(L\) relationships
between societies influence \(U\), since changing a nation's spatial
distance to or shared cultural ancestry with other nations will change
its environmental and cultural traits. We can thus use \(G\) and \(L\)
to model the covariation between \(X\) and \(Y\) induced by \(U\).
Failing to do this and simply estimating the bivariate correlation
between \(X\) and \(Y\) will produce spurious relationships and
residuals that are spatially and culturally non-independent around the
world (see Figure @ref(fig:spatialExample)).

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/dag-1.pdf}
\caption{(\#fig:dag)(ref:dagCaption)}
\end{figure}

\newpage

(ref:plotSim1Caption) \emph{Distribution of cross-national correlations
from simulation study under moderate spatial autocorrelation.} In these
simulations, the strength of spatial autocorrelation is set to 0.5 for
both outcome and predictor variables. For frequentist regression models,
point ranges represent correlation estimates and 95\% confidence
intervals. For Bayesian Gaussian process regression models, point ranges
represent posterior means and 95\% credible intervals. Correlations are
ordered by effect size independently in each panel. Red point ranges
indicate that the slope is ``significant'' (i.e.~the 95\% confidence /
credible interval excludes zero). Black point ranges indicate that the
slope is ``not significant''.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotSim1-1.pdf}
\caption{(\#fig:plotSim1)(ref:plotSim1Caption)}
\end{figure}

\newpage

(ref:plotReplications2Caption) \emph{Reanalysis models fitted to raw
data, for economic development (a) and cultural values (b) studies.}
Data points are labelled using ISO 3166-1 alpha-2 letter country codes.
In all reanalyses, outcome and predictor variables are standardised,
making regression slopes comparable to Pearson's correlation
coefficients. Lines and shaded areas represent posterior median
regression lines and 95\% credible intervals. For models with covariates
(Adamzyck and Pitt 2009; Gelfand et al.~2011), marginal effects are
presented holding all covariates at zero or their reference categories.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotReplications2-1.pdf}
\caption{(\#fig:plotReplications2)(ref:plotReplications2Caption)}
\end{figure}

\newpage

(ref:plotReplications3Caption) \emph{Posterior estimates of Gaussian
process functions mapping spatial autocorrelation onto geographic
distance from our reanalyses of economic development (a) and cultural
values (b) studies.} Estimates are from models additionally controlling
for cultural phylogenetic non-independence. The y-axis represents the
amount of spatial autocorrelation between data points with increasing
distance between those points on the x-axis (logged distance in
kilometres). Lines and shaded areas represent median posterior spatial
autocorrelation functions and 50\% and 95\% credible intervals.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotReplications3-1.pdf}
\caption{(\#fig:plotReplications3)(ref:plotReplications3Caption)}
\end{figure}

\newpage

(ref:plotReplications4Caption) \emph{Posterior estimates of cultural
phylogenetic signal from our reanalyses.} Estimates are from models
additionally controlling for spatial non-independence. Cultural
phylogenetic signal is operationalised as the proportion of
national-level variance explained by linguistic proximity between
nations. Ridges are full posterior distributions, and point ranges are
posterior medians and 95\% credible intervals.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotReplications4-1.pdf}
\caption{(\#fig:plotReplications4)(ref:plotReplications4Caption)}
\end{figure}

\newpage

\hypertarget{supplementary-references}{%
\subsection*{Supplementary References}\label{supplementary-references}}
\addcontentsline{toc}{subsection}{Supplementary References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Brownrigg2018}{}%
1. Brownrigg, R. \emph{maps: Draw geographical maps}. (2018).

\leavevmode\hypertarget{ref-Hijmans2019}{}%
2. Hijmans, R. J. \emph{geosphere: Spherical trigonometry}. (2019).

\leavevmode\hypertarget{ref-Glottolog}{}%
3. Hammarström, H., Forkel, R., Haspelmath, M. \& Bank, S.
\emph{Glottolog 3.0}. (Max Planck Institute for the Science of Human
History, 2017).
doi:\href{https://doi.org/10.5281/zenodo.4061162}{10.5281/zenodo.4061162}.

\leavevmode\hypertarget{ref-Ethnologue}{}%
4. \emph{Ethnologue: Languages of the world}. (SIL International, 2018).

\leavevmode\hypertarget{ref-Eff2008}{}%
5. Eff, E. A. Weight matrices for cultural proximity: Deriving weights
from a language phylogeny. \emph{Structure and Dynamics} \textbf{3},
(2008).

\leavevmode\hypertarget{ref-Burkner2017}{}%
6. Bürkner, P.-C. brms: An R package for Bayesian multilevel models
using Stan. \emph{Journal of Statistical Software} \textbf{80}, 1--28
(2017).
\end{appendix}

### 2022-03-09 manuscript.Rmd
first commit

---
title             : "Cross-national studies have not sufficiently accounted for non-independence between nations"
shorttitle        : "Non-independence between nations"

author: 
  - name          : "Scott Claessens"
    affiliation   : "1"
    role: # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Data curation
      - Formal analysis
      - Investigation
      - Software
      - Visualization
      - Writing - original draft
      - Writing - review & editing
  - name          : "Quentin D. Atkinson"
    affiliation   : "1,2"
    corresponding : yes
    address       : "Floor 2, Building 302, 23 Symonds Street, Auckland, 1010, New Zealand"
    email         : "q.atkinson@auckland.ac.nz"
    role:
      - Conceptualization
      - Funding acquisition
      - Supervision
      - Writing - review & editing

affiliation:
  - id            : "1"
    institution   : "School of Psychology, University of Auckland, Auckland, New Zealand"
  - id            : "2"
    institution   : "Max Planck Institute for the Science of Human History, Jena, Germany"

authornote: |
 This working paper has not yet been peer-reviewed.

abstract: |
 Cross-national analyses can illuminate patternings of national-level outcomes around the world. But since nations are related via spatial proximity and shared cultural ancestry, cross-national analyses often violate statistical assumptions of non-independence, inflating false positive rates. Here, we show that past work in economics and psychology has not sufficiently accounted for non-independence between nations. In a review of the 100 highest-cited cross-national studies of economic development and cultural values, we find that controls for non-independence are rare and have not increased in frequency over the past 30 years. When studies do attempt to control for non-independence, our simulations suggest that the methods used continue to produce false positives. In replications of a subsample of twelve cross-national relationships, we show that half fail to replicate when controlling for spatial and/or cultural non-independence using global proximity matrices. We urge social scientists to sufficiently control for non-independence when reporting cross-national associations.

keywords          : "spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses"
wordcount         : "5549 words"
bibliography      : ["r-references.bib"]
floatsintext      : no
figurelist        : no
figsintext        : yes
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
numbersections    : false
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
header-includes: 
  - \raggedbottom
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
  - \AtBeginEnvironment{tablenotes}{\doublespacing}
  - \captionsetup[table]{font={stretch=1.5}}
  - \captionsetup[figure]{font={stretch=1.5}}
appendix          : "appendix.Rmd"
csl               : nature.csl
urlcolor          : blue
---

```{r setup, include=FALSE}
# !!! for use without tar_make
library(targets)
library(tarchetypes)
library(tidyverse)
source("R/functionsExamplePlot.R")
source("R/functionsSignal.R")
source("R/functionsReview.R")
source("R/functionsSimulation.R")
source("R/functionsReplications.R")
options(tidyverse.quiet = TRUE)
lapply(c("brms", "cowplot", "conleyreg", "countrycode", 
         "dagitty", "geosphere", "ggdag", "ggrepel", "ggtext", 
         "haven", "lmtest", "papaja", "psych", "readxl", 
         "rstan", "sjlabelled", "tidybayes", "tidyverse"), library, character.only = TRUE)
# !!!

knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

Nations are perhaps the single most important population unit structuring modern human life. The nation in which someone is born has a large effect on what they can expect out of life, including their income level[@Caselli2005], life expectancy[@Austin2012], mental health[@Rai2013], subjective well-being[@Diener2009], and educational attainment[@Kirkcaldy2004]. Nations are also among the most important determinants of human cultural variation, with greater cultural similarity within than outside national borders[@White2021].

Given the importance of nations for structuring human behaviour, there is justifiably huge interest in statistical analyses that attempt to predict variation in national-level outcomes of all kinds. At the time of writing, a search in Google Scholar for the term "cross-national study" returned 180,000 hits. The standard practice for cross-national analyses is to conduct bivariate correlations or multiple regressions with individual data points representing different nations. Such analyses widen the scope of social science beyond Western populations[@Henrich2010;@Pollet2014] and have been used to study, among other topics, the causes of variation in the economic wealth of nations[@Benhabib1994;@Comin2010;@LaPorta1997;@Sachs2001], global patterings of cultural norms and values[@Fincher2008;@Gelfand2011;@Hofstede2001;@Inglehart2000], and the universality and diversity of human behaviour and psychology around the world[@Awad2018;@Rhoads2021;@Schulz2019;@Thomson2018].

However, cross-national analyses are complicated by the fact that nations are not statistically independent data points. Unlike independent random samples from a population, nations are related to one another in a number of ways. First, nations that are closer to one another tend to be more similar than distant nations. This phenomenon is known as spatial non-independence[@Kissling2008], and it occurs because nations in close spatial proximity share characteristics due to local cultural diffusion of ideas[@Hewlett2002] and common climates and environments[@Kissling2008]. For example, the neighbouring African nations Zambia and Tanzania have similar levels of terrain ruggedness, which has been used to partially explain their similar levels of economic development[@Nunn2012]. This pattern conforms to Tobler's first law of geography: "everything is related to everything else, but near things are more related than distant things"[@Tobler1970] (p. 236).

Second, nations with shared cultural ancestry tend to be more similar than culturally unrelated nations. This is known as cultural phylogenetic non-independence[@Tylor1889], and occurs because related nations share cultural traits inherited via descent from a common ancestor. Shared cultural ancestry can result in a form of pseudoreplication, whereby multiple instances of the same trait across nations are merely duplicates of the ancestral original. For example, the related island nations New Zealand and Samoa express similar languages and customs due to cultural inheritance from a shared ancestor deep in the evolutionary history of Polynesian societies. More recently, shared ancestry is also crucial in explaining cultural similarities between colonial settlements and their colonisers, even those separated by large geographic distances (e.g. Spain and Argentina).

Spatial and cultural phylogenetic non-independence between nations make cross-national inference challenging. A fundamental assumption of regression analysis is that model residuals should be independently and identically distributed[@Jarque1987]. But without accounting for spatial or cultural non-independence between nations, model residuals can show structure that remains unaccounted for, violating this assumption (Figure \@ref(fig:spatialExample)). Treating nations as independent can thus inflate false positive rates[@Legendre1993], producing spurious "direct" relationships between variables that in fact only indirectly covary due to spatial or cultural connections (see Supplementary Figure \@ref(fig:dag) for an example causal model).

(ref:spatialExampleCaption) _An example of spatial non-independence in cross-national analysis._ National-level outcome and predictor variables for 25 nations are simulated with high levels of spatial non-independence. (top) Scatterplot visualising the data with fitted line from naive regression without accounting for spatial non-independence. Vertical line segments represent model residual errors, coloured by their sign and size. Countries are labelled with ISO 3166-1 alpha-2 letter codes. (bottom) Model residual errors plotted onto a world map. In this example, residuals clearly show spatial structure that remains unaccounted for.

```{r spatialExample, fig.width=6.5, fig.height=6, fig.cap="(ref:spatialExampleCaption)"}
tar_read(spatialExample)
```

Non-independence between data points is widely acknowledged in fields that routinely deal with spatially or culturally structured data. In ecology and sociology, studies with regional-level data use a variety of autoregressive models designed to account for spatial patternings[@Lichstein2002;@Loftin1983]. In anthropology, the Standard Cross-Cultural Sample of 186 cultures was compiled to minimise the confounding effects of non-independence in comparisons of traditional societies[@Murdock1969], though spatial and cultural dependencies are difficult to remove entirely[@Dow2008;@Eff2004]. Anthropologists also borrow phylogenetic comparative methods from evolutionary biology, such as phylogenetic least squares regression[@Symonds2014], when comparing traditional societies, treating culturally related societies in the same way as biologists treat genetically related species (e.g. [@Watts2018;@Atkinson2016]).

At the national level, recent reanalyses have revealed that several cross-national relationships reported in economics and psychology do not hold when controlling for non-independence between nations. One study replicated 25 analyses of "persistence" in economics, in which modern national outcomes are regressed against historical characteristics of those nations, and found that over half of the relationships were attenuated when controlling for spatial non-independence[@Kelly2020]. Another replication study found that many of the widely publicised relationships between national-level pathogen prevalence and political institutions and attitudes fail to hold when controlling for various kinds of non-independence[@Bromham2018]. These reanalyses, and others[@Bromham2021;@Currie2012], raise the question: how widespread a concern is non-independence in studies of national-level outcomes?

To address this question, we consider national-level variables of general interest across the social sciences: economic development and cultural values. These variables are frequently included as both outcomes and predictors in cross-national studies in economics and psychology[@Benhabib1994;@Comin2010;@Fincher2008;@Gelfand2011;@Hofstede2001;@Inglehart2000;@LaPorta1997;@Sachs2001]. First, we show that economic development and cultural values are spatially and culturally non-independent across nations, emphasising the need to control for non-independence. Second, we systematically review the 100 highest-cited cross-national studies of economic development and cultural values and estimate the proportion of cross-national analyses within these articles that account for non-independence between nations. Third, we run simulations to determine whether common methods of dealing with non-independence in the literature sufficiently reduce false positive rates. Fourth, we reanalyse twelve previous cross-national analyses of economic development and cultural values from our systematic review, incorporating global geographic and linguistic proximity matrices to correctly control for spatial and cultural non-independence.

# Results

## National-level economic development and cultural values are spatially and culturally non-independent

In order to motivate our research question, it is important to first show that economic development and cultural values are indeed spatially and culturally non-independent around the world. To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for the Human Development Index[@hdi] and two dimensions of cultural values from the World Values Survey, traditional vs. secular values and survival vs. self-expression values[@Inglehart2000]. For all three of these variables, we found that a substantial proportion of national-level variation was explained by spatial proximity and shared cultural ancestry between nations (Figure \@ref(fig:plotSignal)). Bayes Factors indicated strong evidence that these estimates of geographic and cultural phylogenetic signal differed from zero, aside from the geographic signal for survival values, for which the evidence was equivocal (see Supplementary Results). These findings emphasise the need to account for spatial and cultural non-independence in cross-national analyses of economic development and cultural values.

(ref:plotSignalCaption) _Posterior estimates of geographic and cultural phylogenetic signal for the Human Development Index and Inglehart's traditional and survival values._ Geographic and cultural phylogenetic signal are operationalised as the proportion of national-level variance explained by geographic and linguistic proximity matrices. Grey ridges are full posterior distributions, points are posterior median values, and lines are 50% and 95% credible intervals.

```{r plotSignal, fig.width=7, fig.height=3.5, fig.cap="(ref:plotSignalCaption)"}
tar_read(plotSignal)
```

## Previous cross-national analyses have not sufficiently accounted for non-independence

```{r load review}
tar_load(review)
```

Since economic development and cultural values show evidence of geographic and cultural phylogenetic signal, have cross-national analyses sufficiently accounted for this non-independence? To assess this, we systematically searched the published literature for articles that combined the search terms "economic development" or "values" with the search terms "cross-national", "cross-cultural", or "cross-country". We removed articles that did not report original research, were not relevant to economic development or cultural values, or did not report at least one cross-national analysis. We then retained the 100 articles (50 for economic development, 50 for cultural values) with the highest annual rate of citations. For each of these highly-cited articles, we exhaustively recorded every cross-national analysis reported in the main text (n = `r printnum(as.numeric(nrow(review)), digits = 0)`), identifying in each case whether or not the analysis attempted to control for spatial, cultural, or any other form of non-independence between nations (see Methods for detailed search criteria and coding decisions).

```{r calculate prop 1}
prop <-
  review %>%
  group_by(Review, Citation) %>%
  summarise(Yes = ifelse(sum(ControlNI == "Yes") > 0, 1, 0)) %>%
  group_by(Review) %>%
  summarise(prop = mean(Yes)) %>%
  # bootstrap cis
  mutate(lower = map(prop, function(x) getBootCIProp(x, 50, 1000)[1]), 
         upper = map(prop, function(x) getBootCIProp(x, 50, 1000)[2])) %>%
  unnest(c(lower, upper))
```

The results of our systematic review show that most published articles and cross-national analyses make no attempt to account for statistical non-independence. Figure \@ref(fig:plotReview)a plots the proportion of articles that contain at least one cross-national analysis accounting for non-independence. We find that `r printnum(prop$prop[1]*100, digits = 0)`% of economic development articles contain at least one attempt to control for non-independence (95% bootstrap confidence interval [`r printnum(prop$lower[1])` `r printnum(prop$upper[1])`]), while this proportion decreases to only `r printnum(prop$prop[2]*100, digits = 0)`% for cultural values articles (95% bCI [`r printnum(prop$lower[2])` `r printnum(prop$upper[2])`]). Both kinds of article are most likely to use regional fixed effects (e.g. continent fixed effects) to account for non-independence, but some articles also include controls for spatial distance (e.g. latitude) and shared cultural history (e.g. legal origin fixed effects).

(ref:plotReviewCaption) _Results from systematic review of 100 highly-cited cross-national studies of economic development (red) and cultural values (blue)._ (a) Proportion of articles containing at least one analysis accounting for non-independence, overall and split by common methods of controlling for non-independence. Point ranges represent proportions and 95% bootstrap confidence intervals. (b) Adjusted proportion of individual analyses accounting for non-independence, overall and split by method. Point ranges are posterior medians and 95% credible intervals from Bayesian multilevel logistic regressions. (c) Histogram of publication years for studies of economic development and cultural values. (d) Estimated trend over time for the probability of controlling for non-independence. Lines and shaded areas are posterior median regression lines and 50% credible intervals from a Bayesian multilevel spline model. Region FEs = region fixed effects.

```{r plotReview, fig.width=8, fig.height=5.5, fig.cap="(ref:plotReviewCaption)"}
tar_read(plotReview)
```

```{r calculate prop 2}
tar_load(postRM1)
propEcon <- inv_logit_scaled(postRM1[,,"b_ReviewEconomicdevelopment"])
propVal  <- inv_logit_scaled(postRM1[,,"b_ReviewValues"])
```

Focusing on the full sample of `r printnum(as.numeric(nrow(review)), digits = 0)` analyses, we find that the proportion of individual cross-national analyses accounting for non-independence is even lower (Figure \@ref(fig:plotReview)b). Across `r printnum(as.numeric(sum(review$Review == "Economic development")), digits = 0)` cross-national analyses from studies of economic development, only `r printnum(median(propEcon)*100, digits = 0)`% are estimated to control for non-independence (95% credible interval [`r printnum(quantile(propEcon, 0.025))` `r printnum(quantile(propEcon, 0.975))`]). Similarly, across `r printnum(as.numeric(sum(review$Review == "Values")), digits = 0)` cross-national analyses from studies of cultural values, only `r printnum(median(propVal)*100, digits = 0)`% are estimated to control for non-independence (95% CI [`r printnum(quantile(propVal, 0.025))` `r printnum(quantile(propVal, 0.975))`]).

Since our systematic review goes back as far as 1993 (Figure \@ref(fig:plotReview)c), it is possible that our estimates are being biased by earlier studies, and that controls for non-independence have increased over time with methodological advancements and greater awareness of the issue. To test this possibility, we fitted a time trend to the full sample of analyses with a Bayesian multilevel spline model. However, we found that, for both studies of economic development and cultural values, the estimated probability of controlling for non-independence has remained low since 1993 (Figure \@ref(fig:plotReview)d).

## Common methods of controlling for non-independence produce inflated false positive rates

Our systematic review revealed that most cross-national analyses in the literature do not control for spatial or cultural phylogenetic non-independence. When they do, they tend to include controls like regional fixed effects, latitude, and legal origin fixed effects. Do these methods sufficiently account for statistical non-independence?

To compare the efficacy of different methods in the literature, we conducted a simulation study. We simulated national-level datasets (n = 178 nations) with varying degrees of spatial autocorrelation (i.e. non-independence) for outcome and predictor variables, but with no direct causal relationship between the variables. We then fitted naive regressions without controls to these datasets, as well as regression models with the following controls: latitude, longitude, continent fixed effects, and legal origin fixed effects. Despite not being identified in our systematic review, we also included in our simulation study Conley standard errors, a widely used standard error correction that purportedly accounts for spatial non-independence[@Conley1999;@Conley2010]. Finally, we included a multilevel model of spatial non-independence that has recently been made widely available in statistical software packages[@Burkner2017;@rstan]: Bayesian Gaussian process regression[@McElreath2020;@Neal1998]. This method allows nation random intercepts to covary according to a global geographic proximity matrix. Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero (i.e. falsely infering a relationship when none is present).

```{r calculate false positive rate}
fp <-
  tar_read(olsModel1) %>%
  group_by(lambda, rho) %>%
  summarise(n = sum(sig) / n())
fpConley <-
  tar_read(conleyModel3) %>%
  group_by(lambda, rho) %>%
  summarise(n = sum(sig) / n())
fpGP <-
  tar_read(brmsModel) %>%
  group_by(lambda, rho) %>%
  summarise(n = sum(sig) / n())
```

Figure \@ref(fig:plotSim2) plots the estimated false positive rates from our simulation study, split by different methods and different degrees of spatial autocorrelation (see Supplementary Figure \@ref(fig:plotSim1) for full distribution of effect sizes under moderate autocorrelation). With at least moderate degrees of spatial autocorrelation for both outcome and predictor variables, naive regression models (Figure \@ref(fig:plotSim2) top left) produce false positive rates above chance levels. With strong spatial autocorrelation for both outcomes and predictors, false positive rates reach as high as `r printnum(fp$n[fp$lambda == 0.9 & fp$rho == 0.9]*100, digits = 0)`%.

(ref:plotSim2Caption) _False positive rates for different methods of controlling for non-independence in our simulation study._ For simulated outcome and predictor variables, we systematically varied the strength of spatial autocorrelation, from weak (0.1) to moderate (0.5) to strong (0.9). We simulated 100 datasets per parameter combination and fitted different models to each dataset. False positive rates were operationalised as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero. Point ranges represent proportions and 95% bootstrap confidence intervals, and dashed lines indicate the 5% false positive rate that is expected due to chance. SEs = standard errors.

```{r plotSim2, fig.width=10, fig.height=7, fig.cap="(ref:plotSim2Caption)"}
tar_read(plotSim2)
```

Common methods in the literature do not reduce this high false positive rate. With strong spatial autocorrelation for both outcome and predictor variables, false positive rates remain above 70% when controlling for latitude, longitude, continent fixed effects, and legal origin fixed effects. Applying Conley standard errors with cutoffs of 100 km or 1,000 km also does not reduce false positive rates below 70% under strong spatial autocorrelation. With a cutoff of 10,000 km, the false positive rate under strong spatial autocorrelation reduces to `r printnum(fpConley$n[fpConley$lambda == 0.9 & fpConley$rho == 0.9]*100, digits = 0)`%, though this correction adds bias at weaker levels of spatial autocorrelation, increasing false positive rates above chance levels.

By contrast, Bayesian Gaussian process regression adequately accounts for non-independence across a range of simulation parameters. At moderate levels of spatial autocorrelation on both outcomes and predictors, Gaussian process regression eliminates false positives, such that the false positive rate is no different from chance. With strong spatial autocorrelation on both outcomes and predictors, the false positive rate is reduced to `r printnum(fpGP$n[fpGP$lambda == 0.9 & fpGP$rho == 0.9]*100, digits = 0)`%, substantially less than any other method.

## Key findings in the literature are not robust to reanalysis with more rigorous methods

Our systematic review and simulation study have showed that controls for non-independence are rare in cross-national studies of economic development and cultural values, and when studies do attempt to control for non-independence, the methods used continue to produce false positives. This raises the worrying possibility that the cross-national literature in economics and psychology is populated with spurious relationships.

To diagnose how widespread this issue of spurious cross-national relationships might be, we reanalysed a subset of twelve previous cross-national analyses from our systematic review, sufficiently controlling for spatial and cultural phylogenetic non-independence using global geographic and linguistic proximity matrices. We subsampled six analyses from our economic development review[@Beck2003;@Beck2005;@Bockstette2002;@Easterly2003;@Easterly2007;@Skidmore2002] and six from our cultural values review[@Adamczyk2009;@Alesina2013;@Fincher2008;@Gelfand2011;@Inglehart2000;@Knack1997]. Our choice of analyses was constrained by data availability and whether we were able to initially replicate the original finding. We pre-registered our subsample of analyses before running any control models (https://osf.io/uywx8/). We controlled for non-independence by including (1) a Gaussian process allowing nation random intercepts to covary according to a geographic proximity matrix, and/or (2) nation random intercepts that covaried according to a linguistic proximity matrix (see Supplementary Methods for full models).

Figure \@ref(fig:plotReplications1) visualises the results of our reanalysis. Overall, cross-national correlation effect sizes tended to reduce when controlling for statistical non-independence between nations, sometimes by as much as half of the original effect size. Based on 95% credible intervals, six out of twelve cross-national relationships are attenuated when controlling for non-independence. For the economic development analyses, four out of six cross-national relationships are attenuated when controlling for spatial non-independence. For the cultural values analyses, two out of six are attenuated when controlling for cultural phylogenetic non-independence. Supplementary Figure \@ref(fig:plotReplications2) shows these cross-national correlations plotted against the raw data.

(ref:plotReplications1Caption) _Posterior correlations from our reanalysis of twelve previous cross-national analyses._ For each previous cross-national relationship, we plot the posterior slopes from a naive regression (red), a regression controlling for spatial non-independence (green), a regression controlling for cultural phylogenetic non-independence (blue), and a regression controlling for both spatial and cultural phylogenetic non-independence simultaneously (purple). All outcome and predictor variables are standardised. Most analyses are simple bivariate cross-national correlations, but Gelfand et al. (2011) is a partial correlation controlling for log gross national income and Adamczyk and Pitt (2009) is a multilevel model including several covariates. Point ranges represent posterior medians and 95% credible intervals. GDP = gross domestic product. FLFP = female labour force participation.

```{r plotReplications1, fig.width=8, fig.height=6.5, fig.cap="(ref:plotReplications1Caption)"}
tar_read(plotReplications1)
```

```{r load autocor}
tar_load(c(inglehartSA, skidmoreSA, alesinaSignal, gelfandSignal))
```

To understand why some cross-national correlations were attenuated by controls for non-independence while others were robust, we further explored our fitted models for evidence of spatial and cultural autocorrelation. For each outcome variable, our Gaussian process models provided varying estimates of how quickly spatial autocorrelation declined with distance (Supplementary Figure \@ref(fig:plotReplications3)). For example, in Skidmore and Toya[@Skidmore2002] gross domestic product growth was only moderately spatially autocorrelated at 1000 km distance (posterior median spatial autocorrelation at 1000 km = `r printnum(median(skidmoreSA))`, 95% CI [`r printnum(quantile(skidmoreSA, 0.025))` `r printnum(quantile(skidmoreSA, 0.975))`]), whereas in Inglehart and Baker[@Inglehart2000] traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1000 km = `r printnum(median(inglehartSA))`, 95% CI [`r printnum(quantile(inglehartSA, 0.025))` `r printnum(quantile(inglehartSA, 0.975))`]). We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \@ref(fig:plotReplications4)), with some outcome variables expressing low signal (e.g. Gelfand et al.[@Gelfand2011]; posterior median = `r printnum(median(gelfandSignal))`, 95% CI [`r printnum(quantile(gelfandSignal, 0.025))` `r printnum(quantile(gelfandSignal, 0.975))`]) and others expressing high signal (e.g. Alesina et al.[@Alesina2013]; posterior median = `r printnum(median(alesinaSignal))`, 95% CI [`r printnum(quantile(alesinaSignal, 0.025))` `r printnum(quantile(alesinaSignal, 0.975))`]). Across all analyses, we found that stronger estimates of spatial autocorrelation or cultural phylogenetic signal resulted in a more pronounced reduction in the effect size when controlling for non-independence between nations (Figure \@ref(fig:plotReplications5)).

(ref:plotReplications5Caption) _The estimated degree of spatial and cultural phylogenetic non-independence predicts reductions in effect size in our reanalysis._ (a) Higher estimated degrees of spatial autocorrelation at 1000 km distance predict more pronounced reductions in effect sizes when controlling for non-independence. (b) Higher estimated levels of cultural phylogenetic signal predict more pronounced reductions in effect sizes when controlling for non-independence. In both panels, the y-axis represents the ratio of the effect size when controlling for spatial and cultural non-independence to the original effect size (from naive regression model), and the x-axis represents posterior median model estimates. Regression lines are plotted with 95% confidence intervals.

```{r plotReplications5, fig.width=8, fig.height=4, fig.cap="(ref:plotReplications5Caption)"}
tar_read(plotReplications5)
```

# Discussion

In a systematic literature review and simulation, we found that cross-national studies in economics and psychology rarely account for non-independence between nations, and, when they do, the methods they use continue to produce false positives. In a reanalysis of twelve cross-national correlations, we further showed that neglecting to account for non-independence has resulted in spurious relationships in the published literature, with half of the correlations failing to replicate when controlling for spatial or cultural non-independence with more rigorous methods. These findings suggest that cross-national analyses in economics and psychology should be interpreted with caution until non-independence is sufficiently accounted for.

Non-independence is a concern because economic development and cultural values are highly geographically and culturally structured around the world, as revealed by our initial modelling. These results build on previous work. One previous study suggested that geographic proximity is more important than deep cultural ancestry in explaining the distribution of human development across Eurasian nations[@Sookias2018]. By contrast, with a global sample of nations, we found strong cultural phylogenetic signal, as well as geographic signal, for the Human Development Index. Another previous study found that similarities in the cultural values of nations are predicted by linguistic, but not geographic, distances between those nations[@White2021]. We find this same result for survival vs. self-expression values, but for traditional vs. secular values we find that both linguistic and geographic proximity are important independent predictors of global variation. These findings emphasise the need to account for both spatial and cultural phylogenetic non-independence in cross-national studies of economic development and cultural values.

Yet our systematic literature review and simulation study revealed that the most commonly used controls for non-independence do not sufficiently deal with the issue. In our simulations, controlling for either latitude or longitude did not reduce false positive rates, calling into question controls like distance to the equator in regression models. High false positive rates persisted with Conley standard errors, which have previously been critiqued for being overly sensitive to arbitrary distance cutoffs[@Kelly2020]. The simulation also confirmed the verbal claim that fixed effects for spatial or cultural groupings (e.g. continent or legal origin fixed effects) are insufficient because non-independence still remains within groupings[@Bromham2018]. This logic applies to analyses that control for non-independence by including language family fixed effects (e.g. [@Currie2009]) or separately analysing different regions (e.g. [@Smith2017]). The only method in our simulation that sufficiently reduced the false positive rate was Bayesian Gaussian process regression, though we did not include other possible controls for non-independence, such as conditional autoregressive models[@Lichstein2002] or generalised additive models[@Hastie2017].

Ours is not the first review to show that studies are misapplying statistical methods in ways that inflate false positive rates. For example, other literature reviews have shown that studies in the social sciences tend to use small sample sizes[@Smaldino2016], treat ordinal data as metric[@Liddell2018], incorrectly handle missing values[@Nicholson2017], and ignore best practices in meta-analyses[@Tipton2019]. Why do cross-national studies also rarely account for non-independence? At the institutional level, one possibility is that such practices are incentivised _because_ they inflate false positive rates, increasing the probability of getting published and promoted[@Smaldino2016]. At the individual level, another possibility is that researchers outside of anthropology and ecology are simply not aware of the problem, or believe that the problem does not apply at the national level. Even if researchers appreciate the problem, they might not know of suitable controls or perceive the methods to be too complex.

These institutional- and individual-level barriers can be combatted. First, cross-national replication studies like ours and others[@Bromham2018;@Bromham2021;@Currie2012;@Kelly2020], combined with the methodological reviews included in Registered Reports[@Chambers2021], might change incentive structures and encourage researchers to analyse the world's nations with more rigorous methods. Second, since the issue of non-independence is fundamentally an issue of causal inference (Supplementary Figure \@ref(fig:dag)), more explicit descriptions of causal models could promote controls for non-independence in cross-national research. In our review, economists studying economic development dealt with national-level non-independence more than psychologists studying cultural values, likely because economics studies tend to be lengthy statistical exercises that systematically incorporate or exclude numerous variables in an attempt to infer causation. Third, the recent widespread accessibility of open source statistical software, such as the programming language Stan[@rstan] and the R package _brms_[@Burkner2017], should promote the use of more rigorous methods to control for non-independence. Using _brms_, for example, Bayesian Gaussian process regression is straightforward to conduct, requiring only longitude and latitude values for nations.

Until such changes are implemented and sufficient controls for non-independence are the norm, existing cross-national correlations should be interpreted with caution. In our reanalyses, we found that four out of six cross-national analyses of economic development were attenuated when controlling for spatial non-independence. Three of these were tests of "persistence" hypotheses, studying the effects of historical and environmental conditions --- settler mortality[@Beck2003], wheat-sugar suitability[@Easterly2007], and natural disaster frequency[@Skidmore2002] --- on modern developmental outcomes. A recent reanalysis has also called into question various studies of this ilk[@Kelly2020]. We also found that two out of six cross-national analyses of cultural values were attenuated when controlling for cultural phylogenetic non-independence. It is interesting that the economic development analyses were attenuated with spatial controls, while the cultural values analyses were attenuated by cultural phylogenetic controls. This highlights that certain kinds of non-independence will be more important depending on the particular national-level outcome and predictor variables of interest.

We do not wish to dissuade researchers from conducting cross-national studies. On the contrary, such work promises to deepen our understanding of human populations beyond Western, educated, industrialised, rich, and democratic societies[@Henrich2010], providing the representative samples necessary to test evolutionary and socio-ecological theories of human behaviour[@Pollet2014;@Oishi2013]. But in order to minimise spurious relationships in global datasets, we urge researchers to control for spatial and cultural phylogenetic non-independence when reporting cross-national correlations. Nations are not independent, and our statistical models should reflect this.

# Methods

## Geographic and cultural phylogenetic signal

```{r}
tar_load(c(hdi, wvs))
```

To estimate the degree of spatial and cultural phylogenetic non-independence in economic development and cultural values, we calculated geographic and linguistic signal for global measures of development and values. Our measure of economic development was the Human Development Index[@hdi]. We retrieved a longitudinal dataset capturing human development for `r printnum(length(unique(hdi$Country)))` nations since 1990 (n = `r printnum(as.numeric(nrow(hdi)), digits = 0)`; https://hdr.undp.org/en/content/download-data). Our measures of cultural values were traditional vs. secular values and survival vs. self-expression values from the World Values Survey[@Inglehart2000]. We downloaded the full Integrated Values Survey, which included all waves from the World Values Survey and the European Values Survey, and computed the two dimensions of cultural values following procedures from previous research[@Inglehart2000]. This longitudinal dataset captures values and attitudes for `r printnum(length(unique(wvs$S009)))` nations since 1981 (n = `r printnum(as.numeric(nrow(wvs)), digits = 0)`; https://www.worldvaluessurvey.org/WVSEVStrend.jsp).

To calculate geographic and linguistic signal, we created two proximity matrices for 269 of the world's nations: a geographic proximity matrix and a linguistic proximity matrix. Geographic proximity was converted from logged geodesic distances between nation capital cities. Linguistic proximity was calculated as the cultural proximity between all languages spoken within nations, weighted by speaker percentages (see Supplementary Methods). We included these matrices in Bayesian multilevel models, allowing nation random intercepts to covary according to both geographic and linguistic proximity simultaneously. These models were fitted with the R package _brms_[@Burkner2017] and converged normally ($\hat{R}$ < 1.1). Estimates of geographic and linguistic signal were computed as the proportion of national-level variance in these models explained by geographic and linguistic proximity matrices.

## Systematic literature review

We exported two searches from Web of Science (https://www.webofknowledge.com/) on 27^th^ September 2021, restricting our searches to articles published between 1900 and 2018. The first search was for the terms "economic development" AND ("cross-national" OR "cross-cultural" OR "cross-country"), which returned 965 articles. The second search was for the terms "values" AND ("cross-national" OR "cross-cultural" OR "cross-country"), which returned 6806 articles. Once exported, we ordered the articles by descending number of citations per year since initial publication, using citation counts reported by Web of Science.

We then systematically coded each article, in order, for inclusion in our review. Articles were only included if: (1) they were judged to be relevant to economic development or cultural values; (2) they were an original empirical research article; and (3) they contained at least one analysis with national-level outcome or predictor variables. We stopped when we had included 50 articles for the economic development review and 50 articles for the cultural values review.

Within each included article, we exhaustively coded every individual cross-national analysis reported in the main text. We coded mainly correlation or regression analyses, and explicitly excluded meta-analyses, factor analyses, measurement invariance analyses, multidimensional scaling analyses, hierarchical clustering analyses, multiverse analyses, and scale development / validation analyses. We also excluded analyses that compared only two, three, four, five, or six nations. For each included analysis, we recorded the year, outcome variable, all predictor variables, test statistic, p-value, number of nations, number of data points, model type, if the data were available, and whether and how the analysis attempted to control for non-independence.

We coded common attempts to control for non-independence between nations. These included: (1) any higher-level control variables for spatial regional groupings (e.g. continent fixed effects); (2) any geographic distance control variables (e.g. distance between capital cities, distance from equator, latitude); (3) any control variables capturing shared cultural history (e.g. former colony, legal origin fixed effects, linguistic history, cultural influence); and (4) any other control variables, tests, or approaches that were deemed as attempts to control for non-independence (e.g. eigenvector filtering[@Tiefelsdorf2007], controls for trade-weightings between countries, cross-sectional dependence tests[@DeHoyos2006], separate analyses for subsets of countries). These were coded by the first author.

Once we had compiled our review database, we calculated the proportion of articles attempting to control for non-independence at least once. We also calculated the proportion of articles employing the different types of control listed above at least once: regional fixed effects, distance, shared cultural history, or other. For these proportions, we calculated 95% boostrap confidence intervals with 1000 bootstrap iterations.

For individual analyses, we dealt with the nested nature of the data (analyses nested within articles) by fitting Bayesian multilevel logistic regression models with review type (economic development vs. cultural values) as the sole fixed effect and random intercepts for articles. We fitted these models separately for overall attempts to control for non-independence and split by method type. We report the adjusted proportions with 95% credible intervals. To test for a trend over time, we also fitted a Bayesian multilevel logistic regression with a multigroup spline for year of publication and random intercepts for articles. Bayesian models were fitted with the _brms_ R package[@Burkner2017]. Our priors were informed by prior predictive checks, and all models converged normally ($\hat{R}$ < 1.1).

## Simulations

To simulate data, we used our global geographic distance matrix (see Supplementary Methods) as a weight matrix $W$, scaling the matrix between 0 and 1 and restricting the matrix to 178 nations with available external data (see below). Using this weight matrix, we simulated outcome $Y$ and predictor $X$ variables using the following formulae from a previous simulation study[@Dow1984]:

$$
\begin{aligned}
Y=(1-\lambda W)^{-1} \zeta + \nu \\
X=(1-\rho W)^{-1} \xi + \omega
\end{aligned}
$$

where $\zeta$, $\nu$, $\xi$, and $\omega$ are all standardised normal random variables, and $\lambda$ and $\rho$ are autocorrelation parameters that vary between 0 and 1, reflecting the strength of spatial autocorrelation for outcome and predictor variables, respectively. Importantly, in this simulation, we know that there is no direct causal relationship between $X$ and $Y$ because we simulated the variables independently. Instead, any relationship between the two variables is merely the result of spatial autocorrelation.

We set the autocorrelation parameters to either 0.1 (weak), 0.5 (moderate), or 0.9 (strong). We simulated 100 datasets for each parameter combination, resulting in 900 datasets. Each dataset had 178 rows, with the following associated data for each nation: latitude, longitude, continent (Asia, North America, Europe, Africa, South America, or Oceania), and legal origin (Islamic, Common, French, German, Mixed, or Scandinavian). Global data on legal origins was taken from a previous study[@Klerman2011].

With the resulting simulated datasets, we standardised outcome and predictor variables and fitted nine different models: (1) naive regression without controls, (2) regression with latitude control, (3) regression with longitude control, (4) regression with continent fixed effects, (5) regression with legal origin fixed effects, (6) regression employing Conley standard errors with 100 km cutoff, (7) regression employing Conley standard errors with 1,000 km cutoff, (8) regression employing Conley standard errors with 10,000 km cutoff, and (9) Bayesian regression including a Gaussian process over latitudes and longitudes.

Models employing Conley standard errors required latitude and longitude values, and the cutoffs implied the distance beyond which spatial autocorrelation is assumed to be zero. These models were fitted using the _conleyreg_ R package[@Duben2021]. Bayesian Gaussian process models were fitted using the _brms_ R package[@Burkner2017]. Our choice of priors was based on prior predictive simulation. All models converged normally ($\hat{R}$ < 1.1). Across all model types and parameter combinations, we calculated the false positive rate as the proportion of models that estimated slopes with a 95% confidence / credible interval excluding zero. We calculated 95% bootstrap confidence intervals for these false positive rates with 1000 bootstrap iterations.

## Reanalyses

We searched the individual analyses from our systematic review for cross-national correlations with available data. We included only analyses for which we were able to replicate the original result (i.e. find a cross-national correlation with the same sign and roughly the same effect size). We restricted our search to one analysis per article, and aimed for an even number of analyses for both economic development and cultural values studies. We also ensured that at least one analysis was a multilevel model, with multiple observations per nation.

The twelve analyses that we settled on[@Adamczyk2009;@Alesina2013;@Beck2003;@Beck2005;@Bockstette2002;@Easterly2003;@Easterly2007;@Fincher2008;@Gelfand2011;@Inglehart2000;@Knack1997;@Skidmore2002] were mostly bivariate cross-national correlations, except for two. One analysis[@Gelfand2011] additionally controlled for log gross national income, and another analysis[@Skidmore2002] is a multilevel model including random intercepts for nations and several individual-level and national-level covariates (see Model 5 in original paper). Before running any additional models, we pre-registered these twelve analyses on the Open Science Framework on 25^th^ January 2022 (https://osf.io/uywx8/).

For each individual analysis, we conducted four models: (1) a naive regression replicating the original finding, (2) a regression including a Gaussian process allowing nation random intercepts to covary according to a geographic proximity matrix from latitude and longitude values, (3) a regression including nation random intercepts that covaried according to a linguistic proximity matrix, and (4) a regression including both a geographic Gaussian process and nation random intercepts with linguistic covariance. See Supplementary Methods for full models.

We fitted these models using the _brms_ R package[@Burkner2017]. Our choice of priors was based on prior predictive simulation. All models converged normally ($\hat{R}$ < 1.1), though for some models we resorted to using approximate Gaussian processes[@Solin2020] to reach convergence.

## Reproducibility

All data and code are accesible at our Open Science Framework repository (https://osf.io/uywx8/). We used the _targets_ R package[@Landau2021] to create a reproducible data analysis pipeline and the _papaja_ R package[@Aust2020] to reproducibly generate the manuscript.

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

### 2022-03-09 manuscript.tex
first commit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Cross-national studies have not sufficiently accounted for non-independence between nations},
  pdfauthor={Scott Claessens1 \& Quentin D. Atkinson1,2},
  pdfkeywords={spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses},
  colorlinks=true,
  linkcolor=Maroon,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{Non-independence between nations}
\keywords{spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses\newline\indent Word count: 5549 words}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\raggedbottom
\usepackage{setspace}
\AtBeginEnvironment{tabular}{\singlespacing}
\AtBeginEnvironment{lltable}{\singlespacing}
\AtBeginEnvironment{tablenotes}{\doublespacing}
\captionsetup[table]{font={stretch=1.5}}
\captionsetup[figure]{font={stretch=1.5}}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[shorthands=off,main=english]{babel}
\fi

\title{Cross-national studies have not sufficiently accounted for non-independence between nations}
\author{Scott Claessens\textsuperscript{1} \& Quentin D. Atkinson\textsuperscript{1,2}}
\date{}

\authornote{

This working paper has not yet been peer-reviewed.

The authors made the following contributions. Scott Claessens: Data curation, Formal analysis, Investigation, Software, Visualization, Writing - original draft, Writing - review \& editing; Quentin D. Atkinson: Conceptualization, Funding acquisition, Supervision, Writing - review \& editing.

Correspondence concerning this article should be addressed to Quentin D. Atkinson, Floor 2, Building 302, 23 Symonds Street, Auckland, 1010, New Zealand. E-mail: \href{mailto:q.atkinson@auckland.ac.nz}{\nolinkurl{q.atkinson@auckland.ac.nz}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} School of Psychology, University of Auckland, Auckland, New Zealand\\\textsuperscript{2} Max Planck Institute for the Science of Human History, Jena, Germany}

\abstract{
Cross-national analyses can illuminate patternings of national-level outcomes around the world. But since nations are related via spatial proximity and shared cultural ancestry, cross-national analyses often violate statistical assumptions of non-independence, inflating false positive rates. Here, we show that past work in economics and psychology has not sufficiently accounted for non-independence between nations. In a review of the 100 highest-cited cross-national studies of economic development and cultural values, we find that controls for non-independence are rare and have not increased in frequency over the past 30 years. When studies do attempt to control for non-independence, our simulations suggest that the methods used continue to produce false positives. In replications of a subsample of twelve cross-national relationships, we show that half fail to replicate when controlling for spatial and/or cultural non-independence using global proximity matrices. We urge social scientists to sufficiently control for non-independence when reporting cross-national associations.
}

\begin{document}
\maketitle

Nations are perhaps the single most important population unit structuring modern human life. The nation in which someone is born has a large effect on what they can expect out of life, including their income level\textsuperscript{1}, life expectancy\textsuperscript{2}, mental health\textsuperscript{3}, subjective well-being\textsuperscript{4}, and educational attainment\textsuperscript{5}. Nations are also among the most important determinants of human cultural variation, with greater cultural similarity within than outside national borders\textsuperscript{6}.

Given the importance of nations for structuring human behaviour, there is justifiably huge interest in statistical analyses that attempt to predict variation in national-level outcomes of all kinds. At the time of writing, a search in Google Scholar for the term \enquote{cross-national study} returned 180,000 hits. The standard practice for cross-national analyses is to conduct bivariate correlations or multiple regressions with individual data points representing different nations. Such analyses widen the scope of social science beyond Western populations\textsuperscript{7,8} and have been used to study, among other topics, the causes of variation in the economic wealth of nations\textsuperscript{9--12}, global patterings of cultural norms and values\textsuperscript{13--16}, and the universality and diversity of human behaviour and psychology around the world\textsuperscript{17--20}.

However, cross-national analyses are complicated by the fact that nations are not statistically independent data points. Unlike independent random samples from a population, nations are related to one another in a number of ways. First, nations that are closer to one another tend to be more similar than distant nations. This phenomenon is known as spatial non-independence\textsuperscript{21}, and it occurs because nations in close spatial proximity share characteristics due to local cultural diffusion of ideas\textsuperscript{22} and common climates and environments\textsuperscript{21}. For example, the neighbouring African nations Zambia and Tanzania have similar levels of terrain ruggedness, which has been used to partially explain their similar levels of economic development\textsuperscript{23}. This pattern conforms to Tobler's first law of geography: \enquote{everything is related to everything else, but near things are more related than distant things}\textsuperscript{24} (p.~236).

Second, nations with shared cultural ancestry tend to be more similar than culturally unrelated nations. This is known as cultural phylogenetic non-independence\textsuperscript{25}, and occurs because related nations share cultural traits inherited via descent from a common ancestor. Shared cultural ancestry can result in a form of pseudoreplication, whereby multiple instances of the same trait across nations are merely duplicates of the ancestral original. For example, the related island nations New Zealand and Samoa express similar languages and customs due to cultural inheritance from a shared ancestor deep in the evolutionary history of Polynesian societies. More recently, shared ancestry is also crucial in explaining cultural similarities between colonial settlements and their colonisers, even those separated by large geographic distances (e.g.~Spain and Argentina).

Spatial and cultural phylogenetic non-independence between nations make cross-national inference challenging. A fundamental assumption of regression analysis is that model residuals should be independently and identically distributed\textsuperscript{26}. But without accounting for spatial or cultural non-independence between nations, model residuals can show structure that remains unaccounted for, violating this assumption (Figure \ref{fig:spatialExample}). Treating nations as independent can thus inflate false positive rates\textsuperscript{27}, producing spurious \enquote{direct} relationships between variables that in fact only indirectly covary due to spatial or cultural connections (see Supplementary Figure \ref{fig:dag} for an example causal model).

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/spatialExample-1.pdf}
\caption{\label{fig:spatialExample}\emph{An example of spatial non-independence in cross-national analysis.} National-level outcome and predictor variables for 25 nations are simulated with high levels of spatial non-independence. (top) Scatterplot visualising the data with fitted line from naive regression without accounting for spatial non-independence. Vertical line segments represent model residual errors, coloured by their sign and size. Countries are labelled with ISO 3166-1 alpha-2 letter codes. (bottom) Model residual errors plotted onto a world map. In this example, residuals clearly show spatial structure that remains unaccounted for.}
\end{figure}

Non-independence between data points is widely acknowledged in fields that routinely deal with spatially or culturally structured data. In ecology and sociology, studies with regional-level data use a variety of autoregressive models designed to account for spatial patternings\textsuperscript{28,29}. In anthropology, the Standard Cross-Cultural Sample of 186 cultures was compiled to minimise the confounding effects of non-independence in comparisons of traditional societies\textsuperscript{30}, though spatial and cultural dependencies are difficult to remove entirely\textsuperscript{31,32}. Anthropologists also borrow phylogenetic comparative methods from evolutionary biology, such as phylogenetic least squares regression\textsuperscript{33}, when comparing traditional societies, treating culturally related societies in the same way as biologists treat genetically related species (e.g.~\textsuperscript{34,35}).

At the national level, recent reanalyses have revealed that several cross-national relationships reported in economics and psychology do not hold when controlling for non-independence between nations. One study replicated 25 analyses of \enquote{persistence} in economics, in which modern national outcomes are regressed against historical characteristics of those nations, and found that over half of the relationships were attenuated when controlling for spatial non-independence\textsuperscript{36}. Another replication study found that many of the widely publicised relationships between national-level pathogen prevalence and political institutions and attitudes fail to hold when controlling for various kinds of non-independence\textsuperscript{37}. These reanalyses, and others\textsuperscript{38,39}, raise the question: how widespread a concern is non-independence in studies of national-level outcomes?

To address this question, we consider national-level variables of general interest across the social sciences: economic development and cultural values. These variables are frequently included as both outcomes and predictors in cross-national studies in economics and psychology\textsuperscript{9--16}. First, we show that economic development and cultural values are spatially and culturally non-independent across nations, emphasising the need to control for non-independence. Second, we systematically review the 100 highest-cited cross-national studies of economic development and cultural values and estimate the proportion of cross-national analyses within these articles that account for non-independence between nations. Third, we run simulations to determine whether common methods of dealing with non-independence in the literature sufficiently reduce false positive rates. Fourth, we reanalyse twelve previous cross-national analyses of economic development and cultural values from our systematic review, incorporating global geographic and linguistic proximity matrices to correctly control for spatial and cultural non-independence.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{national-level-economic-development-and-cultural-values-are-spatially-and-culturally-non-independent}{%
\subsection{National-level economic development and cultural values are spatially and culturally non-independent}\label{national-level-economic-development-and-cultural-values-are-spatially-and-culturally-non-independent}}

In order to motivate our research question, it is important to first show that economic development and cultural values are indeed spatially and culturally non-independent around the world. To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for the Human Development Index\textsuperscript{40} and two dimensions of cultural values from the World Values Survey, traditional vs.~secular values and survival vs.~self-expression values\textsuperscript{16}. For all three of these variables, we found that a substantial proportion of national-level variation was explained by spatial proximity and shared cultural ancestry between nations (Figure \ref{fig:plotSignal}). Bayes Factors indicated strong evidence that these estimates of geographic and cultural phylogenetic signal differed from zero, aside from the geographic signal for survival values, for which the evidence was equivocal (see Supplementary Results). These findings emphasise the need to account for spatial and cultural non-independence in cross-national analyses of economic development and cultural values.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotSignal-1.pdf}
\caption{\label{fig:plotSignal}\emph{Posterior estimates of geographic and cultural phylogenetic signal for the Human Development Index and Inglehart's traditional and survival values.} Geographic and cultural phylogenetic signal are operationalised as the proportion of national-level variance explained by geographic and linguistic proximity matrices. Grey ridges are full posterior distributions, points are posterior median values, and lines are 50\% and 95\% credible intervals.}
\end{figure}

\hypertarget{previous-cross-national-analyses-have-not-sufficiently-accounted-for-non-independence}{%
\subsection{Previous cross-national analyses have not sufficiently accounted for non-independence}\label{previous-cross-national-analyses-have-not-sufficiently-accounted-for-non-independence}}

Since economic development and cultural values show evidence of geographic and cultural phylogenetic signal, have cross-national analyses sufficiently accounted for this non-independence? To assess this, we systematically searched the published literature for articles that combined the search terms \enquote{economic development} or \enquote{values} with the search terms \enquote{cross-national}, \enquote{cross-cultural}, or \enquote{cross-country}. We removed articles that did not report original research, were not relevant to economic development or cultural values, or did not report at least one cross-national analysis. We then retained the 100 articles (50 for economic development, 50 for cultural values) with the highest annual rate of citations. For each of these highly-cited articles, we exhaustively recorded every cross-national analysis reported in the main text (n = 4,308), identifying in each case whether or not the analysis attempted to control for spatial, cultural, or any other form of non-independence between nations (see Methods for detailed search criteria and coding decisions).

The results of our systematic review show that most published articles and cross-national analyses make no attempt to account for statistical non-independence. Figure \ref{fig:plotReview}a plots the proportion of articles that contain at least one cross-national analysis accounting for non-independence. We find that 42\% of economic development articles contain at least one attempt to control for non-independence (95\% bootstrap confidence interval {[}0.30 0.54{]}), while this proportion decreases to only 8\% for cultural values articles (95\% bCI {[}0.02 0.16{]}). Both kinds of article are most likely to use regional fixed effects (e.g.~continent fixed effects) to account for non-independence, but some articles also include controls for spatial distance (e.g.~latitude) and shared cultural history (e.g.~legal origin fixed effects).

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotReview-1.pdf}
\caption{\label{fig:plotReview}\emph{Results from systematic review of 100 highly-cited cross-national studies of economic development (red) and cultural values (blue).} (a) Proportion of articles containing at least one analysis accounting for non-independence, overall and split by common methods of controlling for non-independence. Point ranges represent proportions and 95\% bootstrap confidence intervals. (b) Adjusted proportion of individual analyses accounting for non-independence, overall and split by method. Point ranges are posterior medians and 95\% credible intervals from Bayesian multilevel logistic regressions. (c) Histogram of publication years for studies of economic development and cultural values. (d) Estimated trend over time for the probability of controlling for non-independence. Lines and shaded areas are posterior median regression lines and 50\% credible intervals from a Bayesian multilevel spline model. Region FEs = region fixed effects.}
\end{figure}

Focusing on the full sample of 4,308 analyses, we find that the proportion of individual cross-national analyses accounting for non-independence is even lower (Figure \ref{fig:plotReview}b). Across 2,487 cross-national analyses from studies of economic development, only 5\% are estimated to control for non-independence (95\% credible interval {[}0.02 0.14{]}). Similarly, across 1,821 cross-national analyses from studies of cultural values, only 1\% are estimated to control for non-independence (95\% CI {[}0.00 0.02{]}).

Since our systematic review goes back as far as 1993 (Figure \ref{fig:plotReview}c), it is possible that our estimates are being biased by earlier studies, and that controls for non-independence have increased over time with methodological advancements and greater awareness of the issue. To test this possibility, we fitted a time trend to the full sample of analyses with a Bayesian multilevel spline model. However, we found that, for both studies of economic development and cultural values, the estimated probability of controlling for non-independence has remained low since 1993 (Figure \ref{fig:plotReview}d).

\hypertarget{common-methods-of-controlling-for-non-independence-produce-inflated-false-positive-rates}{%
\subsection{Common methods of controlling for non-independence produce inflated false positive rates}\label{common-methods-of-controlling-for-non-independence-produce-inflated-false-positive-rates}}

Our systematic review revealed that most cross-national analyses in the literature do not control for spatial or cultural phylogenetic non-independence. When they do, they tend to include controls like regional fixed effects, latitude, and legal origin fixed effects. Do these methods sufficiently account for statistical non-independence?

To compare the efficacy of different methods in the literature, we conducted a simulation study. We simulated national-level datasets (n = 178 nations) with varying degrees of spatial autocorrelation (i.e.~non-independence) for outcome and predictor variables, but with no direct causal relationship between the variables. We then fitted naive regressions without controls to these datasets, as well as regression models with the following controls: latitude, longitude, continent fixed effects, and legal origin fixed effects. Despite not being identified in our systematic review, we also included in our simulation study Conley standard errors, a widely used standard error correction that purportedly accounts for spatial non-independence\textsuperscript{41,42}. Finally, we included a multilevel model of spatial non-independence that has recently been made widely available in statistical software packages\textsuperscript{43,44}: Bayesian Gaussian process regression\textsuperscript{45,46}. This method allows nation random intercepts to covary according to a global geographic proximity matrix. Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero (i.e.~falsely infering a relationship when none is present).

Figure \ref{fig:plotSim2} plots the estimated false positive rates from our simulation study, split by different methods and different degrees of spatial autocorrelation (see Supplementary Figure \ref{fig:plotSim1} for full distribution of effect sizes under moderate autocorrelation). With at least moderate degrees of spatial autocorrelation for both outcome and predictor variables, naive regression models (Figure \ref{fig:plotSim2} top left) produce false positive rates above chance levels. With strong spatial autocorrelation for both outcomes and predictors, false positive rates reach as high as 81\%.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotSim2-1.pdf}
\caption{\label{fig:plotSim2}\emph{False positive rates for different methods of controlling for non-independence in our simulation study.} For simulated outcome and predictor variables, we systematically varied the strength of spatial autocorrelation, from weak (0.1) to moderate (0.5) to strong (0.9). We simulated 100 datasets per parameter combination and fitted different models to each dataset. False positive rates were operationalised as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero. Point ranges represent proportions and 95\% bootstrap confidence intervals, and dashed lines indicate the 5\% false positive rate that is expected due to chance. SEs = standard errors.}
\end{figure}

Common methods in the literature do not reduce this high false positive rate. With strong spatial autocorrelation for both outcome and predictor variables, false positive rates remain above 70\% when controlling for latitude, longitude, continent fixed effects, and legal origin fixed effects. Applying Conley standard errors with cutoffs of 100 km or 1,000 km also does not reduce false positive rates below 70\% under strong spatial autocorrelation. With a cutoff of 10,000 km, the false positive rate under strong spatial autocorrelation reduces to 57\%, though this correction adds bias at weaker levels of spatial autocorrelation, increasing false positive rates above chance levels.

By contrast, Bayesian Gaussian process regression adequately accounts for non-independence across a range of simulation parameters. At moderate levels of spatial autocorrelation on both outcomes and predictors, Gaussian process regression eliminates false positives, such that the false positive rate is no different from chance. With strong spatial autocorrelation on both outcomes and predictors, the false positive rate is reduced to 30\%, substantially less than any other method.

\hypertarget{key-findings-in-the-literature-are-not-robust-to-reanalysis-with-more-rigorous-methods}{%
\subsection{Key findings in the literature are not robust to reanalysis with more rigorous methods}\label{key-findings-in-the-literature-are-not-robust-to-reanalysis-with-more-rigorous-methods}}

Our systematic review and simulation study have showed that controls for non-independence are rare in cross-national studies of economic development and cultural values, and when studies do attempt to control for non-independence, the methods used continue to produce false positives. This raises the worrying possibility that the cross-national literature in economics and psychology is populated with spurious relationships.

To diagnose how widespread this issue of spurious cross-national relationships might be, we reanalysed a subset of twelve previous cross-national analyses from our systematic review, sufficiently controlling for spatial and cultural phylogenetic non-independence using global geographic and linguistic proximity matrices. We subsampled six analyses from our economic development review\textsuperscript{47--52} and six from our cultural values review\textsuperscript{13,14,16,53--55}. Our choice of analyses was constrained by data availability and whether we were able to initially replicate the original finding. We pre-registered our subsample of analyses before running any control models (\url{https://osf.io/uywx8/}). We controlled for non-independence by including (1) a Gaussian process allowing nation random intercepts to covary according to a geographic proximity matrix, and/or (2) nation random intercepts that covaried according to a linguistic proximity matrix (see Supplementary Methods for full models).

Figure \ref{fig:plotReplications1} visualises the results of our reanalysis. Overall, cross-national correlation effect sizes tended to reduce when controlling for statistical non-independence between nations, sometimes by as much as half of the original effect size. Based on 95\% credible intervals, six out of twelve cross-national relationships are attenuated when controlling for non-independence. For the economic development analyses, four out of six cross-national relationships are attenuated when controlling for spatial non-independence. For the cultural values analyses, two out of six are attenuated when controlling for cultural phylogenetic non-independence. Supplementary Figure \ref{fig:plotReplications2} shows these cross-national correlations plotted against the raw data.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotReplications1-1.pdf}
\caption{\label{fig:plotReplications1}\emph{Posterior correlations from our reanalysis of twelve previous cross-national analyses.} For each previous cross-national relationship, we plot the posterior slopes from a naive regression (red), a regression controlling for spatial non-independence (green), a regression controlling for cultural phylogenetic non-independence (blue), and a regression controlling for both spatial and cultural phylogenetic non-independence simultaneously (purple). All outcome and predictor variables are standardised. Most analyses are simple bivariate cross-national correlations, but Gelfand et al.~(2011) is a partial correlation controlling for log gross national income and Adamczyk and Pitt (2009) is a multilevel model including several covariates. Point ranges represent posterior medians and 95\% credible intervals. GDP = gross domestic product. FLFP = female labour force participation.}
\end{figure}

To understand why some cross-national correlations were attenuated by controls for non-independence while others were robust, we further explored our fitted models for evidence of spatial and cultural autocorrelation. For each outcome variable, our Gaussian process models provided varying estimates of how quickly spatial autocorrelation declined with distance (Supplementary Figure \ref{fig:plotReplications3}). For example, in Skidmore and Toya\textsuperscript{52} gross domestic product growth was only moderately spatially autocorrelated at 1000 km distance (posterior median spatial autocorrelation at 1000 km = 0.43, 95\% CI {[}0.05 0.96{]}), whereas in Inglehart and Baker\textsuperscript{16} traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1000 km = 0.96, 95\% CI {[}0.77 0.99{]}). We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \ref{fig:plotReplications4}), with some outcome variables expressing low signal (e.g.~Gelfand et al.\textsuperscript{14}; posterior median = 0.17, 95\% CI {[}0.00 0.93{]}) and others expressing high signal (e.g.~Alesina et al.\textsuperscript{54}; posterior median = 0.89, 95\% CI {[}0.66 0.98{]}). Across all analyses, we found that stronger estimates of spatial autocorrelation or cultural phylogenetic signal resulted in a more pronounced reduction in the effect size when controlling for non-independence between nations (Figure \ref{fig:plotReplications5}).

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotReplications5-1.pdf}
\caption{\label{fig:plotReplications5}\emph{The estimated degree of spatial and cultural phylogenetic non-independence predicts reductions in effect size in our reanalysis.} (a) Higher estimated degrees of spatial autocorrelation at 1000 km distance predict more pronounced reductions in effect sizes when controlling for non-independence. (b) Higher estimated levels of cultural phylogenetic signal predict more pronounced reductions in effect sizes when controlling for non-independence. In both panels, the y-axis represents the ratio of the effect size when controlling for spatial and cultural non-independence to the original effect size (from naive regression model), and the x-axis represents posterior median model estimates. Regression lines are plotted with 95\% confidence intervals.}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

In a systematic literature review and simulation, we found that cross-national studies in economics and psychology rarely account for non-independence between nations, and, when they do, the methods they use continue to produce false positives. In a reanalysis of twelve cross-national correlations, we further showed that neglecting to account for non-independence has resulted in spurious relationships in the published literature, with half of the correlations failing to replicate when controlling for spatial or cultural non-independence with more rigorous methods. These findings suggest that cross-national analyses in economics and psychology should be interpreted with caution until non-independence is sufficiently accounted for.

Non-independence is a concern because economic development and cultural values are highly geographically and culturally structured around the world, as revealed by our initial modelling. These results build on previous work. One previous study suggested that geographic proximity is more important than deep cultural ancestry in explaining the distribution of human development across Eurasian nations\textsuperscript{56}. By contrast, with a global sample of nations, we found strong cultural phylogenetic signal, as well as geographic signal, for the Human Development Index. Another previous study found that similarities in the cultural values of nations are predicted by linguistic, but not geographic, distances between those nations\textsuperscript{6}. We find this same result for survival vs.~self-expression values, but for traditional vs.~secular values we find that both linguistic and geographic proximity are important independent predictors of global variation. These findings emphasise the need to account for both spatial and cultural phylogenetic non-independence in cross-national studies of economic development and cultural values.

Yet our systematic literature review and simulation study revealed that the most commonly used controls for non-independence do not sufficiently deal with the issue. In our simulations, controlling for either latitude or longitude did not reduce false positive rates, calling into question controls like distance to the equator in regression models. High false positive rates persisted with Conley standard errors, which have previously been critiqued for being overly sensitive to arbitrary distance cutoffs\textsuperscript{36}. The simulation also confirmed the verbal claim that fixed effects for spatial or cultural groupings (e.g.~continent or legal origin fixed effects) are insufficient because non-independence still remains within groupings\textsuperscript{37}. This logic applies to analyses that control for non-independence by including language family fixed effects (e.g.~\textsuperscript{57}) or separately analysing different regions (e.g.~\textsuperscript{58}). The only method in our simulation that sufficiently reduced the false positive rate was Bayesian Gaussian process regression, though we did not include other possible controls for non-independence, such as conditional autoregressive models\textsuperscript{28} or generalised additive models\textsuperscript{59}.

Ours is not the first review to show that studies are misapplying statistical methods in ways that inflate false positive rates. For example, other literature reviews have shown that studies in the social sciences tend to use small sample sizes\textsuperscript{60}, treat ordinal data as metric\textsuperscript{61}, incorrectly handle missing values\textsuperscript{62}, and ignore best practices in meta-analyses\textsuperscript{63}. Why do cross-national studies also rarely account for non-independence? At the institutional level, one possibility is that such practices are incentivised \emph{because} they inflate false positive rates, increasing the probability of getting published and promoted\textsuperscript{60}. At the individual level, another possibility is that researchers outside of anthropology and ecology are simply not aware of the problem, or believe that the problem does not apply at the national level. Even if researchers appreciate the problem, they might not know of suitable controls or perceive the methods to be too complex.

These institutional- and individual-level barriers can be combatted. First, cross-national replication studies like ours and others\textsuperscript{36--39}, combined with the methodological reviews included in Registered Reports\textsuperscript{64}, might change incentive structures and encourage researchers to analyse the world's nations with more rigorous methods. Second, since the issue of non-independence is fundamentally an issue of causal inference (Supplementary Figure \ref{fig:dag}), more explicit descriptions of causal models could promote controls for non-independence in cross-national research. In our review, economists studying economic development dealt with national-level non-independence more than psychologists studying cultural values, likely because economics studies tend to be lengthy statistical exercises that systematically incorporate or exclude numerous variables in an attempt to infer causation. Third, the recent widespread accessibility of open source statistical software, such as the programming language Stan\textsuperscript{44} and the R package \emph{brms}\textsuperscript{43}, should promote the use of more rigorous methods to control for non-independence. Using \emph{brms}, for example, Bayesian Gaussian process regression is straightforward to conduct, requiring only longitude and latitude values for nations.

Until such changes are implemented and sufficient controls for non-independence are the norm, existing cross-national correlations should be interpreted with caution. In our reanalyses, we found that four out of six cross-national analyses of economic development were attenuated when controlling for spatial non-independence. Three of these were tests of \enquote{persistence} hypotheses, studying the effects of historical and environmental conditions --- settler mortality\textsuperscript{47}, wheat-sugar suitability\textsuperscript{51}, and natural disaster frequency\textsuperscript{52} --- on modern developmental outcomes. A recent reanalysis has also called into question various studies of this ilk\textsuperscript{36}. We also found that two out of six cross-national analyses of cultural values were attenuated when controlling for cultural phylogenetic non-independence. It is interesting that the economic development analyses were attenuated with spatial controls, while the cultural values analyses were attenuated by cultural phylogenetic controls. This highlights that certain kinds of non-independence will be more important depending on the particular national-level outcome and predictor variables of interest.

We do not wish to dissuade researchers from conducting cross-national studies. On the contrary, such work promises to deepen our understanding of human populations beyond Western, educated, industrialised, rich, and democratic societies\textsuperscript{7}, providing the representative samples necessary to test evolutionary and socio-ecological theories of human behaviour\textsuperscript{8,65}. But in order to minimise spurious relationships in global datasets, we urge researchers to control for spatial and cultural phylogenetic non-independence when reporting cross-national correlations. Nations are not independent, and our statistical models should reflect this.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{geographic-and-cultural-phylogenetic-signal}{%
\subsection{Geographic and cultural phylogenetic signal}\label{geographic-and-cultural-phylogenetic-signal}}

To estimate the degree of spatial and cultural phylogenetic non-independence in economic development and cultural values, we calculated geographic and linguistic signal for global measures of development and values. Our measure of economic development was the Human Development Index\textsuperscript{40}. We retrieved a longitudinal dataset capturing human development for 189 nations since 1990 (n = 1,512; \url{https://hdr.undp.org/en/content/download-data}). Our measures of cultural values were traditional vs.~secular values and survival vs.~self-expression values from the World Values Survey\textsuperscript{16}. We downloaded the full Integrated Values Survey, which included all waves from the World Values Survey and the European Values Survey, and computed the two dimensions of cultural values following procedures from previous research\textsuperscript{16}. This longitudinal dataset captures values and attitudes for 116 nations since 1981 (n = 645,249; \url{https://www.worldvaluessurvey.org/WVSEVStrend.jsp}).

To calculate geographic and linguistic signal, we created two proximity matrices for 269 of the world's nations: a geographic proximity matrix and a linguistic proximity matrix. Geographic proximity was converted from logged geodesic distances between nation capital cities. Linguistic proximity was calculated as the cultural proximity between all languages spoken within nations, weighted by speaker percentages (see Supplementary Methods). We included these matrices in Bayesian multilevel models, allowing nation random intercepts to covary according to both geographic and linguistic proximity simultaneously. These models were fitted with the R package \emph{brms}\textsuperscript{43} and converged normally (\(\hat{R}\) \textless{} 1.1). Estimates of geographic and linguistic signal were computed as the proportion of national-level variance in these models explained by geographic and linguistic proximity matrices.

\hypertarget{systematic-literature-review}{%
\subsection{Systematic literature review}\label{systematic-literature-review}}

We exported two searches from Web of Science (\url{https://www.webofknowledge.com/}) on 27\textsuperscript{th} September 2021, restricting our searches to articles published between 1900 and 2018. The first search was for the terms \enquote{economic development} AND (\enquote{cross-national} OR \enquote{cross-cultural} OR \enquote{cross-country}), which returned 965 articles. The second search was for the terms \enquote{values} AND (\enquote{cross-national} OR \enquote{cross-cultural} OR \enquote{cross-country}), which returned 6806 articles. Once exported, we ordered the articles by descending number of citations per year since initial publication, using citation counts reported by Web of Science.

We then systematically coded each article, in order, for inclusion in our review. Articles were only included if: (1) they were judged to be relevant to economic development or cultural values; (2) they were an original empirical research article; and (3) they contained at least one analysis with national-level outcome or predictor variables. We stopped when we had included 50 articles for the economic development review and 50 articles for the cultural values review.

Within each included article, we exhaustively coded every individual cross-national analysis reported in the main text. We coded mainly correlation or regression analyses, and explicitly excluded meta-analyses, factor analyses, measurement invariance analyses, multidimensional scaling analyses, hierarchical clustering analyses, multiverse analyses, and scale development / validation analyses. We also excluded analyses that compared only two, three, four, five, or six nations. For each included analysis, we recorded the year, outcome variable, all predictor variables, test statistic, p-value, number of nations, number of data points, model type, if the data were available, and whether and how the analysis attempted to control for non-independence.

We coded common attempts to control for non-independence between nations. These included: (1) any higher-level control variables for spatial regional groupings (e.g.~continent fixed effects); (2) any geographic distance control variables (e.g.~distance between capital cities, distance from equator, latitude); (3) any control variables capturing shared cultural history (e.g.~former colony, legal origin fixed effects, linguistic history, cultural influence); and (4) any other control variables, tests, or approaches that were deemed as attempts to control for non-independence (e.g.~eigenvector filtering\textsuperscript{66}, controls for trade-weightings between countries, cross-sectional dependence tests\textsuperscript{67}, separate analyses for subsets of countries). These were coded by the first author.

Once we had compiled our review database, we calculated the proportion of articles attempting to control for non-independence at least once. We also calculated the proportion of articles employing the different types of control listed above at least once: regional fixed effects, distance, shared cultural history, or other. For these proportions, we calculated 95\% boostrap confidence intervals with 1000 bootstrap iterations.

For individual analyses, we dealt with the nested nature of the data (analyses nested within articles) by fitting Bayesian multilevel logistic regression models with review type (economic development vs.~cultural values) as the sole fixed effect and random intercepts for articles. We fitted these models separately for overall attempts to control for non-independence and split by method type. We report the adjusted proportions with 95\% credible intervals. To test for a trend over time, we also fitted a Bayesian multilevel logistic regression with a multigroup spline for year of publication and random intercepts for articles. Bayesian models were fitted with the \emph{brms} R package\textsuperscript{43}. Our priors were informed by prior predictive checks, and all models converged normally (\(\hat{R}\) \textless{} 1.1).

\hypertarget{simulations}{%
\subsection{Simulations}\label{simulations}}

To simulate data, we used our global geographic proximity matrix (see Supplementary Methods) as a weight matrix \(W\), scaling the matrix between 0 and 1 and restricting the matrix to 178 nations with available external data (see below). Using this weight matrix, we simulated outcome \(Y\) and predictor \(X\) variables using the following formulae from a previous simulation study\textsuperscript{68}:

\[
\begin{aligned}
Y=(1-\lambda W)^{-1} \zeta + \nu \\
X=(1-\rho W)^{-1} \xi + \omega
\end{aligned}
\]

where \(\zeta\), \(\nu\), \(\xi\), and \(\omega\) are all standardised normal random variables, and \(\lambda\) and \(\rho\) are autocorrelation parameters that vary between 0 and 1, reflecting the strength of spatial autocorrelation for outcome and predictor variables, respectively. Importantly, in this simulation, we know that there is no direct causal relationship between \(X\) and \(Y\) because we simulated the variables independently. Instead, any relationship between the two variables is merely the result of spatial autocorrelation.

We set the autocorrelation parameters to either 0.1 (weak), 0.5 (moderate), or 0.9 (strong). We simulated 100 datasets for each parameter combination, resulting in 900 datasets. Each dataset had 178 rows, with the following associated data for each nation: latitude, longitude, continent (Asia, North America, Europe, Africa, South America, or Oceania), and legal origin (Islamic, Common, French, German, Mixed, or Scandinavian). Global data on legal origins was taken from a previous study\textsuperscript{69}.

With the resulting simulated datasets, we standardised outcome and predictor variables and fitted nine different models: (1) naive regression without controls, (2) regression with latitude control, (3) regression with longitude control, (4) regression with continent fixed effects, (5) regression with legal origin fixed effects, (6) regression employing Conley standard errors with 100 km cutoff, (7) regression employing Conley standard errors with 1,000 km cutoff, (8) regression employing Conley standard errors with 10,000 km cutoff, and (9) Bayesian regression including a Gaussian process over latitudes and longitudes.

Models employing Conley standard errors required latitude and longitude values, and the cutoffs implied the distance beyond which spatial autocorrelation is assumed to be zero. These models were fitted using the \emph{conleyreg} R package\textsuperscript{70}. Bayesian Gaussian process models were fitted using the \emph{brms} R package\textsuperscript{43}. Our choice of priors was based on prior predictive simulation. All models converged normally (\(\hat{R}\) \textless{} 1.1). Across all model types and parameter combinations, we calculated the false positive rate as the proportion of models that estimated slopes with a 95\% confidence / credible interval excluding zero. We calculated 95\% bootstrap confidence intervals for these false positive rates with 1000 bootstrap iterations.

\hypertarget{reanalyses}{%
\subsection{Reanalyses}\label{reanalyses}}

We searched the individual analyses from our systematic review for cross-national correlations with available data. We included only analyses for which we were able to replicate the original result (i.e.~find a cross-national correlation with the same sign and roughly the same effect size). We restricted our search to one analysis per article, and aimed for an even number of analyses for both economic development and cultural values studies. We also ensured that at least one analysis was a multilevel model, with multiple observations per nation.

The twelve analyses that we settled on\textsuperscript{13,14,16,47--55} were mostly bivariate cross-national correlations, except for two. One analysis\textsuperscript{14} additionally controlled for log gross national income, and another analysis\textsuperscript{52} is a multilevel model including random intercepts for nations and several individual-level and national-level covariates (see Model 5 in original paper). Before running any additional models, we pre-registered these twelve analyses on the Open Science Framework on 25\textsuperscript{th} January 2022 (\url{https://osf.io/uywx8/}).

For each individual analysis, we conducted four models: (1) a naive regression replicating the original finding, (2) a regression including a Gaussian process allowing nation random intercepts to covary according to a geographic proximity matrix from latitude and longitude values, (3) a regression including nation random intercepts that covaried according to a linguistic proximity matrix, and (4) a regression including both a geographic Gaussian process and nation random intercepts with linguistic covariance. See Supplementary Methods for full models.

We fitted these models using the \emph{brms} R package\textsuperscript{43}. Our choice of priors was based on prior predictive simulation. All models converged normally (\(\hat{R}\) \textless{} 1.1), though for some models we resorted to using approximate Gaussian processes\textsuperscript{71} to reach convergence.

\hypertarget{reproducibility}{%
\subsection{Reproducibility}\label{reproducibility}}

All data and code are accesible at our Open Science Framework repository (\url{https://osf.io/uywx8/}). We used the \emph{targets} R package\textsuperscript{72} to create a reproducible data analysis pipeline and the \emph{papaja} R package\textsuperscript{73} to reproducibly generate the manuscript.

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Caselli2005}{}%
1. Caselli, F. Accounting for cross-country income differences. in (eds. Aghion, P. \& Durlauf, S. N.) vol. 1 679--741 (Elsevier, 2005).

\leavevmode\hypertarget{ref-Austin2012}{}%
2. Austin, K. F. \& McKinney, L. A. Disease, war, hunger, and deprivation: A cross-national investigation of the determinants of life expectancy in less-developed and sub-Saharan African nations. \emph{Sociological Perspectives} \textbf{55}, 421--447 (2012).

\leavevmode\hypertarget{ref-Rai2013}{}%
3. Rai, D., Zitko, P., Jones, K., Lynch, J. \& Araya, R. Country- and individual-level socioeconomic determinants of depression: Multilevel cross-national comparison. \emph{British Journal of Psychiatry} \textbf{202}, 195--203 (2013).

\leavevmode\hypertarget{ref-Diener2009}{}%
4. Diener, E., Diener, M. \& Diener, C. Factors predicting the subjective well-being of nations. in \emph{Culture and well-being: The collected works of Ed Diener} (ed. Diener, E.) 43--70 (Springer Netherlands, 2009). doi:\href{https://doi.org/10.1007/978-90-481-2352-0_3}{10.1007/978-90-481-2352-0\_3}.

\leavevmode\hypertarget{ref-Kirkcaldy2004}{}%
5. Kirkcaldy, B., Furnham, A. \& Siefen, G. The relationship between health efficacy, educational attainment, and well-being among 30 nations. \emph{European Psychologist} \textbf{9}, 107--119 (2004).

\leavevmode\hypertarget{ref-White2021}{}%
6. White, C. J. M., Muthukrishna, M. \& Norenzayan, A. Cultural similarity among coreligionists within and between countries. \emph{Proceedings of the National Academy of Sciences} \textbf{118}, e2109650118 (2021).

\leavevmode\hypertarget{ref-Henrich2010}{}%
7. Henrich, J., Heine, S. J. \& Norenzayan, A. The weirdest people in the world? \emph{Behavioral and Brain Sciences} \textbf{33}, 61--83 (2010).

\leavevmode\hypertarget{ref-Pollet2014}{}%
8. Pollet, T. V., Tybur, J. M., Frankenhuis, W. E. \& Rickard, I. J. What can cross-cultural correlations teach us about human nature? \emph{Human Nature} \textbf{25}, 410--429 (2014).

\leavevmode\hypertarget{ref-Benhabib1994}{}%
9. Benhabib, J. \& Spiegel, M. M. The role of human capital in economic development: Evidence from aggregate cross-country data. \emph{Journal of Monetary Economics} \textbf{34}, 143--173 (1994).

\leavevmode\hypertarget{ref-Comin2010}{}%
10. Comin, D., Easterly, W. \& Gong, E. Was the wealth of nations determined in 1000 BC? \emph{American Economic Journal: Macroeconomics} \textbf{2}, 65--97 (2010).

\leavevmode\hypertarget{ref-LaPorta1997}{}%
11. La Porta, R., Lopez-De-Silanes, F., Shleifer, A. \& Vishny, R. W. Legal determinants of external finance. \emph{The Journal of Finance} \textbf{52}, 1131--1150 (1997).

\leavevmode\hypertarget{ref-Sachs2001}{}%
12. Sachs, J. D. \& Warner, A. M. The curse of natural resources. \emph{European Economic Review} \textbf{45}, 827--838 (2001).

\leavevmode\hypertarget{ref-Fincher2008}{}%
13. Fincher, C. L., Thornhill, R., Murray, D. R. \& Schaller, M. Pathogen prevalence predicts human cross-cultural variability in individualism/collectivism. \emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{275}, 1279--1285 (2008).

\leavevmode\hypertarget{ref-Gelfand2011}{}%
14. Gelfand, M. J. \emph{et al.} Differences between tight and loose cultures: A 33-nation study. \emph{Science} \textbf{332}, 1100--1104 (2011).

\leavevmode\hypertarget{ref-Hofstede2001}{}%
15. Hofstede, G. \emph{Culture's consequences: Comparing values, behaviors, institutions and organizations across nations}. (Sage Publications, 2001).

\leavevmode\hypertarget{ref-Inglehart2000}{}%
16. Inglehart, R. \& Baker, W. E. Modernization, cultural change, and the persistence of traditional values. \emph{American Sociological Review} \textbf{65}, 19--51 (2000).

\leavevmode\hypertarget{ref-Awad2018}{}%
17. Awad, E. \emph{et al.} The moral machine experiment. \emph{Nature} \textbf{563}, 59--64 (2018).

\leavevmode\hypertarget{ref-Rhoads2021}{}%
18. Rhoads, S. A., Gunter, D., Ryan, R. M. \& Marsh, A. A. Global variation in subjective well-being predicts seven forms of altruism. \emph{Psychological Science} \textbf{32}, 1247--1261 (2021).

\leavevmode\hypertarget{ref-Schulz2019}{}%
19. Schulz, J. F., Bahrami-Rad, D., Beauchamp, J. P. \& Henrich, J. The Church, intensive kinship, and global psychological variation. \emph{Science} \textbf{366}, eaau5141 (2019).

\leavevmode\hypertarget{ref-Thomson2018}{}%
20. Thomson, R. \emph{et al.} Relational mobility predicts social behaviors in 39 countries and is tied to historical farming and threat. \emph{Proceedings of the National Academy of Sciences} \textbf{115}, 7521--7526 (2018).

\leavevmode\hypertarget{ref-Kissling2008}{}%
21. Kissling, W. D. \& Carl, G. Spatial autocorrelation and the selection of simultaneous autoregressive models. \emph{Global Ecology and Biogeography} \textbf{17}, 59--71 (2008).

\leavevmode\hypertarget{ref-Hewlett2002}{}%
22. Hewlett, B. S., DeSilvestri, A. \& Guglielmino, C. R. Semes and genes in africa. \emph{Current Anthropology} \textbf{43}, 313--321 (2002).

\leavevmode\hypertarget{ref-Nunn2012}{}%
23. Nunn, N. \& Puga, D. Ruggedness: The blessing of bad geography in Africa. \emph{The Review of Economics and Statistics} \textbf{94}, 20--36 (2012).

\leavevmode\hypertarget{ref-Tobler1970}{}%
24. Tobler, W. R. A computer movie simulating urban growth in the Detroit region. \emph{Economic Geography} \textbf{46}, 234--240 (1970).

\leavevmode\hypertarget{ref-Tylor1889}{}%
25. Tylor, E. B. On a method of investigating the development of institutions; applied to laws of marriage and descent. \emph{The Journal of the Anthropological Institute of Great Britain and Ireland} \textbf{18}, 245--272 (1889).

\leavevmode\hypertarget{ref-Jarque1987}{}%
26. Jarque, C. M. \& Bera, A. K. A test for normality of observations and regression residuals. \emph{International Statistical Review / Revue Internationale de Statistique} \textbf{55}, 163--172 (1987).

\leavevmode\hypertarget{ref-Legendre1993}{}%
27. Legendre, P. Spatial autocorrelation: Trouble or new paradigm? \emph{Ecology} \textbf{74}, 1659--1673 (1993).

\leavevmode\hypertarget{ref-Lichstein2002}{}%
28. Lichstein, J. W., Simons, T. R., Shriner, S. A. \& Franzreb, K. E. Spatial autocorrelation and autoregressive models in ecology. \emph{Ecological Monographs} \textbf{72}, 445--463 (2002).

\leavevmode\hypertarget{ref-Loftin1983}{}%
29. Loftin, C. \& Ward, S. K. A spatial autocorrelation model of the effects of population density on fertility. \emph{American Sociological Review} \textbf{48}, 121--128 (1983).

\leavevmode\hypertarget{ref-Murdock1969}{}%
30. Murdock, G. P. \& White, D. R. Standard cross-cultural sample. \emph{Ethnology} \textbf{8}, 329--369 (1969).

\leavevmode\hypertarget{ref-Dow2008}{}%
31. Dow, M. M. \& Eff, E. A. Global, regional, and local network autocorrelation in the standard cross-cultural sample. \emph{Cross-Cultural Research} \textbf{42}, 148--171 (2008).

\leavevmode\hypertarget{ref-Eff2004}{}%
32. Eff, E. A. Does Mr. Galton still have a problem? Autocorrelation in the standard cross-cultural sample. \emph{World Cultures} \textbf{15}, 153--170 (2004).

\leavevmode\hypertarget{ref-Symonds2014}{}%
33. Symonds, M. R. E. \& Blomberg, S. P. A primer on phylogenetic generalised least squares. in \emph{Modern phylogenetic comparative methods and their application in evolutionary biology: Concepts and practice} (ed. Garamszegi, L. Z.) 105--130 (Springer Berlin Heidelberg, 2014). doi:\href{https://doi.org/10.1007/978-3-662-43550-2_5}{10.1007/978-3-662-43550-2\_5}.

\leavevmode\hypertarget{ref-Watts2018}{}%
34. Watts, J., Sheehan, O., Bulbulia, J., Gray, R. D. \& Atkinson, Q. D. Christianity spread faster in small, politically structured societies. \emph{Nature Human Behaviour} \textbf{2}, 559--564 (2018).

\leavevmode\hypertarget{ref-Atkinson2016}{}%
35. Atkinson, Q. D., Coomber, T., Passmore, S., Greenhill, S. J. \& Kushnick, G. Cultural and environmental predictors of pre-european deforestation on pacific islands. \emph{PLOS ONE} \textbf{11}, 1--15 (2016).

\leavevmode\hypertarget{ref-Kelly2020}{}%
36. Kelly, M. Understanding persistence. \emph{CEPR Discussion Paper No. DP15246} (2020).

\leavevmode\hypertarget{ref-Bromham2018}{}%
37. Bromham, L., Hua, X., Cardillo, M., Schneemann, H. \& Greenhill, S. J. Parasites and politics: Why cross-cultural studies must control for relatedness, proximity and covariation. \emph{Royal Society Open Science} \textbf{5}, 181100 (2018).

\leavevmode\hypertarget{ref-Bromham2021}{}%
38. Bromham, L., Skeels, A., Schneemann, H., Dinnage, R. \& Hua, X. There is little evidence that spicy food in hot countries is an adaptation to reducing infection risk. \emph{Nature Human Behaviour} \textbf{5}, 878--891 (2021).

\leavevmode\hypertarget{ref-Currie2012}{}%
39. Currie, T. E. \& Mace, R. Analyses do not support the parasite-stress theory of human sociality. \emph{Behavioral and Brain Sciences} \textbf{35}, 83--85 (2012).

\leavevmode\hypertarget{ref-hdi}{}%
40. \emph{Human development report}. \url{http://hdr.undp.org/en/composite/HDI}.

\leavevmode\hypertarget{ref-Conley1999}{}%
41. Conley, T. G. GMM estimation with cross sectional dependence. \emph{Journal of Econometrics} \textbf{92}, 1--45 (1999).

\leavevmode\hypertarget{ref-Conley2010}{}%
42. Conley, T. G. Spatial econometrics. in \emph{Microeconometrics} (eds. Durlauf, S. N. \& Blume, L. E.) 303--313 (Palgrave Macmillan, 2010).

\leavevmode\hypertarget{ref-Burkner2017}{}%
43. Bürkner, P.-C. brms: An R package for Bayesian multilevel models using Stan. \emph{Journal of Statistical Software} \textbf{80}, 1--28 (2017).

\leavevmode\hypertarget{ref-rstan}{}%
44. Stan Development Team. RStan: The R interface to Stan. (2020).

\leavevmode\hypertarget{ref-McElreath2020}{}%
45. McElreath, R. \emph{Statistical rethinking: A Bayesian course with examples in R and Stan}. (CRC Press, 2020).

\leavevmode\hypertarget{ref-Neal1998}{}%
46. Neal, R. M. Regression and classification using Gaussian process priors. in \emph{Bayesian statistics} (eds. Bernardo, J. M., Berger, J. O., Dawid, A. P. \& Smith, A. F. M.) vol. 6 475--501 (Oxford University Press).

\leavevmode\hypertarget{ref-Beck2003}{}%
47. Beck, T., Demirgäç-Kunt, A. \& Levine, R. Law, endowments, and finance. \emph{Journal of Financial Economics} \textbf{70}, 137--181 (2003).

\leavevmode\hypertarget{ref-Beck2005}{}%
48. Beck, T., Demirgäç-Kunt, A. \& Levine, R. SMEs, growth, and poverty: Cross-country evidence. \emph{Journal of economic growth} \textbf{10}, 199--229 (2005).

\leavevmode\hypertarget{ref-Bockstette2002}{}%
49. Bockstette, V., Chanda, A. \& Putterman, L. States and markets: The advantage of an early start. \emph{Journal of Economic growth} \textbf{7}, 347--369 (2002).

\leavevmode\hypertarget{ref-Easterly2003}{}%
50. Easterly, W. \& Levine, R. Tropics, germs, and crops: How endowments influence economic development. \emph{Journal of Monetary Economics} \textbf{50}, 3--39 (2003).

\leavevmode\hypertarget{ref-Easterly2007}{}%
51. Easterly, W. Inequality does cause underdevelopment: Insights from a new instrument. \emph{Journal of Development Economics} \textbf{84}, 755--776 (2007).

\leavevmode\hypertarget{ref-Skidmore2002}{}%
52. Skidmore, M. \& Toya, H. Do natural disasters promote long-run growth? \emph{Economic Inquiry} \textbf{40}, 664--687 (2002).

\leavevmode\hypertarget{ref-Adamczyk2009}{}%
53. Adamczyk, A. \& Pitt, C. Shaping attitudes about homosexuality: The role of religion and cultural context. \emph{Social Science Research} \textbf{38}, 338--351 (2009).

\leavevmode\hypertarget{ref-Alesina2013}{}%
54. Alesina, A., Giuliano, P. \& Nunn, N. On the origins of gender roles: Women and the plough. \emph{Quarterly Journal of Economics} \textbf{128}, 469--530 (2013).

\leavevmode\hypertarget{ref-Knack1997}{}%
55. Knack, S. \& Keefer, P. Does social capital have an economic payoff? A cross-country investigation. \emph{The Quarterly Journal of Economics} \textbf{112}, 1251--1288 (1997).

\leavevmode\hypertarget{ref-Sookias2018}{}%
56. Sookias, R. B., Passmore, S. \& Atkinson, Q. D. Deep cultural ancestry and human development indicators across nation states. \emph{Royal Society Open Science} \textbf{5}, 171411 (2018).

\leavevmode\hypertarget{ref-Currie2009}{}%
57. Currie, T. E. \& Mace, R. Political complexity predicts the spread of ethnolinguistic groups. \emph{Proceedings of the National Academy of Sciences} \textbf{106}, 7339--7344 (2009).

\leavevmode\hypertarget{ref-Smith2017}{}%
58. Smith, M. D., Rabbitt, M. P. \& Coleman- Jensen, A. Who are the world's food insecure? New evidence from the Food and Agriculture Organization's Food Insecurity Experience Scale. \emph{World Development} \textbf{93}, 402--412 (2017).

\leavevmode\hypertarget{ref-Hastie2017}{}%
59. Hastie, T. J. \& Tibshirani, R. J. \emph{Generalized additive models}. (Routledge, 2017).

\leavevmode\hypertarget{ref-Smaldino2016}{}%
60. Smaldino, P. E. \& McElreath, R. The natural selection of bad science. \emph{Royal Society Open Science} \textbf{3}, 160384 (2016).

\leavevmode\hypertarget{ref-Liddell2018}{}%
61. Liddell, T. M. \& Kruschke, J. K. Analyzing ordinal data with metric models: What could possibly go wrong? \emph{Journal of Experimental Social Psychology} \textbf{79}, 328--348 (2018).

\leavevmode\hypertarget{ref-Nicholson2017}{}%
62. Nicholson, J. S., Deboeck, P. R. \& Howard, W. Attrition in developmental psychology: A review of modern missing data reporting and practices. \emph{International Journal of Behavioral Development} \textbf{41}, 143--153 (2017).

\leavevmode\hypertarget{ref-Tipton2019}{}%
63. Tipton, E., Pustejovsky, J. E. \& Ahmadi, H. Current practices in meta-regression in psychology, education, and medicine. \emph{Research Synthesis Methods} \textbf{10}, 180--194 (2019).

\leavevmode\hypertarget{ref-Chambers2021}{}%
64. Chambers, C. D. \& Tzavella, L. The past, present and future of Registered Reports. \emph{Nature Human Behaviour} \textbf{6}, 29--42 (2021).

\leavevmode\hypertarget{ref-Oishi2013}{}%
65. Oishi, S. Socioecological psychology. \emph{Annual Review of Psychology} \textbf{65}, 581--609 (2014).

\leavevmode\hypertarget{ref-Tiefelsdorf2007}{}%
66. Tiefelsdorf, M. \& Griffith, D. A. Semiparametric filtering of spatial autocorrelation: The eigenvector approach. \emph{Environment and Planning A: Economy and Space} \textbf{39}, 1193--1221 (2007).

\leavevmode\hypertarget{ref-DeHoyos2006}{}%
67. De Hoyos, R. E. \& Sarafidis, V. Testing for cross-sectional dependence in panel-data models. \emph{The Stata Journal} \textbf{6}, 482--496 (2006).

\leavevmode\hypertarget{ref-Dow1984}{}%
68. Dow, M. M., Burton, M. L., White, D. R. \& Reitz, K. P. Galton's Problem as network autocorrelation. \emph{American Ethnologist} \textbf{11}, 754--770 (1984).

\leavevmode\hypertarget{ref-Klerman2011}{}%
69. Klerman, D. M., Mahoney, P. G., Spamann, H. \& Weinstein, M. I. Legal origin or colonial history? \emph{Journal of Legal Analysis} \textbf{3}, 379--409 (2011).

\leavevmode\hypertarget{ref-Duben2021}{}%
70. Düben, C. \emph{conleyreg: Estimations using Conley standard errors}. (2021).

\leavevmode\hypertarget{ref-Solin2020}{}%
71. Solin, A. \& Särkkä, S. Hilbert space methods for reduced-rank Gaussian process regression. \emph{Statistics and Computing} \textbf{30}, 419--446 (2020).

\leavevmode\hypertarget{ref-Landau2021}{}%
72. Landau, W. M. The targets R package: A dynamic Make-like function-oriented pipeline toolkit for reproducibility and high-performance computing. \emph{Journal of Open Source Software} \textbf{6}, 2959 (2021).

\leavevmode\hypertarget{ref-Aust2020}{}%
73. Aust, F. \& Barth, M. \emph{papaja: Create APA manuscripts with R Markdown}. (2020).

\endgroup

\clearpage
\makeatletter
\efloat@restorefloats
\makeatother

\begin{appendix}
\renewcommand{\appendixname}{\bf{Supplementary Material}}
\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\tablename}{Supplementary Table}
\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
\renewcommand{\theequation}{S\arabic{table}} \setcounter{equation}{0}

\hypertarget{section}{%
\section{}\label{section}}

\hypertarget{supplementary-methods}{%
\subsection{Supplementary Methods}\label{supplementary-methods}}

\hypertarget{calculating-global-geographic-and-linguistic-proximity-matrices}{%
\subsubsection{Calculating global geographic and linguistic proximity
matrices}\label{calculating-global-geographic-and-linguistic-proximity-matrices}}

Geographic distance between two nations was calculated as the logged
geodesic distance between country capital cities (data from the R
package \emph{maps}\textsuperscript{1}) using the R package
\emph{geosphere}\textsuperscript{2}. The geographic proximity matrix was
computed as one minus the log geographic distance matrix scaled between
0 and 1.

Linguistic proximity between two nations was calculated as the cultural
proximity between all languages spoken within those nations, weighted by
speaker percentages. We acquired cultural proximity data by combining
the language family trees provided by Glottolog v3.0\textsuperscript{3}
into one global language tree (undated and unresolved). We calculated
cultural proximity \(s\) between two languages \(j\) and \(k\) as the
distance (in number of nodes traversed) of their most recent common
ancestor \(i\) to the root of the tree, through the formula:

\[
s_{jk} = \frac{n_{r}-n_{i}}{n_{r}}
\]

where \(n_{r}\) is the maximum path length (in number of nodes
traversed) leading to the pan-human root \(r\), and \(n_{i}\) is the
maximum path length leading to node \(i\). We then combined these
proximities with speaker data from Ethnologue 21\textsuperscript{4} and
compared every language spoken within those nations by at least 1
permille of the population, weighted by speaker percentages, through the
formula:

\[
w_{lm} = {\Sigma}{\Sigma}p_{lj}p_{mk}s_{jk}
\]

where \(p_{lj}\) is the percentage of the population in nation \(l\)
speaking language \(j\), \(p_{mk}\) is the percentage of the population
in nation \(m\) speaking language \(k\), and \(s_{jk}\) is the proximity
measure between languages \(j\) and \(k\)\textsuperscript{5}. The
resulting linguistic proximity matrix was also scaled between 0 and 1
before analysis.

\hypertarget{bayesian-models-for-reanalysis}{%
\subsubsection{Bayesian models for
reanalysis}\label{bayesian-models-for-reanalysis}}

We provide model formulae for our reanalyses of cross-national
correlations, for a general bivariate case with standardised outcome
\(Y\) and predictor \(X\) variables. In the naive regression model
without controls for non-independence:

\[
\begin{aligned}
Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
\mu_{i} &= \alpha + \beta X_{i} \\
\alpha &\sim \text{Normal}(0, 0.4) \\
\beta &\sim \text{Normal}(0, 0.4) \\
\sigma &\sim \text{Exponential}(5)
\end{aligned}
\] The priors in this model were arrived at by prior predictive checks,
with wider priors making predictions beyond the scale of standardised
outcome variables and narrower priors being too informative.

\newpage

To control for spatial non-independence, we add a Gaussian process to
this model and feed it a scaled geographic distance matrix \(D\) based
on Euclidean distances between latitude and longitude coordinates. This
distance matrix is computed internally by the R package
\emph{brms}\textsuperscript{6}. The Gaussian process uses an
exponentiated quadratic covariance kernel, the only covariance kernel
currently supported by \emph{brms}. The model formula is:

\[
\begin{aligned}
Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
\mu_{i} &= \alpha + \kappa_{\text{NATION}[i]} + \beta X_{i} \\
\begin{pmatrix}
\kappa_{1} \\ \kappa_{2} \\ \kappa_{3} \\ ... \\ \kappa_{n}
\end{pmatrix} &\sim \text{MVNormal}
\begin{pmatrix}
\begin{pmatrix}
0 \\ 0 \\ 0 \\ ... \\ 0
\end{pmatrix},\textbf{K}
\end{pmatrix}\\
\textbf{K}_{ij} &= sdgp^2 \text{exp} \big (-D_{ij}^2 / (2 lscale^2) \big )\\
\alpha &\sim \text{Normal}(0, 0.4) \\
\beta &\sim \text{Normal}(0, 0.4) \\
\sigma &\sim \text{Exponential}(5) \\
sdgp &\sim \text{Exponential}(5) \\
lscale &\sim \text{InverseGamma}(?,?)
\end{aligned}
\] where \(n\) is the number of nations, and \(D^2_{ij}\) reflects the
squared Euclidean distances between latitude and longitude coordinates
for the \(i\)-th and \(j\)-th nations. Notice that the inverse gamma
prior on \(lscale\) is left undetermined. This is because the
\emph{brms} package intelligently tunes the prior for this parameter
based on the covariates of the Gaussian process (see
https://betanalpha.github.io/assets/case\_studies/gp\_part3/part3.html).

\newpage

To control for cultural phylogenetic non-independence, we manually
specify the covariance structure for nation random intercepts using a
pre-computed linguistic proximity matrix \(L\) (see previous section).
The covariance between two nations is assumed to be linearly
proportional to the linguistic proximity between those nations. This
assumption is justified if we assume that cultural traits evolve via
Brownian motion along a language phylogeny. The non-centered
parameterisation of this model is:

\[
\begin{aligned}
Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
\mu_{i} &= \alpha + z_{\text{NATION}[i]}\sigma_{\alpha}L + \beta X_{i} \\
\alpha &\sim \text{Normal}(0, 0.4) \\
\beta &\sim \text{Normal}(0, 0.4) \\
z_{j} &\sim \text{Normal}(0, 1)\\
\sigma_{\alpha} &\sim \text{Exponential}(5) \\
\sigma &\sim \text{Exponential}(5)
\end{aligned}
\] \newpage

Finally, we can control for spatial and cultural phylogenetic
non-independence simultaneously by including both a Gaussian process
over latitude and longitude coordinates \emph{and} nation random
intercepts that covary according to linguistic proximity. The resulting
model is as follows:

\[
\begin{aligned}
Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
\mu_{i} &= \alpha + \kappa_{\text{NATION}[i]} + z_{\text{NATION}[i]}\sigma_{\alpha}L + \beta X_{i} \\
\begin{pmatrix}
\kappa_{1} \\ \kappa_{2} \\ \kappa_{3} \\ ... \\ \kappa_{n}
\end{pmatrix} &\sim \text{MVNormal}
\begin{pmatrix}
\begin{pmatrix}
0 \\ 0 \\ 0 \\ ... \\ 0
\end{pmatrix},\textbf{K}
\end{pmatrix}\\
\textbf{K}_{ij} &= sdgp^2 \text{exp} \big (-D_{ij}^2 / (2 lscale^2) \big )\\
\alpha &\sim \text{Normal}(0, 0.4) \\
\beta &\sim \text{Normal}(0, 0.4) \\
z_{j} &\sim \text{Normal}(0, 1)\\
\sigma_{\alpha} &\sim \text{Exponential}(5) \\
\sigma &\sim \text{Exponential}(5) \\
sdgp &\sim \text{Exponential}(5) \\
lscale &\sim \text{InverseGamma}(?,?)
\end{aligned}
\]

\newpage

\hypertarget{supplementary-results}{%
\subsection{Supplementary Results}\label{supplementary-results}}

\hypertarget{geographic-and-cultural-phylogenetic-signal-estimates-for-human-development-index-and-ingleharts-value-dimensions}{%
\subsubsection{Geographic and cultural phylogenetic signal estimates for
Human Development Index and Inglehart's value
dimensions}\label{geographic-and-cultural-phylogenetic-signal-estimates-for-human-development-index-and-ingleharts-value-dimensions}}

Controlling for shared cultural ancestry, the proportion of
national-level variance explained by spatial proximity was 0.38 for the
Human Development Index (95\% credible interval {[}0.20 0.60{]};
\(\text{BF}_{\neq0}\) \textgreater{} 100), 0.43 for traditional values
(95\% CI {[}0.18 0.72{]}; \(\text{BF}_{\neq0}\) \textgreater{} 100), and
0.18 for survival values (95\% CI {[}0.00 0.44{]}; \(\text{BF}_{\neq0}\)
= 1.45). Controlling for spatial proximity, the proportion of variance
explained by shared cultural ancestry was 0.60 for the Human Development
Index (95\% CI {[}0.39 0.79{]}; \(\text{BF}_{\neq0}\) \textgreater{}
100), 0.55 for traditional values (95\% CI {[}0.26 0.79{]};
\(\text{BF}_{\neq0}\) \textgreater{} 100), and 0.79 for survival values
(95\% CI {[}0.53 0.97{]}; \(\text{BF}_{\neq0}\) \textgreater{} 100).

\newpage

\hypertarget{supplementary-figures}{%
\subsection{Supplementary Figures}\label{supplementary-figures}}

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/dag-1.pdf}
\caption{\label{fig:dag}\emph{A causal directed acyclic graph of spatial and
cultural phylogenetic non-independence in cross-national studies.} We
are interested in estimating the direct effect of national-level
exposure \(X\) on national-level outcome \(Y\). But these variables are
confounded by their common unobserved cause \(U\). \(U\) is a stand-in
for shared environmental, ecological, and geographic causes
(e.g.~climate, biodiversity, physical topography) and cultural and
institutional causes (e.g.~cultural norms, political systems). In this
causal model, we cannot condition on \(U\) directly, as it is
unobserved. However, geographic \(G\) and linguistic \(L\) relationships
between societies influence \(U\), since changing a nation's spatial
distance to or shared cultural ancestry with other nations will change
its environmental and cultural traits. We can thus use \(G\) and \(L\)
to model the covariation between \(X\) and \(Y\) induced by \(U\).
Failing to do this and simply estimating the bivariate correlation
between \(X\) and \(Y\) will produce spurious relationships and
residuals that are spatially and culturally non-independent around the
world (see Figure \ref{fig:spatialExample}).}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotSim1-1.pdf}
\caption{\label{fig:plotSim1}\emph{Distribution of cross-national correlations
from simulation study under moderate spatial autocorrelation.} In these
simulations, the strength of spatial autocorrelation is set to 0.5 for
both outcome and predictor variables. For frequentist regression models,
point ranges represent correlation estimates and 95\% confidence
intervals. For Bayesian Gaussian process regression models, point ranges
represent posterior means and 95\% credible intervals. Correlations are
ordered by effect size independently in each panel. Red point ranges
indicate that the slope is ``significant'' (i.e.~the 95\% confidence /
credible interval excludes zero). Black point ranges indicate that the
slope is ``not significant''.}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotReplications2-1.pdf}
\caption{\label{fig:plotReplications2}\emph{Reanalysis models fitted to raw
data, for economic development (a) and cultural values (b) studies.}
Data points are labelled using ISO 3166-1 alpha-2 letter country codes.
In all reanalyses, outcome and predictor variables are standardised,
making regression slopes comparable to Pearson's correlation
coefficients. Lines and shaded areas represent posterior median
regression lines and 95\% credible intervals. For models with covariates
(Adamzyck and Pitt 2009; Gelfand et al.~2011), marginal effects are
presented holding all covariates at zero or their reference categories.}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotReplications3-1.pdf}
\caption{\label{fig:plotReplications3}\emph{Posterior estimates of Gaussian
process functions mapping spatial autocorrelation onto geographic
distance from our reanalyses of economic development (a) and cultural
values (b) studies.} Estimates are from models additionally controlling
for cultural phylogenetic non-independence. The y-axis represents the
amount of spatial autocorrelation between data points with increasing
distance between those points on the x-axis (logged distance in
kilometres). Lines and shaded areas represent median posterior spatial
autocorrelation functions and 50\% and 95\% credible intervals.}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotReplications4-1.pdf}
\caption{\label{fig:plotReplications4}\emph{Posterior estimates of cultural
phylogenetic signal from our reanalyses.} Estimates are from models
additionally controlling for spatial non-independence. Cultural
phylogenetic signal is operationalised as the proportion of
national-level variance explained by linguistic proximity between
nations. Ridges are full posterior distributions, and point ranges are
posterior medians and 95\% credible intervals.}
\end{figure}

\newpage

\hypertarget{supplementary-references}{%
\subsection*{Supplementary References}\label{supplementary-references}}
\addcontentsline{toc}{subsection}{Supplementary References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Brownrigg2018}{}%
1. Brownrigg, R. \emph{maps: Draw geographical maps}. (2018).

\leavevmode\hypertarget{ref-Hijmans2019}{}%
2. Hijmans, R. J. \emph{geosphere: Spherical trigonometry}. (2019).

\leavevmode\hypertarget{ref-Glottolog}{}%
3. Hammarström, H., Forkel, R., Haspelmath, M. \& Bank, S.
\emph{Glottolog 3.0}. (Max Planck Institute for the Science of Human
History, 2017).
doi:\href{https://doi.org/10.5281/zenodo.4061162}{10.5281/zenodo.4061162}.

\leavevmode\hypertarget{ref-Ethnologue}{}%
4. \emph{Ethnologue: Languages of the world}. (SIL International, 2018).

\leavevmode\hypertarget{ref-Eff2008}{}%
5. Eff, E. A. Weight matrices for cultural proximity: Deriving weights
from a language phylogeny. \emph{Structure and Dynamics} \textbf{3},
(2008).

\leavevmode\hypertarget{ref-Burkner2017}{}%
6. Bürkner, P.-C. brms: An R package for Bayesian multilevel models
using Stan. \emph{Journal of Statistical Software} \textbf{80}, 1--28
(2017).
\end{appendix}

\end{document}

### 2022-03-09 R/functionsExamplePlot.R

@@ -1,0 +1,76 @@
+# plot showing spatial autocorrelation example on map
+plotSpatialExample <- function(data) {
+  # sample countries
+  d <-
+    data %>%
+    filter(country %in% c("GB","ES","RU","IT","FR",
+                          "DE","CA","US","MX","BR",
+                          "AR","CO","PE","CN","IN",
+                          "UA","JP","DZ","NE","LY",
+                          "AU","NZ","TH","PH","VN")) %>%
+    mutate(Residual = residuals(lm(y ~ x, .)),
+           line = y - Residual)
+  # top plot
+  pA <-
+    ggplot(d) +
+    geom_segment(aes(x = x, xend = x, y = y, yend = line, colour = Residual)) +
+    geom_point(aes(x = x, y = y)) +
+    geom_text_repel(aes(x = x, y = y, label = country), 
+                    segment.colour = NA, size = 3.5, force = 1.5,
+                    nudge_x = ifelse(d$country == "AU", -0.2, ifelse(d$country %in% c("GB","RU"), 0.1, 0)),
+                    nudge_y = ifelse(d$country == "ES", 0.2, ifelse(d$country == "JP", -0.1, 0))) +
+    geom_smooth(aes(x = x, y = y), method = "lm", 
+                formula = y ~ x, se = FALSE, colour = "black") +
+    scale_colour_gradient2(high = "red", low = "blue") +
+    labs(y = "National-level outcome variable",
+         x = "National-level predictor variable") +
+    theme_classic() +
+    theme(legend.position = "none")
+  # bottom plot
+  d$region <- countrycode(d$country, "iso2c", "country.name")
+  d$region <- ifelse(d$region == "United States", "USA", d$region)
+  d$region <- ifelse(d$region == "United Kingdom", "UK", d$region)
+  world <- 
+    map_data("world") %>% 
+    filter(region != "Antarctica") %>%
+    left_join(select(d, region, Residual), by = "region")
+  pB <-
+    ggplot() +
+    geom_map(
+      data = world, map = world,
+      aes(long, lat, map_id = region),
+      fill = "lightgrey"
+    ) +
+    geom_map(
+      data = world[!is.na(world$Residual),], 
+      map = world[!is.na(world$Residual),],
+      aes(long, lat, map_id = region, fill = Residual)
+    ) +
+    scale_fill_gradient2(high = "red", low = "blue") +
+    theme_void()
[…]

### 2022-03-09 R/functionsReplications.R

@@ -1,0 +1,1067 @@
+# custom functions - replications
+
+# load data for Adamczyk 2009
+loadDataAdamczyk2009 <- function(wvs, iso) {
+  wvs %>%
+    # wvs only
+    filter(studyno == 4001) %>%
+    # wave 4
+    filter(S002 == 4) %>%
+    # keep relevant variables
+    select(
+      # country code
+      S009,
+      # survival vs. self-expression values
+      Y002, A008, E025, A165, # does not include F118 as it is dv
+      # disapproval of homosexuality (dependent variable)
+      F118,
+      # covariates
+      A006, # importance of religion
+      F025, # religious denomination
+      X007, # marital status
+      X011, # number of children
+      X025, # education
+      X001, # gender
+      C006, # satisfaction with financial situation
+      X002  # year of birth
+    ) %>%
+    # drop missings
+    drop_na() %>%
+    mutate(
+      # create survival vs. self-expression values composite
+      surv = as.numeric(scale(4-Y002) + scale(A008) + scale(E025) + scale(A165)),
+      # reverse code dependent variable
+      F118 = 11 - F118,
+      # reverse code religion importance
+      A006 = 5 - A006,
+      # 1 = married, 0 = all else
+      X007 = ifelse(X007 == 1, "Married", "Not married"),
+      # birth cohorts
+      X002 = ifelse(X002 %in% 1920:1929, "1920s",
+                    ifelse(X002 %in% 1930:1939, "1930s",
+                           ifelse(X002 %in% 1940:1949, "1940s",
+                                  ifelse(X002 %in% 1950:1959, "1950s",
+                                         ifelse(X002 %in% 1960:1969, "1960s",
+                                                ifelse(X002 %in% 1970:1979, "1970s", "1980s")))))),
+      # principal religious traditions for countries
+      countryReligTrad = ifelse(S009 %in% c("CA","PR","MX","VE","ES",
+                                            "AR","PE","CL","UG","PH"), "Catholic",
+                                ifelse(S009 %in% c("IN"), "Hindu",
+                                       ifelse(S009 %in% c("MD","RS","ME","MK"), "Orthodox Christian",
[…]

### 2022-03-09 R/functionsReview.R

@@ -1,0 +1,192 @@
+# custom functions - review
+
+# load main review
+loadReview <- function(fileReview, fileMethods) {
+  read_excel(fileReview, sheet = "Review") %>%
+    left_join(read_csv(fileMethods), by = "MethodNI")
+}
+
+# fit random effects model to retrieve corrected proportions
+fitReviewMLM <- function(review, outcome) {
+  # fill NAs (analyses that did not control for NI at all) with zeroes
+  review <- 
+    review %>%
+    mutate_at(vars(starts_with("method_")), function(x) ifelse(is.na(x), 0, x))
+  # fit model
+  brm(
+    formula = bf(paste0(outcome, " ~ 0 + Review + (1 | Review:Citation)")),
+    data = review,
+    family = bernoulli,
+    prior = c(prior(normal(0, 1), class = b),
+              prior(exponential(1), class = sd)),
+    cores = 4, seed = 2113
+  )
+}
+
+# spline model
+fitSpline <- function(review) {
+  # prepare data
+  review <-
+    review %>%
+    mutate(
+      # standardise year
+      Year.std = as.numeric(scale(Year)),
+      # as factor
+      Review = factor(Review)
+    )
+  # fit spline model
+  brm(ControlNI ~ 0 + Intercept + s(Year.std, by = Review) + (1 | Review:Citation),
+      data = review, family = bernoulli,
+      prior = c(prior(normal(0, 0.5), class = b),
+                prior(exponential(2), class = sd),
+                prior(exponential(2), class = sds)),
+      cores = 4, seed = 2113, iter = 3000,
+      control = list(adapt_delta = 0.99))
+}
+
+# function to get bootstrapped confidence intervals for proportion
+getBootCIProp <- function(prop, n, nboot) {
+  set.seed(1)
+  p <- c()
+  for (i in 1:nboot) {
+    p[i] <- mean(rbinom(n, 1, prop))
+  }
+  return(quantile(p, c(0.025, 0.975)))
+}
+
+# summary of review results
+plotReviewSummary <- function(review, spline, postRM1, postRM2, 
+                              postRM3, postRM4, postRM5) {
+  # prepare data
+  review <-
+    review %>%
+    mutate_at(vars(starts_with("method_")), function(x) ifelse(is.na(x), 0, x)) %>%
+    mutate(Review = ifelse(Review == "Economic development", "Economic\ndevelopment", "Cultural\nvalues"),
[…]

### 2022-03-09 R/functionsSignal.R

@@ -1,0 +1,163 @@
+# custom functions - signal
+
+# load covariance matrices
+loadCovMat <- function(file, log) {
+  # load distance matrix
+  out <- 
+    read_excel(file, na = "") %>%
+    select(-ISO) %>%
+    as.matrix()
+  rownames(out) <- colnames(out)
+  # log distances?
+  if (log) out <- log(out)
+  # distances between 0 and 1
+  out <- out / max(out)
+  diag(out) <- 0
+  # 1 - distance = proximity (covariance)
+  out <- 1 - out
+  return(out)
+}
+
+# load HDI data
+loadHDIData <- function(fileHDI, fileISOHDI) {
+  # tidy up messy excel sheet
+  read_xlsx(
+    path = fileHDI,
+    sheet = "Table 2",
+    skip = 4
+  ) %>%
+    drop_na("HDI rank") %>%
+    select(c("Country", "1990", "2000", "2010", 
+             "2014", "2015", "2017", "2018", "2019")) %>%
+    # get data in long form
+    pivot_longer(cols = !Country, names_to = "year", values_to = "hdi") %>%
+    filter(hdi != "..") %>%
+    mutate(year = as.numeric(year),
+           hdi = round(as.numeric(hdi), 3)) %>%
+    # add iso-codes
+    left_join(read.csv(fileISOHDI), by = "Country") %>%
+    # complete missings
+    complete(year, nesting(iso, Country)) %>%
+    # fix cote d'ivoire error
+    mutate(iso = ifelse(is.na(iso), "CI", iso))
+}
+
+# signal for hdi
+fitHDISignal <- function(hdi, geoCov, linCov) {
+  # add iso variables for modelling
+  hdi$isoGeo <- hdi$iso
+  hdi$isoLin <- hdi$iso
+  # subset matrices
+  geoCov <- geoCov[rownames(geoCov)[rownames(geoCov) %in% hdi$iso],
+                   colnames(geoCov)[colnames(geoCov) %in% hdi$iso]]
+  linCov <- linCov[rownames(linCov)[rownames(linCov) %in% hdi$iso],
+                   colnames(linCov)[colnames(linCov) %in% hdi$iso]]
+  # fit model
+  brm(hdi ~ 0 + Intercept + (1 | gr(isoGeo, cov = geoCov)) + (1 | gr(isoLin, cov = linCov)) + (1 | iso),
+      data = hdi, data2 = list(geoCov = geoCov, linCov = linCov),
[…]

### 2022-03-09 R/functionsSimulation.R

@@ -1,0 +1,234 @@
+# custom functions - simulation
+
+# load isocode data
+loadISO <- function(fileISO) {
+  read.csv(fileISO) %>%
+    mutate(iso3 = str_remove(Alpha.3.code, " "),
+           iso2 = str_remove(Alpha.2.code, " ")) %>%
+    distinct(iso2, iso3, Latitude..average., Longitude..average.)
+}
+
+# load geographic distance matrix
+loadDistanceMatrix <- function(file, continent, iso, legal) {
+  # load distance matrix
+  out <- 
+    read_excel(file, na = "") %>%
+    select(-ISO) %>%
+    as.matrix()
+  # row and column names identical
+  rownames(out) <- colnames(out)
+  # log
+  out <- log(out)
+  # diagonal = 0
+  diag(out) <- 0
+  # keep only countries with longitude, latitude, continent, and legal origin data
+  legalOriginISO <- iso$iso2[iso$iso3 %in% as.character(legal$countrycode)]
+  out <- out[rownames(out)[rownames(out) %in% iso$iso2],
+             rownames(out)[rownames(out) %in% iso$iso2]]
+  out <- out[rownames(out)[rownames(out) %in% continent$country],
+             rownames(out)[rownames(out) %in% continent$country]]
+  out <- out[rownames(out)[rownames(out) %in% legalOriginISO],
+             rownames(out)[rownames(out) %in% legalOriginISO]]
+  # scale matrix
+  out <- out / max(out)
+  return(out)
+}
+
+# simulate data
+simulateData <- function(distGeo, continent, iso, legal, seed, lambda, rho) {
+  # set seed
+  set.seed(seed)
+  # parameters
+  ### lambda = amount of autocorrelation for dependent variable
+  ### rho = amount of autocorrelation for independent variable
+  # simulate x and y, modified from https://www.jstor.org/stable/644404 (p. 761)
+  n <- nrow(distGeo)
+  x <- ((1 - lambda*distGeo)^-1) %*% rnorm(n)
+  y <- ((1 -    rho*distGeo)^-1) %*% rnorm(n)
+  # add some random noise
+  x <- x + rnorm(n)
+  y <- y + rnorm(n)
+  # standardise x and y
+  x <- as.numeric(scale(x))
+  y <- as.numeric(scale(y))
+  # produce data frame and match longitude, latitude, continent, and legal origin data
+  out <- 
[…]

### 2022-03-09 R/scripts/simulateDAG.R

@@ -1,0 +1,65 @@
+library(brms)
+library(tidyverse)
+
+# simulate DAG to show that method works
+
+# DAG:
+# X -> Y
+# U -> X
+# U -> X
+# D -> U
+# where X = exposure, Y = outcome, U = unobserved confound
+# and D is a distance matrix
+
+# set seed
+set.seed(1)
+
+# sample size
+n <- 20^2
+
+# coordinates
+lon <- rep(seq(0, 1, length.out = sqrt(n)), each = sqrt(n))
+lat <- rep(seq(0, 1, length.out = sqrt(n)), times = sqrt(n))
+
+# distance matrix
+dist <- as.matrix(dist(data.frame(lon = lon, lat = lat)))
+dist <- dist / max(dist)
+
+# strength of autocorrelation
+lambda <- 0.95
+
+# generate spatially autocorrelated unobserved confound
+U <- ((1 - lambda*dist)^-1 %*% rnorm(n)) + rnorm(n)
+U <- as.numeric(scale(U)) # standardise
+
+# U positively predicts X and negatively predicts Y
+X <- as.numeric(scale(rnorm(n, 1*U)))
+Y <- as.numeric(scale(rnorm(n, -1*U)))
+
+# creates spurious bivariate relationship between X and Y
+m1 <- lm(Y ~ X)
+summary(m1)
+
+# residuals are spatially autocorrelated
+r1 <- resid(m1)
+tibble(lat, lon, r1) %>%
+  ggplot(aes(lat, lon, colour = r1)) +
+  geom_point(size = 5) +
+  scale_color_gradient2(high = "red", low = "blue") +
+  theme_classic()
+
+# control for confound U removes Y ~ X relationship
+summary(lm(Y ~ X + U))
+
+# but we can't control for U as it is unobserved!
+# instead, control for spatial autocorrelation
+# with approximate gaussian process
+m2 <- brm(Y ~ X + gp(lat, lon, k = 5, c = 5/4),
+          data = tibble(Y, X, lat, lon),
+          control = list(adapt_delta = 0.99, max_treedepth = 15),
+          cores = 4, seed = 1)
+summary(m2)
+
+# removes association between X and Y by accounting for
+# the covariation between X and Y induced by the
+# spatially autocorrelated confound U

### 2022-03-09 _targets.R

@@ -1,0 +1,404 @@
+library(targets)
+library(tarchetypes)
+library(tidyverse)
+source("R/functionsExamplePlot.R")
+source("R/functionsSignal.R")
+source("R/functionsReview.R")
+source("R/functionsSimulation.R")
+source("R/functionsReplications.R")
+options(tidyverse.quiet = TRUE)
+tar_option_set(packages = c("brms", "cowplot", "conleyreg", "countrycode", 
+                            "dagitty", "geosphere", "ggdag", "ggrepel", "ggtext", 
+                            "haven", "lmtest", "papaja", "psych", "readxl", 
+                            "rstan", "sjlabelled", "tidybayes", "tidyverse"))
+
+# pipeline
+list(
+  
+  #### Visual example of spatial autocorrelation ####
+  
+  tar_target(spatialExample, plotSpatialExample(simData_0.9_0.9[[56]])),
+  
+  #### Causal model ####
+  
+  tar_target(dag, plotDAG()),
+  
+  #### Geographic and linguistic signal ####
+  
+  # files
+  tar_target(fileGeo, "data/networks/1F Population Distance.xlsx", format = "file"),
+  tar_target(fileLin, "data/networks/2F Country Distance 1pml adj.xlsx", format = "file"),
+  tar_target(fileHDI, "data/hdi/2020_statistical_annex_all.xlsx", format = "file"),
+  tar_target(fileISOHDI, "data/hdi/iso.csv", format = "file"),
+  tar_target(fileWVS, "data/wvs/Integrated_values_surveys_1981-2021.sav", format = "file"),
+  # hdi data
+  tar_target(hdi, loadHDIData(fileHDI, fileISOHDI)),
+  # wvs data
+  tar_target(wvs, read_sav(fileWVS)),
+  # covariance matrices
+  tar_target(geoCov, loadCovMat(fileGeo, log = TRUE)),
+  tar_target(linCov, loadCovMat(fileLin, log = FALSE)),
+  # geographic and linguistic signal
+  tar_target(hdiSignal, fitHDISignal(hdi, geoCov, linCov)),
+  tar_target(tradSignal, fitWVSSignal(wvs, outcome = "trad", geoCov, linCov)),
+  tar_target(survSignal, fitWVSSignal(wvs, outcome = "surv", geoCov, linCov)),
+  # posterior samples
[…]

### 2022-03-09 review/wosExports/econDev/wosEconDev.R

@@ -1,0 +1,17 @@
+# paste together wos export files - economic development
+
+library(readxl)
+library(tidyverse)
+
+read_xlsx("savedrecs1.xlsx", col_types = "text") %>%
+  bind_rows(read_xlsx("savedrecs2.xlsx", col_types = "text")) %>%
+  # keep only journal articles
+  filter(PT == "J") %>%
+  # calculate citations per year published
+  mutate(citesPerYear = parse_number(Z9) / (2021 - parse_number(PY))) %>%
+  # arrange by citations per year published
+  arrange(desc(citesPerYear)) %>%
+  # keep relevant columns
+  select(PY, AU, TI, SO, VL, IS, DI, Z9, citesPerYear) %>%
+  # write
+  write_excel_csv(path = "wosEconDev.csv")

### 2022-03-09 review/wosExports/values/wosValues.R

@@ -1,0 +1,23 @@
+# paste together wos export files - values
+
+library(readxl)
+library(tidyverse)
+
+read_xlsx("savedrecs1.xlsx", col_types = "text") %>%
+  bind_rows(read_xlsx("savedrecs2.xlsx", col_types = "text")) %>%
+  bind_rows(read_xlsx("savedrecs3.xlsx", col_types = "text")) %>%
+  bind_rows(read_xlsx("savedrecs4.xlsx", col_types = "text")) %>%
+  bind_rows(read_xlsx("savedrecs5.xlsx", col_types = "text")) %>%
+  bind_rows(read_xlsx("savedrecs6.xlsx", col_types = "text")) %>%
+  bind_rows(read_xlsx("savedrecs7.xlsx", col_types = "text")) %>%
+  bind_rows(read_xlsx("savedrecs8.xlsx", col_types = "text")) %>%
+  # keep only journal articles
+  filter(PT == "J") %>%
+  # calculate citations per year published
+  mutate(citesPerYear = parse_number(Z9) / (2021 - parse_number(PY))) %>%
+  # arrange by citations per year published
+  arrange(desc(citesPerYear)) %>%
+  # keep relevant columns
+  select(PY, AU, TI, SO, VL, IS, DI, Z9, citesPerYear) %>%
+  # write
+  write_excel_csv(path = "wosValues.csv")

### 2022-03-14 manuscript.Rmd
add language family data file

@@
 ---
-title             : "Cross-national studies have not sufficiently accounted for non-independence between nations"
+title             : "Why non-independence matters for cross-national correlations"
 shorttitle        : "Non-independence between nations"
@@
 abstract: |
- Cross-national analyses can illuminate patternings of national-level outcomes around the world.
-But since nations are related via spatial proximity and shared cultural ancestry, cross-national analyses often violate statistical assumptions of non-independence, inflating false positive rates.
-Here, we show that past work in economics and psychology has not sufficiently accounted for non-independence between nations.
-In a review of the 100 highest-cited cross-national studies of economic development and cultural values, we find that controls for non-independence are rare and have not increased in frequency over the past 30 years.
-When studies do attempt to control for non-independence, our simulations suggest that the methods used continue to produce false positives.
-In replications of a subsample of twelve cross-national relationships, we show that half fail to replicate when controlling for spatial and/or cultural non-independence using global proximity matrices.
-We urge social scientists to sufficiently control for non-independence when reporting cross-national associations.
+ Cross-national analyses can illuminate global patternings of national-level outcomes.
+But since nations are connected in various ways, such as via spatial proximity and shared cultural ancestry, cross-national analyses often violate assumptions of non-independence, inflating false positive rates.
+Here, we show that, despite being recognised as an important statistical pitfall for over 200 years, cross-national research in economics and psychology still does not sufficiently account for non-independence.
+In a review of the 100 highest-cited cross-national studies of economic development and values, we find that controls for non-independence are rare and have remained so over the past 30 years.
+When studies do include controls for non-independence, our simulations suggest that commonly used methods continue to produce false positives.
+In reanalyses of twelve cross-national relationships, we show that half are no longer significant after controlling for non-independence using global proximity matrices.
+We urge social scientists to sufficiently control for non-independence in cross-national research.
 
@@
 Given the importance of nations for structuring human behaviour, there is justifiably huge interest in statistical analyses that attempt to predict variation in national-level outcomes of all kinds.
-At the time of writing, a search in Google Scholar for the term "cross-national study" returned 180,000 hits.
+At the time of writing, a search in Web of Science for the terms "cross-national", "cross-cultural", or "cross-country" in titles and abstracts returned approximately 80,000 unique hits.
 The standard practice for cross-national analyses is to conduct bivariate correlations or multiple regressions with individual data points representing different nations.
@@
 First, nations that are closer to one another tend to be more similar than distant nations.
-This phenomenon is known as spatial non-independence[@Kissling2008], and it occurs because nations in close spatial proximity share characteristics due to local cultural diffusion of ideas[@Hewlett2002] and common climates and environments[@Kissling2008].
+This phenomenon is known as spatial non-independence[@Kissling2008], and it occurs because nations in close spatial proximity share characteristics due to local cultural diffusion of ideas[@Hewlett2002] and regional variation in climate and environment[@Kissling2008].
 For example, the neighbouring African nations Zambia and Tanzania have similar levels of terrain ruggedness, which has been used to partially explain their similar levels of economic development[@Nunn2012].
@@
 Shared cultural ancestry can result in a form of pseudoreplication, whereby multiple instances of the same trait across nations are merely duplicates of the ancestral original.
-For example, the related island nations New Zealand and Samoa express similar languages and customs due to cultural inheritance from a shared ancestor deep in the evolutionary history of Polynesian societies.
-More recently, shared ancestry is also crucial in explaining cultural similarities between colonial settlements and their colonisers, even those separated by large geographic distances (e.g.
+For example, despite being separated by over 1500 km of ocean, the related island nations Tonga and Tuvalu express similar languages and customs due to cultural inheritance from a shared ancestor deep in the evolutionary history of Polynesian societies.
+More recently, shared ancestry is also crucial in explaining cultural similarities between colonial settlements and their colonisers (e.g.
 Spain and Argentina).
@@
 A fundamental assumption of regression analysis is that model residuals should be independently and identically distributed[@Jarque1987].
-But without accounting for spatial or cultural non-independence between nations, model residuals can show structure that remains unaccounted for, violating this assumption (Figure \@ref(fig:spatialExample)).
+But without accounting for spatial or cultural non-independence between nations, model residuals can show structure that remains unaccounted for, violating this assumption.
 Treating nations as independent can thus inflate false positive rates[@Legendre1993], producing spurious "direct" relationships between variables that in fact only indirectly covary due to spatial or cultural connections (see Supplementary Figure \@ref(fig:dag) for an example causal model).
-
-(ref:spatialExampleCaption) _An example of spatial non-independence in cross-national analysis._ National-level outcome and predictor variables for 25 nations are simulated with high levels of spatial non-independence.
-(top) Scatterplot visualising the data with fitted line from naive regression without accounting for spatial non-independence.
-Vertical line segments represent model residual errors, coloured by their sign and size.
-Countries are labelled with ISO 3166-1 alpha-2 letter codes.
-(bottom) Model residual errors plotted onto a world map.
-In this example, residuals clearly show spatial structure that remains unaccounted for.
-
-```{r spatialExample, fig.width=6.5, fig.height=6, fig.cap="(ref:spatialExampleCaption)"}
-tar_read(spatialExample)
-```
 
@@
 For all three of these variables, we found that a substantial proportion of national-level variation was explained by spatial proximity and shared cultural ancestry between nations (Figure \@ref(fig:plotSignal)).
+In our models, shared cultural ancestry explained over half of the national-level variation in economic development and cultural values, with spatial proximity explaining most of the remaining national-level variance.
 Bayes Factors indicated strong evidence that these estimates of geographic and cultural phylogenetic signal differed from zero, aside from the geographic signal for survival values, for which the evidence was equivocal (see Supplementary Results).

### 2022-03-14 R/functionsExamplePlot.R

@@ -1,76 +1,0 @@
-# plot showing spatial autocorrelation example on map
-plotSpatialExample <- function(data) {
-  # sample countries
-  d <-
-    data %>%
-    filter(country %in% c("GB","ES","RU","IT","FR",
-                          "DE","CA","US","MX","BR",
-                          "AR","CO","PE","CN","IN",
-                          "UA","JP","DZ","NE","LY",
-                          "AU","NZ","TH","PH","VN")) %>%
-    mutate(Residual = residuals(lm(y ~ x, .)),
-           line = y - Residual)
-  # top plot
-  pA <-
-    ggplot(d) +
-    geom_segment(aes(x = x, xend = x, y = y, yend = line, colour = Residual)) +
-    geom_point(aes(x = x, y = y)) +
-    geom_text_repel(aes(x = x, y = y, label = country), 
-                    segment.colour = NA, size = 3.5, force = 1.5,
-                    nudge_x = ifelse(d$country == "AU", -0.2, ifelse(d$country %in% c("GB","RU"), 0.1, 0)),
-                    nudge_y = ifelse(d$country == "ES", 0.2, ifelse(d$country == "JP", -0.1, 0))) +
-    geom_smooth(aes(x = x, y = y), method = "lm", 
-                formula = y ~ x, se = FALSE, colour = "black") +
-    scale_colour_gradient2(high = "red", low = "blue") +
-    labs(y = "National-level outcome variable",
-         x = "National-level predictor variable") +
-    theme_classic() +
-    theme(legend.position = "none")
-  # bottom plot
-  d$region <- countrycode(d$country, "iso2c", "country.name")
-  d$region <- ifelse(d$region == "United States", "USA", d$region)
-  d$region <- ifelse(d$region == "United Kingdom", "UK", d$region)
-  world <- 
-    map_data("world") %>% 
-    filter(region != "Antarctica") %>%
-    left_join(select(d, region, Residual), by = "region")
-  pB <-
-    ggplot() +
-    geom_map(
-      data = world, map = world,
-      aes(long, lat, map_id = region),
-      fill = "lightgrey"
-    ) +
-    geom_map(
-      data = world[!is.na(world$Residual),], 
-      map = world[!is.na(world$Residual),],
-      aes(long, lat, map_id = region, fill = Residual)
-    ) +
-    scale_fill_gradient2(high = "red", low = "blue") +
-    theme_void()
[…]

### 2022-03-14 _targets.R

@@ -1,7 +1,7 @@
 library(targets)
 library(tarchetypes)
 library(tidyverse)
-source("R/functionsExamplePlot.R")
+source("R/functionsDAG.R")
 source("R/functionsSignal.R")
 source("R/functionsReview.R")
 source("R/functionsSimulation.R")
@@ -14,10 +14,6 @@ tar_option_set(packages = c("brms", "cowplot", "conleyreg", "countrycode",
 
 # pipeline
 list(
-  
-  #### Visual example of spatial autocorrelation ####
-  
-  tar_target(spatialExample, plotSpatialExample(simData_0.9_0.9[[56]])),
   
   #### Causal model ####
   
@@ -85,12 +81,12 @@ list(
   
   # files
   tar_target(fileContinent, "data/countryData/continent.csv", format = "file"),
-  tar_target(fileLegal, "data/countryData/Klerman_etal_LO_v_CO.dta", format = "file"),
+  tar_target(fileLangFam, "data/countryData/langFamily.xlsx", format = "file"),
   tar_target(fileISO, "data/countryData/countries_codes_and_coordinates.csv", format = "file"),
   # continent data (https://datahub.io/JohnSnowLabs/country-and-continent-codes-list#data)
   tar_target(continent, read.csv(fileContinent, na.strings = "")),
-  # legal origins data (https://academic.oup.com/jla/article/3/2/379/899816#supplementary-data)
-  tar_target(legal, read_dta(fileLegal)),
+  # language family data (from glottolog)
+  tar_target(legal, read_xlsx(fileLegal)),
   # isocodes (https://gist.github.com/tadast/8827699)
   tar_target(iso, loadISO(fileISO)),
   # geographic distance matrix

### 2022-03-14 R/functionsSimulation.R
new covariance matrix creation

@@ -8,29 +8,28 @@ loadISO <- function(fileISO) {
     distinct(iso2, iso3, Latitude..average., Longitude..average.)
 }
 
-# load geographic distance matrix
-loadDistanceMatrix <- function(file, continent, iso, legal) {
+# load geographic and linguistic covariance matrices
+loadSimCovMat <- function(file, log, continent, iso, langFam) {
   # load distance matrix
   out <- 
     read_excel(file, na = "") %>%
     select(-ISO) %>%
     as.matrix()
-  # row and column names identical
   rownames(out) <- colnames(out)
-  # log
-  out <- log(out)
-  # diagonal = 0
-  diag(out) <- 0
-  # keep only countries with longitude, latitude, continent, and legal origin data
-  legalOriginISO <- iso$iso2[iso$iso3 %in% as.character(legal$countrycode)]
+  # keep only countries with longitude, latitude, continent, and language family data
   out <- out[rownames(out)[rownames(out) %in% iso$iso2],
              rownames(out)[rownames(out) %in% iso$iso2]]
   out <- out[rownames(out)[rownames(out) %in% continent$country],
              rownames(out)[rownames(out) %in% continent$country]]
-  out <- out[rownames(out)[rownames(out) %in% legalOriginISO],
-             rownames(out)[rownames(out) %in% legalOriginISO]]
-  # scale matrix
+  out <- out[rownames(out)[rownames(out) %in% langFam$iso],
+             rownames(out)[rownames(out) %in% langFam$iso]]
+  # log distances?
+  if (log) out <- log(out)
+  # distances between 0 and 1
   out <- out / max(out)
+  diag(out) <- 0
+  # 1 - distance = proximity (covariance)
+  out <- 1 - out
   return(out)
 }
 

### 2022-03-14 _targets.R

@@ -86,11 +86,12 @@ list(
   # continent data (https://datahub.io/JohnSnowLabs/country-and-continent-codes-list#data)
   tar_target(continent, read.csv(fileContinent, na.strings = "")),
   # language family data (from glottolog)
-  tar_target(legal, read_xlsx(fileLegal)),
+  tar_target(langFam, read_xlsx(fileLangFam)),
   # isocodes (https://gist.github.com/tadast/8827699)
   tar_target(iso, loadISO(fileISO)),
-  # geographic distance matrix
-  tar_target(distGeo, loadDistanceMatrix(fileGeo, continent, iso, legal)),
+  # geographic and linguistic covariance matrices
+  tar_target(simGeoCov, loadSimCovMat(fileGeo, log = TRUE , continent, iso, langFam)),
+  tar_target(simLinCov, loadSimCovMat(fileLin, log = FALSE, continent, iso, langFam)),
   # map simulation over different values of lambda and rho
   tar_map(
     values = expand_grid(lambda = c(0.1, 0.5, 0.9), rho = c(0.1, 0.5, 0.9)),

### 2022-03-14 R/functionsSimulation.R
edit simulation to include linguistic covariance

@@ -34,30 +34,31 @@ loadSimCovMat <- function(file, log, continent, iso, langFam) {
 }
 
 # simulate data
-simulateData <- function(distGeo, continent, iso, legal, seed, lambda, rho) {
+simulateData <- function(covMat, continent, iso, langFam, seed, lambda, rho) {
   # set seed
   set.seed(seed)
   # parameters
-  ### lambda = amount of autocorrelation for dependent variable
-  ### rho = amount of autocorrelation for independent variable
-  # simulate x and y, modified from https://www.jstor.org/stable/644404 (p. 761)
-  n <- nrow(distGeo)
-  x <- ((1 - lambda*distGeo)^-1) %*% rnorm(n)
-  y <- ((1 -    rho*distGeo)^-1) %*% rnorm(n)
-  # add some random noise
-  x <- x + rnorm(n)
-  y <- y + rnorm(n)
+  ### lambda = amount of autocorrelation ("signal") for dependent variable
+  ### rho = amount of autocorrelation ("signal") for independent variable
+  # simulate x and y
+  n <- nrow(covMat)
+  x <- mvrnorm(1, rep(0, n), (rho    / (1 - rho   )) * covMat)
+  y <- mvrnorm(1, rep(0, n), (lambda / (1 - lambda)) * covMat)
+  # add error
+  x <- x + mvrnorm(1, rep(0, n), 1 * diag(n))
+  y <- y + mvrnorm(1, rep(0, n), 1 * diag(n))
   # standardise x and y
   x <- as.numeric(scale(x))
   y <- as.numeric(scale(y))
   # produce data frame and match longitude, latitude, continent, and legal origin data
   out <- 
-    tibble(x = x, y = y, country = rownames(distGeo), seed = seed) %>%
+    tibble(x = x, y = y, country = rownames(covMat), seed = seed) %>%
     left_join(iso, by = c("country" = "iso2")) %>%
     left_join(distinct(continent, country, .keep_all = TRUE), by = "country") %>%
-    left_join(select(legal, countrycode, LO), by = c("iso3" = "countrycode")) %>%
+    left_join(dplyr::select(langFam, iso, langFamily), by = c("country" = "iso")) %>%
     mutate(continent = factor(continent),
-           LO = factor(get_labels(LO)[LO])) %>%
+           langFamily = factor(langFamily),
[…]

### 2022-03-14 _targets.R

@@ -9,8 +9,37 @@ source("R/functionsReplications.R")
 options(tidyverse.quiet = TRUE)
 tar_option_set(packages = c("brms", "cowplot", "conleyreg", "countrycode", 
                             "dagitty", "geosphere", "ggdag", "ggrepel", "ggtext", 
-                            "haven", "lmtest", "papaja", "psych", "readxl", 
+                            "haven", "lmtest", "MASS", "papaja", "psych", "readxl", 
                             "rstan", "sjlabelled", "tidybayes", "tidyverse"))
+
+# targets for simulation (see below)
+simulationTargets <-
+  tar_map(
+    # return nested list from tar_map()
+    unlist = FALSE,
+    # map simulation over different covariance matrices and different values of lambda and rho
+    values = expand_grid(covMat = rlang::syms(c("simGeoCov", "simLinCov")),
+                         lambda = c(0.2, 0.5, 0.8), rho = c(0.2, 0.5, 0.8)),
+    # simulation seed
+    tar_target(seed, 1:5),
+    # simulate data
+    tar_target(simData, simulateData(covMat = covMat, continent, iso, langFam, 
+                                     seed = seed, lambda = lambda, rho = rho), 
+               pattern = map(seed), iteration = "list"),
+    # ols analyses
+    tar_target(olsModel1, fitOLSModel(y ~ x,              data = simData), pattern = map(simData)),
+    tar_target(olsModel2, fitOLSModel(y ~ x + latitude,   data = simData), pattern = map(simData)),
+    tar_target(olsModel3, fitOLSModel(y ~ x + longitude,  data = simData), pattern = map(simData)),
+    tar_target(olsModel4, fitOLSModel(y ~ x + continent,  data = simData), pattern = map(simData)),
+    tar_target(olsModel5, fitOLSModel(y ~ x + langFamily, data = simData), pattern = map(simData)),
+    # conley se analyses
+    tar_target(conleyModel1, fitConleyModel(simData, dist_cutoff = 100), pattern = map(simData)),
+    tar_target(conleyModel2, fitConleyModel(simData, dist_cutoff = 1000), pattern = map(simData)),
[…]

### 2022-03-14 R/functionsSimulation.R
added brms models

@@ -73,6 +73,7 @@ fitOLSModel <- function(formula, data) {
     bX = as.numeric(coef(m)["x"]),
     Q2.5 = confint(m)["x","2.5 %"],
     Q97.5 = confint(m)["x","97.5 %"],
+    sig = (Q2.5 > 0 & Q97.5 > 0) | (Q2.5 < 0 & Q97.5 < 0),
     seed = unique(data$seed),
     rho = unique(data$rho),
     lambda = unique(data$lambda)
@@ -80,25 +81,37 @@ fitOLSModel <- function(formula, data) {
 }
 
 # initialise brms model
-setupBrms <- function(data) {
-  brm(y ~ 0 + Intercept + x + gp(latitude, longitude), data = data,
-      prior = c(prior(normal(0, 0.5), class = b),
-                prior(exponential(5), class = sdgp),
-                prior(exponential(5), class = sigma)),
-      chains = 0)
+setupBrms <- function(data, covMat, type = "") {
+  # formula
+  if (type == "spatial")    f <- bf(y ~ 0 + Intercept + x + gp(latitude, longitude, k = 5, c = 5/4))
+  if (type == "linguistic") f <- bf(y ~ 0 + Intercept + x + (1 | gr(country, cov = covMat)))
+  if (type == "both")       f <- bf(y ~ 0 + Intercept + x + gp(latitude, longitude, k = 5, c = 5/4) +
+                                      (1 | gr(country, cov = covMat)))
+  # priors
+  priors <- c(prior(normal(0, 0.5), class = b),
+              prior(exponential(5), class = sigma))
+  if (type == "spatial")    priors <- c(priors, prior(exponential(5), class = sdgp))
+  if (type == "linguistic") priors <- c(priors, prior(exponential(5), class = sd))
+  if (type == "both")       priors <- c(priors, prior(exponential(5), class = sdgp),
+                                        prior(exponential(5), class = sd))
+  # ensure covariance matrix is positive definite by adding small constant to diagonal
+  diag(covMat) <- diag(covMat) + 1e-06
+  # fit model
+  brm(f, data = data, data2 = list(covMat = covMat), prior = priors, chains = 0)
 }
 
 # fit brms model
 fitBrmsModel <- function(brmsInitial, data) {
   # fit model
[…]

### 2022-03-14 _targets.R

@@ -20,8 +20,6 @@ simulationTargets <-
     # map simulation over different covariance matrices and different values of lambda and rho
     values = expand_grid(covMat = rlang::syms(c("simGeoCov", "simLinCov")),
                          lambda = c(0.2, 0.5, 0.8), rho = c(0.2, 0.5, 0.8)),
-    # simulation seed
-    tar_target(seed, 1:5),
     # simulate data
     tar_target(simData, simulateData(covMat = covMat, continent, iso, langFam, 
                                      seed = seed, lambda = lambda, rho = rho), 
@@ -35,10 +33,11 @@ simulationTargets <-
     # conley se analyses
     tar_target(conleyModel1, fitConleyModel(simData, dist_cutoff = 100), pattern = map(simData)),
     tar_target(conleyModel2, fitConleyModel(simData, dist_cutoff = 1000), pattern = map(simData)),
-    tar_target(conleyModel3, fitConleyModel(simData, dist_cutoff = 10000), pattern = map(simData))
-    # brms analysis
-    #tar_target(brmsInitial, setupBrms(simData[[1]])),
-    #tar_target(brmsModel, fitBrmsModel(brmsInitial, simData), pattern = map(simData))
+    tar_target(conleyModel3, fitConleyModel(simData, dist_cutoff = 10000), pattern = map(simData)),
+    # brms analyses
+    tar_target(brmsModel1, fitBrmsModel(brmsInitial1, simData), pattern = map(simData)),
+    tar_target(brmsModel2, fitBrmsModel(brmsInitial2, simData), pattern = map(simData)),
+    tar_target(brmsModel3, fitBrmsModel(brmsInitial3, simData), pattern = map(simData))
     )
 
 # pipeline
@@ -121,6 +120,12 @@ list(
   # geographic and linguistic covariance matrices
   tar_target(simGeoCov, loadSimCovMat(fileGeo, log = TRUE , continent, iso, langFam)),
   tar_target(simLinCov, loadSimCovMat(fileLin, log = FALSE, continent, iso, langFam)),
+  # initialise brms models
+  tar_target(brmsInitial1, setupBrms(simData_simLinCov_0.8_0.8[[1]], simLinCov, type = "spatial")),
[…]

### 2022-03-15 R/functionsDAG.R
update sim plots

@@ -1,0 +1,20 @@
+# causal model of spatial and cultural phylogenetic non-independence
+plotDAG <- function() {
+  out <-
+    dagify(Y ~ X + U,
+           X ~ U,
+           U ~ L + G,
+           coords = tibble(name = c("X", "Y", "U", "G", "L"),
+                           x = c(1, 2, 1.5, 1, 2),
+                           y = c(1, 1, 0, 0, 0))) %>%
+    ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
+    geom_dag_point(data = . %>% filter(name %in% c("U")),
+                   shape = 1, stroke = 2, color = "black") +
+    geom_dag_text(color = "black", size = 10) +
+    geom_dag_edges(edge_color = "black", edge_width = 2,
+                   arrow_directed = grid::arrow(type = "closed")) +
+    theme_dag_blank()
+  # save
+  ggsave(out, filename = "figures/dag.pdf", height = 3.5, width = 6)
+  return(out)
+}

### 2022-03-15 R/functionsSimulation.R

@@ -50,7 +50,7 @@ simulateData <- function(covMat, continent, iso, langFam, seed, lambda, rho) {
   # standardise x and y
   x <- as.numeric(scale(x))
   y <- as.numeric(scale(y))
-  # produce data frame and match longitude, latitude, continent, and legal origin data
+  # produce data frame and match longitude, latitude, continent, and language family data
   out <- 
     tibble(x = x, y = y, country = rownames(covMat), seed = seed) %>%
     left_join(iso, by = c("country" = "iso2")) %>%
@@ -137,19 +137,24 @@ fitConleyModel <- function(data, dist_cutoff) {
   )
 }
 
-# plot simulation results for medium autocorrelation levels
-plotSimulation1 <- function(olsModel1, olsModel2, olsModel3, olsModel4, olsModel5, 
-                            conleyModel1, conleyModel2, conleyModel3, brmsModel) {
+# plot individual simulation results at medium autocorrelation levels
+plotSimInd <- function(olsModel1, olsModel2, olsModel3, olsModel4, olsModel5,
+                       conleyModel1, conleyModel2, conleyModel3, brmsModel1,
+                       brmsModel2, brmsModel3, type, file) {
   # plotting function
   plotFun <- function(results, title, ylab) {
     results %>%
-      mutate(sig = (Q2.5 > 0 & Q97.5 > 0) | (Q2.5 < 0 & Q97.5 < 0),
-             id = factor(1:nrow(.))) %>%
+      # medium autocorrelation levels
+      filter(rho == 0.5 & lambda == 0.5) %>%
+      # order by slope
+      mutate(id = factor(1:nrow(.))) %>%
       ggplot(aes(x = fct_reorder(id, bX, .desc = TRUE), 
                  y = bX, ymin = Q2.5, ymax = Q97.5, colour = sig)) +
+      # above or below is false positive
       geom_hline(yintercept = 0, linetype = "dashed") +
       geom_pointrange(size = 0.03) +
       scale_y_continuous(name = ylab, limits = c(-1, 1)) +
+      # red = sig, black = not sig
       scale_colour_manual(values = c("#000000", "#FF0000")) +
       ggtitle(title) +
       theme_classic() +
[…]

### 2022-03-15 _targets.R

@@ -152,11 +152,22 @@ list(
   tar_combine(brmsModel3_simGeoCov, simulationTargets$brmsModel3[1:9]  , command = dplyr::bind_rows(!!!.x)),
   tar_combine(brmsModel3_simLinCov, simulationTargets$brmsModel3[10:18], command = dplyr::bind_rows(!!!.x)),
   # plot simulation results
-  tar_target(plotSim1, plotSimulation1(olsModel1_0.5_0.5, olsModel2_0.5_0.5, olsModel3_0.5_0.5, 
-                                       olsModel4_0.5_0.5, olsModel5_0.5_0.5, conleyModel1_0.5_0.5, 
-                                       conleyModel2_0.5_0.5, conleyModel3_0.5_0.5, brmsModel_0.5_0.5)),
-  tar_target(plotSim2, plotSimulation2(olsModel1, olsModel2, olsModel3, olsModel4, olsModel5, 
-                                       conleyModel1, conleyModel2, conleyModel3, brmsModel)),
+  tar_target(plotSim1, plotSimInd(olsModel1_simGeoCov, olsModel2_simGeoCov, olsModel3_simGeoCov, olsModel4_simGeoCov, 
+                                  olsModel5_simGeoCov, conleyModel1_simGeoCov, conleyModel2_simGeoCov, conleyModel3_simGeoCov,
+                                  brmsModel1_simGeoCov, brmsModel2_simGeoCov, brmsModel3_simGeoCov, 
+                                  type = "spatial", file = "figures/simIndGeo.pdf")),
+  tar_target(plotSim2, plotSimInd(olsModel1_simLinCov, olsModel2_simLinCov, olsModel3_simLinCov, olsModel4_simLinCov, 
+                                  olsModel5_simLinCov, conleyModel1_simLinCov, conleyModel2_simLinCov, conleyModel3_simLinCov,
+                                  brmsModel1_simLinCov, brmsModel2_simLinCov, brmsModel3_simLinCov, 
+                                  type = "cultural phylogenetic", file = "figures/simIndLin.pdf")),
+  tar_target(plotSim3, plotSimAll(olsModel1_simGeoCov, olsModel2_simGeoCov, olsModel3_simGeoCov, olsModel4_simGeoCov, 
+                                  olsModel5_simGeoCov, conleyModel1_simGeoCov, conleyModel2_simGeoCov, conleyModel3_simGeoCov,
[…]

### 2022-03-15 _targets.R
scale up to 100 sims

@@ -125,7 +125,7 @@ list(
   tar_target(brmsInitial2, setupBrms(simData_simLinCov_0.8_0.8[[1]], simLinCov, type = "linguistic")),
   tar_target(brmsInitial3, setupBrms(simData_simLinCov_0.8_0.8[[1]], simLinCov, type = "both")),
   # seed for simulation
-  tar_target(seed, 1:2),
+  tar_target(seed, 1:100),
   # simulation
   simulationTargets,
   # combine
@@ -413,9 +413,9 @@ list(
                                                              knackM1a, skidmoreM1a),
                                                         list(adamczykData, alesinaData, beckData1, beckData2, bockstetteData,
                                                              easterlyData1, easterlyData2, fincherData, gelfandData, inglehartData,
-                                                             knackData, skidmoreData))),
+                                                             knackData, skidmoreData)))
   
   #### Manuscript ####
   
-  tar_render(manuscript, "manuscript.Rmd")
+  #tar_render(manuscript, "manuscript.Rmd")
 )

### 2022-03-15 manuscript.Rmd
first manuscript edits (up to methods)

@@
 ---
-title             : "Why non-independence matters for cross-national correlations"
-shorttitle        : "Non-independence between nations"
+title             : "The non-independence of nations and why it matters"
+shorttitle        : "Non-independence of nations"
 
@@
 Given the importance of nations for structuring human behaviour, there is justifiably huge interest in statistical analyses that attempt to predict variation in national-level outcomes of all kinds.
-At the time of writing, a search in Web of Science for the terms "cross-national", "cross-cultural", or "cross-country" in titles and abstracts returned approximately 80,000 unique hits.
+At the time of writing, a search in Web of Science for the terms "cross-national", "cross-cultural", or "cross-country" in titles and abstracts returned 80,000 unique hits.
 The standard practice for cross-national analyses is to conduct bivariate correlations or multiple regressions with individual data points representing different nations.
@@
 Bayes Factors indicated strong evidence that these estimates of geographic and cultural phylogenetic signal differed from zero, aside from the geographic signal for survival values, for which the evidence was equivocal (see Supplementary Results).
-These findings emphasise the need to account for spatial and cultural non-independence in cross-national analyses of economic development and cultural values.
+These findings emphasise the need to account for spatial and cultural phylogenetic non-independence in cross-national analyses of economic development and cultural values.
 
@@
 Our systematic review revealed that most cross-national analyses in the literature do not control for spatial or cultural phylogenetic non-independence.
-When they do, they tend to include controls like regional fixed effects, latitude, and legal origin fixed effects.
+When they do, they tend to include controls like latitude and regional fixed effects.
 Do these methods sufficiently account for statistical non-independence?
@@
 To compare the efficacy of different methods in the literature, we conducted a simulation study.
-We simulated national-level datasets (n = 178 nations) with varying degrees of spatial autocorrelation (i.e.
+We simulated national-level datasets (n = 236 nations) with varying degrees of spatial or cultural phylogenetic autocorrelation (i.e.
 non-independence) for outcome and predictor variables, but with no direct causal relationship between the variables.
-We then fitted naive regressions without controls to these datasets, as well as regression models with the following controls: latitude, longitude, continent fixed effects, and legal origin fixed effects.
-Despite not being identified in our systematic review, we also included in our simulation study Conley standard errors, a widely used standard error correction that purportedly accounts for spatial non-independence[@Conley1999;@Conley2010].
-Finally, we included a multilevel model of spatial non-independence that has recently been made widely available in statistical software packages[@Burkner2017;@rstan]: Bayesian Gaussian process regression[@McElreath2020;@Neal1998].
-This method allows nation random intercepts to covary according to a global geographic proximity matrix.
+We then fitted naive regressions without controls to these datasets, as well as regression models with controls for latitude, longitude, and continent fixed effects.
+Despite not being identified in our systematic review, we also included fixed effects for the language families of the majority-spoken languages in each country, as this control is often used to account for cultural phylogenetic non-independence (e.g.
+[@Currie2009]).
+In addition, we included Conley standard errors, a widely used standard error correction that purportedly accounts for spatial non-independence[@Conley1999;@Conley2010].
+Finally, we included Bayesian multilevel regressions that explicitly model spatial and/or cultural phylogenetic non-independence by allowing nation random intercepts to covary according to global geographic and/or linguistic proximity matrices (see Supplementary Methods).
+To model spatial non-independence we included a Gaussian process[@McElreath2020;@Neal1998] over latitude and longitude values, and to model cultural phylogenetic non-independence we assume that nation random intercepts are correlated proportional to their linguistic proximity[@deVillemereuil2014].
 Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero (i.e.
@@
 ```{r calculate false positive rate}
-fp <-
-  tar_read(olsModel1) %>%
+fpGeo <-
+  tar_read(olsModel1_simGeoCov) %>%
   group_by(lambda, rho) %>%
   summarise(n = sum(sig) / n())
-fpConley <-
-  tar_read(conleyModel3) %>%
+fpLin <-
+  tar_read(olsModel1_simLinCov) %>%
   group_by(lambda, rho) %>%
   summarise(n = sum(sig) / n())
-fpGP <-
-  tar_read(brmsModel) %>%
-  group_by(lambda, rho) %>%
-  summarise(n = sum(sig) / n())
+#fpConley <-
+#  tar_read(conleyModel3) %>%
+#  group_by(lambda, rho) %>%
+#  summarise(n = sum(sig) / n())
+#fpGP <-
+#  tar_read(brmsModel) %>%
+#  group_by(lambda, rho) %>%
+#  summarise(n = sum(sig) / n())
 ```
 
-Figure \@ref(fig:plotSim2) plots the estimated false positive rates from our simulation study, split by different methods and different degrees of spatial autocorrelation (see Supplementary Figure \@ref(fig:plotSim1) for full distribution of effect sizes under moderate autocorrelation).
-With at least moderate degrees of spatial autocorrelation for both outcome and predictor variables, naive regression models (Figure \@ref(fig:plotSim2) top left) produce false positive rates above chance levels.
-With strong spatial autocorrelation for both outcomes and predictors, false positive rates reach as high as `r printnum(fp$n[fp$lambda == 0.9 & fp$rho == 0.9]*100, digits = 0)`%.
+Figures \@ref(fig:plotSim1) and \@ref(fig:plotSim2) plot the estimated false positive rates from our simulation study, split by different methods and different degrees of spatial or cultural phylogenetic autocorrelation (see Supplementary Figures \@ref(fig:plotSim3) and \@ref(fig:plotSim4) for full distributions of effect sizes under moderate autocorrelation).
+For reference, "weak" autocorrelation in our simulation is comparable to the geographic signal for survival values in Figure \@ref(fig:plotSignal), while "moderate" and "strong" levels of autocorrelation are comparable to the cultural phylogenetic signal for traditional and survival values, respectively.
+With at least moderate degrees of spatial or cultural phylogenetic autocorrelation for both outcome and predictor variables, naive regression models produce false positive rates above chance levels.
+With strong spatial autocorrelation for both outcomes and predictors, false positive rates reach as high as `r printnum(fpGeo$n[fpGeo$lambda == 0.9 & fpGeo$rho == 0.9]*100, digits = 0)`%.
+We find the same false positive rate under strong cultural phylogenetic autocorrelation (`r printnum(fpLin$n[fpLin$lambda == 0.9 & fpLin$rho == 0.9]*100, digits = 0)`%`).
 
-(ref:plotSim2Caption) _False positive rates for different methods of controlling for non-independence in our simulation study._ For simulated outcome and predictor variables, we systematically varied the strength of spatial autocorrelation, from weak (0.1) to moderate (0.5) to strong (0.9).
+(ref:plotSim1Caption) _False positive rates for different methods of controlling for spatial non-independence in our simulation study._ For simulated outcome and predictor variables, we systematically varied the strength of spatial autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
 We simulated 100 datasets per parameter combination and fitted different models to each dataset.
@@
 
-```{r plotSim2, fig.width=10, fig.height=7, fig.cap="(ref:plotSim2Caption)"}
+```{r plotSim1, fig.width=9.5, fig.height=5.5, fig.cap="(ref:plotSim1Caption)"}
+tar_read(plotSim1)
+```
+
+(ref:plotSim2Caption) _False positive rates for different methods of controlling for cultural phylogenetic non-independence in our simulation study._ For simulated outcome and predictor variables, we systematically varied the strength of cultural phylogenetic autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+We simulated 100 datasets per parameter combination and fitted different models to each dataset.
+False positive rates were operationalised as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero.
+Point ranges represent proportions and 95% bootstrap confidence intervals, and dashed lines indicate the 5% false positive rate that is expected due to chance.
+SEs = standard errors.
+
+```{r plotSim2, fig.width=9.5, fig.height=5.5, fig.cap="(ref:plotSim2Caption)"}
 tar_read(plotSim2)
@@
 
-Common methods in the literature do not reduce this high false positive rate.
-With strong spatial autocorrelation for both outcome and predictor variables, false positive rates remain above 70% when controlling for latitude, longitude, continent fixed effects, and legal origin fixed effects.
+Common methods in the literature do not reduce these high false positive rates.
+With strong spatial and cultural phylogenetic autocorrelation for both outcome and predictor variables, false positive rates remain above 70% when controlling for latitude, longitude, continent fixed effects, and language family fixed effects.
 Applying Conley standard errors with cutoffs of 100 km or 1,000 km also does not reduce false positive rates below 70% under strong spatial autocorrelation.
-With a cutoff of 10,000 km, the false positive rate under strong spatial autocorrelation reduces to `r printnum(fpConley$n[fpConley$lambda == 0.9 & fpConley$rho == 0.9]*100, digits = 0)`%, though this correction adds bias at weaker levels of spatial autocorrelation, increasing false positive rates above chance levels.
+With a cutoff of 10,000 km, the false positive rate under strong spatial autocorrelation reduces to XX%, though this correction adds bias at weaker levels of spatial autocorrelation, increasing false positive rates above chance levels.
 
-By contrast, Bayesian Gaussian process regression adequately accounts for non-independence across a range of simulation parameters.
+<!-- By contrast, Bayesian Gaussian process regression adequately accounts for non-independence across a range of simulation parameters.
 At moderate levels of spatial autocorrelation on both outcomes and predictors, Gaussian process regression eliminates false positives, such that the false positive rate is no different from chance.
 With strong spatial autocorrelation on both outcomes and predictors, the false positive rate is reduced to `r printnum(fpGP$n[fpGP$lambda == 0.9 & fpGP$rho == 0.9]*100, digits = 0)`%, substantially less than any other method.
+-->
 
@@
 
-To diagnose how widespread this issue of spurious cross-national relationships might be, we reanalysed a subset of twelve previous cross-national analyses from our systematic review, sufficiently controlling for spatial and cultural phylogenetic non-independence using global geographic and linguistic proximity matrices.
+To determine how widespread this issue of spurious cross-national relationships might be, we reanalysed a subset of twelve previous cross-national analyses from our systematic review, sufficiently controlling for spatial and cultural phylogenetic non-independence using global geographic and linguistic proximity matrices.
 We subsampled six analyses from our economic development review[@Beck2003;@Beck2005;@Bockstette2002;@Easterly2003;@Easterly2007;@Skidmore2002] and six from our cultural values review[@Adamczyk2009;@Alesina2013;@Fincher2008;@Gelfand2011;@Inglehart2000;@Knack1997].
@@
 Overall, cross-national correlation effect sizes tended to reduce when controlling for statistical non-independence between nations, sometimes by as much as half of the original effect size.
-Based on 95% credible intervals, six out of twelve cross-national relationships are attenuated when controlling for non-independence.
-For the economic development analyses, four out of six cross-national relationships are attenuated when controlling for spatial non-independence.
-For the cultural values analyses, two out of six are attenuated when controlling for cultural phylogenetic non-independence.
+Overall, after controlling for non-independence, six out of twelve cross-national associations had 95% credible intervals that included zero.
+For the economic development analyses, four out of six cross-national relationships had 95% credible intervals including zero when controlling for spatial non-independence.
+For the cultural values analyses, two out of six cross-national relationships had 95% credible intervals including zero when controlling for cultural phylogenetic non-independence.
 Supplementary Figure \@ref(fig:plotReplications2) shows these cross-national correlations plotted against the raw data.
@@
 We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \@ref(fig:plotReplications4)), with some outcome variables expressing low signal (e.g.
-Gelfand et al.[@Gelfand2011]; posterior median = `r printnum(median(gelfandSignal))`, 95% CI [`r printnum(quantile(gelfandSignal, 0.025))` `r printnum(quantile(gelfandSignal, 0.975))`]) and others expressing high signal (e.g.
-Alesina et al.[@Alesina2013]; posterior median = `r printnum(median(alesinaSignal))`, 95% CI [`r printnum(quantile(alesinaSignal, 0.025))` `r printnum(quantile(alesinaSignal, 0.975))`]).
+tightness[@Gelfand2011]; posterior median = `r printnum(median(gelfandSignal))`, 95% CI [`r printnum(quantile(gelfandSignal, 0.025))` `r printnum(quantile(gelfandSignal, 0.975))`]) and others expressing high signal (e.g.
+female labour force participation[@Alesina2013]; posterior median = `r printnum(median(alesinaSignal))`, 95% CI [`r printnum(quantile(alesinaSignal, 0.025))` `r printnum(quantile(alesinaSignal, 0.975))`]).
 Across all analyses, we found that stronger estimates of spatial autocorrelation or cultural phylogenetic signal resulted in a more pronounced reduction in the effect size when controlling for non-independence between nations (Figure \@ref(fig:plotReplications5)).
@@
 
-Non-independence is a concern because economic development and cultural values are highly geographically and culturally structured around the world, as revealed by our initial modelling.
-These results build on previous work.
-One previous study suggested that geographic proximity is more important than deep cultural ancestry in explaining the distribution of human development across Eurasian nations[@Sookias2018].
-By contrast, with a global sample of nations, we found strong cultural phylogenetic signal, as well as geographic signal, for the Human Development Index.
+Our initial analyses add to and clarify existing evidence of the non-independence of economic and cultural variation among nations.
+One previous study suggested that geographic proximity is more important than deep cultural ancestry in explaining the distribution of human development across Eurasian nations, though the authors noted that their small sample of 44 nations and regional focus limited their statistical power[@Sookias2018].
+Our global sample of nations revealed strong cultural phylogenetic signal, as well as geographic signal, for the Human Development Index.
 Another previous study found that similarities in the cultural values of nations are predicted by linguistic, but not geographic, distances between those nations[@White2021].
@@
 Yet our systematic literature review and simulation study revealed that the most commonly used controls for non-independence do not sufficiently deal with the issue.
-In our simulations, controlling for either latitude or longitude did not reduce false positive rates, calling into question controls like distance to the equator in regression models.
+In our simulations, controlling for either latitude or longitude did not reduce false positive rates.
+This result calls into question controls like distance to the equator to account for non-independence in cross-national regression models, though these controls maybe still be suitable to account for regional or latitudinal variation in ecology, which we did not simulate.
 High false positive rates persisted with Conley standard errors, which have previously been critiqued for being overly sensitive to arbitrary distance cutoffs[@Kelly2020].
 The simulation also confirmed the verbal claim that fixed effects for spatial or cultural groupings (e.g.
-continent or legal origin fixed effects) are insufficient because non-independence still remains within groupings[@Bromham2018].
-This logic applies to analyses that control for non-independence by including language family fixed effects (e.g.
-[@Currie2009]) or separately analysing different regions (e.g.
+continent or language family fixed effects) are insufficient because non-independence still remains within groupings[@Bromham2018].
+This logic further applies to analyses that control for non-independence by separately analysing different regions (e.g.
 [@Smith2017]).
-The only method in our simulation that sufficiently reduced the false positive rate was Bayesian Gaussian process regression, though we did not include other possible controls for non-independence, such as conditional autoregressive models[@Lichstein2002] or generalised additive models[@Hastie2017].
+Across all model types in our simulation, the only methods that sufficiently reduced the false positive rate were the Bayesian multilevel regressions that explicitly modelled spatial and cultural phylogenetic autocorrelation, though we did not include other possible controls for non-independence, such as conditional autoregressive models[@Lichstein2002] or generalised additive models[@Hastie2017].
 
@@
 At the institutional level, one possibility is that such practices are incentivised _because_ they inflate false positive rates, increasing the probability of getting published and promoted[@Smaldino2016].
-At the individual level, another possibility is that researchers outside of anthropology and ecology are simply not aware of the problem, or believe that the problem does not apply at the national level.
+At the individual level, another possibility is that researchers outside of anthropology and ecology are simply not aware of the problem, or believe that the problem does not apply to analyses of nations.
 Even if researchers appreciate the problem, they might not know of suitable controls or perceive the methods to be too complex.
@@
 Until such changes are implemented and sufficient controls for non-independence are the norm, existing cross-national correlations should be interpreted with caution.
-In our reanalyses, we found that four out of six cross-national analyses of economic development were attenuated when controlling for spatial non-independence.
-Three of these were tests of "persistence" hypotheses, studying the effects of historical and environmental conditions --- settler mortality[@Beck2003], wheat-sugar suitability[@Easterly2007], and natural disaster frequency[@Skidmore2002] --- on modern developmental outcomes.
+In our reanalyses, we found that four out of six cross-national correlations with economic development variables had 95% credible intervals that included zero when controlling for spatial non-independence.
+Three of these analyses were tests of "persistence" hypotheses, studying the effects of historical and environmental conditions --- settler mortality[@Beck2003], wheat-sugar suitability[@Easterly2007], and natural disaster frequency[@Skidmore2002] --- on modern developmental outcomes.
 A recent reanalysis has also called into question various studies of this ilk[@Kelly2020].
-We also found that two out of six cross-national analyses of cultural values were attenuated when controlling for cultural phylogenetic non-independence.
-It is interesting that the economic development analyses were attenuated with spatial controls, while the cultural values analyses were attenuated by cultural phylogenetic controls.
+We also found that two out of six cross-national correlations with cultural values variables had 95% credible intervals that included zero when controlling for cultural phylogenetic non-independence.
+It is interesting that the economic development analyses were more affected by the spatial controls, while the cultural values analyses were more affected by the cultural phylogenetic controls.
 This highlights that certain kinds of non-independence will be more important depending on the particular national-level outcome and predictor variables of interest.
@@
 We do not wish to dissuade researchers from conducting cross-national studies.
-On the contrary, such work promises to deepen our understanding of human populations beyond Western, educated, industrialised, rich, and democratic societies[@Henrich2010], providing the representative samples necessary to test evolutionary and socio-ecological theories of human behaviour[@Pollet2014;@Oishi2013].
+On the contrary, such work promises to deepen our understanding of the causes and consequences of economic development and cultural values around the world.
+Moreover, cross-national studies allow social scientists to broaden their scope of study beyond Western populations[@Henrich2010], providing the representative samples necessary to test evolutionary and socio-ecological theories of human behaviour[@Pollet2014;@Oishi2013].
 But in order to minimise spurious relationships in global datasets, we urge researchers to control for spatial and cultural phylogenetic non-independence when reporting cross-national correlations.
-Nations are not independent, and our statistical models should reflect this.
+Nations are not independent, and our statistical models must reflect this.
 

### 2022-03-31 appendix.Rmd
updated manuscript

@@
 cultural norms, political systems).
-In this causal model, we cannot condition on $U$ directly, as it is unobserved.
+In this causal model, we need to condition on $U$ to estimate the direct path from $X$ to $Y$, but we cannot since it is unobserved.
 However, geographic $G$ and linguistic $L$ relationships between societies influence $U$, since changing a nation's spatial distance to or shared cultural ancestry with other nations will change its environmental and cultural traits.
 We can thus use $G$ and $L$ to model the covariation between $X$ and $Y$ induced by $U$.
-Failing to do this and simply estimating the bivariate correlation between $X$ and $Y$ will produce spurious relationships and residuals that are spatially and culturally non-independent around the world (see Figure \@ref(fig:spatialExample)).
+Failing to do this and simply estimating the bivariate correlation between $X$ and $Y$ will produce spurious relationships and residuals that are spatially and culturally non-independent around the world.
 
@@
 
-(ref:plotSim1Caption) _Distribution of cross-national correlations from simulation study under moderate spatial autocorrelation._ In these simulations, the strength of spatial autocorrelation is set to 0.5 for both outcome and predictor variables.
+(ref:plotSim1Caption) _Distribution of cross-national correlations from simulation study under strong spatial autocorrelation._ In these simulations, the strength of spatial autocorrelation is set to 0.8 for both outcome and predictor variables.
 For frequentist regression models, point ranges represent correlation estimates and 95% confidence intervals.
-For Bayesian Gaussian process regression models, point ranges represent posterior means and 95% credible intervals.
+For Bayesian regression models, point ranges represent posterior means and 95% credible intervals.
 Correlations are ordered by effect size independently in each panel.
@@
 
-```{r plotSim1, fig.width=7.5, fig.height=5, fig.cap="(ref:plotSim1Caption)"}
+```{r plotSim1, fig.width=9.5, fig.height=5.5, fig.cap="(ref:plotSim1Caption)"}
 tar_read(plotSim1)
+```
+
+\newpage
+
+(ref:plotSim2Caption) _Distribution of cross-national correlations from simulation study under strong cultural phylogenetic autocorrelation._ In these simulations, the strength of cultural phylogenetic autocorrelation is set to 0.8 for both outcome and predictor variables.
+For frequentist regression models, point ranges represent correlation estimates and 95% confidence intervals.
+For Bayesian regression models, point ranges represent posterior means and 95% credible intervals.
+Correlations are ordered by effect size independently in each panel.
+Red point ranges indicate that the slope is "significant" (i.e.
+the 95% confidence / credible interval excludes zero).
+Black point ranges indicate that the slope is "not significant".
+
+```{r plotSim2, fig.width=9.5, fig.height=5.5, fig.cap="(ref:plotSim2Caption)"}
+tar_read(plotSim2)
 ```
@@
 Cultural phylogenetic signal is operationalised as the proportion of national-level variance explained by linguistic proximity between nations.
-Ridges are full posterior distributions, and point ranges are posterior medians and 95% credible intervals.
+Ridges are full posterior distributions, and points are posterior medians, and lines represent 50% and 95% credible intervals.
 

### 2022-03-31 manuscript.Rmd
updated manuscript

@@
 keywords          : "spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses"
-wordcount         : "5549 words"
+wordcount         : "5500 words"
 bibliography      : ["r-references.bib"]
@@
 ```{r setup, include=FALSE}
-# !!! for use without tar_make
-library(targets)
-library(tarchetypes)
-library(tidyverse)
-source("R/functionsExamplePlot.R")
-source("R/functionsSignal.R")
-source("R/functionsReview.R")
-source("R/functionsSimulation.R")
-source("R/functionsReplications.R")
-options(tidyverse.quiet = TRUE)
-lapply(c("brms", "cowplot", "conleyreg", "countrycode", 
-         "dagitty", "geosphere", "ggdag", "ggrepel", "ggtext", 
-         "haven", "lmtest", "papaja", "psych", "readxl", 
-         "rstan", "sjlabelled", "tidybayes", "tidyverse"), library, character.only = TRUE)
-# !!!
-
+## !!! for use without tar_make
+#library(targets)
+#library(tarchetypes)
+#library(tidyverse)
+#source("R/functionsDAG.R")
+#source("R/functionsSignal.R")
+#source("R/functionsReview.R")
+#source("R/functionsSimulation.R")
+#source("R/functionsReplications.R")
+#options(tidyverse.quiet = TRUE)
+#lapply(c("brms", "cowplot", "conleyreg", "countrycode", 
+#         "dagitty", "geosphere", "ggdag", "ggrepel", "ggtext", 
+#         "haven", "lmtest", "papaja", "psych", "readxl", 
+#         "rstan", "sjlabelled", "tidybayes", "tidyverse"), library, character.only = TRUE)
+## !!!
+#
 knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
@@
 Second, nations with shared cultural ancestry tend to be more similar than culturally unrelated nations.
-This is known as cultural phylogenetic non-independence[@Tylor1889], and occurs because related nations share cultural traits inherited via descent from a common ancestor.
+This is known as cultural phylogenetic non-independence[@Tylor1889;@Naroll1961;@Naroll1965], and occurs because related nations share cultural traits inherited via descent from a common ancestor.
 Shared cultural ancestry can result in a form of pseudoreplication, whereby multiple instances of the same trait across nations are merely duplicates of the ancestral original.
-For example, despite being separated by over 1500 km of ocean, the related island nations Tonga and Tuvalu express similar languages and customs due to cultural inheritance from a shared ancestor deep in the evolutionary history of Polynesian societies.
+For example, despite being separated by over 1500 kilometres of ocean, the related island nations Tonga and Tuvalu express similar languages and customs due to cultural inheritance from a shared ancestor deep in the evolutionary history of Polynesian societies.
 More recently, shared ancestry is also crucial in explaining cultural similarities between colonial settlements and their colonisers (e.g.
-Spain and Argentina).
+Argentina and Spain).
 
@@
 But without accounting for spatial or cultural non-independence between nations, model residuals can show structure that remains unaccounted for, violating this assumption.
-Treating nations as independent can thus inflate false positive rates[@Legendre1993], producing spurious "direct" relationships between variables that in fact only indirectly covary due to spatial or cultural connections (see Supplementary Figure \@ref(fig:dag) for an example causal model).
+Treating nations as independent can thus inflate false positive rates[@Legendre1993], producing spurious "direct" relationships between variables that in fact only indirectly covary due to spatial or cultural connections[@Roberts2013] (see Supplementary Figure \@ref(fig:dag) for an example causal model).
 
@@
 In ecology and sociology, studies with regional-level data use a variety of autoregressive models designed to account for spatial patternings[@Lichstein2002;@Loftin1983].
-In anthropology, the Standard Cross-Cultural Sample of 186 cultures was compiled to minimise the confounding effects of non-independence in comparisons of traditional societies[@Murdock1969], though spatial and cultural dependencies are difficult to remove entirely[@Dow2008;@Eff2004].
+In anthropology, researchers have recognised cultural non-independence as an important statistical pitfall for over 200 years, with issues of cultural pseudoreplication being identified in early comparative studies of marriage practices across societies[@Tylor1889].
+More recently, the Standard Cross-Cultural Sample of 186 cultures was compiled to minimise the confounding effects of non-independence in comparisons of traditional societies[@Murdock1969], though spatial and cultural dependencies are difficult to remove entirely[@Dow2008;@Eff2004].
 Anthropologists also borrow phylogenetic comparative methods from evolutionary biology, such as phylogenetic least squares regression[@Symonds2014], when comparing traditional societies, treating culturally related societies in the same way as biologists treat genetically related species (e.g.
@@
 Another replication study found that many of the widely publicised relationships between national-level pathogen prevalence and political institutions and attitudes fail to hold when controlling for various kinds of non-independence[@Bromham2018].
-These reanalyses, and others[@Bromham2021;@Currie2012], raise the question: how widespread a concern is non-independence in studies of national-level outcomes?
+These reanalyses, and others[@Bromham2021;@Currie2012;@Passmore2022], raise the question: how widespread a concern is non-independence in studies of national-level outcomes?
 
@@
 latitude) and shared cultural history (e.g.
-legal origin fixed effects).
+colony status).
 
@@
 Finally, we included Bayesian multilevel regressions that explicitly model spatial and/or cultural phylogenetic non-independence by allowing nation random intercepts to covary according to global geographic and/or linguistic proximity matrices (see Supplementary Methods).
-To model spatial non-independence we included a Gaussian process[@McElreath2020;@Neal1998] over latitude and longitude values, and to model cultural phylogenetic non-independence we assume that nation random intercepts are correlated proportional to their linguistic proximity[@deVillemereuil2014].
+To model spatial non-independence we included a Gaussian process[@McElreath2020;@Neal1998] over latitude and longitude values, and to model cultural phylogenetic non-independence we assumed that nation random intercepts were correlated proportional to their linguistic proximity[@deVillemereuil2014].
 Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero (i.e.
@@
   summarise(n = sum(sig) / n())
-#fpConley <-
-#  tar_read(conleyModel3) %>%
-#  group_by(lambda, rho) %>%
-#  summarise(n = sum(sig) / n())
-#fpGP <-
-#  tar_read(brmsModel) %>%
-#  group_by(lambda, rho) %>%
-#  summarise(n = sum(sig) / n())
+fpGeoContinent <-
+  tar_read(olsModel4_simGeoCov) %>%
+  group_by(lambda, rho) %>%
+  summarise(n = sum(sig) / n())
+fpGeoSpatial <-
+  tar_read(brmsModel1_simGeoCov) %>%
+  group_by(lambda, rho) %>%
+  summarise(n = sum(sig) / n())
+fpLinLangFam <-
+  tar_read(olsModel5_simLinCov) %>%
+  group_by(lambda, rho) %>%
+  summarise(n = sum(sig) / n())
 ```
 
-Figures \@ref(fig:plotSim1) and \@ref(fig:plotSim2) plot the estimated false positive rates from our simulation study, split by different methods and different degrees of spatial or cultural phylogenetic autocorrelation (see Supplementary Figures \@ref(fig:plotSim3) and \@ref(fig:plotSim4) for full distributions of effect sizes under moderate autocorrelation).
+Figures \@ref(fig:plotSim3) and \@ref(fig:plotSim4) plot the estimated false positive rates from our simulation study, split by different methods and different degrees of spatial or cultural phylogenetic autocorrelation (see Supplementary Figures \@ref(fig:plotSim1) and \@ref(fig:plotSim2) for full distributions of effect sizes under strong autocorrelation).
 For reference, "weak" autocorrelation in our simulation is comparable to the geographic signal for survival values in Figure \@ref(fig:plotSignal), while "moderate" and "strong" levels of autocorrelation are comparable to the cultural phylogenetic signal for traditional and survival values, respectively.
-With at least moderate degrees of spatial or cultural phylogenetic autocorrelation for both outcome and predictor variables, naive regression models produce false positive rates above chance levels.
-With strong spatial autocorrelation for both outcomes and predictors, false positive rates reach as high as `r printnum(fpGeo$n[fpGeo$lambda == 0.9 & fpGeo$rho == 0.9]*100, digits = 0)`%.
-We find the same false positive rate under strong cultural phylogenetic autocorrelation (`r printnum(fpLin$n[fpLin$lambda == 0.9 & fpLin$rho == 0.9]*100, digits = 0)`%`).
 
-(ref:plotSim1Caption) _False positive rates for different methods of controlling for spatial non-independence in our simulation study._ For simulated outcome and predictor variables, we systematically varied the strength of spatial autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+Our simulation study revealed that with at least moderate degrees of spatial or cultural phylogenetic autocorrelation for both outcome and predictor variables, naive regression models produce false positive rates above chance levels.
+This false positive rate increases as the degree of autocorrelation increases.
+With strong spatial autocorrelation for both outcomes and predictors, false positive rates reach as high as `r printnum(fpGeo$n[fpGeo$lambda == 0.8 & fpGeo$rho == 0.8]*100, digits = 0)`%.
+We find a slightly lower false positive rate under strong cultural phylogenetic autocorrelation, though this false positive rate is still greater than expected by chance (`r printnum(fpLin$n[fpLin$lambda == 0.8 & fpLin$rho == 0.8]*100, digits = 0)`%).
+
+(ref:plotSim3Caption) _False positive rates for different methods of controlling for spatial non-independence in our simulation study._ For simulated outcome and predictor variables, we systematically varied the strength of spatial autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
 We simulated 100 datasets per parameter combination and fitted different models to each dataset.
@@
 
-```{r plotSim1, fig.width=9.5, fig.height=5.5, fig.cap="(ref:plotSim1Caption)"}
-tar_read(plotSim1)
+```{r plotSim3, fig.width=11, fig.height=7, fig.cap="(ref:plotSim3Caption)"}
+tar_read(plotSim3)
 ```
 
-(ref:plotSim2Caption) _False positive rates for different methods of controlling for cultural phylogenetic non-independence in our simulation study._ For simulated outcome and predictor variables, we systematically varied the strength of cultural phylogenetic autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+(ref:plotSim4Caption) _False positive rates for different methods of controlling for cultural phylogenetic non-independence in our simulation study._ For simulated outcome and predictor variables, we systematically varied the strength of cultural phylogenetic autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
 We simulated 100 datasets per parameter combination and fitted different models to each dataset.
@@
 
-```{r plotSim2, fig.width=9.5, fig.height=5.5, fig.cap="(ref:plotSim2Caption)"}
-tar_read(plotSim2)
+```{r plotSim4, fig.width=11, fig.height=7, fig.cap="(ref:plotSim4Caption)"}
+tar_read(plotSim4)
 ```
@@
 Common methods in the literature do not reduce these high false positive rates.
-With strong spatial and cultural phylogenetic autocorrelation for both outcome and predictor variables, false positive rates remain above 70% when controlling for latitude, longitude, continent fixed effects, and language family fixed effects.
-Applying Conley standard errors with cutoffs of 100 km or 1,000 km also does not reduce false positive rates below 70% under strong spatial autocorrelation.
-With a cutoff of 10,000 km, the false positive rate under strong spatial autocorrelation reduces to XX%, though this correction adds bias at weaker levels of spatial autocorrelation, increasing false positive rates above chance levels.
+With strong spatial autocorrelation for both outcome and predictor variables, false positive rates remain above 50% when controlling for latitude, longitude, and language family fixed effects (Figure \@ref(fig:plotSim3)).
+Applying Conley standard errors also does not reduce false positive rates below 50% under strong spatial autocorrelation, regardless of the distance cutoff.
+Continent fixed effects are more effective than other frequentist methods, though they continue to produce a false positive rate of `r printnum(fpGeoContinent$n[fpGeoContinent$lambda == 0.8 & fpGeoContinent$rho == 0.8]*100, digits = 0)`% under strong spatial autocorrelation.
+By contrast, Bayesian spatial Gaussian process regression with longitude and latitude values outperforms all other methods.
+This approach eliminates false positives under moderate spatial autocorrelation, such that the false positive rate is no different from chance, and reduces the false positive rate under strong spatial autocorrelation to `r printnum(fpGeoSpatial$n[fpGeoSpatial$lambda == 0.8 & fpGeoSpatial$rho == 0.8]*100, digits = 0)`%.
+Bayesian models that additionally account for linguistic proximity between nations perform equally well, though models with only linguistic covariance continue to produce false positives.
 
-<!-- By contrast, Bayesian Gaussian process regression adequately accounts for non-independence across a range of simulation parameters.
-At moderate levels of spatial autocorrelation on both outcomes and predictors, Gaussian process regression eliminates false positives, such that the false positive rate is no different from chance.
-With strong spatial autocorrelation on both outcomes and predictors, the false positive rate is reduced to `r printnum(fpGP$n[fpGP$lambda == 0.9 & fpGP$rho == 0.9]*100, digits = 0)`%, substantially less than any other method.
--->
+In our simulation of cultural phylogenetic non-independence, we find that most frequentist methods do not reduce false positive rates (Figure \@ref(fig:plotSim4)).
+Controls for latitude and longitude, continent fixed effects, and Conley standard errors do little to change false positive rates.
+Language family fixed effects are slightly more effective than other frequentist methods, though they continue to produce a false positive rate of `r printnum(fpLinLangFam$n[fpLinLangFam$lambda == 0.8 & fpLinLangFam$rho == 0.8]*100, digits = 0)`% under strong cultural phylogenetic autocorrelation.
+By contrast, Bayesian models with random effects covarying according to linguistic proximity completely eliminate false positives across all degrees of cultural phylogenetic autocorrelation.
+Bayesian models that additionally account for geographic proximity between nations perform equally well, though models with only a spatial Gaussian process continue to produce false positives.
 
@@
 
-Our systematic review and simulation study have showed that controls for non-independence are rare in cross-national studies of economic development and cultural values, and when studies do attempt to control for non-independence, the methods used continue to produce false positives.
+Our systematic review and simulation study have shown that controls for non-independence are rare in cross-national studies of economic development and cultural values, and when studies do attempt to control for non-independence, the methods used continue to produce false positives.
 This raises the worrying possibility that the cross-national literature in economics and psychology is populated with spurious relationships.
@@
 Figure \@ref(fig:plotReplications1) visualises the results of our reanalysis.
-Overall, cross-national correlation effect sizes tended to reduce when controlling for statistical non-independence between nations, sometimes by as much as half of the original effect size.
+Cross-national correlation effect sizes tended to reduce when controlling for statistical non-independence between nations, sometimes by as much as half of the original effect size.
 Overall, after controlling for non-independence, six out of twelve cross-national associations had 95% credible intervals that included zero.
@@
 
-Our initial analyses add to and clarify existing evidence of the non-independence of economic and cultural variation among nations.
+Our initial analyses add to and clarify existing evidence regarding the non-independence of economic and cultural variation among nations.
 One previous study suggested that geographic proximity is more important than deep cultural ancestry in explaining the distribution of human development across Eurasian nations, though the authors noted that their small sample of 44 nations and regional focus limited their statistical power[@Sookias2018].
@@
 These institutional- and individual-level barriers can be combatted.
-First, cross-national replication studies like ours and others[@Bromham2018;@Bromham2021;@Currie2012;@Kelly2020], combined with the methodological reviews included in Registered Reports[@Chambers2021], might change incentive structures and encourage researchers to analyse the world's nations with more rigorous methods.
+First, cross-national replication studies like ours and others[@Bromham2018;@Bromham2021;@Currie2012;@Kelly2020;@Passmore2022], combined with the methodological reviews included in Registered Reports[@Chambers2021], might change incentive structures and encourage researchers to analyse the world's nations with more rigorous methods.
 Second, since the issue of non-independence is fundamentally an issue of causal inference (Supplementary Figure \@ref(fig:dag)), more explicit descriptions of causal models could promote controls for non-independence in cross-national research.
@@
 
-To estimate the degree of spatial and cultural phylogenetic non-independence in economic development and cultural values, we calculated geographic and linguistic signal for global measures of development and values.
+To estimate the degree of spatial and cultural phylogenetic non-independence in economic development and cultural values, we calculated geographic and cultural phylogenetic signal for global measures of development and values.
 Our measure of economic development was the Human Development Index[@hdi].
@@
 
-To calculate geographic and linguistic signal, we created two proximity matrices for 269 of the world's nations: a geographic proximity matrix and a linguistic proximity matrix.
+To calculate geographic and cultural phylogenetic signal, we created two proximity matrices for 269 of the world's nations: a geographic proximity matrix and a linguistic proximity matrix.
 Geographic proximity was converted from logged geodesic distances between nation capital cities.
@@
 These models were fitted with the R package _brms_[@Burkner2017] and converged normally ($\hat{R}$ < 1.1).
-Estimates of geographic and linguistic signal were computed as the proportion of national-level variance in these models explained by geographic and linguistic proximity matrices.
+Estimates of geographic and cultural phylogenetic signal were computed as the proportion of national-level variance in these models explained by geographic and linguistic proximity matrices.
 
@@
 
-To simulate data, we used our global geographic distance matrix (see Supplementary Methods) as a weight matrix $W$, scaling the matrix between 0 and 1 and restricting the matrix to 178 nations with available external data (see below).
-Using this weight matrix, we simulated outcome $Y$ and predictor $X$ variables using the following formulae from a previous simulation study[@Dow1984]:
+We simulated data for 236 nations with varying degrees of spatial or cultural phylogenetic signal for outcome $y$ and predictor $x$ using the following generative model:
 
@@
 \begin{aligned}
-Y=(1-\lambda W)^{-1} \zeta + \nu \\
-X=(1-\rho W)^{-1} \xi + \omega
+y &= \alpha_y + \epsilon_y \\
+x &= \alpha_x + \epsilon_x \\
+\alpha_y &\sim \mathcal{N}(0, \sqrt{\lambda} \cdot \Sigma) \\
+\alpha_x &\sim \mathcal{N}(0, \sqrt{\rho} \cdot \Sigma) \\
+\epsilon_y &\sim \mathcal{N}(0, \sqrt{1 - \lambda}) \\
+\epsilon_x &\sim \mathcal{N}(0, \sqrt{1 - \rho})
 \end{aligned}
@@
 
-where $\zeta$, $\nu$, $\xi$, and $\omega$ are all standardised normal random variables, and $\lambda$ and $\rho$ are autocorrelation parameters that vary between 0 and 1, reflecting the strength of spatial autocorrelation for outcome and predictor variables, respectively.
-Importantly, in this simulation, we know that there is no direct causal relationship between $X$ and $Y$ because we simulated the variables independently.
-Instead, any relationship between the two variables is merely the result of spatial autocorrelation.
+where $\Sigma$ is a correlation matrix proportional to either geographic or linguistic proximities between nations, and $\lambda$ and $\rho$ are autocorrelation parameters that represent the expected spatial or cultural phylogenetic signal for outcome and predictor variables, respectively.
+Importantly, in this simulation, we know that there is no direct causal relationship between $y$ and $x$ because we simulate the variables independently.
+Instead, any relationship between the two variables is merely the result of autocorrelation.
 
-We set the autocorrelation parameters to either 0.1 (weak), 0.5 (moderate), or 0.9 (strong).
+We set the autocorrelation parameters to either 0.2 (weak), 0.5 (moderate), or 0.8 (strong).
 We simulated 100 datasets for each parameter combination, resulting in 900 datasets.
-Each dataset had 178 rows, with the following associated data for each nation: latitude, longitude, continent (Asia, North America, Europe, Africa, South America, or Oceania), and legal origin (Islamic, Common, French, German, Mixed, or Scandinavian).
-Global data on legal origins was taken from a previous study[@Klerman2011].
+Each dataset had 236 rows representing different nations, with the following associated data for each nation: latitude, longitude, continent (Africa, Asia, Europe, North America, Oceania, or South America), and language family of the nation's majority spoken language (Afro-Asiatic, Atlantic-Congo, Austroasiatic, Austronesian, Eskimo-Aleut, Indo-European, Japonic, Kartvelian, Koreanic, Mande, Mongolic-Khitan, Nilotic, Nuclear Trans New Guinea, Sino-Tibetan, Tai-Kadai, Tupian, Turkic, or Uralic).
 
-With the resulting simulated datasets, we standardised outcome and predictor variables and fitted nine different models: (1) naive regression without controls, (2) regression with latitude control, (3) regression with longitude control, (4) regression with continent fixed effects, (5) regression with legal origin fixed effects, (6) regression employing Conley standard errors with 100 km cutoff, (7) regression employing Conley standard errors with 1,000 km cutoff, (8) regression employing Conley standard errors with 10,000 km cutoff, and (9) Bayesian regression including a Gaussian process over latitudes and longitudes.
+With the resulting simulated datasets, we standardised outcome and predictor variables and fitted eleven different models: (1) naive regression without controls, (2) regression with latitude control, (3) regression with longitude control, (4) regression with continent fixed effects, (5) regression with language family fixed effects, (6) regression employing Conley standard errors with 100 km cutoff, (7) regression employing Conley standard errors with 1,000 km cutoff, (8) regression employing Conley standard errors with 10,000 km cutoff, (9) Bayesian regression including a Gaussian process over latitudes and longitudes, (10) Bayesian regression including random intercepts covarying according to linguistic proximity, and (11) Bayesian regression including both a Gaussian process over latitudes and longitudes and random intercepts covarying according to linguistic proximity.
 
-Models employing Conley standard errors required latitude and longitude values, and the cutoffs implied the distance beyond which spatial autocorrelation is assumed to be zero.
+Models employing Conley standard errors required latitude and longitude values, and the cutoffs implied the distance beyond which autocorrelation is assumed to be zero.
 These models were fitted using the _conleyreg_ R package[@Duben2021].
-Bayesian Gaussian process models were fitted using the _brms_ R package[@Burkner2017].
+Bayesian models were fitted using the _brms_ R package[@Burkner2017].
 Our choice of priors was based on prior predictive simulation.
@@
 \endgroup
+
+\newpage
+
+# (APPENDIX) Appendix {-}
+
+```{r child = "appendix.Rmd"}
+```

### 2022-03-31 manuscript.tex
updated manuscript

@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
 \PassOptionsToPackage{hyphens}{url}
@@
 %
-\documentclass[
-  english,
-  man,floatsintext]{apa6}
+\documentclass[english,man,floatsintext]{apa6}
 \usepackage{lmodern}
@@
 \usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
+  \usepackage{textcomp} % provides euro and other symbols
+\else % if luatex or xelatex
   \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
 \fi
-% Use upquote if available, for straight quotes in verbatim environments
+% use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
+% use microtype if available
+\IfFileExists{microtype.sty}{%
 \usepackage[]{microtype}
@@
 }{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
 \IfFileExists{parskip.sty}{%
@@
 \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\usepackage{hyperref}
 \hypersetup{
-  pdftitle={Cross-national studies have not sufficiently accounted for non-independence between nations},
+            pdftitle={The non-independence of nations and why it matters},
             pdfauthor={Scott Claessens1 \& Quentin D. Atkinson1,2},
@@
             urlcolor=blue,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
 \usepackage{graphicx,grffile}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{0}
+
+% set default figure placement to htbp
 \makeatletter
@@
 \makeatother
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-% Make \paragraph and \subparagraph free-standing
-\ifx\paragraph\undefined\else
-  \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-  \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
+
 % Manuscript styling
@@
 % \newenvironment{ltable}
-%   {\begin{landscape}\centering\begin{threeparttable}}
-%   {\end{threeparttable}\end{landscape}}
-\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}
+%   {\begin{landscape}\begin{center}\begin{threeparttable}}
+%   {\end{threeparttable}\end{center}\end{landscape}}
+\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}
 
@@
 % \setlength{\parskip}{0pt plus 0pt minus 0pt}
+
+% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
+% See https://github.com/crsh/papaja/issues/292
+\makeatletter
+\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-1em}%
+  {\normalfont\normalsize\bfseries\itshape\typesectitle}}
+
+\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
+  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+  {-\z@\relax}%
+  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
+\makeatother
 
@@
 \makeatother
-\shorttitle{Non-independence between nations}
-\keywords{spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses\newline\indent Word count: 5549 words}
+
+\usepackage{xpatch}
+\makeatletter
+\xapptocmd\appendix
+  {\xapptocmd\section
+    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
+    {}{\InnerPatchFailed}%
+  }
+{}{\PatchFailed}
+\keywords{spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses\newline\indent Word count: 5500 words}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 \captionsetup[figure]{font={stretch=1.5}}
-\ifxetex
-  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[shorthands=off,main=english]{babel}
+\else
+  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
   \usepackage{polyglossia}
   \setmainlanguage[]{english}
-\else
-  \usepackage[shorthands=off,main=english]{babel}
 \fi
 
-\title{Cross-national studies have not sufficiently accounted for non-independence between nations}
+\title{The non-independence of nations and why it matters}
 \author{Scott Claessens\textsuperscript{1} \& Quentin D. Atkinson\textsuperscript{1,2}}
 \date{}
+
+\shorttitle{Non-independence of nations}
 
@@
 
-\abstract{
-Cross-national analyses can illuminate patternings of national-level outcomes around the world.
-But since nations are related via spatial proximity and shared cultural ancestry, cross-national analyses often violate statistical assumptions of non-independence, inflating false positive rates.
-Here, we show that past work in economics and psychology has not sufficiently accounted for non-independence between nations.
-In a review of the 100 highest-cited cross-national studies of economic development and cultural values, we find that controls for non-independence are rare and have not increased in frequency over the past 30 years.
-When studies do attempt to control for non-independence, our simulations suggest that the methods used continue to produce false positives.
-In replications of a subsample of twelve cross-national relationships, we show that half fail to replicate when controlling for spatial and/or cultural non-independence using global proximity matrices.
-We urge social scientists to sufficiently control for non-independence when reporting cross-national associations.
+\abstract{%
+Cross-national analyses can illuminate global patternings of national-level outcomes.
+But since nations are connected in various ways, such as via spatial proximity and shared cultural ancestry, cross-national analyses often violate assumptions of non-independence, inflating false positive rates.
+Here, we show that, despite being recognised as an important statistical pitfall for over 200 years, cross-national research in economics and psychology still does not sufficiently account for non-independence.
+In a review of the 100 highest-cited cross-national studies of economic development and values, we find that controls for non-independence are rare and have remained so over the past 30 years.
+When studies do include controls for non-independence, our simulations suggest that commonly used methods continue to produce false positives.
+In reanalyses of twelve cross-national relationships, we show that half are no longer significant after controlling for non-independence using global proximity matrices.
+We urge social scientists to sufficiently control for non-independence in cross-national research.
 }
@@
 Given the importance of nations for structuring human behaviour, there is justifiably huge interest in statistical analyses that attempt to predict variation in national-level outcomes of all kinds.
-At the time of writing, a search in Google Scholar for the term \enquote{cross-national study} returned 180,000 hits.
+At the time of writing, a search in Web of Science for the terms \enquote{cross-national}, \enquote{cross-cultural}, or \enquote{cross-country} in titles and abstracts returned 80,000 unique hits.
 The standard practice for cross-national analyses is to conduct bivariate correlations or multiple regressions with individual data points representing different nations.
@@
 First, nations that are closer to one another tend to be more similar than distant nations.
-This phenomenon is known as spatial non-independence\textsuperscript{21}, and it occurs because nations in close spatial proximity share characteristics due to local cultural diffusion of ideas\textsuperscript{22} and common climates and environments\textsuperscript{21}.
+This phenomenon is known as spatial non-independence\textsuperscript{21}, and it occurs because nations in close spatial proximity share characteristics due to local cultural diffusion of ideas\textsuperscript{22} and regional variation in climate and environment\textsuperscript{21}.
 For example, the neighbouring African nations Zambia and Tanzania have similar levels of terrain ruggedness, which has been used to partially explain their similar levels of economic development\textsuperscript{23}.
@@
 Second, nations with shared cultural ancestry tend to be more similar than culturally unrelated nations.
-This is known as cultural phylogenetic non-independence\textsuperscript{25}, and occurs because related nations share cultural traits inherited via descent from a common ancestor.
+This is known as cultural phylogenetic non-independence\textsuperscript{25--27}, and occurs because related nations share cultural traits inherited via descent from a common ancestor.
 Shared cultural ancestry can result in a form of pseudoreplication, whereby multiple instances of the same trait across nations are merely duplicates of the ancestral original.
-For example, the related island nations New Zealand and Samoa express similar languages and customs due to cultural inheritance from a shared ancestor deep in the evolutionary history of Polynesian societies.
-More recently, shared ancestry is also crucial in explaining cultural similarities between colonial settlements and their colonisers, even those separated by large geographic distances (e.g.~Spain and Argentina).
+For example, despite being separated by over 1500 kilometres of ocean, the related island nations Tonga and Tuvalu express similar languages and customs due to cultural inheritance from a shared ancestor deep in the evolutionary history of Polynesian societies.
+More recently, shared ancestry is also crucial in explaining cultural similarities between colonial settlements and their colonisers (e.g.~Argentina and Spain).
 
 Spatial and cultural phylogenetic non-independence between nations make cross-national inference challenging.
-A fundamental assumption of regression analysis is that model residuals should be independently and identically distributed\textsuperscript{26}.
-But without accounting for spatial or cultural non-independence between nations, model residuals can show structure that remains unaccounted for, violating this assumption (Figure \ref{fig:spatialExample}).
-Treating nations as independent can thus inflate false positive rates\textsuperscript{27}, producing spurious \enquote{direct} relationships between variables that in fact only indirectly covary due to spatial or cultural connections (see Supplementary Figure \ref{fig:dag} for an example causal model).
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/spatialExample-1.pdf}
-\caption{\label{fig:spatialExample}\emph{An example of spatial non-independence in cross-national analysis.} National-level outcome and predictor variables for 25 nations are simulated with high levels of spatial non-independence.
-(top) Scatterplot visualising the data with fitted line from naive regression without accounting for spatial non-independence.
-Vertical line segments represent model residual errors, coloured by their sign and size.
-Countries are labelled with ISO 3166-1 alpha-2 letter codes.
-(bottom) Model residual errors plotted onto a world map.
-In this example, residuals clearly show spatial structure that remains unaccounted for.}
-\end{figure}
+A fundamental assumption of regression analysis is that model residuals should be independently and identically distributed\textsuperscript{28}.
+But without accounting for spatial or cultural non-independence between nations, model residuals can show structure that remains unaccounted for, violating this assumption.
+Treating nations as independent can thus inflate false positive rates\textsuperscript{29}, producing spurious \enquote{direct} relationships between variables that in fact only indirectly covary due to spatial or cultural connections\textsuperscript{30} (see Supplementary Figure \ref{fig:dag} for an example causal model).
 
 Non-independence between data points is widely acknowledged in fields that routinely deal with spatially or culturally structured data.
-In ecology and sociology, studies with regional-level data use a variety of autoregressive models designed to account for spatial patternings\textsuperscript{28,29}.
-In anthropology, the Standard Cross-Cultural Sample of 186 cultures was compiled to minimise the confounding effects of non-independence in comparisons of traditional societies\textsuperscript{30}, though spatial and cultural dependencies are difficult to remove entirely\textsuperscript{31,32}.
-Anthropologists also borrow phylogenetic comparative methods from evolutionary biology, such as phylogenetic least squares regression\textsuperscript{33}, when comparing traditional societies, treating culturally related societies in the same way as biologists treat genetically related species (e.g.~\textsuperscript{34,35}).
+In ecology and sociology, studies with regional-level data use a variety of autoregressive models designed to account for spatial patternings\textsuperscript{31,32}.
+In anthropology, researchers have recognised cultural non-independence as an important statistical pitfall for over 200 years, with issues of cultural pseudoreplication being identified in early comparative studies of marriage practices across societies\textsuperscript{25}.
+More recently, the Standard Cross-Cultural Sample of 186 cultures was compiled to minimise the confounding effects of non-independence in comparisons of traditional societies\textsuperscript{33}, though spatial and cultural dependencies are difficult to remove entirely\textsuperscript{34,35}.
+Anthropologists also borrow phylogenetic comparative methods from evolutionary biology, such as phylogenetic least squares regression\textsuperscript{36}, when comparing traditional societies, treating culturally related societies in the same way as biologists treat genetically related species (e.g.\textsuperscript{37,38}).
 
 At the national level, recent reanalyses have revealed that several cross-national relationships reported in economics and psychology do not hold when controlling for non-independence between nations.
-One study replicated 25 analyses of \enquote{persistence} in economics, in which modern national outcomes are regressed against historical characteristics of those nations, and found that over half of the relationships were attenuated when controlling for spatial non-independence\textsuperscript{36}.
-Another replication study found that many of the widely publicised relationships between national-level pathogen prevalence and political institutions and attitudes fail to hold when controlling for various kinds of non-independence\textsuperscript{37}.
-These reanalyses, and others\textsuperscript{38,39}, raise the question: how widespread a concern is non-independence in studies of national-level outcomes?
+One study replicated 25 analyses of \enquote{persistence} in economics, in which modern national outcomes are regressed against historical characteristics of those nations, and found that over half of the relationships were attenuated when controlling for spatial non-independence\textsuperscript{39}.
+Another replication study found that many of the widely publicised relationships between national-level pathogen prevalence and political institutions and attitudes fail to hold when controlling for various kinds of non-independence\textsuperscript{40}.
+These reanalyses, and others\textsuperscript{41--43}, raise the question: how widespread a concern is non-independence in studies of national-level outcomes?
 
@@
 In order to motivate our research question, it is important to first show that economic development and cultural values are indeed spatially and culturally non-independent around the world.
-To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for the Human Development Index\textsuperscript{40} and two dimensions of cultural values from the World Values Survey, traditional vs.~secular values and survival vs.~self-expression values\textsuperscript{16}.
+To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for the Human Development Index\textsuperscript{44} and two dimensions of cultural values from the World Values Survey, traditional vs.~secular values and survival vs.~self-expression values\textsuperscript{16}.
 For all three of these variables, we found that a substantial proportion of national-level variation was explained by spatial proximity and shared cultural ancestry between nations (Figure \ref{fig:plotSignal}).
+In our models, shared cultural ancestry explained over half of the national-level variation in economic development and cultural values, with spatial proximity explaining most of the remaining national-level variance.
 Bayes Factors indicated strong evidence that these estimates of geographic and cultural phylogenetic signal differed from zero, aside from the geographic signal for survival values, for which the evidence was equivocal (see Supplementary Results).
-These findings emphasise the need to account for spatial and cultural non-independence in cross-national analyses of economic development and cultural values.
+These findings emphasise the need to account for spatial and cultural phylogenetic non-independence in cross-national analyses of economic development and cultural values.
 
@@
 We find that 42\% of economic development articles contain at least one attempt to control for non-independence (95\% bootstrap confidence interval {[}0.30 0.54{]}), while this proportion decreases to only 8\% for cultural values articles (95\% bCI {[}0.02 0.16{]}).
-Both kinds of article are most likely to use regional fixed effects (e.g.~continent fixed effects) to account for non-independence, but some articles also include controls for spatial distance (e.g.~latitude) and shared cultural history (e.g.~legal origin fixed effects).
+Both kinds of article are most likely to use regional fixed effects (e.g.~continent fixed effects) to account for non-independence, but some articles also include controls for spatial distance (e.g.~latitude) and shared cultural history (e.g.~colony status).
 
@@
 Focusing on the full sample of 4,308 analyses, we find that the proportion of individual cross-national analyses accounting for non-independence is even lower (Figure \ref{fig:plotReview}b).
-Across 2,487 cross-national analyses from studies of economic development, only 5\% are estimated to control for non-independence (95\% credible interval {[}0.02 0.14{]}).
+Across 2,487 cross-national analyses from studies of economic development, only 5\% are estimated to control for non-independence (95\% credible interval {[}0.02 0.15{]}).
 Similarly, across 1,821 cross-national analyses from studies of cultural values, only 1\% are estimated to control for non-independence (95\% CI {[}0.00 0.02{]}).
@@
 To compare the efficacy of different methods in the literature, we conducted a simulation study.
-We simulated national-level datasets (n = 178 nations) with varying degrees of spatial autocorrelation (i.e.~non-independence) for outcome and predictor variables, but with no direct causal relationship between the variables.
-We then fitted naive regressions without controls to these datasets, as well as regression models with the following controls: latitude, longitude, continent fixed effects, and legal origin fixed effects.
-Despite not being identified in our systematic review, we also included in our simulation study Conley standard errors, a widely used standard error correction that purportedly accounts for spatial non-independence\textsuperscript{41,42}.
-Finally, we included a multilevel model of spatial non-independence that has recently been made widely available in statistical software packages\textsuperscript{43,44}: Bayesian Gaussian process regression\textsuperscript{45,46}.
-This method allows nation random intercepts to covary according to a global geographic proximity matrix.
+We simulated national-level datasets (n = 236 nations) with varying degrees of spatial or cultural phylogenetic autocorrelation (i.e.~non-independence) for outcome and predictor variables, but with no direct causal relationship between the variables.
+We then fitted naive regressions without controls to these datasets, as well as regression models with controls for latitude, longitude, and continent fixed effects.
+Despite not being identified in our systematic review, we also included fixed effects for the language families of the majority-spoken languages in each country, as this control is often used to account for cultural phylogenetic non-independence (e.g.\textsuperscript{45}).
+In addition, we included Conley standard errors, a widely used standard error correction that purportedly accounts for spatial non-independence\textsuperscript{46,47}.
+Finally, we included Bayesian multilevel regressions that explicitly model spatial and/or cultural phylogenetic non-independence by allowing nation random intercepts to covary according to global geographic and/or linguistic proximity matrices (see Supplementary Methods).
+To model spatial non-independence we included a Gaussian process\textsuperscript{48,49} over latitude and longitude values, and to model cultural phylogenetic non-independence we assumed that nation random intercepts were correlated proportional to their linguistic proximity\textsuperscript{50}.
 Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero (i.e.~falsely infering a relationship when none is present).
 
-Figure \ref{fig:plotSim2} plots the estimated false positive rates from our simulation study, split by different methods and different degrees of spatial autocorrelation (see Supplementary Figure \ref{fig:plotSim1} for full distribution of effect sizes under moderate autocorrelation).
-With at least moderate degrees of spatial autocorrelation for both outcome and predictor variables, naive regression models (Figure \ref{fig:plotSim2} top left) produce false positive rates above chance levels.
-With strong spatial autocorrelation for both outcomes and predictors, false positive rates reach as high as 81\%.
+Figures \ref{fig:plotSim3} and \ref{fig:plotSim4} plot the estimated false positive rates from our simulation study, split by different methods and different degrees of spatial or cultural phylogenetic autocorrelation (see Supplementary Figures \ref{fig:plotSim1} and \ref{fig:plotSim2} for full distributions of effect sizes under strong autocorrelation).
+For reference, \enquote{weak} autocorrelation in our simulation is comparable to the geographic signal for survival values in Figure \ref{fig:plotSignal}, while \enquote{moderate} and \enquote{strong} levels of autocorrelation are comparable to the cultural phylogenetic signal for traditional and survival values, respectively.
+
+Our simulation study revealed that with at least moderate degrees of spatial or cultural phylogenetic autocorrelation for both outcome and predictor variables, naive regression models produce false positive rates above chance levels.
+This false positive rate increases as the degree of autocorrelation increases.
+With strong spatial autocorrelation for both outcomes and predictors, false positive rates reach as high as 71\%.
+We find a slightly lower false positive rate under strong cultural phylogenetic autocorrelation, though this false positive rate is still greater than expected by chance (38\%).
 
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/plotSim2-1.pdf}
-\caption{\label{fig:plotSim2}\emph{False positive rates for different methods of controlling for non-independence in our simulation study.} For simulated outcome and predictor variables, we systematically varied the strength of spatial autocorrelation, from weak (0.1) to moderate (0.5) to strong (0.9).
+\includegraphics{manuscript_files/figure-latex/plotSim3-1.pdf}
+\caption{\label{fig:plotSim3}\emph{False positive rates for different methods of controlling for spatial non-independence in our simulation study.} For simulated outcome and predictor variables, we systematically varied the strength of spatial autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
 We simulated 100 datasets per parameter combination and fitted different models to each dataset.
@@
 
-Common methods in the literature do not reduce this high false positive rate.
-With strong spatial autocorrelation for both outcome and predictor variables, false positive rates remain above 70\% when controlling for latitude, longitude, continent fixed effects, and legal origin fixed effects.
-Applying Conley standard errors with cutoffs of 100 km or 1,000 km also does not reduce false positive rates below 70\% under strong spatial autocorrelation.
-With a cutoff of 10,000 km, the false positive rate under strong spatial autocorrelation reduces to 57\%, though this correction adds bias at weaker levels of spatial autocorrelation, increasing false positive rates above chance levels.
-
-By contrast, Bayesian Gaussian process regression adequately accounts for non-independence across a range of simulation parameters.
-At moderate levels of spatial autocorrelation on both outcomes and predictors, Gaussian process regression eliminates false positives, such that the false positive rate is no different from chance.
-With strong spatial autocorrelation on both outcomes and predictors, the false positive rate is reduced to 30\%, substantially less than any other method.
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotSim4-1.pdf}
+\caption{\label{fig:plotSim4}\emph{False positive rates for different methods of controlling for cultural phylogenetic non-independence in our simulation study.} For simulated outcome and predictor variables, we systematically varied the strength of cultural phylogenetic autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+We simulated 100 datasets per parameter combination and fitted different models to each dataset.
+False positive rates were operationalised as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero.
+Point ranges represent proportions and 95\% bootstrap confidence intervals, and dashed lines indicate the 5\% false positive rate that is expected due to chance.
+SEs = standard errors.}
+\end{figure}
+
+Common methods in the literature do not reduce these high false positive rates.
+With strong spatial autocorrelation for both outcome and predictor variables, false positive rates remain above 50\% when controlling for latitude, longitude, and language family fixed effects (Figure \ref{fig:plotSim3}).
+Applying Conley standard errors also does not reduce false positive rates below 50\% under strong spatial autocorrelation, regardless of the distance cutoff.
+Continent fixed effects are more effective than other frequentist methods, though they continue to produce a false positive rate of 28\% under strong spatial autocorrelation.
+By contrast, Bayesian spatial Gaussian process regression with longitude and latitude values outperforms all other methods.
+This approach eliminates false positives under moderate spatial autocorrelation, such that the false positive rate is no different from chance, and reduces the false positive rate under strong spatial autocorrelation to 17\%.
+Bayesian models that additionally account for linguistic proximity between nations perform equally well, though models with only linguistic covariance continue to produce false positives.
+
+In our simulation of cultural phylogenetic non-independence, we find that most frequentist methods do not reduce false positive rates (Figure \ref{fig:plotSim4}).
+Controls for latitude and longitude, continent fixed effects, and Conley standard errors do little to change false positive rates.
+Language family fixed effects are slightly more effective than other frequentist methods, though they continue to produce a false positive rate of 29\% under strong cultural phylogenetic autocorrelation.
+By contrast, Bayesian models with random effects covarying according to linguistic proximity completely eliminate false positives across all degrees of cultural phylogenetic autocorrelation.
+Bayesian models that additionally account for geographic proximity between nations perform equally well, though models with only a spatial Gaussian process continue to produce false positives.
 
@@
 
-Our systematic review and simulation study have showed that controls for non-independence are rare in cross-national studies of economic development and cultural values, and when studies do attempt to control for non-independence, the methods used continue to produce false positives.
+Our systematic review and simulation study have shown that controls for non-independence are rare in cross-national studies of economic development and cultural values, and when studies do attempt to control for non-independence, the methods used continue to produce false positives.
 This raises the worrying possibility that the cross-national literature in economics and psychology is populated with spurious relationships.
 
-To diagnose how widespread this issue of spurious cross-national relationships might be, we reanalysed a subset of twelve previous cross-national analyses from our systematic review, sufficiently controlling for spatial and cultural phylogenetic non-independence using global geographic and linguistic proximity matrices.
-We subsampled six analyses from our economic development review\textsuperscript{47--52} and six from our cultural values review\textsuperscript{13,14,16,53--55}.
+To determine how widespread this issue of spurious cross-national relationships might be, we reanalysed a subset of twelve previous cross-national analyses from our systematic review, sufficiently controlling for spatial and cultural phylogenetic non-independence using global geographic and linguistic proximity matrices.
+We subsampled six analyses from our economic development review\textsuperscript{51--56} and six from our cultural values review\textsuperscript{13,14,16,57--59}.
 Our choice of analyses was constrained by data availability and whether we were able to initially replicate the original finding.
@@
 Figure \ref{fig:plotReplications1} visualises the results of our reanalysis.
-Overall, cross-national correlation effect sizes tended to reduce when controlling for statistical non-independence between nations, sometimes by as much as half of the original effect size.
-Based on 95\% credible intervals, six out of twelve cross-national relationships are attenuated when controlling for non-independence.
-For the economic development analyses, four out of six cross-national relationships are attenuated when controlling for spatial non-independence.
-For the cultural values analyses, two out of six are attenuated when controlling for cultural phylogenetic non-independence.
+Cross-national correlation effect sizes tended to reduce when controlling for statistical non-independence between nations, sometimes by as much as half of the original effect size.
+Overall, after controlling for non-independence, six out of twelve cross-national associations had 95\% credible intervals that included zero.
+For the economic development analyses, four out of six cross-national relationships had 95\% credible intervals including zero when controlling for spatial non-independence.
+For the cultural values analyses, two out of six cross-national relationships had 95\% credible intervals including zero when controlling for cultural phylogenetic non-independence.
 Supplementary Figure \ref{fig:plotReplications2} shows these cross-national correlations plotted against the raw data.
@@
 All outcome and predictor variables are standardised.
-Most analyses are simple bivariate cross-national correlations, but Gelfand et al.~(2011) is a partial correlation controlling for log gross national income and Adamczyk and Pitt (2009) is a multilevel model including several covariates.
+Most analyses are simple bivariate cross-national correlations, but Gelfand et al.
+(2011) is a partial correlation controlling for log gross national income and Adamczyk and Pitt (2009) is a multilevel model including several covariates.
 Point ranges represent posterior medians and 95\% credible intervals.
@@
 For each outcome variable, our Gaussian process models provided varying estimates of how quickly spatial autocorrelation declined with distance (Supplementary Figure \ref{fig:plotReplications3}).
-For example, in Skidmore and Toya\textsuperscript{52} gross domestic product growth was only moderately spatially autocorrelated at 1000 km distance (posterior median spatial autocorrelation at 1000 km = 0.43, 95\% CI {[}0.05 0.96{]}), whereas in Inglehart and Baker\textsuperscript{16} traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1000 km = 0.96, 95\% CI {[}0.77 0.99{]}).
-We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \ref{fig:plotReplications4}), with some outcome variables expressing low signal (e.g.~Gelfand et al.\textsuperscript{14}; posterior median = 0.17, 95\% CI {[}0.00 0.93{]}) and others expressing high signal (e.g.~Alesina et al.\textsuperscript{54}; posterior median = 0.89, 95\% CI {[}0.66 0.98{]}).
+For example, in Skidmore and Toya\textsuperscript{56} gross domestic product growth was only moderately spatially autocorrelated at 1000 km distance (posterior median spatial autocorrelation at 1000 km = 0.47, 95\% CI {[}0.06 0.97{]}), whereas in Inglehart and Baker\textsuperscript{16} traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1000 km = 0.96, 95\% CI {[}0.78 0.99{]}).
+We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \ref{fig:plotReplications4}), with some outcome variables expressing low signal (e.g.~tightness\textsuperscript{14}; posterior median = 0.11, 95\% CI {[}0.00 0.82{]}) and others expressing high signal (e.g.~female labour force participation\textsuperscript{58}; posterior median = 0.91, 95\% CI {[}0.67 0.99{]}).
 Across all analyses, we found that stronger estimates of spatial autocorrelation or cultural phylogenetic signal resulted in a more pronounced reduction in the effect size when controlling for non-independence between nations (Figure \ref{fig:plotReplications5}).
@@
 
-Non-independence is a concern because economic development and cultural values are highly geographically and culturally structured around the world, as revealed by our initial modelling.
-These results build on previous work.
-One previous study suggested that geographic proximity is more important than deep cultural ancestry in explaining the distribution of human development across Eurasian nations\textsuperscript{56}.
-By contrast, with a global sample of nations, we found strong cultural phylogenetic signal, as well as geographic signal, for the Human Development Index.
+Our initial analyses add to and clarify existing evidence regarding the non-independence of economic and cultural variation among nations.
+One previous study suggested that geographic proximity is more important than deep cultural ancestry in explaining the distribution of human development across Eurasian nations, though the authors noted that their small sample of 44 nations and regional focus limited their statistical power\textsuperscript{60}.
+Our global sample of nations revealed strong cultural phylogenetic signal, as well as geographic signal, for the Human Development Index.
 Another previous study found that similarities in the cultural values of nations are predicted by linguistic, but not geographic, distances between those nations\textsuperscript{6}.
@@
 Yet our systematic literature review and simulation study revealed that the most commonly used controls for non-independence do not sufficiently deal with the issue.
-In our simulations, controlling for either latitude or longitude did not reduce false positive rates, calling into question controls like distance to the equator in regression models.
-High false positive rates persisted with Conley standard errors, which have previously been critiqued for being overly sensitive to arbitrary distance cutoffs\textsuperscript{36}.
-The simulation also confirmed the verbal claim that fixed effects for spatial or cultural groupings (e.g.~continent or legal origin fixed effects) are insufficient because non-independence still remains within groupings\textsuperscript{37}.
-This logic applies to analyses that control for non-independence by including language family fixed effects (e.g.~\textsuperscript{57}) or separately analysing different regions (e.g.~\textsuperscript{58}).
-The only method in our simulation that sufficiently reduced the false positive rate was Bayesian Gaussian process regression, though we did not include other possible controls for non-independence, such as conditional autoregressive models\textsuperscript{28} or generalised additive models\textsuperscript{59}.
+In our simulations, controlling for either latitude or longitude did not reduce false positive rates.
+This result calls into question controls like distance to the equator to account for non-independence in cross-national regression models, though these controls maybe still be suitable to account for regional or latitudinal variation in ecology, which we did not simulate.
+High false positive rates persisted with Conley standard errors, which have previously been critiqued for being overly sensitive to arbitrary distance cutoffs\textsuperscript{39}.
+The simulation also confirmed the verbal claim that fixed effects for spatial or cultural groupings (e.g.~continent or language family fixed effects) are insufficient because non-independence still remains within groupings\textsuperscript{40}.
+This logic further applies to analyses that control for non-independence by separately analysing different regions (e.g.\textsuperscript{61}).
+Across all model types in our simulation, the only methods that sufficiently reduced the false positive rate were the Bayesian multilevel regressions that explicitly modelled spatial and cultural phylogenetic autocorrelation, though we did not include other possible controls for non-independence, such as conditional autoregressive models\textsuperscript{31} or generalised additive models\textsuperscript{62}.
 
 Ours is not the first review to show that studies are misapplying statistical methods in ways that inflate false positive rates.
-For example, other literature reviews have shown that studies in the social sciences tend to use small sample sizes\textsuperscript{60}, treat ordinal data as metric\textsuperscript{61}, incorrectly handle missing values\textsuperscript{62}, and ignore best practices in meta-analyses\textsuperscript{63}.
+For example, other literature reviews have shown that studies in the social sciences tend to use small sample sizes\textsuperscript{63}, treat ordinal data as metric\textsuperscript{64}, incorrectly handle missing values\textsuperscript{65}, and ignore best practices in meta-analyses\textsuperscript{66}.
 Why do cross-national studies also rarely account for non-independence?
-At the institutional level, one possibility is that such practices are incentivised \emph{because} they inflate false positive rates, increasing the probability of getting published and promoted\textsuperscript{60}.
-At the individual level, another possibility is that researchers outside of anthropology and ecology are simply not aware of the problem, or believe that the problem does not apply at the national level.
+At the institutional level, one possibility is that such practices are incentivised \emph{because} they inflate false positive rates, increasing the probability of getting published and promoted\textsuperscript{63}.
+At the individual level, another possibility is that researchers outside of anthropology and ecology are simply not aware of the problem, or believe that the problem does not apply to analyses of nations.
 Even if researchers appreciate the problem, they might not know of suitable controls or perceive the methods to be too complex.
@@
 These institutional- and individual-level barriers can be combatted.
-First, cross-national replication studies like ours and others\textsuperscript{36--39}, combined with the methodological reviews included in Registered Reports\textsuperscript{64}, might change incentive structures and encourage researchers to analyse the world's nations with more rigorous methods.
+First, cross-national replication studies like ours and others\textsuperscript{39--43}, combined with the methodological reviews included in Registered Reports\textsuperscript{67}, might change incentive structures and encourage researchers to analyse the world's nations with more rigorous methods.
 Second, since the issue of non-independence is fundamentally an issue of causal inference (Supplementary Figure \ref{fig:dag}), more explicit descriptions of causal models could promote controls for non-independence in cross-national research.
 In our review, economists studying economic development dealt with national-level non-independence more than psychologists studying cultural values, likely because economics studies tend to be lengthy statistical exercises that systematically incorporate or exclude numerous variables in an attempt to infer causation.
-Third, the recent widespread accessibility of open source statistical software, such as the programming language Stan\textsuperscript{44} and the R package \emph{brms}\textsuperscript{43}, should promote the use of more rigorous methods to control for non-independence.
+Third, the recent widespread accessibility of open source statistical software, such as the programming language Stan\textsuperscript{68} and the R package \emph{brms}\textsuperscript{69}, should promote the use of more rigorous methods to control for non-independence.
 Using \emph{brms}, for example, Bayesian Gaussian process regression is straightforward to conduct, requiring only longitude and latitude values for nations.
@@
 Until such changes are implemented and sufficient controls for non-independence are the norm, existing cross-national correlations should be interpreted with caution.
-In our reanalyses, we found that four out of six cross-national analyses of economic development were attenuated when controlling for spatial non-independence.
-Three of these were tests of \enquote{persistence} hypotheses, studying the effects of historical and environmental conditions --- settler mortality\textsuperscript{47}, wheat-sugar suitability\textsuperscript{51}, and natural disaster frequency\textsuperscript{52} --- on modern developmental outcomes.
-A recent reanalysis has also called into question various studies of this ilk\textsuperscript{36}.
-We also found that two out of six cross-national analyses of cultural values were attenuated when controlling for cultural phylogenetic non-independence.
-It is interesting that the economic development analyses were attenuated with spatial controls, while the cultural values analyses were attenuated by cultural phylogenetic controls.
+In our reanalyses, we found that four out of six cross-national correlations with economic development variables had 95\% credible intervals that included zero when controlling for spatial non-independence.
+Three of these analyses were tests of \enquote{persistence} hypotheses, studying the effects of historical and environmental conditions --- settler mortality\textsuperscript{51}, wheat-sugar suitability\textsuperscript{55}, and natural disaster frequency\textsuperscript{56} --- on modern developmental outcomes.
+A recent reanalysis has also called into question various studies of this ilk\textsuperscript{39}.
+We also found that two out of six cross-national correlations with cultural values variables had 95\% credible intervals that included zero when controlling for cultural phylogenetic non-independence.
+It is interesting that the economic development analyses were more affected by the spatial controls, while the cultural values analyses were more affected by the cultural phylogenetic controls.
 This highlights that certain kinds of non-independence will be more important depending on the particular national-level outcome and predictor variables of interest.
@@
 We do not wish to dissuade researchers from conducting cross-national studies.
-On the contrary, such work promises to deepen our understanding of human populations beyond Western, educated, industrialised, rich, and democratic societies\textsuperscript{7}, providing the representative samples necessary to test evolutionary and socio-ecological theories of human behaviour\textsuperscript{8,65}.
+On the contrary, such work promises to deepen our understanding of the causes and consequences of economic development and cultural values around the world.
+Moreover, cross-national studies allow social scientists to broaden their scope of study beyond Western populations\textsuperscript{7}, providing the representative samples necessary to test evolutionary and socio-ecological theories of human behaviour\textsuperscript{8,70}.
 But in order to minimise spurious relationships in global datasets, we urge researchers to control for spatial and cultural phylogenetic non-independence when reporting cross-national correlations.
-Nations are not independent, and our statistical models should reflect this.
+Nations are not independent, and our statistical models must reflect this.
 
@@
 
-To estimate the degree of spatial and cultural phylogenetic non-independence in economic development and cultural values, we calculated geographic and linguistic signal for global measures of development and values.
-Our measure of economic development was the Human Development Index\textsuperscript{40}.
+To estimate the degree of spatial and cultural phylogenetic non-independence in economic development and cultural values, we calculated geographic and cultural phylogenetic signal for global measures of development and values.
+Our measure of economic development was the Human Development Index\textsuperscript{44}.
 We retrieved a longitudinal dataset capturing human development for 189 nations since 1990 (n = 1,512; \url{https://hdr.undp.org/en/content/download-data}).
@@
 We included these matrices in Bayesian multilevel models, allowing nation random intercepts to covary according to both geographic and linguistic proximity simultaneously.
-These models were fitted with the R package \emph{brms}\textsuperscript{43} and converged normally (\(\hat{R}\) \textless{} 1.1).
-Estimates of geographic and linguistic signal were computed as the proportion of national-level variance in these models explained by geographic and linguistic proximity matrices.
+These models were fitted with the R package \emph{brms}\textsuperscript{69} and converged normally (\(\hat{R}\) \textless{} 1.1).
+Estimates of geographic and cultural phylogenetic signal were computed as the proportion of national-level variance in these models explained by geographic and linguistic proximity matrices.
 
@@
 We coded common attempts to control for non-independence between nations.
-These included: (1) any higher-level control variables for spatial regional groupings (e.g.~continent fixed effects); (2) any geographic distance control variables (e.g.~distance between capital cities, distance from equator, latitude); (3) any control variables capturing shared cultural history (e.g.~former colony, legal origin fixed effects, linguistic history, cultural influence); and (4) any other control variables, tests, or approaches that were deemed as attempts to control for non-independence (e.g.~eigenvector filtering\textsuperscript{66}, controls for trade-weightings between countries, cross-sectional dependence tests\textsuperscript{67}, separate analyses for subsets of countries).
+These included: (1) any higher-level control variables for spatial regional groupings (e.g.~continent fixed effects); (2) any geographic distance control variables (e.g.~distance between capital cities, distance from equator, latitude); (3) any control variables capturing shared cultural history (e.g.~former colony, legal origin fixed effects, linguistic history, cultural influence); and (4) any other control variables, tests, or approaches that were deemed as attempts to control for non-independence (e.g.~eigenvector filtering\textsuperscript{71}, controls for trade-weightings between countries, cross-sectional dependence tests\textsuperscript{72}, separate analyses for subsets of countries).
 These were coded by the first author.
@@
 To test for a trend over time, we also fitted a Bayesian multilevel logistic regression with a multigroup spline for year of publication and random intercepts for articles.
-Bayesian models were fitted with the \emph{brms} R package\textsuperscript{43}.
+Bayesian models were fitted with the \emph{brms} R package\textsuperscript{69}.
 Our priors were informed by prior predictive checks, and all models converged normally (\(\hat{R}\) \textless{} 1.1).
@@
 
-To simulate data, we used our global geographic proximity matrix (see Supplementary Methods) as a weight matrix \(W\), scaling the matrix between 0 and 1 and restricting the matrix to 178 nations with available external data (see below).
-Using this weight matrix, we simulated outcome \(Y\) and predictor \(X\) variables using the following formulae from a previous simulation study\textsuperscript{68}:
+We simulated data for 236 nations with varying degrees of spatial or cultural phylogenetic signal for outcome \(y\) and predictor \(x\) using the following generative model:
 
@@
 
-where \(\zeta\), \(\nu\), \(\xi\), and \(\omega\) are all standardised normal random variables, and \(\lambda\) and \(\rho\) are autocorrelation parameters that vary between 0 and 1, reflecting the strength of spatial autocorrelation for outcome and predictor variables, respectively.
-Importantly, in this simulation, we know that there is no direct causal relationship between \(X\) and \(Y\) because we simulated the variables independently.
-Instead, any relationship between the two variables is merely the result of spatial autocorrelation.
-
-We set the autocorrelation parameters to either 0.1 (weak), 0.5 (moderate), or 0.9 (strong).
+where \(\Sigma\) is a correlation matrix proportional to either geographic or linguistic proximities between nations, and \(\lambda\) and \(\rho\) are autocorrelation parameters that represent the expected spatial or cultural phylogenetic signal for outcome and predictor variables, respectively.
+Importantly, in this simulation, we know that there is no direct causal relationship between \(y\) and \(x\) because we simulate the variables independently.
+Instead, any relationship between the two variables is merely the result of autocorrelation.
+
+We set the autocorrelation parameters to either 0.2 (weak), 0.5 (moderate), or 0.8 (strong).
 We simulated 100 datasets for each parameter combination, resulting in 900 datasets.
-Each dataset had 178 rows, with the following associated data for each nation: latitude, longitude, continent (Asia, North America, Europe, Africa, South America, or Oceania), and legal origin (Islamic, Common, French, German, Mixed, or Scandinavian).
-Global data on legal origins was taken from a previous study\textsuperscript{69}.
-
-With the resulting simulated datasets, we standardised outcome and predictor variables and fitted nine different models: (1) naive regression without controls, (2) regression with latitude control, (3) regression with longitude control, (4) regression with continent fixed effects, (5) regression with legal origin fixed effects, (6) regression employing Conley standard errors with 100 km cutoff, (7) regression employing Conley standard errors with 1,000 km cutoff, (8) regression employing Conley standard errors with 10,000 km cutoff, and (9) Bayesian regression including a Gaussian process over latitudes and longitudes.
-
-Models employing Conley standard errors required latitude and longitude values, and the cutoffs implied the distance beyond which spatial autocorrelation is assumed to be zero.
-These models were fitted using the \emph{conleyreg} R package\textsuperscript{70}.
-Bayesian Gaussian process models were fitted using the \emph{brms} R package\textsuperscript{43}.
+Each dataset had 236 rows representing different nations, with the following associated data for each nation: latitude, longitude, continent (Africa, Asia, Europe, North America, Oceania, or South America), and language family of the nation's majority spoken language (Afro-Asiatic, Atlantic-Congo, Austroasiatic, Austronesian, Eskimo-Aleut, Indo-European, Japonic, Kartvelian, Koreanic, Mande, Mongolic-Khitan, Nilotic, Nuclear Trans New Guinea, Sino-Tibetan, Tai-Kadai, Tupian, Turkic, or Uralic).
+
+With the resulting simulated datasets, we standardised outcome and predictor variables and fitted eleven different models: (1) naive regression without controls, (2) regression with latitude control, (3) regression with longitude control, (4) regression with continent fixed effects, (5) regression with language family fixed effects, (6) regression employing Conley standard errors with 100 km cutoff, (7) regression employing Conley standard errors with 1,000 km cutoff, (8) regression employing Conley standard errors with 10,000 km cutoff, (9) Bayesian regression including a Gaussian process over latitudes and longitudes, (10) Bayesian regression including random intercepts covarying according to linguistic proximity, and (11) Bayesian regression including both a Gaussian process over latitudes and longitudes and random intercepts covarying according to linguistic proximity.
+
+Models employing Conley standard errors required latitude and longitude values, and the cutoffs implied the distance beyond which autocorrelation is assumed to be zero.
+These models were fitted using the \emph{conleyreg} R package\textsuperscript{73}.
+Bayesian models were fitted using the \emph{brms} R package\textsuperscript{69}.
 Our choice of priors was based on prior predictive simulation.
@@
 
-The twelve analyses that we settled on\textsuperscript{13,14,16,47--55} were mostly bivariate cross-national correlations, except for two.
-One analysis\textsuperscript{14} additionally controlled for log gross national income, and another analysis\textsuperscript{52} is a multilevel model including random intercepts for nations and several individual-level and national-level covariates (see Model 5 in original paper).
+The twelve analyses that we settled on\textsuperscript{13,14,16,51--59} were mostly bivariate cross-national correlations, except for two.
+One analysis\textsuperscript{14} additionally controlled for log gross national income, and another analysis\textsuperscript{56} is a multilevel model including random intercepts for nations and several individual-level and national-level covariates (see Model 5 in original paper).
 Before running any additional models, we pre-registered these twelve analyses on the Open Science Framework on 25\textsuperscript{th} January 2022 (\url{https://osf.io/uywx8/}).
@@
 
-We fitted these models using the \emph{brms} R package\textsuperscript{43}.
+We fitted these models using the \emph{brms} R package\textsuperscript{69}.
 Our choice of priors was based on prior predictive simulation.
-All models converged normally (\(\hat{R}\) \textless{} 1.1), though for some models we resorted to using approximate Gaussian processes\textsuperscript{71} to reach convergence.
+All models converged normally (\(\hat{R}\) \textless{} 1.1), though for some models we resorted to using approximate Gaussian processes\textsuperscript{74} to reach convergence.
 
@@
 All data and code are accesible at our Open Science Framework repository (\url{https://osf.io/uywx8/}).
-We used the \emph{targets} R package\textsuperscript{72} to create a reproducible data analysis pipeline and the \emph{papaja} R package\textsuperscript{73} to reproducibly generate the manuscript.
+We used the \emph{targets} R package\textsuperscript{75} to create a reproducible data analysis pipeline and the \emph{papaja} R package\textsuperscript{76} to reproducibly generate the manuscript.
 
@@
 
+\leavevmode\hypertarget{ref-Naroll1961}{}%
+26. Naroll, R. Two solutions to Galton's Problem. \emph{Philosophy of Science} \textbf{28}, 15--39 (1961).
+
+\leavevmode\hypertarget{ref-Naroll1965}{}%
+27. Naroll, R. Galton's Problem: The logic of cross-cultural analysis. \emph{Social Research} \textbf{32}, 428--451 (1965).
+
 \leavevmode\hypertarget{ref-Jarque1987}{}%
-26.
+28.
 Jarque, C.
@@
 \leavevmode\hypertarget{ref-Legendre1993}{}%
-27. Legendre, P. Spatial autocorrelation: Trouble or new paradigm? \emph{Ecology} \textbf{74}, 1659--1673 (1993).
+29. Legendre, P. Spatial autocorrelation: Trouble or new paradigm? \emph{Ecology} \textbf{74}, 1659--1673 (1993).
+
+\leavevmode\hypertarget{ref-Roberts2013}{}%
+30. Roberts, J., Seán AND Winters. Linguistic diversity and traffic accidents: Lessons from statistical studies of cultural traits. \emph{PLOS ONE} \textbf{8}, 1--13 (2013).
 
 \leavevmode\hypertarget{ref-Lichstein2002}{}%
-28. Lichstein, J. W., Simons, T. R., Shriner, S. A. \& Franzreb, K. E. Spatial autocorrelation and autoregressive models in ecology. \emph{Ecological Monographs} \textbf{72}, 445--463 (2002).
+31. Lichstein, J. W., Simons, T. R., Shriner, S. A. \& Franzreb, K. E. Spatial autocorrelation and autoregressive models in ecology. \emph{Ecological Monographs} \textbf{72}, 445--463 (2002).
 
 \leavevmode\hypertarget{ref-Loftin1983}{}%
-29. Loftin, C. \& Ward, S. K. A spatial autocorrelation model of the effects of population density on fertility. \emph{American Sociological Review} \textbf{48}, 121--128 (1983).
+32. Loftin, C. \& Ward, S. K. A spatial autocorrelation model of the effects of population density on fertility. \emph{American Sociological Review} \textbf{48}, 121--128 (1983).
 
 \leavevmode\hypertarget{ref-Murdock1969}{}%
-30. Murdock, G. P. \& White, D. R. Standard cross-cultural sample. \emph{Ethnology} \textbf{8}, 329--369 (1969).
+33. Murdock, G. P. \& White, D. R. Standard cross-cultural sample. \emph{Ethnology} \textbf{8}, 329--369 (1969).
 
 \leavevmode\hypertarget{ref-Dow2008}{}%
-31. Dow, M. M. \& Eff, E. A. Global, regional, and local network autocorrelation in the standard cross-cultural sample. \emph{Cross-Cultural Research} \textbf{42}, 148--171 (2008).
+34. Dow, M. M. \& Eff, E. A. Global, regional, and local network autocorrelation in the standard cross-cultural sample. \emph{Cross-Cultural Research} \textbf{42}, 148--171 (2008).
 
 \leavevmode\hypertarget{ref-Eff2004}{}%
-32. Eff, E. A. Does Mr. Galton still have a problem? Autocorrelation in the standard cross-cultural sample. \emph{World Cultures} \textbf{15}, 153--170 (2004).
+35. Eff, E. A. Does Mr. Galton still have a problem? Autocorrelation in the standard cross-cultural sample. \emph{World Cultures} \textbf{15}, 153--170 (2004).
 
 \leavevmode\hypertarget{ref-Symonds2014}{}%
-33.
+36.
 Symonds, M.
@@
 \leavevmode\hypertarget{ref-Watts2018}{}%
-34.
+37.
 Watts, J., Sheehan, O., Bulbulia, J., Gray, R.
@@
 \leavevmode\hypertarget{ref-Atkinson2016}{}%
-35.
+38.
 Atkinson, Q.
@@
 \leavevmode\hypertarget{ref-Kelly2020}{}%
-36. Kelly, M. Understanding persistence. \emph{CEPR Discussion Paper No. DP15246} (2020).
+39. Kelly, M. Understanding persistence. \emph{CEPR Discussion Paper No. DP15246} (2020).
 
 \leavevmode\hypertarget{ref-Bromham2018}{}%
-37.
+40.
 Bromham, L., Hua, X., Cardillo, M., Schneemann, H.
@@
 \leavevmode\hypertarget{ref-Bromham2021}{}%
-38.
+41.
 Bromham, L., Skeels, A., Schneemann, H., Dinnage, R.
@@
 \leavevmode\hypertarget{ref-Currie2012}{}%
-39. Currie, T. E. \& Mace, R. Analyses do not support the parasite-stress theory of human sociality. \emph{Behavioral and Brain Sciences} \textbf{35}, 83--85 (2012).
+42. Currie, T. E. \& Mace, R. Analyses do not support the parasite-stress theory of human sociality. \emph{Behavioral and Brain Sciences} \textbf{35}, 83--85 (2012).
+
+\leavevmode\hypertarget{ref-Passmore2022}{}%
+43.
+Passmore, S.
+\& Watts, J.
+WEIRD people and the Western Church: Who made whom?
+\emph{Religion, Brain \& Behavior} 1--58 (2022) doi:\href{https://doi.org/10.1080/2153599X.2021.1991459}{10.1080/2153599X.2021.1991459}.
 
 \leavevmode\hypertarget{ref-hdi}{}%
-40. \emph{Human development report}. \url{http://hdr.undp.org/en/composite/HDI}.
+44. \emph{Human development report}. \url{http://hdr.undp.org/en/composite/HDI}.
+
+\leavevmode\hypertarget{ref-Currie2009}{}%
+45. Currie, T. E. \& Mace, R. Political complexity predicts the spread of ethnolinguistic groups. \emph{Proceedings of the National Academy of Sciences} \textbf{106}, 7339--7344 (2009).
 
 \leavevmode\hypertarget{ref-Conley1999}{}%
-41. Conley, T. G. GMM estimation with cross sectional dependence. \emph{Journal of Econometrics} \textbf{92}, 1--45 (1999).
+46. Conley, T. GMM estimation with cross sectional dependence. \emph{Journal of Econometrics} \textbf{92}, 1--45 (1999).
 
 \leavevmode\hypertarget{ref-Conley2010}{}%
-42. Conley, T. G. Spatial econometrics. in \emph{Microeconometrics} (eds. Durlauf, S. N. \& Blume, L. E.) 303--313 (Palgrave Macmillan, 2010).
-
-\leavevmode\hypertarget{ref-Burkner2017}{}%
-43. Bürkner, P.-C. brms: An R package for Bayesian multilevel models using Stan. \emph{Journal of Statistical Software} \textbf{80}, 1--28 (2017).
-
-\leavevmode\hypertarget{ref-rstan}{}%
-44. Stan Development Team. RStan: The R interface to Stan. (2020).
+47. Conley, T. G. Spatial econometrics. in \emph{Microeconometrics} (eds. Durlauf, S. N. \& Blume, L. E.) 303--313 (Palgrave Macmillan, 2010).
 
 \leavevmode\hypertarget{ref-McElreath2020}{}%
-45. McElreath, R. \emph{Statistical rethinking: A Bayesian course with examples in R and Stan}. (CRC Press, 2020).
+48. McElreath, R. \emph{Statistical rethinking: A Bayesian course with examples in R and Stan}. (CRC Press, 2020).
 
 \leavevmode\hypertarget{ref-Neal1998}{}%
-46.
+49.
 Neal, R.
@@
 
+\leavevmode\hypertarget{ref-deVillemereuil2014}{}%
+50.
+Villemereuil, P.
+de \& Nakagawa, S.
+General quantitative genetic methods for comparative biology.
+in \emph{Modern phylogenetic comparative methods and their application in evolutionary biology: Concepts and practice} (ed.
+Garamszegi, L.
+Z.) 287--303 (Springer Berlin Heidelberg, 2014).
+doi:\href{https://doi.org/10.1007/978-3-662-43550-2_11}{10.1007/978-3-662-43550-2\_11}.
+
 \leavevmode\hypertarget{ref-Beck2003}{}%
-47. Beck, T., Demirgäç-Kunt, A. \& Levine, R. Law, endowments, and finance. \emph{Journal of Financial Economics} \textbf{70}, 137--181 (2003).
+51. Beck, T., Demirgäç-Kunt, A. \& Levine, R. Law, endowments, and finance. \emph{Journal of Financial Economics} \textbf{70}, 137--181 (2003).
 
 \leavevmode\hypertarget{ref-Beck2005}{}%
-48. Beck, T., Demirgäç-Kunt, A. \& Levine, R. SMEs, growth, and poverty: Cross-country evidence. \emph{Journal of economic growth} \textbf{10}, 199--229 (2005).
+52. Beck, T., Demirgäç-Kunt, A. \& Levine, R. SMEs, growth, and poverty: Cross-country evidence. \emph{Journal of economic growth} \textbf{10}, 199--229 (2005).
 
 \leavevmode\hypertarget{ref-Bockstette2002}{}%
-49. Bockstette, V., Chanda, A. \& Putterman, L. States and markets: The advantage of an early start. \emph{Journal of Economic growth} \textbf{7}, 347--369 (2002).
+53. Bockstette, V., Chanda, A. \& Putterman, L. States and markets: The advantage of an early start. \emph{Journal of Economic growth} \textbf{7}, 347--369 (2002).
 
 \leavevmode\hypertarget{ref-Easterly2003}{}%
-50. Easterly, W. \& Levine, R. Tropics, germs, and crops: How endowments influence economic development. \emph{Journal of Monetary Economics} \textbf{50}, 3--39 (2003).
+54. Easterly, W. \& Levine, R. Tropics, germs, and crops: How endowments influence economic development. \emph{Journal of Monetary Economics} \textbf{50}, 3--39 (2003).
 
 \leavevmode\hypertarget{ref-Easterly2007}{}%
-51. Easterly, W. Inequality does cause underdevelopment: Insights from a new instrument. \emph{Journal of Development Economics} \textbf{84}, 755--776 (2007).
+55. Easterly, W. Inequality does cause underdevelopment: Insights from a new instrument. \emph{Journal of Development Economics} \textbf{84}, 755--776 (2007).
 
 \leavevmode\hypertarget{ref-Skidmore2002}{}%
-52. Skidmore, M. \& Toya, H. Do natural disasters promote long-run growth? \emph{Economic Inquiry} \textbf{40}, 664--687 (2002).
+56. Skidmore, M. \& Toya, H. Do natural disasters promote long-run growth? \emph{Economic Inquiry} \textbf{40}, 664--687 (2002).
 
 \leavevmode\hypertarget{ref-Adamczyk2009}{}%
-53. Adamczyk, A. \& Pitt, C. Shaping attitudes about homosexuality: The role of religion and cultural context. \emph{Social Science Research} \textbf{38}, 338--351 (2009).
+57. Adamczyk, A. \& Pitt, C. Shaping attitudes about homosexuality: The role of religion and cultural context. \emph{Social Science Research} \textbf{38}, 338--351 (2009).
 
 \leavevmode\hypertarget{ref-Alesina2013}{}%
-54. Alesina, A., Giuliano, P. \& Nunn, N. On the origins of gender roles: Women and the plough. \emph{Quarterly Journal of Economics} \textbf{128}, 469--530 (2013).
+58. Alesina, A., Giuliano, P. \& Nunn, N. On the origins of gender roles: Women and the plough. \emph{Quarterly Journal of Economics} \textbf{128}, 469--530 (2013).
 
 \leavevmode\hypertarget{ref-Knack1997}{}%
-55. Knack, S. \& Keefer, P. Does social capital have an economic payoff? A cross-country investigation. \emph{The Quarterly Journal of Economics} \textbf{112}, 1251--1288 (1997).
+59. Knack, S. \& Keefer, P. Does social capital have an economic payoff? A cross-country investigation. \emph{The Quarterly Journal of Economics} \textbf{112}, 1251--1288 (1997).
 
 \leavevmode\hypertarget{ref-Sookias2018}{}%
-56. Sookias, R. B., Passmore, S. \& Atkinson, Q. D. Deep cultural ancestry and human development indicators across nation states. \emph{Royal Society Open Science} \textbf{5}, 171411 (2018).
-
-\leavevmode\hypertarget{ref-Currie2009}{}%
-57. Currie, T. E. \& Mace, R. Political complexity predicts the spread of ethnolinguistic groups. \emph{Proceedings of the National Academy of Sciences} \textbf{106}, 7339--7344 (2009).
+60. Sookias, R. B., Passmore, S. \& Atkinson, Q. D. Deep cultural ancestry and human development indicators across nation states. \emph{Royal Society Open Science} \textbf{5}, 171411 (2018).
 
 \leavevmode\hypertarget{ref-Smith2017}{}%
-58.
+61.
 Smith, M.
@@
 \leavevmode\hypertarget{ref-Hastie2017}{}%
-59. Hastie, T. J. \& Tibshirani, R. J. \emph{Generalized additive models}. (Routledge, 2017).
+62. Hastie, T. J. \& Tibshirani, R. J. \emph{Generalized additive models}. (Routledge, 2017).
 
 \leavevmode\hypertarget{ref-Smaldino2016}{}%
-60. Smaldino, P. E. \& McElreath, R. The natural selection of bad science. \emph{Royal Society Open Science} \textbf{3}, 160384 (2016).
+63. Smaldino, P. E. \& McElreath, R. The natural selection of bad science. \emph{Royal Society Open Science} \textbf{3}, 160384 (2016).
 
 \leavevmode\hypertarget{ref-Liddell2018}{}%
-61. Liddell, T. M. \& Kruschke, J. K. Analyzing ordinal data with metric models: What could possibly go wrong? \emph{Journal of Experimental Social Psychology} \textbf{79}, 328--348 (2018).
+64. Liddell, T. M. \& Kruschke, J. K. Analyzing ordinal data with metric models: What could possibly go wrong? \emph{Journal of Experimental Social Psychology} \textbf{79}, 328--348 (2018).
 
 \leavevmode\hypertarget{ref-Nicholson2017}{}%
-62.
+65.
 Nicholson, J.
@@
 \leavevmode\hypertarget{ref-Tipton2019}{}%
-63. Tipton, E., Pustejovsky, J. E. \& Ahmadi, H. Current practices in meta-regression in psychology, education, and medicine. \emph{Research Synthesis Methods} \textbf{10}, 180--194 (2019).
+66. Tipton, E., Pustejovsky, J. E. \& Ahmadi, H. Current practices in meta-regression in psychology, education, and medicine. \emph{Research Synthesis Methods} \textbf{10}, 180--194 (2019).
 
 \leavevmode\hypertarget{ref-Chambers2021}{}%
-64. Chambers, C. D. \& Tzavella, L. The past, present and future of Registered Reports. \emph{Nature Human Behaviour} \textbf{6}, 29--42 (2021).
+67. Chambers, C. D. \& Tzavella, L. The past, present and future of Registered Reports. \emph{Nature Human Behaviour} \textbf{6}, 29--42 (2021).
+
+\leavevmode\hypertarget{ref-rstan}{}%
+68. Stan Development Team. RStan: The R interface to Stan. (2020).
+
+\leavevmode\hypertarget{ref-Burkner2017}{}%
+69. Bürkner, P.-C. brms: An R package for Bayesian multilevel models using Stan. \emph{Journal of Statistical Software} \textbf{80}, 1--28 (2017).
 
 \leavevmode\hypertarget{ref-Oishi2013}{}%
-65. Oishi, S. Socioecological psychology. \emph{Annual Review of Psychology} \textbf{65}, 581--609 (2014).
+70. Oishi, S. Socioecological psychology. \emph{Annual Review of Psychology} \textbf{65}, 581--609 (2014).
 
 \leavevmode\hypertarget{ref-Tiefelsdorf2007}{}%
-66.
+71.
 Tiefelsdorf, M.
@@
 \leavevmode\hypertarget{ref-DeHoyos2006}{}%
-67. De Hoyos, R. E. \& Sarafidis, V. Testing for cross-sectional dependence in panel-data models. \emph{The Stata Journal} \textbf{6}, 482--496 (2006).
-
-\leavevmode\hypertarget{ref-Dow1984}{}%
-68. Dow, M. M., Burton, M. L., White, D. R. \& Reitz, K. P. Galton's Problem as network autocorrelation. \emph{American Ethnologist} \textbf{11}, 754--770 (1984).
-
-\leavevmode\hypertarget{ref-Klerman2011}{}%
-69. Klerman, D. M., Mahoney, P. G., Spamann, H. \& Weinstein, M. I. Legal origin or colonial history? \emph{Journal of Legal Analysis} \textbf{3}, 379--409 (2011).
+72. De Hoyos, R. E. \& Sarafidis, V. Testing for cross-sectional dependence in panel-data models. \emph{The Stata Journal} \textbf{6}, 482--496 (2006).
 
 \leavevmode\hypertarget{ref-Duben2021}{}%
-70. Düben, C. \emph{conleyreg: Estimations using Conley standard errors}. (2021).
+73. Düben, C. \emph{conleyreg: Estimations using Conley standard errors}. (2021).
 
 \leavevmode\hypertarget{ref-Solin2020}{}%
-71. Solin, A. \& Särkkä, S. Hilbert space methods for reduced-rank Gaussian process regression. \emph{Statistics and Computing} \textbf{30}, 419--446 (2020).
+74. Solin, A. \& Särkkä, S. Hilbert space methods for reduced-rank Gaussian process regression. \emph{Statistics and Computing} \textbf{30}, 419--446 (2020).
 
 \leavevmode\hypertarget{ref-Landau2021}{}%
-72.
+75.
 Landau, W.
@@
 \leavevmode\hypertarget{ref-Aust2020}{}%
-73. Aust, F. \& Barth, M. \emph{papaja: Create APA manuscripts with R Markdown}. (2020).
+76. Aust, F. \& Barth, M. \emph{papaja: Create APA manuscripts with R Markdown}. (2020).
+
+\leavevmode\hypertarget{ref-Brownrigg2018}{}%
+77. Brownrigg, R. \emph{maps: Draw geographical maps}. (2018).
+
+\leavevmode\hypertarget{ref-Hijmans2019}{}%
+78. Hijmans, R. J. \emph{geosphere: Spherical trigonometry}. (2019).
+
+\leavevmode\hypertarget{ref-Glottolog}{}%
+79.
+Hammarström, H., Forkel, R., Haspelmath, M.
+\& Bank, S.
+\emph{Glottolog 3.0}.
+(Max Planck Institute for the Science of Human History, 2017).
+doi:\href{https://doi.org/10.5281/zenodo.4061162}{10.5281/zenodo.4061162}.
+
+\leavevmode\hypertarget{ref-Ethnologue}{}%
+80. \emph{Ethnologue: Languages of the world}. (SIL International, 2018).
+
+\leavevmode\hypertarget{ref-Eff2008}{}%
+81. Eff, E. A. Weight matrices for cultural proximity: Deriving weights from a language phylogeny. \emph{Structure and Dynamics} \textbf{3}, (2008).
 
@@
 
-\clearpage
-\makeatletter
-\efloat@restorefloats
-\makeatother
-
-\begin{appendix}
+\newpage
+
+\hypertarget{appendix-appendix}{%
+\appendix}
+
 \renewcommand{\appendixname}{\bf{Supplementary Material}}
@@
 \hypertarget{calculating-global-geographic-and-linguistic-proximity-matrices}{%
-\subsubsection{Calculating global geographic and linguistic proximity
-matrices}\label{calculating-global-geographic-and-linguistic-proximity-matrices}}
-
-Geographic distance between two nations was calculated as the logged
-geodesic distance between country capital cities (data from the R
-package \emph{maps}\textsuperscript{1}) using the R package
-\emph{geosphere}\textsuperscript{2}. The geographic proximity matrix was
-computed as one minus the log geographic distance matrix scaled between
-0 and 1.
-
-Linguistic proximity between two nations was calculated as the cultural
-proximity between all languages spoken within those nations, weighted by
-speaker percentages. We acquired cultural proximity data by combining
-the language family trees provided by Glottolog v3.0\textsuperscript{3}
-into one global language tree (undated and unresolved). We calculated
-cultural proximity \(s\) between two languages \(j\) and \(k\) as the
-distance (in number of nodes traversed) of their most recent common
-ancestor \(i\) to the root of the tree, through the formula:
+\subsubsection{Calculating global geographic and linguistic proximity matrices}\label{calculating-global-geographic-and-linguistic-proximity-matrices}}
+
+Geographic distance between two nations was calculated as the logged geodesic distance between country capital cities (data from the R package \emph{maps}\textsuperscript{77}) using the R package \emph{geosphere}\textsuperscript{78}.
+The geographic proximity matrix was computed as one minus the log geographic distance matrix scaled between 0 and 1.
+
+Linguistic proximity between two nations was calculated as the cultural proximity between all languages spoken within those nations, weighted by speaker percentages.
+We acquired cultural proximity data by combining the language family trees provided by Glottolog v3.0\textsuperscript{79} into one global language tree (undated and unresolved).
+We calculated cultural proximity \(s\) between two languages \(j\) and \(k\) as the distance (in number of nodes traversed) of their most recent common ancestor \(i\) to the root of the tree, through the formula:
 
@@
 
-where \(n_{r}\) is the maximum path length (in number of nodes
-traversed) leading to the pan-human root \(r\), and \(n_{i}\) is the
-maximum path length leading to node \(i\). We then combined these
-proximities with speaker data from Ethnologue 21\textsuperscript{4} and
-compared every language spoken within those nations by at least 1
-permille of the population, weighted by speaker percentages, through the
-formula:
+where \(n_{r}\) is the maximum path length (in number of nodes traversed) leading to the pan-human root \(r\), and \(n_{i}\) is the maximum path length leading to node \(i\).
+We then combined these proximities with speaker data from Ethnologue 21\textsuperscript{80} and compared every language spoken within those nations by at least 1 permille of the population, weighted by speaker percentages, through the formula:
 
@@
 
-where \(p_{lj}\) is the percentage of the population in nation \(l\)
-speaking language \(j\), \(p_{mk}\) is the percentage of the population
-in nation \(m\) speaking language \(k\), and \(s_{jk}\) is the proximity
-measure between languages \(j\) and \(k\)\textsuperscript{5}. The
-resulting linguistic proximity matrix was also scaled between 0 and 1
-before analysis.
+where \(p_{lj}\) is the percentage of the population in nation \(l\) speaking language \(j\), \(p_{mk}\) is the percentage of the population in nation \(m\) speaking language \(k\), and \(s_{jk}\) is the proximity measure between languages \(j\) and \(k\)\textsuperscript{81}.
+The resulting linguistic proximity matrix was also scaled between 0 and 1 before analysis.
 
 \hypertarget{bayesian-models-for-reanalysis}{%
-\subsubsection{Bayesian models for
-reanalysis}\label{bayesian-models-for-reanalysis}}
-
-We provide model formulae for our reanalyses of cross-national
-correlations, for a general bivariate case with standardised outcome
-\(Y\) and predictor \(X\) variables. In the naive regression model
-without controls for non-independence:
+\subsubsection{Bayesian models for reanalysis}\label{bayesian-models-for-reanalysis}}
+
+We provide model formulae for our reanalyses of cross-national correlations, for a general bivariate case with standardised outcome \(Y\) and predictor \(X\) variables.
+In the naive regression model without controls for non-independence:
 
@@
 \end{aligned}
-\] The priors in this model were arrived at by prior predictive checks,
-with wider priors making predictions beyond the scale of standardised
-outcome variables and narrower priors being too informative.
-
-\newpage
-
-To control for spatial non-independence, we add a Gaussian process to
-this model and feed it a scaled geographic distance matrix \(D\) based
-on Euclidean distances between latitude and longitude coordinates. This
-distance matrix is computed internally by the R package
-\emph{brms}\textsuperscript{6}. The Gaussian process uses an
-exponentiated quadratic covariance kernel, the only covariance kernel
-currently supported by \emph{brms}. The model formula is:
+\]
+The priors in this model were arrived at by prior predictive checks, with wider priors making predictions beyond the scale of standardised outcome variables and narrower priors being too informative.
+
+\newpage
+
+To control for spatial non-independence, we add a Gaussian process to this model and feed it a scaled geographic distance matrix \(D\) based on Euclidean distances between latitude and longitude coordinates.
+This distance matrix is computed internally by the R package \emph{brms}\textsuperscript{69}.
+The Gaussian process uses an exponentiated quadratic covariance kernel, the only covariance kernel currently supported by \emph{brms}.
+The model formula is:
 
@@
 \end{aligned}
-\] where \(n\) is the number of nations, and \(D^2_{ij}\) reflects the
-squared Euclidean distances between latitude and longitude coordinates
-for the \(i\)-th and \(j\)-th nations. Notice that the inverse gamma
-prior on \(lscale\) is left undetermined. This is because the
-\emph{brms} package intelligently tunes the prior for this parameter
-based on the covariates of the Gaussian process (see
-https://betanalpha.github.io/assets/case\_studies/gp\_part3/part3.html).
-
-\newpage
-
-To control for cultural phylogenetic non-independence, we manually
-specify the covariance structure for nation random intercepts using a
-pre-computed linguistic proximity matrix \(L\) (see previous section).
-The covariance between two nations is assumed to be linearly
-proportional to the linguistic proximity between those nations. This
-assumption is justified if we assume that cultural traits evolve via
-Brownian motion along a language phylogeny. The non-centered
-parameterisation of this model is:
+\]
+where \(n\) is the number of nations, and \(D^2_{ij}\) reflects the squared Euclidean distances between latitude and longitude coordinates for the \(i\)-th and \(j\)-th nations.
+Notice that the inverse gamma prior on \(lscale\) is left undetermined.
+This is because the \emph{brms} package intelligently tunes the prior for this parameter based on the covariates of the Gaussian process (see \url{https://betanalpha.github.io/assets/case_studies/gp_part3/part3.html}).
+
+\newpage
+
+To control for cultural phylogenetic non-independence, we manually specify the covariance structure for nation random intercepts using a pre-computed linguistic proximity matrix \(L\) (see previous section).
+The covariance between two nations is assumed to be linearly proportional to the linguistic proximity between those nations.
+This assumption is justified if we assume that cultural traits evolve via Brownian motion along a language phylogeny.
+The non-centered parameterisation of this model is:
 
@@
 \end{aligned}
-\] \newpage
-
-Finally, we can control for spatial and cultural phylogenetic
-non-independence simultaneously by including both a Gaussian process
-over latitude and longitude coordinates \emph{and} nation random
-intercepts that covary according to linguistic proximity. The resulting
-model is as follows:
+\]
+\newpage
+
+Finally, we can control for spatial and cultural phylogenetic non-independence simultaneously by including both a Gaussian process over latitude and longitude coordinates \emph{and} nation random intercepts that covary according to linguistic proximity.
+The resulting model is as follows:
 
@@
 \hypertarget{geographic-and-cultural-phylogenetic-signal-estimates-for-human-development-index-and-ingleharts-value-dimensions}{%
-\subsubsection{Geographic and cultural phylogenetic signal estimates for
-Human Development Index and Inglehart's value
-dimensions}\label{geographic-and-cultural-phylogenetic-signal-estimates-for-human-development-index-and-ingleharts-value-dimensions}}
-
-Controlling for shared cultural ancestry, the proportion of
-national-level variance explained by spatial proximity was 0.38 for the
-Human Development Index (95\% credible interval {[}0.20 0.60{]};
-\(\text{BF}_{\neq0}\) \textgreater{} 100), 0.43 for traditional values
-(95\% CI {[}0.18 0.72{]}; \(\text{BF}_{\neq0}\) \textgreater{} 100), and
-0.18 for survival values (95\% CI {[}0.00 0.44{]}; \(\text{BF}_{\neq0}\)
-= 1.45). Controlling for spatial proximity, the proportion of variance
-explained by shared cultural ancestry was 0.60 for the Human Development
-Index (95\% CI {[}0.39 0.79{]}; \(\text{BF}_{\neq0}\) \textgreater{}
-100), 0.55 for traditional values (95\% CI {[}0.26 0.79{]};
-\(\text{BF}_{\neq0}\) \textgreater{} 100), and 0.79 for survival values
-(95\% CI {[}0.53 0.97{]}; \(\text{BF}_{\neq0}\) \textgreater{} 100).
+\subsubsection{Geographic and cultural phylogenetic signal estimates for Human Development Index and Inglehart's value dimensions}\label{geographic-and-cultural-phylogenetic-signal-estimates-for-human-development-index-and-ingleharts-value-dimensions}}
+
+Controlling for shared cultural ancestry, the proportion of national-level variance explained by spatial proximity was 0.37 for the Human Development Index (95\% credible interval {[}0.19 0.60{]}; \(\text{BF}_{\neq0}\) \textgreater{} 100), 0.43 for traditional values (95\% CI {[}0.19 0.73{]}; \(\text{BF}_{\neq0}\) \textgreater{} 100), and 0.18 for survival values (95\% CI {[}0.01 0.44{]}; \(\text{BF}_{\neq0}\) = 1.77).
+Controlling for spatial proximity, the proportion of variance explained by shared cultural ancestry was 0.61 for the Human Development Index (95\% CI {[}0.39 0.80{]}; \(\text{BF}_{\neq0}\) \textgreater{} 100), 0.55 for traditional values (95\% CI {[}0.26 0.79{]}; \(\text{BF}_{\neq0}\) \textgreater{} 100), and 0.79 for survival values (95\% CI {[}0.54 0.96{]}; \(\text{BF}_{\neq0}\) \textgreater{} 100).
 
@@
 \includegraphics{manuscript_files/figure-latex/dag-1.pdf}
-\caption{\label{fig:dag}\emph{A causal directed acyclic graph of spatial and
-cultural phylogenetic non-independence in cross-national studies.} We
-are interested in estimating the direct effect of national-level
-exposure \(X\) on national-level outcome \(Y\). But these variables are
-confounded by their common unobserved cause \(U\). \(U\) is a stand-in
-for shared environmental, ecological, and geographic causes
-(e.g.~climate, biodiversity, physical topography) and cultural and
-institutional causes (e.g.~cultural norms, political systems). In this
-causal model, we cannot condition on \(U\) directly, as it is
-unobserved. However, geographic \(G\) and linguistic \(L\) relationships
-between societies influence \(U\), since changing a nation's spatial
-distance to or shared cultural ancestry with other nations will change
-its environmental and cultural traits. We can thus use \(G\) and \(L\)
-to model the covariation between \(X\) and \(Y\) induced by \(U\).
-Failing to do this and simply estimating the bivariate correlation
-between \(X\) and \(Y\) will produce spurious relationships and
-residuals that are spatially and culturally non-independent around the
-world (see Figure \ref{fig:spatialExample}).}
+\caption{\label{fig:dag}\emph{A causal directed acyclic graph of spatial and cultural phylogenetic non-independence in cross-national studies.} We are interested in estimating the direct effect of national-level exposure \(X\) on national-level outcome \(Y\).
+But these variables are confounded by their common unobserved cause \(U\).
+\(U\) is a stand-in for shared environmental, ecological, and geographic causes (e.g.~climate, biodiversity, physical topography) and cultural and institutional causes (e.g.~cultural norms, political systems).
+In this causal model, we need to condition on \(U\) to estimate the direct path from \(X\) to \(Y\), but we cannot since it is unobserved.
+However, geographic \(G\) and linguistic \(L\) relationships between societies influence \(U\), since changing a nation's spatial distance to or shared cultural ancestry with other nations will change its environmental and cultural traits.
+We can thus use \(G\) and \(L\) to model the covariation between \(X\) and \(Y\) induced by \(U\).
+Failing to do this and simply estimating the bivariate correlation between \(X\) and \(Y\) will produce spurious relationships and residuals that are spatially and culturally non-independent around the world.}
 \end{figure}
@@
 \includegraphics{manuscript_files/figure-latex/plotSim1-1.pdf}
-\caption{\label{fig:plotSim1}\emph{Distribution of cross-national correlations
-from simulation study under moderate spatial autocorrelation.} In these
-simulations, the strength of spatial autocorrelation is set to 0.5 for
-both outcome and predictor variables. For frequentist regression models,
-point ranges represent correlation estimates and 95\% confidence
-intervals. For Bayesian Gaussian process regression models, point ranges
-represent posterior means and 95\% credible intervals. Correlations are
-ordered by effect size independently in each panel. Red point ranges
-indicate that the slope is ``significant'' (i.e.~the 95\% confidence /
-credible interval excludes zero). Black point ranges indicate that the
-slope is ``not significant''.}
+\caption{\label{fig:plotSim1}\emph{Distribution of cross-national correlations from simulation study under strong spatial autocorrelation.} In these simulations, the strength of spatial autocorrelation is set to 0.8 for both outcome and predictor variables.
+For frequentist regression models, point ranges represent correlation estimates and 95\% confidence intervals.
+For Bayesian regression models, point ranges represent posterior means and 95\% credible intervals.
+Correlations are ordered by effect size independently in each panel.
+Red point ranges indicate that the slope is \enquote{significant} (i.e.~the 95\% confidence / credible interval excludes zero).
+Black point ranges indicate that the slope is \enquote{not significant}.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotSim2-1.pdf}
+\caption{\label{fig:plotSim2}\emph{Distribution of cross-national correlations from simulation study under strong cultural phylogenetic autocorrelation.} In these simulations, the strength of cultural phylogenetic autocorrelation is set to 0.8 for both outcome and predictor variables.
+For frequentist regression models, point ranges represent correlation estimates and 95\% confidence intervals.
+For Bayesian regression models, point ranges represent posterior means and 95\% credible intervals.
+Correlations are ordered by effect size independently in each panel.
+Red point ranges indicate that the slope is \enquote{significant} (i.e.~the 95\% confidence / credible interval excludes zero).
+Black point ranges indicate that the slope is \enquote{not significant}.}
 \end{figure}
@@
 \includegraphics{manuscript_files/figure-latex/plotReplications2-1.pdf}
-\caption{\label{fig:plotReplications2}\emph{Reanalysis models fitted to raw
-data, for economic development (a) and cultural values (b) studies.}
-Data points are labelled using ISO 3166-1 alpha-2 letter country codes.
-In all reanalyses, outcome and predictor variables are standardised,
-making regression slopes comparable to Pearson's correlation
-coefficients. Lines and shaded areas represent posterior median
-regression lines and 95\% credible intervals. For models with covariates
-(Adamzyck and Pitt 2009; Gelfand et al.~2011), marginal effects are
-presented holding all covariates at zero or their reference categories.}
+\caption{\label{fig:plotReplications2}\emph{Reanalysis models fitted to raw data, for economic development (a) and cultural values (b) studies.} Data points are labelled using ISO 3166-1 alpha-2 letter country codes.
+In all reanalyses, outcome and predictor variables are standardised, making regression slopes comparable to Pearson's correlation coefficients.
+Lines and shaded areas represent posterior median regression lines and 95\% credible intervals.
+For models with covariates (Adamzyck and Pitt 2009; Gelfand et al.~2011), marginal effects are presented holding all covariates at zero or their reference categories.}
 \end{figure}
@@
 \includegraphics{manuscript_files/figure-latex/plotReplications3-1.pdf}
-\caption{\label{fig:plotReplications3}\emph{Posterior estimates of Gaussian
-process functions mapping spatial autocorrelation onto geographic
-distance from our reanalyses of economic development (a) and cultural
-values (b) studies.} Estimates are from models additionally controlling
-for cultural phylogenetic non-independence. The y-axis represents the
-amount of spatial autocorrelation between data points with increasing
-distance between those points on the x-axis (logged distance in
-kilometres). Lines and shaded areas represent median posterior spatial
-autocorrelation functions and 50\% and 95\% credible intervals.}
+\caption{\label{fig:plotReplications3}\emph{Posterior estimates of Gaussian process functions mapping spatial autocorrelation onto geographic distance from our reanalyses of economic development (a) and cultural values (b) studies.} Estimates are from models additionally controlling for cultural phylogenetic non-independence.
+The y-axis represents the amount of spatial autocorrelation between data points with increasing distance between those points on the x-axis (logged distance in kilometres).
+Lines and shaded areas represent median posterior spatial autocorrelation functions and 50\% and 95\% credible intervals.}
 \end{figure}
@@
 \includegraphics{manuscript_files/figure-latex/plotReplications4-1.pdf}
-\caption{\label{fig:plotReplications4}\emph{Posterior estimates of cultural
-phylogenetic signal from our reanalyses.} Estimates are from models
-additionally controlling for spatial non-independence. Cultural
-phylogenetic signal is operationalised as the proportion of
-national-level variance explained by linguistic proximity between
-nations. Ridges are full posterior distributions, and point ranges are
-posterior medians and 95\% credible intervals.}
+\caption{\label{fig:plotReplications4}\emph{Posterior estimates of cultural phylogenetic signal from our reanalyses.} Estimates are from models additionally controlling for spatial non-independence.
+Cultural phylogenetic signal is operationalised as the proportion of national-level variance explained by linguistic proximity between nations.
+Ridges are full posterior distributions, and points are posterior medians, and lines represent 50\% and 95\% credible intervals.}
 \end{figure}
@@
 \hypertarget{supplementary-references}{%
-\subsection*{Supplementary References}\label{supplementary-references}}
-\addcontentsline{toc}{subsection}{Supplementary References}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Brownrigg2018}{}%
-1. Brownrigg, R. \emph{maps: Draw geographical maps}. (2018).
-
-\leavevmode\hypertarget{ref-Hijmans2019}{}%
-2. Hijmans, R. J. \emph{geosphere: Spherical trigonometry}. (2019).
-
-\leavevmode\hypertarget{ref-Glottolog}{}%
-3. Hammarström, H., Forkel, R., Haspelmath, M. \& Bank, S.
-\emph{Glottolog 3.0}. (Max Planck Institute for the Science of Human
-History, 2017).
-doi:\href{https://doi.org/10.5281/zenodo.4061162}{10.5281/zenodo.4061162}.
-
-\leavevmode\hypertarget{ref-Ethnologue}{}%
-4. \emph{Ethnologue: Languages of the world}. (SIL International, 2018).
-
-\leavevmode\hypertarget{ref-Eff2008}{}%
-5. Eff, E. A. Weight matrices for cultural proximity: Deriving weights
-from a language phylogeny. \emph{Structure and Dynamics} \textbf{3},
-(2008).
-
-\leavevmode\hypertarget{ref-Burkner2017}{}%
-6. Bürkner, P.-C. brms: An R package for Bayesian multilevel models
-using Stan. \emph{Journal of Statistical Software} \textbf{80}, 1--28
-(2017).
-\end{appendix}
+\subsection{Supplementary References}\label{supplementary-references}}
 

### 2022-03-31 R/functionsReplications.R

@@ -778,18 +778,18 @@ plotReplicationResults1 <- function(slopeList) {
   # plot
   out <-
     tibble(
-      paper = rep(c("*Adamczyk and Pitt (2009)*<br>Homosexuality disapproval ~ Survival values",
-                    "*Alesina et al. (2013)*<br>FLFP ~ Traditional plough use",
-                    "*Beck et al. (2003)*<br>Stock market development ~ Settler mortality",
-                    "*Beck et al. (2005)*<br>GDP ~ SME sector share",
-                    "*Bockstette et al. (2002)*<br>GDP growth ~ State antiquity",
-                    "*Easterly and Levine (2003)*<br>Log GDP ~ Institutional development",
-                    "*Easterly (2007)*<br>Gini ~ Log wheat sugar ratio",
-                    "*Fincher et al. (2008)*<br>Individualism ~ Hist. pathogen prevalence",
-                    "*Gelfand et al. (2011)*<br>Tightness ~ Nature disaster vulnerability",
-                    "*Inglehart and Baker (2000)*<br>Traditional values ~ % industrial sector",
-                    "*Knack and Keefer (1997)*<br>Confidence in institutions ~ % trusting",
-                    "*Skidmore and Toya (2002)*<br>GDP growth ~ Log number of natural disasters"), each = 4),
+      paper = rep(c("*Adamczyk and Pitt (2009)*, n = 33<br>Homosexuality disapproval ~ Survival values",
+                    "*Alesina et al. (2013)*, n = 75<br>FLFP ~ Traditional plough use",
+                    "*Beck et al. (2003)*, n = 69<br>Stock market development ~ Settler mortality",
+                    "*Beck et al. (2005)*, n = 45<br>GDP ~ SME sector share",
+                    "*Bockstette et al. (2002)*, n = 103<br>GDP growth ~ State antiquity",
+                    "*Easterly and Levine (2003)*, n = 63<br>Log GDP ~ Institutional development",
+                    "*Easterly (2007)*, n = 98<br>Gini ~ Log wheat sugar ratio",
+                    "*Fincher et al. (2008)*, n = 66<br>Individualism ~ Hist. pathogen prevalence",
+                    "*Gelfand et al. (2011)*, n = 28<br>Tightness ~ Nature disaster vulnerability",
+                    "*Inglehart and Baker (2000)*, n = 38<br>Traditional values ~ % industrial sector",
[…]

### 2022-03-31 R/functionsSignal.R

@@ -5,7 +5,7 @@ loadCovMat <- function(file, log) {
   # load distance matrix
   out <- 
     read_excel(file, na = "") %>%
-    select(-ISO) %>%
+    dplyr::select(-ISO) %>%
     as.matrix()
   rownames(out) <- colnames(out)
   # log distances?
@@ -27,8 +27,8 @@ loadHDIData <- function(fileHDI, fileISOHDI) {
     skip = 4
   ) %>%
     drop_na("HDI rank") %>%
-    select(c("Country", "1990", "2000", "2010", 
-             "2014", "2015", "2017", "2018", "2019")) %>%
+    dplyr::select(c("Country", "1990", "2000", "2010", 
+                    "2014", "2015", "2017", "2018", "2019")) %>%
     # get data in long form
     pivot_longer(cols = !Country, names_to = "year", values_to = "hdi") %>%
     filter(hdi != "..") %>%
@@ -67,7 +67,7 @@ fitWVSSignal <- function(wvs, outcome = "", geoCov, linCov) {
   wvs <-
     wvs %>%
     # select vars
-    select(
+    dplyr::select(
       # country code
       S009,
       # traditional vs. secular-rational values

### 2022-03-31 R/functionsSimulation.R

@@ -9,11 +9,11 @@ loadISO <- function(fileISO) {
 }
 
 # load geographic and linguistic covariance matrices
-loadSimCovMat <- function(file, log, continent, iso, langFam) {
+loadSimCovMat <- function(file, continent, iso, langFam) {
   # load distance matrix
   out <- 
     read_excel(file, na = "") %>%
-    select(-ISO) %>%
+    dplyr::select(-ISO) %>%
     as.matrix()
   rownames(out) <- colnames(out)
   # keep only countries with longitude, latitude, continent, and language family data
@@ -23,36 +23,58 @@ loadSimCovMat <- function(file, log, continent, iso, langFam) {
              rownames(out)[rownames(out) %in% continent$country]]
   out <- out[rownames(out)[rownames(out) %in% langFam$iso],
              rownames(out)[rownames(out) %in% langFam$iso]]
-  # log distances?
-  if (log) out <- log(out)
   # distances between 0 and 1
   out <- out / max(out)
   diag(out) <- 0
   # 1 - distance = proximity (covariance)
   out <- 1 - out
+  # ensure covariance matrix is positive definite by adding small constant to diagonal
+  diag(out) <- diag(out) + 1e-06
   return(out)
 }
 
+# brms model to simulate data
+fitSimulationModel <- function(covMat, lambda, rho, iter) {
+  # signal for dependent variable = lambda
+  # signal for independent variable = rho
+  signalY <- lambda
+  signalX <- rho
+  # calculate sd and sigma values, assuming total variance sd^2 + sigma^2 = 1
+  # signal = sd^2 / (sd^2 + sigma^2)
+  # signal = sd^2 / 1
+  # sd = sqrt(signal)
+  sdY <- sqrt(signalY)
+  sdX <- sqrt(signalX)
+  # sd^2 + sigma^2 = 1
+  # sigma = sqrt(1 - sd^2)
+  sigmaY <- sqrt(1 - sdY^2)
+  sigmaX <- sqrt(1 - sdX^2)
+  # mock data for simulation (just to feed to stan so model can run)
+  d <- data.frame(y = rnorm(nrow(covMat)), x = rnorm(nrow(covMat)), id = rownames(covMat))
+  # simulation
+  mSim <- brm(bf(y ~ 0 + (1 | gr(id, cov = covMat))) +
+              bf(x ~ 0 + (1 | gr(id, cov = covMat))) + set_rescor(FALSE),
+              data = d, data2 = list(covMat = covMat),
[…]

### 2022-03-31 R/scripts/testSim.R

@@ -1,0 +1,72 @@
+library(brms)
+
+n <- 300
+
+# create proximity matrix
+distMat <- as.matrix(dist(rnorm(n)))
+distMat <- distMat / max(distMat)
+rownames(distMat) <- colnames(distMat)
+proxMat <- 1 - distMat
+
+# mock data
+d <- data.frame(y = rnorm(n), x = rnorm(n), id = 1:n)
+
+# signal for each variable
+signalY <- 0.2
+signalX <- 0.5
+
+# calculate sd and sigma values, assuming total variance sd^2 + sigma^2 = 1
+# signal = sd^2 / sd^2 + sigma^2
+# signal = sd^2 / 1
+# sd = sqrt(signal)
+sdY <- sqrt(signalY)
+sdX <- sqrt(signalX)
+# sd^2 + sigma^2 = 1
+# sigma = sqrt(1 - sd^2)
+sigmaY <- sqrt(1 - sdY^2)
+sigmaX <- sqrt(1 - sdX^2)
+
+# "data"
+d <- data.frame(y = rnorm(n), x = rnorm(n), id = 1:n)
+
+# simulation
+mSim <- brm(bf(y ~ 0 + (1 | gr(id, cov = proxMat))) +
+            bf(x ~ 0 + (1 | gr(id, cov = proxMat))) + set_rescor(FALSE),
+            data = d, data2 = list(proxMat = proxMat),
+            prior = c(prior_string(paste0("constant(", sdY, ")"), class = "sd", resp = "y"),
+                      prior_string(paste0("constant(", sigmaY, ")"), class = "sigma", resp = "y"),
+                      prior_string(paste0("constant(", sdX, ")"), class = "sd", resp = "x"),
+                      prior_string(paste0("constant(", sigmaX, ")"), class = "sigma", resp = "x")),
+            sample_prior = "only", iter = 1, chains = 1)
+
+# simulate data
+out <- posterior_predict(mSim)
+
+# standardise y and x
+y <- as.numeric(scale(out[,,"y"]))
+x <- as.numeric(scale(out[,,"x"]))
+
+# data frame
+d <- data.frame(y, x, id = 1:n)
+
+# get signal back
+m1 <- brm(y ~ 1 + (1 | gr(id, cov = proxMat)),
+          data = d, data2 = list(proxMat = proxMat),
+          prior = c(prior(normal(0, 0.5), class = Intercept),
+                    prior(exponential(1), class = sd),
[…]

### 2022-03-31 _targets.R

@@ -9,7 +9,7 @@ source("R/functionsReplications.R")
 options(tidyverse.quiet = TRUE)
 tar_option_set(packages = c("brms", "cowplot", "conleyreg", "countrycode", 
                             "dagitty", "geosphere", "ggdag", "ggrepel", "ggtext", 
-                            "haven", "lmtest", "MASS", "papaja", "psych", "readxl", 
+                            "haven", "lmtest", "papaja", "psych", "readxl", 
                             "rstan", "sjlabelled", "tidybayes", "tidyverse"))
 
 # targets for simulation (see below)
@@ -20,10 +20,12 @@ simulationTargets <-
     # map simulation over different covariance matrices and different values of lambda and rho
     values = expand_grid(covMat = rlang::syms(c("simGeoCov", "simLinCov")),
                          lambda = c(0.2, 0.5, 0.8), rho = c(0.2, 0.5, 0.8)),
+    # simulation model
+    tar_target(simModel, fitSimulationModel(covMat, lambda, rho, iter)),
     # simulate data
-    tar_target(simData, simulateData(covMat = covMat, continent, iso, langFam, 
-                                     seed = seed, lambda = lambda, rho = rho), 
-               pattern = map(seed), iteration = "list"),
+    tar_target(simData, simulateData(simModel, covMat, continent, iso, 
+                                     langFam, iter, rho, lambda), 
+               pattern = map(iter), iteration = "list"),
     # ols analyses
     tar_target(olsModel1, fitOLSModel(y ~ x,              data = simData), pattern = map(simData)),
     tar_target(olsModel2, fitOLSModel(y ~ x + latitude,   data = simData), pattern = map(simData)),
@@ -58,7 +60,7 @@ list(
   # hdi data
   tar_target(hdi, loadHDIData(fileHDI, fileISOHDI)),
   # wvs data
-  tar_target(wvs, read_sav(fileWVS)),
+  tar_target(wvs, haven::read_sav(fileWVS, encoding = "latin1")),
   # covariance matrices
   tar_target(geoCov, loadCovMat(fileGeo, log = TRUE)),
   tar_target(linCov, loadCovMat(fileLin, log = FALSE)),
@@ -118,14 +120,14 @@ list(
   # isocodes (https://gist.github.com/tadast/8827699)
   tar_target(iso, loadISO(fileISO)),
   # geographic and linguistic covariance matrices
-  tar_target(simGeoCov, loadSimCovMat(fileGeo, log = TRUE , continent, iso, langFam)),
[…]

### 2022-04-18 appendix.Rmd
pre-submission manuscript edits

@@
+---
+output:
+  pdf_document: default
+  html_document: default
+---
 \renewcommand{\appendixname}{\bf{Supplementary Material}}
@@
 
-Geographic distance between two nations was calculated as the logged geodesic distance between country capital cities (data from the R package _maps_[@Brownrigg2018]) using the R package _geosphere_[@Hijmans2019].
+Geographic distance between two nations was calculated as the logged geodesic distance between country capital cities (data from the R package _maps_; Brownrigg, 2018) using the R package _geosphere_ (Hijmans, 2019).
 The geographic proximity matrix was computed as one minus the log geographic distance matrix scaled between 0 and 1.
@@
 Linguistic proximity between two nations was calculated as the cultural proximity between all languages spoken within those nations, weighted by speaker percentages.
-We acquired cultural proximity data by combining the language family trees provided by Glottolog v3.0[@Glottolog] into one global language tree (undated and unresolved).
+We acquired cultural proximity data by combining the language family trees provided by Glottolog v3.0 (Hammarström et al., 2017) into one global language tree (undated and unresolved).
 We calculated cultural proximity $s$ between two languages $j$ and $k$ as the distance (in number of nodes traversed) of their most recent common ancestor $i$ to the root of the tree, through the formula:
@@
 where $n_{r}$ is the maximum path length (in number of nodes traversed) leading to the pan-human root $r$, and $n_{i}$ is the maximum path length leading to node $i$.
-We then combined these proximities with speaker data from Ethnologue 21[@Ethnologue] and compared every language spoken within those nations by at least 1 permille of the population, weighted by speaker percentages, through the formula:
+We then combined these proximities with speaker data from Ethnologue 21 (Ethnologue, 2018) and compared every language spoken within those nations by at least 1 permille of the population, weighted by speaker percentages, through the formula:
 
@@
 
-where $p_{lj}$ is the percentage of the population in nation $l$ speaking language $j$, $p_{mk}$ is the percentage of the population in nation $m$ speaking language $k$, and $s_{jk}$ is the proximity measure between languages $j$ and $k$[@Eff2008].
+where $p_{lj}$ is the percentage of the population in nation $l$ speaking language $j$, $p_{mk}$ is the percentage of the population in nation $m$ speaking language $k$, and $s_{jk}$ is the proximity measure between languages $j$ and $k$ (Eff, 2008).
 The resulting linguistic proximity matrix was also scaled between 0 and 1 before analysis.
@@
 To control for spatial non-independence, we add a Gaussian process to this model and feed it a scaled geographic distance matrix $D$ based on Euclidean distances between latitude and longitude coordinates.
-This distance matrix is computed internally by the R package _brms_[@Burkner2017].
+This distance matrix is computed internally by the R package _brms_ (Bürkner, 2017).
 The Gaussian process uses an exponentiated quadratic covariance kernel, the only covariance kernel currently supported by _brms_.
@@
 ## Supplementary References
+
+Brownrigg, R. _maps: Draw geographical maps._ (2018).
+
+Bürkner, P.-C. brms: An R package for Bayesian multilevel models using Stan. _Journal of Statistical Software_ **80**, 1---28 (2017).
+
+Eff, E. A. Weight matrices for cultural proximity: Deriving weights from a language phylogeny. _Structure and Dynamics_ **3**, (2008).
+
+_Ethnologue: Languages of the world._ (SIL International, 2018).
+
+Hammarström, H., Forkel, R., Haspelmath, M.
+& Bank, S.
+_Glottolog 3.0._ (Max Planck Institute for the Science of Human History, 2017).
+doi:[10.5281/zenodo.4061162](https://doi.org/10.5281/zenodo.4061162)
+
+Hijmans, R. J. _geosphere: Spherical trigonometry._ (2019).

### 2022-04-18 manuscript.Rmd
pre-submission manuscript edits

@@
   - id            : "2"
-    institution   : "Max Planck Institute for the Science of Human History, Jena, Germany"
+    institution   : "Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany"
 
@@
 abstract: |
- Cross-national analyses can illuminate global patternings of national-level outcomes.
-But since nations are connected in various ways, such as via spatial proximity and shared cultural ancestry, cross-national analyses often violate assumptions of non-independence, inflating false positive rates.
+ Cross-national analyses test hypotheses about the drivers of global variation in national outcomes.
+However, since nations are connected in various ways, such as via spatial proximity and shared cultural ancestry, cross-national analyses often violate assumptions of non-independence, inflating false positive rates.
 Here, we show that, despite being recognised as an important statistical pitfall for over 200 years, cross-national research in economics and psychology still does not sufficiently account for non-independence.
-In a review of the 100 highest-cited cross-national studies of economic development and values, we find that controls for non-independence are rare and have remained so over the past 30 years.
+In a review of the 100 highest-cited cross-national studies of economic development and values, we find that controls for non-independence are rare.
 When studies do include controls for non-independence, our simulations suggest that commonly used methods continue to produce false positives.
@@
 keywords          : "spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses"
-wordcount         : "5500 words"
+wordcount         : "5459 words"
 bibliography      : ["r-references.bib"]
@@
 ```{r setup, include=FALSE}
-## !!! for use without tar_make
-#library(targets)
-#library(tarchetypes)
-#library(tidyverse)
-#source("R/functionsDAG.R")
-#source("R/functionsSignal.R")
-#source("R/functionsReview.R")
-#source("R/functionsSimulation.R")
-#source("R/functionsReplications.R")
-#options(tidyverse.quiet = TRUE)
-#lapply(c("brms", "cowplot", "conleyreg", "countrycode", 
-#         "dagitty", "geosphere", "ggdag", "ggrepel", "ggtext", 
-#         "haven", "lmtest", "papaja", "psych", "readxl", 
-#         "rstan", "sjlabelled", "tidybayes", "tidyverse"), library, character.only = TRUE)
-## !!!
-#
 knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
@@
 
-Given the importance of nations for structuring human behaviour, there is justifiably huge interest in statistical analyses that attempt to predict variation in national-level outcomes of all kinds.
+Given the importance of nations for structuring human behaviour, there is justifiably huge interest in statistical analyses that attempt to predict variation in national outcomes of all kinds.
 At the time of writing, a search in Web of Science for the terms "cross-national", "cross-cultural", or "cross-country" in titles and abstracts returned 80,000 unique hits.
@@
 Shared cultural ancestry can result in a form of pseudoreplication, whereby multiple instances of the same trait across nations are merely duplicates of the ancestral original.
-For example, despite being separated by over 1500 kilometres of ocean, the related island nations Tonga and Tuvalu express similar languages and customs due to cultural inheritance from a shared ancestor deep in the evolutionary history of Polynesian societies.
-More recently, shared ancestry is also crucial in explaining cultural similarities between colonial settlements and their colonisers (e.g.
+For example, despite being separated by over 1,500 kilometres of ocean, the related island nations Tonga and Tuvalu share similar languages and customs due to cultural inheritance from a common Polynesian population dating back more than 1,000 years.
+More recently, shared ancestry explains a myriad of cultural similarities between colonial settlements and their colonisers (e.g.
 Argentina and Spain).
@@
 In order to motivate our research question, it is important to first show that economic development and cultural values are indeed spatially and culturally non-independent around the world.
-To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for the Human Development Index[@hdi] and two dimensions of cultural values from the World Values Survey, traditional vs.
+To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for the Human Development Index[@hdi] and two primary dimensions of cultural values from the World Values Survey, traditional vs.
 secular values and survival vs.
@@
 
-Since economic development and cultural values show evidence of geographic and cultural phylogenetic signal, have cross-national analyses sufficiently accounted for this non-independence?
+Given that economic development and cultural values show evidence of geographic and cultural phylogenetic signal, have cross-national analyses sufficiently accounted for this non-independence?
 To assess this, we systematically searched the published literature for articles that combined the search terms "economic development" or "values" with the search terms "cross-national", "cross-cultural", or "cross-country".
@@
 Finally, we included Bayesian multilevel regressions that explicitly model spatial and/or cultural phylogenetic non-independence by allowing nation random intercepts to covary according to global geographic and/or linguistic proximity matrices (see Supplementary Methods).
-To model spatial non-independence we included a Gaussian process[@McElreath2020;@Neal1998] over latitude and longitude values, and to model cultural phylogenetic non-independence we assumed that nation random intercepts were correlated proportional to their linguistic proximity[@deVillemereuil2014].
+To model spatial non-independence we included a Gaussian process[@McElreath2020;@Neal1998] over latitude and longitude values, and to model cultural phylogenetic non-independence we assumed that nation random intercepts were correlated in proportion to their linguistic proximity[@deVillemereuil2014].
 Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero (i.e.
@@
 
-Our systematic review and simulation study have shown that controls for non-independence are rare in cross-national studies of economic development and cultural values, and when studies do attempt to control for non-independence, the methods used continue to produce false positives.
+Our systematic review and simulation study have shown that controls for non-independence are rare in cross-national studies of economic development and cultural values, and when studies do attempt to control for non-independence, the methods typically used are likely to continue to produce false positives.
 This raises the worrying possibility that the cross-national literature in economics and psychology is populated with spurious relationships.
@@
 For each outcome variable, our Gaussian process models provided varying estimates of how quickly spatial autocorrelation declined with distance (Supplementary Figure \@ref(fig:plotReplications3)).
-For example, in Skidmore and Toya[@Skidmore2002] gross domestic product growth was only moderately spatially autocorrelated at 1000 km distance (posterior median spatial autocorrelation at 1000 km = `r printnum(median(skidmoreSA))`, 95% CI [`r printnum(quantile(skidmoreSA, 0.025))` `r printnum(quantile(skidmoreSA, 0.975))`]), whereas in Inglehart and Baker[@Inglehart2000] traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1000 km = `r printnum(median(inglehartSA))`, 95% CI [`r printnum(quantile(inglehartSA, 0.025))` `r printnum(quantile(inglehartSA, 0.975))`]).
+For example, in Skidmore and Toya[@Skidmore2002] gross domestic product growth was only moderately spatially autocorrelated at 1,000 km distance (posterior median spatial autocorrelation at 1,000 km = `r printnum(median(skidmoreSA))`, 95% CI [`r printnum(quantile(skidmoreSA, 0.025))` `r printnum(quantile(skidmoreSA, 0.975))`]), whereas in Inglehart and Baker[@Inglehart2000] traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1,000 km = `r printnum(median(inglehartSA))`, 95% CI [`r printnum(quantile(inglehartSA, 0.025))` `r printnum(quantile(inglehartSA, 0.975))`]).
 We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \@ref(fig:plotReplications4)), with some outcome variables expressing low signal (e.g.
@@
 
-(ref:plotReplications5Caption) _The estimated degree of spatial and cultural phylogenetic non-independence predicts reductions in effect size in our reanalysis._ (a) Higher estimated degrees of spatial autocorrelation at 1000 km distance predict more pronounced reductions in effect sizes when controlling for non-independence.
+(ref:plotReplications5Caption) _The estimated degree of spatial and cultural phylogenetic non-independence predicts reductions in effect size in our reanalysis._ (a) Higher estimated degrees of spatial autocorrelation at 1,000 km distance predict more pronounced reductions in effect sizes when controlling for non-independence.
 (b) Higher estimated levels of cultural phylogenetic signal predict more pronounced reductions in effect sizes when controlling for non-independence.
@@
 
-Yet our systematic literature review and simulation study revealed that the most commonly used controls for non-independence do not sufficiently deal with the issue.
+Crucially, our systematic literature review and simulation study revealed that the most commonly used controls for non-independence do not sufficiently deal with the issue.
 In our simulations, controlling for either latitude or longitude did not reduce false positive rates.
-This result calls into question controls like distance to the equator to account for non-independence in cross-national regression models, though these controls maybe still be suitable to account for regional or latitudinal variation in ecology, which we did not simulate.
+This result calls into question controls like distance to the equator to account for non-independence in cross-national regression models, though these controls may still be suitable to account for regional or latitudinal variation in ecology, which we did not simulate.
 High false positive rates persisted with Conley standard errors, which have previously been critiqued for being overly sensitive to arbitrary distance cutoffs[@Kelly2020].
-The simulation also confirmed the verbal claim that fixed effects for spatial or cultural groupings (e.g.
+The simulation also confirmed the assertion that fixed effects for spatial or cultural groupings (e.g.
 continent or language family fixed effects) are insufficient because non-independence still remains within groupings[@Bromham2018].
@@
 Ours is not the first review to show that studies are misapplying statistical methods in ways that inflate false positive rates.
-For example, other literature reviews have shown that studies in the social sciences tend to use small sample sizes[@Smaldino2016], treat ordinal data as metric[@Liddell2018], incorrectly handle missing values[@Nicholson2017], and ignore best practices in meta-analyses[@Tipton2019].
+For example, other literature reviews have shown that studies in the social sciences tend to use small samples of participants[@Smaldino2016], treat ordinal data as metric[@Liddell2018], incorrectly handle missing values[@Nicholson2017], and ignore best practices in meta-analyses[@Tipton2019].
 Why do cross-national studies also rarely account for non-independence?
-At the institutional level, one possibility is that such practices are incentivised _because_ they inflate false positive rates, increasing the probability of getting published and promoted[@Smaldino2016].
+At the institutional level, one possibility is that such practices are incentivised _because_ they generate statistically significant relationships, which increase the probability that a study is published[@Smaldino2016].
 At the individual level, another possibility is that researchers outside of anthropology and ecology are simply not aware of the problem, or believe that the problem does not apply to analyses of nations.
@@
 We also found that two out of six cross-national correlations with cultural values variables had 95% credible intervals that included zero when controlling for cultural phylogenetic non-independence.
-It is interesting that the economic development analyses were more affected by the spatial controls, while the cultural values analyses were more affected by the cultural phylogenetic controls.
-This highlights that certain kinds of non-independence will be more important depending on the particular national-level outcome and predictor variables of interest.
 
 We do not wish to dissuade researchers from conducting cross-national studies.
-On the contrary, such work promises to deepen our understanding of the causes and consequences of economic development and cultural values around the world.
+On the contrary, such work promises to deepen understanding of our world, including the causes and consequences of economic development and cultural values.
 Moreover, cross-national studies allow social scientists to broaden their scope of study beyond Western populations[@Henrich2010], providing the representative samples necessary to test evolutionary and socio-ecological theories of human behaviour[@Pollet2014;@Oishi2013].
@@
 We also calculated the proportion of articles employing the different types of control listed above at least once: regional fixed effects, distance, shared cultural history, or other.
-For these proportions, we calculated 95% boostrap confidence intervals with 1000 bootstrap iterations.
+For these proportions, we calculated 95% boostrap confidence intervals with 1,000 bootstrap iterations.
 
@@
 Across all model types and parameter combinations, we calculated the false positive rate as the proportion of models that estimated slopes with a 95% confidence / credible interval excluding zero.
-We calculated 95% bootstrap confidence intervals for these false positive rates with 1000 bootstrap iterations.
+We calculated 95% bootstrap confidence intervals for these false positive rates with 1,000 bootstrap iterations.
 
@@
 
-For each individual analysis, we conducted four models: (1) a naive regression replicating the original finding, (2) a regression including a Gaussian process allowing nation random intercepts to covary according to a geographic proximity matrix from latitude and longitude values, (3) a regression including nation random intercepts that covaried according to a linguistic proximity matrix, and (4) a regression including both a geographic Gaussian process and nation random intercepts with linguistic covariance.
+For each individual analysis, we ran four models: (1) a naive regression replicating the original finding, (2) a regression including a Gaussian process allowing nation random intercepts to covary according to a geographic proximity matrix from latitude and longitude values, (3) a regression including nation random intercepts that covaried according to a linguistic proximity matrix, and (4) a regression including both a geographic Gaussian process and nation random intercepts with linguistic covariance.
 See Supplementary Methods for full models.

### 2022-04-18 manuscript.tex
pre-submission manuscript edits

@@
 {}{\PatchFailed}
-\keywords{spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses\newline\indent Word count: 5500 words}
+\keywords{spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses\newline\indent Word count: 5459 words}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} School of Psychology, University of Auckland, Auckland, New Zealand\\\textsuperscript{2} Max Planck Institute for the Science of Human History, Jena, Germany}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} School of Psychology, University of Auckland, Auckland, New Zealand\\\textsuperscript{2} Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany}
 
 \abstract{%
-Cross-national analyses can illuminate global patternings of national-level outcomes.
-But since nations are connected in various ways, such as via spatial proximity and shared cultural ancestry, cross-national analyses often violate assumptions of non-independence, inflating false positive rates.
+Cross-national analyses test hypotheses about the drivers of global variation in national outcomes.
+However, since nations are connected in various ways, such as via spatial proximity and shared cultural ancestry, cross-national analyses often violate assumptions of non-independence, inflating false positive rates.
 Here, we show that, despite being recognised as an important statistical pitfall for over 200 years, cross-national research in economics and psychology still does not sufficiently account for non-independence.
-In a review of the 100 highest-cited cross-national studies of economic development and values, we find that controls for non-independence are rare and have remained so over the past 30 years.
+In a review of the 100 highest-cited cross-national studies of economic development and values, we find that controls for non-independence are rare.
 When studies do include controls for non-independence, our simulations suggest that commonly used methods continue to produce false positives.
@@
 
-Given the importance of nations for structuring human behaviour, there is justifiably huge interest in statistical analyses that attempt to predict variation in national-level outcomes of all kinds.
+Given the importance of nations for structuring human behaviour, there is justifiably huge interest in statistical analyses that attempt to predict variation in national outcomes of all kinds.
 At the time of writing, a search in Web of Science for the terms \enquote{cross-national}, \enquote{cross-cultural}, or \enquote{cross-country} in titles and abstracts returned 80,000 unique hits.
@@
 Shared cultural ancestry can result in a form of pseudoreplication, whereby multiple instances of the same trait across nations are merely duplicates of the ancestral original.
-For example, despite being separated by over 1500 kilometres of ocean, the related island nations Tonga and Tuvalu express similar languages and customs due to cultural inheritance from a shared ancestor deep in the evolutionary history of Polynesian societies.
-More recently, shared ancestry is also crucial in explaining cultural similarities between colonial settlements and their colonisers (e.g.~Argentina and Spain).
+For example, despite being separated by over 1,500 kilometres of ocean, the related island nations Tonga and Tuvalu share similar languages and customs due to cultural inheritance from a common Polynesian population dating back more than 1,000 years.
+More recently, shared ancestry explains a myriad of cultural similarities between colonial settlements and their colonisers (e.g.~Argentina and Spain).
 
@@
 In order to motivate our research question, it is important to first show that economic development and cultural values are indeed spatially and culturally non-independent around the world.
-To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for the Human Development Index\textsuperscript{44} and two dimensions of cultural values from the World Values Survey, traditional vs.~secular values and survival vs.~self-expression values\textsuperscript{16}.
+To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for the Human Development Index\textsuperscript{44} and two primary dimensions of cultural values from the World Values Survey, traditional vs.~secular values and survival vs.~self-expression values\textsuperscript{16}.
 For all three of these variables, we found that a substantial proportion of national-level variation was explained by spatial proximity and shared cultural ancestry between nations (Figure \ref{fig:plotSignal}).
@@
 
-Since economic development and cultural values show evidence of geographic and cultural phylogenetic signal, have cross-national analyses sufficiently accounted for this non-independence?
+Given that economic development and cultural values show evidence of geographic and cultural phylogenetic signal, have cross-national analyses sufficiently accounted for this non-independence?
 To assess this, we systematically searched the published literature for articles that combined the search terms \enquote{economic development} or \enquote{values} with the search terms \enquote{cross-national}, \enquote{cross-cultural}, or \enquote{cross-country}.
@@
 Finally, we included Bayesian multilevel regressions that explicitly model spatial and/or cultural phylogenetic non-independence by allowing nation random intercepts to covary according to global geographic and/or linguistic proximity matrices (see Supplementary Methods).
-To model spatial non-independence we included a Gaussian process\textsuperscript{48,49} over latitude and longitude values, and to model cultural phylogenetic non-independence we assumed that nation random intercepts were correlated proportional to their linguistic proximity\textsuperscript{50}.
+To model spatial non-independence we included a Gaussian process\textsuperscript{48,49} over latitude and longitude values, and to model cultural phylogenetic non-independence we assumed that nation random intercepts were correlated in proportion to their linguistic proximity\textsuperscript{50}.
 Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero (i.e.~falsely infering a relationship when none is present).
@@
 For each outcome variable, our Gaussian process models provided varying estimates of how quickly spatial autocorrelation declined with distance (Supplementary Figure \ref{fig:plotReplications3}).
-For example, in Skidmore and Toya\textsuperscript{56} gross domestic product growth was only moderately spatially autocorrelated at 1000 km distance (posterior median spatial autocorrelation at 1000 km = 0.47, 95\% CI {[}0.06 0.97{]}), whereas in Inglehart and Baker\textsuperscript{16} traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1000 km = 0.96, 95\% CI {[}0.78 0.99{]}).
+For example, in Skidmore and Toya\textsuperscript{56} gross domestic product growth was only moderately spatially autocorrelated at 1,000 km distance (posterior median spatial autocorrelation at 1,000 km = 0.47, 95\% CI {[}0.06 0.97{]}), whereas in Inglehart and Baker\textsuperscript{16} traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1,000 km = 0.96, 95\% CI {[}0.78 0.99{]}).
 We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \ref{fig:plotReplications4}), with some outcome variables expressing low signal (e.g.~tightness\textsuperscript{14}; posterior median = 0.11, 95\% CI {[}0.00 0.82{]}) and others expressing high signal (e.g.~female labour force participation\textsuperscript{58}; posterior median = 0.91, 95\% CI {[}0.67 0.99{]}).
@@
 \includegraphics{manuscript_files/figure-latex/plotReplications5-1.pdf}
-\caption{\label{fig:plotReplications5}\emph{The estimated degree of spatial and cultural phylogenetic non-independence predicts reductions in effect size in our reanalysis.} (a) Higher estimated degrees of spatial autocorrelation at 1000 km distance predict more pronounced reductions in effect sizes when controlling for non-independence.
+\caption{\label{fig:plotReplications5}\emph{The estimated degree of spatial and cultural phylogenetic non-independence predicts reductions in effect size in our reanalysis.} (a) Higher estimated degrees of spatial autocorrelation at 1,000 km distance predict more pronounced reductions in effect sizes when controlling for non-independence.
 (b) Higher estimated levels of cultural phylogenetic signal predict more pronounced reductions in effect sizes when controlling for non-independence.
@@
 
-Yet our systematic literature review and simulation study revealed that the most commonly used controls for non-independence do not sufficiently deal with the issue.
+Crucially, our systematic literature review and simulation study revealed that the most commonly used controls for non-independence do not sufficiently deal with the issue.
 In our simulations, controlling for either latitude or longitude did not reduce false positive rates.
-This result calls into question controls like distance to the equator to account for non-independence in cross-national regression models, though these controls maybe still be suitable to account for regional or latitudinal variation in ecology, which we did not simulate.
+This result calls into question controls like distance to the equator to account for non-independence in cross-national regression models, though these controls may still be suitable to account for regional or latitudinal variation in ecology, which we did not simulate.
 High false positive rates persisted with Conley standard errors, which have previously been critiqued for being overly sensitive to arbitrary distance cutoffs\textsuperscript{39}.
-The simulation also confirmed the verbal claim that fixed effects for spatial or cultural groupings (e.g.~continent or language family fixed effects) are insufficient because non-independence still remains within groupings\textsuperscript{40}.
+The simulation also confirmed the assertion that fixed effects for spatial or cultural groupings (e.g.~continent or language family fixed effects) are insufficient because non-independence still remains within groupings\textsuperscript{40}.
 This logic further applies to analyses that control for non-independence by separately analysing different regions (e.g.\textsuperscript{61}).
@@
 Ours is not the first review to show that studies are misapplying statistical methods in ways that inflate false positive rates.
-For example, other literature reviews have shown that studies in the social sciences tend to use small sample sizes\textsuperscript{63}, treat ordinal data as metric\textsuperscript{64}, incorrectly handle missing values\textsuperscript{65}, and ignore best practices in meta-analyses\textsuperscript{66}.
+For example, other literature reviews have shown that studies in the social sciences tend to use small samples of participants\textsuperscript{63}, treat ordinal data as metric\textsuperscript{64}, incorrectly handle missing values\textsuperscript{65}, and ignore best practices in meta-analyses\textsuperscript{66}.
 Why do cross-national studies also rarely account for non-independence?
-At the institutional level, one possibility is that such practices are incentivised \emph{because} they inflate false positive rates, increasing the probability of getting published and promoted\textsuperscript{63}.
+At the institutional level, one possibility is that such practices are incentivised \emph{because} they generate statistically significant relationships, which increase the probability that a study is published\textsuperscript{63}.
 At the individual level, another possibility is that researchers outside of anthropology and ecology are simply not aware of the problem, or believe that the problem does not apply to analyses of nations.
@@
 We also found that two out of six cross-national correlations with cultural values variables had 95\% credible intervals that included zero when controlling for cultural phylogenetic non-independence.
-It is interesting that the economic development analyses were more affected by the spatial controls, while the cultural values analyses were more affected by the cultural phylogenetic controls.
-This highlights that certain kinds of non-independence will be more important depending on the particular national-level outcome and predictor variables of interest.
 
 We do not wish to dissuade researchers from conducting cross-national studies.
-On the contrary, such work promises to deepen our understanding of the causes and consequences of economic development and cultural values around the world.
+On the contrary, such work promises to deepen understanding of our world, including the causes and consequences of economic development and cultural values.
 Moreover, cross-national studies allow social scientists to broaden their scope of study beyond Western populations\textsuperscript{7}, providing the representative samples necessary to test evolutionary and socio-ecological theories of human behaviour\textsuperscript{8,70}.
@@
 We also calculated the proportion of articles employing the different types of control listed above at least once: regional fixed effects, distance, shared cultural history, or other.
-For these proportions, we calculated 95\% boostrap confidence intervals with 1000 bootstrap iterations.
+For these proportions, we calculated 95\% boostrap confidence intervals with 1,000 bootstrap iterations.
 
@@
 Across all model types and parameter combinations, we calculated the false positive rate as the proportion of models that estimated slopes with a 95\% confidence / credible interval excluding zero.
-We calculated 95\% bootstrap confidence intervals for these false positive rates with 1000 bootstrap iterations.
+We calculated 95\% bootstrap confidence intervals for these false positive rates with 1,000 bootstrap iterations.
 
@@
 
-\leavevmode\hypertarget{ref-Brownrigg2018}{}%
-77. Brownrigg, R. \emph{maps: Draw geographical maps}. (2018).
-
-\leavevmode\hypertarget{ref-Hijmans2019}{}%
-78. Hijmans, R. J. \emph{geosphere: Spherical trigonometry}. (2019).
-
-\leavevmode\hypertarget{ref-Glottolog}{}%
-79.
-Hammarström, H., Forkel, R., Haspelmath, M.
-\& Bank, S.
-\emph{Glottolog 3.0}.
-(Max Planck Institute for the Science of Human History, 2017).
-doi:\href{https://doi.org/10.5281/zenodo.4061162}{10.5281/zenodo.4061162}.
-
-\leavevmode\hypertarget{ref-Ethnologue}{}%
-80. \emph{Ethnologue: Languages of the world}. (SIL International, 2018).
-
-\leavevmode\hypertarget{ref-Eff2008}{}%
-81. Eff, E. A. Weight matrices for cultural proximity: Deriving weights from a language phylogeny. \emph{Structure and Dynamics} \textbf{3}, (2008).
-
 \endgroup
@@
 
-Geographic distance between two nations was calculated as the logged geodesic distance between country capital cities (data from the R package \emph{maps}\textsuperscript{77}) using the R package \emph{geosphere}\textsuperscript{78}.
+Geographic distance between two nations was calculated as the logged geodesic distance between country capital cities (data from the R package \emph{maps}; Brownrigg, 2018) using the R package \emph{geosphere} (Hijmans, 2019).
 The geographic proximity matrix was computed as one minus the log geographic distance matrix scaled between 0 and 1.
@@
 Linguistic proximity between two nations was calculated as the cultural proximity between all languages spoken within those nations, weighted by speaker percentages.
-We acquired cultural proximity data by combining the language family trees provided by Glottolog v3.0\textsuperscript{79} into one global language tree (undated and unresolved).
+We acquired cultural proximity data by combining the language family trees provided by Glottolog v3.0 (Hammarström et al., 2017) into one global language tree (undated and unresolved).
 We calculated cultural proximity \(s\) between two languages \(j\) and \(k\) as the distance (in number of nodes traversed) of their most recent common ancestor \(i\) to the root of the tree, through the formula:
@@
 where \(n_{r}\) is the maximum path length (in number of nodes traversed) leading to the pan-human root \(r\), and \(n_{i}\) is the maximum path length leading to node \(i\).
-We then combined these proximities with speaker data from Ethnologue 21\textsuperscript{80} and compared every language spoken within those nations by at least 1 permille of the population, weighted by speaker percentages, through the formula:
+We then combined these proximities with speaker data from Ethnologue 21 (Ethnologue, 2018) and compared every language spoken within those nations by at least 1 permille of the population, weighted by speaker percentages, through the formula:
 
@@
 
-where \(p_{lj}\) is the percentage of the population in nation \(l\) speaking language \(j\), \(p_{mk}\) is the percentage of the population in nation \(m\) speaking language \(k\), and \(s_{jk}\) is the proximity measure between languages \(j\) and \(k\)\textsuperscript{81}.
+where \(p_{lj}\) is the percentage of the population in nation \(l\) speaking language \(j\), \(p_{mk}\) is the percentage of the population in nation \(m\) speaking language \(k\), and \(s_{jk}\) is the proximity measure between languages \(j\) and \(k\) (Eff, 2008).
 The resulting linguistic proximity matrix was also scaled between 0 and 1 before analysis.
@@
 To control for spatial non-independence, we add a Gaussian process to this model and feed it a scaled geographic distance matrix \(D\) based on Euclidean distances between latitude and longitude coordinates.
-This distance matrix is computed internally by the R package \emph{brms}\textsuperscript{69}.
+This distance matrix is computed internally by the R package \emph{brms} (Bürkner, 2017).
 The Gaussian process uses an exponentiated quadratic covariance kernel, the only covariance kernel currently supported by \emph{brms}.
@@
 
+Brownrigg, R. \emph{maps: Draw geographical maps.} (2018).
+
+Bürkner, P.-C. brms: An R package for Bayesian multilevel models using Stan. \emph{Journal of Statistical Software} \textbf{80}, 1---28 (2017).
+
+Eff, E. A. Weight matrices for cultural proximity: Deriving weights from a language phylogeny. \emph{Structure and Dynamics} \textbf{3}, (2008).
+
+\emph{Ethnologue: Languages of the world.} (SIL International, 2018).
+
+Hammarström, H., Forkel, R., Haspelmath, M.
+\& Bank, S.
+\emph{Glottolog 3.0.} (Max Planck Institute for the Science of Human History, 2017).
+\url{doi:\%5B10.5281/zenodo.4061162\%5D(https://doi.org/10.5281/zenodo.4061162)}
+
+Hijmans, R. J. \emph{geosphere: Spherical trigonometry.} (2019).
+
 \end{document}

### 2022-04-29 vignette.Rmd
add vignette

---
title: "How to control for non-independence in cross-national analyses: An R tutorial"
author: "Scott Claessens"
date: "29/04/2022"
output: md_document
---

```{r echo=F}
options(width = 300)
```

I recently posted a pre-print on PsyArxiv entitled ["The non-independence of nations and why it matters"](https://psyarxiv.com/m6bsn/). In the paper, myself and Quentin Atkinson argue that many cross-national correlations in the social sciences are flawed because they do not adequately take into account various forms of non-independence between nations. Nations are linked to one another via spatial proximity and shared cultural ancestry, and failing to account for these dependencies can increase the chance of false positives.

That's the issue: but how to rectify it? In the paper, we suggest some Bayesian modelling techniques that researchers can use to overcome the problem. However, we did not go into these techniques in detail. So I thought it would be useful to provide a brief R tutorial on the methods we used in the paper, so that interested researchers can apply them to their own cross-national datasets.

In this blog post, I will walk through methods of dealing with (1) spatial non-independence, and (2) cultural phylogenetic non-independence. In each case, I will use simulated data, show that a naive cross-national correlation produces an inflated effect size, and then present an alternative method that controls for non-independence. The techniques I will use are not the only ways of dealing with non-independence, but they are the techniques I learned through folks like Richard McElreath and Paul-Christian Bürkner.

For this tutorial, we will need the following R packages: _tidyverse_ for plotting and data wrangling, _brms_ for fitting Bayesian models, _ggcorrplot_ for plotting proximity matrices, _ggrepel_ for neatly plotting country labels, _readxl_ for reading in Excel files, and _targets_ for extracting objects from our reproducible analysis pipeline.

```{r warning=FALSE, error=FALSE, message=FALSE}
library(tidyverse)
library(brms)
library(ggcorrplot)
library(ggrepel)
library(readxl)
library(targets)
```

The data and code for this vignette can be found on GitHub: https://github.com/ScottClaessens/crossNationalCorrelations

# Spatial non-independence

## Spatially non-independent simulated data

For this demonstration, we load a subset of OECD nations from one of the many simulated datasets from our paper. This dataset contains a standardised predictor variable (X) and a standardised outcome variable (Y), both of which are strongly spatially autocorrelated.

```{r echo=FALSE}
# load data from targets cache
spatialSimData <- tar_read(simData_simGeoCov_0.8_0.8_1728b220)
# only OECD countries for example, country iso codes
oecd <- c("AT","AU","BE","CA","CH","CL","CZ","DE","DK",
          "EE","ES","FI","FR","GB","GR","HU","IE","IL",
          "IS","IT","JP","KR","LT","LU","LV","MX","NL",
          "NO","NZ","PL","PT","SE","SI","SK","TR","US")
spatialSimData <- filter(spatialSimData, country %in% oecd)
# write data to file
write.csv(file = "data/vignette/spatialSimData.csv", 
          spatialSimData, row.names = FALSE)
```

```{r}
# load dataset from file
spatialSimData <- read.csv("data/vignette/spatialSimData.csv")
```

We can plot the relationship between the nation-level predictor and outcome variables.

```{r warning=FALSE, message=FALSE}
ggplot(spatialSimData, aes(x, y, label = country)) +
  geom_point() +
  geom_text_repel() +
  geom_smooth(method = "lm") +
  theme_classic()
```

The relationship between the simulated variables appears to be positive. But is this just due to spatial non-independence? From the plot, it looks like this might be the case. For example, neighbouring countries, such as Sweden (SE) and Finland (FI), have very similar values on both variables, while the distant countries Australia (AU) and New Zealand (NZ) are far off in the top right hand corner of the plot.

## Fit naive correlation

We can easily get the naive Pearson's correlation between the national-level variables.

```{r}
cor(spatialSimData$y, spatialSimData$x) %>% round(2)
```
We can also get this same correlation from fitting a Bayesian regression. This is where we use the R package _brms_. We set some regularising priors on the intercept, slope, and residual variance in this model.

```{r message=FALSE, warning=FALSE, error=FALSE}
# fit naive bayesian regression
m1.1 <- brm(y ~ x, data = spatialSimData, 
            cores = 4, seed = 2113,
            # regularising priors
            prior = c(prior(normal(0, 1), class = Intercept),
                      prior(normal(0, 1), class = b),
                      prior(exponential(2), class = sigma)))
```

```{r}
# get model summary
summary(m1.1)
```

Because both variables are standardised, the slope is comparable to the Pearson's correlation reported above. The model converged normally, and the 95% credible interval for the slope is above zero. In other words, this is a "significantly positive" relationship.

## Fit correlation with Gaussian Process

But is spatial non-independence ruining our inference? To find out, we fit another model including a Gaussian Process over latitudes and longitudes for nations. A Gaussian Process adds a random intercept for each nation, and these random intercepts are allowed to covary according to the distance between latitude-longitude coordinates. For a more detailed discussion of this method of dealing with spatial autocorrelation, see Section 14.5 in Richard McElreath's book Statistical Rethinking (second edition), or his 2022 lecture on the topic [here](https://www.youtube.com/watch?v=PIuqxOBJqLU).

We can include Gaussian Processes in _brms_ models using the `gp()` function. This function takes a variable --- or multiple variables, in our case --- and, under the hood, computes a normalised distance matrix between all the cases for that variable. It then estimates a covariance function specifying how these distances relate to the covariance between nations. If the data are spatially autocorrelated, the model will estimate strong spatial covariance between nations, which could "soak up" much of the relationship we saw in the previous section.

We fit this model by adding a `gp()` term to our formula.

```{r message=FALSE, warning=FALSE, error=FALSE}
# fit Bayesian regression with Gaussian Process
m1.2 <- brm(y ~ x + gp(latitude, longitude), 
            data = spatialSimData, cores = 4, seed = 2113,
            # regularising priors
            prior = c(prior(normal(0, 1), class = Intercept),
                      prior(normal(0, 1), class = b),
                      prior(exponential(2), class = sdgp),
                      prior(exponential(2), class = sigma)),
            # tune the mcmc sampler
            control = list(adapt_delta = 0.99))
```

```{r}
# get model summary
summary(m1.2)
```

As expected, the slope has reduced in size, and the 95% credible interval now includes zero. This suggests that the positive relationship we initially found in our naive regression was confounded by spatial non-independence.

# Cultural phylogenetic non-independence

## Culturally non-independent simulated data

It is also possible for nations to be culturally non-independent --- that is, they share common influences from shared cultural ancestors. For example, even though the United Kingdom and New Zealand are on opposite sides of the planet, they are culturally similar due to the effects of British colonialism. This source of non-independence can be just as important as spatial proximity, especially when we are interested in cultural norms and institutions.

As before, we load a subset of OECD nations from a simulated dataset containing a standardised predictor variable (X) and a standardised outcome variable (Y). Both of these variables are strongly culturally autocorrelated.

```{r echo=FALSE}
# load data from targets cache
culturalSimData <- tar_read(simData_simLinCov_0.8_0.8_0fb442ad)
# only OECD countries
culturalSimData <- filter(culturalSimData, country %in% oecd)
# write data to file
write.csv(file = "data/vignette/culturalSimData.csv", 
          culturalSimData, row.names = FALSE)
```

```{r}
# load dataset from file
culturalSimData <- read.csv("data/vignette/culturalSimData.csv")
```

Let's plot the relationship between the nation-level predictor and outcome variables.

```{r warning=FALSE, message=FALSE}
ggplot(culturalSimData, aes(x, y, label = country)) +
  geom_point() +
  geom_text_repel() +
  geom_smooth(method = "lm") +
  theme_classic()
```

There's a positive relationship between the variables. But a close look at the plot suggests some cultural clustering. For example, the English-speaking nations Australia (AU), New Zealand (NZ), the United States (US), and the United Kingdom (GB) are all towards the left of the plot, whereas nations speaking more distantly related languages, such as Japan (JP) and Greece (GR), are further to the right.

## Fit naive correlation

As before, we can get the naive Pearson's correlation.

```{r}
cor(culturalSimData$y, culturalSimData$x) %>% round(2)
```

And we can estimate this correlation in a naive Bayesian regression model.

```{r message=FALSE, warning=FALSE, error=FALSE}
# fit naive bayesian regression
m2.1 <- brm(y ~ x, data = culturalSimData, 
            cores = 4, seed = 2113,
            # regularising priors
            prior = c(prior(normal(0, 1), class = Intercept),
                      prior(normal(0, 1), class = b),
                      prior(exponential(2), class = sigma)))
```

```{r}
# get model summary
summary(m2.1)
```

The 95% credible interval for the slope is above zero, suggesting that there is a "significantly positive" relationship.

## Fit correlation with covarying random effects

"Cultural similarity" does not have a fixed coordinate system like longitude and latitude, making Gaussian Processes tricky in _brms_. Instead, we can use another technique that allows nation random intercepts to covary, but we _specify the covariance matrix in advance rather than estimate it in the model_. This will allow us to control for cultural non-independence.

What covariance matrix should we specify in advance? The matrix that we use in the paper is the linguistic proximity between nations, weighted by the proportion of speakers of each language in each nation. "Linguistic proximity" here means how related two languages are on a phylogeny of languages. We just average this measure over all languages spoken in each nation, weighted by speaker percentages. This gives us a sense of how "culturally similar" different nations are.

It is simpler to just show the matrix. Below, we load the full linguistic distance matrix, subset it to just our OECD nations, and convert it to a linguistic proximity matrix.

```{r}
# load full linguistic distance matrix
linDist <- 
  read_xlsx("data/networks/2F Country Distance 1pml adj.xlsx") %>%
  select(-ISO) %>%
  as.matrix()

# rows and columns identical in symmetrical matrix
rownames(linDist) <- colnames(linDist)

# subset to only OECD countries
linDist <- linDist[oecd,oecd]

# nation's distance from itself is zero
diag(linDist) <- 0

# normalise between 0 and 1
linDist <- linDist / max(linDist)

# convert to linguistic proximity
linProx <- 1 - linDist
```

We can then plot this `linProx` object, to get a sense of the cultural relatedness of different nations.

```{r fig.height=8,fig.width=8}
ggcorrplot(linProx, hc.order = TRUE, type = "lower") +
  theme(legend.title = element_blank())
```

From this plot, we can see that majority English-speaking nations have high linguistic proximity (i.e. cultural similarity), as do Scandinavian nations, nations in Western Europe, and nations in Eastern Europe.

We can use this linguistic proximity matrix in our modelling. In _brms_, we can specify in advance how our random effects should be correlated by using the `cov` argument in the `gr()` function. The only restrictions are that the matrix that we feed in must have row and column names that match our variable, and that the matrix is positive definite. To learn more about this method, see [this vignette on phylogenetic regression](https://cran.r-project.org/web/packages/brms/vignettes/brms_phylogenetics.html) from Paul-Christian Bürkner.

```{r message=FALSE, warning=FALSE, error=FALSE}
# fit bayesian regression with covarying random effects
m2.2 <- brm(y ~ x + (1 | gr(country, cov = linProx)), 
            data = culturalSimData, cores = 4, seed = 2113,
            # pre-specified covariance matrix must be passed to brms via data2 argument
            data2 = list(linProx = linProx),
            # regularising priors
            prior = c(prior(normal(0, 1), class = Intercept),
                      prior(normal(0, 1), class = b),
                      prior(exponential(2), class = sd),
                      prior(exponential(2), class = sigma)),
            # tune mcmc sampler
            control = list(adapt_delta = 0.999))
```

```{r}
# get model summary
summary(m2.2)
```

Note that there were a number of divergent transitions, even after we increased `adapt_delta`, and the number of effective samples are not great. But the Rhat values are all beneath 1.05, suggesting that this model is reasonable.

As expected, we find that the positive relationship between our variables goes away when we control for cultural phylogenetic non-independence between nations.

# Applying these methods in your research

There are a number of advantages to using Gaussian Processes and covarying random effects to deal with various forms of non-independence, beyond what I've outlined here. Both methods are possible to use when we have multiple data points per nation. Both methods can also be used for any conceivable likelihood --- logistic regression for binary variables, cumulative link regression for ordinal variables, beta regression for variables bounded between 0 and 1, etc. All of this is possible using the _brms_ package.

Happy cross-national modelling!

### 2022-04-29 _targets.R

@@ -419,5 +419,9 @@ list(
   
   #### Manuscript ####
   
-  tar_render(manuscript, "manuscript.Rmd")
+  tar_render(manuscript, "manuscript.Rmd"),
+  
+  #### Vignette ####
+  
+  tar_render(vignette, "vignette.Rmd")
 )

### 2022-10-17 R/functionsSignal.R
added tightness data

@@ -129,47 +129,6 @@ fitHDISignal <- function(hdi, geoCov, linCov) {
       control = list(adapt_delta = 0.9))
 }
 
-# signal for wvs
-fitWVSSignal <- function(wvs, outcome = "", geoCov, linCov) {
-  wvs <-
-    wvs %>%
-    # select vars
-    dplyr::select(
-      # country code
-      S009,
-      # traditional vs. secular-rational values
-      F063, Y003, F120, G006, E018,
-      # survival vs. self-expression values
-      Y002, A008, F118, E025, A165
-    ) %>%
-      # remove missings
-      drop_na() %>%
-      filter(Y003 != -5)
-  # get cultural map
-  # https://www.worldvaluessurvey.org/WVSContents.jsp?CMSID=Findings
-  m <- pca(select(wvs, -S009), nfactors = 2, rotate = "varimax")
-  # summarise and prepare for modelling
-  wvs <-
-    wvs %>%
-    mutate(trad = as.numeric(scale(m$scores[,2])),
-           surv = as.numeric(scale(m$scores[,1])),
-           isoGeo = S009, 
-           isoLin = S009)
-  # subset matrices
-  geoCov <- geoCov[rownames(geoCov)[rownames(geoCov) %in% wvs$S009],
-                   colnames(geoCov)[colnames(geoCov) %in% wvs$S009]]
-  linCov <- linCov[rownames(linCov)[rownames(linCov) %in% wvs$S009],
-                   colnames(linCov)[colnames(linCov) %in% wvs$S009]]
-  # fit model
-  bf1 <- bf(paste0(outcome, " ~ 0 + Intercept + (1 | gr(isoGeo, cov = geoCov)) + (1 | gr(isoLin, cov = linCov)) + (1 | S009)"))
-  brm(formula = bf1, data = wvs, data2 = list(geoCov = geoCov, linCov = linCov),
-      prior = c(prior(normal(0, 0.1), class = b),
-                prior(exponential(8), class = sd),
-                prior(exponential(8), class = sigma)),
-      seed = 2113, iter = 2000, cores = 4, sample_prior = "yes",
-      control = list(adapt_delta = 0.95, max_treedepth = 15))
-}
-
 # signal for gdp
 fitGDPSignal <- function(gdp, geoCov, linCov) {
   # add iso variables for modelling
@@ -228,6 +187,67 @@ fitGiniSignal <- function(gini, geoCov, linCov) {
[…]

### 2022-10-17 _targets.R

@@ -63,37 +63,53 @@ list(
   tar_target(fileGDP, "data/gdp/API_NY.GDP.PCAP.CD_DS2_en_csv_v2_4666475.csv", format = "file"),
   tar_target(fileGDPGrowth, "data/gdpGrowth/API_NY.GDP.PCAP.KD.ZG_DS2_en_csv_v2_4666256.csv", format = "file"),
   tar_target(fileGini, "data/gini/API_SI.POV.GINI_DS2_en_csv_v2_4666600.csv", format = "file"),
+  tar_target(fileTightness, "data/tightness/Tightness_Scores.xlsx", format = "file"),
   # isocodes (https://gist.github.com/tadast/8827699)
   tar_target(iso, loadISO(fileISO)),
-  # load datasets
+  # load datasets for signal models
   tar_target(hdi, loadHDIData(fileHDI, fileISOHDI)),
   tar_target(wvs, haven::read_sav(fileWVS, encoding = "latin1")),
   tar_target(gdp, loadGDPData(fileGDP, iso)),
   tar_target(gdpGrowth, loadGDPGrowthData(fileGDPGrowth, iso)),
   tar_target(gini, loadGiniData(fileGini, iso)),
+  tar_target(tightness, read_xlsx(fileTightness)),
   # covariance matrices
   tar_target(geoCov, loadCovMat(fileGeo, log = TRUE)),
   tar_target(linCov, loadCovMat(fileLin, log = FALSE)),
-  # geographic and linguistic signal
-  tar_target(hdiSignal, fitHDISignal(hdi, geoCov, linCov)),
-  tar_target(tradSignal, fitWVSSignal(wvs, outcome = "trad", geoCov, linCov)),
-  tar_target(survSignal, fitWVSSignal(wvs, outcome = "surv", geoCov, linCov)),
-  tar_target(gdpSignal, fitGDPSignal(gdp, geoCov, linCov)),
-  tar_target(gdpGrowthSignal, fitGDPGrowthSignal(gdpGrowth, geoCov, linCov)),
-  tar_target(giniSignal, fitGiniSignal(gini, geoCov, linCov)),
+  # fit geographic and linguistic signal models
+  tar_target(signalHDI,   fitHDISignal(hdi, geoCov, linCov)),
+  tar_target(signalGDP,   fitGDPSignal(gdp, geoCov, linCov)),
[…]

### 2022-10-17 R/functionsSignal.R
add individualism data

@@ -250,6 +250,28 @@ fitTightnessSignal <- function(tightness, geoCov, linCov) {
       control = list(adapt_delta = 0.9999, max_treedepth = 15))
 }
 
+# signal for individualism
+fitIndividualismSignal <- function(fincherData, geoCov, linCov) {
+  # add iso variables for modelling
+  fincherData$isoGeo <- fincherData$iso
+  fincherData$isoLin <- fincherData$iso
+  # subset matrices
+  geoCov <- geoCov[rownames(geoCov)[rownames(geoCov) %in% fincherData$iso],
+                   colnames(geoCov)[colnames(geoCov) %in% fincherData$iso]]
+  linCov <- linCov[rownames(linCov)[rownames(linCov) %in% fincherData$iso],
+                   colnames(linCov)[colnames(linCov) %in% fincherData$iso]]
+  # outcome variable between 0 and 1
+  fincherData <- mutate(fincherData, individualismHofstede = individualismHofstede / 100)
+  # fit model
+  brm(individualismHofstede ~ 0 + Intercept + (1 | gr(isoGeo, cov = geoCov)) + (1 | gr(isoLin, cov = linCov)),
+      data = fincherData, data2 = list(geoCov = geoCov, linCov = linCov),
+      prior = c(prior(normal(0.5, 0.1), class = b),
+                prior(exponential(8), class = sd),
+                prior(exponential(8), class = sigma)),
+      iter = 6000, seed = 2113, cores = 4, sample_prior = "yes",
+      control = list(adapt_delta = 0.9999, max_treedepth = 15))
+}
+
 # plot signal
 plotGeoLinSignal <- function(postHDI, postTrad, postSurv) {
   # plotting function

### 2022-10-17 _targets.R

@@ -84,6 +84,7 @@ list(
   tar_target(signalTrad,  fitWVSSignal(wvs, outcome = "trad", geoCov, linCov)),
   tar_target(signalSurv,  fitWVSSignal(wvs, outcome = "surv", geoCov, linCov)),
   tar_target(signalTight, fitTightnessSignal(tightness, geoCov, linCov)),
+  tar_target(signalInd,   fitIndividualismSignal(fincherData, geoCov, linCov)),
   # posterior samples
   tar_target(postHDI,   as_draws_array(signalHDI, variable = "^sd_", regex = TRUE)),
   tar_target(postGDP,   as_draws_array(signalGDP, variable = "^sd_", regex = TRUE)),
@@ -92,6 +93,7 @@ list(
   tar_target(postTrad,  as_draws_array(signalTrad, variable = "^sd_", regex = TRUE)),
   tar_target(postSurv,  as_draws_array(signalSurv, variable = "^sd_", regex = TRUE)),
   tar_target(postTight, as_draws_array(signalTight, variable = c("^sd_", "sigma"), regex = TRUE)),
+  tar_target(postInd,   as_draws_array(signalInd, variable = c("^sd_", "sigma"), regex = TRUE)),
   # calculate geographic signal
   tar_target(geoHDI,   hypothesis(signalHDI,   "(sd_isoGeo__Intercept^2/(sd_isoGeo__Intercept^2+sd_isoLin__Intercept^2+sd_iso__Intercept^2))=0",  class = NULL)),
   tar_target(geoGDP,   hypothesis(signalGDP,   "(sd_isoGeo__Intercept^2/(sd_isoGeo__Intercept^2+sd_isoLin__Intercept^2+sd_iso2__Intercept^2))=0",  class = NULL)),
@@ -100,6 +102,7 @@ list(
   tar_target(geoTrad,  hypothesis(signalTrad,  "(sd_isoGeo__Intercept^2/(sd_isoGeo__Intercept^2+sd_isoLin__Intercept^2+sd_S009__Intercept^2))=0", class = NULL)),
   tar_target(geoSurv,  hypothesis(signalSurv,  "(sd_isoGeo__Intercept^2/(sd_isoGeo__Intercept^2+sd_isoLin__Intercept^2+sd_S009__Intercept^2))=0", class = NULL)),
[…]

### 2022-10-18 R/functionsSimulation.R
update code for simulation figures

@@ -230,7 +230,7 @@ plotSimInd <- function(olsModel1, olsModel2, olsModel3, olsModel4, olsModel5,
   plotFun <- function(results, title, ylab) {
     results %>%
       # strong autocorrelation levels
-      filter(rho == 0.8 & lambda == 0.8) %>%
+      filter(rho == 0.8 & lambda == 0.8 & r == 0) %>%
       # order by slope
       mutate(id = factor(1:nrow(.))) %>%
       ggplot(aes(x = fct_reorder(id, bX, .desc = TRUE), 
@@ -286,6 +286,7 @@ plotSimAll <- function(olsModel1, olsModel2, olsModel3, olsModel4, olsModel5,
   # plotting function
   plotFun <- function(data, title, ylab, xlab) {
     data %>%
+      filter(r == 0) %>%
       group_by(lambda, rho) %>%
       summarise(prop = sum(sig) / n(), n = n()) %>%
       mutate(lower = map2(prop, n, function(x, y) getBootCIProp(x, y, 1000)[1]), 

### 2022-10-18 _targets.R

@@ -185,29 +185,29 @@ list(
   tar_combine(conleyModel1_simLinCov, simulationTargets$conleyModel1[37:72], command = dplyr::bind_rows(!!!.x)),
   tar_combine(conleyModel2_simGeoCov, simulationTargets$conleyModel2[1:36],  command = dplyr::bind_rows(!!!.x)),
   tar_combine(conleyModel2_simLinCov, simulationTargets$conleyModel2[37:72], command = dplyr::bind_rows(!!!.x)),
-  tar_combine(brmsModel1_simGeoCov,   simulationTargets$brmsModel1[1:36],    command = dplyr::bind_rows(!!!.x)),
-  tar_combine(brmsModel1_simLinCov,   simulationTargets$brmsModel1[37:72],   command = dplyr::bind_rows(!!!.x)),
-  tar_combine(brmsModel2_simGeoCov,   simulationTargets$brmsModel2[1:36],    command = dplyr::bind_rows(!!!.x)),
-  tar_combine(brmsModel2_simLinCov,   simulationTargets$brmsModel2[37:72],   command = dplyr::bind_rows(!!!.x)),
-  tar_combine(brmsModel3_simGeoCov,   simulationTargets$brmsModel3[1:36],    command = dplyr::bind_rows(!!!.x)),
-  tar_combine(brmsModel3_simLinCov,   simulationTargets$brmsModel3[37:72],   command = dplyr::bind_rows(!!!.x)),
+  #tar_combine(brmsModel1_simGeoCov,   simulationTargets$brmsModel1[1:36],    command = dplyr::bind_rows(!!!.x)),
+  #tar_combine(brmsModel1_simLinCov,   simulationTargets$brmsModel1[37:72],   command = dplyr::bind_rows(!!!.x)),
+  #tar_combine(brmsModel2_simGeoCov,   simulationTargets$brmsModel2[1:36],    command = dplyr::bind_rows(!!!.x)),
+  #tar_combine(brmsModel2_simLinCov,   simulationTargets$brmsModel2[37:72],   command = dplyr::bind_rows(!!!.x)),
+  #tar_combine(brmsModel3_simGeoCov,   simulationTargets$brmsModel3[1:36],    command = dplyr::bind_rows(!!!.x)),
+  #tar_combine(brmsModel3_simLinCov,   simulationTargets$brmsModel3[37:72],   command = dplyr::bind_rows(!!!.x)),
[…]

### 2022-10-20 manuscript.Rmd
include full posteriors in effect size analyses and update figure

@@
 ```{r load autocor}
-tar_load(c(inglehartSA, skidmoreSA, alesinaSignal, gelfandSignal))
+tar_load(c(inglehartSA1, skidmoreSA1, alesinaSignal1, gelfandSignal1))
 ```
@@
 For each outcome variable, our Gaussian process models provided varying estimates of how quickly spatial autocorrelation declined with distance (Supplementary Figure \@ref(fig:plotReplications3)).
-For example, in Skidmore and Toya[@Skidmore2002] gross domestic product growth was only moderately spatially autocorrelated at 1,000 km distance (posterior median spatial autocorrelation at 1,000 km = `r printnum(median(skidmoreSA))`, 95% CI [`r printnum(quantile(skidmoreSA, 0.025))` `r printnum(quantile(skidmoreSA, 0.975))`]), whereas in Inglehart and Baker[@Inglehart2000] traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1,000 km = `r printnum(median(inglehartSA))`, 95% CI [`r printnum(quantile(inglehartSA, 0.025))` `r printnum(quantile(inglehartSA, 0.975))`]).
+For example, in Skidmore and Toya[@Skidmore2002] gross domestic product growth was only moderately spatially autocorrelated at 1,000 km distance (posterior median spatial autocorrelation at 1,000 km = `r printnum(median(skidmoreSA1))`, 95% CI [`r printnum(quantile(skidmoreSA1, 0.025))` `r printnum(quantile(skidmoreSA1, 0.975))`]), whereas in Inglehart and Baker[@Inglehart2000] traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1,000 km = `r printnum(median(inglehartSA1))`, 95% CI [`r printnum(quantile(inglehartSA1, 0.025))` `r printnum(quantile(inglehartSA1, 0.975))`]).
 We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \@ref(fig:plotReplications4)), with some outcome variables expressing low signal (e.g.
-tightness[@Gelfand2011]; posterior median = `r printnum(median(gelfandSignal))`, 95% CI [`r printnum(quantile(gelfandSignal, 0.025))` `r printnum(quantile(gelfandSignal, 0.975))`]) and others expressing high signal (e.g.
-female labour force participation[@Alesina2013]; posterior median = `r printnum(median(alesinaSignal))`, 95% CI [`r printnum(quantile(alesinaSignal, 0.025))` `r printnum(quantile(alesinaSignal, 0.975))`]).
+tightness[@Gelfand2011]; posterior median = `r printnum(median(gelfandSignal1))`, 95% CI [`r printnum(quantile(gelfandSignal1, 0.025))` `r printnum(quantile(gelfandSignal1, 0.975))`]) and others expressing high signal (e.g.
+female labour force participation[@Alesina2013]; posterior median = `r printnum(median(alesinaSignal1))`, 95% CI [`r printnum(quantile(alesinaSignal1, 0.025))` `r printnum(quantile(alesinaSignal1, 0.975))`]).
 Across all analyses, we found that stronger estimates of spatial autocorrelation or cultural phylogenetic signal resulted in a more pronounced reduction in the effect size when controlling for non-independence between nations (Figure \@ref(fig:plotReplications5)).

### 2022-10-20 R/functionsReplications.R

@@ -771,6 +771,74 @@ fitModelSkidmore2002 <- function(skidmoreData, linCov, control) {
         prior = priors, control = list(adapt_delta = 0.999))
 }
 
+# calculate spatial autocorrelation at 1000km distance
+getSpatialAutocorrelation1000km <- function(model, data) {
+  # get posterior
+  post <- posterior_samples(model, pars = c("lscale_gpLongitude..average.Latitude..average."))
+  # calculate maximum distance in dataset
+  data <- data %>% group_by(iso2lin) %>% summarise_if(is.numeric, mean)
+  maxDist <- max(as.matrix(distm(data[data$iso2lin %in% model$data$iso2lin, c("Longitude..average.","Latitude..average.")])))
+  # get spatial autocorrelation at 1000km distance
+  out <- exp(-(1 / (2 * post$lscale_gpLongitude..average.Latitude..average.^2)) * (1000*1000 / maxDist)^2)
+  return(out)
+}
+
+# calculate cultural phylogenetic signal
+getCulturalPhylogeneticSignal <- function(model) {
+  # get posterior
+  post <- posterior_samples(model, pars = c("sd_", "sigma"))
+  # get total variance
+  total <- post$sd_iso2lin__Intercept^2 + post$sigma^2
+  # include additional random effect in total variance if necessary
+  if ("sd_S009__Intercept" %in% names(post)) total <- total + post$sd_S009__Intercept^2
+  # calculate cultural phylogenetic signal as ICC
+  signal <- post$sd_iso2lin__Intercept^2 / total
+  return(signal)
+}
+
+# ratio of original to "control" effect size
+getRatioEffectSize <- function(modelOriginal, modelControl) {
+  # get posterior samples
+  postOriginal <- posterior_samples(modelOriginal)
+  postControl  <- posterior_samples(modelControl)
+  # calculate posterior eff size ratio
+  if (nrow(fixef(modelOriginal)) == 29) {
+    ratio <- postControl[,29] / median(postOriginal[,29])
+  } else {
+    ratio <- postControl[,2] / median(postOriginal[,2])
+  }
+  return(ratio)
+}
+
+# fit effect size ratio model
+fitEffSizeRatioModel <- function(formula, ratioList, saList, signalList, nsamps = 1000) {
+  # set seed
+  set.seed(2113)
+  # get indexes to sample iterations
[…]

### 2022-10-20 _targets.R

@@ -185,28 +185,28 @@ list(
   tar_combine(conleyModel1_simLinCov, simulationTargets$conleyModel1[37:72], command = dplyr::bind_rows(!!!.x)),
   tar_combine(conleyModel2_simGeoCov, simulationTargets$conleyModel2[1:36],  command = dplyr::bind_rows(!!!.x)),
   tar_combine(conleyModel2_simLinCov, simulationTargets$conleyModel2[37:72], command = dplyr::bind_rows(!!!.x)),
-  #tar_combine(brmsModel1_simGeoCov,   simulationTargets$brmsModel1[1:36],    command = dplyr::bind_rows(!!!.x)),
-  #tar_combine(brmsModel1_simLinCov,   simulationTargets$brmsModel1[37:72],   command = dplyr::bind_rows(!!!.x)),
-  #tar_combine(brmsModel2_simGeoCov,   simulationTargets$brmsModel2[1:36],    command = dplyr::bind_rows(!!!.x)),
-  #tar_combine(brmsModel2_simLinCov,   simulationTargets$brmsModel2[37:72],   command = dplyr::bind_rows(!!!.x)),
-  #tar_combine(brmsModel3_simGeoCov,   simulationTargets$brmsModel3[1:36],    command = dplyr::bind_rows(!!!.x)),
-  #tar_combine(brmsModel3_simLinCov,   simulationTargets$brmsModel3[37:72],   command = dplyr::bind_rows(!!!.x)),
+  tar_combine(brmsModel1_simGeoCov,   simulationTargets$brmsModel1[1:36],    command = dplyr::bind_rows(!!!.x)),
+  tar_combine(brmsModel1_simLinCov,   simulationTargets$brmsModel1[37:72],   command = dplyr::bind_rows(!!!.x)),
+  tar_combine(brmsModel2_simGeoCov,   simulationTargets$brmsModel2[1:36],    command = dplyr::bind_rows(!!!.x)),
+  tar_combine(brmsModel2_simLinCov,   simulationTargets$brmsModel2[37:72],   command = dplyr::bind_rows(!!!.x)),
+  tar_combine(brmsModel3_simGeoCov,   simulationTargets$brmsModel3[1:36],    command = dplyr::bind_rows(!!!.x)),
+  tar_combine(brmsModel3_simLinCov,   simulationTargets$brmsModel3[37:72],   command = dplyr::bind_rows(!!!.x)),
   # plot simulation results
[…]

### 2022-11-20 _targets.R
rerun with fewer targets

@@ -162,7 +162,7 @@ list(
   tar_target(brmsInitial2, setupBrms(simData_simLinCov_0.8_0.8_0$simData[[1]], simLinCov, type = "linguistic")),
   tar_target(brmsInitial3, setupBrms(simData_simLinCov_0.8_0.8_0$simData[[1]], simLinCov, type = "both")),
   # sim iterations
-  tar_target(iter, 2),
+  tar_target(iter, 100),
   # simulation
   simulationTargets,
   # combine

### 2022-11-20 run.R

@@ -1,1 +1,1 @@
-targets::tar_make_clustermq(workers = 150L)
+targets::tar_make_clustermq(workers = 5L)

### 2022-11-21 R/functionsReplications.R
create tables

@@ -853,6 +853,70 @@ fitEffSizeRatioModel <- function(formula, ratioList, saList, signalList, nsamps 
   return(out)
 }
 
+# table of deviations from original analyses
+makeTableDeviations <- function() {
+  tibble(
+    Analysis = c(
+      "Adamczyk and Pitt (2009)", "Alesina et al. (2013)", "Beck et al. (2003)", 
+      "Beck et al. (2005)", "Bockstette et al. (2002)", "Easterly and Levine (2003)", 
+      "Easterly (2007)", "Fincher et al. (2008)", "Gelfand et al. (2011)", 
+      "Inglehart and Baker (2000)", "Knack and Keefer (1997)", "Skidmore and Toya (2002)"
+      ),
+    `Original N` = c("33", ">177", "70", "45", "94", "72", "118", "68", "30", "65", "28", "89"),
+    `Reanalysis N` = c("33", "75", "69", "45", "103", "63", "98", "67", "28", "38", "28", "89"),
+    `Reason for deviation` = c(
+      "N/A",
+      paste0("Original analysis used imputed values for traditional plough use which were ",
+             "unavailable to us. For our reanalysis, we collected data from Tables A4 and A9 ",
+             "and ran our regression on the 75 nations with available data for female labour ",
+             "force participation in 2000 and traditional plough use."),
+      paste0("We removed one nation (Zaire) as it was not present in our linguistic distance matrix."),
+      "N/A",
+      paste0("After manually linking state history data from Appendix A with world development ",
+             "indicators, we had complete data for 103 nations."),
+      paste0("After manually linking log GDP per capita 1995 data from Appendix Table A2 in Acemoglu ",
+             "et al. (2001) with an institutions index self-constructed from world governance indicators, ",
+             "we only had complete data for 63 nations."),
+      paste0("After manually linking log wheat-sugar ratio data from Appendix A with longitudinal Gini ",
+             "data, we only had complete data for 98 nations."),
+      "We combined England and Northern Ireland into a single nation (United Kingdom) for the reanalysis.",
+      paste0("We combined East and West Germany into a single nation (Germany) and removed a high ",
+             "leverage point with a substantially higher disaster rate than other nations (Venezuela)."),
[…]

### 2022-11-21 R/functionsReview.R

@@ -305,3 +305,13 @@ plotReviewArticle <- function(review, yearArticle, ifArticle) {
   ggsave(out, filename = "figures/reviewArticle.pdf", height = 6, width = 6.5)
   return(out)
 }
+
+# make table with all articles in review
+makeTableArticlesReview <- function(review) {
+  review %>%
+    select(Review, Citation, CitationsPerYear) %>%
+    distinct() %>%
+    rename(Reference = Citation, `Citations per year` = CitationsPerYear) %>%
+    mutate(Review = c("Cultural values", rep("", times = 49),
+                      "Economic development", rep("", times = 49)))
+}

### 2022-11-21 R/functionsSignal.R

@@ -319,3 +319,49 @@ plotGeoLinSignal <- function(geoHDI, geoGDP, geoGrow, geoGini,
   ggsave(out, filename = "figures/signal.pdf", width = 7, height = 5)
   return(out)
 }
+
+# make table of signal values
+makeTableGeoLinSignal <- function(geoHDI, geoGDP, geoGrow, geoGini,
+                                  geoTrad, geoSurv, geoTight, geoInd,
+                                  linHDI, linGDP, linGrow, linGini,
+                                  linTrad, linSurv, linTight, linInd) {
+  # tidy function
+  tidySignal <- function(signal) {
+    paste0(
+      format(round(signal$hypothesis$Estimate, 2), nsmall = 2), ", 95% CI [",
+      format(round(signal$hypothesis$CI.Lower, 2), nsmall = 2), ", ",
+      format(round(signal$hypothesis$CI.Upper, 2), nsmall = 2), "], BF ",
+      ifelse(
+        signal$hypothesis$Evid.Ratio <= 0, "> 100",
+        ifelse(
+          1 / signal$hypothesis$Evid.Ratio > 100, "> 100",
+          paste0("= ", format(round(1 / signal$hypothesis$Evid.Ratio, 2), nsmall = 2)))
+        )
+      )
+  }
+  # make table
+  tibble(
+    Outcome = c("Human Development Index", "GDP per capita", "GDP per capita growth", "Gini index",
+                "Traditional values", "Survival values", "Tightness", "Individualism"),
+    `Geographic signal` = c(
+      tidySignal(geoHDI),
+      tidySignal(geoGDP),
+      tidySignal(geoGrow),
+      tidySignal(geoGini),
+      tidySignal(geoTrad),
+      tidySignal(geoSurv),
+      tidySignal(geoTight),
+      tidySignal(geoInd)
+    ),
+    `Cultural phylogenetic signal` = c(
+      tidySignal(linHDI),
+      tidySignal(linGDP),
+      tidySignal(linGrow),
+      tidySignal(linGini),
+      tidySignal(linTrad),
+      tidySignal(linSurv),
+      tidySignal(linTight),
+      tidySignal(linInd)
+    )
+  )
+}

### 2022-11-21 R/functionsSimulation.R

@@ -241,6 +241,38 @@ fitBrmsModel <- function(brmsInitial, data) {
     select(-c(simData, model))
 }
 
+# plot simulation results across all autocorrelation levels
+makeTableSimAll <- function(olsModel1, olsModel2, olsModel3, olsModel4, olsModel5,
+                            olsModel6, conleyModel1, conleyModel2, brmsModel1,
+                            brmsModel2, brmsModel3) {
+  # list of outputs
+  l <- list(olsModel1, olsModel2, olsModel3, olsModel4, olsModel5,
+            olsModel6, conleyModel1, conleyModel2, brmsModel1,
+            brmsModel2, brmsModel3)
+  # get percentage
+  getPerc <- function(x, Rho, Lambda) {
+    x %>% 
+      filter(rho == Rho & lambda == Lambda & r == 0) %>% 
+      pull(sig) %>% 
+      sum()
+  }
+  tibble(
+    Model = c("No control", "Latitude", "Longitude", "Continent",
+              "Language family", "Mean 2000km radius", "Conley SEs spatial",
+              "Conley SEs genetic", "Bayesian spatial", "Bayesian linguistic",
+              "Bayesian spatial & linguistic"),
+    `Rho = 0.2\nLambda = 0.2` = unlist(lapply(l, getPerc, Rho = 0.2, Lambda = 0.2)),
+    `Rho = 0.2\nLambda = 0.5` = unlist(lapply(l, getPerc, Rho = 0.2, Lambda = 0.5)),
+    `Rho = 0.2\nLambda = 0.8` = unlist(lapply(l, getPerc, Rho = 0.2, Lambda = 0.8)),
+    `Rho = 0.5\nLambda = 0.2` = unlist(lapply(l, getPerc, Rho = 0.5, Lambda = 0.2)),
+    `Rho = 0.5\nLambda = 0.5` = unlist(lapply(l, getPerc, Rho = 0.5, Lambda = 0.5)),
+    `Rho = 0.5\nLambda = 0.8` = unlist(lapply(l, getPerc, Rho = 0.5, Lambda = 0.8)),
+    `Rho = 0.8\nLambda = 0.2` = unlist(lapply(l, getPerc, Rho = 0.8, Lambda = 0.2)),
+    `Rho = 0.8\nLambda = 0.5` = unlist(lapply(l, getPerc, Rho = 0.8, Lambda = 0.5)),
[…]

### 2022-11-21 _targets.R

@@ -111,6 +111,11 @@ list(
                                           geoTrad, geoSurv, geoTight, geoInd,
                                           linHDI, linGDP, linGrow, linGini,
                                           linTrad, linSurv, linTight, linInd)),
+  # table
+  tar_target(tableSignal, makeTableGeoLinSignal(geoHDI, geoGDP, geoGrow, geoGini,
+                                                geoTrad, geoSurv, geoTight, geoInd,
+                                                linHDI, linGDP, linGrow, linGini,
+                                                linTrad, linSurv, linTight, linInd)),
   
   #### Review ####
   
@@ -141,6 +146,8 @@ list(
   tar_target(plotReview1, plotReviewArticle(review, yearArticle, ifArticle)), 
   tar_target(plotReview2, plotReviewAnalysis(review, yearAnalysis, ifAnalysis, 
                                              postRM1, postRM2, postRM3, postRM4, postRM5)),
+  # table of articles in review
+  tar_target(tableArticlesReview, makeTableArticlesReview(review)),
   
   #### Simulation ####
   
@@ -188,6 +195,13 @@ list(
   tar_combine(brmsModel2_simLinCov,   simulationTargets$brmsModel2[37:72],   command = dplyr::bind_rows(!!!.x)),
   tar_combine(brmsModel3_simGeoCov,   simulationTargets$brmsModel3[1:36],    command = dplyr::bind_rows(!!!.x)),
   tar_combine(brmsModel3_simLinCov,   simulationTargets$brmsModel3[37:72],   command = dplyr::bind_rows(!!!.x)),
+  # tables of simulation results
+  tar_target(tableSim1, makeTableSimAll(olsModel1_simGeoCov, olsModel2_simGeoCov, olsModel3_simGeoCov, olsModel4_simGeoCov, 
+                                        olsModel5_simGeoCov, olsModel6_simGeoCov, conleyModel1_simGeoCov, conleyModel2_simGeoCov,
+                                        brmsModel1_simGeoCov, brmsModel2_simGeoCov, brmsModel3_simGeoCov)),
+  tar_target(tableSim2, makeTableSimAll(olsModel1_simLinCov, olsModel2_simLinCov, olsModel3_simLinCov, olsModel4_simLinCov, 
+                                        olsModel5_simLinCov, olsModel6_simLinCov, conleyModel1_simLinCov, conleyModel2_simLinCov,
[…]

### 2022-11-21 vignette.Rmd
update vignette

@@
 options(width = 300)
+oecd <- c("AT","AU","BE","CA","CH","CL","CZ","DE","DK",
+          "EE","ES","FI","FR","GB","GR","HU","IE","IL",
+          "IS","IT","JP","KR","LT","LU","LV","MX","NL",
+          "NO","NZ","PL","PT","SE","SI","SK","TR","US")
 ```
@@
 This dataset contains a standardised predictor variable (X) and a standardised outcome variable (Y), both of which are strongly spatially autocorrelated.
-
-```{r echo=FALSE}
-# load data from targets cache
-spatialSimData <- tar_read(simData_simGeoCov_0.8_0.8_1728b220)
-# only OECD countries for example, country iso codes
-oecd <- c("AT","AU","BE","CA","CH","CL","CZ","DE","DK",
-          "EE","ES","FI","FR","GB","GR","HU","IE","IL",
-          "IS","IT","JP","KR","LT","LU","LV","MX","NL",
-          "NO","NZ","PL","PT","SE","SI","SK","TR","US")
-spatialSimData <- filter(spatialSimData, country %in% oecd)
-# write data to file
-write.csv(file = "data/vignette/spatialSimData.csv", 
-          spatialSimData, row.names = FALSE)
-```
 
@@
 Both of these variables are strongly culturally autocorrelated.
-
-```{r echo=FALSE}
-# load data from targets cache
-culturalSimData <- tar_read(simData_simLinCov_0.8_0.8_0fb442ad)
-# only OECD countries
-culturalSimData <- filter(culturalSimData, country %in% oecd)
-# write data to file
-write.csv(file = "data/vignette/culturalSimData.csv", 
-          culturalSimData, row.names = FALSE)
-```
 

### 2022-11-21 _targets.R

@@ -580,7 +580,7 @@ list(
   
   #### Vignette ####
   
-  #tar_render(vignette, "vignette.Rmd"),
+  tar_render(vignette, "vignette.Rmd"),
   
   #### Session Info ####
   

### 2022-11-28 appendix.Rmd
fix broken link in appendix
fix mean 2000km radius method in sim; start updating manuscript

@@
----
-output:
-  pdf_document: default
-  html_document: default
----
 \renewcommand{\appendixname}{\bf{Supplementary Material}}
@@
 
-## Supplementary Results
-
-### Geographic and cultural phylogenetic signal estimates for Human Development Index and Inglehart's value dimensions
-
-```{r load signal}
-tar_load(c(hdiGeo, hdiLin, tradGeo, tradLin, survGeo, survLin))
-```
-
-Controlling for shared cultural ancestry, the proportion of national-level variance explained by spatial proximity was `r printnum(hdiGeo$hypothesis$Estimate, digits = 2)` for the Human Development Index (95% credible interval [`r printnum(hdiGeo$hypothesis$CI.Lower, digits = 2)` `r printnum(hdiGeo$hypothesis$CI.Upper, digits = 2)`]; $\text{BF}_{\neq0}$ > 100), `r printnum(tradGeo$hypothesis$Estimate, digits = 2)` for traditional values (95% CI [`r printnum(tradGeo$hypothesis$CI.Lower, digits = 2)` `r printnum(tradGeo$hypothesis$CI.Upper, digits = 2)`]; $\text{BF}_{\neq0}$ > 100), and `r printnum(survGeo$hypothesis$Estimate, digits = 2)` for survival values (95% CI [`r printnum(survGeo$hypothesis$CI.Lower, digits = 2)` `r printnum(survGeo$hypothesis$CI.Upper, digits = 2)`]; $\text{BF}_{\neq0}$ = `r printnum(1 / survGeo$hypothesis$Evid.Ratio, digits = 2)`).
-Controlling for spatial proximity, the proportion of variance explained by shared cultural ancestry was `r printnum(hdiLin$hypothesis$Estimate, digits = 2)` for the Human Development Index (95% CI [`r printnum(hdiLin$hypothesis$CI.Lower, digits = 2)` `r printnum(hdiLin$hypothesis$CI.Upper, digits = 2)`]; $\text{BF}_{\neq0}$ > 100), `r printnum(tradLin$hypothesis$Estimate, digits = 2)` for traditional values (95% CI [`r printnum(tradLin$hypothesis$CI.Lower, digits = 2)` `r printnum(tradLin$hypothesis$CI.Upper, digits = 2)`]; $\text{BF}_{\neq0}$ > 100), and `r printnum(survLin$hypothesis$Estimate, digits = 2)` for survival values (95% CI [`r printnum(survLin$hypothesis$CI.Lower, digits = 2)` `r printnum(survLin$hypothesis$CI.Upper, digits = 2)`]; $\text{BF}_{\neq0}$ > 100).
-
-\newpage
-
 ## Supplementary Figures
@@
 
-```{r dag, fig.width=6, fig.height=3.5, fig.cap="(ref:dagCaption)"}
+```{r dag, fig.width=6, fig.height=3.5, fig.cap="(ref:dagCaption)", fig.pos='H', out.width="\\textwidth"}
 tar_read(dag)
+```
+
+\newpage
+
+(ref:plotReview2Caption) _Analysis-level results from systematic review of 100 highly-cited cross-national studies of economic development (red) and cultural values (blue)._ (a) Proportion of analyses accounting for non-independence, overall and split by common methods of controlling for non-independence.
+(b) The association between journal impact factor and the probability that an analysis accounts for non-independence.
+(c) Estimated trend over time for the probability that an analysis accounts for non-independence.
+Point ranges represent estimated proportions and 95% credible intervals.
+Lines and shaded areas are posterior median regression lines and 50% credible intervals from Bayesian multilevel models.
+Region FEs = region fixed effects.
+
+```{r plotReview2, fig.width=6.5, fig.height=6, fig.cap="(ref:plotReview2Caption)", fig.pos='H', out.width="\\textwidth"}
+tar_read(plotReview2)
 ```
@@
 
-```{r plotSim1, fig.width=9.5, fig.height=5.5, fig.cap="(ref:plotSim1Caption)"}
+```{r plotSim1, fig.width=9.5, fig.height=5.5, fig.cap="(ref:plotSim1Caption)", fig.pos='H', out.width="\\textwidth"}
 tar_read(plotSim1)
@@
 
-```{r plotSim2, fig.width=9.5, fig.height=5.5, fig.cap="(ref:plotSim2Caption)"}
+```{r plotSim2, fig.width=9.5, fig.height=5.5, fig.cap="(ref:plotSim2Caption)", fig.pos='H', out.width="\\textwidth"}
 tar_read(plotSim2)
@@
 
-```{r plotReplications2, fig.width=8, fig.height=8, fig.cap="(ref:plotReplications2Caption)"}
+```{r plotReplications2, fig.width=8, fig.height=8, fig.cap="(ref:plotReplications2Caption)", fig.pos='H', out.width="\\textwidth"}
 tar_read(plotReplications2)
@@
 
-```{r plotReplications3, fig.width=8.2, fig.height=6, fig.cap="(ref:plotReplications3Caption)"}
+```{r plotReplications3, fig.width=8.2, fig.height=6, fig.cap="(ref:plotReplications3Caption)", fig.pos='H', out.width="\\textwidth"}
 tar_read(plotReplications3)
@@
 
-```{r plotReplications4, fig.width=5, fig.height=5, fig.cap="(ref:plotReplications4Caption)"}
+```{r plotReplications4, fig.width=5, fig.height=5, fig.cap="(ref:plotReplications4Caption)", fig.pos='H', out.width="\\textwidth"}
 tar_read(plotReplications4)
+```
+
+\newpage
+
+## Supplementary Tables
+
+(ref:tableSignalCaption) Geographic and cultural phylogenetic signal results for economic development and cultural values variables.
+Signal estimates reflect the proportion of national-level variance explained by geographic and linguistic covariance matrices.
+Bayes Factors (BF) reflect support for the hypothesis that the signal estimate differs from zero.
+HDI = Human Development Index; GDPpc = gross domestic product per capita.
+
+```{r tableSignal, echo=F, warning=F, message=F}
+apa_table(
+  tar_read(tableSignal),
+  booktabs = TRUE,
+  placement = "H",
+  font_size = "small",
+  caption = "(ref:tableSignalCaption)"
+)
+```
+
+\newpage
+
+(ref:tableArticlesReviewCaption) List of 100 papers included in literature review, sorted by annual rate of citations since publication.
+
+```{r tableArticlesReview, echo=F, warning=F, message=F}
+apa_table(
+  tar_read(tableArticlesReview),
+  booktabs = TRUE,
+  placement = "H",
+  font_size = "small",
+  longtable = TRUE,
+  align = c("m{2cm}", "m{11cm}", "m{2cm}"),
+  caption = "(ref:tableArticlesReviewCaption)"
+)
 ```
@@
 _Glottolog 3.0._ (Max Planck Institute for the Science of Human History, 2017).
-doi:[10.5281/zenodo.4061162](https://doi.org/10.5281/zenodo.4061162)
+doi: [10.5281/zenodo.4061162](https://doi.org/10.5281/zenodo.4061162)
 

### 2022-11-28 manuscript.Rmd
fix mean 2000km radius method in sim; start updating manuscript

@@
 In a review of the 100 highest-cited cross-national studies of economic development and values, we find that controls for non-independence are rare.
-When studies do include controls for non-independence, our simulations suggest that commonly used methods continue to produce false positives.
-In reanalyses of twelve cross-national relationships, we show that half are no longer significant after controlling for non-independence using global proximity matrices.
+When studies do control for non-independence, our simulations suggest that most commonly used methods are insufficient for reducing false positives in non-independent data.
+In reanalyses of twelve previous cross-national correlations, half of the effect sizes show no association after controlling for non-independence using global proximity matrices.
 We urge social scientists to sufficiently control for non-independence in cross-national research.
@@
 keywords          : "spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses"
-wordcount         : "5459 words"
+wordcount         : "xxxx words"
 bibliography      : ["r-references.bib"]
-floatsintext      : no
+floatsintext      : yes
 figurelist        : no
@@
 Given the importance of nations for structuring human behaviour, there is justifiably huge interest in statistical analyses that attempt to predict variation in national outcomes of all kinds.
-At the time of writing, a search in Web of Science for the terms "cross-national", "cross-cultural", or "cross-country" in titles and abstracts returned 80,000 unique hits.
+At the time of writing, a search in Web of Science for the term "cross-national" in titles or abstracts returned over 13,000 unique hits.
 The standard practice for cross-national analyses is to conduct bivariate correlations or multiple regressions with individual data points representing different nations.
@@
 Shared cultural ancestry can result in a form of pseudoreplication, whereby multiple instances of the same trait across nations are merely duplicates of the ancestral original.
-For example, despite being separated by over 1,500 kilometres of ocean, the related island nations Tonga and Tuvalu share similar languages and customs due to cultural inheritance from a common Polynesian population dating back more than 1,000 years.
+For example, the related island nations Tonga and Tuvalu share similar languages and customs due to cultural inheritance from a common Polynesian population dating back more than 1,000 years.
 More recently, shared ancestry explains a myriad of cultural similarities between colonial settlements and their colonisers (e.g.
 Argentina and Spain).
+Importantly, these deep cultural connections between nations often span large geographic distances around the world.
+Tonga and Tuvalu share cultural traits despite being separated by over 1,500 kilometres of ocean, and Argentina and Spain remain culturally similar today despite their location on two separate continents.
+Shared cultural ancestry must therefore be considered independently of spatial proximity in the study of nations.
 
@@
 
-Non-independence between data points is widely acknowledged in fields that routinely deal with spatially or culturally structured data.
+Non-independence is widely acknowledged in fields that routinely deal with spatially or culturally structured data.
 In ecology and sociology, studies with regional-level data use a variety of autoregressive models designed to account for spatial patternings[@Lichstein2002;@Loftin1983].
 In anthropology, researchers have recognised cultural non-independence as an important statistical pitfall for over 200 years, with issues of cultural pseudoreplication being identified in early comparative studies of marriage practices across societies[@Tylor1889].
-More recently, the Standard Cross-Cultural Sample of 186 cultures was compiled to minimise the confounding effects of non-independence in comparisons of traditional societies[@Murdock1969], though spatial and cultural dependencies are difficult to remove entirely[@Dow2008;@Eff2004].
-Anthropologists also borrow phylogenetic comparative methods from evolutionary biology, such as phylogenetic least squares regression[@Symonds2014], when comparing traditional societies, treating culturally related societies in the same way as biologists treat genetically related species (e.g.
+In the twentieth century, anthropologists began to emphasise that human societies do not develop independently, but rather exist in a globally interconnected "world system" linked by shared history and cultural ancestry [@Ferguson1997].
+Researchers compiled the Standard Cross-Cultural Sample of 186 cultures to minimise the confounding effects of this non-independence in comparisons of human societies[@Murdock1969], though spatial and cultural dependencies are difficult to remove entirely[@Dow2008;@Eff2004].
+Today, anthropologists borrow phylogenetic comparative methods from evolutionary biology, such as phylogenetic least squares regression[@Symonds2014], when comparing societies, treating culturally related societies in the same way as biologists treat genetically related species (e.g.
 [@Watts2018;@Atkinson2016]).
@@
 At the national level, recent reanalyses have revealed that several cross-national relationships reported in economics and psychology do not hold when controlling for non-independence between nations.
-One study replicated 25 analyses of "persistence" in economics, in which modern national outcomes are regressed against historical characteristics of those nations, and found that over half of the relationships were attenuated when controlling for spatial non-independence[@Kelly2020].
+One working paper replicated 25 analyses of "persistence" in economics, in which modern national outcomes are regressed against historical characteristics of those nations, and found that over half of the relationships were attenuated when controlling for spatial non-independence[@Kelly2020].
 Another replication study found that many of the widely publicised relationships between national-level pathogen prevalence and political institutions and attitudes fail to hold when controlling for various kinds of non-independence[@Bromham2018].
@@
 
-In order to motivate our research question, it is important to first show that economic development and cultural values are indeed spatially and culturally non-independent around the world.
-To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for the Human Development Index[@hdi] and two primary dimensions of cultural values from the World Values Survey, traditional vs.
+In order to motivate our research question, it is important to first quantify the degree of spatial and cultural non-independence for economic development and cultural values around the world.
+If these variables are only weakly non-independent, then the issue might be safe to ignore.
+However, if they are strongly non-independent, then there is a possibility that non-independence could be confounding cross-national inferences.
+
+To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for a range of economic development and cultural values variables that are widely used in the literature.
+For economic development, we focused on the Human Development Index[@hdi], gross domestic product per capita, gross domestic product per capita growth, and the Gini index of income inequality.
+For cultural values, we focused on two primary dimensions of cultural values from the World Values Survey, traditional vs.
 secular values and survival vs.
-self-expression values[@Inglehart2000].
-For all three of these variables, we found that a substantial proportion of national-level variation was explained by spatial proximity and shared cultural ancestry between nations (Figure \@ref(fig:plotSignal)).
-In our models, shared cultural ancestry explained over half of the national-level variation in economic development and cultural values, with spatial proximity explaining most of the remaining national-level variance.
-Bayes Factors indicated strong evidence that these estimates of geographic and cultural phylogenetic signal differed from zero, aside from the geographic signal for survival values, for which the evidence was equivocal (see Supplementary Results).
+self-expression values[@Inglehart2000], as well as cultural tightness[@Gelfand2011] and individualism[@Hofstede2001].
+
+For all of these variables, we found that a substantial proportion of national-level variation was explained by spatial proximity and/or shared cultural ancestry between nations (Figure \@ref(fig:plotSignal); Supplementary Table \@ref(tab:tableSignal)).
+Signal estimates were often markedly strong, with spatial proximity and shared cultural ancestry frequently explaining over half of the national-level variation.
+For spatial proximity, Bayes Factors indicated strong evidence that the geographic signal estimates differed from zero for all economic development variables and traditional values.
+However, the evidence was only equivocal for survival values and individualism, and strong evidence was found that the geographic signal estimate for tightness was equal to zero.
+For shared cultural ancestry, Bayes Factors indicated strong evidence that the cultural phylogenetic signal estimates differed from zero for all economic development and cultural values variables except for gross domestic product per capita growth, for which the evidence was equivocal.
 These findings emphasise the need to account for spatial and cultural phylogenetic non-independence in cross-national analyses of economic development and cultural values.
 
-(ref:plotSignalCaption) _Posterior estimates of geographic and cultural phylogenetic signal for the Human Development Index and Inglehart's traditional and survival values._ Geographic and cultural phylogenetic signal are operationalised as the proportion of national-level variance explained by geographic and linguistic proximity matrices.
-Grey ridges are full posterior distributions, points are posterior median values, and lines are 50% and 95% credible intervals.
+(ref:plotSignalCaption) _Posterior estimates of geographic and cultural phylogenetic signal for a range of economic development and cultural values variables._ Geographic and cultural phylogenetic signal are operationalised as the proportion of national-level variance explained by geographic and linguistic proximity matrices.
+Grey ridges are full posterior distributions, points are posterior median values, and black lines are 50% and 95% credible intervals.
+HDI = Human Development Index; GDP = gross domestic product.
 
-```{r plotSignal, fig.width=7, fig.height=3.5, fig.cap="(ref:plotSignalCaption)"}
+```{r plotSignal, fig.width=7, fig.height=5, fig.cap="(ref:plotSignalCaption)"}
 tar_read(plotSignal)
@@
 We removed articles that did not report original research, were not relevant to economic development or cultural values, or did not report at least one cross-national analysis.
-We then retained the 100 articles (50 for economic development, 50 for cultural values) with the highest annual rate of citations.
+We then retained the 100 articles (50 for economic development, 50 for cultural values) with the highest annual rate of citations (Supplementary Table \@ref(tab:tableArticlesReview)).
 For each of these highly-cited articles, we exhaustively recorded every cross-national analysis reported in the main text (n = `r printnum(as.numeric(nrow(review)), digits = 0)`), identifying in each case whether or not the analysis attempted to control for spatial, cultural, or any other form of non-independence between nations (see Methods for detailed search criteria and coding decisions).
 
-```{r calculate prop 1}
+```{r calculate prop}
+# article-level
 prop <-
@@
   unnest(c(lower, upper))
+# analysis-level
+tar_load(postRM1)
+propEcon <- inv_logit_scaled(postRM1[,,"b_ReviewEconomicdevelopment"])
+propVal  <- inv_logit_scaled(postRM1[,,"b_ReviewValues"])
 ```
 
-The results of our systematic review show that most published articles and cross-national analyses make no attempt to account for statistical non-independence.
-Figure \@ref(fig:plotReview)a plots the proportion of articles that contain at least one cross-national analysis accounting for non-independence.
+The results of our systematic review show that most published articles containing cross-national analyses make no attempt to account for statistical non-independence.
+Figure \@ref(fig:plotReview1)a plots the proportion of articles that contain at least one cross-national analysis accounting for non-independence.
 We find that `r printnum(prop$prop[1]*100, digits = 0)`% of economic development articles contain at least one attempt to control for non-independence (95% bootstrap confidence interval [`r printnum(prop$lower[1])` `r printnum(prop$upper[1])`]), while this proportion decreases to only `r printnum(prop$prop[2]*100, digits = 0)`% for cultural values articles (95% bCI [`r printnum(prop$lower[2])` `r printnum(prop$upper[2])`]).
@@
 colony status).
+These proportions are even lower when focusing on the full sample of `r printnum(as.numeric(nrow(review)), digits = 0)` analyses: only `r printnum(median(propEcon)*100, digits = 0)`% (95% credible interval [`r printnum(quantile(propEcon, 0.025))` `r printnum(quantile(propEcon, 0.975))`]) of individual economic development analyses and `r printnum(median(propVal)*100, digits = 0)`% (95% CI [`r printnum(quantile(propVal, 0.025))` `r printnum(quantile(propVal, 0.975))`]) of individual cultural values analyses are estimated to control for non-independence (Supplementary Figure \@ref(fig:plotReview2)).
 
-(ref:plotReviewCaption) _Results from systematic review of 100 highly-cited cross-national studies of economic development (red) and cultural values (blue)._ (a) Proportion of articles containing at least one analysis accounting for non-independence, overall and split by common methods of controlling for non-independence.
+(ref:plotReview1Caption) _Results from systematic review of 100 highly-cited cross-national studies of economic development (red) and cultural values (blue)._ (a) Proportion of articles containing at least one analysis accounting for non-independence, overall and split by common methods of controlling for non-independence.
+(b) The association between journal impact factor and the probability that an article contains at least one analysis accounting for non-independence.
+(c) Estimated trend over time for the probability that an article contains at least one analysis accounting for non-independence.
 Point ranges represent proportions and 95% bootstrap confidence intervals.
-(b) Adjusted proportion of individual analyses accounting for non-independence, overall and split by method.
-Point ranges are posterior medians and 95% credible intervals from Bayesian multilevel logistic regressions.
-(c) Histogram of publication years for studies of economic development and cultural values.
-(d) Estimated trend over time for the probability of controlling for non-independence.
-Lines and shaded areas are posterior median regression lines and 50% credible intervals from a Bayesian multilevel spline model.
+Lines and shaded areas are posterior median regression lines and 50% credible intervals from Bayesian multilevel models.
 Region FEs = region fixed effects.
 
-```{r plotReview, fig.width=8, fig.height=5.5, fig.cap="(ref:plotReviewCaption)"}
-tar_read(plotReview)
+```{r plotReview1, fig.width=6.5, fig.height=6, fig.cap="(ref:plotReview1Caption)"}
+tar_read(plotReview1)
 ```
 
-```{r calculate prop 2}
-tar_load(postRM1)
-propEcon <- inv_logit_scaled(postRM1[,,"b_ReviewEconomicdevelopment"])
-propVal  <- inv_logit_scaled(postRM1[,,"b_ReviewValues"])
+```{r load article models, echo=F}
+tar_load(yearArticle)
+tar_load(ifArticle)
 ```
 
-Focusing on the full sample of `r printnum(as.numeric(nrow(review)), digits = 0)` analyses, we find that the proportion of individual cross-national analyses accounting for non-independence is even lower (Figure \@ref(fig:plotReview)b).
-Across `r printnum(as.numeric(sum(review$Review == "Economic development")), digits = 0)` cross-national analyses from studies of economic development, only `r printnum(median(propEcon)*100, digits = 0)`% are estimated to control for non-independence (95% credible interval [`r printnum(quantile(propEcon, 0.025))` `r printnum(quantile(propEcon, 0.975))`]).
-Similarly, across `r printnum(as.numeric(sum(review$Review == "Values")), digits = 0)` cross-national analyses from studies of cultural values, only `r printnum(median(propVal)*100, digits = 0)`% are estimated to control for non-independence (95% CI [`r printnum(quantile(propVal, 0.025))` `r printnum(quantile(propVal, 0.975))`]).
+While our review contains a range of articles from low- and high-impact journals, an anonymous reviewer suggested that our estimates could be biased downwards by analyses published in lower impact outlets with more relaxed standards for issues like non-independence.
+It is also possible that, since our systematic review goes back as far as 1993, our estimates are being biased downwards by earlier studies, and that controls for non-independence have increased over time with methodological advancements and greater awareness of the issue.
+To test these possible explanations for our low estimates, we fitted Bayesian logistic regression models to the data from the review, including log journal impact factor and publication year as separate predictors.
+Interestingly, we found that, contrary to the explanation above, studies from higher impact journals were _less_ likely to include at least one control for non-independence than studies from lower impact journals, both for studies of economic development (b = `r printnum(fixef(ifArticle)["ReviewEconomicdevelopment:logIF","Estimate"])`, 95% CI [`r printnum(fixef(ifArticle)["ReviewEconomicdevelopment:logIF","Q2.5"])` `r printnum(fixef(ifArticle)["ReviewEconomicdevelopment:logIF","Q97.5"])`]) and for studies of cultural values (b = `r printnum(fixef(ifArticle)["ReviewValues:logIF","Estimate"])`, 95% CI [`r printnum(fixef(ifArticle)["ReviewValues:logIF","Q2.5"])` `r printnum(fixef(ifArticle)["ReviewValues:logIF","Q97.5"])`]; Figure \@ref(fig:plotReview1)b).
+Moreover, splines revealed no relationship between publication year and the probability of including at least one control for non-independence, both for studies of economic development (b = `r printnum(fixef(yearArticle)["sYear.std:ReviewEconomicdevelopment_1","Estimate"])`, 95% CI [`r printnum(fixef(yearArticle)["sYear.std:ReviewEconomicdevelopment_1","Q2.5"])` `r printnum(fixef(yearArticle)["sYear.std:ReviewEconomicdevelopment_1","Q97.5"])`]) and for studies of cultural values (b = `r printnum(fixef(yearArticle)["sYear.std:ReviewValues_1","Estimate"])`, 95% CI [`r printnum(fixef(yearArticle)["sYear.std:ReviewValues_1","Q2.5"])` `r printnum(fixef(yearArticle)["sYear.std:ReviewValues_1","Q97.5"])`]; Figure \@ref(fig:plotReview1)c).
 
-Since our systematic review goes back as far as 1993 (Figure \@ref(fig:plotReview)c), it is possible that our estimates are being biased by earlier studies, and that controls for non-independence have increased over time with methodological advancements and greater awareness of the issue.
-To test this possibility, we fitted a time trend to the full sample of analyses with a Bayesian multilevel spline model.
-However, we found that, for both studies of economic development and cultural values, the estimated probability of controlling for non-independence has remained low since 1993 (Figure \@ref(fig:plotReview)d).
-
-## Common methods of controlling for non-independence produce inflated false positive rates
+## Common methods of controlling for non-independence are insufficient for reducing false positive rates in non-independent data
 
@@
 We then fitted naive regressions without controls to these datasets, as well as regression models with controls for latitude, longitude, and continent fixed effects.
-Despite not being identified in our systematic review, we also included fixed effects for the language families of the majority-spoken languages in each country, as this control is often used to account for cultural phylogenetic non-independence (e.g.
-[@Currie2009]).
-In addition, we included Conley standard errors, a widely used standard error correction that purportedly accounts for spatial non-independence[@Conley1999;@Conley2010].
-Finally, we included Bayesian multilevel regressions that explicitly model spatial and/or cultural phylogenetic non-independence by allowing nation random intercepts to covary according to global geographic and/or linguistic proximity matrices (see Supplementary Methods).
-To model spatial non-independence we included a Gaussian process[@McElreath2020;@Neal1998] over latitude and longitude values, and to model cultural phylogenetic non-independence we assumed that nation random intercepts were correlated in proportion to their linguistic proximity[@deVillemereuil2014].
-Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero (i.e.
-falsely infering a relationship when none is present).
+Despite not being identified in our systematic review, we also included other methods that are often used in the literature to account for non-independence.
+Additional spatial controls included the mean of the predictor variable within a surrounding 2000km radius (e.g.
+[@Schulz2022]) and Conley standard errors[@Conley1999;@Conley2010] based on geographic distances between nations (e.g.
+[@Schulz2019;@Schulz2022]).
+Additional cultural controls included fixed effects for the language families of the majority-spoken languages in each nation (e.g.
+[@Currie2009]) and Conley standard errors based on genetic distances between nations (e.g.
+[@Schulz2019;@Schulz2022]).
+These frequentist approaches attempt to account for non-independence by holding geographic location constant (latitude, longitude), discarding between-region variation and exploiting only local variation (continent fixed effects, mean of surrounding 2000km), or correcting standard errors for autocorrelation post-hoc while leaving model coefficients unchanged (Conley standard errors).
+
+Beyond frequentist approaches, we also fitted Bayesian multilevel regressions that explicitly model spatial and/or cultural phylogenetic non-independence by allowing nations to covary according to geographic and/or linguistic proximity matrices.
+Geographic proximity between nations is calculated from inverse distances between longitude and latitude coordinates.
+Linguistic proximity between nations is calculated from a global phylogenetic tree that represents hierarchical relationships of genealogical descent for all languages in the world.
+For each pair of nations, we calculate inverse phylogenetic distances (i.e.
+number of branches separating two taxa) between all languages spoken in that nation pair and produce an average "linguistic proximity" score weighted by the percentages of speakers within those nations.
+To include the resulting geographic and linguistic proximity matrices in our models, we included a Gaussian process[@McElreath2020;@Neal1998] over latitude and longitude values and/or assumed that nation random intercepts were correlated in proportion to their linguistic proximity[@deVillemereuil2014].
+These approaches attempt to account for non-independence by modelling the covariance between nations that is induced by their geographic or linguistic connections.
 
@@
 Figures \@ref(fig:plotSim3) and \@ref(fig:plotSim4) plot the estimated false positive rates from our simulation study, split by different methods and different degrees of spatial or cultural phylogenetic autocorrelation (see Supplementary Figures \@ref(fig:plotSim1) and \@ref(fig:plotSim2) for full distributions of effect sizes under strong autocorrelation).
-For reference, "weak" autocorrelation in our simulation is comparable to the geographic signal for survival values in Figure \@ref(fig:plotSignal), while "moderate" and "strong" levels of autocorrelation are comparable to the cultural phylogenetic signal for traditional and survival values, respectively.
+Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero (i.e.
+falsely infering a relationship when none is present).
+For reference, "weak" autocorrelation in our simulation is comparable to the geographic signal for survival values in Figure \@ref(fig:plotSignal) (i.e.
+20% of the national-level variance is explained by non-independence), while "moderate" and "strong" levels of autocorrelation are comparable to the cultural phylogenetic signal for traditional and survival values, respectively (i.e.
+50% and 80% of the national-level variance is explained by non-independence).
 
@@
 (2011) is a partial correlation controlling for log gross national income and Adamczyk and Pitt (2009) is a multilevel model including several covariates.
-Point ranges represent posterior medians and 95% credible intervals.
+Points and black lines represent posterior medians and 95% credible intervals.
 GDP = gross domestic product.
@@
 former colony, legal origin fixed effects, linguistic history, cultural influence); and (4) any other control variables, tests, or approaches that were deemed as attempts to control for non-independence (e.g.
-eigenvector filtering[@Tiefelsdorf2007], controls for trade-weightings between countries, cross-sectional dependence tests[@DeHoyos2006], separate analyses for subsets of countries).
+eigenvector filtering[@Tiefelsdorf2007], controls for trade-weightings between nations, cross-sectional dependence tests[@DeHoyos2006], separate analyses for subsets of nations).
 These were coded by the first author.

### 2022-11-28 manuscript.tex
fix broken link in appendix
fix mean 2000km radius method in sim; start updating manuscript

@@
-\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
 \PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
 %
-\documentclass[english,man,floatsintext]{apa6}
+\documentclass[
+  man,floatsintext]{apa6}
+\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
   \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provides euro and other symbols
-\else % if luatex or xelatex
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
   \usepackage{unicode-math}
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
+\IfFileExists{microtype.sty}{% use microtype if available
   \usepackage[]{microtype}
@@
 }{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
   \IfFileExists{parskip.sty}{%
@@
     \setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
 \usepackage{xcolor}
-\usepackage{hyperref}
-\hypersetup{
-            pdftitle={The non-independence of nations and why it matters},
-            pdfauthor={Scott Claessens1 \& Quentin D. Atkinson1,2},
-            pdfkeywords={spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses},
-            colorlinks=true,
-            linkcolor=Maroon,
-            filecolor=Maroon,
-            citecolor=Blue,
-            urlcolor=blue,
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{graphicx,grffile}
+\usepackage{graphicx}
 \makeatletter
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\ifLuaTeX
+\usepackage[bidi=basic]{babel}
+\else
+\usepackage[bidi=default]{babel}
+\fi
+\babelprovide[main,import]{english}
+% get rid of language-specific shorthands (see #6817):
+\let\LanguageShortHands\languageshorthands
+\def\languageshorthands#1{}
 % Manuscript styling
@@
 % \newenvironment{ltable}
-%   {\begin{landscape}\begin{center}\begin{threeparttable}}
-%   {\end{threeparttable}\end{center}\end{landscape}}
-\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}
+%   {\begin{landscape}\centering\begin{threeparttable}}
+%   {\end{threeparttable}\end{landscape}}
+\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}
 
@@
 {}{\PatchFailed}
-\keywords{spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses\newline\indent Word count: 5459 words}
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
-\DeclareDelayedFloatFlavor{lltable}{table}
-\DeclareDelayedFloatFlavor*{longtable}{table}
-\makeatletter
-\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
-\makeatother
+\keywords{spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses\newline\indent Word count: xxxx words}
 \usepackage{lineno}
@@
 \captionsetup[figure]{font={stretch=1.5}}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[shorthands=off,main=english]{babel}
-\else
-  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
-  \usepackage{polyglossia}
-  \setmainlanguage[]{english}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
 \fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={The non-independence of nations and why it matters},
+  pdfauthor={Scott Claessens1 \& Quentin D. Atkinson1,2},
+  pdflang={en-EN},
+  pdfkeywords={spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses},
+  colorlinks=true,
+  linkcolor={Maroon},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={blue},
+  pdfcreator={LaTeX via pandoc}}
 
@@
 Given the importance of nations for structuring human behaviour, there is justifiably huge interest in statistical analyses that attempt to predict variation in national outcomes of all kinds.
-At the time of writing, a search in Web of Science for the terms \enquote{cross-national}, \enquote{cross-cultural}, or \enquote{cross-country} in titles and abstracts returned 80,000 unique hits.
+At the time of writing, a search in Web of Science for the term ``cross-national'' in titles or abstracts returned over 13,000 unique hits.
 The standard practice for cross-national analyses is to conduct bivariate correlations or multiple regressions with individual data points representing different nations.
@@
 For example, the neighbouring African nations Zambia and Tanzania have similar levels of terrain ruggedness, which has been used to partially explain their similar levels of economic development\textsuperscript{23}.
-This pattern conforms to Tobler's first law of geography: \enquote{everything is related to everything else, but near things are more related than distant things}\textsuperscript{24} (p.~236).
+This pattern conforms to Tobler's first law of geography: ``everything is related to everything else, but near things are more related than distant things''\textsuperscript{24} (p.~236).
 
@@
 Shared cultural ancestry can result in a form of pseudoreplication, whereby multiple instances of the same trait across nations are merely duplicates of the ancestral original.
-For example, despite being separated by over 1,500 kilometres of ocean, the related island nations Tonga and Tuvalu share similar languages and customs due to cultural inheritance from a common Polynesian population dating back more than 1,000 years.
+For example, the related island nations Tonga and Tuvalu share similar languages and customs due to cultural inheritance from a common Polynesian population dating back more than 1,000 years.
 More recently, shared ancestry explains a myriad of cultural similarities between colonial settlements and their colonisers (e.g.~Argentina and Spain).
+Importantly, these deep cultural connections between nations often span large geographic distances around the world.
+Tonga and Tuvalu share cultural traits despite being separated by over 1,500 kilometres of ocean, and Argentina and Spain remain culturally similar today despite their location on two separate continents.
+Shared cultural ancestry must therefore be considered independently of spatial proximity in the study of nations.
 
@@
 But without accounting for spatial or cultural non-independence between nations, model residuals can show structure that remains unaccounted for, violating this assumption.
-Treating nations as independent can thus inflate false positive rates\textsuperscript{29}, producing spurious \enquote{direct} relationships between variables that in fact only indirectly covary due to spatial or cultural connections\textsuperscript{30} (see Supplementary Figure \ref{fig:dag} for an example causal model).
-
-Non-independence between data points is widely acknowledged in fields that routinely deal with spatially or culturally structured data.
+Treating nations as independent can thus inflate false positive rates\textsuperscript{29}, producing spurious ``direct'' relationships between variables that in fact only indirectly covary due to spatial or cultural connections\textsuperscript{30} (see Supplementary Figure \ref{fig:dag} for an example causal model).
+
+Non-independence is widely acknowledged in fields that routinely deal with spatially or culturally structured data.
 In ecology and sociology, studies with regional-level data use a variety of autoregressive models designed to account for spatial patternings\textsuperscript{31,32}.
 In anthropology, researchers have recognised cultural non-independence as an important statistical pitfall for over 200 years, with issues of cultural pseudoreplication being identified in early comparative studies of marriage practices across societies\textsuperscript{25}.
-More recently, the Standard Cross-Cultural Sample of 186 cultures was compiled to minimise the confounding effects of non-independence in comparisons of traditional societies\textsuperscript{33}, though spatial and cultural dependencies are difficult to remove entirely\textsuperscript{34,35}.
-Anthropologists also borrow phylogenetic comparative methods from evolutionary biology, such as phylogenetic least squares regression\textsuperscript{36}, when comparing traditional societies, treating culturally related societies in the same way as biologists treat genetically related species (e.g.\textsuperscript{37,38}).
+In the twentieth century, anthropologists began to emphasise that human societies do not develop independently, but rather exist in a globally interconnected ``world system'' linked by shared history and cultural ancestry\textsuperscript{33}.
+Researchers compiled the Standard Cross-Cultural Sample of 186 cultures to minimise the confounding effects of this non-independence in comparisons of human societies\textsuperscript{34}, though spatial and cultural dependencies are difficult to remove entirely\textsuperscript{35,36}.
+Today, anthropologists borrow phylogenetic comparative methods from evolutionary biology, such as phylogenetic least squares regression\textsuperscript{37}, when comparing societies, treating culturally related societies in the same way as biologists treat genetically related species (e.g.\textsuperscript{38,39}).
 
 At the national level, recent reanalyses have revealed that several cross-national relationships reported in economics and psychology do not hold when controlling for non-independence between nations.
-One study replicated 25 analyses of \enquote{persistence} in economics, in which modern national outcomes are regressed against historical characteristics of those nations, and found that over half of the relationships were attenuated when controlling for spatial non-independence\textsuperscript{39}.
-Another replication study found that many of the widely publicised relationships between national-level pathogen prevalence and political institutions and attitudes fail to hold when controlling for various kinds of non-independence\textsuperscript{40}.
-These reanalyses, and others\textsuperscript{41--43}, raise the question: how widespread a concern is non-independence in studies of national-level outcomes?
+One working paper replicated 25 analyses of ``persistence'' in economics, in which modern national outcomes are regressed against historical characteristics of those nations, and found that over half of the relationships were attenuated when controlling for spatial non-independence\textsuperscript{40}.
+Another replication study found that many of the widely publicised relationships between national-level pathogen prevalence and political institutions and attitudes fail to hold when controlling for various kinds of non-independence\textsuperscript{41}.
+These reanalyses, and others\textsuperscript{42--44}, raise the question: how widespread a concern is non-independence in studies of national-level outcomes?
 
@@
 
-In order to motivate our research question, it is important to first show that economic development and cultural values are indeed spatially and culturally non-independent around the world.
-To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for the Human Development Index\textsuperscript{44} and two primary dimensions of cultural values from the World Values Survey, traditional vs.~secular values and survival vs.~self-expression values\textsuperscript{16}.
-For all three of these variables, we found that a substantial proportion of national-level variation was explained by spatial proximity and shared cultural ancestry between nations (Figure \ref{fig:plotSignal}).
-In our models, shared cultural ancestry explained over half of the national-level variation in economic development and cultural values, with spatial proximity explaining most of the remaining national-level variance.
-Bayes Factors indicated strong evidence that these estimates of geographic and cultural phylogenetic signal differed from zero, aside from the geographic signal for survival values, for which the evidence was equivocal (see Supplementary Results).
+In order to motivate our research question, it is important to first quantify the degree of spatial and cultural non-independence for economic development and cultural values around the world.
+If these variables are only weakly non-independent, then the issue might be safe to ignore.
+However, if they are strongly non-independent, then there is a possibility that non-independence could be confounding cross-national inferences.
+
+To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for a range of economic development and cultural values variables that are widely used in the literature.
+For economic development, we focused on the Human Development Index\textsuperscript{45}, gross domestic product per capita, gross domestic product per capita growth, and the Gini index of income inequality.
+For cultural values, we focused on two primary dimensions of cultural values from the World Values Survey, traditional vs.~secular values and survival vs.~self-expression values\textsuperscript{16}, as well as cultural tightness\textsuperscript{14} and individualism\textsuperscript{15}.
+
+For all of these variables, we found that a substantial proportion of national-level variation was explained by spatial proximity and/or shared cultural ancestry between nations (Figure \ref{fig:plotSignal}; Supplementary Table \ref{tab:tableSignal}).
+Signal estimates were often markedly strong, with spatial proximity and shared cultural ancestry frequently explaining over half of the national-level variation.
+For spatial proximity, Bayes Factors indicated strong evidence that the geographic signal estimates differed from zero for all economic development variables and traditional values.
+However, the evidence was only equivocal for survival values and individualism, and strong evidence was found that the geographic signal estimate for tightness was equal to zero.
+For shared cultural ancestry, Bayes Factors indicated strong evidence that the cultural phylogenetic signal estimates differed from zero for all economic development and cultural values variables except for gross domestic product per capita growth, for which the evidence was equivocal.
 These findings emphasise the need to account for spatial and cultural phylogenetic non-independence in cross-national analyses of economic development and cultural values.
@@
 \includegraphics{manuscript_files/figure-latex/plotSignal-1.pdf}
-\caption{\label{fig:plotSignal}\emph{Posterior estimates of geographic and cultural phylogenetic signal for the Human Development Index and Inglehart's traditional and survival values.} Geographic and cultural phylogenetic signal are operationalised as the proportion of national-level variance explained by geographic and linguistic proximity matrices.
-Grey ridges are full posterior distributions, points are posterior median values, and lines are 50\% and 95\% credible intervals.}
+\caption{\label{fig:plotSignal}\emph{Posterior estimates of geographic and cultural phylogenetic signal for a range of economic development and cultural values variables.} Geographic and cultural phylogenetic signal are operationalised as the proportion of national-level variance explained by geographic and linguistic proximity matrices.
+Grey ridges are full posterior distributions, points are posterior median values, and black lines are 50\% and 95\% credible intervals.
+HDI = Human Development Index; GDP = gross domestic product.}
 \end{figure}
@@
 Given that economic development and cultural values show evidence of geographic and cultural phylogenetic signal, have cross-national analyses sufficiently accounted for this non-independence?
-To assess this, we systematically searched the published literature for articles that combined the search terms \enquote{economic development} or \enquote{values} with the search terms \enquote{cross-national}, \enquote{cross-cultural}, or \enquote{cross-country}.
+To assess this, we systematically searched the published literature for articles that combined the search terms ``economic development'' or ``values'' with the search terms ``cross-national'', ``cross-cultural'', or ``cross-country''.
 We removed articles that did not report original research, were not relevant to economic development or cultural values, or did not report at least one cross-national analysis.
-We then retained the 100 articles (50 for economic development, 50 for cultural values) with the highest annual rate of citations.
+We then retained the 100 articles (50 for economic development, 50 for cultural values) with the highest annual rate of citations (Supplementary Table \ref{tab:tableArticlesReview}).
 For each of these highly-cited articles, we exhaustively recorded every cross-national analysis reported in the main text (n = 4,308), identifying in each case whether or not the analysis attempted to control for spatial, cultural, or any other form of non-independence between nations (see Methods for detailed search criteria and coding decisions).
 
-The results of our systematic review show that most published articles and cross-national analyses make no attempt to account for statistical non-independence.
-Figure \ref{fig:plotReview}a plots the proportion of articles that contain at least one cross-national analysis accounting for non-independence.
+The results of our systematic review show that most published articles containing cross-national analyses make no attempt to account for statistical non-independence.
+Figure \ref{fig:plotReview1}a plots the proportion of articles that contain at least one cross-national analysis accounting for non-independence.
 We find that 42\% of economic development articles contain at least one attempt to control for non-independence (95\% bootstrap confidence interval {[}0.30 0.54{]}), while this proportion decreases to only 8\% for cultural values articles (95\% bCI {[}0.02 0.16{]}).
 Both kinds of article are most likely to use regional fixed effects (e.g.~continent fixed effects) to account for non-independence, but some articles also include controls for spatial distance (e.g.~latitude) and shared cultural history (e.g.~colony status).
+These proportions are even lower when focusing on the full sample of 4,308 analyses: only 5\% (95\% credible interval {[}0.02 0.13{]}) of individual economic development analyses and 1\% (95\% CI {[}0.00 0.02{]}) of individual cultural values analyses are estimated to control for non-independence (Supplementary Figure \ref{fig:plotReview2}).
 
@@
 \centering
-\includegraphics{manuscript_files/figure-latex/plotReview-1.pdf}
-\caption{\label{fig:plotReview}\emph{Results from systematic review of 100 highly-cited cross-national studies of economic development (red) and cultural values (blue).} (a) Proportion of articles containing at least one analysis accounting for non-independence, overall and split by common methods of controlling for non-independence.
+\includegraphics{manuscript_files/figure-latex/plotReview1-1.pdf}
+\caption{\label{fig:plotReview1}\emph{Results from systematic review of 100 highly-cited cross-national studies of economic development (red) and cultural values (blue).} (a) Proportion of articles containing at least one analysis accounting for non-independence, overall and split by common methods of controlling for non-independence.
+(b) The association between journal impact factor and the probability that an article contains at least one analysis accounting for non-independence.
+(c) Estimated trend over time for the probability that an article contains at least one analysis accounting for non-independence.
 Point ranges represent proportions and 95\% bootstrap confidence intervals.
-(b) Adjusted proportion of individual analyses accounting for non-independence, overall and split by method.
-Point ranges are posterior medians and 95\% credible intervals from Bayesian multilevel logistic regressions.
-(c) Histogram of publication years for studies of economic development and cultural values.
-(d) Estimated trend over time for the probability of controlling for non-independence.
-Lines and shaded areas are posterior median regression lines and 50\% credible intervals from a Bayesian multilevel spline model.
+Lines and shaded areas are posterior median regression lines and 50\% credible intervals from Bayesian multilevel models.
 Region FEs = region fixed effects.}
@@
 
-Focusing on the full sample of 4,308 analyses, we find that the proportion of individual cross-national analyses accounting for non-independence is even lower (Figure \ref{fig:plotReview}b).
-Across 2,487 cross-national analyses from studies of economic development, only 5\% are estimated to control for non-independence (95\% credible interval {[}0.02 0.15{]}).
-Similarly, across 1,821 cross-national analyses from studies of cultural values, only 1\% are estimated to control for non-independence (95\% CI {[}0.00 0.02{]}).
-
-Since our systematic review goes back as far as 1993 (Figure \ref{fig:plotReview}c), it is possible that our estimates are being biased by earlier studies, and that controls for non-independence have increased over time with methodological advancements and greater awareness of the issue.
-To test this possibility, we fitted a time trend to the full sample of analyses with a Bayesian multilevel spline model.
-However, we found that, for both studies of economic development and cultural values, the estimated probability of controlling for non-independence has remained low since 1993 (Figure \ref{fig:plotReview}d).
-
-\hypertarget{common-methods-of-controlling-for-non-independence-produce-inflated-false-positive-rates}{%
-\subsection{Common methods of controlling for non-independence produce inflated false positive rates}\label{common-methods-of-controlling-for-non-independence-produce-inflated-false-positive-rates}}
+While our review contains a range of articles from low- and high-impact journals, an anonymous reviewer suggested that our estimates could be biased downwards by analyses published in lower impact outlets with more relaxed standards for issues like non-independence.
+It is also possible that, since our systematic review goes back as far as 1993, our estimates are being biased downwards by earlier studies, and that controls for non-independence have increased over time with methodological advancements and greater awareness of the issue.
+To test these possible explanations for our low estimates, we fitted Bayesian logistic regression models to the data from the review, including log journal impact factor and publication year as separate predictors.
+Interestingly, we found that, contrary to the explanation above, studies from higher impact journals were \emph{less} likely to include at least one control for non-independence than studies from lower impact journals, both for studies of economic development (b = -0.38, 95\% CI {[}-0.87 0.09{]}) and for studies of cultural values (b = -0.56, 95\% CI {[}-1.10 -0.05{]}; Figure \ref{fig:plotReview1}b).
+Moreover, splines revealed no relationship between publication year and the probability of including at least one control for non-independence, both for studies of economic development (b = -0.13, 95\% CI {[}-1.10 0.82{]}) and for studies of cultural values (b = -0.02, 95\% CI {[}-1.00 0.96{]}; Figure \ref{fig:plotReview1}c).
+
+\hypertarget{common-methods-of-controlling-for-non-independence-are-insufficient-for-reducing-false-positive-rates-in-non-independent-data}{%
+\subsection{Common methods of controlling for non-independence are insufficient for reducing false positive rates in non-independent data}\label{common-methods-of-controlling-for-non-independence-are-insufficient-for-reducing-false-positive-rates-in-non-independent-data}}
 
@@
 We then fitted naive regressions without controls to these datasets, as well as regression models with controls for latitude, longitude, and continent fixed effects.
-Despite not being identified in our systematic review, we also included fixed effects for the language families of the majority-spoken languages in each country, as this control is often used to account for cultural phylogenetic non-independence (e.g.\textsuperscript{45}).
-In addition, we included Conley standard errors, a widely used standard error correction that purportedly accounts for spatial non-independence\textsuperscript{46,47}.
-Finally, we included Bayesian multilevel regressions that explicitly model spatial and/or cultural phylogenetic non-independence by allowing nation random intercepts to covary according to global geographic and/or linguistic proximity matrices (see Supplementary Methods).
-To model spatial non-independence we included a Gaussian process\textsuperscript{48,49} over latitude and longitude values, and to model cultural phylogenetic non-independence we assumed that nation random intercepts were correlated in proportion to their linguistic proximity\textsuperscript{50}.
+Despite not being identified in our systematic review, we also included other methods that are often used in the literature to account for non-independence.
+Additional spatial controls included the mean of the predictor variable within a surrounding 2000km radius (e.g.\textsuperscript{46}) and Conley standard errors\textsuperscript{47,48} based on geographic distances between nations (e.g.\textsuperscript{46,\textbf{Schulz2021?}}).
+Additional cultural controls included fixed effects for the language families of the majority-spoken languages in each nation (e.g.\textsuperscript{49}) and Conley standard errors based on genetic distances between nations (e.g.\textsuperscript{46,\textbf{Schulz2021?}}).
+These frequentist approaches attempt to account for non-independence by holding geographic location constant (latitude, longitude), discarding between-region variation and exploiting only local variation (continent fixed effects, mean of surrounding 2000km), or correcting standard errors for autocorrelation post-hoc while leaving model coefficients unchanged (Conley standard errors).
+
+Beyond frequentist approaches, we also fitted Bayesian multilevel regressions that explicitly model spatial and/or cultural phylogenetic non-independence by allowing nations to covary according to geographic and/or linguistic proximity matrices.
+Geographic proximity between nations is calculated from inverse distances between longitude and latitude coordinates.
+Linguistic proximity between nations is calculated from a global phylogenetic tree that represents hierarchical relationships of genealogical descent for all languages in the world.
+For each pair of nations, we calculate inverse phylogenetic distances (i.e.~number of branches separating two taxa) between all languages spoken in that nation pair and produce an average ``linguistic proximity'' score weighted by the percentages of speakers within those nations.
+To include the resulting geographic and linguistic proximity matrices in our models, we included a Gaussian process\textsuperscript{50,51} over latitude and longitude values and/or assumed that nation random intercepts were correlated in proportion to their linguistic proximity\textsuperscript{52}.
+These approaches attempt to account for non-independence by modelling the covariance between nations that is induced by their geographic or linguistic connections.
+
+Figures \ref{fig:plotSim3} and \ref{fig:plotSim4} plot the estimated false positive rates from our simulation study, split by different methods and different degrees of spatial or cultural phylogenetic autocorrelation (see Supplementary Figures \ref{fig:plotSim1} and \ref{fig:plotSim2} for full distributions of effect sizes under strong autocorrelation).
 Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero (i.e.~falsely infering a relationship when none is present).
-
-Figures \ref{fig:plotSim3} and \ref{fig:plotSim4} plot the estimated false positive rates from our simulation study, split by different methods and different degrees of spatial or cultural phylogenetic autocorrelation (see Supplementary Figures \ref{fig:plotSim1} and \ref{fig:plotSim2} for full distributions of effect sizes under strong autocorrelation).
-For reference, \enquote{weak} autocorrelation in our simulation is comparable to the geographic signal for survival values in Figure \ref{fig:plotSignal}, while \enquote{moderate} and \enquote{strong} levels of autocorrelation are comparable to the cultural phylogenetic signal for traditional and survival values, respectively.
+For reference, ``weak'' autocorrelation in our simulation is comparable to the geographic signal for survival values in Figure \ref{fig:plotSignal} (i.e.~20\% of the national-level variance is explained by non-independence), while ``moderate'' and ``strong'' levels of autocorrelation are comparable to the cultural phylogenetic signal for traditional and survival values, respectively (i.e.~50\% and 80\% of the national-level variance is explained by non-independence).
 
@@
 This false positive rate increases as the degree of autocorrelation increases.
-With strong spatial autocorrelation for both outcomes and predictors, false positive rates reach as high as 71\%.
-We find a slightly lower false positive rate under strong cultural phylogenetic autocorrelation, though this false positive rate is still greater than expected by chance (38\%).
+With strong spatial autocorrelation for both outcomes and predictors, false positive rates reach as high as 78\%.
+We find a slightly lower false positive rate under strong cultural phylogenetic autocorrelation, though this false positive rate is still greater than expected by chance (39\%).
 
@@
 Applying Conley standard errors also does not reduce false positive rates below 50\% under strong spatial autocorrelation, regardless of the distance cutoff.
-Continent fixed effects are more effective than other frequentist methods, though they continue to produce a false positive rate of 28\% under strong spatial autocorrelation.
+Continent fixed effects are more effective than other frequentist methods, though they continue to produce a false positive rate of 56\% under strong spatial autocorrelation.
 By contrast, Bayesian spatial Gaussian process regression with longitude and latitude values outperforms all other methods.
-This approach eliminates false positives under moderate spatial autocorrelation, such that the false positive rate is no different from chance, and reduces the false positive rate under strong spatial autocorrelation to 17\%.
+This approach eliminates false positives under moderate spatial autocorrelation, such that the false positive rate is no different from chance, and reduces the false positive rate under strong spatial autocorrelation to 54\%.
 Bayesian models that additionally account for linguistic proximity between nations perform equally well, though models with only linguistic covariance continue to produce false positives.
@@
 Controls for latitude and longitude, continent fixed effects, and Conley standard errors do little to change false positive rates.
-Language family fixed effects are slightly more effective than other frequentist methods, though they continue to produce a false positive rate of 29\% under strong cultural phylogenetic autocorrelation.
+Language family fixed effects are slightly more effective than other frequentist methods, though they continue to produce a false positive rate of 39\% under strong cultural phylogenetic autocorrelation.
 By contrast, Bayesian models with random effects covarying according to linguistic proximity completely eliminate false positives across all degrees of cultural phylogenetic autocorrelation.
@@
 To determine how widespread this issue of spurious cross-national relationships might be, we reanalysed a subset of twelve previous cross-national analyses from our systematic review, sufficiently controlling for spatial and cultural phylogenetic non-independence using global geographic and linguistic proximity matrices.
-We subsampled six analyses from our economic development review\textsuperscript{51--56} and six from our cultural values review\textsuperscript{13,14,16,57--59}.
+We subsampled six analyses from our economic development review\textsuperscript{53--58} and six from our cultural values review\textsuperscript{13,14,16,59--61}.
 Our choice of analyses was constrained by data availability and whether we were able to initially replicate the original finding.
@@
 All outcome and predictor variables are standardised.
-Most analyses are simple bivariate cross-national correlations, but Gelfand et al.
-(2011) is a partial correlation controlling for log gross national income and Adamczyk and Pitt (2009) is a multilevel model including several covariates.
-Point ranges represent posterior medians and 95\% credible intervals.
+Most analyses are simple bivariate cross-national correlations, but Gelfand et al.~(2011) is a partial correlation controlling for log gross national income and Adamczyk and Pitt (2009) is a multilevel model including several covariates.
+Points and black lines represent posterior medians and 95\% credible intervals.
 GDP = gross domestic product.
@@
 For each outcome variable, our Gaussian process models provided varying estimates of how quickly spatial autocorrelation declined with distance (Supplementary Figure \ref{fig:plotReplications3}).
-For example, in Skidmore and Toya\textsuperscript{56} gross domestic product growth was only moderately spatially autocorrelated at 1,000 km distance (posterior median spatial autocorrelation at 1,000 km = 0.47, 95\% CI {[}0.06 0.97{]}), whereas in Inglehart and Baker\textsuperscript{16} traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1,000 km = 0.96, 95\% CI {[}0.78 0.99{]}).
-We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \ref{fig:plotReplications4}), with some outcome variables expressing low signal (e.g.~tightness\textsuperscript{14}; posterior median = 0.11, 95\% CI {[}0.00 0.82{]}) and others expressing high signal (e.g.~female labour force participation\textsuperscript{58}; posterior median = 0.91, 95\% CI {[}0.67 0.99{]}).
+For example, in Skidmore and Toya\textsuperscript{58} gross domestic product growth was only moderately spatially autocorrelated at 1,000 km distance (posterior median spatial autocorrelation at 1,000 km = 0.42, 95\% CI {[}0.07 0.90{]}), whereas in Inglehart and Baker\textsuperscript{16} traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1,000 km = 0.96, 95\% CI {[}0.81 0.99{]}).
+We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \ref{fig:plotReplications4}), with some outcome variables expressing low signal (e.g.~tightness\textsuperscript{14}; posterior median = 0.15, 95\% CI {[}0.00 0.85{]}) and others expressing high signal (e.g.~female labour force participation\textsuperscript{60}; posterior median = 0.89, 95\% CI {[}0.63 0.98{]}).
 Across all analyses, we found that stronger estimates of spatial autocorrelation or cultural phylogenetic signal resulted in a more pronounced reduction in the effect size when controlling for non-independence between nations (Figure \ref{fig:plotReplications5}).
@@
 Our initial analyses add to and clarify existing evidence regarding the non-independence of economic and cultural variation among nations.
-One previous study suggested that geographic proximity is more important than deep cultural ancestry in explaining the distribution of human development across Eurasian nations, though the authors noted that their small sample of 44 nations and regional focus limited their statistical power\textsuperscript{60}.
+One previous study suggested that geographic proximity is more important than deep cultural ancestry in explaining the distribution of human development across Eurasian nations, though the authors noted that their small sample of 44 nations and regional focus limited their statistical power\textsuperscript{62}.
 Our global sample of nations revealed strong cultural phylogenetic signal, as well as geographic signal, for the Human Development Index.
@@
 This result calls into question controls like distance to the equator to account for non-independence in cross-national regression models, though these controls may still be suitable to account for regional or latitudinal variation in ecology, which we did not simulate.
-High false positive rates persisted with Conley standard errors, which have previously been critiqued for being overly sensitive to arbitrary distance cutoffs\textsuperscript{39}.
-The simulation also confirmed the assertion that fixed effects for spatial or cultural groupings (e.g.~continent or language family fixed effects) are insufficient because non-independence still remains within groupings\textsuperscript{40}.
-This logic further applies to analyses that control for non-independence by separately analysing different regions (e.g.\textsuperscript{61}).
-Across all model types in our simulation, the only methods that sufficiently reduced the false positive rate were the Bayesian multilevel regressions that explicitly modelled spatial and cultural phylogenetic autocorrelation, though we did not include other possible controls for non-independence, such as conditional autoregressive models\textsuperscript{31} or generalised additive models\textsuperscript{62}.
+High false positive rates persisted with Conley standard errors, which have previously been critiqued for being overly sensitive to arbitrary distance cutoffs\textsuperscript{40}.
+The simulation also confirmed the assertion that fixed effects for spatial or cultural groupings (e.g.~continent or language family fixed effects) are insufficient because non-independence still remains within groupings\textsuperscript{41}.
+This logic further applies to analyses that control for non-independence by separately analysing different regions (e.g.\textsuperscript{63}).
+Across all model types in our simulation, the only methods that sufficiently reduced the false positive rate were the Bayesian multilevel regressions that explicitly modelled spatial and cultural phylogenetic autocorrelation, though we did not include other possible controls for non-independence, such as conditional autoregressive models\textsuperscript{31} or generalised additive models\textsuperscript{64}.
 
 Ours is not the first review to show that studies are misapplying statistical methods in ways that inflate false positive rates.
-For example, other literature reviews have shown that studies in the social sciences tend to use small samples of participants\textsuperscript{63}, treat ordinal data as metric\textsuperscript{64}, incorrectly handle missing values\textsuperscript{65}, and ignore best practices in meta-analyses\textsuperscript{66}.
+For example, other literature reviews have shown that studies in the social sciences tend to use small samples of participants\textsuperscript{65}, treat ordinal data as metric\textsuperscript{66}, incorrectly handle missing values\textsuperscript{67}, and ignore best practices in meta-analyses\textsuperscript{68}.
 Why do cross-national studies also rarely account for non-independence?
-At the institutional level, one possibility is that such practices are incentivised \emph{because} they generate statistically significant relationships, which increase the probability that a study is published\textsuperscript{63}.
+At the institutional level, one possibility is that such practices are incentivised \emph{because} they generate statistically significant relationships, which increase the probability that a study is published\textsuperscript{65}.
 At the individual level, another possibility is that researchers outside of anthropology and ecology are simply not aware of the problem, or believe that the problem does not apply to analyses of nations.
@@
 These institutional- and individual-level barriers can be combatted.
-First, cross-national replication studies like ours and others\textsuperscript{39--43}, combined with the methodological reviews included in Registered Reports\textsuperscript{67}, might change incentive structures and encourage researchers to analyse the world's nations with more rigorous methods.
+First, cross-national replication studies like ours and others\textsuperscript{40--44}, combined with the methodological reviews included in Registered Reports\textsuperscript{69}, might change incentive structures and encourage researchers to analyse the world's nations with more rigorous methods.
 Second, since the issue of non-independence is fundamentally an issue of causal inference (Supplementary Figure \ref{fig:dag}), more explicit descriptions of causal models could promote controls for non-independence in cross-national research.
 In our review, economists studying economic development dealt with national-level non-independence more than psychologists studying cultural values, likely because economics studies tend to be lengthy statistical exercises that systematically incorporate or exclude numerous variables in an attempt to infer causation.
-Third, the recent widespread accessibility of open source statistical software, such as the programming language Stan\textsuperscript{68} and the R package \emph{brms}\textsuperscript{69}, should promote the use of more rigorous methods to control for non-independence.
+Third, the recent widespread accessibility of open source statistical software, such as the programming language Stan\textsuperscript{70} and the R package \emph{brms}\textsuperscript{71}, should promote the use of more rigorous methods to control for non-independence.
 Using \emph{brms}, for example, Bayesian Gaussian process regression is straightforward to conduct, requiring only longitude and latitude values for nations.
@@
 In our reanalyses, we found that four out of six cross-national correlations with economic development variables had 95\% credible intervals that included zero when controlling for spatial non-independence.
-Three of these analyses were tests of \enquote{persistence} hypotheses, studying the effects of historical and environmental conditions --- settler mortality\textsuperscript{51}, wheat-sugar suitability\textsuperscript{55}, and natural disaster frequency\textsuperscript{56} --- on modern developmental outcomes.
-A recent reanalysis has also called into question various studies of this ilk\textsuperscript{39}.
+Three of these analyses were tests of ``persistence'' hypotheses, studying the effects of historical and environmental conditions --- settler mortality\textsuperscript{53}, wheat-sugar suitability\textsuperscript{57}, and natural disaster frequency\textsuperscript{58} --- on modern developmental outcomes.
+A recent reanalysis has also called into question various studies of this ilk\textsuperscript{40}.
 We also found that two out of six cross-national correlations with cultural values variables had 95\% credible intervals that included zero when controlling for cultural phylogenetic non-independence.
@@
 On the contrary, such work promises to deepen understanding of our world, including the causes and consequences of economic development and cultural values.
-Moreover, cross-national studies allow social scientists to broaden their scope of study beyond Western populations\textsuperscript{7}, providing the representative samples necessary to test evolutionary and socio-ecological theories of human behaviour\textsuperscript{8,70}.
+Moreover, cross-national studies allow social scientists to broaden their scope of study beyond Western populations\textsuperscript{7}, providing the representative samples necessary to test evolutionary and socio-ecological theories of human behaviour\textsuperscript{8,72}.
 But in order to minimise spurious relationships in global datasets, we urge researchers to control for spatial and cultural phylogenetic non-independence when reporting cross-national correlations.
@@
 To estimate the degree of spatial and cultural phylogenetic non-independence in economic development and cultural values, we calculated geographic and cultural phylogenetic signal for global measures of development and values.
-Our measure of economic development was the Human Development Index\textsuperscript{44}.
+Our measure of economic development was the Human Development Index\textsuperscript{45}.
 We retrieved a longitudinal dataset capturing human development for 189 nations since 1990 (n = 1,512; \url{https://hdr.undp.org/en/content/download-data}).
@@
 We included these matrices in Bayesian multilevel models, allowing nation random intercepts to covary according to both geographic and linguistic proximity simultaneously.
-These models were fitted with the R package \emph{brms}\textsuperscript{69} and converged normally (\(\hat{R}\) \textless{} 1.1).
+These models were fitted with the R package \emph{brms}\textsuperscript{71} and converged normally (\(\hat{R}\) \textless{} 1.1).
 Estimates of geographic and cultural phylogenetic signal were computed as the proportion of national-level variance in these models explained by geographic and linguistic proximity matrices.
@@
 We exported two searches from Web of Science (\url{https://www.webofknowledge.com/}) on 27\textsuperscript{th} September 2021, restricting our searches to articles published between 1900 and 2018.
-The first search was for the terms \enquote{economic development} AND (\enquote{cross-national} OR \enquote{cross-cultural} OR \enquote{cross-country}), which returned 965 articles.
-The second search was for the terms \enquote{values} AND (\enquote{cross-national} OR \enquote{cross-cultural} OR \enquote{cross-country}), which returned 6806 articles.
+The first search was for the terms ``economic development'' AND (``cross-national'' OR ``cross-cultural'' OR ``cross-country''), which returned 965 articles.
+The second search was for the terms ``values'' AND (``cross-national'' OR ``cross-cultural'' OR ``cross-country''), which returned 6806 articles.
 Once exported, we ordered the articles by descending number of citations per year since initial publication, using citation counts reported by Web of Science.
@@
 We coded common attempts to control for non-independence between nations.
-These included: (1) any higher-level control variables for spatial regional groupings (e.g.~continent fixed effects); (2) any geographic distance control variables (e.g.~distance between capital cities, distance from equator, latitude); (3) any control variables capturing shared cultural history (e.g.~former colony, legal origin fixed effects, linguistic history, cultural influence); and (4) any other control variables, tests, or approaches that were deemed as attempts to control for non-independence (e.g.~eigenvector filtering\textsuperscript{71}, controls for trade-weightings between countries, cross-sectional dependence tests\textsuperscript{72}, separate analyses for subsets of countries).
+These included: (1) any higher-level control variables for spatial regional groupings (e.g.~continent fixed effects); (2) any geographic distance control variables (e.g.~distance between capital cities, distance from equator, latitude); (3) any control variables capturing shared cultural history (e.g.~former colony, legal origin fixed effects, linguistic history, cultural influence); and (4) any other control variables, tests, or approaches that were deemed as attempts to control for non-independence (e.g.~eigenvector filtering\textsuperscript{73}, controls for trade-weightings between nations, cross-sectional dependence tests\textsuperscript{74}, separate analyses for subsets of nations).
 These were coded by the first author.
@@
 To test for a trend over time, we also fitted a Bayesian multilevel logistic regression with a multigroup spline for year of publication and random intercepts for articles.
-Bayesian models were fitted with the \emph{brms} R package\textsuperscript{69}.
+Bayesian models were fitted with the \emph{brms} R package\textsuperscript{71}.
 Our priors were informed by prior predictive checks, and all models converged normally (\(\hat{R}\) \textless{} 1.1).
@@
 Models employing Conley standard errors required latitude and longitude values, and the cutoffs implied the distance beyond which autocorrelation is assumed to be zero.
-These models were fitted using the \emph{conleyreg} R package\textsuperscript{73}.
-Bayesian models were fitted using the \emph{brms} R package\textsuperscript{69}.
+These models were fitted using the \emph{conleyreg} R package\textsuperscript{75}.
+Bayesian models were fitted using the \emph{brms} R package\textsuperscript{71}.
 Our choice of priors was based on prior predictive simulation.
@@
 
-The twelve analyses that we settled on\textsuperscript{13,14,16,51--59} were mostly bivariate cross-national correlations, except for two.
-One analysis\textsuperscript{14} additionally controlled for log gross national income, and another analysis\textsuperscript{56} is a multilevel model including random intercepts for nations and several individual-level and national-level covariates (see Model 5 in original paper).
+The twelve analyses that we settled on\textsuperscript{13,14,16,53--61} were mostly bivariate cross-national correlations, except for two.
+One analysis\textsuperscript{14} additionally controlled for log gross national income, and another analysis\textsuperscript{58} is a multilevel model including random intercepts for nations and several individual-level and national-level covariates (see Model 5 in original paper).
 Before running any additional models, we pre-registered these twelve analyses on the Open Science Framework on 25\textsuperscript{th} January 2022 (\url{https://osf.io/uywx8/}).
@@
 
-We fitted these models using the \emph{brms} R package\textsuperscript{69}.
+We fitted these models using the \emph{brms} R package\textsuperscript{71}.
 Our choice of priors was based on prior predictive simulation.
-All models converged normally (\(\hat{R}\) \textless{} 1.1), though for some models we resorted to using approximate Gaussian processes\textsuperscript{74} to reach convergence.
+All models converged normally (\(\hat{R}\) \textless{} 1.1), though for some models we resorted to using approximate Gaussian processes\textsuperscript{76} to reach convergence.
 
@@
 All data and code are accesible at our Open Science Framework repository (\url{https://osf.io/uywx8/}).
-We used the \emph{targets} R package\textsuperscript{75} to create a reproducible data analysis pipeline and the \emph{papaja} R package\textsuperscript{76} to reproducibly generate the manuscript.
+We used the \emph{targets} R package\textsuperscript{77} to create a reproducible data analysis pipeline and the \emph{papaja} R package\textsuperscript{78} to reproducibly generate the manuscript.
 
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Caselli2005}{}%
-1. Caselli, F. Accounting for cross-country income differences. in (eds. Aghion, P. \& Durlauf, S. N.) vol. 1 679--741 (Elsevier, 2005).
-
-\leavevmode\hypertarget{ref-Austin2012}{}%
-2.
-Austin, K.
+\begin{CSLReferences}{0}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-Caselli2005}{}}%
+\CSLLeftMargin{1. }%
+\CSLRightInline{Caselli, F.
+\href{https://doi.org/10.1016/S1574-0684(05)01009-9}{Accounting for cross-country income differences}.
+in (eds.
+Aghion, P.
+\& Durlauf, S.
+N.) vol.
+1 679--741 (Elsevier, 2005).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Austin2012}{}}%
+\CSLLeftMargin{2. }%
+\CSLRightInline{Austin, K.
 F.
@@
 A.
-Disease, war, hunger, and deprivation: A cross-national investigation of the determinants of life expectancy in less-developed and sub-Saharan African nations.
-\emph{Sociological Perspectives} \textbf{55}, 421--447 (2012).
-
-\leavevmode\hypertarget{ref-Rai2013}{}%
-3.
-Rai, D., Zitko, P., Jones, K., Lynch, J.
+\href{https://doi.org/10.1525/sop.2012.55.3.421}{Disease, war, hunger, and deprivation: A cross-national investigation of the determinants of life expectancy in less-developed and sub-{S}aharan {A}frican nations}.
+\emph{Sociological Perspectives} \textbf{55}, 421--447 (2012).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Rai2013}{}}%
+\CSLLeftMargin{3. }%
+\CSLRightInline{Rai, D., Zitko, P., Jones, K., Lynch, J.
 \& Araya, R.
-Country- and individual-level socioeconomic determinants of depression: Multilevel cross-national comparison.
-\emph{British Journal of Psychiatry} \textbf{202}, 195--203 (2013).
-
-\leavevmode\hypertarget{ref-Diener2009}{}%
-4.
-Diener, E., Diener, M.
+\href{https://doi.org/10.1192/bjp.bp.112.112482}{Country- and individual-level socioeconomic determinants of depression: Multilevel cross-national comparison}.
+\emph{British Journal of Psychiatry} \textbf{202}, 195--203 (2013).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Diener2009}{}}%
+\CSLLeftMargin{4. }%
+\CSLRightInline{Diener, E., Diener, M.
 \& Diener, C.
 Factors predicting the subjective well-being of nations.
-in \emph{Culture and well-being: The collected works of Ed Diener} (ed.
+in \emph{Culture and well-being: The collected works of {E}d {D}iener} (ed.
 Diener, E.) 43--70 (Springer Netherlands, 2009).
-doi:\href{https://doi.org/10.1007/978-90-481-2352-0_3}{10.1007/978-90-481-2352-0\_3}.
-
-\leavevmode\hypertarget{ref-Kirkcaldy2004}{}%
-5. Kirkcaldy, B., Furnham, A. \& Siefen, G. The relationship between health efficacy, educational attainment, and well-being among 30 nations. \emph{European Psychologist} \textbf{9}, 107--119 (2004).
-
-\leavevmode\hypertarget{ref-White2021}{}%
-6.
-White, C.
+doi:\href{https://doi.org/10.1007/978-90-481-2352-0_3}{10.1007/978-90-481-2352-0\_3}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Kirkcaldy2004}{}}%
+\CSLLeftMargin{5. }%
+\CSLRightInline{Kirkcaldy, B., Furnham, A.
+\& Siefen, G.
+\href{https://doi.org/10.1027/1016-9040.9.2.107}{The relationship between health efficacy, educational attainment, and well-being among 30 nations}.
+\emph{European Psychologist} \textbf{9}, 107--119 (2004).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-White2021}{}}%
+\CSLLeftMargin{6. }%
+\CSLRightInline{White, C.
 J.
@@
 \& Norenzayan, A.
-Cultural similarity among coreligionists within and between countries.
-\emph{Proceedings of the National Academy of Sciences} \textbf{118}, e2109650118 (2021).
-
-\leavevmode\hypertarget{ref-Henrich2010}{}%
-7. Henrich, J., Heine, S. J. \& Norenzayan, A. The weirdest people in the world? \emph{Behavioral and Brain Sciences} \textbf{33}, 61--83 (2010).
-
-\leavevmode\hypertarget{ref-Pollet2014}{}%
-8. Pollet, T. V., Tybur, J. M., Frankenhuis, W. E. \& Rickard, I. J. What can cross-cultural correlations teach us about human nature? \emph{Human Nature} \textbf{25}, 410--429 (2014).
-
-\leavevmode\hypertarget{ref-Benhabib1994}{}%
-9. Benhabib, J. \& Spiegel, M. M. The role of human capital in economic development: Evidence from aggregate cross-country data. \emph{Journal of Monetary Economics} \textbf{34}, 143--173 (1994).
-
-\leavevmode\hypertarget{ref-Comin2010}{}%
-10. Comin, D., Easterly, W. \& Gong, E. Was the wealth of nations determined in 1000 BC? \emph{American Economic Journal: Macroeconomics} \textbf{2}, 65--97 (2010).
-
-\leavevmode\hypertarget{ref-LaPorta1997}{}%
-11. La Porta, R., Lopez-De-Silanes, F., Shleifer, A. \& Vishny, R. W. Legal determinants of external finance. \emph{The Journal of Finance} \textbf{52}, 1131--1150 (1997).
-
-\leavevmode\hypertarget{ref-Sachs2001}{}%
-12. Sachs, J. D. \& Warner, A. M. The curse of natural resources. \emph{European Economic Review} \textbf{45}, 827--838 (2001).
-
-\leavevmode\hypertarget{ref-Fincher2008}{}%
-13.
-Fincher, C.
+\href{https://doi.org/10.1073/pnas.2109650118}{Cultural similarity among coreligionists within and between countries}.
+\emph{Proceedings of the National Academy of Sciences} \textbf{118}, e2109650118 (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Henrich2010}{}}%
+\CSLLeftMargin{7. }%
+\CSLRightInline{Henrich, J., Heine, S.
+J.
+\& Norenzayan, A.
+\href{https://doi.org/10.1017/S0140525X0999152X}{The weirdest people in the world?} \emph{Behavioral and Brain Sciences} \textbf{33}, 61--83 (2010).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Pollet2014}{}}%
+\CSLLeftMargin{8. }%
+\CSLRightInline{Pollet, T.
+V., Tybur, J.
+M., Frankenhuis, W.
+E.
+\& Rickard, I.
+J.
+\href{https://doi.org/10.1007/s12110-014-9206-3}{What can cross-cultural correlations teach us about human nature?} \emph{Human Nature} \textbf{25}, 410--429 (2014).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Benhabib1994}{}}%
+\CSLLeftMargin{9. }%
+\CSLRightInline{Benhabib, J.
+\& Spiegel, M.
+M.
+\href{https://doi.org/10.1016/0304-3932(94)90047-7}{The role of human capital in economic development: Evidence from aggregate cross-country data}.
+\emph{Journal of Monetary Economics} \textbf{34}, 143--173 (1994).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Comin2010}{}}%
+\CSLLeftMargin{10. }%
+\CSLRightInline{Comin, D., Easterly, W.
+\& Gong, E.
+\href{http://www.jstor.org/stable/25760309}{Was the wealth of nations determined in 1000 {BC}?} \emph{American Economic Journal: Macroeconomics} \textbf{2}, 65--97 (2010).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-LaPorta1997}{}}%
+\CSLLeftMargin{11. }%
+\CSLRightInline{La Porta, R., Lopez-De-Silanes, F., Shleifer, A.
+\& Vishny, R.
+W.
+\href{https://doi.org/10.1111/j.1540-6261.1997.tb02727.x}{Legal determinants of external finance}.
+\emph{The Journal of Finance} \textbf{52}, 1131--1150 (1997).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Sachs2001}{}}%
+\CSLLeftMargin{12. }%
+\CSLRightInline{Sachs, J. D. \& Warner, A. M. \href{https://doi.org/10.1016/S0014-2921(01)00125-8}{The curse of natural resources}. \emph{European Economic Review} \textbf{45}, 827--838 (2001).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Fincher2008}{}}%
+\CSLLeftMargin{13. }%
+\CSLRightInline{Fincher, C.
 L., Thornhill, R., Murray, D.
@@
 \& Schaller, M.
-Pathogen prevalence predicts human cross-cultural variability in individualism/collectivism.
-\emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{275}, 1279--1285 (2008).
-
-\leavevmode\hypertarget{ref-Gelfand2011}{}%
-14. Gelfand, M. J. \emph{et al.} Differences between tight and loose cultures: A 33-nation study. \emph{Science} \textbf{332}, 1100--1104 (2011).
-
-\leavevmode\hypertarget{ref-Hofstede2001}{}%
-15. Hofstede, G. \emph{Culture's consequences: Comparing values, behaviors, institutions and organizations across nations}. (Sage Publications, 2001).
-
-\leavevmode\hypertarget{ref-Inglehart2000}{}%
-16. Inglehart, R. \& Baker, W. E. Modernization, cultural change, and the persistence of traditional values. \emph{American Sociological Review} \textbf{65}, 19--51 (2000).
-
-\leavevmode\hypertarget{ref-Awad2018}{}%
-17. Awad, E. \emph{et al.} The moral machine experiment. \emph{Nature} \textbf{563}, 59--64 (2018).
-
-\leavevmode\hypertarget{ref-Rhoads2021}{}%
-18. Rhoads, S. A., Gunter, D., Ryan, R. M. \& Marsh, A. A. Global variation in subjective well-being predicts seven forms of altruism. \emph{Psychological Science} \textbf{32}, 1247--1261 (2021).
-
-\leavevmode\hypertarget{ref-Schulz2019}{}%
-19. Schulz, J. F., Bahrami-Rad, D., Beauchamp, J. P. \& Henrich, J. The Church, intensive kinship, and global psychological variation. \emph{Science} \textbf{366}, eaau5141 (2019).
-
-\leavevmode\hypertarget{ref-Thomson2018}{}%
-20.
-Thomson, R.
-\emph{et al.} Relational mobility predicts social behaviors in 39 countries and is tied to historical farming and threat.
-\emph{Proceedings of the National Academy of Sciences} \textbf{115}, 7521--7526 (2018).
-
-\leavevmode\hypertarget{ref-Kissling2008}{}%
-21. Kissling, W. D. \& Carl, G. Spatial autocorrelation and the selection of simultaneous autoregressive models. \emph{Global Ecology and Biogeography} \textbf{17}, 59--71 (2008).
-
-\leavevmode\hypertarget{ref-Hewlett2002}{}%
-22. Hewlett, B. S., DeSilvestri, A. \& Guglielmino, C. R. Semes and genes in africa. \emph{Current Anthropology} \textbf{43}, 313--321 (2002).
-
-\leavevmode\hypertarget{ref-Nunn2012}{}%
-23. Nunn, N. \& Puga, D. Ruggedness: The blessing of bad geography in Africa. \emph{The Review of Economics and Statistics} \textbf{94}, 20--36 (2012).
-
-\leavevmode\hypertarget{ref-Tobler1970}{}%
-24. Tobler, W. R. A computer movie simulating urban growth in the Detroit region. \emph{Economic Geography} \textbf{46}, 234--240 (1970).
-
-\leavevmode\hypertarget{ref-Tylor1889}{}%
-25.
-Tylor, E.
+\href{https://doi.org/10.1098/rspb.2008.0094}{Pathogen prevalence predicts human cross-cultural variability in individualism/collectivism}.
+\emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{275}, 1279--1285 (2008).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Gelfand2011}{}}%
+\CSLLeftMargin{14. }%
+\CSLRightInline{Gelfand, M.
+J.
+\emph{et al.} \href{https://doi.org/10.1126/science.1197754}{Differences between tight and loose cultures: A 33-nation study}.
+\emph{Science} \textbf{332}, 1100--1104 (2011).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Hofstede2001}{}}%
+\CSLLeftMargin{15. }%
+\CSLRightInline{Hofstede, G. \emph{Culture's consequences: Comparing values, behaviors, institutions and organizations across nations}. (Sage Publications, 2001).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Inglehart2000}{}}%
+\CSLLeftMargin{16. }%
+\CSLRightInline{Inglehart, R.
+\& Baker, W.
+E.
+\href{https://doi.org/10.2307/2657288}{Modernization, cultural change, and the persistence of traditional values}.
+\emph{American Sociological Review} \textbf{65}, 19--51 (2000).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Awad2018}{}}%
+\CSLLeftMargin{17. }%
+\CSLRightInline{Awad, E. \emph{et al.} \href{https://doi.org/10.1038/s41586-018-0637-6}{The moral machine experiment}. \emph{Nature} \textbf{563}, 59--64 (2018).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Rhoads2021}{}}%
+\CSLLeftMargin{18. }%
+\CSLRightInline{Rhoads, S.
+A., Gunter, D., Ryan, R.
+M.
+\& Marsh, A.
+A.
+\href{https://doi.org/10.1177/0956797621994767}{Global variation in subjective well-being predicts seven forms of altruism}.
+\emph{Psychological Science} \textbf{32}, 1247--1261 (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Schulz2019}{}}%
+\CSLLeftMargin{19. }%
+\CSLRightInline{Schulz, J.
+F., Bahrami-Rad, D., Beauchamp, J.
+P.
+\& Henrich, J.
+\href{https://doi.org/10.1126/science.aau5141}{The {C}hurch, intensive kinship, and global psychological variation}.
+\emph{Science} \textbf{366}, eaau5141 (2019).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Thomson2018}{}}%
+\CSLLeftMargin{20. }%
+\CSLRightInline{Thomson, R.
+\emph{et al.} \href{https://doi.org/10.1073/pnas.1713191115}{Relational mobility predicts social behaviors in 39 countries and is tied to historical farming and threat}.
+\emph{Proceedings of the National Academy of Sciences} \textbf{115}, 7521--7526 (2018).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Kissling2008}{}}%
+\CSLLeftMargin{21. }%
+\CSLRightInline{Kissling, W.
+D.
+\& Carl, G.
+\href{https://doi.org/10.1111/j.1466-8238.2007.00334.x}{Spatial autocorrelation and the selection of simultaneous autoregressive models}.
+\emph{Global Ecology and Biogeography} \textbf{17}, 59--71 (2008).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Hewlett2002}{}}%
+\CSLLeftMargin{22. }%
+\CSLRightInline{Hewlett, B. S., DeSilvestri, A. \& Guglielmino, C. R. \href{https://doi.org/10.1086/339379}{Semes and genes in africa}. \emph{Current Anthropology} \textbf{43}, 313--321 (2002).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Nunn2012}{}}%
+\CSLLeftMargin{23. }%
+\CSLRightInline{Nunn, N.
+\& Puga, D.
+\href{https://doi.org/10.1162/REST_a_00161}{Ruggedness: The blessing of bad geography in {A}frica}.
+\emph{The Review of Economics and Statistics} \textbf{94}, 20--36 (2012).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Tobler1970}{}}%
+\CSLLeftMargin{24. }%
+\CSLRightInline{Tobler, W. R. \href{https://doi.org/10.2307/143141}{A computer movie simulating urban growth in the {D}etroit region}. \emph{Economic Geography} \textbf{46}, 234--240 (1970).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Tylor1889}{}}%
+\CSLLeftMargin{25. }%
+\CSLRightInline{Tylor, E.
 B.
-On a method of investigating the development of institutions; applied to laws of marriage and descent.
-\emph{The Journal of the Anthropological Institute of Great Britain and Ireland} \textbf{18}, 245--272 (1889).
-
-\leavevmode\hypertarget{ref-Naroll1961}{}%
-26. Naroll, R. Two solutions to Galton's Problem. \emph{Philosophy of Science} \textbf{28}, 15--39 (1961).
-
-\leavevmode\hypertarget{ref-Naroll1965}{}%
-27. Naroll, R. Galton's Problem: The logic of cross-cultural analysis. \emph{Social Research} \textbf{32}, 428--451 (1965).
-
-\leavevmode\hypertarget{ref-Jarque1987}{}%
-28.
-Jarque, C.
+\href{https://doi.org/10.2307/2842423}{On a method of investigating the development of institutions; applied to laws of marriage and descent}.
+\emph{The Journal of the Anthropological Institute of Great Britain and Ireland} \textbf{18}, 245--272 (1889).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Naroll1961}{}}%
+\CSLLeftMargin{26. }%
+\CSLRightInline{Naroll, R. \href{https://doi.org/10.1086/287778}{Two solutions to {G}alton's {P}roblem}. \emph{Philosophy of Science} \textbf{28}, 15--39 (1961).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Naroll1965}{}}%
+\CSLLeftMargin{27. }%
+\CSLRightInline{Naroll, R. \href{http://www.jstor.org/stable/40969817}{Galton's {P}roblem: The logic of cross-cultural analysis}. \emph{Social Research} \textbf{32}, 428--451 (1965).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Jarque1987}{}}%
+\CSLLeftMargin{28. }%
+\CSLRightInline{Jarque, C.
 M.
@@
 K.
-A test for normality of observations and regression residuals.
-\emph{International Statistical Review / Revue Internationale de Statistique} \textbf{55}, 163--172 (1987).
-
-\leavevmode\hypertarget{ref-Legendre1993}{}%
-29. Legendre, P. Spatial autocorrelation: Trouble or new paradigm? \emph{Ecology} \textbf{74}, 1659--1673 (1993).
-
-\leavevmode\hypertarget{ref-Roberts2013}{}%
-30. Roberts, J., Seán AND Winters. Linguistic diversity and traffic accidents: Lessons from statistical studies of cultural traits. \emph{PLOS ONE} \textbf{8}, 1--13 (2013).
-
-\leavevmode\hypertarget{ref-Lichstein2002}{}%
-31. Lichstein, J. W., Simons, T. R., Shriner, S. A. \& Franzreb, K. E. Spatial autocorrelation and autoregressive models in ecology. \emph{Ecological Monographs} \textbf{72}, 445--463 (2002).
-
-\leavevmode\hypertarget{ref-Loftin1983}{}%
-32. Loftin, C. \& Ward, S. K. A spatial autocorrelation model of the effects of population density on fertility. \emph{American Sociological Review} \textbf{48}, 121--128 (1983).
-
-\leavevmode\hypertarget{ref-Murdock1969}{}%
-33. Murdock, G. P. \& White, D. R. Standard cross-cultural sample. \emph{Ethnology} \textbf{8}, 329--369 (1969).
-
-\leavevmode\hypertarget{ref-Dow2008}{}%
-34. Dow, M. M. \& Eff, E. A. Global, regional, and local network autocorrelation in the standard cross-cultural sample. \emph{Cross-Cultural Research} \textbf{42}, 148--171 (2008).
-
-\leavevmode\hypertarget{ref-Eff2004}{}%
-35. Eff, E. A. Does Mr. Galton still have a problem? Autocorrelation in the standard cross-cultural sample. \emph{World Cultures} \textbf{15}, 153--170 (2004).
-
-\leavevmode\hypertarget{ref-Symonds2014}{}%
-36.
-Symonds, M.
+\href{https://doi.org/10.2307/1403192}{A test for normality of observations and regression residuals}.
+\emph{International Statistical Review / Revue Internationale de Statistique} \textbf{55}, 163--172 (1987).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Legendre1993}{}}%
+\CSLLeftMargin{29. }%
+\CSLRightInline{Legendre, P. \href{https://doi.org/10.2307/1939924}{Spatial autocorrelation: Trouble or new paradigm?} \emph{Ecology} \textbf{74}, 1659--1673 (1993).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Roberts2013}{}}%
+\CSLLeftMargin{30. }%
+\CSLRightInline{Roberts, J., Seán AND Winters.
+\href{https://doi.org/10.1371/journal.pone.0070902}{Linguistic diversity and traffic accidents: Lessons from statistical studies of cultural traits}.
+\emph{PLOS ONE} \textbf{8}, 1--13 (2013).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Lichstein2002}{}}%
+\CSLLeftMargin{31. }%
+\CSLRightInline{Lichstein, J.
+W., Simons, T.
+R., Shriner, S.
+A.
+\& Franzreb, K.
+E.
+\href{https://doi.org/10.1890/0012-9615(2002)072\%5B0445:SAAAMI\%5D2.0.CO;2}{Spatial autocorrelation and autoregressive models in ecology}.
+\emph{Ecological Monographs} \textbf{72}, 445--463 (2002).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Loftin1983}{}}%
+\CSLLeftMargin{32. }%
+\CSLRightInline{Loftin, C.
+\& Ward, S.
+K.
+\href{https://doi.org/10.2307/2095150}{A spatial autocorrelation model of the effects of population density on fertility}.
+\emph{American Sociological Review} \textbf{48}, 121--128 (1983).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Ferguson1997}{}}%
+\CSLLeftMargin{33. }%
+\CSLRightInline{Ferguson, J. Anthropology and its evil twin. in (eds. Cooper, F. \& Packard, R.) 150--175 (University of California Press Berkeley, 1997).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Murdock1969}{}}%
+\CSLLeftMargin{34. }%
+\CSLRightInline{Murdock, G. P. \& White, D. R. \href{https://doi.org/10.2307/3772907}{Standard cross-cultural sample}. \emph{Ethnology} \textbf{8}, 329--369 (1969).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Dow2008}{}}%
+\CSLLeftMargin{35. }%
+\CSLRightInline{Dow, M.
+M.
+\& Eff, E.
+A.
+\href{https://doi.org/10.1177/1069397107311186}{Global, regional, and local network autocorrelation in the standard cross-cultural sample}.
+\emph{Cross-Cultural Research} \textbf{42}, 148--171 (2008).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Eff2004}{}}%
+\CSLLeftMargin{36. }%
+\CSLRightInline{Eff, E. A. Does {M}r. {G}alton still have a problem? Autocorrelation in the standard cross-cultural sample. \emph{World Cultures} \textbf{15}, 153--170 (2004).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Symonds2014}{}}%
+\CSLLeftMargin{37. }%
+\CSLRightInline{Symonds, M.
 R.
@@
 Z.) 105--130 (Springer Berlin Heidelberg, 2014).
-doi:\href{https://doi.org/10.1007/978-3-662-43550-2_5}{10.1007/978-3-662-43550-2\_5}.
-
-\leavevmode\hypertarget{ref-Watts2018}{}%
-37.
-Watts, J., Sheehan, O., Bulbulia, J., Gray, R.
+doi:\href{https://doi.org/10.1007/978-3-662-43550-2_5}{10.1007/978-3-662-43550-2\_5}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Watts2018}{}}%
+\CSLLeftMargin{38. }%
+\CSLRightInline{Watts, J., Sheehan, O., Bulbulia, J., Gray, R.
 D.
@@
 D.
-Christianity spread faster in small, politically structured societies.
-\emph{Nature Human Behaviour} \textbf{2}, 559--564 (2018).
-
-\leavevmode\hypertarget{ref-Atkinson2016}{}%
-38.
-Atkinson, Q.
+\href{https://doi.org/10.1038/s41562-018-0379-3}{Christianity spread faster in small, politically structured societies}.
+\emph{Nature Human Behaviour} \textbf{2}, 559--564 (2018).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Atkinson2016}{}}%
+\CSLLeftMargin{39. }%
+\CSLRightInline{Atkinson, Q.
 D., Coomber, T., Passmore, S., Greenhill, S.
@@
 \& Kushnick, G.
-Cultural and environmental predictors of pre-european deforestation on pacific islands.
-\emph{PLOS ONE} \textbf{11}, 1--15 (2016).
-
-\leavevmode\hypertarget{ref-Kelly2020}{}%
-39. Kelly, M. Understanding persistence. \emph{CEPR Discussion Paper No. DP15246} (2020).
-
-\leavevmode\hypertarget{ref-Bromham2018}{}%
-40.
-Bromham, L., Hua, X., Cardillo, M., Schneemann, H.
+\href{https://doi.org/10.1371/journal.pone.0156340}{Cultural and environmental predictors of pre-european deforestation on pacific islands}.
+\emph{PLOS ONE} \textbf{11}, 1--15 (2016).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Kelly2020}{}}%
+\CSLLeftMargin{40. }%
+\CSLRightInline{Kelly, M. \href{http://ssrn.com/abstract=3688200}{Understanding persistence}. \emph{CEPR Discussion Paper No. DP15246} (2020).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bromham2018}{}}%
+\CSLLeftMargin{41. }%
+\CSLRightInline{Bromham, L., Hua, X., Cardillo, M., Schneemann, H.
 \& Greenhill, S.
 J.
-Parasites and politics: Why cross-cultural studies must control for relatedness, proximity and covariation.
-\emph{Royal Society Open Science} \textbf{5}, 181100 (2018).
-
-\leavevmode\hypertarget{ref-Bromham2021}{}%
-41.
-Bromham, L., Skeels, A., Schneemann, H., Dinnage, R.
+\href{https://doi.org/10.1098/rsos.181100}{Parasites and politics: Why cross-cultural studies must control for relatedness, proximity and covariation}.
+\emph{Royal Society Open Science} \textbf{5}, 181100 (2018).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bromham2021}{}}%
+\CSLLeftMargin{42. }%
+\CSLRightInline{Bromham, L., Skeels, A., Schneemann, H., Dinnage, R.
 \& Hua, X.
-There is little evidence that spicy food in hot countries is an adaptation to reducing infection risk.
-\emph{Nature Human Behaviour} \textbf{5}, 878--891 (2021).
-
-\leavevmode\hypertarget{ref-Currie2012}{}%
-42. Currie, T. E. \& Mace, R. Analyses do not support the parasite-stress theory of human sociality. \emph{Behavioral and Brain Sciences} \textbf{35}, 83--85 (2012).
-
-\leavevmode\hypertarget{ref-Passmore2022}{}%
-43.
-Passmore, S.
+\href{https://doi.org/10.1038/s41562-020-01039-8}{There is little evidence that spicy food in hot countries is an adaptation to reducing infection risk}.
+\emph{Nature Human Behaviour} \textbf{5}, 878--891 (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Currie2012}{}}%
+\CSLLeftMargin{43. }%
+\CSLRightInline{Currie, T.
+E.
+\& Mace, R.
+\href{https://doi.org/10.1017/S0140525X11000963}{Analyses do not support the parasite-stress theory of human sociality}.
+\emph{Behavioral and Brain Sciences} \textbf{35}, 83--85 (2012).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Passmore2022}{}}%
+\CSLLeftMargin{44. }%
+\CSLRightInline{Passmore, S.
 \& Watts, J.
-WEIRD people and the Western Church: Who made whom?
-\emph{Religion, Brain \& Behavior} 1--58 (2022) doi:\href{https://doi.org/10.1080/2153599X.2021.1991459}{10.1080/2153599X.2021.1991459}.
-
-\leavevmode\hypertarget{ref-hdi}{}%
-44. \emph{Human development report}. \url{http://hdr.undp.org/en/composite/HDI}.
-
-\leavevmode\hypertarget{ref-Currie2009}{}%
-45. Currie, T. E. \& Mace, R. Political complexity predicts the spread of ethnolinguistic groups. \emph{Proceedings of the National Academy of Sciences} \textbf{106}, 7339--7344 (2009).
-
-\leavevmode\hypertarget{ref-Conley1999}{}%
-46. Conley, T. GMM estimation with cross sectional dependence. \emph{Journal of Econometrics} \textbf{92}, 1--45 (1999).
-
-\leavevmode\hypertarget{ref-Conley2010}{}%
-47. Conley, T. G. Spatial econometrics. in \emph{Microeconometrics} (eds. Durlauf, S. N. \& Blume, L. E.) 303--313 (Palgrave Macmillan, 2010).
-
-\leavevmode\hypertarget{ref-McElreath2020}{}%
-48. McElreath, R. \emph{Statistical rethinking: A Bayesian course with examples in R and Stan}. (CRC Press, 2020).
-
-\leavevmode\hypertarget{ref-Neal1998}{}%
-49.
-Neal, R.
+{WEIRD} people and the {W}estern {C}hurch: Who made whom?
+\emph{Religion, Brain \& Behavior} 1--58 (2022) doi:\href{https://doi.org/10.1080/2153599X.2021.1991459}{10.1080/2153599X.2021.1991459}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hdi}{}}%
+\CSLLeftMargin{45. }%
+\CSLRightInline{\emph{Human development report}. \url{http://hdr.undp.org/en/composite/HDI}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Schulz2022}{}}%
+\CSLLeftMargin{46. }%
+\CSLRightInline{Schulz, J. F. \href{https://doi.org/10.1093/ej/ueac027}{Kin networks and institutional development}. \emph{The Economic Journal} \textbf{132}, 2578--2613 (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Conley1999}{}}%
+\CSLLeftMargin{47. }%
+\CSLRightInline{Conley, T. G. \href{https://doi.org/10.1016/S0304-4076(98)00084-0}{{GMM} estimation with cross sectional dependence}. \emph{Journal of Econometrics} \textbf{92}, 1--45 (1999).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Conley2010}{}}%
+\CSLLeftMargin{48. }%
+\CSLRightInline{Conley, T. G. Spatial econometrics. in \emph{Microeconometrics} (eds. Durlauf, S. N. \& Blume, L. E.) 303--313 (Palgrave Macmillan, 2010).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Currie2009}{}}%
+\CSLLeftMargin{49. }%
+\CSLRightInline{Currie, T.
+E.
+\& Mace, R.
+\href{https://doi.org/10.1073/pnas.0804698106}{Political complexity predicts the spread of ethnolinguistic groups}.
+\emph{Proceedings of the National Academy of Sciences} \textbf{106}, 7339--7344 (2009).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-McElreath2020}{}}%
+\CSLLeftMargin{50. }%
+\CSLRightInline{McElreath, R. \emph{Statistical rethinking: A {B}ayesian course with examples in {R} and {Stan}}. (CRC Press, 2020).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Neal1998}{}}%
+\CSLLeftMargin{51. }%
+\CSLRightInline{Neal, R.
 M.
-Regression and classification using Gaussian process priors.
+Regression and classification using {Gaussian} process priors.
 in \emph{Bayesian statistics} (eds.
@@
 M.) vol.
-6 475--501 (Oxford University Press).
-
-\leavevmode\hypertarget{ref-deVillemereuil2014}{}%
-50.
-Villemereuil, P.
+6 475--501 (Oxford University Press).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-deVillemereuil2014}{}}%
+\CSLLeftMargin{52. }%
+\CSLRightInline{Villemereuil, P.
 de \& Nakagawa, S.
@@
 Z.) 287--303 (Springer Berlin Heidelberg, 2014).
-doi:\href{https://doi.org/10.1007/978-3-662-43550-2_11}{10.1007/978-3-662-43550-2\_11}.
-
-\leavevmode\hypertarget{ref-Beck2003}{}%
-51. Beck, T., Demirgäç-Kunt, A. \& Levine, R. Law, endowments, and finance. \emph{Journal of Financial Economics} \textbf{70}, 137--181 (2003).
-
-\leavevmode\hypertarget{ref-Beck2005}{}%
-52. Beck, T., Demirgäç-Kunt, A. \& Levine, R. SMEs, growth, and poverty: Cross-country evidence. \emph{Journal of economic growth} \textbf{10}, 199--229 (2005).
-
-\leavevmode\hypertarget{ref-Bockstette2002}{}%
-53. Bockstette, V., Chanda, A. \& Putterman, L. States and markets: The advantage of an early start. \emph{Journal of Economic growth} \textbf{7}, 347--369 (2002).
-
-\leavevmode\hypertarget{ref-Easterly2003}{}%
-54. Easterly, W. \& Levine, R. Tropics, germs, and crops: How endowments influence economic development. \emph{Journal of Monetary Economics} \textbf{50}, 3--39 (2003).
-
-\leavevmode\hypertarget{ref-Easterly2007}{}%
-55. Easterly, W. Inequality does cause underdevelopment: Insights from a new instrument. \emph{Journal of Development Economics} \textbf{84}, 755--776 (2007).
-
-\leavevmode\hypertarget{ref-Skidmore2002}{}%
-56. Skidmore, M. \& Toya, H. Do natural disasters promote long-run growth? \emph{Economic Inquiry} \textbf{40}, 664--687 (2002).
-
-\leavevmode\hypertarget{ref-Adamczyk2009}{}%
-57. Adamczyk, A. \& Pitt, C. Shaping attitudes about homosexuality: The role of religion and cultural context. \emph{Social Science Research} \textbf{38}, 338--351 (2009).
-
-\leavevmode\hypertarget{ref-Alesina2013}{}%
-58. Alesina, A., Giuliano, P. \& Nunn, N. On the origins of gender roles: Women and the plough. \emph{Quarterly Journal of Economics} \textbf{128}, 469--530 (2013).
-
-\leavevmode\hypertarget{ref-Knack1997}{}%
-59. Knack, S. \& Keefer, P. Does social capital have an economic payoff? A cross-country investigation. \emph{The Quarterly Journal of Economics} \textbf{112}, 1251--1288 (1997).
-
-\leavevmode\hypertarget{ref-Sookias2018}{}%
-60. Sookias, R. B., Passmore, S. \& Atkinson, Q. D. Deep cultural ancestry and human development indicators across nation states. \emph{Royal Society Open Science} \textbf{5}, 171411 (2018).
-
-\leavevmode\hypertarget{ref-Smith2017}{}%
-61.
-Smith, M.
+doi:\href{https://doi.org/10.1007/978-3-662-43550-2_11}{10.1007/978-3-662-43550-2\_11}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Beck2003}{}}%
+\CSLLeftMargin{53. }%
+\CSLRightInline{Beck, T., Demirgäç-Kunt, A.
+\& Levine, R.
+\href{https://doi.org/10.1016/S0304-405X(03)00144-2}{Law, endowments, and finance}.
+\emph{Journal of Financial Economics} \textbf{70}, 137--181 (2003).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Beck2005}{}}%
+\CSLLeftMargin{54. }%
+\CSLRightInline{Beck, T., Demirgäç-Kunt, A. \& Levine, R. {SME}s, growth, and poverty: Cross-country evidence. \emph{Journal of economic growth} \textbf{10}, 199--229 (2005).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bockstette2002}{}}%
+\CSLLeftMargin{55. }%
+\CSLRightInline{Bockstette, V., Chanda, A. \& Putterman, L. States and markets: The advantage of an early start. \emph{Journal of Economic growth} \textbf{7}, 347--369 (2002).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Easterly2003}{}}%
+\CSLLeftMargin{56. }%
+\CSLRightInline{Easterly, W.
+\& Levine, R.
+\href{https://doi.org/10.1016/S0304-3932(02)00200-3}{Tropics, germs, and crops: How endowments influence economic development}.
+\emph{Journal of Monetary Economics} \textbf{50}, 3--39 (2003).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Easterly2007}{}}%
+\CSLLeftMargin{57. }%
+\CSLRightInline{Easterly, W.
+\href{https://doi.org/10.1016/j.jdeveco.2006.11.002}{Inequality does cause underdevelopment: Insights from a new instrument}.
+\emph{Journal of Development Economics} \textbf{84}, 755--776 (2007).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Skidmore2002}{}}%
+\CSLLeftMargin{58. }%
+\CSLRightInline{Skidmore, M. \& Toya, H. \href{https://doi.org/10.1093/ei/40.4.664}{Do natural disasters promote long-run growth?} \emph{Economic Inquiry} \textbf{40}, 664--687 (2002).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Adamczyk2009}{}}%
+\CSLLeftMargin{59. }%
+\CSLRightInline{Adamczyk, A.
+\& Pitt, C.
+\href{https://doi.org/10.1016/j.ssresearch.2009.01.002}{Shaping attitudes about homosexuality: The role of religion and cultural context}.
+\emph{Social Science Research} \textbf{38}, 338--351 (2009).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Alesina2013}{}}%
+\CSLLeftMargin{60. }%
+\CSLRightInline{Alesina, A., Giuliano, P. \& Nunn, N. On the origins of gender roles: Women and the plough. \emph{Quarterly Journal of Economics} \textbf{128}, 469--530 (2013).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Knack1997}{}}%
+\CSLLeftMargin{61. }%
+\CSLRightInline{Knack, S.
+\& Keefer, P.
+\href{https://doi.org/10.1162/003355300555475}{Does social capital have an economic payoff?
+A cross-country investigation}.
+\emph{The Quarterly Journal of Economics} \textbf{112}, 1251--1288 (1997).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Sookias2018}{}}%
+\CSLLeftMargin{62. }%
+\CSLRightInline{Sookias, R.
+B., Passmore, S.
+\& Atkinson, Q.
+D.
+\href{https://doi.org/10.1098/rsos.171411}{Deep cultural ancestry and human development indicators across nation states}.
+\emph{Royal Society Open Science} \textbf{5}, 171411 (2018).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Smith2017}{}}%
+\CSLLeftMargin{63. }%
+\CSLRightInline{Smith, M.
 D., Rabbitt, M.
@@
 \& Coleman- Jensen, A.
-Who are the world's food insecure?
-New evidence from the Food and Agriculture Organization's Food Insecurity Experience Scale.
-\emph{World Development} \textbf{93}, 402--412 (2017).
-
-\leavevmode\hypertarget{ref-Hastie2017}{}%
-62. Hastie, T. J. \& Tibshirani, R. J. \emph{Generalized additive models}. (Routledge, 2017).
-
-\leavevmode\hypertarget{ref-Smaldino2016}{}%
-63. Smaldino, P. E. \& McElreath, R. The natural selection of bad science. \emph{Royal Society Open Science} \textbf{3}, 160384 (2016).
-
-\leavevmode\hypertarget{ref-Liddell2018}{}%
-64. Liddell, T. M. \& Kruschke, J. K. Analyzing ordinal data with metric models: What could possibly go wrong? \emph{Journal of Experimental Social Psychology} \textbf{79}, 328--348 (2018).
-
-\leavevmode\hypertarget{ref-Nicholson2017}{}%
-65.
-Nicholson, J.
+\href{https://doi.org/10.1016/j.worlddev.2017.01.006}{Who are the world's food insecure?
+New evidence from the {F}ood and {A}griculture {O}rganization's {F}ood {I}nsecurity {E}xperience {S}cale}.
+\emph{World Development} \textbf{93}, 402--412 (2017).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Hastie2017}{}}%
+\CSLLeftMargin{64. }%
+\CSLRightInline{Hastie, T. J. \& Tibshirani, R. J. \emph{Generalized additive models}. (Routledge, 2017).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Smaldino2016}{}}%
+\CSLLeftMargin{65. }%
+\CSLRightInline{Smaldino, P. E. \& McElreath, R. \href{https://doi.org/10.1098/rsos.160384}{The natural selection of bad science}. \emph{Royal Society Open Science} \textbf{3}, 160384 (2016).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Liddell2018}{}}%
+\CSLLeftMargin{66. }%
+\CSLRightInline{Liddell, T.
+M.
+\& Kruschke, J.
+K.
+\href{https://doi.org/10.1016/j.jesp.2018.08.009}{Analyzing ordinal data with metric models: What could possibly go wrong?} \emph{Journal of Experimental Social Psychology} \textbf{79}, 328--348 (2018).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Nicholson2017}{}}%
+\CSLLeftMargin{67. }%
+\CSLRightInline{Nicholson, J.
 S., Deboeck, P.
@@
 \& Howard, W.
-Attrition in developmental psychology: A review of modern missing data reporting and practices.
-\emph{International Journal of Behavioral Development} \textbf{41}, 143--153 (2017).
-
-\leavevmode\hypertarget{ref-Tipton2019}{}%
-66. Tipton, E., Pustejovsky, J. E. \& Ahmadi, H. Current practices in meta-regression in psychology, education, and medicine. \emph{Research Synthesis Methods} \textbf{10}, 180--194 (2019).
-
-\leavevmode\hypertarget{ref-Chambers2021}{}%
-67. Chambers, C. D. \& Tzavella, L. The past, present and future of Registered Reports. \emph{Nature Human Behaviour} \textbf{6}, 29--42 (2021).
-
-\leavevmode\hypertarget{ref-rstan}{}%
-68. Stan Development Team. RStan: The R interface to Stan. (2020).
-
-\leavevmode\hypertarget{ref-Burkner2017}{}%
-69. Bürkner, P.-C. brms: An R package for Bayesian multilevel models using Stan. \emph{Journal of Statistical Software} \textbf{80}, 1--28 (2017).
-
-\leavevmode\hypertarget{ref-Oishi2013}{}%
-70. Oishi, S. Socioecological psychology. \emph{Annual Review of Psychology} \textbf{65}, 581--609 (2014).
-
-\leavevmode\hypertarget{ref-Tiefelsdorf2007}{}%
-71.
-Tiefelsdorf, M.
+\href{https://doi.org/10.1177/0165025415618275}{Attrition in developmental psychology: A review of modern missing data reporting and practices}.
+\emph{International Journal of Behavioral Development} \textbf{41}, 143--153 (2017).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Tipton2019}{}}%
+\CSLLeftMargin{68. }%
+\CSLRightInline{Tipton, E., Pustejovsky, J.
+E.
+\& Ahmadi, H.
+\href{https://doi.org/10.1002/jrsm.1339}{Current practices in meta-regression in psychology, education, and medicine}.
+\emph{Research Synthesis Methods} \textbf{10}, 180--194 (2019).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Chambers2021}{}}%
+\CSLLeftMargin{69. }%
+\CSLRightInline{Chambers, C. D. \& Tzavella, L. The past, present and future of {Registered Reports}. \emph{Nature Human Behaviour} \textbf{6}, 29--42 (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-rstan}{}}%
+\CSLLeftMargin{70. }%
+\CSLRightInline{Stan Development Team. \href{http://mc-stan.org/}{{RStan}: The {R} interface to {Stan}}. (2020).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Burkner2017}{}}%
+\CSLLeftMargin{71. }%
+\CSLRightInline{Bürkner, P.-C.
+\href{https://doi.org/10.18637/jss.v080.i01}{{brms}: An {R} package for {Bayesian} multilevel models using {Stan}}.
+\emph{Journal of Statistical Software} \textbf{80}, 1--28 (2017).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Oishi2013}{}}%
+\CSLLeftMargin{72. }%
+\CSLRightInline{Oishi, S. \href{https://doi.org/10.1146/annurev-psych-030413-152156}{Socioecological psychology}. \emph{Annual Review of Psychology} \textbf{65}, 581--609 (2014).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Tiefelsdorf2007}{}}%
+\CSLLeftMargin{73. }%
+\CSLRightInline{Tiefelsdorf, M.
 \& Griffith, D.
 A.
-Semiparametric filtering of spatial autocorrelation: The eigenvector approach.
-\emph{Environment and Planning A: Economy and Space} \textbf{39}, 1193--1221 (2007).
-
-\leavevmode\hypertarget{ref-DeHoyos2006}{}%
-72. De Hoyos, R. E. \& Sarafidis, V. Testing for cross-sectional dependence in panel-data models. \emph{The Stata Journal} \textbf{6}, 482--496 (2006).
-
-\leavevmode\hypertarget{ref-Duben2021}{}%
-73. Düben, C. \emph{conleyreg: Estimations using Conley standard errors}. (2021).
-
-\leavevmode\hypertarget{ref-Solin2020}{}%
-74. Solin, A. \& Särkkä, S. Hilbert space methods for reduced-rank Gaussian process regression. \emph{Statistics and Computing} \textbf{30}, 419--446 (2020).
-
-\leavevmode\hypertarget{ref-Landau2021}{}%
-75.
-Landau, W.
+\href{https://doi.org/10.1068/a37378}{Semiparametric filtering of spatial autocorrelation: The eigenvector approach}.
+\emph{Environment and Planning A: Economy and Space} \textbf{39}, 1193--1221 (2007).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-DeHoyos2006}{}}%
+\CSLLeftMargin{74. }%
+\CSLRightInline{De Hoyos, R.
+E.
+\& Sarafidis, V.
+\href{https://doi.org/10.1177/1536867X0600600403}{Testing for cross-sectional dependence in panel-data models}.
+\emph{The Stata Journal} \textbf{6}, 482--496 (2006).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Duben2021}{}}%
+\CSLLeftMargin{75. }%
+\CSLRightInline{Düben, C. \emph{\href{https://CRAN.R-project.org/package=conleyreg}{{conleyreg}: Estimations using {C}onley standard errors}}. (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Solin2020}{}}%
+\CSLLeftMargin{76. }%
+\CSLRightInline{Solin, A. \& Särkkä, S. Hilbert space methods for reduced-rank {G}aussian process regression. \emph{Statistics and Computing} \textbf{30}, 419--446 (2020).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
+\CSLLeftMargin{77. }%
+\CSLRightInline{Landau, W.
 M.
-The targets R package: A dynamic Make-like function-oriented pipeline toolkit for reproducibility and high-performance computing.
-\emph{Journal of Open Source Software} \textbf{6}, 2959 (2021).
-
-\leavevmode\hypertarget{ref-Aust2020}{}%
-76. Aust, F. \& Barth, M. \emph{papaja: Create APA manuscripts with R Markdown}. (2020).
+\href{https://doi.org/10.21105/joss.02959}{The targets {R} package: A dynamic {M}ake-like function-oriented pipeline toolkit for reproducibility and high-performance computing}.
+\emph{Journal of Open Source Software} \textbf{6}, 2959 (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Aust2020}{}}%
+\CSLLeftMargin{78. }%
+\CSLRightInline{Aust, F. \& Barth, M. \emph{\href{https://github.com/crsh/papaja}{{papaja}: {Create} {APA} manuscripts with {R Markdown}}}. (2020).}
+
+\end{CSLReferences}
 
@@
 
-\hypertarget{supplementary-results}{%
-\subsection{Supplementary Results}\label{supplementary-results}}
-
-\hypertarget{geographic-and-cultural-phylogenetic-signal-estimates-for-human-development-index-and-ingleharts-value-dimensions}{%
-\subsubsection{Geographic and cultural phylogenetic signal estimates for Human Development Index and Inglehart's value dimensions}\label{geographic-and-cultural-phylogenetic-signal-estimates-for-human-development-index-and-ingleharts-value-dimensions}}
-
-Controlling for shared cultural ancestry, the proportion of national-level variance explained by spatial proximity was 0.37 for the Human Development Index (95\% credible interval {[}0.19 0.60{]}; \(\text{BF}_{\neq0}\) \textgreater{} 100), 0.43 for traditional values (95\% CI {[}0.19 0.73{]}; \(\text{BF}_{\neq0}\) \textgreater{} 100), and 0.18 for survival values (95\% CI {[}0.01 0.44{]}; \(\text{BF}_{\neq0}\) = 1.77).
-Controlling for spatial proximity, the proportion of variance explained by shared cultural ancestry was 0.61 for the Human Development Index (95\% CI {[}0.39 0.80{]}; \(\text{BF}_{\neq0}\) \textgreater{} 100), 0.55 for traditional values (95\% CI {[}0.26 0.79{]}; \(\text{BF}_{\neq0}\) \textgreater{} 100), and 0.79 for survival values (95\% CI {[}0.54 0.96{]}; \(\text{BF}_{\neq0}\) \textgreater{} 100).
-
-\newpage
-
 \hypertarget{supplementary-figures}{%
@@
 
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/dag-1.pdf}
-\caption{\label{fig:dag}\emph{A causal directed acyclic graph of spatial and cultural phylogenetic non-independence in cross-national studies.} We are interested in estimating the direct effect of national-level exposure \(X\) on national-level outcome \(Y\).
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/dag-1} \caption{\emph{A causal directed acyclic graph of spatial and cultural phylogenetic non-independence in cross-national studies.} We are interested in estimating the direct effect of national-level exposure \(X\) on national-level outcome \(Y\).
 But these variables are confounded by their common unobserved cause \(U\).
@@
 We can thus use \(G\) and \(L\) to model the covariation between \(X\) and \(Y\) induced by \(U\).
-Failing to do this and simply estimating the bivariate correlation between \(X\) and \(Y\) will produce spurious relationships and residuals that are spatially and culturally non-independent around the world.}
+Failing to do this and simply estimating the bivariate correlation between \(X\) and \(Y\) will produce spurious relationships and residuals that are spatially and culturally non-independent around the world.}\label{fig:dag}
 \end{figure}
@@
 
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/plotSim1-1.pdf}
-\caption{\label{fig:plotSim1}\emph{Distribution of cross-national correlations from simulation study under strong spatial autocorrelation.} In these simulations, the strength of spatial autocorrelation is set to 0.8 for both outcome and predictor variables.
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotReview2-1} \caption{\emph{Analysis-level results from systematic review of 100 highly-cited cross-national studies of economic development (red) and cultural values (blue).} (a) Proportion of analyses accounting for non-independence, overall and split by common methods of controlling for non-independence.
+(b) The association between journal impact factor and the probability that an analysis accounts for non-independence.
+(c) Estimated trend over time for the probability that an analysis accounts for non-independence.
+Point ranges represent estimated proportions and 95\% credible intervals.
+Lines and shaded areas are posterior median regression lines and 50\% credible intervals from Bayesian multilevel models.
+Region FEs = region fixed effects.}\label{fig:plotReview2}
+\end{figure}
+
+\newpage
+
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotSim1-1} \caption{\emph{Distribution of cross-national correlations from simulation study under strong spatial autocorrelation.} In these simulations, the strength of spatial autocorrelation is set to 0.8 for both outcome and predictor variables.
 For frequentist regression models, point ranges represent correlation estimates and 95\% confidence intervals.
@@
 Correlations are ordered by effect size independently in each panel.
-Red point ranges indicate that the slope is \enquote{significant} (i.e.~the 95\% confidence / credible interval excludes zero).
-Black point ranges indicate that the slope is \enquote{not significant}.}
+Red point ranges indicate that the slope is ``significant'' (i.e.~the 95\% confidence / credible interval excludes zero).
+Black point ranges indicate that the slope is ``not significant''.}\label{fig:plotSim1}
 \end{figure}
@@
 
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/plotSim2-1.pdf}
-\caption{\label{fig:plotSim2}\emph{Distribution of cross-national correlations from simulation study under strong cultural phylogenetic autocorrelation.} In these simulations, the strength of cultural phylogenetic autocorrelation is set to 0.8 for both outcome and predictor variables.
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotSim2-1} \caption{\emph{Distribution of cross-national correlations from simulation study under strong cultural phylogenetic autocorrelation.} In these simulations, the strength of cultural phylogenetic autocorrelation is set to 0.8 for both outcome and predictor variables.
 For frequentist regression models, point ranges represent correlation estimates and 95\% confidence intervals.
@@
 Correlations are ordered by effect size independently in each panel.
-Red point ranges indicate that the slope is \enquote{significant} (i.e.~the 95\% confidence / credible interval excludes zero).
-Black point ranges indicate that the slope is \enquote{not significant}.}
+Red point ranges indicate that the slope is ``significant'' (i.e.~the 95\% confidence / credible interval excludes zero).
+Black point ranges indicate that the slope is ``not significant''.}\label{fig:plotSim2}
 \end{figure}
@@
 
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/plotReplications2-1.pdf}
-\caption{\label{fig:plotReplications2}\emph{Reanalysis models fitted to raw data, for economic development (a) and cultural values (b) studies.} Data points are labelled using ISO 3166-1 alpha-2 letter country codes.
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotReplications2-1} \caption{\emph{Reanalysis models fitted to raw data, for economic development (a) and cultural values (b) studies.} Data points are labelled using ISO 3166-1 alpha-2 letter country codes.
 In all reanalyses, outcome and predictor variables are standardised, making regression slopes comparable to Pearson's correlation coefficients.
 Lines and shaded areas represent posterior median regression lines and 95\% credible intervals.
-For models with covariates (Adamzyck and Pitt 2009; Gelfand et al.~2011), marginal effects are presented holding all covariates at zero or their reference categories.}
+For models with covariates (Adamzyck and Pitt 2009; Gelfand et al.~2011), marginal effects are presented holding all covariates at zero or their reference categories.}\label{fig:plotReplications2}
 \end{figure}
@@
 
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/plotReplications3-1.pdf}
-\caption{\label{fig:plotReplications3}\emph{Posterior estimates of Gaussian process functions mapping spatial autocorrelation onto geographic distance from our reanalyses of economic development (a) and cultural values (b) studies.} Estimates are from models additionally controlling for cultural phylogenetic non-independence.
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotReplications3-1} \caption{\emph{Posterior estimates of Gaussian process functions mapping spatial autocorrelation onto geographic distance from our reanalyses of economic development (a) and cultural values (b) studies.} Estimates are from models additionally controlling for cultural phylogenetic non-independence.
 The y-axis represents the amount of spatial autocorrelation between data points with increasing distance between those points on the x-axis (logged distance in kilometres).
-Lines and shaded areas represent median posterior spatial autocorrelation functions and 50\% and 95\% credible intervals.}
+Lines and shaded areas represent median posterior spatial autocorrelation functions and 50\% and 95\% credible intervals.}\label{fig:plotReplications3}
 \end{figure}
@@
 
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/plotReplications4-1.pdf}
-\caption{\label{fig:plotReplications4}\emph{Posterior estimates of cultural phylogenetic signal from our reanalyses.} Estimates are from models additionally controlling for spatial non-independence.
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotReplications4-1} \caption{\emph{Posterior estimates of cultural phylogenetic signal from our reanalyses.} Estimates are from models additionally controlling for spatial non-independence.
 Cultural phylogenetic signal is operationalised as the proportion of national-level variance explained by linguistic proximity between nations.
-Ridges are full posterior distributions, and points are posterior medians, and lines represent 50\% and 95\% credible intervals.}
+Ridges are full posterior distributions, and points are posterior medians, and lines represent 50\% and 95\% credible intervals.}\label{fig:plotReplications4}
 \end{figure}
+
+\newpage
+
+\hypertarget{supplementary-tables}{%
+\subsection{Supplementary Tables}\label{supplementary-tables}}
+
+\begin{table}[H]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:tableSignal}Geographic and cultural phylogenetic signal results for economic development and cultural values variables.
+Signal estimates reflect the proportion of national-level variance explained by geographic and linguistic covariance matrices.
+Bayes Factors (BF) reflect support for the hypothesis that the signal estimate differs from zero.
+HDI = Human Development Index; GDPpc = gross domestic product per capita.}
+
+\small{
+
+\begin{tabular}{lll}
+\toprule
+Outcome & \multicolumn{1}{c}{Geographic signal} & \multicolumn{1}{c}{Cultural phylogenetic signal}\\
+\midrule
+HDI & 0.37, 95\% CI [0.18, 0.59], BF > 100 & 0.62, 95\% CI [0.40, 0.80], BF > 100\\
+GDPpc & 0.42, 95\% CI [0.20, 0.66], BF > 100 & 0.56, 95\% CI [0.33, 0.78], BF > 100\\
+GDPpc growth & 0.65, 95\% CI [0.09, 0.98], BF = 16.79 & 0.26, 95\% CI [0.00, 0.70], BF = 1.16\\
+Gini index & 0.74, 95\% CI [0.48, 0.97], BF > 100 & 0.25, 95\% CI [0.02, 0.51], BF = 3.97\\
+Traditional values & 0.44, 95\% CI [0.17, 0.76], BF > 100 & 0.54, 95\% CI [0.23, 0.79], BF > 100\\
+Survival values & 0.20, 95\% CI [0.01, 0.45], BF = 1.96 & 0.78, 95\% CI [0.53, 0.96], BF > 100\\
+Tightness & 0.09, 95\% CI [0.00, 0.32], BF = 0.33 & 0.84, 95\% CI [0.61, 0.97], BF > 100\\
+Individualism & 0.26, 95\% CI [0.00, 0.66], BF = 1.97 & 0.69, 95\% CI [0.29, 0.96], BF > 100\\
+\bottomrule
+\end{tabular}
+
+}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\newpage
+
+\begin{center}
+\begin{ThreePartTable}
+
+\small{
+
+\begin{longtable}{m{2cm}m{11cm}m{2cm}}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:tableArticlesReview}List of 100 papers included in literature review, sorted by annual rate of citations since publication.}\\
+\toprule
+Review & \multicolumn{1}{c}{Reference} & \multicolumn{1}{c}{Citations per year}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:tableArticlesReview} continued}}\\
+\toprule
+Review & \multicolumn{1}{c}{Reference} & \multicolumn{1}{c}{Citations per year}\\
+\midrule
+\endhead
+Cultural values & Knack, S., \& Keefer, P. (1997). Does social capital have an economic payoff? A cross-country investigation. The Quarterly journal of economics, 112(4), 1251-1288. & 121.50\\
+ & Inglehart, R., \& Baker, W. E. (2000). Modernization, cultural change, and the persistence of traditional values. American sociological review, 19-51. & 119.33\\
+ & Gelfand, M.
+J., Raver, J.
+L., Nishii, L., Leslie, L.
+M., Lun, J., Lim, B.
+C., ...
+\& Yamaguchi, S.
+(2011).
+Differences between tight and loose cultures: A 33-nation study. science, 332(6033), 1100-1104.
+& 91.33\\
+ & Alesina, A., Giuliano, P., \& Nunn, N. (2013). On the origins of gender roles: Women and the plough. The quarterly journal of economics, 128(2), 469-530. & 57.12\\
+ & Schmitt, D.
+P., Realo, A., Voracek, M., \& Allik, J.
+(2008).
+Why can't a man be more like a woman?
+Sex differences in Big Five personality traits across 55 cultures. Journal of personality and social psychology, 94(1), 168.
+& 49.15\\
+ & Hofstede, G., \& McCrae, R. R. (2004). Personality and culture revisited: Linking traits and dimensions of culture. Cross-cultural research, 38(1), 52-88. & 35.53\\
+ & Delhey, J., Newton, K., \& Welzel, C. (2011). How general is trust in “most people”? Solving the radius of trust problem. American Sociological Review, 76(5), 786-807. & 35.10\\
+ & Fincher, C.
+L., Thornhill, R., Murray, D.
+R., \& Schaller, M.
+(2008).
+Pathogen prevalence predicts human cross-cultural variability in individualism/collectivism. Proceedings of the Royal Society B: Biological Sciences, 275(1640), 1279-1285.
+& 33.38\\
+ & Schneider, S. L. (2008). Anti-immigrant attitudes in Europe: Outgroup size and perceived ethnic threat. European Sociological Review, 24(1), 53-67. & 30.54\\
+ & Franzen, A., \& Meyer, R. (2010). Environmental attitudes in cross-national perspective: A multilevel analysis of the ISSP 1993 and 2000. European sociological review, 26(2), 219-234. & 30.27\\
+ & Steenkamp, J.
+B.
+E., Ter Hofstede, F., \& Wedel, M.
+(1999).
+A cross-national investigation into the individual and national cultural antecedents of consumer innovativeness. Journal of marketing, 63(2), 55-69.
+& 29.32\\
+ & Paxton, P. (2002). Social capital and democracy: An interdependent relationship. American sociological review, 254-277. & 29.21\\
+ & Santos, H. C., Varnum, M. E., \& Grossmann, I. (2017). Global increases in individualism. Psychological science, 28(9), 1228-1239. & 28.25\\
+ & Van Zanten, B.
+T., Van Berkel, D.
+B., Meentemeyer, R.
+K., Smith, J.
+W., Tieskens, K.
+F., \& Verburg, P.
+H.
+(2016).
+Continental-scale quantification of landscape values using social media data. Proceedings of the National Academy of Sciences, 113(46), 12974-12979.
+& 27.00\\
+ & Beugelsdijk, S., \& Welzel, C. (2018). Dimensions and dynamics of national culture: Synthesizing Hofstede with Inglehart. Journal of cross-cultural psychology, 49(10), 1469-1505. & 24.00\\
+ & Stephan, U., \& Uhlaner, L.
+M.
+(2010).
+Performance-based vs socially supportive culture: A cross-national study of descriptive norms and entrepreneurship. Journal of International Business Studies, 41(8), 1347-1364.
+& 23.91\\
+ & Dakhli, M., \& De Clercq, D. (2004). Human capital, social capital, and innovation: a multi-country study. Entrepreneurship \& regional development, 16(2), 107-128. & 23.53\\
+ & Adamczyk, A., \& Pitt, C. (2009). Shaping attitudes about homosexuality: The role of religion and cultural context. Social Science Research, 38(2), 338-351. & 21.75\\
+ & Mayda, A. M., \& Rodrik, D. (2005). Why are some people (and countries) more protectionist than others?. European Economic Review, 49(6), 1393-1430. & 21.69\\
+ & Johnson, T., Kulesa, P., Cho, Y.
+I., \& Shavitt, S.
+(2005).
+The relation between culture and response styles: Evidence from 19 countries. Journal of Cross-cultural psychology, 36(2), 264-277.
+& 20.62\\
+ & Tam, K.
+P., \& Chan, H.
+W.
+(2018).
+Generalized trust narrows the gap between environmental concern and pro-environmental behavior: Multilevel evidence. Global Environmental Change, 48, 182-194.
+& 20.00\\
+ & Kuppens, P., Realo, A., \& Diener, E.
+(2008).
+The role of positive and negative emotions in life satisfaction judgment across nations. Journal of personality and social psychology, 95(1), 66.
+& 19.62\\
+ & Tam, K.
+P., \& Chan, H.
+W.
+(2017).
+Environmental concern has a weaker association with pro-environmental behavior in some societies than others: A cross-cultural psychology perspective. Journal of Environmental Psychology, 53, 213-223.
+& 18.75\\
+ & Miska, C., Szőcs, I., \& Schiffinger, M. (2018). Culture’s effects on corporate sustainability practices: A multi-domain and multi-level view. Journal of World Business, 53(2), 263-279. & 17.67\\
+ & Oishi, S., Diener, E., Lucas, R.
+E., \& Suh, E.
+M.
+(2009).
+Cross-cultural variations in predictors of life satisfaction: Perspectives from needs and values.
+In Culture and well-being (pp.
+109-127).
+Springer, Dordrecht.
+& 16.00\\
+ & Malka, A., Soto, C.
+J., Inzlicht, M., \& Lelkes, Y.
+(2014).
+Do needs for security and certainty predict cultural and economic conservatism?
+A cross-national analysis. Journal of personality and social psychology, 106(6), 1031.
+& 16.00\\
+ & Eom, K., Kim, H.
+S., Sherman, D.
+K., \& Ishii, K.
+(2016).
+Cultural variability in the link between environmental concern and support for environmental action. Psychological Science, 27(10), 1331-1339.
+& 15.80\\
+ & Larsen, C. A. (2008). The institutional logic of welfare attitudes: How welfare regimes influence public support. Comparative political studies, 41(2), 145-168. & 15.62\\
+ & Sortheix, F. M., \& Schwartz, S. H. (2017). Values that underlie and undermine well–being: Variability across countries. European Journal of Personality, 31(2), 187-201. & 15.50\\
+ & Van der Meer, T., \& Hakhverdian, A.
+(2017).
+Political trust as the evaluation of process and performance: A cross-national study of 42 European countries. Political Studies, 65(1), 81-102.
+& 15.25\\
+ & Resick, C.
+J., Hanges, P.
+J., Dickson, M.
+W., \& Mitchelson, J.
+K.
+(2006).
+A cross-cultural examination of the endorsement of ethical leadership. Journal of Business Ethics, 63(4), 345-359.
+& 15.20\\
+ & Murray, D.
+R., \& Schaller, M.
+(2010).
+Historical prevalence of infectious diseases within 230 geopolitical regions: A tool for investigating origins of culture. Journal of Cross-Cultural Psychology, 41(1), 99-108.
+& 15.00\\
+ & Welzel, C., Inglehart, R., \& Kligemann, H. D. (2003). The theory of human development: A cross‐cultural analysis. European Journal of Political Research, 42(3), 341-379. & 14.94\\
+ & Löckenhoff, C.
+E., De Fruyt, F., Terracciano, A., McCrae, R.
+R., De Bolle, M., Costa, P.
+T., ...
+\& Yik, M.
+(2009).
+Perceptions of aging across 26 cultures and their culture-level associates. Psychology and aging, 24(4), 941.
+& 14.75\\
+ & Fulmer, C.
+A., Gelfand, M.
+J., Kruglanski, A.
+W., Kim-Prieto, C., Diener, E., Pierro, A., \& Higgins, E.
+T.
+(2010).
+On “feeling right” in cultural contexts: How person-culture match affects self-esteem and subjective well-being. Psychological Science, 21(11), 1563-1569.
+& 14.64\\
+ & Gifford, R., Scannell, L., Kormos, C., Smolova, L., Biel, A., Boncu, S., ...
+\& Uzzell, D.
+(2009).
+Temporal pessimism and spatial optimism in environmental assessments: An 18-nation study. Journal of environmental psychology, 29(1), 1-12.
+& 14.25\\
+ & Tranter, B., \& Booth, K. (2015). Scepticism in a changing climate: A cross-national study. Global Environmental Change, 33, 154-164. & 14.17\\
+ & Pisano, I., \& Lubell, M. (2017). Environmental behavior in cross-national perspective: A multilevel analysis of 30 countries. Environment and Behavior, 49(1), 31-58. & 13.75\\
+ & Kesler, C., \& Bloemraad, I.
+(2010).
+Does immigration erode social capital?
+The conditional effects of immigration-generated diversity on trust, membership, and participation across 19 countries, 1981–2000. Canadian Journal of Political Science/Revue canadienne de science politique, 43(2), 319-347.
+& 13.73\\
+ & Gelissen, J. (2007). Explaining popular support for environmental protection: A multilevel analysis of 50 nations. Environment and behavior, 39(3), 392-415. & 13.71\\
+ & Thorisdottir, H., Jost, J.
+T., Liviatan, I., \& Shrout, P.
+E.
+(2007).
+Psychological needs and values underlying left-right political orientation: Cross-national evidence from Eastern and Western Europe. Public Opinion Quarterly, 71(2), 175-203.
+& 13.50\\
+ & Hörisch, J., Kollat, J., \& Brieger, S.
+A.
+(2017).
+What influences environmental entrepreneurship?
+A multilevel analysis of the determinants of entrepreneurs’ environmental orientation. Small Business Economics, 48(1), 47-69.
+& 13.50\\
+ & Fungáčová, Z., Hasan, I., \& Weill, L. (2019). Trust in banks. Journal of Economic Behavior \& Organization, 157, 452-476. & 13.50\\
+ & Cuddy, A.
+J., Wolf, E.
+B., Glick, P., Crotty, S., Chong, J., \& Norton, M.
+I.
+(2015).
+Men as cultural ideals: Cultural values moderate gender stereotype content. Journal of personality and social psychology, 109(4), 622.
+& 13.17\\
+ & Mata, R., Josef, A. K., \& Hertwig, R. (2016). Propensity for risk taking across the life span and around the globe. Psychological science, 27(2), 231-243. & 13.00\\
+ & Venaik, S., \& Brewer, P. (2010). Avoiding uncertainty in Hofstede and GLOBE. Journal of international business studies, 41(8), 1294-1315. & 12.55\\
+ & Schimmack, U., Oishi, S., \& Diener, E.
+(2005).
+Individualism: A valid and important dimension of cultural differences between nations. Personality and Social Psychology Review, 9(1), 17-31.
+& 12.31\\
+ & Kvaløy, B., Finseraas, H., \& Listhaug, O. (2012). The publics’ concern for global warming: A cross-national study of 47 countries. Journal of Peace Research, 49(1), 11-22. & 12.00\\
+ & Mikucka, M., Sarracino, F., \& Dubrow, J.
+K.
+(2017).
+When does economic growth improve life satisfaction?
+Multilevel analysis of the roles of social trust and income inequality in 46 countries, 1981–2012. World Development, 93, 447-459.
+& 12.00\\
+ & Miyamoto, Y., Yoo, J., Levine, C.
+S., Park, J., Boylan, J.
+M., Sims, T., ...
+\& Ryff, C.
+D.
+(2018).
+Culture and social hierarchy: Self-and other-oriented correlates of socioeconomic status across cultures. Journal of Personality and Social Psychology, 115(3), 427.
+& 12.00\\
+Economic development & Inglehart, R., \& Baker, W. E. (2000). Modernization, cultural change, and the persistence of traditional values. American sociological review, 19-51. & 119.33\\
+ & King, R. G., \& Levine, R. (1993). Finance and growth: Schumpeter might be right. The quarterly journal of economics, 108(3), 717-737. & 90.00\\
+ & Treisman, D. (2000). The causes of corruption: a cross-national study. Journal of public economics, 76(3), 399-457. & 73.67\\
+ & Selden, T.
+M., \& Song, D.
+(1994).
+Environmental quality and development: is there a Kuznets curve for air pollution emissions?. Journal of Environmental Economics and management, 27(2), 147-162.
+& 49.07\\
+ & Benhabib, J., \& Spiegel, M. M. (1994). The role of human capital in economic development evidence from aggregate cross-country data. Journal of Monetary economics, 34(2), 143-173. & 47.89\\
+ & Acemoglu, D., Johnson, S., Robinson, J. A., \& Yared, P. (2008). Income and democracy. American Economic Review, 98(3), 808-42. & 44.62\\
+ & Easterly, W., \& Levine, R. (2003). Tropics, germs, and crops: how endowments influence economic development. Journal of monetary economics, 50(1), 3-39. & 36.10\\
+ & Wimmer, A., Cederman, L. E., \& Min, B. (2009). Ethnic politics and armed conflict: A configurational analysis of a new global data set. American sociological review, 74(2), 316-337. & 34.42\\
+ & Redding, S., \& Venables, A. J. (2004). Economic geography and international inequality. Journal of international Economics, 62(1), 53-82. & 32.65\\
+ & Liobikienė, G., Mandravickaitė, J., \& Bernatonienė, J.
+(2016).
+Theory of planned behavior approach to understand the green purchasing behavior in the EU: A cross-cultural study. Ecological Economics, 125, 38-46.
+& 31.00\\
+ & McCrae, R. R., \& Terracciano, A. (2005). Personality profiles of cultures: aggregate personality traits. Journal of personality and social psychology, 89(3), 407. & 26.12\\
+ & Faccio, M. (2010). Differences between politically connected and nonconnected firms: A cross‐country analysis. Financial management, 39(3), 905-928. & 25.73\\
+ & Ginarte, J. C., \& Park, W. G. (1997). Determinants of patent rights: A cross-national study. Research policy, 26(3), 283-301. & 25.08\\
+ & Beugelsdijk, S., \& Welzel, C. (2018). Dimensions and dynamics of national culture: Synthesizing Hofstede with Inglehart. Journal of cross-cultural psychology, 49(10), 1469-1505. & 24.00\\
+ & Jorgenson, A. K., \& Clark, B. (2012). Are the economy and the environment decoupling? A comparative international study, 1960–2005. American Journal of Sociology, 118(1), 1-44. & 23.00\\
+ & Durlauf, S. N., \& Johnson, P. A. (1995). Multiple regimes and cross‐country growth behaviour. Journal of applied econometrics, 10(4), 365-384. & 22.88\\
+ & Easterly, W. (2007). Inequality does cause underdevelopment: Insights from a new instrument. Journal of development economics, 84(2), 755-776. & 21.93\\
+ & Fuwa, M. (2004). Macro-level gender inequality and the division of household labor in 22 countries. American sociological review, 69(6), 751-767. & 21.24\\
+ & Smith, M.
+D., Rabbitt, M.
+P., \& Coleman-Jensen, A.
+(2017).
+Who are the world’s food insecure?
+New evidence from the Food and Agriculture Organization’s food insecurity experience scale. World Development, 93, 402-412.
+& 20.75\\
+ & Beck, T., Demirgüç-Kunt, A., \& Levine, R. (2003). Law, endowments, and finance. Journal of financial Economics, 70(2), 137-181. & 20.28\\
+ & Schreinemachers, P., \& Tipraqsa, P. (2012). Agricultural pesticides and land use intensification in high, middle and low income countries. Food policy, 37(6), 616-626. & 20.11\\
+ & Poe, S.
+C., Tate, C.
+N., \& Keith, L.
+C.
+(1999).
+Repression of the human right to personal integrity revisited: A global cross-national study covering the years 1976–1993. International studies quarterly, 43(2), 291-313.
+& 19.36\\
+ & Bloom, D. E., Sachs, J. D., Collier, P., \& Udry, C. (1998). Geography, demography, and economic growth in Africa. Brookings papers on economic activity, 1998(2), 207-295. & 18.61\\
+ & Skidmore, M., \& Toya, H. (2002). Do natural disasters promote long‐run growth?. Economic inquiry, 40(4), 664-687. & 18.47\\
+ & Kurtz, M. J., \& Schrank, A. (2007). Growth and governance: Models, measures, and mechanisms. The Journal of Politics, 69(2), 538-554. & 17.53\\
+ & Klasen, S., \& Lamanna, F.
+(2009).
+The impact of gender inequality in education and employment on economic growth: new evidence for a panel of countries. Feminist economics, 15(3), 91-132.
+& 17.25\\
+ & Welzel, C., Inglehart, R., \& Kligemann, H. D. (2003). The theory of human development: A cross‐cultural analysis. European Journal of Political Research, 42(3), 341-379. & 14.94\\
+ & Brülhart, M., \& Sbergami, F. (2009). Agglomeration and growth: Cross-country evidence. Journal of Urban Economics, 65(1), 48-63. & 14.33\\
+ & Beck, T., Demirguc-Kunt, A., \& Levine, R. (2005). SMEs, growth, and poverty: Cross-country evidence. Journal of economic growth, 10(3), 199-229. & 14.12\\
+ & Ehrhardt‐Martinez, K., Crenshaw, E.
+M., \& Jenkins, J.
+C.
+(2002).
+Deforestation and the environmental Kuznets curve: A cross‐national investigation of intervening mechanisms. Social Science Quarterly, 83(1), 226-243.
+& 13.05\\
+ & Rai, D., Zitko, P., Jones, K., Lynch, J., \& Araya, R.
+(2013).
+Country-and individual-level socioeconomic determinants of depression: multilevel cross-national comparison. The British Journal of Psychiatry, 202(3), 195-203.
+& 12.75\\
+ & Holzner, M. (2011). Tourism and economic development: The beach disease?. Tourism Management, 32(4), 922-933. & 12.70\\
+ & Calderón, C., Moral‐Benito, E., \& Servén, L. (2015). Is infrastructure capital productive? A dynamic heterogeneous approach. Journal of Applied Econometrics, 30(2), 177-198. & 12.67\\
+ & Roskam, A.
+J.
+R., Kunst, A.
+E., Van Oyen, H., Demarest, S., Klumbiene, J., Regidor, E., ...
+\& additional participants to the study.
+(2010).
+Comparative appraisal of educational inequalities in overweight and obesity among adults in 19 European countries. International journal of epidemiology, 39(2), 392-404.
+& 12.64\\
+ & Curtis, J. E., Baer, D. E., \& Grabb, E. G. (2001). Nations of joiners: Explaining voluntary association membership in democratic societies. American Sociological Review, 783-805. & 12.40\\
+ & Adams, S., \& Klobodu, E. K. M. (2018). Financial development and environmental degradation: does political regime matter?. Journal of Cleaner Production, 197, 1472-1479. & 12.33\\
+ & Qian, Y.
+(2007).
+Do national patent laws stimulate domestic innovation in a global patenting environment?
+A cross-country analysis of pharmaceutical patent protection, 1978–2002. The Review of Economics and Statistics, 89(3), 436-453.
+& 12.29\\
+ & Gaddis, I., \& Klasen, S. (2014). Economic development, structural change, and women’s labor force participation. Journal of Population Economics, 27(3), 639-681. & 11.86\\
+ & Bockstette, V., Chanda, A., \& Putterman, L. (2002). States and markets: The advantage of an early start. Journal of Economic growth, 7(4), 347-369. & 11.84\\
+ & Levels, M., Dronkers, J., \& Kraaykamp, G.
+(2008).
+Immigrant children's educational achievement in western countries: origin, destination, and community effects on mathematical performance. American Sociological Review, 73(5), 835-853.
+& 11.62\\
+ & Klasen, S.
+(2002).
+Low schooling for girls, slower growth for all?
+Cross‐country evidence on the effect of gender inequality in education on economic development. The World Bank Economic Review, 16(3), 345-373.
+& 11.42\\
+ & Knudsen, K., \& Wærness, K. (2008). National context and spouses’ housework in 34 countries. European Sociological Review, 24(1), 97-113. & 11.15\\
+ & Rigobon, R., \& Rodrik, D. (2005). Rule of law, democracy, openness, and income: Estimating the interrelationships. Economics of transition, 13(3), 533-564. & 11.00\\
+ & Beladi, H., Chao, C. C., Ee, M. S., \& Hollas, D. (2019). Does medical tourism promote economic growth? A cross-country analysis. Journal of Travel Research, 58(1), 121-135. & 11.00\\
+ & Roe, M.
+J., \& Siegel, J.
+I.
+(2011).
+Political instability: Effects on financial development, roots in the severity of economic inequality. Journal of Comparative Economics, 39(3), 279-309.
+& 10.90\\
+ & Roberts, J. T., \& Grimes, P. E. (1997). Carbon intensity and economic development 1962–1991: A brief exploration of the environmental Kuznets curve. World development, 25(2), 191-198. & 10.71\\
+ & Reinsberg, B., Kentikelenis, A., Stubbs, T., \& King, L.
+(2019).
+The world system and the hollowing out of state capacity: How structural adjustment programs affect bureaucratic quality in developing countries. American Journal of Sociology, 124(4), 1222-1257.
+& 10.50\\
+ & Jones-Smith, J.
+C., Gordon-Larsen, P., Siddiqi, A., \& Popkin, B.
+M.
+(2011).
+Cross-national comparisons of time trends in overweight inequality by socioeconomic status among women using repeated cross-sectional surveys from 37 developing countries, 1989–2007. American journal of epidemiology, 173(6), 667-675.
+& 10.40\\
+ & Demirgüç-Kunt, A., \& Levine, R. (1996). Stock market development and financial intermediaries: stylized facts. The World Bank Economic Review, 10(2), 291-321. & 10.36\\
+ & Sethi, D., \& Acharya, D. (2018). Financial inclusion and economic growth linkage: Some cross country evidence. Journal of Financial Economic Policy. & 10.33\\
+\bottomrule
+\end{longtable}
+
+}
+
+\end{ThreePartTable}
+\end{center}
 
@@
 \emph{Glottolog 3.0.} (Max Planck Institute for the Science of Human History, 2017).
-\url{doi:\%5B10.5281/zenodo.4061162\%5D(https://doi.org/10.5281/zenodo.4061162)}
+doi: \href{https://doi.org/10.5281/zenodo.4061162}{10.5281/zenodo.4061162}
 

### 2022-11-28 R/functionsSignal.R

@@ -341,7 +341,7 @@ makeTableGeoLinSignal <- function(geoHDI, geoGDP, geoGrow, geoGini,
   }
   # make table
   tibble(
-    Outcome = c("Human Development Index", "GDP per capita", "GDP per capita growth", "Gini index",
+    Outcome = c("HDI", "GDPpc", "GDPpc growth", "Gini index",
                 "Traditional values", "Survival values", "Tightness", "Individualism"),
     `Geographic signal` = c(
       tidySignal(geoHDI),

### 2022-11-28 R/functionsSimulation.R

@@ -101,14 +101,14 @@ simulateData <- function(simModel, covMat, continent, iso, langFam,
              rho = rho, lambda = lambda, r = r) %>%
       rename(latitude = Latitude..average.,
              longitude = Longitude..average.)
-    # add outcome variable averages for nations within a 2000km radius
+    # add predictor variable averages for nations within a 2000km radius
     # first get distance matrix in km
     distMat <- distm(as.matrix(out %>% dplyr::select(longitude, latitude))) / 1000
     # are two countries within a 2000km radius? (not including oneself)
     radiusMat <- distMat < 2000
     diag(radiusMat) <- FALSE
-    # for each nation, get outcome variable averages for surrounding nations
-    out$surroundingMean2000km <- apply(radiusMat, 1, function(x) mean(out$y[x]))
+    # for each nation, get predictor variable averages for surrounding nations
+    out$surroundingMean2000km <- apply(radiusMat, 1, function(i) mean(out$x[i]))
     return(out)
   }
   # map over iter

### 2022-11-28 _targets.R

@@ -14,7 +14,8 @@ options(
 tar_option_set(packages = c("brms", "cowplot", "conleyreg", "countrycode", 
                             "dagitty", "geosphere", "ggdag", "ggrepel", "ggtext", 
                             "haven", "lmtest", "papaja", "psych", "readxl", 
-                            "rstan", "sf", "sjlabelled", "tidybayes", "tidyverse"))
+                            "rstan", "sf", "sjlabelled", "tidybayes", "tidyverse"),
+               memory = "transient", garbage_collection = TRUE)
 
 # targets for simulation (see below)
 simulationTargets <-
@@ -576,7 +577,7 @@ list(
   
   #### Manuscript ####
   
-  #tar_render(manuscript, "manuscript.Rmd"),
+  tar_render(manuscript, "manuscript.Rmd"),
   
   #### Vignette ####
   

### 2022-11-28 run.R

@@ -1,1 +1,1 @@
-targets::tar_make_clustermq(workers = 5L)
+targets::tar_make_clustermq(workers = 3L)

### 2022-12-21 appendix.Rmd
resubmission

@@
 
-Geographic distance between two nations was calculated as the logged geodesic distance between country capital cities (data from the R package _maps_; Brownrigg, 2018) using the R package _geosphere_ (Hijmans, 2019).
+Geographic distance between two nations was calculated as the logged geodesic distance between nation capital cities (data from the R package _maps_; Brownrigg, 2018) using the R package _geosphere_ (Hijmans, 2019).
 The geographic proximity matrix was computed as one minus the log geographic distance matrix scaled between 0 and 1.
@@
 The covariance between two nations is assumed to be linearly proportional to the linguistic proximity between those nations.
-This assumption is justified if we assume that cultural traits evolve via Brownian motion along a language phylogeny.
+This assumption is justified if we assume that cultural traits evolve neutrally via Brownian motion along a language phylogeny.
+Although this is a conservative assumption, we follow this approach here rather than another Gaussian Process for several reasons: (1) to avoid model non-convergence when later including a spatial Gaussian Process in the same model, (2) to work around the lack of a coordinate system for linguistic distances, as opposed to latitude and longitude values for geographic distance, and (3) to showcase different ways that researchers can allow nations to covary to control for non-independence.
+
 The non-centered parameterisation of this model is:
@@
 
+## Supplementary Results
+
+### Further exploration of twelve reanalyses
+
+```{r load autocor}
+tar_load(c(inglehartSA1, skidmoreSA1, alesinaSignal1, knackSignal1, effSizeRatioModel1, effSizeRatioModel2))
+```
+
+To understand why some cross-national correlations were attenuated by controls for non-independence while others were robust, we further explored our fitted models for evidence of spatial and cultural autocorrelation.
+For each outcome variable, our Gaussian process models provided varying estimates of how quickly spatial autocorrelation declined with distance (Supplementary Figure \@ref(fig:plotReplications3)).
+For example, in Skidmore and Toya[@Skidmore2002] gross domestic product growth was only moderately spatially autocorrelated at 1,000 km distance (posterior median spatial autocorrelation at 1,000 km = `r printnum(median(skidmoreSA1))`, 95% CI [`r printnum(quantile(skidmoreSA1, 0.025))` `r printnum(quantile(skidmoreSA1, 0.975))`]), whereas in Inglehart and Baker[@Inglehart2000] traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1,000 km = `r printnum(median(inglehartSA1))`, 95% CI [`r printnum(quantile(inglehartSA1, 0.025))` `r printnum(quantile(inglehartSA1, 0.975))`]).
+We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \@ref(fig:plotReplications4)), with some outcome variables expressing low signal (e.g.
+confidence in institutions[@Knack1997]; posterior median = `r printnum(median(knackSignal1))`, 95% CI [`r printnum(quantile(knackSignal1, 0.025))` `r printnum(quantile(knackSignal1, 0.975))`]) and others expressing high signal (e.g.
+female labour force participation[@Alesina2013]; posterior median = `r printnum(median(alesinaSignal1))`, 95% CI [`r printnum(quantile(alesinaSignal1, 0.025))` `r printnum(quantile(alesinaSignal1, 0.975))`]).
+Exploratory regressions provided suggestive evidence that stronger estimates of spatial autocorrelation or cultural phylogenetic signal resulted in a more pronounced reduction in the effect size when controlling for non-independence between nations (Supplementary Figure \@ref(fig:plotReplications5)).
+However, these negative slopes for spatial autocorrelation (b = `r printnum(fixef(effSizeRatioModel1)["corAt1000km","Estimate"])`, 95% CI [`r printnum(fixef(effSizeRatioModel1)["corAt1000km","Q2.5"])` `r printnum(fixef(effSizeRatioModel1)["corAt1000km","Q97.5"])`]) and cultural phylogenetic signal (b = `r printnum(fixef(effSizeRatioModel2)["signal","Estimate"])`, 95% CI [`r printnum(fixef(effSizeRatioModel2)["signal","Q2.5"])` `r printnum(fixef(effSizeRatioModel2)["signal","Q97.5"])`]) were both very uncertain, due to the small number of analyses and the posterior uncertainty in effect sizes and estimates of non-independence.
+
+\newpage
+
 ## Supplementary Figures
@@
 climate, biodiversity, physical topography) and cultural and institutional causes (e.g.
-cultural norms, political systems).
+cultural norms, technologies, and institutions).
 In this causal model, we need to condition on $U$ to estimate the direct path from $X$ to $Y$, but we cannot since it is unobserved.
@@
 Lines and shaded areas are posterior median regression lines and 50% credible intervals from Bayesian multilevel models.
+Histograms represent relative counts for individual analyses that did (top) or did not (bottom) account for non-independence.
 Region FEs = region fixed effects.
@@
 
-(ref:plotSim1Caption) _Distribution of cross-national correlations from simulation study under strong spatial autocorrelation._ In these simulations, the strength of spatial autocorrelation is set to 0.8 for both outcome and predictor variables.
+(ref:plotSim1Caption) _Distribution of cross-national correlations from simulation study under strong spatial autocorrelation._ In these simulations, the strength of spatial autocorrelation is set to 0.8 for both outcome and predictor variables, and the true correlation is set to 0.
 For frequentist regression models, point ranges represent correlation estimates and 95% confidence intervals.
@@
 Correlations are ordered by effect size independently in each panel.
-Red point ranges indicate that the slope is "significant" (i.e.
-the 95% confidence / credible interval excludes zero).
-Black point ranges indicate that the slope is "not significant".
+Red point ranges indicate that the 95% confidence / credible interval excludes zero.
+Black point ranges indicate that the 95% confidence / credible interval includes zero.
 
@@
 
-(ref:plotSim2Caption) _Distribution of cross-national correlations from simulation study under strong cultural phylogenetic autocorrelation._ In these simulations, the strength of cultural phylogenetic autocorrelation is set to 0.8 for both outcome and predictor variables.
+(ref:plotSim2Caption) _Distribution of cross-national correlations from simulation study under strong cultural phylogenetic autocorrelation._ In these simulations, the strength of cultural phylogenetic autocorrelation is set to 0.8 for both outcome and predictor variables, and the true correlation is set to 0.
 For frequentist regression models, point ranges represent correlation estimates and 95% confidence intervals.
@@
 tar_read(plotSim2)
+```
+
+\newpage
+
+(ref:plotPowerGeo01Caption) _Statistical power estimates for different methods of controlling for spatial non-independence in our simulation study, assuming a small true effect size (r = 0.1)._ For simulated outcome and predictor variables, we systematically varied the strength of spatial autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+We simulated 100 datasets per parameter combination assuming a small true correlation between variables (_r_ = 0.1) and fitted different models to each dataset.
+Statistical power was operationalised as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero.
+Point ranges represent proportions and 95% bootstrap confidence intervals, and dashed lines indicate 80% power.
+SEs = standard errors.
+
+```{r plotPowerGeo01, fig.width=11, fig.height=7, fig.cap="(ref:plotPowerGeo01Caption)", fig.pos='H', out.width="\\textwidth"}
+tar_read(plotPowerGeo0.1)
+```
+
+\newpage
+
+(ref:plotPowerGeo03Caption) _Statistical power estimates for different methods of controlling for spatial non-independence in our simulation study, assuming a medium true effect size (r = 0.3)._ For simulated outcome and predictor variables, we systematically varied the strength of spatial autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+We simulated 100 datasets per parameter combination assuming a medium true correlation between variables (_r_ = 0.3) and fitted different models to each dataset.
+Statistical power was operationalised as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero.
+Point ranges represent proportions and 95% bootstrap confidence intervals, and dashed lines indicate 80% power.
+SEs = standard errors.
+
+```{r plotPowerGeo03, fig.width=11, fig.height=7, fig.cap="(ref:plotPowerGeo03Caption)", fig.pos='H', out.width="\\textwidth"}
+tar_read(plotPowerGeo0.3)
+```
+
+\newpage
+
+(ref:plotPowerGeo05Caption) _Statistical power estimates for different methods of controlling for spatial non-independence in our simulation study, assuming a large true effect size (r = 0.5)._ For simulated outcome and predictor variables, we systematically varied the strength of spatial autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+We simulated 100 datasets per parameter combination assuming a large true correlation between variables (_r_ = 0.5) and fitted different models to each dataset.
+Statistical power was operationalised as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero.
+Point ranges represent proportions and 95% bootstrap confidence intervals, and dashed lines indicate 80% power.
+SEs = standard errors.
+
+```{r plotPowerGeo05, fig.width=11, fig.height=7, fig.cap="(ref:plotPowerGeo05Caption)", fig.pos='H', out.width="\\textwidth"}
+tar_read(plotPowerGeo0.5)
+```
+
+\newpage
+
+(ref:plotPowerLin01Caption) _Statistical power estimates for different methods of controlling for cultural phylogenetic non-independence in our simulation study, assuming a small true effect size (r = 0.1)._ For simulated outcome and predictor variables, we systematically varied the strength of cultural phylogenetic autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+We simulated 100 datasets per parameter combination assuming a small true correlation between variables (_r_ = 0.1) and fitted different models to each dataset.
+Statistical power was operationalised as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero.
+Point ranges represent proportions and 95% bootstrap confidence intervals, and dashed lines indicate 80% power.
+SEs = standard errors.
+
+```{r plotPowerLin01, fig.width=11, fig.height=7, fig.cap="(ref:plotPowerLin01Caption)", fig.pos='H', out.width="\\textwidth"}
+tar_read(plotPowerLin0.1)
+```
+
+\newpage
+
+(ref:plotPowerLin03Caption) _Statistical power estimates for different methods of controlling for cultural phylogenetic non-independence in our simulation study, assuming a medium true effect size (r = 0.3)._ For simulated outcome and predictor variables, we systematically varied the strength of cultural phylogenetic autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+We simulated 100 datasets per parameter combination assuming a medium true correlation between variables (_r_ = 0.3) and fitted different models to each dataset.
+Statistical power was operationalised as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero.
+Point ranges represent proportions and 95% bootstrap confidence intervals, and dashed lines indicate 80% power.
+SEs = standard errors.
+
+```{r plotPowerLin03, fig.width=11, fig.height=7, fig.cap="(ref:plotPowerLin03Caption)", fig.pos='H', out.width="\\textwidth"}
+tar_read(plotPowerLin0.3)
+```
+
+\newpage
+
+(ref:plotPowerLin05Caption) _Statistical power estimates for different methods of controlling for cultural phylogenetic non-independence in our simulation study, assuming a large true effect size (r = 0.5)._ For simulated outcome and predictor variables, we systematically varied the strength of cultural phylogenetic autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+We simulated 100 datasets per parameter combination assuming a large true correlation between variables (_r_ = 0.5) and fitted different models to each dataset.
+Statistical power was operationalised as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero.
+Point ranges represent proportions and 95% bootstrap confidence intervals, and dashed lines indicate 80% power.
+SEs = standard errors.
+
+```{r plotPowerLin05, fig.width=11, fig.height=7, fig.cap="(ref:plotPowerLin05Caption)", fig.pos='H', out.width="\\textwidth"}
+tar_read(plotPowerLin0.5)
 ```
@@
 
-(ref:plotReplications3Caption) _Posterior estimates of Gaussian process functions mapping spatial autocorrelation onto geographic distance from our reanalyses of economic development (a) and cultural values (b) studies._ Estimates are from models additionally controlling for cultural phylogenetic non-independence.
-The y-axis represents the amount of spatial autocorrelation between data points with increasing distance between those points on the x-axis (logged distance in kilometres).
+(ref:plotReplications3Caption) _Posterior estimates of Gaussian process functions mapping spatial autocorrelation onto geographic distance from our reanalyses of economic development (a) and cultural values (b) studies._ The y-axis represents the amount of spatial autocorrelation between data points with increasing distance between those points on the x-axis (logged distance in kilometres).
 Lines and shaded areas represent median posterior spatial autocorrelation functions and 50% and 95% credible intervals.
@@
 
-(ref:plotReplications4Caption) _Posterior estimates of cultural phylogenetic signal from our reanalyses._ Estimates are from models additionally controlling for spatial non-independence.
-Cultural phylogenetic signal is operationalised as the proportion of national-level variance explained by linguistic proximity between nations.
+(ref:plotReplications4Caption) _Posterior estimates of cultural phylogenetic signal from our reanalyses._ Cultural phylogenetic signal is operationalised as the proportion of national-level variance explained by linguistic proximity between nations.
 Ridges are full posterior distributions, and points are posterior medians, and lines represent 50% and 95% credible intervals.
@@
 tar_read(plotReplications4)
+```
+
+\newpage
+
+(ref:plotReplications5Caption) _The estimated degree of spatial and cultural phylogenetic non-independence predicts reductions in effect size in our reanalysis._ (a) Higher estimated degrees of spatial autocorrelation at 1,000 km distance predict more pronounced reductions in effect sizes when controlling for non-independence.
+(b) Higher estimated levels of cultural phylogenetic signal predict more pronounced reductions in effect sizes when controlling for non-independence.
+In both panels, the y-axis represents the ratio of the effect size when controlling for spatial and cultural non-independence to the original effect size (from naive regression model), and the x-axis represents posterior median model estimates.
+Regression lines are plotted with 95% credible intervals.
+
+```{r plotReplications5, fig.width=8, fig.height=4, fig.cap="(ref:plotReplications5Caption)"}
+tar_read(plotReplications5)
 ```
@@
   placement = "H",
-  font_size = "small",
+  font_size = "footnotesize",
   longtable = TRUE,
@@
   caption = "(ref:tableArticlesReviewCaption)"
+)
+```
+
+\newpage
+
+(ref:tableSim1Caption) False positive rates in simulation with varying levels of spatial autocorrelation for the predictor variable ($\rho$) and for the outcome variable ($\lambda$), assuming that the true correlation is 0.
+Numbers represent the total number of analyses, out of 100, that estimated a slope with a 95% confidence / credible interval excluding zero, falsely inferring a relationship when none is present.
+
+```{r tableSim1, echo=F, warning=F, message=F}
+apa_table(
+  tar_read(tableSim1),
+  booktabs = TRUE,
+  landscape = TRUE,
+  escape = FALSE,
+  col_spanners = list(`$\\rho$ = 0.2` = c(2,4),
+                      `$\\rho$ = 0.5` = c(5,7),
+                      `$\\rho$ = 0.8` = c(8,10)),
+  col.names = c("Model", rep(paste0("$\\lambda$ = ", 
+                                    c(0.2, 0.5, 0.8)), times = 3)),
+  placement = "H",
+  font_size = "small",
+  caption = "(ref:tableSim1Caption)"
+)
+```
+
+\newpage
+
+(ref:tableSim2Caption) False positive rates in simulation with varying levels of cultural phylogenetic autocorrelation for the predictor variable ($\rho$) and for the outcome variable ($\lambda$), assuming that the true correlation is 0.
+Numbers represent the total number of analyses, out of 100, that estimated a slope with a 95% confidence / credible interval excluding zero, falsely inferring a relationship when none is present.
+
+```{r tableSim2, echo=F, warning=F, message=F}
+apa_table(
+  tar_read(tableSim2),
+  booktabs = TRUE,
+  landscape = TRUE,
+  escape = FALSE,
+  col_spanners = list(`$\\rho$ = 0.2` = c(2,4),
+                      `$\\rho$ = 0.5` = c(5,7),
+                      `$\\rho$ = 0.8` = c(8,10)),
+  col.names = c("Model", rep(paste0("$\\lambda$ = ", 
+                                    c(0.2, 0.5, 0.8)), times = 3)),
+  placement = "H",
+  font_size = "small",
+  caption = "(ref:tableSim2Caption)"
+)
+```
+
+\newpage
+
+(ref:tableReplicationsCaption) Estimated cross-national correlations from our reanalysis of twelve previous cross-national analyses.
+Numbers are posterior median slopes (equivalent to Pearson's r correlation coefficients) with 95% credible intervals.
+
+```{r tableReplications, echo=F, warning=F, message=F}
+apa_table(
+  tar_read(tableReplications),
+  booktabs = TRUE,
+  placement = "H",
+  font_size = "footnotesize",
+  align = "r",
+  landscape = TRUE,
+  caption = "(ref:tableReplicationsCaption)"
+)
+```
+
+\newpage
+
+(ref:tableDeviationsCaption) List of sample size deviations between original analyses and reanalyses.
+
+```{r tableDeviations, echo=F, warning=F, message=F}
+apa_table(
+  tar_read(tableDeviations),
+  booktabs = TRUE,
+  placement = "H",
+  font_size = "footnotesize",
+  longtable = TRUE,
+  align = c("m{4cm}", "m{2cm}", "m{2cm}", "m{7cm}"),
+  caption = "(ref:tableDeviationsCaption)"
 )

### 2022-12-21 manuscript.Rmd
resubmission

@@
     role: # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
+      - Conceptualization
       - Data curation
@@
       - Writing - review & editing
+  - name          : "Thanos Kyritsis"
+    affiliation   : "1"
+    role:
+      - Data curation
+      - Writing - review & editing
   - name          : "Quentin D. Atkinson"
     affiliation   : "1,2"
-    corresponding : yes
-    address       : "Floor 2, Building 302, 23 Symonds Street, Auckland, 1010, New Zealand"
-    email         : "q.atkinson@auckland.ac.nz"
+    #corresponding : yes
+    #address       : "Floor 2, Building 302, 23 Symonds Street, Auckland, 1010, New Zealand"
+    #email         : "q.atkinson@auckland.ac.nz"
     role:
       - Conceptualization
+      - Data curation
       - Funding acquisition
@@
 
-authornote: |
- This working paper has not yet been peer-reviewed.
+authornote: "This working paper has not yet been peer-reviewed.
+\\par Correspondence concerning this article should be addressed to Scott Claessens (scott.claessens@gmail.com) and Quentin D.
+Atkinson (q.atkinson@auckland.ac.nz), Floor 2, Building 302, 23 Symonds Street, Auckland, 1010, New Zealand."
 
 abstract: |
- Cross-national analyses test hypotheses about the drivers of global variation in national outcomes.
+ Cross-national analyses test hypotheses about the drivers of variation in national outcomes.
 However, since nations are connected in various ways, such as via spatial proximity and shared cultural ancestry, cross-national analyses often violate assumptions of non-independence, inflating false positive rates.
@@
 When studies do control for non-independence, our simulations suggest that most commonly used methods are insufficient for reducing false positives in non-independent data.
-In reanalyses of twelve previous cross-national correlations, half of the effect sizes show no association after controlling for non-independence using global proximity matrices.
+In reanalyses of twelve previous cross-national correlations, half of the estimates are compatible with no association after controlling for non-independence using global proximity matrices.
 We urge social scientists to sufficiently control for non-independence in cross-national research.
@@
 keywords          : "spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses"
-wordcount         : "xxxx words"
+wordcount         : "4892 words excl. Methods (Methods: 1877 words)"
 bibliography      : ["r-references.bib"]
@@
 The standard practice for cross-national analyses is to conduct bivariate correlations or multiple regressions with individual data points representing different nations.
-Such analyses widen the scope of social science beyond Western populations[@Henrich2010;@Pollet2014] and have been used to study, among other topics, the causes of variation in the economic wealth of nations[@Benhabib1994;@Comin2010;@LaPorta1997;@Sachs2001], global patterings of cultural norms and values[@Fincher2008;@Gelfand2011;@Hofstede2001;@Inglehart2000], and the universality and diversity of human behaviour and psychology around the world[@Awad2018;@Rhoads2021;@Schulz2019;@Thomson2018].
+Such analyses widen the scope of social science beyond Western populations[@Henrich2010;@Pollet2014] and have been used to study, among other topics, the causes of variation in the economic wealth of nations[@Benhabib1994;@Comin2010;@LaPorta1997;@Sachs2001], global patternings of cultural norms and values[@Fincher2008;@Gelfand2011;@Hofstede2001;@Inglehart2000], and the universality and diversity of human behaviour and psychology around the world[@Awad2018;@Rhoads2021;@Schulz2019;@Thomson2018].
 
@@
 For example, the related island nations Tonga and Tuvalu share similar languages and customs due to cultural inheritance from a common Polynesian population dating back more than 1,000 years.
-More recently, shared ancestry explains a myriad of cultural similarities between colonial settlements and their colonisers (e.g.
-Argentina and Spain).
+More recently, shared ancestry explains similarities in institutions, norms, technologies, and values between colonial settlements and their colonisers (e.g.
+English, French, Spanish, and Portuguese settlements of the Americas).
 Importantly, these deep cultural connections between nations often span large geographic distances around the world.
-Tonga and Tuvalu share cultural traits despite being separated by over 1,500 kilometres of ocean, and Argentina and Spain remain culturally similar today despite their location on two separate continents.
+Tonga and Tuvalu share cultural traits despite being separated by over 1,500 kilometres of ocean, and South American and European nations remain culturally similar today despite their locations on two separate continents.
 Shared cultural ancestry must therefore be considered independently of spatial proximity in the study of nations.
@@
 In the twentieth century, anthropologists began to emphasise that human societies do not develop independently, but rather exist in a globally interconnected "world system" linked by shared history and cultural ancestry [@Ferguson1997].
-Researchers compiled the Standard Cross-Cultural Sample of 186 cultures to minimise the confounding effects of this non-independence in comparisons of human societies[@Murdock1969], though spatial and cultural dependencies are difficult to remove entirely[@Dow2008;@Eff2004].
+In order to minimise the confounding effects of this non-independence in comparisons of human societies, researchers compiled the Standard Cross-Cultural Sample of 186 cultures which deliberately avoids sampling closely related cultures[@Murdock1969], though it is difficult to completely remove spatial and cultural dependencies[@Dow2008;@Eff2004] and the smaller sample of cultures reduces statistical power.
 Today, anthropologists borrow phylogenetic comparative methods from evolutionary biology, such as phylogenetic least squares regression[@Symonds2014], when comparing societies, treating culturally related societies in the same way as biologists treat genetically related species (e.g.
@@
 These variables are frequently included as both outcomes and predictors in cross-national studies in economics and psychology[@Benhabib1994;@Comin2010;@Fincher2008;@Gelfand2011;@Hofstede2001;@Inglehart2000;@LaPorta1997;@Sachs2001].
-First, we show that economic development and cultural values are spatially and culturally non-independent across nations, emphasising the need to control for non-independence.
-Second, we systematically review the 100 highest-cited cross-national studies of economic development and cultural values and estimate the proportion of cross-national analyses within these articles that account for non-independence between nations.
+First, we demonstrate that economic development and cultural values are spatially and culturally non-independent across nations, emphasising the need to control for non-independence.
+Second, we systematically review the 100 highest-cited cross-national studies of economic development and cultural values and determine baseline rates of controlling for non-independence in the literature.
 Third, we run simulations to determine whether common methods of dealing with non-independence in the literature sufficiently reduce false positive rates.
-Fourth, we reanalyse twelve previous cross-national analyses of economic development and cultural values from our systematic review, incorporating global geographic and linguistic proximity matrices to correctly control for spatial and cultural non-independence.
+Fourth, we reanalyse twelve previous cross-national analyses of economic development and cultural values from our systematic review, incorporating global geographic and linguistic proximity matrices as controls for spatial and cultural non-independence.
 
@@
 In order to motivate our research question, it is important to first quantify the degree of spatial and cultural non-independence for economic development and cultural values around the world.
-If these variables are only weakly non-independent, then the issue might be safe to ignore.
-However, if they are strongly non-independent, then there is a possibility that non-independence could be confounding cross-national inferences.
+If these variables are independent or only weakly non-independent, then the issue might be safe to ignore.
+However, if they are more strongly non-independent, then there is a possibility that non-independence could be confounding cross-national inferences.
 
-To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for a range of economic development and cultural values variables that are widely used in the literature.
-For economic development, we focused on the Human Development Index[@hdi], gross domestic product per capita, gross domestic product per capita growth, and the Gini index of income inequality.
+To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for a range of economic development and cultural values variables.
+For economic development, we focused on the Human Development Index[@hdi], gross domestic product per capita, annual gross domestic product per capita growth, and the Gini index of income inequality.
 For cultural values, we focused on two primary dimensions of cultural values from the World Values Survey, traditional vs.
@@
 self-expression values[@Inglehart2000], as well as cultural tightness[@Gelfand2011] and individualism[@Hofstede2001].
+These variables are not intended to be a comprehensive list of all national-level variables included in cross-national research, but rather an illustrative set of variables that are widely used in the literature.
 
-For all of these variables, we found that a substantial proportion of national-level variation was explained by spatial proximity and/or shared cultural ancestry between nations (Figure \@ref(fig:plotSignal); Supplementary Table \@ref(tab:tableSignal)).
+For all of these variables, we found that a substantial proportion of national-level variation was explained by spatial proximity and/or shared cultural ancestry between nations (Figure \@ref(fig:plotSignal); see Supplementary Table \@ref(tab:tableSignal) for numerical results).
 Signal estimates were often markedly strong, with spatial proximity and shared cultural ancestry frequently explaining over half of the national-level variation.
@@
   group_by(Review, Citation) %>%
-  summarise(Yes = ifelse(sum(ControlNI == "Yes") > 0, 1, 0)) %>%
+  summarise(Yes = ifelse(sum(ControlNI == "Yes") > 0, 1, 0), .groups = "drop") %>%
   group_by(Review) %>%
-  summarise(prop = mean(Yes)) %>%
+  summarise(prop = mean(Yes), .groups = "drop") %>%
   # bootstrap cis
@@
 Point ranges represent proportions and 95% bootstrap confidence intervals.
-Lines and shaded areas are posterior median regression lines and 50% credible intervals from Bayesian multilevel models.
+Lines and shaded areas are posterior median regression lines and 50% credible intervals from Bayesian logistic regression models.
+Dots represent raw counts of individual articles that did (top) or did not (bottom) account for non-independence.
 Region FEs = region fixed effects.
 
-```{r plotReview1, fig.width=6.5, fig.height=6, fig.cap="(ref:plotReview1Caption)"}
+```{r plotReview1, fig.width=7, fig.height=6, fig.cap="(ref:plotReview1Caption)"}
 tar_read(plotReview1)
@@
 
-While our review contains a range of articles from low- and high-impact journals, an anonymous reviewer suggested that our estimates could be biased downwards by analyses published in lower impact outlets with more relaxed standards for issues like non-independence.
+While our review contains articles from journals with a range of impact factors, an anonymous reviewer suggested that our estimates could be biased downwards by analyses published in lower impact outlets with more relaxed standards for issues like non-independence.
 It is also possible that, since our systematic review goes back as far as 1993, our estimates are being biased downwards by earlier studies, and that controls for non-independence have increased over time with methodological advancements and greater awareness of the issue.
-To test these possible explanations for our low estimates, we fitted Bayesian logistic regression models to the data from the review, including log journal impact factor and publication year as separate predictors.
-Interestingly, we found that, contrary to the explanation above, studies from higher impact journals were _less_ likely to include at least one control for non-independence than studies from lower impact journals, both for studies of economic development (b = `r printnum(fixef(ifArticle)["ReviewEconomicdevelopment:logIF","Estimate"])`, 95% CI [`r printnum(fixef(ifArticle)["ReviewEconomicdevelopment:logIF","Q2.5"])` `r printnum(fixef(ifArticle)["ReviewEconomicdevelopment:logIF","Q97.5"])`]) and for studies of cultural values (b = `r printnum(fixef(ifArticle)["ReviewValues:logIF","Estimate"])`, 95% CI [`r printnum(fixef(ifArticle)["ReviewValues:logIF","Q2.5"])` `r printnum(fixef(ifArticle)["ReviewValues:logIF","Q97.5"])`]; Figure \@ref(fig:plotReview1)b).
+To test these possible explanations for our low estimates, we fitted logistic regression models to the data from the review, including log journal impact factor and publication year as separate predictors.
+Interestingly, we found that, if anything, studies from higher impact journals were _less_ likely to include at least one control for non-independence than studies from lower impact journals, both for studies of economic development (b = `r printnum(fixef(ifArticle)["ReviewEconomicdevelopment:logIF","Estimate"])`, 95% CI [`r printnum(fixef(ifArticle)["ReviewEconomicdevelopment:logIF","Q2.5"])` `r printnum(fixef(ifArticle)["ReviewEconomicdevelopment:logIF","Q97.5"])`]) and for studies of cultural values (b = `r printnum(fixef(ifArticle)["ReviewValues:logIF","Estimate"])`, 95% CI [`r printnum(fixef(ifArticle)["ReviewValues:logIF","Q2.5"])` `r printnum(fixef(ifArticle)["ReviewValues:logIF","Q97.5"])`]; Figure \@ref(fig:plotReview1)b).
 Moreover, splines revealed no relationship between publication year and the probability of including at least one control for non-independence, both for studies of economic development (b = `r printnum(fixef(yearArticle)["sYear.std:ReviewEconomicdevelopment_1","Estimate"])`, 95% CI [`r printnum(fixef(yearArticle)["sYear.std:ReviewEconomicdevelopment_1","Q2.5"])` `r printnum(fixef(yearArticle)["sYear.std:ReviewEconomicdevelopment_1","Q97.5"])`]) and for studies of cultural values (b = `r printnum(fixef(yearArticle)["sYear.std:ReviewValues_1","Estimate"])`, 95% CI [`r printnum(fixef(yearArticle)["sYear.std:ReviewValues_1","Q2.5"])` `r printnum(fixef(yearArticle)["sYear.std:ReviewValues_1","Q97.5"])`]; Figure \@ref(fig:plotReview1)c).
@@
 [@Schulz2019;@Schulz2022]).
-These frequentist approaches attempt to account for non-independence by holding geographic location constant (latitude, longitude), discarding between-region variation and exploiting only local variation (continent fixed effects, mean of surrounding 2000km), or correcting standard errors for autocorrelation post-hoc while leaving model coefficients unchanged (Conley standard errors).
+These fixed effect approaches attempt to account for non-independence by holding geographic location constant (latitude, longitude), discarding between-region variation and exploiting only local variation (continent fixed effects, mean of surrounding 2000km), or correcting standard errors for autocorrelation post-hoc while leaving model coefficients unchanged (Conley standard errors).
 
-Beyond frequentist approaches, we also fitted Bayesian multilevel regressions that explicitly model spatial and/or cultural phylogenetic non-independence by allowing nations to covary according to geographic and/or linguistic proximity matrices.
+Beyond fixed effect approaches, we also fitted Bayesian random effects regressions that explicitly model spatial and/or cultural phylogenetic non-independence by allowing nations to covary according to geographic and/or linguistic proximity matrices.
 Geographic proximity between nations is calculated from inverse distances between longitude and latitude coordinates.
 Linguistic proximity between nations is calculated from a global phylogenetic tree that represents hierarchical relationships of genealogical descent for all languages in the world.
-For each pair of nations, we calculate inverse phylogenetic distances (i.e.
-number of branches separating two taxa) between all languages spoken in that nation pair and produce an average "linguistic proximity" score weighted by the percentages of speakers within those nations.
+For each pair of nations, we calculated inverse phylogenetic distances (i.e.
+number of branches separating two taxa) between all languages spoken in that nation pair and produced an average "linguistic proximity" score weighted by the percentages of speakers within those nations.
 To include the resulting geographic and linguistic proximity matrices in our models, we included a Gaussian process[@McElreath2020;@Neal1998] over latitude and longitude values and/or assumed that nation random intercepts were correlated in proportion to their linguistic proximity[@deVillemereuil2014].
-These approaches attempt to account for non-independence by modelling the covariance between nations that is induced by their geographic or linguistic connections.
+These random effects approaches attempt to account for non-independence by modelling the covariance between nations that is induced by their geographic or linguistic connections.
 
@@
   tar_read(olsModel1_simGeoCov) %>%
+  filter(r == 0) %>%
   group_by(lambda, rho) %>%
-  summarise(n = sum(sig) / n())
+  summarise(n = sum(sig) / n(), .groups = "drop")
 fpLin <-
   tar_read(olsModel1_simLinCov) %>%
+  filter(r == 0) %>%
   group_by(lambda, rho) %>%
-  summarise(n = sum(sig) / n())
+  summarise(n = sum(sig) / n(), .groups = "drop")
 fpGeoContinent <-
   tar_read(olsModel4_simGeoCov) %>%
+  filter(r == 0) %>%
   group_by(lambda, rho) %>%
-  summarise(n = sum(sig) / n())
+  summarise(n = sum(sig) / n(), .groups = "drop")
+fpGeoRadius <-
+  tar_read(olsModel6_simGeoCov) %>%
+  filter(r == 0) %>%
+  group_by(lambda, rho) %>%
+  summarise(n = sum(sig) / n(), .groups = "drop")
 fpGeoSpatial <-
   tar_read(brmsModel1_simGeoCov) %>%
+  filter(r == 0) %>%
   group_by(lambda, rho) %>%
-  summarise(n = sum(sig) / n())
+  summarise(n = sum(sig) / n(), .groups = "drop")
 fpLinLangFam <-
   tar_read(olsModel5_simLinCov) %>%
+  filter(r == 0) %>%
   group_by(lambda, rho) %>%
-  summarise(n = sum(sig) / n())
+  summarise(n = sum(sig) / n(), .groups = "drop")
 ```
 
-Figures \@ref(fig:plotSim3) and \@ref(fig:plotSim4) plot the estimated false positive rates from our simulation study, split by different methods and different degrees of spatial or cultural phylogenetic autocorrelation (see Supplementary Figures \@ref(fig:plotSim1) and \@ref(fig:plotSim2) for full distributions of effect sizes under strong autocorrelation).
-Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero (i.e.
-falsely infering a relationship when none is present).
+Figures \@ref(fig:plotSim3) and \@ref(fig:plotSim4) plot the estimated false positive rates from our simulation study, split by different methods and different degrees of spatial or cultural phylogenetic autocorrelation (see Supplementary Tables \@ref(tab:tableSim1) and \@ref(tab:tableSim2) for numerical results and Supplementary Figures \@ref(fig:plotSim1) and \@ref(fig:plotSim2) for full distributions of effect sizes under strong autocorrelation).
+Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero, falsely inferring a relationship when none is present.
 For reference, "weak" autocorrelation in our simulation is comparable to the geographic signal for survival values in Figure \@ref(fig:plotSignal) (i.e.
@@
 
-Common methods in the literature do not reduce these high false positive rates.
+Most methods common in the literature do not reduce these high false positive rates.
 With strong spatial autocorrelation for both outcome and predictor variables, false positive rates remain above 50% when controlling for latitude, longitude, and language family fixed effects (Figure \@ref(fig:plotSim3)).
-Applying Conley standard errors also does not reduce false positive rates below 50% under strong spatial autocorrelation, regardless of the distance cutoff.
-Continent fixed effects are more effective than other frequentist methods, though they continue to produce a false positive rate of `r printnum(fpGeoContinent$n[fpGeoContinent$lambda == 0.8 & fpGeoContinent$rho == 0.8]*100, digits = 0)`% under strong spatial autocorrelation.
-By contrast, Bayesian spatial Gaussian process regression with longitude and latitude values outperforms all other methods.
-This approach eliminates false positives under moderate spatial autocorrelation, such that the false positive rate is no different from chance, and reduces the false positive rate under strong spatial autocorrelation to `r printnum(fpGeoSpatial$n[fpGeoSpatial$lambda == 0.8 & fpGeoSpatial$rho == 0.8]*100, digits = 0)`%.
-Bayesian models that additionally account for linguistic proximity between nations perform equally well, though models with only linguistic covariance continue to produce false positives.
+Similarly, Conley standard error corrections based on spatial and genetic distances do not reduce false positive rates below 40% under strong spatial autocorrelation.
+The most effective fixed effects methods are continent fixed effects, which continue to produce a false positive rate of `r printnum(fpGeoContinent$n[fpGeoContinent$lambda == 0.8 & fpGeoContinent$rho == 0.8]*100, digits = 0)`% under strong spatial autocorrelation, and controlling for the mean of the predictor variable within a 2000km radius, which eliminates false positives under even strong spatial autocorrelation (`r printnum(fpGeoRadius$n[fpGeoRadius$lambda == 0.8 & fpGeoRadius$rho == 0.8]*100, digits = 0)`%).
+However, additional simulations revealed that these reductions in false positive rates come at the cost of lower statistical power (see Supplementary Figures \@ref(fig:plotPowerGeo01) - \@ref(fig:plotPowerGeo05)).
+In additional simulations where the true relationship between the predictor and outcome variable was known, power analyses showed that both continent fixed effects and the 2000km radius control had less than 80% power to detect moderate true correlations (_r_ = 0.3) under strong spatial autocorrelation.
 
-In our simulation of cultural phylogenetic non-independence, we find that most frequentist methods do not reduce false positive rates (Figure \@ref(fig:plotSim4)).
-Controls for latitude and longitude, continent fixed effects, and Conley standard errors do little to change false positive rates.
-Language family fixed effects are slightly more effective than other frequentist methods, though they continue to produce a false positive rate of `r printnum(fpLinLangFam$n[fpLinLangFam$lambda == 0.8 & fpLinLangFam$rho == 0.8]*100, digits = 0)`% under strong cultural phylogenetic autocorrelation.
-By contrast, Bayesian models with random effects covarying according to linguistic proximity completely eliminate false positives across all degrees of cultural phylogenetic autocorrelation.
-Bayesian models that additionally account for geographic proximity between nations perform equally well, though models with only a spatial Gaussian process continue to produce false positives.
+By contrast, Bayesian spatial Gaussian process regression with longitude and latitude strikes a balance between reducing false positives and retaining high statistical power to detect true effects.
+This approach reduces false positives to `r printnum(fpGeoSpatial$n[fpGeoSpatial$lambda == 0.5 & fpGeoSpatial$rho == 0.5]*100, digits = 0)`% under moderate spatial autocorrelation and `r printnum(fpGeoSpatial$n[fpGeoSpatial$lambda == 0.8 & fpGeoSpatial$rho == 0.8]*100, digits = 0)`% under strong spatial autocorrelation.
+Random effects models that additionally account for linguistic proximity between nations perform equally well, though models with only linguistic covariance continue to produce false positives.
+False positives are not completely eliminated with these random effects models.
+Nevertheless, these methods have at least 80% power to detect moderate (_r_ = 0.3) and large (_r_ = 0.5) true correlations between variables under all levels of spatial autocorrelation (see Supplementary Figures \@ref(fig:plotPowerGeo01) - \@ref(fig:plotPowerGeo05)).
+
+In our simulation of cultural phylogenetic non-independence, we find that none of the fixed effects methods reduce false positive rates (Figure \@ref(fig:plotSim4)).
+Controls for latitude, longitude, continent fixed effects, the mean of the predictor variable in a 2000km radius, and Conley standard error corrections based on spatial and genetic distances do little to change false positive rates.
+Even language family fixed effects continue to produce a false positive rate of `r printnum(fpLinLangFam$n[fpLinLangFam$lambda == 0.8 & fpLinLangFam$rho == 0.8]*100, digits = 0)`% under strong cultural phylogenetic autocorrelation.
+By contrast, models with random effects covarying according to linguistic proximity completely eliminate false positives across all degrees of cultural phylogenetic autocorrelation.
+This approach is also the only method that is able to detect large true correlations (_r_ = 0.5) with at least 80% power (see Supplementary Figures \@ref(fig:plotPowerLin01) - \@ref(fig:plotPowerLin05)).
+Random effects models that additionally account for geographic proximity between nations perform equally well, though models with only a spatial Gaussian process continue to produce false positives.
 
@@
 
-Our systematic review and simulation study have shown that controls for non-independence are rare in cross-national studies of economic development and cultural values, and when studies do attempt to control for non-independence, the methods typically used are likely to continue to produce false positives.
+Our systematic review and simulation study have shown that controls for non-independence are rare in cross-national studies of economic development and cultural values, and when studies do attempt to control for non-independence, the methods typically used are unable to sufficiently reduce false positive rates.
 This raises the worrying possibility that the cross-national literature in economics and psychology is populated with spurious relationships.
 
-To determine how widespread this issue of spurious cross-national relationships might be, we reanalysed a subset of twelve previous cross-national analyses from our systematic review, sufficiently controlling for spatial and cultural phylogenetic non-independence using global geographic and linguistic proximity matrices.
-We subsampled six analyses from our economic development review[@Beck2003;@Beck2005;@Bockstette2002;@Easterly2003;@Easterly2007;@Skidmore2002] and six from our cultural values review[@Adamczyk2009;@Alesina2013;@Fincher2008;@Gelfand2011;@Inglehart2000;@Knack1997].
-Our choice of analyses was constrained by data availability and whether we were able to initially replicate the original finding.
-We pre-registered our subsample of analyses before running any control models (https://osf.io/uywx8/).
+To determine how widespread this issue of spurious cross-national relationships might be, we reanalysed a subset of twelve previous cross-national analyses from our systematic review, controlling for spatial and cultural phylogenetic non-independence using global geographic and linguistic proximity matrices.
+Out of the 100 papers included in our systematic review, primary or secondary data were publicly available for `r printnum(review %>% group_by(Review, Citation) %>% summarise(DataAvailable = ifelse(mean(DataAvailable == "Yes") > 0, TRUE, FALSE), .groups = "drop") %>% pull(DataAvailable) %>% sum())` papers.
+We attempted to replicate key statistically significant cross-national correlations from these papers --- mostly initial bivariate regression specifications without covariates (see Methods) --- and stopped when we had sampled a set of twelve analyses for which we were able to replicate the original result.
+The final set included six analyses from our economic development review[@Beck2003;@Beck2005;@Bockstette2002;@Easterly2003;@Easterly2007;@Skidmore2002] and six from our cultural values review[@Adamczyk2009;@Alesina2013;@Fincher2008;@Gelfand2011;@Inglehart2000;@Knack1997] that had available data and were able to be replicated.
+We pre-registered this set before running any control models (https://osf.io/uywx8/).
 We controlled for non-independence by including (1) a Gaussian process allowing nation random intercepts to covary according to a geographic proximity matrix, and/or (2) nation random intercepts that covaried according to a linguistic proximity matrix (see Supplementary Methods for full models).
 
-Figure \@ref(fig:plotReplications1) visualises the results of our reanalysis.
+Figure \@ref(fig:plotReplications1) visualises the results of our reanalysis (see Supplementary Table \@ref(tab:tableReplications) for numerical results).
 Cross-national correlation effect sizes tended to reduce when controlling for statistical non-independence between nations, sometimes by as much as half of the original effect size.
@@
 (2011) is a partial correlation controlling for log gross national income and Adamczyk and Pitt (2009) is a multilevel model including several covariates.
-Points and black lines represent posterior medians and 95% credible intervals.
+Coloured ridges are full posterior distributions, and points and black lines represent posterior medians and 95% credible intervals.
 GDP = gross domestic product.
@@
 
-```{r load autocor}
-tar_load(c(inglehartSA1, skidmoreSA1, alesinaSignal1, gelfandSignal1))
-```
-
-To understand why some cross-national correlations were attenuated by controls for non-independence while others were robust, we further explored our fitted models for evidence of spatial and cultural autocorrelation.
-For each outcome variable, our Gaussian process models provided varying estimates of how quickly spatial autocorrelation declined with distance (Supplementary Figure \@ref(fig:plotReplications3)).
-For example, in Skidmore and Toya[@Skidmore2002] gross domestic product growth was only moderately spatially autocorrelated at 1,000 km distance (posterior median spatial autocorrelation at 1,000 km = `r printnum(median(skidmoreSA1))`, 95% CI [`r printnum(quantile(skidmoreSA1, 0.025))` `r printnum(quantile(skidmoreSA1, 0.975))`]), whereas in Inglehart and Baker[@Inglehart2000] traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1,000 km = `r printnum(median(inglehartSA1))`, 95% CI [`r printnum(quantile(inglehartSA1, 0.025))` `r printnum(quantile(inglehartSA1, 0.975))`]).
-We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \@ref(fig:plotReplications4)), with some outcome variables expressing low signal (e.g.
-tightness[@Gelfand2011]; posterior median = `r printnum(median(gelfandSignal1))`, 95% CI [`r printnum(quantile(gelfandSignal1, 0.025))` `r printnum(quantile(gelfandSignal1, 0.975))`]) and others expressing high signal (e.g.
-female labour force participation[@Alesina2013]; posterior median = `r printnum(median(alesinaSignal1))`, 95% CI [`r printnum(quantile(alesinaSignal1, 0.025))` `r printnum(quantile(alesinaSignal1, 0.975))`]).
-Across all analyses, we found that stronger estimates of spatial autocorrelation or cultural phylogenetic signal resulted in a more pronounced reduction in the effect size when controlling for non-independence between nations (Figure \@ref(fig:plotReplications5)).
-
-(ref:plotReplications5Caption) _The estimated degree of spatial and cultural phylogenetic non-independence predicts reductions in effect size in our reanalysis._ (a) Higher estimated degrees of spatial autocorrelation at 1,000 km distance predict more pronounced reductions in effect sizes when controlling for non-independence.
-(b) Higher estimated levels of cultural phylogenetic signal predict more pronounced reductions in effect sizes when controlling for non-independence.
-In both panels, the y-axis represents the ratio of the effect size when controlling for spatial and cultural non-independence to the original effect size (from naive regression model), and the x-axis represents posterior median model estimates.
-Regression lines are plotted with 95% confidence intervals.
-
-```{r plotReplications5, fig.width=8, fig.height=4, fig.cap="(ref:plotReplications5Caption)"}
-tar_read(plotReplications5)
-```
-
 # Discussion
 
-In a systematic literature review and simulation, we found that cross-national studies in economics and psychology rarely account for non-independence between nations, and, when they do, the methods they use continue to produce false positives.
-In a reanalysis of twelve cross-national correlations, we further showed that neglecting to account for non-independence has resulted in spurious relationships in the published literature, with half of the correlations failing to replicate when controlling for spatial or cultural non-independence with more rigorous methods.
+In a systematic literature review and simulation, we found that cross-national studies in economics and psychology rarely account for non-independence between nations, and, when they do, the methods they use are insufficient to reduce false positives in non-independent data.
+In a reanalysis of twelve cross-national correlations, we further showed that neglecting to account for non-independence has resulted in potentially spurious relationships in the published literature, with half of the correlations failing to replicate when controlling for spatial or cultural non-independence with more rigorous methods.
 These findings suggest that cross-national analyses in economics and psychology should be interpreted with caution until non-independence is sufficiently accounted for.
 
-Our initial analyses add to and clarify existing evidence regarding the non-independence of economic and cultural variation among nations.
+Our initial analyses add to and clarify existing evidence regarding the degree of non-independence for national-level economic and cultural variables.
 One previous study suggested that geographic proximity is more important than deep cultural ancestry in explaining the distribution of human development across Eurasian nations, though the authors noted that their small sample of 44 nations and regional focus limited their statistical power[@Sookias2018].
-Our global sample of nations revealed strong cultural phylogenetic signal, as well as geographic signal, for the Human Development Index.
+By contrast, our global samples of over 160 nations revealed strong cultural phylogenetic signal, as well as geographic signal, for the Human Development Index, GDP per capita, and the Gini index of inequality.
 Another previous study found that similarities in the cultural values of nations are predicted by linguistic, but not geographic, distances between those nations[@White2021].
 We find this same result for survival vs.
-self-expression values, but for traditional vs.
+self-expression values, cultural tightness, and individualism, but for traditional vs.
 secular values we find that both linguistic and geographic proximity are important independent predictors of global variation.
@@
 In our simulations, controlling for either latitude or longitude did not reduce false positive rates.
-This result calls into question controls like distance to the equator to account for non-independence in cross-national regression models, though these controls may still be suitable to account for regional or latitudinal variation in ecology, which we did not simulate.
-High false positive rates persisted with Conley standard errors, which have previously been critiqued for being overly sensitive to arbitrary distance cutoffs[@Kelly2020].
+This result calls into question the use of controls like distance to the equator to account for non-independence in cross-national regression models, though these controls may still be suitable to account for regional or latitudinal variation in climate, ecology, and natural threats (e.g.
+pathogens), which we did not simulate.
+High false positive rates persisted with Conley standard errors, despite recent claims that these standard error corrections are sufficient to deal with spatial non-independence[@Voth2021].
 The simulation also confirmed the assertion that fixed effects for spatial or cultural groupings (e.g.
@@
 [@Smith2017]).
-Across all model types in our simulation, the only methods that sufficiently reduced the false positive rate were the Bayesian multilevel regressions that explicitly modelled spatial and cultural phylogenetic autocorrelation, though we did not include other possible controls for non-independence, such as conditional autoregressive models[@Lichstein2002] or generalised additive models[@Hastie2017].
+Controlling for the mean of the predictor variable within a 2000km radius[@Schulz2022] eliminated false positive rates in spatially autocorrelated data, but had reduced statistical power to detect true associations.
+Across all model types in our simulation, the only methods that markedly reduced false positive rates while retaining high statistical power were the random effects models with covariance matrices.
+The important advantage of these models is not that they are Bayesian _per se_ (any of the approaches used here could be implemented in a Bayesian framework), but rather that they explicitly model covariance as a function of spatial or cultural distance.
+
+There are other approaches to controlling for non-independence that we did not include in our simulation.
+For example, conditional autoregressive models[@Lichstein2002] and generalised additive models[@Hastie2017] are approaches that can be applied in both frequentist and Bayesian frameworks.
+There are also alternative ways to operationalise cultural distances between nations, including metrics like cultural fixation indices (e.g.
+cultural $F_{ST}$[@Muthukrishna2020;@White2021]), covariance based on genetic distances[@Spolaore2009], and phylogenetic distances between religious traditions.
+These metrics allow for cultural ancestry effects beyond ethnolinguistic genealogies, though some may produce less interpretable results than language ancestry.
+For example, unlike language, genes are strongly correlated with geography[@Handley2007;@Novembre2008] and their global patternings are consistent with different underlying processes, such as deep cultural inheritance and more recent gene flow between populations.
+Future work should explore the interpretability of these methods and determine whether they are sufficient to reduce false positive rates in spatially and culturally non-independent data.
+In addition to spatial proximity and shared cultural ancestry, we did not simulate other sources of non-independence that potentially exist in real cross-national datasets, such as modern connections between nations due to flows of people and information (e.g.
+flight networks, social media networks) and shared histories of colonialism and capitalist modes of production that have shaped today's global landscape[@Trouillot2016;@Wolf2010].
+Additional controls will be required to ensure that these sources of non-independence do not confound cross-national inferences.
 
@@
 At the institutional level, one possibility is that such practices are incentivised _because_ they generate statistically significant relationships, which increase the probability that a study is published[@Smaldino2016].
-At the individual level, another possibility is that researchers outside of anthropology and ecology are simply not aware of the problem, or believe that the problem does not apply to analyses of nations.
+Indeed, we found that controls for non-independence were _less_ common among articles published in high impact journals, suggesting that researchers are rewarded for such practices.
+At the individual level, another possibility is that researchers outside of anthropology and ecology are less aware of the problem, or believe that the problem does not apply to analyses of nations.
 Even if researchers appreciate the problem, they might not know of suitable controls or perceive the methods to be too complex.
@@
 First, cross-national replication studies like ours and others[@Bromham2018;@Bromham2021;@Currie2012;@Kelly2020;@Passmore2022], combined with the methodological reviews included in Registered Reports[@Chambers2021], might change incentive structures and encourage researchers to analyse the world's nations with more rigorous methods.
-Second, since the issue of non-independence is fundamentally an issue of causal inference (Supplementary Figure \@ref(fig:dag)), more explicit descriptions of causal models could promote controls for non-independence in cross-national research.
-In our review, economists studying economic development dealt with national-level non-independence more than psychologists studying cultural values, likely because economics studies tend to be lengthy statistical exercises that systematically incorporate or exclude numerous variables in an attempt to infer causation.
+Second, more explicit descriptions of causal models could promote controls for non-independence by clearly outlining the nature of confounding and the sources of autocorrelation in cross-national data[@Deffner2022].
+The causal model outlined in Supplementary Figure \@ref(fig:dag) is a useful example, but individual studies must outline their own particular causal assumptions, which may include further sources of non-independence and confounding variables to control for (e.g.
+post-communist status, colony status).
+These causal models can then be used to design tailored statistical estimation strategies.
+Indeed, in our review, economists studying economic development dealt with national-level non-independence more than psychologists studying cultural values, likely because economics studies tend to be lengthy statistical exercises that systematically incorporate or exclude numerous variables in an attempt to infer causation.
 Third, the recent widespread accessibility of open source statistical software, such as the programming language Stan[@rstan] and the R package _brms_[@Burkner2017], should promote the use of more rigorous methods to control for non-independence.
 Using _brms_, for example, Bayesian Gaussian process regression is straightforward to conduct, requiring only longitude and latitude values for nations.
+We have provided an online tutorial to help researchers apply these methods to their own cross-cultural datasets (https://scottclaessens.github.io/blog/2022/crossnational/).
 
@@
 Three of these analyses were tests of "persistence" hypotheses, studying the effects of historical and environmental conditions --- settler mortality[@Beck2003], wheat-sugar suitability[@Easterly2007], and natural disaster frequency[@Skidmore2002] --- on modern developmental outcomes.
-A recent reanalysis has also called into question various studies of this ilk[@Kelly2020].
+These results are in line with a recent paper that has called into question various studies of this ilk[@Kelly2020] and contradict claims that the concern over spatial non-independence has been overblown in these studies[@Voth2021].
 We also found that two out of six cross-national correlations with cultural values variables had 95% credible intervals that included zero when controlling for cultural phylogenetic non-independence.
+Overall, these reanalyses show that the problem of statistical non-independence applies to a wider range of national-level variables than those identified by previous work, such as parasite stress and democratic outcomes[@Bromham2018], raising the worrying possibility that the current cross-national literature is populated with spurious relationships.
+Future work should expand our set of reanalyses to determine the extent of this problem in the literature.
 
@@
 ```{r}
-tar_load(c(hdi, wvs))
+tar_load(c(hdi, gdp, gdpGrowth, gini, wvs, tightness, fincherData))
 ```
@@
 To estimate the degree of spatial and cultural phylogenetic non-independence in economic development and cultural values, we calculated geographic and cultural phylogenetic signal for global measures of development and values.
-Our measure of economic development was the Human Development Index[@hdi].
-We retrieved a longitudinal dataset capturing human development for `r printnum(length(unique(hdi$Country)))` nations since 1990 (n = `r printnum(as.numeric(nrow(hdi)), digits = 0)`; https://hdr.undp.org/en/content/download-data).
-Our measures of cultural values were traditional vs.
+For economic development variables, we retrieved longitudinal data on the Human Development Index[@hdi] (`r min(hdi$year)` - `r max(hdi$year)`; n = `r printnum(length(unique(hdi$Country)))` nations), gross domestic product per capita (`r min(gdp$year)` - `r max(gdp$year)`; n = `r printnum(length(unique(gdp$iso2)))` nations), annual percentage growth in gross domestic product per capita (`r min(gdpGrowth$year)` - `r max(gdpGrowth$year)`; n = `r printnum(length(unique(gdpGrowth$iso2)))` nations), and the Gini coefficient of income inequality (`r min(gini$year)` - `r max(gini$year)`; n = `r printnum(length(unique(gini$iso2)))` nations).
+Human development data were retrieved from the United Nations Development Programme (https://hdr.undp.org/en/content/download-data) and data for all other economic development variables were retrieved from the World Bank (https://data.worldbank.org/).
+For cultural values variables, we retrieved longitudinal data on traditional vs.
 secular values and survival vs.
-self-expression values from the World Values Survey[@Inglehart2000].
+self-expression values from the World Values Survey[@Inglehart2000] (1981 - 2019; n = `r printnum(length(unique(wvs$S009)))` nations).
 We downloaded the full Integrated Values Survey, which included all waves from the World Values Survey and the European Values Survey, and computed the two dimensions of cultural values following procedures from previous research[@Inglehart2000].
-This longitudinal dataset captures values and attitudes for `r printnum(length(unique(wvs$S009)))` nations since 1981 (n = `r printnum(as.numeric(nrow(wvs)), digits = 0)`; https://www.worldvaluessurvey.org/WVSEVStrend.jsp).
+Additionally, we retrieved cross-sectional data on cultural tightness (n = `r printnum(nrow(tightness))` nations) and individualism (n = `r printnum(nrow(fincherData))` nations) from previous work[@Gelfand2021;@Fincher2008].
 
@@
 We also excluded analyses that compared only two, three, four, five, or six nations.
-For each included analysis, we recorded the year, outcome variable, all predictor variables, test statistic, p-value, number of nations, number of data points, model type, if the data were available, and whether and how the analysis attempted to control for non-independence.
+For each included analysis, we recorded the year, impact factor of the journal (retrieved from https://jcr.clarivate.com/jcr/home), outcome variable, all predictor variables, test statistic, p-value, number of nations, number of data points, model type, if the data were available, and whether and how the analysis attempted to control for non-independence.
 
@@
 We also calculated the proportion of articles employing the different types of control listed above at least once: regional fixed effects, distance, shared cultural history, or other.
-For these proportions, we calculated 95% boostrap confidence intervals with 1,000 bootstrap iterations.
+For these proportions, we calculated 95% bootstrap confidence intervals with 1,000 bootstrap iterations.
+Additionally, we predicted the probability of an article attempting to control for non-independence at least once using Bayesian logistic regression, including in separate models log journal impact factor and year of publication as linear and spline predictors, respectively.
 
@@
 We fitted these models separately for overall attempts to control for non-independence and split by method type.
-We report the adjusted proportions with 95% credible intervals.
-To test for a trend over time, we also fitted a Bayesian multilevel logistic regression with a multigroup spline for year of publication and random intercepts for articles.
-Bayesian models were fitted with the _brms_ R package[@Burkner2017].
+We report the adjusted proportions with 95% credible intervals (see Results and Supplementary Figure \@ref(fig:plotReview2)).
+Additionally, we predicted the probability of an analysis attempting to control for non-independence using Bayesian multilevel logistic regression with random intercepts for articles.
+In separate models, we included log journal impact factor and year of publication as linear and spline predictors, respectively.
+All Bayesian models were fitted with the _brms_ R package[@Burkner2017].
 Our priors were informed by prior predictive checks, and all models converged normally ($\hat{R}$ < 1.1).
@@
 
-We simulated data for 236 nations with varying degrees of spatial or cultural phylogenetic signal for outcome $y$ and predictor $x$ using the following generative model:
+```{r, echo=FALSE}
+tar_load(geneticDistances)
+```
 
-$$
-\begin{aligned}
-y &= \alpha_y + \epsilon_y \\
-x &= \alpha_x + \epsilon_x \\
-\alpha_y &\sim \mathcal{N}(0, \sqrt{\lambda} \cdot \Sigma) \\
-\alpha_x &\sim \mathcal{N}(0, \sqrt{\rho} \cdot \Sigma) \\
-\epsilon_y &\sim \mathcal{N}(0, \sqrt{1 - \lambda}) \\
-\epsilon_x &\sim \mathcal{N}(0, \sqrt{1 - \rho})
-\end{aligned}
-$$
+We simulated data for 236 nations $i$ with varying degrees of spatial or cultural phylogenetic signal for outcome $y$ and predictor $x$ using the following generative model:
 
-where $\Sigma$ is a correlation matrix proportional to either geographic or linguistic proximities between nations, and $\lambda$ and $\rho$ are autocorrelation parameters that represent the expected spatial or cultural phylogenetic signal for outcome and predictor variables, respectively.
-Importantly, in this simulation, we know that there is no direct causal relationship between $y$ and $x$ because we simulate the variables independently.
+\begin{gather*}
+\begin{bmatrix}y_i\\x_i \end{bmatrix} \sim \text{MVNormal}
+\begin{pmatrix}\begin{bmatrix}\alpha_y\\\alpha_x \end{bmatrix},\textbf{S}\end{pmatrix}\\
+\alpha_y \sim \text{Normal}(0, \sqrt{\lambda} \cdot \Sigma) \\
+\alpha_x \sim \text{Normal}(0, \sqrt{\rho} \cdot \Sigma) \\
+\textbf{S} = 
+\begin{pmatrix}\sqrt{1 - \lambda} & 0 \\ 0 & \sqrt{1 - \rho} \end{pmatrix}
+\begin{pmatrix}1 & r \\ r & 1\end{pmatrix}
+\begin{pmatrix}\sqrt{1 - \lambda} & 0 \\ 0 & \sqrt{1 - \rho} \end{pmatrix}
+\\
+\end{gather*}
+
+where $\Sigma$ is a correlation matrix proportional to either geographic or linguistic proximities between nations, $\lambda$ and $\rho$ are autocorrelation parameters that represent the expected spatial or cultural phylogenetic signal for outcome and predictor variables, respectively, and $r$ is the true cross-national correlation between the variables after accounting for autocorrelation.
+Importantly, when $r = 0$ in this simulation, we know that there is no direct causal relationship between $y$ and $x$.
 Instead, any relationship between the two variables is merely the result of autocorrelation.
 
-We set the autocorrelation parameters to either 0.2 (weak), 0.5 (moderate), or 0.8 (strong).
-We simulated 100 datasets for each parameter combination, resulting in 900 datasets.
-Each dataset had 236 rows representing different nations, with the following associated data for each nation: latitude, longitude, continent (Africa, Asia, Europe, North America, Oceania, or South America), and language family of the nation's majority spoken language (Afro-Asiatic, Atlantic-Congo, Austroasiatic, Austronesian, Eskimo-Aleut, Indo-European, Japonic, Kartvelian, Koreanic, Mande, Mongolic-Khitan, Nilotic, Nuclear Trans New Guinea, Sino-Tibetan, Tai-Kadai, Tupian, Turkic, or Uralic).
+We set the autocorrelation parameters $\lambda$ and $\rho$ to either 0.2 (weak), 0.5 (moderate), or 0.8 (strong).
+We also initially set the true cross-national correlation to 0 in order to determine false positive rates, and then additionally set $r$ to 0.1 (small effect), 0.3 (medium effect), and 0.5 (large effect) in order to determine statistical power to detect true effects.
+For each parameter combination, we simulated 100 datasets, resulting in 3600 datasets.
+Each dataset had 236 rows representing different nations, with the following associated data for each nation: latitude, longitude, continent (Africa, Asia, Europe, North America, Oceania, or South America), language family of the nation's majority spoken language (Afro-Asiatic, Atlantic-Congo, Austroasiatic, Austronesian, Eskimo-Aleut, Indo-European, Japonic, Kartvelian, Koreanic, Mande, Mongolic-Khitan, Nilotic, Nuclear Trans New Guinea, Sino-Tibetan, Tai-Kadai, Tupian, Turkic, or Uralic), the mean of the predictor variable within a 2000km radius, and coordinates for genetic distances from a previous study[@Schulz2019] (only available for `r printnum(geneticDistances %>% drop_na() %>% nrow())` nations).
 
-With the resulting simulated datasets, we standardised outcome and predictor variables and fitted eleven different models: (1) naive regression without controls, (2) regression with latitude control, (3) regression with longitude control, (4) regression with continent fixed effects, (5) regression with language family fixed effects, (6) regression employing Conley standard errors with 100 km cutoff, (7) regression employing Conley standard errors with 1,000 km cutoff, (8) regression employing Conley standard errors with 10,000 km cutoff, (9) Bayesian regression including a Gaussian process over latitudes and longitudes, (10) Bayesian regression including random intercepts covarying according to linguistic proximity, and (11) Bayesian regression including both a Gaussian process over latitudes and longitudes and random intercepts covarying according to linguistic proximity.
+With the resulting simulated datasets, we standardised outcome and predictor variables and fitted eleven different models: (1) naive regression without controls, (2) regression with latitude control, (3) regression with longitude control, (4) regression with continent fixed effects, (5) regression with language family fixed effects, (6) regression controlling for the mean of the predictor variable in a 2000km radius, (7) regression employing Conley standard errors based on geographic distances, (8) regression employing Conley standard errors based on genetic distances, (9) Bayesian regression including a Gaussian process over latitudes and longitudes, (10) Bayesian regression including random intercepts covarying according to linguistic proximity, and (11) Bayesian regression including both a Gaussian process over latitudes and longitudes and random intercepts covarying according to linguistic proximity.
 
-Models employing Conley standard errors required latitude and longitude values, and the cutoffs implied the distance beyond which autocorrelation is assumed to be zero.
+Models employing Conley standard errors either required latitude and longitude values or coordinates for genetic distances.
+To determine distance cutoffs, we employed an approach recommended in previous work[@Colella2019]: we fitted models with a range of feasible distance cutoffs and retained the model with the largest standard error for the slope parameter.
 These models were fitted using the _conleyreg_ R package[@Duben2021].
@@
 All models converged normally ($\hat{R}$ < 1.1).
-Across all model types and parameter combinations, we calculated the false positive rate as the proportion of models that estimated slopes with a 95% confidence / credible interval excluding zero.
-We calculated 95% bootstrap confidence intervals for these false positive rates with 1,000 bootstrap iterations.
+Across all model types and parameter combinations, we calculated the false positive rate as the proportion of models that estimated slopes with a 95% confidence / credible interval excluding zero when $r = 0$.
+We calculated statistical power as the proportion of models that estimated slopes with a 95% confidence / credible interval excluding zero when $r > 0$.
+We calculated 95% bootstrap confidence intervals for these false positive rates and statistical power estimates with 1,000 bootstrap iterations.
 
@@
 
-We searched the individual analyses from our systematic review for cross-national correlations with available data.
-We included only analyses for which we were able to replicate the original result (i.e.
+We searched the individual analyses from our systematic review for statistically significant cross-national correlations with available primary or secondary data.
+We restricted our search to one analysis per paper, and searched until we had a set of twelve analyses, six from economic development papers and six from cultural values papers, for which we were able to replicate the original result (i.e.
 find a cross-national correlation with the same sign and roughly the same effect size).
-We restricted our search to one analysis per article, and aimed for an even number of analyses for both economic development and cultural values studies.
 We also ensured that at least one analysis was a multilevel model, with multiple observations per nation.
@@
 Before running any additional models, we pre-registered these twelve analyses on the Open Science Framework on 25^th^ January 2022 (https://osf.io/uywx8/).
+We endeavoured to keep the sample sizes of our reanalyses as close to the original analyses as possible, though there were some deviations (see Supplementary Table \@ref(tab:tableDeviations)).
+Despite these slight deviations from the original analyses, all models reported in Figure \@ref(fig:plotReplications1) are fitted to the same number of data points, meaning that any changes in effect sizes are solely due to controlling for non-independence.
 
@@
 
-All data and code are accesible at our Open Science Framework repository (https://osf.io/uywx8/).
+All data and code are accessible at our Open Science Framework repository (https://osf.io/uywx8/).
 We used the _targets_ R package[@Landau2021] to create a reproducible data analysis pipeline and the _papaja_ R package[@Aust2020] to reproducibly generate the manuscript.
+
+\newpage
+
+# Acknowledgements
+
+This work was supported by a Royal Society of New Zealand Marsden grant (20-UOA123) to QDA.
+
+# Author Contributions
+
+SC and QDA conceived of and designed the study.
+SC curated the data, produced all code for analysis and visualisation, and wrote the original draft of the manuscript.
+TK and QDA developed and compiled the geographic and linguistic distance matrices.
+QDA provided funding and input on manuscript preparation and revision.
+All authors reviewed and edited the final draft of the manuscript.
+
+# Competing Interests
+
+The authors declare no competing interests.
+
+# Research Transparency and Reproducibility
+
+All data and code to reproduce the statistical analyses in this manuscript can be found on the Open Science Framework: https://osf.io/uywx8/
 
@@
 \begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
 <div id="refs" custom-style="Bibliography"></div>

### 2022-12-21 manuscript.tex
resubmission

@@
 {}{\PatchFailed}
-\keywords{spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses\newline\indent Word count: xxxx words}
+\keywords{spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses\newline\indent Word count: 4892 words excl. Methods (Methods: 1877 words)}
 \usepackage{lineno}
@@
   pdftitle={The non-independence of nations and why it matters},
-  pdfauthor={Scott Claessens1 \& Quentin D. Atkinson1,2},
+  pdfauthor={Scott Claessens1, Thanos Kyritsis1, \& Quentin D. Atkinson1,2},
   pdflang={en-EN},
@@
 \title{The non-independence of nations and why it matters}
-\author{Scott Claessens\textsuperscript{1} \& Quentin D. Atkinson\textsuperscript{1,2}}
+\author{Scott Claessens\textsuperscript{1}, Thanos Kyritsis\textsuperscript{1}, \& Quentin D. Atkinson\textsuperscript{1,2}}
 \date{}
@@
 
+\par
+
+Correspondence concerning this article should be addressed to Scott Claessens (\href{mailto:scott.claessens@gmail.com}{\nolinkurl{scott.claessens@gmail.com}}) and Quentin D.
+Atkinson (\href{mailto:q.atkinson@auckland.ac.nz}{\nolinkurl{q.atkinson@auckland.ac.nz}}), Floor 2, Building 302, 23 Symonds Street, Auckland, 1010, New Zealand.
+
 The authors made the following contributions.
-Scott Claessens: Data curation, Formal analysis, Investigation, Software, Visualization, Writing - original draft, Writing - review \& editing; Quentin D.
-Atkinson: Conceptualization, Funding acquisition, Supervision, Writing - review \& editing.
-
-Correspondence concerning this article should be addressed to Quentin D.
-Atkinson, Floor 2, Building 302, 23 Symonds Street, Auckland, 1010, New Zealand.
-E-mail: \href{mailto:q.atkinson@auckland.ac.nz}{\nolinkurl{q.atkinson@auckland.ac.nz}}
+Scott Claessens: Conceptualization, Data curation, Formal analysis, Investigation, Software, Visualization, Writing - original draft, Writing - review \& editing; Thanos Kyritsis: Data curation, Writing - review \& editing; Quentin D.
+Atkinson: Conceptualization, Data curation, Funding acquisition, Supervision, Writing - review \& editing.
 
@@
 \abstract{%
-Cross-national analyses test hypotheses about the drivers of global variation in national outcomes.
+Cross-national analyses test hypotheses about the drivers of variation in national outcomes.
 However, since nations are connected in various ways, such as via spatial proximity and shared cultural ancestry, cross-national analyses often violate assumptions of non-independence, inflating false positive rates.
@@
 The standard practice for cross-national analyses is to conduct bivariate correlations or multiple regressions with individual data points representing different nations.
-Such analyses widen the scope of social science beyond Western populations\textsuperscript{7,8} and have been used to study, among other topics, the causes of variation in the economic wealth of nations\textsuperscript{9--12}, global patterings of cultural norms and values\textsuperscript{13--16}, and the universality and diversity of human behaviour and psychology around the world\textsuperscript{17--20}.
+Such analyses widen the scope of social science beyond Western populations\textsuperscript{7,8} and have been used to study, among other topics, the causes of variation in the economic wealth of nations\textsuperscript{9--12}, global patternings of cultural norms and values\textsuperscript{13--16}, and the universality and diversity of human behaviour and psychology around the world\textsuperscript{17--20}.
 
@@
 For example, the related island nations Tonga and Tuvalu share similar languages and customs due to cultural inheritance from a common Polynesian population dating back more than 1,000 years.
-More recently, shared ancestry explains a myriad of cultural similarities between colonial settlements and their colonisers (e.g.~Argentina and Spain).
+More recently, shared ancestry explains similarities in institutions, norms, technologies, and values between colonial settlements and their colonisers (e.g.~English, French, Spanish, and Portuguese settlements of the Americas).
 Importantly, these deep cultural connections between nations often span large geographic distances around the world.
-Tonga and Tuvalu share cultural traits despite being separated by over 1,500 kilometres of ocean, and Argentina and Spain remain culturally similar today despite their location on two separate continents.
+Tonga and Tuvalu share cultural traits despite being separated by over 1,500 kilometres of ocean, and South American and European nations remain culturally similar today despite their locations on two separate continents.
 Shared cultural ancestry must therefore be considered independently of spatial proximity in the study of nations.
@@
 In the twentieth century, anthropologists began to emphasise that human societies do not develop independently, but rather exist in a globally interconnected ``world system'' linked by shared history and cultural ancestry\textsuperscript{33}.
-Researchers compiled the Standard Cross-Cultural Sample of 186 cultures to minimise the confounding effects of this non-independence in comparisons of human societies\textsuperscript{34}, though spatial and cultural dependencies are difficult to remove entirely\textsuperscript{35,36}.
+In order to minimise the confounding effects of this non-independence in comparisons of human societies, researchers compiled the Standard Cross-Cultural Sample of 186 cultures which deliberately avoids sampling closely related cultures\textsuperscript{34}, though it is difficult to completely remove spatial and cultural dependencies\textsuperscript{35,36} and the smaller sample of cultures reduces statistical power.
 Today, anthropologists borrow phylogenetic comparative methods from evolutionary biology, such as phylogenetic least squares regression\textsuperscript{37}, when comparing societies, treating culturally related societies in the same way as biologists treat genetically related species (e.g.\textsuperscript{38,39}).
@@
 These variables are frequently included as both outcomes and predictors in cross-national studies in economics and psychology\textsuperscript{9--16}.
-First, we show that economic development and cultural values are spatially and culturally non-independent across nations, emphasising the need to control for non-independence.
-Second, we systematically review the 100 highest-cited cross-national studies of economic development and cultural values and estimate the proportion of cross-national analyses within these articles that account for non-independence between nations.
+First, we demonstrate that economic development and cultural values are spatially and culturally non-independent across nations, emphasising the need to control for non-independence.
+Second, we systematically review the 100 highest-cited cross-national studies of economic development and cultural values and determine baseline rates of controlling for non-independence in the literature.
 Third, we run simulations to determine whether common methods of dealing with non-independence in the literature sufficiently reduce false positive rates.
-Fourth, we reanalyse twelve previous cross-national analyses of economic development and cultural values from our systematic review, incorporating global geographic and linguistic proximity matrices to correctly control for spatial and cultural non-independence.
+Fourth, we reanalyse twelve previous cross-national analyses of economic development and cultural values from our systematic review, incorporating global geographic and linguistic proximity matrices as controls for spatial and cultural non-independence.
 
@@
 In order to motivate our research question, it is important to first quantify the degree of spatial and cultural non-independence for economic development and cultural values around the world.
-If these variables are only weakly non-independent, then the issue might be safe to ignore.
-However, if they are strongly non-independent, then there is a possibility that non-independence could be confounding cross-national inferences.
-
-To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for a range of economic development and cultural values variables that are widely used in the literature.
-For economic development, we focused on the Human Development Index\textsuperscript{45}, gross domestic product per capita, gross domestic product per capita growth, and the Gini index of income inequality.
+If these variables are independent or only weakly non-independent, then the issue might be safe to ignore.
+However, if they are more strongly non-independent, then there is a possibility that non-independence could be confounding cross-national inferences.
+
+To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for a range of economic development and cultural values variables.
+For economic development, we focused on the Human Development Index\textsuperscript{45}, gross domestic product per capita, annual gross domestic product per capita growth, and the Gini index of income inequality.
 For cultural values, we focused on two primary dimensions of cultural values from the World Values Survey, traditional vs.~secular values and survival vs.~self-expression values\textsuperscript{16}, as well as cultural tightness\textsuperscript{14} and individualism\textsuperscript{15}.
-
-For all of these variables, we found that a substantial proportion of national-level variation was explained by spatial proximity and/or shared cultural ancestry between nations (Figure \ref{fig:plotSignal}; Supplementary Table \ref{tab:tableSignal}).
+These variables are not intended to be a comprehensive list of all national-level variables included in cross-national research, but rather an illustrative set of variables that are widely used in the literature.
+
+For all of these variables, we found that a substantial proportion of national-level variation was explained by spatial proximity and/or shared cultural ancestry between nations (Figure \ref{fig:plotSignal}; see Supplementary Table \ref{tab:tableSignal} for numerical results).
 Signal estimates were often markedly strong, with spatial proximity and shared cultural ancestry frequently explaining over half of the national-level variation.
@@
 Point ranges represent proportions and 95\% bootstrap confidence intervals.
-Lines and shaded areas are posterior median regression lines and 50\% credible intervals from Bayesian multilevel models.
+Lines and shaded areas are posterior median regression lines and 50\% credible intervals from Bayesian logistic regression models.
+Dots represent raw counts of individual articles that did (top) or did not (bottom) account for non-independence.
 Region FEs = region fixed effects.}
@@
 
-While our review contains a range of articles from low- and high-impact journals, an anonymous reviewer suggested that our estimates could be biased downwards by analyses published in lower impact outlets with more relaxed standards for issues like non-independence.
+While our review contains articles from journals with a range of impact factors, an anonymous reviewer suggested that our estimates could be biased downwards by analyses published in lower impact outlets with more relaxed standards for issues like non-independence.
 It is also possible that, since our systematic review goes back as far as 1993, our estimates are being biased downwards by earlier studies, and that controls for non-independence have increased over time with methodological advancements and greater awareness of the issue.
-To test these possible explanations for our low estimates, we fitted Bayesian logistic regression models to the data from the review, including log journal impact factor and publication year as separate predictors.
-Interestingly, we found that, contrary to the explanation above, studies from higher impact journals were \emph{less} likely to include at least one control for non-independence than studies from lower impact journals, both for studies of economic development (b = -0.38, 95\% CI {[}-0.87 0.09{]}) and for studies of cultural values (b = -0.56, 95\% CI {[}-1.10 -0.05{]}; Figure \ref{fig:plotReview1}b).
+To test these possible explanations for our low estimates, we fitted logistic regression models to the data from the review, including log journal impact factor and publication year as separate predictors.
+Interestingly, we found that, if anything, studies from higher impact journals were \emph{less} likely to include at least one control for non-independence than studies from lower impact journals, both for studies of economic development (b = -0.38, 95\% CI {[}-0.87 0.09{]}) and for studies of cultural values (b = -0.56, 95\% CI {[}-1.10 -0.05{]}; Figure \ref{fig:plotReview1}b).
 Moreover, splines revealed no relationship between publication year and the probability of including at least one control for non-independence, both for studies of economic development (b = -0.13, 95\% CI {[}-1.10 0.82{]}) and for studies of cultural values (b = -0.02, 95\% CI {[}-1.00 0.96{]}; Figure \ref{fig:plotReview1}c).
@@
 Despite not being identified in our systematic review, we also included other methods that are often used in the literature to account for non-independence.
-Additional spatial controls included the mean of the predictor variable within a surrounding 2000km radius (e.g.\textsuperscript{46}) and Conley standard errors\textsuperscript{47,48} based on geographic distances between nations (e.g.\textsuperscript{46,\textbf{Schulz2021?}}).
-Additional cultural controls included fixed effects for the language families of the majority-spoken languages in each nation (e.g.\textsuperscript{49}) and Conley standard errors based on genetic distances between nations (e.g.\textsuperscript{46,\textbf{Schulz2021?}}).
-These frequentist approaches attempt to account for non-independence by holding geographic location constant (latitude, longitude), discarding between-region variation and exploiting only local variation (continent fixed effects, mean of surrounding 2000km), or correcting standard errors for autocorrelation post-hoc while leaving model coefficients unchanged (Conley standard errors).
-
-Beyond frequentist approaches, we also fitted Bayesian multilevel regressions that explicitly model spatial and/or cultural phylogenetic non-independence by allowing nations to covary according to geographic and/or linguistic proximity matrices.
+Additional spatial controls included the mean of the predictor variable within a surrounding 2000km radius (e.g.\textsuperscript{46}) and Conley standard errors\textsuperscript{47,48} based on geographic distances between nations (e.g.\textsuperscript{19,46}).
+Additional cultural controls included fixed effects for the language families of the majority-spoken languages in each nation (e.g.\textsuperscript{49}) and Conley standard errors based on genetic distances between nations (e.g.\textsuperscript{19,46}).
+These fixed effect approaches attempt to account for non-independence by holding geographic location constant (latitude, longitude), discarding between-region variation and exploiting only local variation (continent fixed effects, mean of surrounding 2000km), or correcting standard errors for autocorrelation post-hoc while leaving model coefficients unchanged (Conley standard errors).
+
+Beyond fixed effect approaches, we also fitted Bayesian random effects regressions that explicitly model spatial and/or cultural phylogenetic non-independence by allowing nations to covary according to geographic and/or linguistic proximity matrices.
 Geographic proximity between nations is calculated from inverse distances between longitude and latitude coordinates.
 Linguistic proximity between nations is calculated from a global phylogenetic tree that represents hierarchical relationships of genealogical descent for all languages in the world.
-For each pair of nations, we calculate inverse phylogenetic distances (i.e.~number of branches separating two taxa) between all languages spoken in that nation pair and produce an average ``linguistic proximity'' score weighted by the percentages of speakers within those nations.
+For each pair of nations, we calculated inverse phylogenetic distances (i.e.~number of branches separating two taxa) between all languages spoken in that nation pair and produced an average ``linguistic proximity'' score weighted by the percentages of speakers within those nations.
 To include the resulting geographic and linguistic proximity matrices in our models, we included a Gaussian process\textsuperscript{50,51} over latitude and longitude values and/or assumed that nation random intercepts were correlated in proportion to their linguistic proximity\textsuperscript{52}.
-These approaches attempt to account for non-independence by modelling the covariance between nations that is induced by their geographic or linguistic connections.
-
-Figures \ref{fig:plotSim3} and \ref{fig:plotSim4} plot the estimated false positive rates from our simulation study, split by different methods and different degrees of spatial or cultural phylogenetic autocorrelation (see Supplementary Figures \ref{fig:plotSim1} and \ref{fig:plotSim2} for full distributions of effect sizes under strong autocorrelation).
-Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero (i.e.~falsely infering a relationship when none is present).
+These random effects approaches attempt to account for non-independence by modelling the covariance between nations that is induced by their geographic or linguistic connections.
+
+Figures \ref{fig:plotSim3} and \ref{fig:plotSim4} plot the estimated false positive rates from our simulation study, split by different methods and different degrees of spatial or cultural phylogenetic autocorrelation (see Supplementary Tables \ref{tab:tableSim1} and \ref{tab:tableSim2} for numerical results and Supplementary Figures \ref{fig:plotSim1} and \ref{fig:plotSim2} for full distributions of effect sizes under strong autocorrelation).
+Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero, falsely inferring a relationship when none is present.
 For reference, ``weak'' autocorrelation in our simulation is comparable to the geographic signal for survival values in Figure \ref{fig:plotSignal} (i.e.~20\% of the national-level variance is explained by non-independence), while ``moderate'' and ``strong'' levels of autocorrelation are comparable to the cultural phylogenetic signal for traditional and survival values, respectively (i.e.~50\% and 80\% of the national-level variance is explained by non-independence).
@@
 This false positive rate increases as the degree of autocorrelation increases.
-With strong spatial autocorrelation for both outcomes and predictors, false positive rates reach as high as 78\%.
-We find a slightly lower false positive rate under strong cultural phylogenetic autocorrelation, though this false positive rate is still greater than expected by chance (39\%).
+With strong spatial autocorrelation for both outcomes and predictors, false positive rates reach as high as 77\%.
+We find a slightly lower false positive rate under strong cultural phylogenetic autocorrelation, though this false positive rate is still greater than expected by chance (36\%).
 
@@
 
-Common methods in the literature do not reduce these high false positive rates.
+Most methods common in the literature do not reduce these high false positive rates.
 With strong spatial autocorrelation for both outcome and predictor variables, false positive rates remain above 50\% when controlling for latitude, longitude, and language family fixed effects (Figure \ref{fig:plotSim3}).
-Applying Conley standard errors also does not reduce false positive rates below 50\% under strong spatial autocorrelation, regardless of the distance cutoff.
-Continent fixed effects are more effective than other frequentist methods, though they continue to produce a false positive rate of 56\% under strong spatial autocorrelation.
-By contrast, Bayesian spatial Gaussian process regression with longitude and latitude values outperforms all other methods.
-This approach eliminates false positives under moderate spatial autocorrelation, such that the false positive rate is no different from chance, and reduces the false positive rate under strong spatial autocorrelation to 54\%.
-Bayesian models that additionally account for linguistic proximity between nations perform equally well, though models with only linguistic covariance continue to produce false positives.
-
-In our simulation of cultural phylogenetic non-independence, we find that most frequentist methods do not reduce false positive rates (Figure \ref{fig:plotSim4}).
-Controls for latitude and longitude, continent fixed effects, and Conley standard errors do little to change false positive rates.
-Language family fixed effects are slightly more effective than other frequentist methods, though they continue to produce a false positive rate of 39\% under strong cultural phylogenetic autocorrelation.
-By contrast, Bayesian models with random effects covarying according to linguistic proximity completely eliminate false positives across all degrees of cultural phylogenetic autocorrelation.
-Bayesian models that additionally account for geographic proximity between nations perform equally well, though models with only a spatial Gaussian process continue to produce false positives.
+Similarly, Conley standard error corrections based on spatial and genetic distances do not reduce false positive rates below 40\% under strong spatial autocorrelation.
+The most effective fixed effects methods are continent fixed effects, which continue to produce a false positive rate of 35\% under strong spatial autocorrelation, and controlling for the mean of the predictor variable within a 2000km radius, which eliminates false positives under even strong spatial autocorrelation (6\%).
+However, additional simulations revealed that these reductions in false positive rates come at the cost of lower statistical power (see Supplementary Figures \ref{fig:plotPowerGeo01} - \ref{fig:plotPowerGeo05}).
+In additional simulations where the true relationship between the predictor and outcome variable was known, power analyses showed that both continent fixed effects and the 2000km radius control had less than 80\% power to detect moderate true correlations (\emph{r} = 0.3) under strong spatial autocorrelation.
+
+By contrast, Bayesian spatial Gaussian process regression with longitude and latitude strikes a balance between reducing false positives and retaining high statistical power to detect true effects.
+This approach reduces false positives to 15\% under moderate spatial autocorrelation and 23\% under strong spatial autocorrelation.
+Random effects models that additionally account for linguistic proximity between nations perform equally well, though models with only linguistic covariance continue to produce false positives.
+False positives are not completely eliminated with these random effects models.
+Nevertheless, these methods have at least 80\% power to detect moderate (\emph{r} = 0.3) and large (\emph{r} = 0.5) true correlations between variables under all levels of spatial autocorrelation (see Supplementary Figures \ref{fig:plotPowerGeo01} - \ref{fig:plotPowerGeo05}).
+
+In our simulation of cultural phylogenetic non-independence, we find that none of the fixed effects methods reduce false positive rates (Figure \ref{fig:plotSim4}).
+Controls for latitude, longitude, continent fixed effects, the mean of the predictor variable in a 2000km radius, and Conley standard error corrections based on spatial and genetic distances do little to change false positive rates.
+Even language family fixed effects continue to produce a false positive rate of 32\% under strong cultural phylogenetic autocorrelation.
+By contrast, models with random effects covarying according to linguistic proximity completely eliminate false positives across all degrees of cultural phylogenetic autocorrelation.
+This approach is also the only method that is able to detect large true correlations (\emph{r} = 0.5) with at least 80\% power (see Supplementary Figures \ref{fig:plotPowerLin01} - \ref{fig:plotPowerLin05}).
+Random effects models that additionally account for geographic proximity between nations perform equally well, though models with only a spatial Gaussian process continue to produce false positives.
 
@@
 
-Our systematic review and simulation study have shown that controls for non-independence are rare in cross-national studies of economic development and cultural values, and when studies do attempt to control for non-independence, the methods typically used are likely to continue to produce false positives.
+Our systematic review and simulation study have shown that controls for non-independence are rare in cross-national studies of economic development and cultural values, and when studies do attempt to control for non-independence, the methods typically used are unable to sufficiently reduce false positive rates.
 This raises the worrying possibility that the cross-national literature in economics and psychology is populated with spurious relationships.
 
-To determine how widespread this issue of spurious cross-national relationships might be, we reanalysed a subset of twelve previous cross-national analyses from our systematic review, sufficiently controlling for spatial and cultural phylogenetic non-independence using global geographic and linguistic proximity matrices.
-We subsampled six analyses from our economic development review\textsuperscript{53--58} and six from our cultural values review\textsuperscript{13,14,16,59--61}.
-Our choice of analyses was constrained by data availability and whether we were able to initially replicate the original finding.
-We pre-registered our subsample of analyses before running any control models (\url{https://osf.io/uywx8/}).
+To determine how widespread this issue of spurious cross-national relationships might be, we reanalysed a subset of twelve previous cross-national analyses from our systematic review, controlling for spatial and cultural phylogenetic non-independence using global geographic and linguistic proximity matrices.
+Out of the 100 papers included in our systematic review, primary or secondary data were publicly available for 47 papers.
+We attempted to replicate key statistically significant cross-national correlations from these papers --- mostly initial bivariate regression specifications without covariates (see Methods) --- and stopped when we had sampled a set of twelve analyses for which we were able to replicate the original result.
+The final set included six analyses from our economic development review\textsuperscript{53--58} and six from our cultural values review\textsuperscript{13,14,16,59--61} that had available data and were able to be replicated.
+We pre-registered this set before running any control models (\url{https://osf.io/uywx8/}).
 We controlled for non-independence by including (1) a Gaussian process allowing nation random intercepts to covary according to a geographic proximity matrix, and/or (2) nation random intercepts that covaried according to a linguistic proximity matrix (see Supplementary Methods for full models).
 
-Figure \ref{fig:plotReplications1} visualises the results of our reanalysis.
+Figure \ref{fig:plotReplications1} visualises the results of our reanalysis (see Supplementary Table \ref{tab:tableReplications} for numerical results).
 Cross-national correlation effect sizes tended to reduce when controlling for statistical non-independence between nations, sometimes by as much as half of the original effect size.
@@
 Most analyses are simple bivariate cross-national correlations, but Gelfand et al.~(2011) is a partial correlation controlling for log gross national income and Adamczyk and Pitt (2009) is a multilevel model including several covariates.
-Points and black lines represent posterior medians and 95\% credible intervals.
+Coloured ridges are full posterior distributions, and points and black lines represent posterior medians and 95\% credible intervals.
 GDP = gross domestic product.
@@
 
-To understand why some cross-national correlations were attenuated by controls for non-independence while others were robust, we further explored our fitted models for evidence of spatial and cultural autocorrelation.
-For each outcome variable, our Gaussian process models provided varying estimates of how quickly spatial autocorrelation declined with distance (Supplementary Figure \ref{fig:plotReplications3}).
-For example, in Skidmore and Toya\textsuperscript{58} gross domestic product growth was only moderately spatially autocorrelated at 1,000 km distance (posterior median spatial autocorrelation at 1,000 km = 0.42, 95\% CI {[}0.07 0.90{]}), whereas in Inglehart and Baker\textsuperscript{16} traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1,000 km = 0.96, 95\% CI {[}0.81 0.99{]}).
-We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \ref{fig:plotReplications4}), with some outcome variables expressing low signal (e.g.~tightness\textsuperscript{14}; posterior median = 0.15, 95\% CI {[}0.00 0.85{]}) and others expressing high signal (e.g.~female labour force participation\textsuperscript{60}; posterior median = 0.89, 95\% CI {[}0.63 0.98{]}).
-Across all analyses, we found that stronger estimates of spatial autocorrelation or cultural phylogenetic signal resulted in a more pronounced reduction in the effect size when controlling for non-independence between nations (Figure \ref{fig:plotReplications5}).
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/plotReplications5-1.pdf}
-\caption{\label{fig:plotReplications5}\emph{The estimated degree of spatial and cultural phylogenetic non-independence predicts reductions in effect size in our reanalysis.} (a) Higher estimated degrees of spatial autocorrelation at 1,000 km distance predict more pronounced reductions in effect sizes when controlling for non-independence.
-(b) Higher estimated levels of cultural phylogenetic signal predict more pronounced reductions in effect sizes when controlling for non-independence.
-In both panels, the y-axis represents the ratio of the effect size when controlling for spatial and cultural non-independence to the original effect size (from naive regression model), and the x-axis represents posterior median model estimates.
-Regression lines are plotted with 95\% confidence intervals.}
-\end{figure}
-
 \hypertarget{discussion}{%
@@
 
-In a systematic literature review and simulation, we found that cross-national studies in economics and psychology rarely account for non-independence between nations, and, when they do, the methods they use continue to produce false positives.
-In a reanalysis of twelve cross-national correlations, we further showed that neglecting to account for non-independence has resulted in spurious relationships in the published literature, with half of the correlations failing to replicate when controlling for spatial or cultural non-independence with more rigorous methods.
+In a systematic literature review and simulation, we found that cross-national studies in economics and psychology rarely account for non-independence between nations, and, when they do, the methods they use are insufficient to reduce false positives in non-independent data.
+In a reanalysis of twelve cross-national correlations, we further showed that neglecting to account for non-independence has resulted in potentially spurious relationships in the published literature, with half of the correlations failing to replicate when controlling for spatial or cultural non-independence with more rigorous methods.
 These findings suggest that cross-national analyses in economics and psychology should be interpreted with caution until non-independence is sufficiently accounted for.
 
-Our initial analyses add to and clarify existing evidence regarding the non-independence of economic and cultural variation among nations.
+Our initial analyses add to and clarify existing evidence regarding the degree of non-independence for national-level economic and cultural variables.
 One previous study suggested that geographic proximity is more important than deep cultural ancestry in explaining the distribution of human development across Eurasian nations, though the authors noted that their small sample of 44 nations and regional focus limited their statistical power\textsuperscript{62}.
-Our global sample of nations revealed strong cultural phylogenetic signal, as well as geographic signal, for the Human Development Index.
+By contrast, our global samples of over 160 nations revealed strong cultural phylogenetic signal, as well as geographic signal, for the Human Development Index, GDP per capita, and the Gini index of inequality.
 Another previous study found that similarities in the cultural values of nations are predicted by linguistic, but not geographic, distances between those nations\textsuperscript{6}.
-We find this same result for survival vs.~self-expression values, but for traditional vs.~secular values we find that both linguistic and geographic proximity are important independent predictors of global variation.
+We find this same result for survival vs.~self-expression values, cultural tightness, and individualism, but for traditional vs.~secular values we find that both linguistic and geographic proximity are important independent predictors of global variation.
 These findings emphasise the need to account for both spatial and cultural phylogenetic non-independence in cross-national studies of economic development and cultural values.
@@
 In our simulations, controlling for either latitude or longitude did not reduce false positive rates.
-This result calls into question controls like distance to the equator to account for non-independence in cross-national regression models, though these controls may still be suitable to account for regional or latitudinal variation in ecology, which we did not simulate.
-High false positive rates persisted with Conley standard errors, which have previously been critiqued for being overly sensitive to arbitrary distance cutoffs\textsuperscript{40}.
+This result calls into question the use of controls like distance to the equator to account for non-independence in cross-national regression models, though these controls may still be suitable to account for regional or latitudinal variation in climate, ecology, and natural threats (e.g.~pathogens), which we did not simulate.
+High false positive rates persisted with Conley standard errors, despite recent claims that these standard error corrections are sufficient to deal with spatial non-independence\textsuperscript{63}.
 The simulation also confirmed the assertion that fixed effects for spatial or cultural groupings (e.g.~continent or language family fixed effects) are insufficient because non-independence still remains within groupings\textsuperscript{41}.
-This logic further applies to analyses that control for non-independence by separately analysing different regions (e.g.\textsuperscript{63}).
-Across all model types in our simulation, the only methods that sufficiently reduced the false positive rate were the Bayesian multilevel regressions that explicitly modelled spatial and cultural phylogenetic autocorrelation, though we did not include other possible controls for non-independence, such as conditional autoregressive models\textsuperscript{31} or generalised additive models\textsuperscript{64}.
+This logic further applies to analyses that control for non-independence by separately analysing different regions (e.g.\textsuperscript{64}).
+Controlling for the mean of the predictor variable within a 2000km radius\textsuperscript{46} eliminated false positive rates in spatially autocorrelated data, but had reduced statistical power to detect true associations.
+Across all model types in our simulation, the only methods that markedly reduced false positive rates while retaining high statistical power were the random effects models with covariance matrices.
+The important advantage of these models is not that they are Bayesian \emph{per se} (any of the approaches used here could be implemented in a Bayesian framework), but rather that they explicitly model covariance as a function of spatial or cultural distance.
+
+There are other approaches to controlling for non-independence that we did not include in our simulation.
+For example, conditional autoregressive models\textsuperscript{31} and generalised additive models\textsuperscript{65} are approaches that can be applied in both frequentist and Bayesian frameworks.
+There are also alternative ways to operationalise cultural distances between nations, including metrics like cultural fixation indices (e.g.~cultural \(F_{ST}\)\textsuperscript{6,66}), covariance based on genetic distances\textsuperscript{67}, and phylogenetic distances between religious traditions.
+These metrics allow for cultural ancestry effects beyond ethnolinguistic genealogies, though some may produce less interpretable results than language ancestry.
+For example, unlike language, genes are strongly correlated with geography\textsuperscript{68,69} and their global patternings are consistent with different underlying processes, such as deep cultural inheritance and more recent gene flow between populations.
+Future work should explore the interpretability of these methods and determine whether they are sufficient to reduce false positive rates in spatially and culturally non-independent data.
+In addition to spatial proximity and shared cultural ancestry, we did not simulate other sources of non-independence that potentially exist in real cross-national datasets, such as modern connections between nations due to flows of people and information (e.g.~flight networks, social media networks) and shared histories of colonialism and capitalist modes of production that have shaped today's global landscape\textsuperscript{70,71}.
+Additional controls will be required to ensure that these sources of non-independence do not confound cross-national inferences.
 
 Ours is not the first review to show that studies are misapplying statistical methods in ways that inflate false positive rates.
-For example, other literature reviews have shown that studies in the social sciences tend to use small samples of participants\textsuperscript{65}, treat ordinal data as metric\textsuperscript{66}, incorrectly handle missing values\textsuperscript{67}, and ignore best practices in meta-analyses\textsuperscript{68}.
+For example, other literature reviews have shown that studies in the social sciences tend to use small samples of participants\textsuperscript{72}, treat ordinal data as metric\textsuperscript{73}, incorrectly handle missing values\textsuperscript{74}, and ignore best practices in meta-analyses\textsuperscript{75}.
 Why do cross-national studies also rarely account for non-independence?
-At the institutional level, one possibility is that such practices are incentivised \emph{because} they generate statistically significant relationships, which increase the probability that a study is published\textsuperscript{65}.
-At the individual level, another possibility is that researchers outside of anthropology and ecology are simply not aware of the problem, or believe that the problem does not apply to analyses of nations.
+At the institutional level, one possibility is that such practices are incentivised \emph{because} they generate statistically significant relationships, which increase the probability that a study is published\textsuperscript{72}.
+Indeed, we found that controls for non-independence were \emph{less} common among articles published in high impact journals, suggesting that researchers are rewarded for such practices.
+At the individual level, another possibility is that researchers outside of anthropology and ecology are less aware of the problem, or believe that the problem does not apply to analyses of nations.
 Even if researchers appreciate the problem, they might not know of suitable controls or perceive the methods to be too complex.
@@
 These institutional- and individual-level barriers can be combatted.
-First, cross-national replication studies like ours and others\textsuperscript{40--44}, combined with the methodological reviews included in Registered Reports\textsuperscript{69}, might change incentive structures and encourage researchers to analyse the world's nations with more rigorous methods.
-Second, since the issue of non-independence is fundamentally an issue of causal inference (Supplementary Figure \ref{fig:dag}), more explicit descriptions of causal models could promote controls for non-independence in cross-national research.
-In our review, economists studying economic development dealt with national-level non-independence more than psychologists studying cultural values, likely because economics studies tend to be lengthy statistical exercises that systematically incorporate or exclude numerous variables in an attempt to infer causation.
-Third, the recent widespread accessibility of open source statistical software, such as the programming language Stan\textsuperscript{70} and the R package \emph{brms}\textsuperscript{71}, should promote the use of more rigorous methods to control for non-independence.
+First, cross-national replication studies like ours and others\textsuperscript{40--44}, combined with the methodological reviews included in Registered Reports\textsuperscript{76}, might change incentive structures and encourage researchers to analyse the world's nations with more rigorous methods.
+Second, more explicit descriptions of causal models could promote controls for non-independence by clearly outlining the nature of confounding and the sources of autocorrelation in cross-national data\textsuperscript{77}.
+The causal model outlined in Supplementary Figure \ref{fig:dag} is a useful example, but individual studies must outline their own particular causal assumptions, which may include further sources of non-independence and confounding variables to control for (e.g.~post-communist status, colony status).
+These causal models can then be used to design tailored statistical estimation strategies.
+Indeed, in our review, economists studying economic development dealt with national-level non-independence more than psychologists studying cultural values, likely because economics studies tend to be lengthy statistical exercises that systematically incorporate or exclude numerous variables in an attempt to infer causation.
+Third, the recent widespread accessibility of open source statistical software, such as the programming language Stan\textsuperscript{78} and the R package \emph{brms}\textsuperscript{79}, should promote the use of more rigorous methods to control for non-independence.
 Using \emph{brms}, for example, Bayesian Gaussian process regression is straightforward to conduct, requiring only longitude and latitude values for nations.
+We have provided an online tutorial to help researchers apply these methods to their own cross-cultural datasets (\url{https://scottclaessens.github.io/blog/2022/crossnational/}).
 
@@
 Three of these analyses were tests of ``persistence'' hypotheses, studying the effects of historical and environmental conditions --- settler mortality\textsuperscript{53}, wheat-sugar suitability\textsuperscript{57}, and natural disaster frequency\textsuperscript{58} --- on modern developmental outcomes.
-A recent reanalysis has also called into question various studies of this ilk\textsuperscript{40}.
+These results are in line with a recent paper that has called into question various studies of this ilk\textsuperscript{40} and contradict claims that the concern over spatial non-independence has been overblown in these studies\textsuperscript{63}.
 We also found that two out of six cross-national correlations with cultural values variables had 95\% credible intervals that included zero when controlling for cultural phylogenetic non-independence.
+Overall, these reanalyses show that the problem of statistical non-independence applies to a wider range of national-level variables than those identified by previous work, such as parasite stress and democratic outcomes\textsuperscript{41}, raising the worrying possibility that the current cross-national literature is populated with spurious relationships.
+Future work should expand our set of reanalyses to determine the extent of this problem in the literature.
 
@@
 On the contrary, such work promises to deepen understanding of our world, including the causes and consequences of economic development and cultural values.
-Moreover, cross-national studies allow social scientists to broaden their scope of study beyond Western populations\textsuperscript{7}, providing the representative samples necessary to test evolutionary and socio-ecological theories of human behaviour\textsuperscript{8,72}.
+Moreover, cross-national studies allow social scientists to broaden their scope of study beyond Western populations\textsuperscript{7}, providing the representative samples necessary to test evolutionary and socio-ecological theories of human behaviour\textsuperscript{8,80}.
 But in order to minimise spurious relationships in global datasets, we urge researchers to control for spatial and cultural phylogenetic non-independence when reporting cross-national correlations.
@@
 To estimate the degree of spatial and cultural phylogenetic non-independence in economic development and cultural values, we calculated geographic and cultural phylogenetic signal for global measures of development and values.
-Our measure of economic development was the Human Development Index\textsuperscript{45}.
-We retrieved a longitudinal dataset capturing human development for 189 nations since 1990 (n = 1,512; \url{https://hdr.undp.org/en/content/download-data}).
-Our measures of cultural values were traditional vs.~secular values and survival vs.~self-expression values from the World Values Survey\textsuperscript{16}.
+For economic development variables, we retrieved longitudinal data on the Human Development Index\textsuperscript{45} (1990 - 2019; n = 189 nations), gross domestic product per capita (1960 - 2021; n = 209 nations), annual percentage growth in gross domestic product per capita (1961 - 2021; n = 208 nations), and the Gini coefficient of income inequality (1967 - 2021; n = 167 nations).
+Human development data were retrieved from the United Nations Development Programme (\url{https://hdr.undp.org/en/content/download-data}) and data for all other economic development variables were retrieved from the World Bank (\url{https://data.worldbank.org/}).
+For cultural values variables, we retrieved longitudinal data on traditional vs.~secular values and survival vs.~self-expression values from the World Values Survey\textsuperscript{16} (1981 - 2019; n = 116 nations).
 We downloaded the full Integrated Values Survey, which included all waves from the World Values Survey and the European Values Survey, and computed the two dimensions of cultural values following procedures from previous research\textsuperscript{16}.
-This longitudinal dataset captures values and attitudes for 116 nations since 1981 (n = 645,249; \url{https://www.worldvaluessurvey.org/WVSEVStrend.jsp}).
+Additionally, we retrieved cross-sectional data on cultural tightness (n = 57 nations) and individualism (n = 97 nations) from previous work\textsuperscript{13,81}.
 
@@
 We included these matrices in Bayesian multilevel models, allowing nation random intercepts to covary according to both geographic and linguistic proximity simultaneously.
-These models were fitted with the R package \emph{brms}\textsuperscript{71} and converged normally (\(\hat{R}\) \textless{} 1.1).
+These models were fitted with the R package \emph{brms}\textsuperscript{79} and converged normally (\(\hat{R}\) \textless{} 1.1).
 Estimates of geographic and cultural phylogenetic signal were computed as the proportion of national-level variance in these models explained by geographic and linguistic proximity matrices.
@@
 We also excluded analyses that compared only two, three, four, five, or six nations.
-For each included analysis, we recorded the year, outcome variable, all predictor variables, test statistic, p-value, number of nations, number of data points, model type, if the data were available, and whether and how the analysis attempted to control for non-independence.
+For each included analysis, we recorded the year, impact factor of the journal (retrieved from \url{https://jcr.clarivate.com/jcr/home}), outcome variable, all predictor variables, test statistic, p-value, number of nations, number of data points, model type, if the data were available, and whether and how the analysis attempted to control for non-independence.
 
 We coded common attempts to control for non-independence between nations.
-These included: (1) any higher-level control variables for spatial regional groupings (e.g.~continent fixed effects); (2) any geographic distance control variables (e.g.~distance between capital cities, distance from equator, latitude); (3) any control variables capturing shared cultural history (e.g.~former colony, legal origin fixed effects, linguistic history, cultural influence); and (4) any other control variables, tests, or approaches that were deemed as attempts to control for non-independence (e.g.~eigenvector filtering\textsuperscript{73}, controls for trade-weightings between nations, cross-sectional dependence tests\textsuperscript{74}, separate analyses for subsets of nations).
+These included: (1) any higher-level control variables for spatial regional groupings (e.g.~continent fixed effects); (2) any geographic distance control variables (e.g.~distance between capital cities, distance from equator, latitude); (3) any control variables capturing shared cultural history (e.g.~former colony, legal origin fixed effects, linguistic history, cultural influence); and (4) any other control variables, tests, or approaches that were deemed as attempts to control for non-independence (e.g.~eigenvector filtering\textsuperscript{82}, controls for trade-weightings between nations, cross-sectional dependence tests\textsuperscript{83}, separate analyses for subsets of nations).
 These were coded by the first author.
@@
 We also calculated the proportion of articles employing the different types of control listed above at least once: regional fixed effects, distance, shared cultural history, or other.
-For these proportions, we calculated 95\% boostrap confidence intervals with 1,000 bootstrap iterations.
+For these proportions, we calculated 95\% bootstrap confidence intervals with 1,000 bootstrap iterations.
+Additionally, we predicted the probability of an article attempting to control for non-independence at least once using Bayesian logistic regression, including in separate models log journal impact factor and year of publication as linear and spline predictors, respectively.
 
@@
 We fitted these models separately for overall attempts to control for non-independence and split by method type.
-We report the adjusted proportions with 95\% credible intervals.
-To test for a trend over time, we also fitted a Bayesian multilevel logistic regression with a multigroup spline for year of publication and random intercepts for articles.
-Bayesian models were fitted with the \emph{brms} R package\textsuperscript{71}.
+We report the adjusted proportions with 95\% credible intervals (see Results and Supplementary Figure \ref{fig:plotReview2}).
+Additionally, we predicted the probability of an analysis attempting to control for non-independence using Bayesian multilevel logistic regression with random intercepts for articles.
+In separate models, we included log journal impact factor and year of publication as linear and spline predictors, respectively.
+All Bayesian models were fitted with the \emph{brms} R package\textsuperscript{79}.
 Our priors were informed by prior predictive checks, and all models converged normally (\(\hat{R}\) \textless{} 1.1).
@@
 
-We simulated data for 236 nations with varying degrees of spatial or cultural phylogenetic signal for outcome \(y\) and predictor \(x\) using the following generative model:
-
-\[
-\begin{aligned}
-y &= \alpha_y + \epsilon_y \\
-x &= \alpha_x + \epsilon_x \\
-\alpha_y &\sim \mathcal{N}(0, \sqrt{\lambda} \cdot \Sigma) \\
-\alpha_x &\sim \mathcal{N}(0, \sqrt{\rho} \cdot \Sigma) \\
-\epsilon_y &\sim \mathcal{N}(0, \sqrt{1 - \lambda}) \\
-\epsilon_x &\sim \mathcal{N}(0, \sqrt{1 - \rho})
-\end{aligned}
-\]
-
-where \(\Sigma\) is a correlation matrix proportional to either geographic or linguistic proximities between nations, and \(\lambda\) and \(\rho\) are autocorrelation parameters that represent the expected spatial or cultural phylogenetic signal for outcome and predictor variables, respectively.
-Importantly, in this simulation, we know that there is no direct causal relationship between \(y\) and \(x\) because we simulate the variables independently.
+We simulated data for 236 nations \(i\) with varying degrees of spatial or cultural phylogenetic signal for outcome \(y\) and predictor \(x\) using the following generative model:
+
+\begin{gather*}
+\begin{bmatrix}y_i\\x_i \end{bmatrix} \sim \text{MVNormal}
+\begin{pmatrix}\begin{bmatrix}\alpha_y\\\alpha_x \end{bmatrix},\textbf{S}\end{pmatrix}\\
+\alpha_y \sim \text{Normal}(0, \sqrt{\lambda} \cdot \Sigma) \\
+\alpha_x \sim \text{Normal}(0, \sqrt{\rho} \cdot \Sigma) \\
+\textbf{S} = 
+\begin{pmatrix}\sqrt{1 - \lambda} & 0 \\ 0 & \sqrt{1 - \rho} \end{pmatrix}
+\begin{pmatrix}1 & r \\ r & 1\end{pmatrix}
+\begin{pmatrix}\sqrt{1 - \lambda} & 0 \\ 0 & \sqrt{1 - \rho} \end{pmatrix}
+\\
+\end{gather*}
+
+where \(\Sigma\) is a correlation matrix proportional to either geographic or linguistic proximities between nations, \(\lambda\) and \(\rho\) are autocorrelation parameters that represent the expected spatial or cultural phylogenetic signal for outcome and predictor variables, respectively, and \(r\) is the true cross-national correlation between the variables after accounting for autocorrelation.
+Importantly, when \(r = 0\) in this simulation, we know that there is no direct causal relationship between \(y\) and \(x\).
 Instead, any relationship between the two variables is merely the result of autocorrelation.
 
-We set the autocorrelation parameters to either 0.2 (weak), 0.5 (moderate), or 0.8 (strong).
-We simulated 100 datasets for each parameter combination, resulting in 900 datasets.
-Each dataset had 236 rows representing different nations, with the following associated data for each nation: latitude, longitude, continent (Africa, Asia, Europe, North America, Oceania, or South America), and language family of the nation's majority spoken language (Afro-Asiatic, Atlantic-Congo, Austroasiatic, Austronesian, Eskimo-Aleut, Indo-European, Japonic, Kartvelian, Koreanic, Mande, Mongolic-Khitan, Nilotic, Nuclear Trans New Guinea, Sino-Tibetan, Tai-Kadai, Tupian, Turkic, or Uralic).
-
-With the resulting simulated datasets, we standardised outcome and predictor variables and fitted eleven different models: (1) naive regression without controls, (2) regression with latitude control, (3) regression with longitude control, (4) regression with continent fixed effects, (5) regression with language family fixed effects, (6) regression employing Conley standard errors with 100 km cutoff, (7) regression employing Conley standard errors with 1,000 km cutoff, (8) regression employing Conley standard errors with 10,000 km cutoff, (9) Bayesian regression including a Gaussian process over latitudes and longitudes, (10) Bayesian regression including random intercepts covarying according to linguistic proximity, and (11) Bayesian regression including both a Gaussian process over latitudes and longitudes and random intercepts covarying according to linguistic proximity.
-
-Models employing Conley standard errors required latitude and longitude values, and the cutoffs implied the distance beyond which autocorrelation is assumed to be zero.
-These models were fitted using the \emph{conleyreg} R package\textsuperscript{75}.
-Bayesian models were fitted using the \emph{brms} R package\textsuperscript{71}.
+We set the autocorrelation parameters \(\lambda\) and \(\rho\) to either 0.2 (weak), 0.5 (moderate), or 0.8 (strong).
+We also initially set the true cross-national correlation to 0 in order to determine false positive rates, and then additionally set \(r\) to 0.1 (small effect), 0.3 (medium effect), and 0.5 (large effect) in order to determine statistical power to detect true effects.
+For each parameter combination, we simulated 100 datasets, resulting in 3600 datasets.
+Each dataset had 236 rows representing different nations, with the following associated data for each nation: latitude, longitude, continent (Africa, Asia, Europe, North America, Oceania, or South America), language family of the nation's majority spoken language (Afro-Asiatic, Atlantic-Congo, Austroasiatic, Austronesian, Eskimo-Aleut, Indo-European, Japonic, Kartvelian, Koreanic, Mande, Mongolic-Khitan, Nilotic, Nuclear Trans New Guinea, Sino-Tibetan, Tai-Kadai, Tupian, Turkic, or Uralic), the mean of the predictor variable within a 2000km radius, and coordinates for genetic distances from a previous study\textsuperscript{19} (only available for 177 nations).
+
+With the resulting simulated datasets, we standardised outcome and predictor variables and fitted eleven different models: (1) naive regression without controls, (2) regression with latitude control, (3) regression with longitude control, (4) regression with continent fixed effects, (5) regression with language family fixed effects, (6) regression controlling for the mean of the predictor variable in a 2000km radius, (7) regression employing Conley standard errors based on geographic distances, (8) regression employing Conley standard errors based on genetic distances, (9) Bayesian regression including a Gaussian process over latitudes and longitudes, (10) Bayesian regression including random intercepts covarying according to linguistic proximity, and (11) Bayesian regression including both a Gaussian process over latitudes and longitudes and random intercepts covarying according to linguistic proximity.
+
+Models employing Conley standard errors either required latitude and longitude values or coordinates for genetic distances.
+To determine distance cutoffs, we employed an approach recommended in previous work\textsuperscript{84}: we fitted models with a range of feasible distance cutoffs and retained the model with the largest standard error for the slope parameter.
+These models were fitted using the \emph{conleyreg} R package\textsuperscript{85}.
+Bayesian models were fitted using the \emph{brms} R package\textsuperscript{79}.
 Our choice of priors was based on prior predictive simulation.
 All models converged normally (\(\hat{R}\) \textless{} 1.1).
-Across all model types and parameter combinations, we calculated the false positive rate as the proportion of models that estimated slopes with a 95\% confidence / credible interval excluding zero.
-We calculated 95\% bootstrap confidence intervals for these false positive rates with 1,000 bootstrap iterations.
+Across all model types and parameter combinations, we calculated the false positive rate as the proportion of models that estimated slopes with a 95\% confidence / credible interval excluding zero when \(r = 0\).
+We calculated statistical power as the proportion of models that estimated slopes with a 95\% confidence / credible interval excluding zero when \(r > 0\).
+We calculated 95\% bootstrap confidence intervals for these false positive rates and statistical power estimates with 1,000 bootstrap iterations.
 
@@
 
-We searched the individual analyses from our systematic review for cross-national correlations with available data.
-We included only analyses for which we were able to replicate the original result (i.e.~find a cross-national correlation with the same sign and roughly the same effect size).
-We restricted our search to one analysis per article, and aimed for an even number of analyses for both economic development and cultural values studies.
+We searched the individual analyses from our systematic review for statistically significant cross-national correlations with available primary or secondary data.
+We restricted our search to one analysis per paper, and searched until we had a set of twelve analyses, six from economic development papers and six from cultural values papers, for which we were able to replicate the original result (i.e.~find a cross-national correlation with the same sign and roughly the same effect size).
 We also ensured that at least one analysis was a multilevel model, with multiple observations per nation.
@@
 Before running any additional models, we pre-registered these twelve analyses on the Open Science Framework on 25\textsuperscript{th} January 2022 (\url{https://osf.io/uywx8/}).
+We endeavoured to keep the sample sizes of our reanalyses as close to the original analyses as possible, though there were some deviations (see Supplementary Table \ref{tab:tableDeviations}).
+Despite these slight deviations from the original analyses, all models reported in Figure \ref{fig:plotReplications1} are fitted to the same number of data points, meaning that any changes in effect sizes are solely due to controlling for non-independence.
 
@@
 
-We fitted these models using the \emph{brms} R package\textsuperscript{71}.
+We fitted these models using the \emph{brms} R package\textsuperscript{79}.
 Our choice of priors was based on prior predictive simulation.
-All models converged normally (\(\hat{R}\) \textless{} 1.1), though for some models we resorted to using approximate Gaussian processes\textsuperscript{76} to reach convergence.
+All models converged normally (\(\hat{R}\) \textless{} 1.1), though for some models we resorted to using approximate Gaussian processes\textsuperscript{86} to reach convergence.
 
@@
 
-All data and code are accesible at our Open Science Framework repository (\url{https://osf.io/uywx8/}).
-We used the \emph{targets} R package\textsuperscript{77} to create a reproducible data analysis pipeline and the \emph{papaja} R package\textsuperscript{78} to reproducibly generate the manuscript.
+All data and code are accessible at our Open Science Framework repository (\url{https://osf.io/uywx8/}).
+We used the \emph{targets} R package\textsuperscript{87} to create a reproducible data analysis pipeline and the \emph{papaja} R package\textsuperscript{88} to reproducibly generate the manuscript.
+
+\newpage
+
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+This work was supported by a Royal Society of New Zealand Marsden grant (20-UOA123) to QDA.
+
+\hypertarget{author-contributions}{%
+\section{Author Contributions}\label{author-contributions}}
+
+SC and QDA conceived of and designed the study.
+SC curated the data, produced all code for analysis and visualisation, and wrote the original draft of the manuscript.
+TK and QDA developed and compiled the geographic and linguistic distance matrices.
+QDA provided funding and input on manuscript preparation and revision.
+All authors reviewed and edited the final draft of the manuscript.
+
+\hypertarget{competing-interests}{%
+\section{Competing Interests}\label{competing-interests}}
+
+The authors declare no competing interests.
+
+\hypertarget{research-transparency-and-reproducibility}{%
+\section{Research Transparency and Reproducibility}\label{research-transparency-and-reproducibility}}
+
+All data and code to reproduce the statistical analyses in this manuscript can be found on the Open Science Framework: \url{https://osf.io/uywx8/}
 
@@
 \begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
 
@@
 \CSLLeftMargin{22. }%
-\CSLRightInline{Hewlett, B. S., DeSilvestri, A. \& Guglielmino, C. R. \href{https://doi.org/10.1086/339379}{Semes and genes in africa}. \emph{Current Anthropology} \textbf{43}, 313--321 (2002).}
+\CSLRightInline{Hewlett, B. S., DeSilvestri, A. \& Guglielmino, C. R. \href{https://doi.org/10.1086/339379}{Semes and genes in {Africa}}. \emph{Current Anthropology} \textbf{43}, 313--321 (2002).}
 
@@
 \& Kushnick, G.
-\href{https://doi.org/10.1371/journal.pone.0156340}{Cultural and environmental predictors of pre-european deforestation on pacific islands}.
+\href{https://doi.org/10.1371/journal.pone.0156340}{Cultural and environmental predictors of pre-{European} deforestation on {Pacific} islands}.
 \emph{PLOS ONE} \textbf{11}, 1--15 (2016).}
@@
 \CSLLeftMargin{54. }%
-\CSLRightInline{Beck, T., Demirgäç-Kunt, A. \& Levine, R. {SME}s, growth, and poverty: Cross-country evidence. \emph{Journal of economic growth} \textbf{10}, 199--229 (2005).}
+\CSLRightInline{Beck, T., Demirgäç-Kunt, A. \& Levine, R. {SME}s, growth, and poverty: Cross-country evidence. \emph{Journal of Economic Growth} \textbf{10}, 199--229 (2005).}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Voth2021}{}}%
+\CSLLeftMargin{63. }%
+\CSLRightInline{Voth, H.-J.
+Persistence -- myth and mystery.
+in \emph{The handbook of historical economics} (eds.
+Bisin, A.
+\& Federico, G.) 243--267 (Academic Press, 2021).
+doi:\href{https://doi.org/10.1016/B978-0-12-815874-6.00015-0}{10.1016/B978-0-12-815874-6.00015-0}.}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Smith2017}{}}%
-\CSLLeftMargin{63. }%
+\CSLLeftMargin{64. }%
 \CSLRightInline{Smith, M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Hastie2017}{}}%
-\CSLLeftMargin{64. }%
+\CSLLeftMargin{65. }%
 \CSLRightInline{Hastie, T. J. \& Tibshirani, R. J. \emph{Generalized additive models}. (Routledge, 2017).}
 
+\leavevmode\vadjust pre{\hypertarget{ref-Muthukrishna2020}{}}%
+\CSLLeftMargin{66. }%
+\CSLRightInline{Muthukrishna, M.
+\emph{et al.} \href{https://doi.org/10.1177/0956797620916782}{Beyond western, educated, industrial, rich, and democratic (WEIRD) psychology: Measuring and mapping scales of cultural and psychological distance}.
+\emph{Psychological Science} \textbf{31}, 678--701 (2020).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Spolaore2009}{}}%
+\CSLLeftMargin{67. }%
+\CSLRightInline{Spolaore, E. \& Wacziarg, R. The diffusion of development. \emph{The Quarterly Journal of Economics} \textbf{124}, 469--529 (2009).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Handley2007}{}}%
+\CSLLeftMargin{68. }%
+\CSLRightInline{Handley, L. J. L., Manica, A., Goudet, J. \& Balloux, F. Going the distance: Human population genetics in a clinal world. \emph{TRENDS in Genetics} \textbf{23}, 432--439 (2007).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Novembre2008}{}}%
+\CSLLeftMargin{69. }%
+\CSLRightInline{Novembre, J. \emph{et al.} Genes mirror geography within europe. \emph{Nature} \textbf{456}, 98--101 (2008).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Trouillot2016}{}}%
+\CSLLeftMargin{70. }%
+\CSLRightInline{Trouillot, M. \emph{Global transformations: Anthropology and the modern world}. (Springer, 2016).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wolf2010}{}}%
+\CSLLeftMargin{71. }%
+\CSLRightInline{Wolf, E. R. Europe and the people without history. \emph{University of California Press} (2010).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Smaldino2016}{}}%
-\CSLLeftMargin{65. }%
+\CSLLeftMargin{72. }%
 \CSLRightInline{Smaldino, P. E. \& McElreath, R. \href{https://doi.org/10.1098/rsos.160384}{The natural selection of bad science}. \emph{Royal Society Open Science} \textbf{3}, 160384 (2016).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Liddell2018}{}}%
-\CSLLeftMargin{66. }%
+\CSLLeftMargin{73. }%
 \CSLRightInline{Liddell, T.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Nicholson2017}{}}%
-\CSLLeftMargin{67. }%
+\CSLLeftMargin{74. }%
 \CSLRightInline{Nicholson, J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Tipton2019}{}}%
-\CSLLeftMargin{68. }%
+\CSLLeftMargin{75. }%
 \CSLRightInline{Tipton, E., Pustejovsky, J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Chambers2021}{}}%
-\CSLLeftMargin{69. }%
+\CSLLeftMargin{76. }%
 \CSLRightInline{Chambers, C. D. \& Tzavella, L. The past, present and future of {Registered Reports}. \emph{Nature Human Behaviour} \textbf{6}, 29--42 (2021).}
 
+\leavevmode\vadjust pre{\hypertarget{ref-Deffner2022}{}}%
+\CSLLeftMargin{77. }%
+\CSLRightInline{Deffner, D., Rohrer, J.
+M.
+\& McElreath, R.
+\href{https://doi.org/10.1177/25152459221106366}{A causal framework for cross-cultural generalizability}.
+\emph{Advances in Methods and Practices in Psychological Science} \textbf{5}, 25152459221106366 (2022).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-rstan}{}}%
-\CSLLeftMargin{70. }%
+\CSLLeftMargin{78. }%
 \CSLRightInline{Stan Development Team. \href{http://mc-stan.org/}{{RStan}: The {R} interface to {Stan}}. (2020).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Burkner2017}{}}%
-\CSLLeftMargin{71. }%
+\CSLLeftMargin{79. }%
 \CSLRightInline{Bürkner, P.-C.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Oishi2013}{}}%
-\CSLLeftMargin{72. }%
+\CSLLeftMargin{80. }%
 \CSLRightInline{Oishi, S. \href{https://doi.org/10.1146/annurev-psych-030413-152156}{Socioecological psychology}. \emph{Annual Review of Psychology} \textbf{65}, 581--609 (2014).}
 
+\leavevmode\vadjust pre{\hypertarget{ref-Gelfand2021}{}}%
+\CSLLeftMargin{81. }%
+\CSLRightInline{Gelfand, M.
+J.
+\emph{et al.} \href{https://doi.org/10.1016/S2542-5196(20)30301-6}{The relationship between cultural tightness--looseness and {COVID-19} cases and deaths: A global analysis}.
+\emph{The Lancet Planetary Health} \textbf{5}, e135--e144 (2021).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Tiefelsdorf2007}{}}%
-\CSLLeftMargin{73. }%
+\CSLLeftMargin{82. }%
 \CSLRightInline{Tiefelsdorf, M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-DeHoyos2006}{}}%
-\CSLLeftMargin{74. }%
+\CSLLeftMargin{83. }%
 \CSLRightInline{De Hoyos, R.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Colella2019}{}}%
+\CSLLeftMargin{84. }%
+\CSLRightInline{Colella, F., Lalive, R., Sakalli, S.
+O.
+\& Thoenig, M.
+Inference with arbitrary clustering.
+\emph{IZA Discussion Paper No.
+12584} (2019) doi:\href{https://doi.org/10.2139/ssrn.3449578}{10.2139/ssrn.3449578}.}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Duben2021}{}}%
-\CSLLeftMargin{75. }%
+\CSLLeftMargin{85. }%
 \CSLRightInline{Düben, C. \emph{\href{https://CRAN.R-project.org/package=conleyreg}{{conleyreg}: Estimations using {C}onley standard errors}}. (2021).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Solin2020}{}}%
-\CSLLeftMargin{76. }%
+\CSLLeftMargin{86. }%
 \CSLRightInline{Solin, A. \& Särkkä, S. Hilbert space methods for reduced-rank {G}aussian process regression. \emph{Statistics and Computing} \textbf{30}, 419--446 (2020).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
-\CSLLeftMargin{77. }%
+\CSLLeftMargin{87. }%
 \CSLRightInline{Landau, W.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Aust2020}{}}%
-\CSLLeftMargin{78. }%
+\CSLLeftMargin{88. }%
 \CSLRightInline{Aust, F. \& Barth, M. \emph{\href{https://github.com/crsh/papaja}{{papaja}: {Create} {APA} manuscripts with {R Markdown}}}. (2020).}
@@
 
-Geographic distance between two nations was calculated as the logged geodesic distance between country capital cities (data from the R package \emph{maps}; Brownrigg, 2018) using the R package \emph{geosphere} (Hijmans, 2019).
+Geographic distance between two nations was calculated as the logged geodesic distance between nation capital cities (data from the R package \emph{maps}; Brownrigg, 2018) using the R package \emph{geosphere} (Hijmans, 2019).
 The geographic proximity matrix was computed as one minus the log geographic distance matrix scaled between 0 and 1.
@@
 
+\hypertarget{supplementary-results}{%
+\subsection{Supplementary Results}\label{supplementary-results}}
+
+\hypertarget{further-exploration-of-twelve-reanalyses}{%
+\subsubsection{Further exploration of twelve reanalyses}\label{further-exploration-of-twelve-reanalyses}}
+
+To understand why some cross-national correlations were attenuated by controls for non-independence while others were robust, we further explored our fitted models for evidence of spatial and cultural autocorrelation.
+For each outcome variable, our Gaussian process models provided varying estimates of how quickly spatial autocorrelation declined with distance (Supplementary Figure \ref{fig:plotReplications3}).
+For example, in Skidmore and Toya\textsuperscript{58} gross domestic product growth was only moderately spatially autocorrelated at 1,000 km distance (posterior median spatial autocorrelation at 1,000 km = 0.42, 95\% CI {[}0.07 0.90{]}), whereas in Inglehart and Baker\textsuperscript{16} traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1,000 km = 0.96, 95\% CI {[}0.81 0.99{]}).
+We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \ref{fig:plotReplications4}), with some outcome variables expressing low signal (e.g.~confidence in institutions\textsuperscript{61}; posterior median = 0.07, 95\% CI {[}0.00 0.53{]}) and others expressing high signal (e.g.~female labour force participation\textsuperscript{60}; posterior median = 0.89, 95\% CI {[}0.63 0.98{]}).
+Exploratory regressions provided suggestive evidence that stronger estimates of spatial autocorrelation or cultural phylogenetic signal resulted in a more pronounced reduction in the effect size when controlling for non-independence between nations (Supplementary Figure \ref{fig:plotReplications5}).
+However, these negative slopes for spatial autocorrelation (b = -0.19, 95\% CI {[}-1.33 1.04{]}) and cultural phylogenetic signal (b = -0.28, 95\% CI {[}-1.26 0.70{]}) were both very uncertain, due to the small number of analyses and the posterior uncertainty in effect sizes and estimates of non-independence.
+
+\newpage
+
 \hypertarget{supplementary-figures}{%
@@
 But these variables are confounded by their common unobserved cause \(U\).
-\(U\) is a stand-in for shared environmental, ecological, and geographic causes (e.g.~climate, biodiversity, physical topography) and cultural and institutional causes (e.g.~cultural norms, political systems).
+\(U\) is a stand-in for shared environmental, ecological, and geographic causes (e.g.~climate, biodiversity, physical topography) and cultural and institutional causes (e.g.~cultural norms, technologies, and institutions).
 In this causal model, we need to condition on \(U\) to estimate the direct path from \(X\) to \(Y\), but we cannot since it is unobserved.
@@
 Lines and shaded areas are posterior median regression lines and 50\% credible intervals from Bayesian multilevel models.
+Histograms represent relative counts for individual analyses that did (top) or did not (bottom) account for non-independence.
 Region FEs = region fixed effects.}\label{fig:plotReview2}
@@
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotSim1-1} \caption{\emph{Distribution of cross-national correlations from simulation study under strong spatial autocorrelation.} In these simulations, the strength of spatial autocorrelation is set to 0.8 for both outcome and predictor variables.
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotSim1-1} \caption{\emph{Distribution of cross-national correlations from simulation study under strong spatial autocorrelation.} In these simulations, the strength of spatial autocorrelation is set to 0.8 for both outcome and predictor variables, and the true correlation is set to 0.
 For frequentist regression models, point ranges represent correlation estimates and 95\% confidence intervals.
@@
 Correlations are ordered by effect size independently in each panel.
-Red point ranges indicate that the slope is ``significant'' (i.e.~the 95\% confidence / credible interval excludes zero).
-Black point ranges indicate that the slope is ``not significant''.}\label{fig:plotSim1}
+Red point ranges indicate that the 95\% confidence / credible interval excludes zero.
+Black point ranges indicate that the 95\% confidence / credible interval includes zero.}\label{fig:plotSim1}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotSim2-1} \caption{\emph{Distribution of cross-national correlations from simulation study under strong cultural phylogenetic autocorrelation.} In these simulations, the strength of cultural phylogenetic autocorrelation is set to 0.8 for both outcome and predictor variables.
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotSim2-1} \caption{\emph{Distribution of cross-national correlations from simulation study under strong cultural phylogenetic autocorrelation.} In these simulations, the strength of cultural phylogenetic autocorrelation is set to 0.8 for both outcome and predictor variables, and the true correlation is set to 0.
 For frequentist regression models, point ranges represent correlation estimates and 95\% confidence intervals.
@@
 Correlations are ordered by effect size independently in each panel.
-Red point ranges indicate that the slope is ``significant'' (i.e.~the 95\% confidence / credible interval excludes zero).
-Black point ranges indicate that the slope is ``not significant''.}\label{fig:plotSim2}
+Red point ranges indicate that the 95\% confidence / credible interval excludes zero.
+Black point ranges indicate that the 95\% confidence / credible interval includes zero.}\label{fig:plotSim2}
+\end{figure}
+
+\newpage
+
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotPowerGeo01-1} \caption{\emph{Statistical power estimates for different methods of controlling for spatial non-independence in our simulation study, assuming a small true effect size (r = 0.1).} For simulated outcome and predictor variables, we systematically varied the strength of spatial autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+We simulated 100 datasets per parameter combination assuming a small true correlation between variables (\emph{r} = 0.1) and fitted different models to each dataset.
+Statistical power was operationalised as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero.
+Point ranges represent proportions and 95\% bootstrap confidence intervals, and dashed lines indicate 80\% power.
+SEs = standard errors.}\label{fig:plotPowerGeo01}
+\end{figure}
+
+\newpage
+
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotPowerGeo03-1} \caption{\emph{Statistical power estimates for different methods of controlling for spatial non-independence in our simulation study, assuming a medium true effect size (r = 0.3).} For simulated outcome and predictor variables, we systematically varied the strength of spatial autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+We simulated 100 datasets per parameter combination assuming a medium true correlation between variables (\emph{r} = 0.3) and fitted different models to each dataset.
+Statistical power was operationalised as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero.
+Point ranges represent proportions and 95\% bootstrap confidence intervals, and dashed lines indicate 80\% power.
+SEs = standard errors.}\label{fig:plotPowerGeo03}
+\end{figure}
+
+\newpage
+
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotPowerGeo05-1} \caption{\emph{Statistical power estimates for different methods of controlling for spatial non-independence in our simulation study, assuming a large true effect size (r = 0.5).} For simulated outcome and predictor variables, we systematically varied the strength of spatial autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+We simulated 100 datasets per parameter combination assuming a large true correlation between variables (\emph{r} = 0.5) and fitted different models to each dataset.
+Statistical power was operationalised as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero.
+Point ranges represent proportions and 95\% bootstrap confidence intervals, and dashed lines indicate 80\% power.
+SEs = standard errors.}\label{fig:plotPowerGeo05}
+\end{figure}
+
+\newpage
+
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotPowerLin01-1} \caption{\emph{Statistical power estimates for different methods of controlling for cultural phylogenetic non-independence in our simulation study, assuming a small true effect size (r = 0.1).} For simulated outcome and predictor variables, we systematically varied the strength of cultural phylogenetic autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+We simulated 100 datasets per parameter combination assuming a small true correlation between variables (\emph{r} = 0.1) and fitted different models to each dataset.
+Statistical power was operationalised as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero.
+Point ranges represent proportions and 95\% bootstrap confidence intervals, and dashed lines indicate 80\% power.
+SEs = standard errors.}\label{fig:plotPowerLin01}
+\end{figure}
+
+\newpage
+
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotPowerLin03-1} \caption{\emph{Statistical power estimates for different methods of controlling for cultural phylogenetic non-independence in our simulation study, assuming a medium true effect size (r = 0.3).} For simulated outcome and predictor variables, we systematically varied the strength of cultural phylogenetic autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+We simulated 100 datasets per parameter combination assuming a medium true correlation between variables (\emph{r} = 0.3) and fitted different models to each dataset.
+Statistical power was operationalised as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero.
+Point ranges represent proportions and 95\% bootstrap confidence intervals, and dashed lines indicate 80\% power.
+SEs = standard errors.}\label{fig:plotPowerLin03}
+\end{figure}
+
+\newpage
+
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotPowerLin05-1} \caption{\emph{Statistical power estimates for different methods of controlling for cultural phylogenetic non-independence in our simulation study, assuming a large true effect size (r = 0.5).} For simulated outcome and predictor variables, we systematically varied the strength of cultural phylogenetic autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+We simulated 100 datasets per parameter combination assuming a large true correlation between variables (\emph{r} = 0.5) and fitted different models to each dataset.
+Statistical power was operationalised as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero.
+Point ranges represent proportions and 95\% bootstrap confidence intervals, and dashed lines indicate 80\% power.
+SEs = standard errors.}\label{fig:plotPowerLin05}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotReplications3-1} \caption{\emph{Posterior estimates of Gaussian process functions mapping spatial autocorrelation onto geographic distance from our reanalyses of economic development (a) and cultural values (b) studies.} Estimates are from models additionally controlling for cultural phylogenetic non-independence.
-The y-axis represents the amount of spatial autocorrelation between data points with increasing distance between those points on the x-axis (logged distance in kilometres).
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotReplications3-1} \caption{\emph{Posterior estimates of Gaussian process functions mapping spatial autocorrelation onto geographic distance from our reanalyses of economic development (a) and cultural values (b) studies.} The y-axis represents the amount of spatial autocorrelation between data points with increasing distance between those points on the x-axis (logged distance in kilometres).
 Lines and shaded areas represent median posterior spatial autocorrelation functions and 50\% and 95\% credible intervals.}\label{fig:plotReplications3}
@@
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotReplications4-1} \caption{\emph{Posterior estimates of cultural phylogenetic signal from our reanalyses.} Estimates are from models additionally controlling for spatial non-independence.
-Cultural phylogenetic signal is operationalised as the proportion of national-level variance explained by linguistic proximity between nations.
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotReplications4-1} \caption{\emph{Posterior estimates of cultural phylogenetic signal from our reanalyses.} Cultural phylogenetic signal is operationalised as the proportion of national-level variance explained by linguistic proximity between nations.
 Ridges are full posterior distributions, and points are posterior medians, and lines represent 50\% and 95\% credible intervals.}\label{fig:plotReplications4}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotReplications5-1.pdf}
+\caption{\label{fig:plotReplications5}\emph{The estimated degree of spatial and cultural phylogenetic non-independence predicts reductions in effect size in our reanalysis.} (a) Higher estimated degrees of spatial autocorrelation at 1,000 km distance predict more pronounced reductions in effect sizes when controlling for non-independence.
+(b) Higher estimated levels of cultural phylogenetic signal predict more pronounced reductions in effect sizes when controlling for non-independence.
+In both panels, the y-axis represents the ratio of the effect size when controlling for spatial and cultural non-independence to the original effect size (from naive regression model), and the x-axis represents posterior median model estimates.
+Regression lines are plotted with 95\% credible intervals.}
 \end{figure}
@@
 
-\small{
+\footnotesize{
 
@@
 
+\begin{lltable}
+
+\small{
+
+\begin{longtable}{llllllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:tableSim1}False positive rates in simulation with varying levels of spatial autocorrelation for the predictor variable (\(\rho\)) and for the outcome variable (\(\lambda\)), assuming that the true correlation is 0.
+Numbers represent the total number of analyses, out of 100, that estimated a slope with a 95\% confidence / credible interval excluding zero, falsely inferring a relationship when none is present.}\\
+\toprule
+ & \multicolumn{3}{c}{$\rho$ = 0.2} & \multicolumn{3}{c}{$\rho$ = 0.5} & \multicolumn{3}{c}{$\rho$ = 0.8} \\
+\cmidrule(r){2-4} \cmidrule(r){5-7} \cmidrule(r){8-10}
+Model & $\lambda$ = 0.2 & $\lambda$ = 0.5 & $\lambda$ = 0.8 & $\lambda$ = 0.2 & $\lambda$ = 0.5 & $\lambda$ = 0.8 & $\lambda$ = 0.2 & $\lambda$ = 0.5 & $\lambda$ = 0.8\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:tableSim1} continued}}\\
+\toprule
+ & \multicolumn{3}{c}{$\rho$ = 0.2} & \multicolumn{3}{c}{$\rho$ = 0.5} & \multicolumn{3}{c}{$\rho$ = 0.8} \\
+\cmidrule(r){2-4} \cmidrule(r){5-7} \cmidrule(r){8-10}
+Model & $\lambda$ = 0.2 & $\lambda$ = 0.5 & $\lambda$ = 0.8 & $\lambda$ = 0.2 & $\lambda$ = 0.5 & $\lambda$ = 0.8 & $\lambda$ = 0.2 & $\lambda$ = 0.5 & $\lambda$ = 0.8\\
+\midrule
+\endhead
+No control & 18 & 27 & 41 & 27 & 48 & 61 & 40 & 61 & 77\\
+Latitude & 19 & 25 & 39 & 25 & 44 & 60 & 33 & 58 & 71\\
+Longitude & 14 & 20 & 27 & 22 & 34 & 48 & 29 & 49 & 62\\
+Continent & 10 & 13 & 16 & 13 & 18 & 24 & 13 & 25 & 35\\
+Language family & 12 & 17 & 29 & 20 & 33 & 46 & 27 & 48 & 58\\
+Mean 2000km radius & 9 & 12 & 19 & 5 & 9 & 13 & 5 & 5 & 6\\
+Conley SEs spatial & 13 & 13 & 14 & 18 & 23 & 25 & 22 & 37 & 43\\
+Conley SEs genetic & 17 & 22 & 27 & 20 & 34 & 51 & 28 & 49 & 66\\
+Bayesian spatial & 14 & 13 & 11 & 13 & 15 & 12 & 16 & 17 & 23\\
+Bayesian linguistic & 12 & 20 & 21 & 21 & 26 & 39 & 28 & 42 & 49\\
+Bayesian spatial \& linguistic & 11 & 13 & 9 & 14 & 13 & 12 & 17 & 18 & 21\\
+\bottomrule
+\end{longtable}
+
+}
+
+\end{lltable}
+
+\newpage
+
+\begin{lltable}
+
+\small{
+
+\begin{longtable}{llllllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:tableSim2}False positive rates in simulation with varying levels of cultural phylogenetic autocorrelation for the predictor variable (\(\rho\)) and for the outcome variable (\(\lambda\)), assuming that the true correlation is 0.
+Numbers represent the total number of analyses, out of 100, that estimated a slope with a 95\% confidence / credible interval excluding zero, falsely inferring a relationship when none is present.}\\
+\toprule
+ & \multicolumn{3}{c}{$\rho$ = 0.2} & \multicolumn{3}{c}{$\rho$ = 0.5} & \multicolumn{3}{c}{$\rho$ = 0.8} \\
+\cmidrule(r){2-4} \cmidrule(r){5-7} \cmidrule(r){8-10}
+Model & $\lambda$ = 0.2 & $\lambda$ = 0.5 & $\lambda$ = 0.8 & $\lambda$ = 0.2 & $\lambda$ = 0.5 & $\lambda$ = 0.8 & $\lambda$ = 0.2 & $\lambda$ = 0.5 & $\lambda$ = 0.8\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:tableSim2} continued}}\\
+\toprule
+ & \multicolumn{3}{c}{$\rho$ = 0.2} & \multicolumn{3}{c}{$\rho$ = 0.5} & \multicolumn{3}{c}{$\rho$ = 0.8} \\
+\cmidrule(r){2-4} \cmidrule(r){5-7} \cmidrule(r){8-10}
+Model & $\lambda$ = 0.2 & $\lambda$ = 0.5 & $\lambda$ = 0.8 & $\lambda$ = 0.2 & $\lambda$ = 0.5 & $\lambda$ = 0.8 & $\lambda$ = 0.2 & $\lambda$ = 0.5 & $\lambda$ = 0.8\\
+\midrule
+\endhead
+No control & 7 & 14 & 15 & 8 & 19 & 30 & 15 & 26 & 36\\
+Latitude & 7 & 10 & 16 & 10 & 19 & 27 & 15 & 22 & 35\\
+Longitude & 5 & 12 & 14 & 9 & 16 & 26 & 13 & 26 & 35\\
+Continent & 5 & 6 & 10 & 3 & 11 & 21 & 8 & 20 & 29\\
+Language family & 7 & 9 & 12 & 9 & 14 & 22 & 10 & 20 & 32\\
+Mean 2000km radius & 5 & 8 & 12 & 2 & 12 & 21 & 4 & 16 & 25\\
+Conley SEs spatial & 2 & 6 & 7 & 2 & 9 & 15 & 6 & 14 & 20\\
+Conley SEs genetic & 5 & 11 & 13 & 8 & 13 & 21 & 8 & 17 & 26\\
+Bayesian spatial & 7 & 12 & 12 & 9 & 17 & 25 & 14 & 24 & 32\\
+Bayesian linguistic & 5 & 9 & 11 & 6 & 7 & 7 & 4 & 5 & 5\\
+Bayesian spatial \& linguistic & 3 & 10 & 11 & 5 & 8 & 8 & 4 & 5 & 5\\
+\bottomrule
+\end{longtable}
+
+}
+
+\end{lltable}
+
+\newpage
+
+\begin{lltable}
+
+\footnotesize{
+
+\begin{longtable}{rrrrr}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:tableReplications}Estimated cross-national correlations from our reanalysis of twelve previous cross-national analyses.
+Numbers are posterior median slopes (equivalent to Pearson's r correlation coefficients) with 95\% credible intervals.}\\
+\toprule
+Analysis & \multicolumn{1}{c}{No control} & \multicolumn{1}{c}{Spatial control} & \multicolumn{1}{c}{Cultural control} & \multicolumn{1}{c}{Both controls}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:tableReplications} continued}}\\
+\toprule
+Analysis & \multicolumn{1}{c}{No control} & \multicolumn{1}{c}{Spatial control} & \multicolumn{1}{c}{Cultural control} & \multicolumn{1}{c}{Both controls}\\
+\midrule
+\endhead
+Adamczyk and Pitt (2009) & 0.20, 95\% CI [ 0.07,\ \ 0.33] & 0.20, 95\% CI [ 0.07,\ \ 0.34] & 0.19, 95\% CI [ 0.05,\ \ 0.33] & 0.20, 95\% CI [ 0.06,\ \ 0.33]\\
+Alesina et al. (2013) & -0.21, 95\% CI [-0.42,\ \ 0.01] & -0.21, 95\% CI [-0.43,\ \ 0.01] & -0.07, 95\% CI [-0.26,\ \ 0.11] & -0.07, 95\% CI [-0.25,\ \ 0.11]\\
+Beck et al. (2003) & -0.50, 95\% CI [-0.68, -0.30] & -0.15, 95\% CI [-0.38,\ \ 0.02] & -0.56, 95\% CI [-0.77, -0.33] & -0.54, 95\% CI [-0.76, -0.30]\\
+Beck et al. (2005) & 0.38, 95\% CI [ 0.12,\ \ 0.64] & 0.25, 95\% CI [-0.01,\ \ 0.50] & 0.29, 95\% CI [ 0.10,\ \ 0.49] & 0.29, 95\% CI [ 0.09,\ \ 0.48]\\
+Bockstette et al. (2002) & 0.43, 95\% CI [ 0.25,\ \ 0.59] & 0.41, 95\% CI [ 0.19,\ \ 0.59] & 0.37, 95\% CI [ 0.18,\ \ 0.57] & 0.37, 95\% CI [ 0.15,\ \ 0.56]\\
+Easterly and Levine (2003) & 0.75, 95\% CI [ 0.60,\ \ 0.90] & 0.76, 95\% CI [ 0.57,\ \ 0.93] & 0.65, 95\% CI [ 0.47,\ \ 0.82] & 0.66, 95\% CI [ 0.47,\ \ 0.84]\\
+Easterly (2007) & -0.46, 95\% CI [-0.63, -0.29] & -0.18, 95\% CI [-0.36,\ \ 0.01] & -0.25, 95\% CI [-0.41, -0.09] & -0.19, 95\% CI [-0.36, -0.01]\\
+Fincher et al. (2008) & -0.64, 95\% CI [-0.82, -0.46] & -0.59, 95\% CI [-0.80, -0.21] & -0.50, 95\% CI [-0.71, -0.27] & -0.47, 95\% CI [-0.69, -0.18]\\
+Gelfand et al. (2011) & 0.54, 95\% CI [ 0.20,\ \ 0.85] & 0.54, 95\% CI [ 0.19,\ \ 0.86] & 0.49, 95\% CI [ 0.14,\ \ 0.83] & 0.50, 95\% CI [ 0.15,\ \ 0.83]\\
+Inglehart and Baker (2000) & 0.48, 95\% CI [ 0.21,\ \ 0.74] & 0.27, 95\% CI [ 0.11,\ \ 0.44] & 0.36, 95\% CI [ 0.11,\ \ 0.60] & 0.28, 95\% CI [ 0.10,\ \ 0.46]\\
+Knack and Keefer (1997) & 0.36, 95\% CI [ 0.05,\ \ 0.67] & 0.36, 95\% CI [ 0.04,\ \ 0.66] & 0.28, 95\% CI [-0.10,\ \ 0.63] & 0.29, 95\% CI [-0.09,\ \ 0.62]\\
+Skidmore and Toya (2002) & 0.30, 95\% CI [ 0.10,\ \ 0.49] & 0.20, 95\% CI [ 0.00,\ \ 0.41] & 0.21, 95\% CI [ 0.03,\ \ 0.40] & 0.21, 95\% CI [ 0.01,\ \ 0.41]\\
+\bottomrule
+\end{longtable}
+
+}
+
+\end{lltable}
+
+\newpage
+
+\begin{center}
+\begin{ThreePartTable}
+
+\footnotesize{
+
+\begin{longtable}{m{4cm}m{2cm}m{2cm}m{7cm}}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:tableDeviations}List of sample size deviations between original analyses and reanalyses.}\\
+\toprule
+Analysis & \multicolumn{1}{c}{Original N} & \multicolumn{1}{c}{Reanalysis N} & \multicolumn{1}{c}{Reason for deviation}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:tableDeviations} continued}}\\
+\toprule
+Analysis & \multicolumn{1}{c}{Original N} & \multicolumn{1}{c}{Reanalysis N} & \multicolumn{1}{c}{Reason for deviation}\\
+\midrule
+\endhead
+Adamczyk and Pitt (2009) & 33 & 33 & N/A\\
+Alesina et al.
+(2013) & >177 & 75 & Original analysis used imputed values for traditional plough use which were unavailable to us.
+For our reanalysis, we collected data from Tables A4 and A9 and ran our regression on the 75 nations with available data for female labour force participation in 2000 and traditional plough use.\\
+Beck et al. (2003) & 70 & 69 & We removed one nation (Zaire) as it was not present in our linguistic distance matrix.\\
+Beck et al. (2005) & 45 & 45 & N/A\\
+Bockstette et al. (2002) & 94 & 103 & After manually linking state history data from Appendix A with world development indicators, we had complete data for 103 nations.\\
+Easterly and Levine (2003) & 72 & 63 & After manually linking log GDP per capita 1995 data from Appendix Table A2 in Acemoglu et al.
+(2001) with an institutions index self-constructed from world governance indicators, we only had complete data for 63 nations.\\
+Easterly (2007) & 118 & 98 & After manually linking log wheat-sugar ratio data from Appendix A with longitudinal Gini data, we only had complete data for 98 nations.\\
+Fincher et al. (2008) & 68 & 67 & We combined England and Northern Ireland into a single nation (United Kingdom) for the reanalysis.\\
+Gelfand et al.
+(2011) & 30 & 28 & We combined East and West Germany into a single nation (Germany) and removed a high leverage point with a substantially higher disaster rate than other nations (Venezuela).\\
+Inglehart and Baker (2000) & 65 & 38 & After manually linking traditional values data from the World Values Survey with data on percentage total employment in industry from World Development Indicators, we only had complete data for 38 nations.\\
+Knack and Keefer (1997) & 28 & 28 & N/A\\
+Skidmore and Toya (2002) & 89 & 89 & N/A\\
+\bottomrule
+\end{longtable}
+
+}
+
+\end{ThreePartTable}
+\end{center}
+
+\newpage
+
 \hypertarget{supplementary-references}{%

### 2022-12-21 R/functionsReplications.R

@@ -896,7 +896,6 @@ makeTableReplications <- function(slopeListA, slopeListB, slopeListC, slopeListD
   # get median and 95% CI
   printMed95CI <- function(x) {
     paste0(
-      ifelse(round(median(x), 2) >= 0, " ", ""),
       format(round(median(x), 2), nsmall = 2), ", 95% CI [",
       ifelse(round(quantile(x, 0.025), 2) >= 0, " ", ""),
       format(round(quantile(x, 0.025), 2), nsmall = 2), ", ",

### 2022-12-21 R/functionsReview.R

@@ -164,19 +164,60 @@ plotReviewAnalysis <- function(review, yearAnalysis, ifAnalysis,
           axis.ticks.y = element_blank(),
           axis.line.y = element_blank())
   # impact factor results
+  pBdata <- plot(conditional_effects(ifAnalysis, resolution = 500, prob = 0.50), plot = FALSE)[[2]]$data
   pB <-
-    plot(conditional_effects(ifAnalysis, resolution = 500, prob = 0.50), plot = FALSE)[[2]] +
+    ggplot() +
+    stat_dots(
+      data = mutate(
+        review, 
+        ControlNI = ifelse(ControlNI == "Yes", 1, 0),
+        side = ifelse(ControlNI == 0, "top", "bottom"),
+        Review = ifelse(Review == "Cultural\nvalues", "Values", "Economic development")
+        ),
+      aes(y = ControlNI, x = IF, side = side, fill = Review, colour = Review),
+      scale = 0.5
+    ) +
+    geom_ribbon(
+      data = pBdata, 
+      aes(x = IF, ymin = lower__, ymax = upper__, fill = Review),
+      alpha = 0.4
+      ) +
+    geom_line(
+      data = pBdata,
+      aes(x = IF, y = estimate__, colour = Review)
+      ) +
     scale_y_continuous(name = "Probability of analysis\ncontrolling for non-independence", 
-                       limits = c(0, 1)) +
+                       limits = c(0, 1), expand = c(0, 0)) +
     scale_x_log10(name = "Journal impact factor") +
     theme_classic()
   # years published
   years <- c(1995, 2000, 2005, 2010, 2015, 2020)
   # spline results
+  pCdata <- plot(conditional_effects(yearAnalysis, prob = 0.50), plot = FALSE)[[3]]$data
   pC <-
-    plot(conditional_effects(yearAnalysis, prob = 0.50), plot = FALSE)[[3]] +
+    ggplot() +
+    stat_dots(
+      data = mutate(
+        review, 
+        ControlNI = ifelse(ControlNI == "Yes", 1, 0),
+        side = ifelse(ControlNI == 0, "top", "bottom"),
+        Review = ifelse(Review == "Cultural\nvalues", "Values", "Economic development"),
+        Year.std = as.numeric(scale(Year))
+      ),
+      aes(y = ControlNI, x = Year.std, side = side, fill = Review, colour = Review),
+      scale = 0.5
+    ) +
+    geom_ribbon(
+      data = pCdata, 
[…]

### 2022-12-21 R/functionsSimulation.R

@@ -261,15 +261,15 @@ makeTableSimAll <- function(olsModel1, olsModel2, olsModel3, olsModel4, olsModel
               "Language family", "Mean 2000km radius", "Conley SEs spatial",
               "Conley SEs genetic", "Bayesian spatial", "Bayesian linguistic",
               "Bayesian spatial & linguistic"),
-    `Rho = 0.2\nLambda = 0.2` = unlist(lapply(l, getPerc, Rho = 0.2, Lambda = 0.2)),
-    `Rho = 0.2\nLambda = 0.5` = unlist(lapply(l, getPerc, Rho = 0.2, Lambda = 0.5)),
-    `Rho = 0.2\nLambda = 0.8` = unlist(lapply(l, getPerc, Rho = 0.2, Lambda = 0.8)),
-    `Rho = 0.5\nLambda = 0.2` = unlist(lapply(l, getPerc, Rho = 0.5, Lambda = 0.2)),
-    `Rho = 0.5\nLambda = 0.5` = unlist(lapply(l, getPerc, Rho = 0.5, Lambda = 0.5)),
-    `Rho = 0.5\nLambda = 0.8` = unlist(lapply(l, getPerc, Rho = 0.5, Lambda = 0.8)),
-    `Rho = 0.8\nLambda = 0.2` = unlist(lapply(l, getPerc, Rho = 0.8, Lambda = 0.2)),
-    `Rho = 0.8\nLambda = 0.5` = unlist(lapply(l, getPerc, Rho = 0.8, Lambda = 0.5)),
-    `Rho = 0.8\nLambda = 0.8` = unlist(lapply(l, getPerc, Rho = 0.8, Lambda = 0.8))
+    `Rho = 0.2 Lambda = 0.2` = unlist(lapply(l, getPerc, Rho = 0.2, Lambda = 0.2)),
+    `Rho = 0.2 Lambda = 0.5` = unlist(lapply(l, getPerc, Rho = 0.2, Lambda = 0.5)),
+    `Rho = 0.2 Lambda = 0.8` = unlist(lapply(l, getPerc, Rho = 0.2, Lambda = 0.8)),
+    `Rho = 0.5 Lambda = 0.2` = unlist(lapply(l, getPerc, Rho = 0.5, Lambda = 0.2)),
[…]

### 2022-12-21 manuscript.Rmd
add ethics statement

@@
 
-# Research Transparency and Reproducibility
+# Data Availability
 
-All data and code to reproduce the statistical analyses in this manuscript can be found on the Open Science Framework: https://osf.io/uywx8/
+All data can be found on the Open Science Framework: https://osf.io/uywx8/
+
+# Code Availability
+
+All code to reproduce the statistical analyses in this manuscript can be found on the Open Science Framework: https://osf.io/uywx8/
+
+# Ethical Approval
+
+Ethical approval was not required for the study as we analysed only secondary and simulated data.
 

### 2022-12-21 manuscript.tex
add ethics statement

@@
 
-\hypertarget{research-transparency-and-reproducibility}{%
-\section{Research Transparency and Reproducibility}\label{research-transparency-and-reproducibility}}
-
-All data and code to reproduce the statistical analyses in this manuscript can be found on the Open Science Framework: \url{https://osf.io/uywx8/}
+\hypertarget{data-availability}{%
+\section{Data Availability}\label{data-availability}}
+
+All data can be found on the Open Science Framework: \url{https://osf.io/uywx8/}
+
+\hypertarget{code-availability}{%
+\section{Code Availability}\label{code-availability}}
+
+All code to reproduce the statistical analyses in this manuscript can be found on the Open Science Framework: \url{https://osf.io/uywx8/}
+
+\hypertarget{ethical-approval}{%
+\section{Ethical Approval}\label{ethical-approval}}
+
+Ethical approval was not required for the study as we analysed only secondary and simulated data.
 

### 2023-04-04 manuscript.Rmd
second resubmission

@@
   - \captionsetup[figure]{font={stretch=1.5}}
+  - \nolinenumbers
 appendix          : "appendix.Rmd"
@@
 ---
+
+\linenumbers
 
@@
 For example, conditional autoregressive models[@Lichstein2002] and generalised additive models[@Hastie2017] are approaches that can be applied in both frequentist and Bayesian frameworks.
-There are also alternative ways to operationalise cultural distances between nations, including metrics like cultural fixation indices (e.g.
+There are also alternative ways to operationalise cultural distances between nations beyond linguistic distances, including metrics like cultural fixation indices (e.g.
 cultural $F_{ST}$[@Muthukrishna2020;@White2021]), covariance based on genetic distances[@Spolaore2009], and phylogenetic distances between religious traditions.
-These metrics allow for cultural ancestry effects beyond ethnolinguistic genealogies, though some may produce less interpretable results than language ancestry.
-For example, unlike language, genes are strongly correlated with geography[@Handley2007;@Novembre2008] and their global patternings are consistent with different underlying processes, such as deep cultural inheritance and more recent gene flow between populations.
-Future work should explore the interpretability of these methods and determine whether they are sufficient to reduce false positive rates in spatially and culturally non-independent data.
+We see merit in each of these approaches and the use of one over another will depend on the specific question at hand.
+We decided to focus on linguistic distances in this study since language is a tangible socially-learned trait that has previously been used to successfully track the effects of deep cultural ancestry on modern national outcomes[@Kyritsis2022;@Matthews2016;@Sookias2018].
+Future work should explore whether other approaches are sufficient to reduce false positive rates in spatially and culturally non-independent data.
 In addition to spatial proximity and shared cultural ancestry, we did not simulate other sources of non-independence that potentially exist in real cross-national datasets, such as modern connections between nations due to flows of people and information (e.g.
@@
 Until such changes are implemented and sufficient controls for non-independence are the norm, existing cross-national correlations should be interpreted with caution.
-In our reanalyses, we found that four out of six cross-national correlations with economic development variables had 95% credible intervals that included zero when controlling for spatial non-independence.
-Three of these analyses were tests of "persistence" hypotheses, studying the effects of historical and environmental conditions --- settler mortality[@Beck2003], wheat-sugar suitability[@Easterly2007], and natural disaster frequency[@Skidmore2002] --- on modern developmental outcomes.
-These results are in line with a recent paper that has called into question various studies of this ilk[@Kelly2020] and contradict claims that the concern over spatial non-independence has been overblown in these studies[@Voth2021].
-We also found that two out of six cross-national correlations with cultural values variables had 95% credible intervals that included zero when controlling for cultural phylogenetic non-independence.
-Overall, these reanalyses show that the problem of statistical non-independence applies to a wider range of national-level variables than those identified by previous work, such as parasite stress and democratic outcomes[@Bromham2018], raising the worrying possibility that the current cross-national literature is populated with spurious relationships.
+In our reanalyses, we found that half of the cross-national correlations had 95% credible intervals that included zero when controlling for spatial and/or cultural phylogenetic non-independence.
+While these results are in line with previous reanalyses[@Kelly2020], we note that we are unable to outright reject the claims from these studies, since we only reanalysed the first bivariate regression specifications presented in the papers[@Voth2021].
+More detailed sets of reanalyses would be required to comprehensively challenge the claims from these specific papers.
+Nevertheless, these reanalyses do show, more broadly, that the problem of statistical non-independence applies to a wider range of national-level variables than those identified by previous work, such as parasite stress and democratic outcomes[@Bromham2018].
+Moreover, given our finding that most studies in the current cross-national literature do not deal with non-independence at all, our reanalyses raise the worrying possibility that this literature is populated with spurious relationships.
 Future work should expand our set of reanalyses to determine the extent of this problem in the literature.
@@
 \newpage
+\nolinenumbers
 

### 2023-04-04 manuscript.tex
second resubmission

@@
 \captionsetup[figure]{font={stretch=1.5}}
+\nolinenumbers
 \ifLuaTeX
@@
 
+\linenumbers
+
 Nations are perhaps the single most important population unit structuring modern human life.
@@
 For example, conditional autoregressive models\textsuperscript{31} and generalised additive models\textsuperscript{65} are approaches that can be applied in both frequentist and Bayesian frameworks.
-There are also alternative ways to operationalise cultural distances between nations, including metrics like cultural fixation indices (e.g.~cultural \(F_{ST}\)\textsuperscript{6,66}), covariance based on genetic distances\textsuperscript{67}, and phylogenetic distances between religious traditions.
-These metrics allow for cultural ancestry effects beyond ethnolinguistic genealogies, though some may produce less interpretable results than language ancestry.
-For example, unlike language, genes are strongly correlated with geography\textsuperscript{68,69} and their global patternings are consistent with different underlying processes, such as deep cultural inheritance and more recent gene flow between populations.
-Future work should explore the interpretability of these methods and determine whether they are sufficient to reduce false positive rates in spatially and culturally non-independent data.
+There are also alternative ways to operationalise cultural distances between nations beyond linguistic distances, including metrics like cultural fixation indices (e.g.~cultural \(F_{ST}\)\textsuperscript{6,66}), covariance based on genetic distances\textsuperscript{67}, and phylogenetic distances between religious traditions.
+We see merit in each of these approaches and the use of one over another will depend on the specific question at hand.
+We decided to focus on linguistic distances in this study since language is a tangible socially-learned trait that has previously been used to successfully track the effects of deep cultural ancestry on modern national outcomes\textsuperscript{62,68,69}.
+Future work should explore whether other approaches are sufficient to reduce false positive rates in spatially and culturally non-independent data.
 In addition to spatial proximity and shared cultural ancestry, we did not simulate other sources of non-independence that potentially exist in real cross-national datasets, such as modern connections between nations due to flows of people and information (e.g.~flight networks, social media networks) and shared histories of colonialism and capitalist modes of production that have shaped today's global landscape\textsuperscript{70,71}.
@@
 Until such changes are implemented and sufficient controls for non-independence are the norm, existing cross-national correlations should be interpreted with caution.
-In our reanalyses, we found that four out of six cross-national correlations with economic development variables had 95\% credible intervals that included zero when controlling for spatial non-independence.
-Three of these analyses were tests of ``persistence'' hypotheses, studying the effects of historical and environmental conditions --- settler mortality\textsuperscript{53}, wheat-sugar suitability\textsuperscript{57}, and natural disaster frequency\textsuperscript{58} --- on modern developmental outcomes.
-These results are in line with a recent paper that has called into question various studies of this ilk\textsuperscript{40} and contradict claims that the concern over spatial non-independence has been overblown in these studies\textsuperscript{63}.
-We also found that two out of six cross-national correlations with cultural values variables had 95\% credible intervals that included zero when controlling for cultural phylogenetic non-independence.
-Overall, these reanalyses show that the problem of statistical non-independence applies to a wider range of national-level variables than those identified by previous work, such as parasite stress and democratic outcomes\textsuperscript{41}, raising the worrying possibility that the current cross-national literature is populated with spurious relationships.
+In our reanalyses, we found that half of the cross-national correlations had 95\% credible intervals that included zero when controlling for spatial and/or cultural phylogenetic non-independence.
+While these results are in line with previous reanalyses\textsuperscript{40}, we note that we are unable to outright reject the claims from these studies, since we only reanalysed the first bivariate regression specifications presented in the papers\textsuperscript{63}.
+More detailed sets of reanalyses would be required to comprehensively challenge the claims from these specific papers.
+Nevertheless, these reanalyses do show, more broadly, that the problem of statistical non-independence applies to a wider range of national-level variables than those identified by previous work, such as parasite stress and democratic outcomes\textsuperscript{41}.
+Moreover, given our finding that most studies in the current cross-national literature do not deal with non-independence at all, our reanalyses raise the worrying possibility that this literature is populated with spurious relationships.
 Future work should expand our set of reanalyses to determine the extent of this problem in the literature.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Handley2007}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-Kyritsis2022}{}}%
 \CSLLeftMargin{68. }%
-\CSLRightInline{Handley, L. J. L., Manica, A., Goudet, J. \& Balloux, F. Going the distance: Human population genetics in a clinal world. \emph{TRENDS in Genetics} \textbf{23}, 432--439 (2007).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Novembre2008}{}}%
+\CSLRightInline{Kyritsis, T., Matthews, L.
+J., Welch, D.
+\& Atkinson, Q.
+D.
+\href{https://doi.org/10.1017/ehs.2022.40}{Shared cultural ancestry predicts the global diffusion of democracy}.
+\emph{Evolutionary Human Sciences} \textbf{4}, e42 (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Matthews2016}{}}%
 \CSLLeftMargin{69. }%
-\CSLRightInline{Novembre, J. \emph{et al.} Genes mirror geography within europe. \emph{Nature} \textbf{456}, 98--101 (2008).}
+\CSLRightInline{Matthews, L.
+J., Passmore, S., Richard, P.
+M., Gray, R.
+D.
+\& Atkinson, Q.
+D.
+\href{https://doi.org/10.1371/journal.pone.0152979}{Shared cultural history as a predictor of political and economic changes among nation states}.
+\emph{PLOS ONE} \textbf{11}, 1--18 (2016).}
 
@@
 \renewcommand{\theequation}{S\arabic{table}} \setcounter{equation}{0}
+\pagenumbering{gobble}
 

### 2023-07-10 appendix.Rmd
second resubmission
third resubmission

@@
-\renewcommand{\appendixname}{\bf{Supplementary Material}}
 \renewcommand{\figurename}{Supplementary Figure}
 \renewcommand{\tablename}{Supplementary Table}
-\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
-\renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
-\renewcommand{\theequation}{S\arabic{table}} \setcounter{equation}{0}
+\renewcommand{\thefigure}{\arabic{figure}} \setcounter{figure}{0}
+\renewcommand{\thetable}{\arabic{table}} \setcounter{table}{0}
+\renewcommand{\theequation}{S\arabic{equation}} \setcounter{equation}{0}
 
-#
+# **Supplementary Information**
+
+\setcounter{page}{1}
+\centering
+\noindent \small Cross-national analyses require additional controls to account for the non-independence of nations \newline
+\hspace*{1cm} \small Scott Claessens\textsuperscript{1}, Thanos Kyritsis\textsuperscript{1}, & Quentin D. Atkinson\textsuperscript{1,2} \newline
+\raggedright
+\noindent \footnotesize \textsuperscript{1} School of Psychology, University of Auckland, New Zealand \newline
+\noindent \footnotesize \textsuperscript{2} School of Anthropology and Museum Ethnography, University of Oxford, United Kingdom
+\normalsize
+\newpage
 
 ## Supplementary Methods
-
-### Calculating global geographic and linguistic proximity matrices
-
-Geographic distance between two nations was calculated as the logged geodesic distance between nation capital cities (data from the R package _maps_; Brownrigg, 2018) using the R package _geosphere_ (Hijmans, 2019).
-The geographic proximity matrix was computed as one minus the log geographic distance matrix scaled between 0 and 1.
-
-Linguistic proximity between two nations was calculated as the cultural proximity between all languages spoken within those nations, weighted by speaker percentages.
-We acquired cultural proximity data by combining the language family trees provided by Glottolog v3.0 (Hammarström et al., 2017) into one global language tree (undated and unresolved).
-We calculated cultural proximity $s$ between two languages $j$ and $k$ as the distance (in number of nodes traversed) of their most recent common ancestor $i$ to the root of the tree, through the formula:
-
-$$
-s_{jk} = \frac{n_{r}-n_{i}}{n_{r}}
-$$
-
-where $n_{r}$ is the maximum path length (in number of nodes traversed) leading to the pan-human root $r$, and $n_{i}$ is the maximum path length leading to node $i$.
-We then combined these proximities with speaker data from Ethnologue 21 (Ethnologue, 2018) and compared every language spoken within those nations by at least 1 permille of the population, weighted by speaker percentages, through the formula:
-
-$$
-w_{lm} = {\Sigma}{\Sigma}p_{lj}p_{mk}s_{jk}
-$$
-
-where $p_{lj}$ is the percentage of the population in nation $l$ speaking language $j$, $p_{mk}$ is the percentage of the population in nation $m$ speaking language $k$, and $s_{jk}$ is the proximity measure between languages $j$ and $k$ (Eff, 2008).
-The resulting linguistic proximity matrix was also scaled between 0 and 1 before analysis.
 
@@
 
-$$
-\begin{aligned}
+\begin{align}
 Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
-\mu_{i} &= \alpha + \beta X_{i} \\
-\alpha &\sim \text{Normal}(0, 0.4) \\
-\beta &\sim \text{Normal}(0, 0.4) \\
-\sigma &\sim \text{Exponential}(5)
-\end{aligned}
-$$
+\mu_{i} &= \alpha + \beta X_{i} \nonumber \\
+\alpha &\sim \text{Normal}(0, 0.4) \nonumber \\
+\beta &\sim \text{Normal}(0, 0.4) \nonumber \\
+\sigma &\sim \text{Exponential}(5) \nonumber
+\end{align}
+
 The priors in this model were arrived at by prior predictive checks, with wider priors making predictions beyond the scale of standardised outcome variables and narrower priors being too informative.
 
-\newpage
-
 To control for spatial non-independence, we add a Gaussian process to this model and feed it a scaled geographic distance matrix $D$ based on Euclidean distances between latitude and longitude coordinates.
-This distance matrix is computed internally by the R package _brms_ (Bürkner, 2017).
+This distance matrix is computed internally by the _brms_ R package^1^.
 The Gaussian process uses an exponentiated quadratic covariance kernel, the only covariance kernel currently supported by _brms_.
@@
 
-$$
-\begin{aligned}
+\begin{align}
 Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
-\mu_{i} &= \alpha + \kappa_{\text{NATION}[i]} + \beta X_{i} \\
+\mu_{i} &= \alpha + \kappa_{\text{NATION}[i]} + \beta X_{i} \nonumber \\
 \begin{pmatrix}
-\kappa_{1} \\ \kappa_{2} \\ \kappa_{3} \\ ... \\ \kappa_{n}
+\kappa_{1} \\ \kappa_{2} \\  ... \\ \kappa_{n}
 \end{pmatrix} &\sim \text{MVNormal}
@@
 \begin{pmatrix}
-0 \\ 0 \\ 0 \\ ... \\ 0
+0 \\ 0 \\ ... \\ 0
 \end{pmatrix},\textbf{K}
-\end{pmatrix}\\
-\textbf{K}_{ij} &= sdgp^2 \text{exp} \big (-D_{ij}^2 / (2 lscale^2) \big )\\
-\alpha &\sim \text{Normal}(0, 0.4) \\
-\beta &\sim \text{Normal}(0, 0.4) \\
-\sigma &\sim \text{Exponential}(5) \\
-sdgp &\sim \text{Exponential}(5) \\
-lscale &\sim \text{InverseGamma}(?,?)
-\end{aligned}
-$$
+\end{pmatrix} \nonumber \\
+\textbf{K}_{ij} &= sdgp^2 \text{exp} \big (-D_{ij}^2 / (2 lscale^2) \big ) \nonumber \\
+\alpha &\sim \text{Normal}(0, 0.4) \nonumber \\
+\beta &\sim \text{Normal}(0, 0.4) \nonumber \\
+\sigma &\sim \text{Exponential}(5) \nonumber \\
+sdgp &\sim \text{Exponential}(5) \nonumber \\
+lscale &\sim \text{InverseGamma}(?,?) \nonumber 
+\end{align}
+
 where $n$ is the number of nations, and $D^2_{ij}$ reflects the squared Euclidean distances between latitude and longitude coordinates for the $i$-th and $j$-th nations.
@@
 This is because the _brms_ package intelligently tunes the prior for this parameter based on the covariates of the Gaussian process (see https://betanalpha.github.io/assets/case_studies/gp_part3/part3.html).
-
-\newpage
 
@@
 
-$$
-\begin{aligned}
+\begin{align}
 Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
-\mu_{i} &= \alpha + z_{\text{NATION}[i]}\sigma_{\alpha}L + \beta X_{i} \\
-\alpha &\sim \text{Normal}(0, 0.4) \\
-\beta &\sim \text{Normal}(0, 0.4) \\
-z_{j} &\sim \text{Normal}(0, 1)\\
-\sigma_{\alpha} &\sim \text{Exponential}(5) \\
-\sigma &\sim \text{Exponential}(5)
-\end{aligned}
-$$
-\newpage
+\mu_{i} &= \alpha + z_{\text{NATION}[i]}\sigma_{\alpha}L + \beta X_{i} \nonumber \\
+\alpha &\sim \text{Normal}(0, 0.4) \nonumber \\
+\beta &\sim \text{Normal}(0, 0.4) \nonumber \\
+z_{j} &\sim \text{Normal}(0, 1) \nonumber \\
+\sigma_{\alpha} &\sim \text{Exponential}(5) \nonumber \\
+\sigma &\sim \text{Exponential}(5) \nonumber 
+\end{align}
 
@@
 
-$$
-\begin{aligned}
+\begin{align}
 Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
-\mu_{i} &= \alpha + \kappa_{\text{NATION}[i]} + z_{\text{NATION}[i]}\sigma_{\alpha}L + \beta X_{i} \\
+\mu_{i} &= \alpha + \kappa_{\text{NATION}[i]} + z_{\text{NATION}[i]}\sigma_{\alpha}L + \beta X_{i} \nonumber \\
 \begin{pmatrix}
-\kappa_{1} \\ \kappa_{2} \\ \kappa_{3} \\ ... \\ \kappa_{n}
+\kappa_{1} \\ \kappa_{2} \\ ... \\ \kappa_{n}
 \end{pmatrix} &\sim \text{MVNormal}
@@
 \begin{pmatrix}
-0 \\ 0 \\ 0 \\ ... \\ 0
+0 \\ 0 \\ ... \\ 0
 \end{pmatrix},\textbf{K}
-\end{pmatrix}\\
-\textbf{K}_{ij} &= sdgp^2 \text{exp} \big (-D_{ij}^2 / (2 lscale^2) \big )\\
-\alpha &\sim \text{Normal}(0, 0.4) \\
-\beta &\sim \text{Normal}(0, 0.4) \\
-z_{j} &\sim \text{Normal}(0, 1)\\
-\sigma_{\alpha} &\sim \text{Exponential}(5) \\
-\sigma &\sim \text{Exponential}(5) \\
-sdgp &\sim \text{Exponential}(5) \\
-lscale &\sim \text{InverseGamma}(?,?)
-\end{aligned}
-$$
-
-\newpage
-
-## Supplementary Results
-
-### Further exploration of twelve reanalyses
-
-```{r load autocor}
-tar_load(c(inglehartSA1, skidmoreSA1, alesinaSignal1, knackSignal1, effSizeRatioModel1, effSizeRatioModel2))
-```
-
-To understand why some cross-national correlations were attenuated by controls for non-independence while others were robust, we further explored our fitted models for evidence of spatial and cultural autocorrelation.
-For each outcome variable, our Gaussian process models provided varying estimates of how quickly spatial autocorrelation declined with distance (Supplementary Figure \@ref(fig:plotReplications3)).
-For example, in Skidmore and Toya[@Skidmore2002] gross domestic product growth was only moderately spatially autocorrelated at 1,000 km distance (posterior median spatial autocorrelation at 1,000 km = `r printnum(median(skidmoreSA1))`, 95% CI [`r printnum(quantile(skidmoreSA1, 0.025))` `r printnum(quantile(skidmoreSA1, 0.975))`]), whereas in Inglehart and Baker[@Inglehart2000] traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1,000 km = `r printnum(median(inglehartSA1))`, 95% CI [`r printnum(quantile(inglehartSA1, 0.025))` `r printnum(quantile(inglehartSA1, 0.975))`]).
-We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \@ref(fig:plotReplications4)), with some outcome variables expressing low signal (e.g.
-confidence in institutions[@Knack1997]; posterior median = `r printnum(median(knackSignal1))`, 95% CI [`r printnum(quantile(knackSignal1, 0.025))` `r printnum(quantile(knackSignal1, 0.975))`]) and others expressing high signal (e.g.
-female labour force participation[@Alesina2013]; posterior median = `r printnum(median(alesinaSignal1))`, 95% CI [`r printnum(quantile(alesinaSignal1, 0.025))` `r printnum(quantile(alesinaSignal1, 0.975))`]).
-Exploratory regressions provided suggestive evidence that stronger estimates of spatial autocorrelation or cultural phylogenetic signal resulted in a more pronounced reduction in the effect size when controlling for non-independence between nations (Supplementary Figure \@ref(fig:plotReplications5)).
-However, these negative slopes for spatial autocorrelation (b = `r printnum(fixef(effSizeRatioModel1)["corAt1000km","Estimate"])`, 95% CI [`r printnum(fixef(effSizeRatioModel1)["corAt1000km","Q2.5"])` `r printnum(fixef(effSizeRatioModel1)["corAt1000km","Q97.5"])`]) and cultural phylogenetic signal (b = `r printnum(fixef(effSizeRatioModel2)["signal","Estimate"])`, 95% CI [`r printnum(fixef(effSizeRatioModel2)["signal","Q2.5"])` `r printnum(fixef(effSizeRatioModel2)["signal","Q97.5"])`]) were both very uncertain, due to the small number of analyses and the posterior uncertainty in effect sizes and estimates of non-independence.
+\end{pmatrix} \nonumber \\
+\textbf{K}_{ij} &= sdgp^2 \text{exp} \big (-D_{ij}^2 / (2 lscale^2) \big ) \nonumber \\
+\alpha &\sim \text{Normal}(0, 0.4) \nonumber \\
+\beta &\sim \text{Normal}(0, 0.4) \nonumber \\
+z_{j} &\sim \text{Normal}(0, 1) \nonumber \\
+\sigma_{\alpha} &\sim \text{Exponential}(5) \nonumber \\
+\sigma &\sim \text{Exponential}(5) \nonumber \\
+sdgp &\sim \text{Exponential}(5) \nonumber \\
+lscale &\sim \text{InverseGamma}(?,?) \nonumber 
+\end{align}
 
@@
 
-(ref:plotReview2Caption) _Analysis-level results from systematic review of 100 highly-cited cross-national studies of economic development (red) and cultural values (blue)._ (a) Proportion of analyses accounting for non-independence, overall and split by common methods of controlling for non-independence.
+(ref:plotReview2Caption) _Analysis-level results from literature review of 100 highly-cited cross-national studies of economic development (red) and cultural values (blue)._ (a) Proportion of analyses accounting for non-independence, overall and split by common methods of controlling for non-independence.
+Points represent posterior median proportions and ranges represent equal-tailed 95% credible intervals (n = 4308 observations).
 (b) The association between journal impact factor and the probability that an analysis accounts for non-independence.
 (c) Estimated trend over time for the probability that an analysis accounts for non-independence.
-Point ranges represent estimated proportions and 95% credible intervals.
-Lines and shaded areas are posterior median regression lines and 50% credible intervals from Bayesian multilevel models.
+Lines and shaded areas are posterior median regression lines and equal-tailed 50% credible intervals from Bayesian multilevel models (n = 4308 observations).
 Histograms represent relative counts for individual analyses that did (top) or did not (bottom) account for non-independence.
@@
 (ref:plotSim1Caption) _Distribution of cross-national correlations from simulation study under strong spatial autocorrelation._ In these simulations, the strength of spatial autocorrelation is set to 0.8 for both outcome and predictor variables, and the true correlation is set to 0.
-For frequentist regression models, point ranges represent correlation estimates and 95% confidence intervals.
-For Bayesian regression models, point ranges represent posterior means and 95% credible intervals.
+For frequentist regression models, points represent correlation estimates and ranges represent two-tailed 95% confidence intervals (n = 236 observations).
+For Bayesian regression models, points represent posterior means and ranges represent equal-tailed 95% credible intervals (n = 236 observations).
 Correlations are ordered by effect size independently in each panel.
@@
 Black point ranges indicate that the 95% confidence / credible interval includes zero.
+SEs = standard errors.
 
@@
 (ref:plotSim2Caption) _Distribution of cross-national correlations from simulation study under strong cultural phylogenetic autocorrelation._ In these simulations, the strength of cultural phylogenetic autocorrelation is set to 0.8 for both outcome and predictor variables, and the true correlation is set to 0.
-For frequentist regression models, point ranges represent correlation estimates and 95% confidence intervals.
-For Bayesian regression models, point ranges represent posterior means and 95% credible intervals.
+For frequentist regression models, points represent correlation estimates and ranges represent two-tailed 95% confidence intervals (n = 236 observations).
+For Bayesian regression models, points represent posterior means and ranges represent equal-tailed 95% credible intervals (n = 236 observations).
 Correlations are ordered by effect size independently in each panel.
@@
 We simulated 100 datasets per parameter combination assuming a small true correlation between variables (_r_ = 0.1) and fitted different models to each dataset.
-Statistical power was operationalised as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero.
-Point ranges represent proportions and 95% bootstrap confidence intervals, and dashed lines indicate 80% power.
+Statistical power was operationalised as the proportion of models that estimated a slope with a two-tailed 95% confidence / credible interval excluding zero.
+Points represent raw proportions of models with slope 95% confidence / credible intervals excluding zero, ranges represent two-tailed 95% bootstrap confidence intervals (n = 1000 bootstrap samples), and dashed lines indicate 80% power.
+Colours indicate whether the strength of autocorrelation for the predictor variable is 0.2 (red), 0.5, (green) or 0.8 (blue).
 SEs = standard errors.
 
-```{r plotPowerGeo01, fig.width=11, fig.height=7, fig.cap="(ref:plotPowerGeo01Caption)", fig.pos='H', out.width="\\textwidth"}
+```{r plotPowerGeo01, fig.width=7, fig.height=5.5, fig.cap="(ref:plotPowerGeo01Caption)", fig.pos='H', out.width="\\textwidth"}
 tar_read(plotPowerGeo0.1)
@@
 We simulated 100 datasets per parameter combination assuming a medium true correlation between variables (_r_ = 0.3) and fitted different models to each dataset.
-Statistical power was operationalised as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero.
-Point ranges represent proportions and 95% bootstrap confidence intervals, and dashed lines indicate 80% power.
+Statistical power was operationalised as the proportion of models that estimated a slope with a two-tailed 95% confidence / credible interval excluding zero.
+Points represent raw proportions of models with slope 95% confidence / credible intervals excluding zero, ranges represent two-tailed 95% bootstrap confidence intervals (n = 1000 bootstrap samples), and dashed lines indicate 80% power.
+Colours indicate whether the strength of autocorrelation for the predictor variable is 0.2 (red), 0.5, (green) or 0.8 (blue).
 SEs = standard errors.
 
-```{r plotPowerGeo03, fig.width=11, fig.height=7, fig.cap="(ref:plotPowerGeo03Caption)", fig.pos='H', out.width="\\textwidth"}
+```{r plotPowerGeo03, fig.width=7, fig.height=5.5, fig.cap="(ref:plotPowerGeo03Caption)", fig.pos='H', out.width="\\textwidth"}
 tar_read(plotPowerGeo0.3)
@@
 We simulated 100 datasets per parameter combination assuming a large true correlation between variables (_r_ = 0.5) and fitted different models to each dataset.
-Statistical power was operationalised as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero.
-Point ranges represent proportions and 95% bootstrap confidence intervals, and dashed lines indicate 80% power.
+Statistical power was operationalised as the proportion of models that estimated a slope with a two-tailed 95% confidence / credible interval excluding zero.
+Points represent raw proportions of models with slope 95% confidence / credible intervals excluding zero, ranges represent two-tailed 95% bootstrap confidence intervals (n = 1000 bootstrap samples), and dashed lines indicate 80% power.
+Colours indicate whether the strength of autocorrelation for the predictor variable is 0.2 (red), 0.5, (green) or 0.8 (blue).
 SEs = standard errors.
 
-```{r plotPowerGeo05, fig.width=11, fig.height=7, fig.cap="(ref:plotPowerGeo05Caption)", fig.pos='H', out.width="\\textwidth"}
+```{r plotPowerGeo05, fig.width=7, fig.height=5.5, fig.cap="(ref:plotPowerGeo05Caption)", fig.pos='H', out.width="\\textwidth"}
 tar_read(plotPowerGeo0.5)
@@
 We simulated 100 datasets per parameter combination assuming a small true correlation between variables (_r_ = 0.1) and fitted different models to each dataset.
-Statistical power was operationalised as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero.
-Point ranges represent proportions and 95% bootstrap confidence intervals, and dashed lines indicate 80% power.
+Statistical power was operationalised as the proportion of models that estimated a slope with a two-tailed 95% confidence / credible interval excluding zero.
+Points represent raw proportions of models with slope 95% confidence / credible intervals excluding zero, ranges represent two-tailed 95% bootstrap confidence intervals (n = 1000 bootstrap samples), and dashed lines indicate 80% power.
+Colours indicate whether the strength of autocorrelation for the predictor variable is 0.2 (red), 0.5, (green) or 0.8 (blue).
 SEs = standard errors.
 
-```{r plotPowerLin01, fig.width=11, fig.height=7, fig.cap="(ref:plotPowerLin01Caption)", fig.pos='H', out.width="\\textwidth"}
+```{r plotPowerLin01, fig.width=7, fig.height=5.5, fig.cap="(ref:plotPowerLin01Caption)", fig.pos='H', out.width="\\textwidth"}
 tar_read(plotPowerLin0.1)
@@
 We simulated 100 datasets per parameter combination assuming a medium true correlation between variables (_r_ = 0.3) and fitted different models to each dataset.
-Statistical power was operationalised as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero.
-Point ranges represent proportions and 95% bootstrap confidence intervals, and dashed lines indicate 80% power.
+Statistical power was operationalised as the proportion of models that estimated a slope with a two-tailed 95% confidence / credible interval excluding zero.
+Points represent raw proportions of models with slope 95% confidence / credible intervals excluding zero, ranges represent two-tailed 95% bootstrap confidence intervals (n = 1000 bootstrap samples), and dashed lines indicate 80% power.
+Colours indicate whether the strength of autocorrelation for the predictor variable is 0.2 (red), 0.5, (green) or 0.8 (blue).
 SEs = standard errors.
 
-```{r plotPowerLin03, fig.width=11, fig.height=7, fig.cap="(ref:plotPowerLin03Caption)", fig.pos='H', out.width="\\textwidth"}
+```{r plotPowerLin03, fig.width=7, fig.height=5.5, fig.cap="(ref:plotPowerLin03Caption)", fig.pos='H', out.width="\\textwidth"}
 tar_read(plotPowerLin0.3)
@@
 We simulated 100 datasets per parameter combination assuming a large true correlation between variables (_r_ = 0.5) and fitted different models to each dataset.
-Statistical power was operationalised as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero.
-Point ranges represent proportions and 95% bootstrap confidence intervals, and dashed lines indicate 80% power.
+Statistical power was operationalised as the proportion of models that estimated a slope with a two-tailed 95% confidence / credible interval excluding zero.
+Points represent raw proportions of models with slope 95% confidence / credible intervals excluding zero, ranges represent two-tailed 95% bootstrap confidence intervals (n = 1000 bootstrap samples), and dashed lines indicate 80% power.
+Colours indicate whether the strength of autocorrelation for the predictor variable is 0.2 (red), 0.5, (green) or 0.8 (blue).
 SEs = standard errors.
 
-```{r plotPowerLin05, fig.width=11, fig.height=7, fig.cap="(ref:plotPowerLin05Caption)", fig.pos='H', out.width="\\textwidth"}
+```{r plotPowerLin05, fig.width=7, fig.height=5.5, fig.cap="(ref:plotPowerLin05Caption)", fig.pos='H', out.width="\\textwidth"}
 tar_read(plotPowerLin0.5)
@@
 In all reanalyses, outcome and predictor variables are standardised, making regression slopes comparable to Pearson's correlation coefficients.
-Lines and shaded areas represent posterior median regression lines and 95% credible intervals.
+Lines and shaded areas represent posterior median regression lines and equal-tailed 95% credible intervals.
+Colours indicate predictions from models with no control (red), spatial control only (green), cultural control only (blue), or both spatial and cultural control (purple).
 For models with covariates (Adamzyck and Pitt 2009; Gelfand et al.
 2011), marginal effects are presented holding all covariates at zero or their reference categories.
+GDP = gross domestic product; SME = small and medium-sized enterprise.
 
@@
 (ref:plotReplications3Caption) _Posterior estimates of Gaussian process functions mapping spatial autocorrelation onto geographic distance from our reanalyses of economic development (a) and cultural values (b) studies._ The y-axis represents the amount of spatial autocorrelation between data points with increasing distance between those points on the x-axis (logged distance in kilometres).
-Lines and shaded areas represent median posterior spatial autocorrelation functions and 50% and 95% credible intervals.
+Lines and shaded areas represent median posterior spatial autocorrelation functions and equal-tailed 50% and 95% credible intervals.
 
@@
 (ref:plotReplications4Caption) _Posterior estimates of cultural phylogenetic signal from our reanalyses._ Cultural phylogenetic signal is operationalised as the proportion of national-level variance explained by linguistic proximity between nations.
-Ridges are full posterior distributions, and points are posterior medians, and lines represent 50% and 95% credible intervals.
+Ridges are full posterior distributions, points are posterior medians, and lines represent equal-tailed 50% and 95% credible intervals.
+Numbers of observations from the models are as follows, from top to bottom: n = 69, n = 45, n = 103, n = 98, n = 63, n = 89, n = 33, n = 75, n = 67, n = 28, n = 38, and n = 28.
 
@@
 In both panels, the y-axis represents the ratio of the effect size when controlling for spatial and cultural non-independence to the original effect size (from naive regression model), and the x-axis represents posterior median model estimates.
-Regression lines are plotted with 95% credible intervals.
+Regression lines are plotted with equal-tailed 95% credible intervals.
 
@@
 
-(ref:tableArticlesReviewCaption) List of 100 papers included in literature review, sorted by annual rate of citations since publication.
-
-```{r tableArticlesReview, echo=F, warning=F, message=F}
-apa_table(
-  tar_read(tableArticlesReview),
-  booktabs = TRUE,
-  placement = "H",
-  font_size = "footnotesize",
-  longtable = TRUE,
-  align = c("m{2cm}", "m{11cm}", "m{2cm}"),
-  caption = "(ref:tableArticlesReviewCaption)"
-)
-```
-
-\newpage
-
 (ref:tableSim1Caption) False positive rates in simulation with varying levels of spatial autocorrelation for the predictor variable ($\rho$) and for the outcome variable ($\lambda$), assuming that the true correlation is 0.
-Numbers represent the total number of analyses, out of 100, that estimated a slope with a 95% confidence / credible interval excluding zero, falsely inferring a relationship when none is present.
+Numbers represent the total number of analyses, out of 100, that estimated a slope with a two-tailed 95% confidence / credible interval excluding zero, falsely inferring a relationship when none is present.
 
@@
 (ref:tableSim2Caption) False positive rates in simulation with varying levels of cultural phylogenetic autocorrelation for the predictor variable ($\rho$) and for the outcome variable ($\lambda$), assuming that the true correlation is 0.
-Numbers represent the total number of analyses, out of 100, that estimated a slope with a 95% confidence / credible interval excluding zero, falsely inferring a relationship when none is present.
+Numbers represent the total number of analyses, out of 100, that estimated a slope with a two-tailed 95% confidence / credible interval excluding zero, falsely inferring a relationship when none is present.
 
@@
 (ref:tableReplicationsCaption) Estimated cross-national correlations from our reanalysis of twelve previous cross-national analyses.
-Numbers are posterior median slopes (equivalent to Pearson's r correlation coefficients) with 95% credible intervals.
+Numbers are posterior median slopes (equivalent to Pearson's r correlation coefficients) with equal-tailed 95% credible intervals.
 
@@
   placement = "H",
-  font_size = "footnotesize",
+  font_size = "scriptsize",
   longtable = TRUE,
-  align = c("m{4cm}", "m{2cm}", "m{2cm}", "m{7cm}"),
+  align = c("m{2cm}", "m{2cm}", "m{2cm}", "m{9cm}"),
   caption = "(ref:tableDeviationsCaption)"
@@
 
-Brownrigg, R. _maps: Draw geographical maps._ (2018).
-
-Bürkner, P.-C. brms: An R package for Bayesian multilevel models using Stan. _Journal of Statistical Software_ **80**, 1---28 (2017).
-
-Eff, E. A. Weight matrices for cultural proximity: Deriving weights from a language phylogeny. _Structure and Dynamics_ **3**, (2008).
-
-_Ethnologue: Languages of the world._ (SIL International, 2018).
-
-Hammarström, H., Forkel, R., Haspelmath, M.
-& Bank, S.
-_Glottolog 3.0._ (Max Planck Institute for the Science of Human History, 2017).
-doi: [10.5281/zenodo.4061162](https://doi.org/10.5281/zenodo.4061162)
-
-Hijmans, R. J. _geosphere: Spherical trigonometry._ (2019).
+1. Bürkner, P.-C. brms: An R package for Bayesian multilevel models using Stan. _Journal of Statistical Software_ **80**, 1–28 (2017).

### 2023-07-10 manuscript.Rmd
third resubmission

@@
 ---
-title             : "The non-independence of nations and why it matters"
+title             : "Cross-national analyses require additional controls to account for the non-independence of nations"
 shorttitle        : "Non-independence of nations"
@@
     affiliation   : "1"
-    role: # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
-      - Conceptualization
-      - Data curation
-      - Formal analysis
-      - Investigation
-      - Software
-      - Visualization
-      - Writing - original draft
-      - Writing - review & editing
   - name          : "Thanos Kyritsis"
     affiliation   : "1"
-    role:
-      - Data curation
-      - Writing - review & editing
   - name          : "Quentin D. Atkinson"
     affiliation   : "1,2"
-    #corresponding : yes
-    #address       : "Floor 2, Building 302, 23 Symonds Street, Auckland, 1010, New Zealand"
-    #email         : "q.atkinson@auckland.ac.nz"
-    role:
-      - Conceptualization
-      - Data curation
-      - Funding acquisition
-      - Supervision
-      - Writing - review & editing
 
@@
   - id            : "1"
-    institution   : "School of Psychology, University of Auckland, Auckland, New Zealand"
+    institution   : "\\footnotesize School of Psychology, University of Auckland, Auckland, New Zealand"
   - id            : "2"
-    institution   : "Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany"
+    institution   : "\\footnotesize School of Anthropology and Museum Ethnography, University of Oxford, Oxford, United Kingdom"
 
-authornote: "This working paper has not yet been peer-reviewed.
-\\par Correspondence concerning this article should be addressed to Scott Claessens (scott.claessens@gmail.com) and Quentin D.
-Atkinson (q.atkinson@auckland.ac.nz), Floor 2, Building 302, 23 Symonds Street, Auckland, 1010, New Zealand."
+authornote: "Correspondence concerning this article should be addressed to Scott Claessens (scott.claessens@gmail.com) and Quentin D.
+Atkinson (q.atkinson@auckland.ac.nz), Floor 2, Building 302, 23 Symonds Street, Auckland, 1010, New Zealand.
+\\newline \\indent The authors made the following contributions.
+Scott Claessens: Conceptualization, Data curation, Formal analysis, Investigation, Software, Visualization, Writing - original draft, Writing - review & editing; Thanos Kyritsis: Data curation, Writing - review & editing; Quentin D.
+Atkinson: Conceptualization, Data curation, Funding acquisition, Supervision, Writing - review & editing."
 
@@
 keywords          : "spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses"
-wordcount         : "4892 words excl. Methods (Methods: 1877 words)"
+wordcount         : "5102 words excl. Methods (Methods: 2163 words)"
 bibliography      : ["r-references.bib"]
@@
   - \captionsetup[table]{font={stretch=1.5}}
-  - \captionsetup[figure]{font={stretch=1.5}}
+  - \captionsetup[figure]{font={stretch=1,small}}
   - \nolinenumbers
-appendix          : "appendix.Rmd"
 csl               : nature.csl
@@
 
-Nations are perhaps the single most important population unit structuring modern human life.
+Nations are an important population unit structuring modern human life.
 The nation in which someone is born has a large effect on what they can expect out of life, including their income level[@Caselli2005], life expectancy[@Austin2012], mental health[@Rai2013], subjective well-being[@Diener2009], and educational attainment[@Kirkcaldy2004].
@@
 This is known as cultural phylogenetic non-independence[@Tylor1889;@Naroll1961;@Naroll1965], and occurs because related nations share cultural traits inherited via descent from a common ancestor.
-Shared cultural ancestry can result in a form of pseudoreplication, whereby multiple instances of the same trait across nations are merely duplicates of the ancestral original.
-For example, the related island nations Tonga and Tuvalu share similar languages and customs due to cultural inheritance from a common Polynesian population dating back more than 1,000 years.
+Shared cultural ancestry can result in a form of pseudoreplication, whereby multiple instances of the same trait across nations are merely duplicates of the ancestral original[@Tylor1889;@Murdock1969].
+For example, the related island nations Tonga and Tuvalu share similar languages and customs due to cultural inheritance from a common Polynesian population dating back more than 1,000 years[@Gray2009;@Green1999].
 More recently, shared ancestry explains similarities in institutions, norms, technologies, and values between colonial settlements and their colonisers (e.g.
@@
 But without accounting for spatial or cultural non-independence between nations, model residuals can show structure that remains unaccounted for, violating this assumption.
-Treating nations as independent can thus inflate false positive rates[@Legendre1993], producing spurious "direct" relationships between variables that in fact only indirectly covary due to spatial or cultural connections[@Roberts2013] (see Supplementary Figure \@ref(fig:dag) for an example causal model).
+Treating nations as independent can thus inflate false positive rates[@Legendre1993], producing spurious relationships between variables that in fact only indirectly covary due to spatial or cultural connections[@Roberts2013] (see Supplementary Figure \@ref(fig:dag) for an example causal model).
 
@@
 In anthropology, researchers have recognised cultural non-independence as an important statistical pitfall for over 200 years, with issues of cultural pseudoreplication being identified in early comparative studies of marriage practices across societies[@Tylor1889].
-In the twentieth century, anthropologists began to emphasise that human societies do not develop independently, but rather exist in a globally interconnected "world system" linked by shared history and cultural ancestry [@Ferguson1997].
+In the twentieth century, anthropologists began to emphasise that human societies do not develop independently, but rather exist in a globally interconnected world system linked by shared history and cultural ancestry [@Ferguson1997].
 In order to minimise the confounding effects of this non-independence in comparisons of human societies, researchers compiled the Standard Cross-Cultural Sample of 186 cultures which deliberately avoids sampling closely related cultures[@Murdock1969], though it is difficult to completely remove spatial and cultural dependencies[@Dow2008;@Eff2004] and the smaller sample of cultures reduces statistical power.
@@
 At the national level, recent reanalyses have revealed that several cross-national relationships reported in economics and psychology do not hold when controlling for non-independence between nations.
-One working paper replicated 25 analyses of "persistence" in economics, in which modern national outcomes are regressed against historical characteristics of those nations, and found that over half of the relationships were attenuated when controlling for spatial non-independence[@Kelly2020].
+One working paper replicated 25 analyses of persistence in economics, in which modern national outcomes are regressed against historical characteristics of those nations, and found that over half of the relationships were attenuated when controlling for spatial non-independence[@Kelly2020].
 Another replication study found that many of the widely publicised relationships between national-level pathogen prevalence and political institutions and attitudes fail to hold when controlling for various kinds of non-independence[@Bromham2018].
@@
 First, we demonstrate that economic development and cultural values are spatially and culturally non-independent across nations, emphasising the need to control for non-independence.
-Second, we systematically review the 100 highest-cited cross-national studies of economic development and cultural values and determine baseline rates of controlling for non-independence in the literature.
+Second, we review the 100 highest-cited cross-national studies of economic development and cultural values and determine baseline rates of controlling for non-independence in the literature.
 Third, we run simulations to determine whether common methods of dealing with non-independence in the literature sufficiently reduce false positive rates.
-Fourth, we reanalyse twelve previous cross-national analyses of economic development and cultural values from our systematic review, incorporating global geographic and linguistic proximity matrices as controls for spatial and cultural non-independence.
+Fourth, we reanalyse twelve previous cross-national analyses of economic development and cultural values from our literature review, incorporating global geographic and linguistic proximity matrices as controls for spatial and cultural non-independence.
 
@@
 For all of these variables, we found that a substantial proportion of national-level variation was explained by spatial proximity and/or shared cultural ancestry between nations (Figure \@ref(fig:plotSignal); see Supplementary Table \@ref(tab:tableSignal) for numerical results).
-Signal estimates were often markedly strong, with spatial proximity and shared cultural ancestry frequently explaining over half of the national-level variation.
+Signal estimates were often strong, with spatial proximity and shared cultural ancestry frequently explaining over half of the national-level variation.
 For spatial proximity, Bayes Factors indicated strong evidence that the geographic signal estimates differed from zero for all economic development variables and traditional values.
@@
 
-(ref:plotSignalCaption) _Posterior estimates of geographic and cultural phylogenetic signal for a range of economic development and cultural values variables._ Geographic and cultural phylogenetic signal are operationalised as the proportion of national-level variance explained by geographic and linguistic proximity matrices.
-Grey ridges are full posterior distributions, points are posterior median values, and black lines are 50% and 95% credible intervals.
+(ref:plotSignalCaption) Posterior estimates of geographic and cultural phylogenetic signal for a range of economic development and cultural values variables.
+Geographic and cultural phylogenetic signal are operationalised as the proportion of national-level variance explained by geographic and linguistic proximity matrices.
+Grey ridges are full posterior distributions, points are posterior median values, and black lines are equal-tailed 50% and 95% credible intervals.
+Number of observations from these models are as follows: HDI (n = 1449), GDP per capita (n = 10289), GDP per capita growth (n = 9755), Gini index (n = 1826), traditional values (n = 277676), survival values (n = 277676), tightness (n = 57), and individualism (n = 67).
 HDI = Human Development Index; GDP = gross domestic product.
@@
 Given that economic development and cultural values show evidence of geographic and cultural phylogenetic signal, have cross-national analyses sufficiently accounted for this non-independence?
-To assess this, we systematically searched the published literature for articles that combined the search terms "economic development" or "values" with the search terms "cross-national", "cross-cultural", or "cross-country".
+To assess this, we searched the published literature for articles that combined the search terms "economic development" or "values" with the search terms "cross-national", "cross-cultural", or "cross-country".
 We removed articles that did not report original research, were not relevant to economic development or cultural values, or did not report at least one cross-national analysis.
-We then retained the 100 articles (50 for economic development, 50 for cultural values) with the highest annual rate of citations (Supplementary Table \@ref(tab:tableArticlesReview)).
+We then retained the 100 articles (50 for economic development, 50 for cultural values) with the highest annual rate of citations (see Supplementary Data).
 For each of these highly-cited articles, we exhaustively recorded every cross-national analysis reported in the main text (n = `r printnum(as.numeric(nrow(review)), digits = 0)`), identifying in each case whether or not the analysis attempted to control for spatial, cultural, or any other form of non-independence between nations (see Methods for detailed search criteria and coding decisions).
@@
 
-The results of our systematic review show that most published articles containing cross-national analyses make no attempt to account for statistical non-independence.
+The results of our literature review show that most published articles containing cross-national analyses make no attempt to account for statistical non-independence.
 Figure \@ref(fig:plotReview1)a plots the proportion of articles that contain at least one cross-national analysis accounting for non-independence.
-We find that `r printnum(prop$prop[1]*100, digits = 0)`% of economic development articles contain at least one attempt to control for non-independence (95% bootstrap confidence interval [`r printnum(prop$lower[1])` `r printnum(prop$upper[1])`]), while this proportion decreases to only `r printnum(prop$prop[2]*100, digits = 0)`% for cultural values articles (95% bCI [`r printnum(prop$lower[2])` `r printnum(prop$upper[2])`]).
+We find that `r printnum(prop$prop[1]*100, digits = 0)`% of economic development articles contain at least one attempt to control for non-independence (two-tailed 95% bootstrap confidence interval [`r printnum(prop$lower[1])` `r printnum(prop$upper[1])`]), while this proportion decreases to only `r printnum(prop$prop[2]*100, digits = 0)`% for cultural values articles (two-tailed 95% bCI [`r printnum(prop$lower[2])` `r printnum(prop$upper[2])`]).
 Both kinds of article are most likely to use regional fixed effects (e.g.
@@
 colony status).
-These proportions are even lower when focusing on the full sample of `r printnum(as.numeric(nrow(review)), digits = 0)` analyses: only `r printnum(median(propEcon)*100, digits = 0)`% (95% credible interval [`r printnum(quantile(propEcon, 0.025))` `r printnum(quantile(propEcon, 0.975))`]) of individual economic development analyses and `r printnum(median(propVal)*100, digits = 0)`% (95% CI [`r printnum(quantile(propVal, 0.025))` `r printnum(quantile(propVal, 0.975))`]) of individual cultural values analyses are estimated to control for non-independence (Supplementary Figure \@ref(fig:plotReview2)).
+These proportions are even lower when focusing on the full sample of `r printnum(as.numeric(nrow(review)), digits = 0)` analyses: only `r printnum(median(propEcon)*100, digits = 0)`% (equal-tailed 95% credible interval [`r printnum(quantile(propEcon, 0.025))` `r printnum(quantile(propEcon, 0.975))`]) of individual economic development analyses and `r printnum(median(propVal)*100, digits = 0)`% (equal-tailed 95% CI [`r printnum(quantile(propVal, 0.025))` `r printnum(quantile(propVal, 0.975))`]) of individual cultural values analyses are estimated to control for non-independence (Supplementary Figure \@ref(fig:plotReview2)).
 
-(ref:plotReview1Caption) _Results from systematic review of 100 highly-cited cross-national studies of economic development (red) and cultural values (blue)._ (a) Proportion of articles containing at least one analysis accounting for non-independence, overall and split by common methods of controlling for non-independence.
+(ref:plotReview1Caption) Results from literature review of 100 highly-cited cross-national studies of economic development (red) and cultural values (blue).
+(a) Proportion of articles containing at least one analysis accounting for non-independence, overall and split by common methods of controlling for non-independence.
+Points represent raw proportions of articles and ranges represent two-tailed 95% bootstrap confidence intervals (n = 1000 bootstrap samples).
 (b) The association between journal impact factor and the probability that an article contains at least one analysis accounting for non-independence.
 (c) Estimated trend over time for the probability that an article contains at least one analysis accounting for non-independence.
-Point ranges represent proportions and 95% bootstrap confidence intervals.
-Lines and shaded areas are posterior median regression lines and 50% credible intervals from Bayesian logistic regression models.
+Lines and shaded areas are posterior median regression lines and equal-tailed 50% credible intervals from Bayesian logistic regression models (n = 100 observations).
 Dots represent raw counts of individual articles that did (top) or did not (bottom) account for non-independence.
@@
 
-While our review contains articles from journals with a range of impact factors, an anonymous reviewer suggested that our estimates could be biased downwards by analyses published in lower impact outlets with more relaxed standards for issues like non-independence.
-It is also possible that, since our systematic review goes back as far as 1993, our estimates are being biased downwards by earlier studies, and that controls for non-independence have increased over time with methodological advancements and greater awareness of the issue.
+While our review contains articles from journals with a range of impact factors, our estimates could be biased downwards by analyses published in lower impact outlets with more relaxed standards for issues like non-independence.
+It is also possible that, since our literature review goes back as far as 1993, our estimates are being biased downwards by earlier studies, and that controls for non-independence have increased over time with methodological advancements and greater awareness of the issue.
 To test these possible explanations for our low estimates, we fitted logistic regression models to the data from the review, including log journal impact factor and publication year as separate predictors.
-Interestingly, we found that, if anything, studies from higher impact journals were _less_ likely to include at least one control for non-independence than studies from lower impact journals, both for studies of economic development (b = `r printnum(fixef(ifArticle)["ReviewEconomicdevelopment:logIF","Estimate"])`, 95% CI [`r printnum(fixef(ifArticle)["ReviewEconomicdevelopment:logIF","Q2.5"])` `r printnum(fixef(ifArticle)["ReviewEconomicdevelopment:logIF","Q97.5"])`]) and for studies of cultural values (b = `r printnum(fixef(ifArticle)["ReviewValues:logIF","Estimate"])`, 95% CI [`r printnum(fixef(ifArticle)["ReviewValues:logIF","Q2.5"])` `r printnum(fixef(ifArticle)["ReviewValues:logIF","Q97.5"])`]; Figure \@ref(fig:plotReview1)b).
-Moreover, splines revealed no relationship between publication year and the probability of including at least one control for non-independence, both for studies of economic development (b = `r printnum(fixef(yearArticle)["sYear.std:ReviewEconomicdevelopment_1","Estimate"])`, 95% CI [`r printnum(fixef(yearArticle)["sYear.std:ReviewEconomicdevelopment_1","Q2.5"])` `r printnum(fixef(yearArticle)["sYear.std:ReviewEconomicdevelopment_1","Q97.5"])`]) and for studies of cultural values (b = `r printnum(fixef(yearArticle)["sYear.std:ReviewValues_1","Estimate"])`, 95% CI [`r printnum(fixef(yearArticle)["sYear.std:ReviewValues_1","Q2.5"])` `r printnum(fixef(yearArticle)["sYear.std:ReviewValues_1","Q97.5"])`]; Figure \@ref(fig:plotReview1)c).
+Interestingly, we found that, if anything, studies from higher impact journals were less likely to include at least one control for non-independence than studies from lower impact journals, both for studies of economic development (b = `r printnum(fixef(ifArticle)["ReviewEconomicdevelopment:logIF","Estimate"])`, equal-tailed 95% CI [`r printnum(fixef(ifArticle)["ReviewEconomicdevelopment:logIF","Q2.5"])` `r printnum(fixef(ifArticle)["ReviewEconomicdevelopment:logIF","Q97.5"])`]) and for studies of cultural values (b = `r printnum(fixef(ifArticle)["ReviewValues:logIF","Estimate"])`, equal-tailed 95% CI [`r printnum(fixef(ifArticle)["ReviewValues:logIF","Q2.5"])` `r printnum(fixef(ifArticle)["ReviewValues:logIF","Q97.5"])`]; Figure \@ref(fig:plotReview1)b).
+Moreover, splines revealed no relationship between publication year and the probability of including at least one control for non-independence, both for studies of economic development (b = `r printnum(fixef(yearArticle)["sYear.std:ReviewEconomicdevelopment_1","Estimate"])`, equal-tailed 95% CI [`r printnum(fixef(yearArticle)["sYear.std:ReviewEconomicdevelopment_1","Q2.5"])` `r printnum(fixef(yearArticle)["sYear.std:ReviewEconomicdevelopment_1","Q97.5"])`]) and for studies of cultural values (b = `r printnum(fixef(yearArticle)["sYear.std:ReviewValues_1","Estimate"])`, equal-tailed 95% CI [`r printnum(fixef(yearArticle)["sYear.std:ReviewValues_1","Q2.5"])` `r printnum(fixef(yearArticle)["sYear.std:ReviewValues_1","Q97.5"])`]; Figure \@ref(fig:plotReview1)c).
 
@@
 
-Our systematic review revealed that most cross-national analyses in the literature do not control for spatial or cultural phylogenetic non-independence.
+Our literature review revealed that most cross-national analyses in the literature do not control for spatial or cultural phylogenetic non-independence.
 When they do, they tend to include controls like latitude and regional fixed effects.
@@
 We then fitted naive regressions without controls to these datasets, as well as regression models with controls for latitude, longitude, and continent fixed effects.
-Despite not being identified in our systematic review, we also included other methods that are often used in the literature to account for non-independence.
+Despite not being identified in our literature review, we also included other methods that are often used in the literature to account for non-independence.
 Additional spatial controls included the mean of the predictor variable within a surrounding 2000km radius (e.g.
@@
 For each pair of nations, we calculated inverse phylogenetic distances (i.e.
-number of branches separating two taxa) between all languages spoken in that nation pair and produced an average "linguistic proximity" score weighted by the percentages of speakers within those nations.
+number of branches separating two taxa) between all languages spoken in that nation pair and produced an average linguistic proximity score weighted by the percentages of speakers within those nations.
 To include the resulting geographic and linguistic proximity matrices in our models, we included a Gaussian process[@McElreath2020;@Neal1998] over latitude and longitude values and/or assumed that nation random intercepts were correlated in proportion to their linguistic proximity[@deVillemereuil2014].
@@
 Figures \@ref(fig:plotSim3) and \@ref(fig:plotSim4) plot the estimated false positive rates from our simulation study, split by different methods and different degrees of spatial or cultural phylogenetic autocorrelation (see Supplementary Tables \@ref(tab:tableSim1) and \@ref(tab:tableSim2) for numerical results and Supplementary Figures \@ref(fig:plotSim1) and \@ref(fig:plotSim2) for full distributions of effect sizes under strong autocorrelation).
-Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero, falsely inferring a relationship when none is present.
-For reference, "weak" autocorrelation in our simulation is comparable to the geographic signal for survival values in Figure \@ref(fig:plotSignal) (i.e.
-20% of the national-level variance is explained by non-independence), while "moderate" and "strong" levels of autocorrelation are comparable to the cultural phylogenetic signal for traditional and survival values, respectively (i.e.
+Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a two-tailed 95% confidence / credible interval excluding zero, falsely inferring a relationship when none is present.
+For reference, weak autocorrelation in our simulation is comparable to the geographic signal for survival values in Figure \@ref(fig:plotSignal) (i.e.
+20% of the national-level variance is explained by non-independence), while moderate and strong levels of autocorrelation are comparable to the cultural phylogenetic signal for traditional and survival values, respectively (i.e.
 50% and 80% of the national-level variance is explained by non-independence).
@@
 
-(ref:plotSim3Caption) _False positive rates for different methods of controlling for spatial non-independence in our simulation study._ For simulated outcome and predictor variables, we systematically varied the strength of spatial autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+(ref:plotSim3Caption) False positive rates for different methods of controlling for spatial non-independence in our simulation study.
+For simulated outcome and predictor variables, we systematically varied the strength of spatial autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
 We simulated 100 datasets per parameter combination and fitted different models to each dataset.
-False positive rates were operationalised as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero.
-Point ranges represent proportions and 95% bootstrap confidence intervals, and dashed lines indicate the 5% false positive rate that is expected due to chance.
+False positive rates were operationalised as the proportion of models that estimated a slope with a two-tailed 95% confidence / credible interval excluding zero.
+Points represent raw proportions of false positive models, ranges represent two-tailed 95% bootstrap confidence intervals (n = 1000 bootstrap samples), and dashed lines indicate the 5% false positive rate that is expected due to chance.
+Colours indicate whether the strength of autocorrelation for the predictor variable is 0.2 (red), 0.5, (green) or 0.8 (blue).
 SEs = standard errors.
 
-```{r plotSim3, fig.width=11, fig.height=7, fig.cap="(ref:plotSim3Caption)"}
+```{r plotSim3, fig.width=7, fig.height=5.5, fig.cap="(ref:plotSim3Caption)"}
 tar_read(plotSim3)
@@
 
-(ref:plotSim4Caption) _False positive rates for different methods of controlling for cultural phylogenetic non-independence in our simulation study._ For simulated outcome and predictor variables, we systematically varied the strength of cultural phylogenetic autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+(ref:plotSim4Caption) False positive rates for different methods of controlling for cultural phylogenetic non-independence in our simulation study.
+For simulated outcome and predictor variables, we systematically varied the strength of cultural phylogenetic autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
 We simulated 100 datasets per parameter combination and fitted different models to each dataset.
-False positive rates were operationalised as the proportion of models that estimated a slope with a 95% confidence / credible interval excluding zero.
-Point ranges represent proportions and 95% bootstrap confidence intervals, and dashed lines indicate the 5% false positive rate that is expected due to chance.
+False positive rates were operationalised as the proportion of models that estimated a slope with a two-tailed 95% confidence / credible interval excluding zero.
+Points represent raw proportions of false positive models, ranges represent two-tailed 95% bootstrap confidence intervals (n = 1000 bootstrap samples), and dashed lines indicate the 5% false positive rate that is expected due to chance.
+Colours indicate whether the strength of autocorrelation for the predictor variable is 0.2 (red), 0.5, (green) or 0.8 (blue).
 SEs = standard errors.
 
-```{r plotSim4, fig.width=11, fig.height=7, fig.cap="(ref:plotSim4Caption)"}
+```{r plotSim4, fig.width=7, fig.height=5.5, fig.cap="(ref:plotSim4Caption)"}
 tar_read(plotSim4)
@@
 However, additional simulations revealed that these reductions in false positive rates come at the cost of lower statistical power (see Supplementary Figures \@ref(fig:plotPowerGeo01) - \@ref(fig:plotPowerGeo05)).
-In additional simulations where the true relationship between the predictor and outcome variable was known, power analyses showed that both continent fixed effects and the 2000km radius control had less than 80% power to detect moderate true correlations (_r_ = 0.3) under strong spatial autocorrelation.
+In additional simulations where the true relationship between the predictor and outcome variable was known, power analyses showed that both continent fixed effects and the 2000km radius control had less than 80% power to detect moderate true correlations (r = 0.3) under strong spatial autocorrelation.
 
@@
 False positives are not completely eliminated with these random effects models.
-Nevertheless, these methods have at least 80% power to detect moderate (_r_ = 0.3) and large (_r_ = 0.5) true correlations between variables under all levels of spatial autocorrelation (see Supplementary Figures \@ref(fig:plotPowerGeo01) - \@ref(fig:plotPowerGeo05)).
+Nevertheless, these methods have at least 80% power to detect moderate (r = 0.3) and large (r = 0.5) true correlations between variables under all levels of spatial autocorrelation (see Supplementary Figures \@ref(fig:plotPowerGeo01) - \@ref(fig:plotPowerGeo05)).
 
@@
 By contrast, models with random effects covarying according to linguistic proximity completely eliminate false positives across all degrees of cultural phylogenetic autocorrelation.
-This approach is also the only method that is able to detect large true correlations (_r_ = 0.5) with at least 80% power (see Supplementary Figures \@ref(fig:plotPowerLin01) - \@ref(fig:plotPowerLin05)).
+This approach is also the only method that is able to detect large true correlations (r = 0.5) with at least 80% power (see Supplementary Figures \@ref(fig:plotPowerLin01) - \@ref(fig:plotPowerLin05)).
 Random effects models that additionally account for geographic proximity between nations perform equally well, though models with only a spatial Gaussian process continue to produce false positives.
@@
 
-Our systematic review and simulation study have shown that controls for non-independence are rare in cross-national studies of economic development and cultural values, and when studies do attempt to control for non-independence, the methods typically used are unable to sufficiently reduce false positive rates.
+Our literature review and simulation study have shown that controls for non-independence are rare in cross-national studies of economic development and cultural values, and when studies do attempt to control for non-independence, the methods typically used are unable to sufficiently reduce false positive rates.
 This raises the worrying possibility that the cross-national literature in economics and psychology is populated with spurious relationships.
 
-To determine how widespread this issue of spurious cross-national relationships might be, we reanalysed a subset of twelve previous cross-national analyses from our systematic review, controlling for spatial and cultural phylogenetic non-independence using global geographic and linguistic proximity matrices.
-Out of the 100 papers included in our systematic review, primary or secondary data were publicly available for `r printnum(review %>% group_by(Review, Citation) %>% summarise(DataAvailable = ifelse(mean(DataAvailable == "Yes") > 0, TRUE, FALSE), .groups = "drop") %>% pull(DataAvailable) %>% sum())` papers.
+To determine how widespread this issue of spurious cross-national relationships might be, we reanalysed a subset of twelve previous cross-national analyses from our literature review, controlling for spatial and cultural phylogenetic non-independence using global geographic and linguistic proximity matrices.
+Out of the 100 papers included in our literature review, primary or secondary data were publicly available for `r printnum(review %>% group_by(Review, Citation) %>% summarise(DataAvailable = ifelse(mean(DataAvailable == "Yes") > 0, TRUE, FALSE), .groups = "drop") %>% pull(DataAvailable) %>% sum())` papers.
 We attempted to replicate key statistically significant cross-national correlations from these papers --- mostly initial bivariate regression specifications without covariates (see Methods) --- and stopped when we had sampled a set of twelve analyses for which we were able to replicate the original result.
@@
 
-(ref:plotReplications1Caption) _Posterior correlations from our reanalysis of twelve previous cross-national analyses._ For each previous cross-national relationship, we plot the posterior slopes from a naive regression (red), a regression controlling for spatial non-independence (green), a regression controlling for cultural phylogenetic non-independence (blue), and a regression controlling for both spatial and cultural phylogenetic non-independence simultaneously (purple).
+(ref:plotReplications1Caption) Posterior correlations from our reanalysis of twelve previous cross-national analyses.
+For each previous cross-national relationship, we plot the posterior slopes from a naive regression (red), a regression controlling for spatial non-independence (green), a regression controlling for cultural phylogenetic non-independence (blue), and a regression controlling for both spatial and cultural phylogenetic non-independence simultaneously (purple).
 All outcome and predictor variables are standardised.
@@
 (2011) is a partial correlation controlling for log gross national income and Adamczyk and Pitt (2009) is a multilevel model including several covariates.
-Coloured ridges are full posterior distributions, and points and black lines represent posterior medians and 95% credible intervals.
+Coloured ridges are full posterior distributions, and points and black lines represent posterior medians and equal-tailed 95% credible intervals.
+Numbers of observations from the models are as follows, from top to bottom: n = 69, n = 45, n = 103, n = 98, n = 63, n = 89, n = 33, n = 75, n = 67, n = 28, n = 38, and n = 28.
 GDP = gross domestic product.
@@
 
+```{r load autocor}
+tar_load(c(inglehartSA1, skidmoreSA1, alesinaSignal1, knackSignal1, effSizeRatioModel1, effSizeRatioModel2))
+```
+
+To understand why some cross-national correlations were attenuated by controls for non-independence while others were robust, we further explored our fitted models for evidence of spatial and cultural autocorrelation.
+For each outcome variable, our Gaussian process models provided varying estimates of how quickly spatial autocorrelation declined with distance (Supplementary Figure \@ref(fig:plotReplications3)).
+For example, in Skidmore and Toya[@Skidmore2002] gross domestic product growth was only moderately spatially autocorrelated at 1,000 km distance (posterior median spatial autocorrelation at 1,000 km = `r printnum(median(skidmoreSA1))`, equal-tailed 95% CI [`r printnum(quantile(skidmoreSA1, 0.025))` `r printnum(quantile(skidmoreSA1, 0.975))`]), whereas in Inglehart and Baker[@Inglehart2000] traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1,000 km = `r printnum(median(inglehartSA1))`, 95% CI [`r printnum(quantile(inglehartSA1, 0.025))` `r printnum(quantile(inglehartSA1, 0.975))`]).
+We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \@ref(fig:plotReplications4)), with some outcome variables expressing low signal (e.g.
+confidence in institutions[@Knack1997]; posterior median = `r printnum(median(knackSignal1))`, 95% CI [`r printnum(quantile(knackSignal1, 0.025))` `r printnum(quantile(knackSignal1, 0.975))`]) and others expressing high signal (e.g.
+female labour force participation[@Alesina2013]; posterior median = `r printnum(median(alesinaSignal1))`, 95% CI [`r printnum(quantile(alesinaSignal1, 0.025))` `r printnum(quantile(alesinaSignal1, 0.975))`]).
+Exploratory regressions provided suggestive evidence that stronger estimates of spatial autocorrelation or cultural phylogenetic signal resulted in a more pronounced reduction in the effect size when controlling for non-independence between nations (Supplementary Figure \@ref(fig:plotReplications5)).
+However, these negative slopes for spatial autocorrelation (b = `r printnum(fixef(effSizeRatioModel1)["corAt1000km","Estimate"])`, 95% CI [`r printnum(fixef(effSizeRatioModel1)["corAt1000km","Q2.5"])` `r printnum(fixef(effSizeRatioModel1)["corAt1000km","Q97.5"])`]) and cultural phylogenetic signal (b = `r printnum(fixef(effSizeRatioModel2)["signal","Estimate"])`, 95% CI [`r printnum(fixef(effSizeRatioModel2)["signal","Q2.5"])` `r printnum(fixef(effSizeRatioModel2)["signal","Q97.5"])`]) were both very uncertain, due to the small number of analyses and the posterior uncertainty in effect sizes and estimates of non-independence.
+
 # Discussion
 
-In a systematic literature review and simulation, we found that cross-national studies in economics and psychology rarely account for non-independence between nations, and, when they do, the methods they use are insufficient to reduce false positives in non-independent data.
+In a literature review and simulation, we found that cross-national studies in economics and psychology rarely account for non-independence between nations, and, when they do, the methods they use are insufficient to reduce false positives in non-independent data.
 In a reanalysis of twelve cross-national correlations, we further showed that neglecting to account for non-independence has resulted in potentially spurious relationships in the published literature, with half of the correlations failing to replicate when controlling for spatial or cultural non-independence with more rigorous methods.
@@
 
-Crucially, our systematic literature review and simulation study revealed that the most commonly used controls for non-independence do not sufficiently deal with the issue.
+Crucially, our literature review and simulation study revealed that the most commonly used controls for non-independence do not sufficiently deal with the issue.
 In our simulations, controlling for either latitude or longitude did not reduce false positive rates.
@@
 Controlling for the mean of the predictor variable within a 2000km radius[@Schulz2022] eliminated false positive rates in spatially autocorrelated data, but had reduced statistical power to detect true associations.
-Across all model types in our simulation, the only methods that markedly reduced false positive rates while retaining high statistical power were the random effects models with covariance matrices.
-The important advantage of these models is not that they are Bayesian _per se_ (any of the approaches used here could be implemented in a Bayesian framework), but rather that they explicitly model covariance as a function of spatial or cultural distance.
+Across all model types in our simulation, the only methods that reduced false positive rates while retaining high statistical power were the random effects models with covariance matrices.
+The important advantage of these models is not that they are Bayesian per se (any of the approaches used here could be implemented in a Bayesian framework), but rather that they explicitly model covariance as a function of spatial or cultural distance.
 
@@
 Why do cross-national studies also rarely account for non-independence?
-At the institutional level, one possibility is that such practices are incentivised _because_ they generate statistically significant relationships, which increase the probability that a study is published[@Smaldino2016].
-Indeed, we found that controls for non-independence were _less_ common among articles published in high impact journals, suggesting that researchers are rewarded for such practices.
+At the institutional level, one possibility is that such practices are incentivised because they generate statistically significant relationships, which increase the probability that a study is published[@Smaldino2016].
+Indeed, we found that controls for non-independence were less common among articles published in high impact journals, suggesting that researchers are rewarded for such practices.
 At the individual level, another possibility is that researchers outside of anthropology and ecology are less aware of the problem, or believe that the problem does not apply to analyses of nations.
@@
 Indeed, in our review, economists studying economic development dealt with national-level non-independence more than psychologists studying cultural values, likely because economics studies tend to be lengthy statistical exercises that systematically incorporate or exclude numerous variables in an attempt to infer causation.
-Third, the recent widespread accessibility of open source statistical software, such as the programming language Stan[@rstan] and the R package _brms_[@Burkner2017], should promote the use of more rigorous methods to control for non-independence.
-Using _brms_, for example, Bayesian Gaussian process regression is straightforward to conduct, requiring only longitude and latitude values for nations.
+Third, the recent widespread accessibility of open source statistical software, such as the programming language Stan[@rstan] and the R package brms[@Burkner2017], should promote the use of more rigorous methods to control for non-independence.
+Using brms, for example, Bayesian Gaussian process regression is straightforward to conduct, requiring only longitude and latitude values for nations.
 We have provided an online tutorial to help researchers apply these methods to their own cross-cultural datasets (https://scottclaessens.github.io/blog/2022/crossnational/).
@@
 Until such changes are implemented and sufficient controls for non-independence are the norm, existing cross-national correlations should be interpreted with caution.
-In our reanalyses, we found that half of the cross-national correlations had 95% credible intervals that included zero when controlling for spatial and/or cultural phylogenetic non-independence.
+In our reanalyses, we found that half of the cross-national correlations had equal-tailed 95% credible intervals that included zero when controlling for spatial and/or cultural phylogenetic non-independence.
 While these results are in line with previous reanalyses[@Kelly2020], we note that we are unable to outright reject the claims from these studies, since we only reanalysed the first bivariate regression specifications presented in the papers[@Voth2021].
@@
 
-## Geographic and cultural phylogenetic signal
+## Geographic and cultural phylogenetic signal for measures of economic development and cultural values
 
@@
 To calculate geographic and cultural phylogenetic signal, we created two proximity matrices for 269 of the world's nations: a geographic proximity matrix and a linguistic proximity matrix.
-Geographic proximity was converted from logged geodesic distances between nation capital cities.
-Linguistic proximity was calculated as the cultural proximity between all languages spoken within nations, weighted by speaker percentages (see Supplementary Methods).
+Geographic distance between two nations was calculated as the logged geodesic distance between nation capital cities (data from the maps R package[@Brownrigg2018]) using the geosphere R package[@Hijmans2019].
+The geographic proximity matrix was computed as one minus the log geographic distance matrix scaled between 0 and 1.
+Linguistic proximity between two nations was calculated as the cultural proximity between all languages spoken within those nations, weighted by speaker percentages.
+We acquired cultural proximity data by combining the language family trees provided by Glottolog v3.0[@Glottolog] into one global language tree (undated and unresolved).
+We calculated cultural proximity $s$ between two languages $j$ and $k$ as the distance (in number of nodes traversed) of their most recent common ancestor $i$ to the root of the tree, through the formula:
+
+\begin{gather}
+s_{jk} = \frac{n_{r}-n_{i}}{n_{r}}
+\end{gather}
+
+where $n_{r}$ is the maximum path length (in number of nodes traversed) leading to the pan-human root $r$, and $n_{i}$ is the maximum path length leading to node $i$.
+We then combined these proximities with speaker data from Ethnologue 21[@Ethnologue] and compared every language spoken within those nations by at least 1 permille of the population, weighted by speaker percentages, through the formula:
+
+\begin{gather}
+w_{lm} = {\Sigma}{\Sigma}p_{lj}p_{mk}s_{jk}
+\end{gather}
+
+where $p_{lj}$ is the percentage of the population in nation $l$ speaking language $j$, $p_{mk}$ is the percentage of the population in nation $m$ speaking language $k$, and $s_{jk}$ is the proximity measure between languages $j$ and $k$[@Eff2008].
+This calculation resulted in a linguistic proximity matrix with values between 0 and 1.
+
 We included these matrices in Bayesian multilevel models, allowing nation random intercepts to covary according to both geographic and linguistic proximity simultaneously.
-These models were fitted with the R package _brms_[@Burkner2017] and converged normally ($\hat{R}$ < 1.1).
+These models were fitted with the R package brms[@Burkner2017] and converged normally ($\hat{R}$ < 1.1).
+The assumptions of these models were met: residuals were approximately normally distributed, though this was not formally tested.
 Estimates of geographic and cultural phylogenetic signal were computed as the proportion of national-level variance in these models explained by geographic and linguistic proximity matrices.
 
-## Systematic literature review
+## Literature review
 
@@
 
-We then systematically coded each article, in order, for inclusion in our review.
+We then coded each article, in order, for inclusion in our review.
 Articles were only included if: (1) they were judged to be relevant to economic development or cultural values; (2) they were an original empirical research article; and (3) they contained at least one analysis with national-level outcome or predictor variables.
@@
 We also calculated the proportion of articles employing the different types of control listed above at least once: regional fixed effects, distance, shared cultural history, or other.
-For these proportions, we calculated 95% bootstrap confidence intervals with 1,000 bootstrap iterations.
+For these proportions, we calculated two-tailed 95% bootstrap confidence intervals with 1,000 bootstrap iterations.
 Additionally, we predicted the probability of an article attempting to control for non-independence at least once using Bayesian logistic regression, including in separate models log journal impact factor and year of publication as linear and spline predictors, respectively.
@@
 We fitted these models separately for overall attempts to control for non-independence and split by method type.
-We report the adjusted proportions with 95% credible intervals (see Results and Supplementary Figure \@ref(fig:plotReview2)).
+We report the adjusted proportions with equal-tailed 95% credible intervals (see Results and Supplementary Figure \@ref(fig:plotReview2)).
 Additionally, we predicted the probability of an analysis attempting to control for non-independence using Bayesian multilevel logistic regression with random intercepts for articles.
 In separate models, we included log journal impact factor and year of publication as linear and spline predictors, respectively.
-All Bayesian models were fitted with the _brms_ R package[@Burkner2017].
+All Bayesian models were fitted with the brms R package[@Burkner2017].
 Our priors were informed by prior predictive checks, and all models converged normally ($\hat{R}$ < 1.1).
+The assumptions of these models (i.e., binary nested data) were met.
 
@@
 
-\begin{gather*}
+\begin{gather}
 \begin{bmatrix}y_i\\x_i \end{bmatrix} \sim \text{MVNormal}
 \begin{pmatrix}\begin{bmatrix}\alpha_y\\\alpha_x \end{bmatrix},\textbf{S}\end{pmatrix}\\
-\alpha_y \sim \text{Normal}(0, \sqrt{\lambda} \cdot \Sigma) \\
-\alpha_x \sim \text{Normal}(0, \sqrt{\rho} \cdot \Sigma) \\
+\alpha_y \sim \text{Normal}(0, \sqrt{\lambda} \cdot \Sigma) \nonumber \\
+\alpha_x \sim \text{Normal}(0, \sqrt{\rho} \cdot \Sigma) \nonumber \\
 \textbf{S} = 
@@
 \begin{pmatrix}1 & r \\ r & 1\end{pmatrix}
-\begin{pmatrix}\sqrt{1 - \lambda} & 0 \\ 0 & \sqrt{1 - \rho} \end{pmatrix}
-\\
-\end{gather*}
+\begin{pmatrix}\sqrt{1 - \lambda} & 0 \\ 0 & \sqrt{1 - \rho} \end{pmatrix} \nonumber
+\end{gather}
 
@@
 To determine distance cutoffs, we employed an approach recommended in previous work[@Colella2019]: we fitted models with a range of feasible distance cutoffs and retained the model with the largest standard error for the slope parameter.
-These models were fitted using the _conleyreg_ R package[@Duben2021].
-Bayesian models were fitted using the _brms_ R package[@Burkner2017].
+These models were fitted using the conleyreg R package[@Duben2021].
+Bayesian models were fitted using the brms R package[@Burkner2017].
 Our choice of priors was based on prior predictive simulation.
 All models converged normally ($\hat{R}$ < 1.1).
-Across all model types and parameter combinations, we calculated the false positive rate as the proportion of models that estimated slopes with a 95% confidence / credible interval excluding zero when $r = 0$.
+Across all model types and parameter combinations, we calculated the false positive rate as the proportion of models that estimated slopes with a two-tailed 95% confidence / credible interval excluding zero when $r = 0$.
 We calculated statistical power as the proportion of models that estimated slopes with a 95% confidence / credible interval excluding zero when $r > 0$.
-We calculated 95% bootstrap confidence intervals for these false positive rates and statistical power estimates with 1,000 bootstrap iterations.
+We calculated two-tailed 95% bootstrap confidence intervals for these false positive rates and statistical power estimates with 1,000 bootstrap iterations.
 
@@
 
-We searched the individual analyses from our systematic review for statistically significant cross-national correlations with available primary or secondary data.
+We searched the individual analyses from our literature review for statistically significant cross-national correlations with available primary or secondary data.
 We restricted our search to one analysis per paper, and searched until we had a set of twelve analyses, six from economic development papers and six from cultural values papers, for which we were able to replicate the original result (i.e.
@@
 One analysis[@Gelfand2011] additionally controlled for log gross national income, and another analysis[@Skidmore2002] is a multilevel model including random intercepts for nations and several individual-level and national-level covariates (see Model 5 in original paper).
-Before running any additional models, we pre-registered these twelve analyses on the Open Science Framework on 25^th^ January 2022 (https://osf.io/uywx8/).
+Before running any additional models, we pre-registered these twelve analyses on the Open Science Framework on 25^th^ January 2022 (https://osf.io/u8tbf).
 We endeavoured to keep the sample sizes of our reanalyses as close to the original analyses as possible, though there were some deviations (see Supplementary Table \@ref(tab:tableDeviations)).
@@
 
-We fitted these models using the _brms_ R package[@Burkner2017].
+We fitted these models using the brms R package[@Burkner2017].
 Our choice of priors was based on prior predictive simulation.
 All models converged normally ($\hat{R}$ < 1.1), though for some models we resorted to using approximate Gaussian processes[@Solin2020] to reach convergence.
+The assumptions of these models were met: residuals were approximately normally distributed, though this was not formally tested.
 
@@
 
-All data and code are accessible at our Open Science Framework repository (https://osf.io/uywx8/).
-We used the _targets_ R package[@Landau2021] to create a reproducible data analysis pipeline and the _papaja_ R package[@Aust2020] to reproducibly generate the manuscript.
+All data and code are accessible on GitHub (https://github.com/ScottClaessens/crossNationalCorrelations).
+We used the targets R package[@Landau2021] to create a reproducible data analysis pipeline and the papaja R package[@Aust2020] to reproducibly generate the manuscript.
+
+## Ethics
+
+Ethical approval was not required for the study as we analysed only secondary and simulated data. We did not apply for ethical approval through an ethics board.
 
@@
 
-# Acknowledgements
-
-This work was supported by a Royal Society of New Zealand Marsden grant (20-UOA123) to QDA.
-
-# Author Contributions
-
-SC and QDA conceived of and designed the study.
-SC curated the data, produced all code for analysis and visualisation, and wrote the original draft of the manuscript.
-TK and QDA developed and compiled the geographic and linguistic distance matrices.
-QDA provided funding and input on manuscript preparation and revision.
-All authors reviewed and edited the final draft of the manuscript.
-
-# Competing Interests
-
-The authors declare no competing interests.
-
 # Data Availability
 
-All data can be found on the Open Science Framework: https://osf.io/uywx8/
+All data can be found on GitHub (https://github.com/ScottClaessens/crossNationalCorrelations).
+Data on human development were retrieved from the United Nations Development Programme (https://hdr.undp.org/en/content/download-data).
+Data on GDP per capita, annual GDP per capita growth, and the Gini coefficient were retrieved from the World Bank (https://data.worldbank.org/).
+Data on traditional vs.
+secular values and survival vs.
+self-expression values were retrieved from the World Values Survey (https://www.worldvaluessurvey.org/wvs.jsp).
+Data on cultural tightness were retrieved from the OSF repository for Gelfand et al.
+2021 (https://osf.io/47pe8/).
+The review data generated in this study are provided in the Supplementary Data file.
+All other datasets (e.g., for replications) were retrieved from tables and supplementary tables directly from papers cited in the main text[@Adamczyk2009;@Alesina2013;@Beck2003;@Beck2005;@Bockstette2002;@Easterly2003;@Easterly2007;@Fincher2008;@Gelfand2011;@Inglehart2000;@Knack1997;@Skidmore2002] - these datasets can be found in our GitHub repository.
 
@@
 
-All code to reproduce the statistical analyses in this manuscript can be found on the Open Science Framework: https://osf.io/uywx8/
-
-# Ethical Approval
-
-Ethical approval was not required for the study as we analysed only secondary and simulated data.
+All code to reproduce the statistical analyses in this manuscript can be found on GitHub (https://github.com/ScottClaessens/crossNationalCorrelations).
 
@@
 
-# (APPENDIX) Appendix {-}
+# Acknowledgements
+
+This work was supported by a Royal Society of New Zealand Marsden grant (20-UOA123) to QDA.
+
+# Author Contributions Statement
+
+SC and QDA conceived of and designed the study.
+SC curated the data, produced all code for analysis and visualisation, and wrote the original draft of the manuscript.
+TK and QDA developed and compiled the geographic and linguistic distance matrices.
+QDA provided funding and input on manuscript preparation and revision.
+All authors reviewed and edited the final draft of the manuscript.
+
+# Competing Interests Statement
+
+The authors declare no competing interests.
+
+\newpage
+\vspace*{60mm}
 

### 2023-07-10 manuscript.tex
third resubmission

@@
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math}
+  \usepackage{unicode-math} % this also loads fontspec
   \defaultfontfeatures{Scale=MatchLowercase}
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
 \fi
@@
 {}{\PatchFailed}
-\keywords{spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses\newline\indent Word count: 4892 words excl. Methods (Methods: 1877 words)}
+\keywords{spatial non-independence; cultural non-independence; cross-national analyses; simulations; replications; reanalyses\newline\indent Word count: 5102 words excl. Methods (Methods: 2163 words)}
 \usepackage{lineno}
@@
 \captionsetup[table]{font={stretch=1.5}}
-\captionsetup[figure]{font={stretch=1.5}}
+\captionsetup[figure]{font={stretch=1,small}}
 \nolinenumbers
@@
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
+\urlstyle{same}
 \hypersetup{
-  pdftitle={The non-independence of nations and why it matters},
+  pdftitle={Cross-national analyses require additional controls to account for the non-independence of nations},
   pdfauthor={Scott Claessens1, Thanos Kyritsis1, \& Quentin D. Atkinson1,2},
@@
 
-\title{The non-independence of nations and why it matters}
+\title{Cross-national analyses require additional controls to account for the non-independence of nations}
 \author{Scott Claessens\textsuperscript{1}, Thanos Kyritsis\textsuperscript{1}, \& Quentin D. Atkinson\textsuperscript{1,2}}
@@
 
-\authornote{
-
-This working paper has not yet been peer-reviewed.
-
-\par
-
-Correspondence concerning this article should be addressed to Scott Claessens (\href{mailto:scott.claessens@gmail.com}{\nolinkurl{scott.claessens@gmail.com}}) and Quentin D.
+\authornote{Correspondence concerning this article should be addressed to Scott Claessens (\href{mailto:scott.claessens@gmail.com}{\nolinkurl{scott.claessens@gmail.com}}) and Quentin D.
 Atkinson (\href{mailto:q.atkinson@auckland.ac.nz}{\nolinkurl{q.atkinson@auckland.ac.nz}}), Floor 2, Building 302, 23 Symonds Street, Auckland, 1010, New Zealand.
-
-The authors made the following contributions.
+\newline \indent The authors made the following contributions.
 Scott Claessens: Conceptualization, Data curation, Formal analysis, Investigation, Software, Visualization, Writing - original draft, Writing - review \& editing; Thanos Kyritsis: Data curation, Writing - review \& editing; Quentin D.
-Atkinson: Conceptualization, Data curation, Funding acquisition, Supervision, Writing - review \& editing.
-
-}
-
-\affiliation{\vspace{0.5cm}\textsuperscript{1} School of Psychology, University of Auckland, Auckland, New Zealand\\\textsuperscript{2} Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany}
+Atkinson: Conceptualization, Data curation, Funding acquisition, Supervision, Writing - review \& editing.}
+
+\affiliation{\vspace{0.5cm}\textsuperscript{1} \footnotesize School of Psychology, University of Auckland, Auckland, New Zealand\\\textsuperscript{2} \footnotesize School of Anthropology and Museum Ethnography, University of Oxford, Oxford, United Kingdom}
 
@@
 
-Nations are perhaps the single most important population unit structuring modern human life.
+Nations are an important population unit structuring modern human life.
 The nation in which someone is born has a large effect on what they can expect out of life, including their income level\textsuperscript{1}, life expectancy\textsuperscript{2}, mental health\textsuperscript{3}, subjective well-being\textsuperscript{4}, and educational attainment\textsuperscript{5}.
@@
 This is known as cultural phylogenetic non-independence\textsuperscript{25--27}, and occurs because related nations share cultural traits inherited via descent from a common ancestor.
-Shared cultural ancestry can result in a form of pseudoreplication, whereby multiple instances of the same trait across nations are merely duplicates of the ancestral original.
-For example, the related island nations Tonga and Tuvalu share similar languages and customs due to cultural inheritance from a common Polynesian population dating back more than 1,000 years.
+Shared cultural ancestry can result in a form of pseudoreplication, whereby multiple instances of the same trait across nations are merely duplicates of the ancestral original\textsuperscript{25,28}.
+For example, the related island nations Tonga and Tuvalu share similar languages and customs due to cultural inheritance from a common Polynesian population dating back more than 1,000 years\textsuperscript{29,30}.
 More recently, shared ancestry explains similarities in institutions, norms, technologies, and values between colonial settlements and their colonisers (e.g.~English, French, Spanish, and Portuguese settlements of the Americas).
@@
 Spatial and cultural phylogenetic non-independence between nations make cross-national inference challenging.
-A fundamental assumption of regression analysis is that model residuals should be independently and identically distributed\textsuperscript{28}.
+A fundamental assumption of regression analysis is that model residuals should be independently and identically distributed\textsuperscript{31}.
 But without accounting for spatial or cultural non-independence between nations, model residuals can show structure that remains unaccounted for, violating this assumption.
-Treating nations as independent can thus inflate false positive rates\textsuperscript{29}, producing spurious ``direct'' relationships between variables that in fact only indirectly covary due to spatial or cultural connections\textsuperscript{30} (see Supplementary Figure \ref{fig:dag} for an example causal model).
+Treating nations as independent can thus inflate false positive rates\textsuperscript{32}, producing spurious relationships between variables that in fact only indirectly covary due to spatial or cultural connections\textsuperscript{33} (see Supplementary Figure \ref{fig:dag} for an example causal model).
 
 Non-independence is widely acknowledged in fields that routinely deal with spatially or culturally structured data.
-In ecology and sociology, studies with regional-level data use a variety of autoregressive models designed to account for spatial patternings\textsuperscript{31,32}.
+In ecology and sociology, studies with regional-level data use a variety of autoregressive models designed to account for spatial patternings\textsuperscript{34,35}.
 In anthropology, researchers have recognised cultural non-independence as an important statistical pitfall for over 200 years, with issues of cultural pseudoreplication being identified in early comparative studies of marriage practices across societies\textsuperscript{25}.
-In the twentieth century, anthropologists began to emphasise that human societies do not develop independently, but rather exist in a globally interconnected ``world system'' linked by shared history and cultural ancestry\textsuperscript{33}.
-In order to minimise the confounding effects of this non-independence in comparisons of human societies, researchers compiled the Standard Cross-Cultural Sample of 186 cultures which deliberately avoids sampling closely related cultures\textsuperscript{34}, though it is difficult to completely remove spatial and cultural dependencies\textsuperscript{35,36} and the smaller sample of cultures reduces statistical power.
-Today, anthropologists borrow phylogenetic comparative methods from evolutionary biology, such as phylogenetic least squares regression\textsuperscript{37}, when comparing societies, treating culturally related societies in the same way as biologists treat genetically related species (e.g.\textsuperscript{38,39}).
+In the twentieth century, anthropologists began to emphasise that human societies do not develop independently, but rather exist in a globally interconnected world system linked by shared history and cultural ancestry\textsuperscript{36}.
+In order to minimise the confounding effects of this non-independence in comparisons of human societies, researchers compiled the Standard Cross-Cultural Sample of 186 cultures which deliberately avoids sampling closely related cultures\textsuperscript{28}, though it is difficult to completely remove spatial and cultural dependencies\textsuperscript{37,38} and the smaller sample of cultures reduces statistical power.
+Today, anthropologists borrow phylogenetic comparative methods from evolutionary biology, such as phylogenetic least squares regression\textsuperscript{39}, when comparing societies, treating culturally related societies in the same way as biologists treat genetically related species (e.g.\textsuperscript{40,41}).
 
 At the national level, recent reanalyses have revealed that several cross-national relationships reported in economics and psychology do not hold when controlling for non-independence between nations.
-One working paper replicated 25 analyses of ``persistence'' in economics, in which modern national outcomes are regressed against historical characteristics of those nations, and found that over half of the relationships were attenuated when controlling for spatial non-independence\textsuperscript{40}.
-Another replication study found that many of the widely publicised relationships between national-level pathogen prevalence and political institutions and attitudes fail to hold when controlling for various kinds of non-independence\textsuperscript{41}.
-These reanalyses, and others\textsuperscript{42--44}, raise the question: how widespread a concern is non-independence in studies of national-level outcomes?
+One working paper replicated 25 analyses of persistence in economics, in which modern national outcomes are regressed against historical characteristics of those nations, and found that over half of the relationships were attenuated when controlling for spatial non-independence\textsuperscript{42}.
+Another replication study found that many of the widely publicised relationships between national-level pathogen prevalence and political institutions and attitudes fail to hold when controlling for various kinds of non-independence\textsuperscript{43}.
+These reanalyses, and others\textsuperscript{44--46}, raise the question: how widespread a concern is non-independence in studies of national-level outcomes?
 
@@
 To this end, we used Bayesian multilevel models to simultaneously estimate geographic and cultural phylogenetic signal for a range of economic development and cultural values variables.
-For economic development, we focused on the Human Development Index\textsuperscript{45}, gross domestic product per capita, annual gross domestic product per capita growth, and the Gini index of income inequality.
+For economic development, we focused on the Human Development Index\textsuperscript{47}, gross domestic product per capita, annual gross domestic product per capita growth, and the Gini index of income inequality.
 For cultural values, we focused on two primary dimensions of cultural values from the World Values Survey, traditional vs.~secular values and survival vs.~self-expression values\textsuperscript{16}, as well as cultural tightness\textsuperscript{14} and individualism\textsuperscript{15}.
@@
 For all of these variables, we found that a substantial proportion of national-level variation was explained by spatial proximity and/or shared cultural ancestry between nations (Figure \ref{fig:plotSignal}; see Supplementary Table \ref{tab:tableSignal} for numerical results).
-Signal estimates were often markedly strong, with spatial proximity and shared cultural ancestry frequently explaining over half of the national-level variation.
+Signal estimates were often strong, with spatial proximity and shared cultural ancestry frequently explaining over half of the national-level variation.
 For spatial proximity, Bayes Factors indicated strong evidence that the geographic signal estimates differed from zero for all economic development variables and traditional values.
@@
 \includegraphics{manuscript_files/figure-latex/plotSignal-1.pdf}
-\caption{\label{fig:plotSignal}\emph{Posterior estimates of geographic and cultural phylogenetic signal for a range of economic development and cultural values variables.} Geographic and cultural phylogenetic signal are operationalised as the proportion of national-level variance explained by geographic and linguistic proximity matrices.
-Grey ridges are full posterior distributions, points are posterior median values, and black lines are 50\% and 95\% credible intervals.
+\caption{\label{fig:plotSignal}Posterior estimates of geographic and cultural phylogenetic signal for a range of economic development and cultural values variables.
+Geographic and cultural phylogenetic signal are operationalised as the proportion of national-level variance explained by geographic and linguistic proximity matrices.
+Grey ridges are full posterior distributions, points are posterior median values, and black lines are equal-tailed 50\% and 95\% credible intervals.
+Number of observations from these models are as follows: HDI (n = 1449), GDP per capita (n = 10289), GDP per capita growth (n = 9755), Gini index (n = 1826), traditional values (n = 277676), survival values (n = 277676), tightness (n = 57), and individualism (n = 67).
 HDI = Human Development Index; GDP = gross domestic product.}
@@
 Given that economic development and cultural values show evidence of geographic and cultural phylogenetic signal, have cross-national analyses sufficiently accounted for this non-independence?
-To assess this, we systematically searched the published literature for articles that combined the search terms ``economic development'' or ``values'' with the search terms ``cross-national'', ``cross-cultural'', or ``cross-country''.
+To assess this, we searched the published literature for articles that combined the search terms ``economic development'' or ``values'' with the search terms ``cross-national'', ``cross-cultural'', or ``cross-country''.
 We removed articles that did not report original research, were not relevant to economic development or cultural values, or did not report at least one cross-national analysis.
-We then retained the 100 articles (50 for economic development, 50 for cultural values) with the highest annual rate of citations (Supplementary Table \ref{tab:tableArticlesReview}).
+We then retained the 100 articles (50 for economic development, 50 for cultural values) with the highest annual rate of citations (see Supplementary Data).
 For each of these highly-cited articles, we exhaustively recorded every cross-national analysis reported in the main text (n = 4,308), identifying in each case whether or not the analysis attempted to control for spatial, cultural, or any other form of non-independence between nations (see Methods for detailed search criteria and coding decisions).
 
-The results of our systematic review show that most published articles containing cross-national analyses make no attempt to account for statistical non-independence.
+The results of our literature review show that most published articles containing cross-national analyses make no attempt to account for statistical non-independence.
 Figure \ref{fig:plotReview1}a plots the proportion of articles that contain at least one cross-national analysis accounting for non-independence.
-We find that 42\% of economic development articles contain at least one attempt to control for non-independence (95\% bootstrap confidence interval {[}0.30 0.54{]}), while this proportion decreases to only 8\% for cultural values articles (95\% bCI {[}0.02 0.16{]}).
+We find that 42\% of economic development articles contain at least one attempt to control for non-independence (two-tailed 95\% bootstrap confidence interval {[}0.30 0.54{]}), while this proportion decreases to only 8\% for cultural values articles (two-tailed 95\% bCI {[}0.02 0.16{]}).
 Both kinds of article are most likely to use regional fixed effects (e.g.~continent fixed effects) to account for non-independence, but some articles also include controls for spatial distance (e.g.~latitude) and shared cultural history (e.g.~colony status).
-These proportions are even lower when focusing on the full sample of 4,308 analyses: only 5\% (95\% credible interval {[}0.02 0.13{]}) of individual economic development analyses and 1\% (95\% CI {[}0.00 0.02{]}) of individual cultural values analyses are estimated to control for non-independence (Supplementary Figure \ref{fig:plotReview2}).
+These proportions are even lower when focusing on the full sample of 4,308 analyses: only 5\% (equal-tailed 95\% credible interval {[}0.02 0.13{]}) of individual economic development analyses and 1\% (equal-tailed 95\% CI {[}0.00 0.02{]}) of individual cultural values analyses are estimated to control for non-independence (Supplementary Figure \ref{fig:plotReview2}).
 
@@
 \includegraphics{manuscript_files/figure-latex/plotReview1-1.pdf}
-\caption{\label{fig:plotReview1}\emph{Results from systematic review of 100 highly-cited cross-national studies of economic development (red) and cultural values (blue).} (a) Proportion of articles containing at least one analysis accounting for non-independence, overall and split by common methods of controlling for non-independence.
+\caption{\label{fig:plotReview1}Results from literature review of 100 highly-cited cross-national studies of economic development (red) and cultural values (blue).
+(a) Proportion of articles containing at least one analysis accounting for non-independence, overall and split by common methods of controlling for non-independence.
+Points represent raw proportions of articles and ranges represent two-tailed 95\% bootstrap confidence intervals (n = 1000 bootstrap samples).
 (b) The association between journal impact factor and the probability that an article contains at least one analysis accounting for non-independence.
 (c) Estimated trend over time for the probability that an article contains at least one analysis accounting for non-independence.
-Point ranges represent proportions and 95\% bootstrap confidence intervals.
-Lines and shaded areas are posterior median regression lines and 50\% credible intervals from Bayesian logistic regression models.
+Lines and shaded areas are posterior median regression lines and equal-tailed 50\% credible intervals from Bayesian logistic regression models (n = 100 observations).
 Dots represent raw counts of individual articles that did (top) or did not (bottom) account for non-independence.
@@
 
-While our review contains articles from journals with a range of impact factors, an anonymous reviewer suggested that our estimates could be biased downwards by analyses published in lower impact outlets with more relaxed standards for issues like non-independence.
-It is also possible that, since our systematic review goes back as far as 1993, our estimates are being biased downwards by earlier studies, and that controls for non-independence have increased over time with methodological advancements and greater awareness of the issue.
+While our review contains articles from journals with a range of impact factors, our estimates could be biased downwards by analyses published in lower impact outlets with more relaxed standards for issues like non-independence.
+It is also possible that, since our literature review goes back as far as 1993, our estimates are being biased downwards by earlier studies, and that controls for non-independence have increased over time with methodological advancements and greater awareness of the issue.
 To test these possible explanations for our low estimates, we fitted logistic regression models to the data from the review, including log journal impact factor and publication year as separate predictors.
-Interestingly, we found that, if anything, studies from higher impact journals were \emph{less} likely to include at least one control for non-independence than studies from lower impact journals, both for studies of economic development (b = -0.38, 95\% CI {[}-0.87 0.09{]}) and for studies of cultural values (b = -0.56, 95\% CI {[}-1.10 -0.05{]}; Figure \ref{fig:plotReview1}b).
-Moreover, splines revealed no relationship between publication year and the probability of including at least one control for non-independence, both for studies of economic development (b = -0.13, 95\% CI {[}-1.10 0.82{]}) and for studies of cultural values (b = -0.02, 95\% CI {[}-1.00 0.96{]}; Figure \ref{fig:plotReview1}c).
+Interestingly, we found that, if anything, studies from higher impact journals were less likely to include at least one control for non-independence than studies from lower impact journals, both for studies of economic development (b = -0.38, equal-tailed 95\% CI {[}-0.87 0.09{]}) and for studies of cultural values (b = -0.56, equal-tailed 95\% CI {[}-1.10 -0.05{]}; Figure \ref{fig:plotReview1}b).
+Moreover, splines revealed no relationship between publication year and the probability of including at least one control for non-independence, both for studies of economic development (b = -0.13, equal-tailed 95\% CI {[}-1.10 0.82{]}) and for studies of cultural values (b = -0.02, equal-tailed 95\% CI {[}-1.00 0.96{]}; Figure \ref{fig:plotReview1}c).
 
@@
 We then fitted naive regressions without controls to these datasets, as well as regression models with controls for latitude, longitude, and continent fixed effects.
-Despite not being identified in our systematic review, we also included other methods that are often used in the literature to account for non-independence.
-Additional spatial controls included the mean of the predictor variable within a surrounding 2000km radius (e.g.\textsuperscript{46}) and Conley standard errors\textsuperscript{47,48} based on geographic distances between nations (e.g.\textsuperscript{19,46}).
-Additional cultural controls included fixed effects for the language families of the majority-spoken languages in each nation (e.g.\textsuperscript{49}) and Conley standard errors based on genetic distances between nations (e.g.\textsuperscript{19,46}).
+Despite not being identified in our literature review, we also included other methods that are often used in the literature to account for non-independence.
+Additional spatial controls included the mean of the predictor variable within a surrounding 2000km radius (e.g.\textsuperscript{48}) and Conley standard errors\textsuperscript{49,50} based on geographic distances between nations (e.g.\textsuperscript{19,48}).
+Additional cultural controls included fixed effects for the language families of the majority-spoken languages in each nation (e.g.\textsuperscript{51}) and Conley standard errors based on genetic distances between nations (e.g.\textsuperscript{19,48}).
 These fixed effect approaches attempt to account for non-independence by holding geographic location constant (latitude, longitude), discarding between-region variation and exploiting only local variation (continent fixed effects, mean of surrounding 2000km), or correcting standard errors for autocorrelation post-hoc while leaving model coefficients unchanged (Conley standard errors).
@@
 Linguistic proximity between nations is calculated from a global phylogenetic tree that represents hierarchical relationships of genealogical descent for all languages in the world.
-For each pair of nations, we calculated inverse phylogenetic distances (i.e.~number of branches separating two taxa) between all languages spoken in that nation pair and produced an average ``linguistic proximity'' score weighted by the percentages of speakers within those nations.
-To include the resulting geographic and linguistic proximity matrices in our models, we included a Gaussian process\textsuperscript{50,51} over latitude and longitude values and/or assumed that nation random intercepts were correlated in proportion to their linguistic proximity\textsuperscript{52}.
+For each pair of nations, we calculated inverse phylogenetic distances (i.e.~number of branches separating two taxa) between all languages spoken in that nation pair and produced an average linguistic proximity score weighted by the percentages of speakers within those nations.
+To include the resulting geographic and linguistic proximity matrices in our models, we included a Gaussian process\textsuperscript{52,53} over latitude and longitude values and/or assumed that nation random intercepts were correlated in proportion to their linguistic proximity\textsuperscript{54}.
 These random effects approaches attempt to account for non-independence by modelling the covariance between nations that is induced by their geographic or linguistic connections.
@@
 Figures \ref{fig:plotSim3} and \ref{fig:plotSim4} plot the estimated false positive rates from our simulation study, split by different methods and different degrees of spatial or cultural phylogenetic autocorrelation (see Supplementary Tables \ref{tab:tableSim1} and \ref{tab:tableSim2} for numerical results and Supplementary Figures \ref{fig:plotSim1} and \ref{fig:plotSim2} for full distributions of effect sizes under strong autocorrelation).
-Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero, falsely inferring a relationship when none is present.
-For reference, ``weak'' autocorrelation in our simulation is comparable to the geographic signal for survival values in Figure \ref{fig:plotSignal} (i.e.~20\% of the national-level variance is explained by non-independence), while ``moderate'' and ``strong'' levels of autocorrelation are comparable to the cultural phylogenetic signal for traditional and survival values, respectively (i.e.~50\% and 80\% of the national-level variance is explained by non-independence).
+Across all model types, false positive rates were measured as the proportion of models that estimated a slope with a two-tailed 95\% confidence / credible interval excluding zero, falsely inferring a relationship when none is present.
+For reference, weak autocorrelation in our simulation is comparable to the geographic signal for survival values in Figure \ref{fig:plotSignal} (i.e.~20\% of the national-level variance is explained by non-independence), while moderate and strong levels of autocorrelation are comparable to the cultural phylogenetic signal for traditional and survival values, respectively (i.e.~50\% and 80\% of the national-level variance is explained by non-independence).
 
@@
 \includegraphics{manuscript_files/figure-latex/plotSim3-1.pdf}
-\caption{\label{fig:plotSim3}\emph{False positive rates for different methods of controlling for spatial non-independence in our simulation study.} For simulated outcome and predictor variables, we systematically varied the strength of spatial autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+\caption{\label{fig:plotSim3}False positive rates for different methods of controlling for spatial non-independence in our simulation study.
+For simulated outcome and predictor variables, we systematically varied the strength of spatial autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
 We simulated 100 datasets per parameter combination and fitted different models to each dataset.
-False positive rates were operationalised as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero.
-Point ranges represent proportions and 95\% bootstrap confidence intervals, and dashed lines indicate the 5\% false positive rate that is expected due to chance.
+False positive rates were operationalised as the proportion of models that estimated a slope with a two-tailed 95\% confidence / credible interval excluding zero.
+Points represent raw proportions of false positive models, ranges represent two-tailed 95\% bootstrap confidence intervals (n = 1000 bootstrap samples), and dashed lines indicate the 5\% false positive rate that is expected due to chance.
+Colours indicate whether the strength of autocorrelation for the predictor variable is 0.2 (red), 0.5, (green) or 0.8 (blue).
 SEs = standard errors.}
@@
 \includegraphics{manuscript_files/figure-latex/plotSim4-1.pdf}
-\caption{\label{fig:plotSim4}\emph{False positive rates for different methods of controlling for cultural phylogenetic non-independence in our simulation study.} For simulated outcome and predictor variables, we systematically varied the strength of cultural phylogenetic autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
+\caption{\label{fig:plotSim4}False positive rates for different methods of controlling for cultural phylogenetic non-independence in our simulation study.
+For simulated outcome and predictor variables, we systematically varied the strength of cultural phylogenetic autocorrelation, from weak (0.2) to moderate (0.5) to strong (0.8).
 We simulated 100 datasets per parameter combination and fitted different models to each dataset.
-False positive rates were operationalised as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero.
-Point ranges represent proportions and 95\% bootstrap confidence intervals, and dashed lines indicate the 5\% false positive rate that is expected due to chance.
+False positive rates were operationalised as the proportion of models that estimated a slope with a two-tailed 95\% confidence / credible interval excluding zero.
+Points represent raw proportions of false positive models, ranges represent two-tailed 95\% bootstrap confidence intervals (n = 1000 bootstrap samples), and dashed lines indicate the 5\% false positive rate that is expected due to chance.
+Colours indicate whether the strength of autocorrelation for the predictor variable is 0.2 (red), 0.5, (green) or 0.8 (blue).
 SEs = standard errors.}
@@
 However, additional simulations revealed that these reductions in false positive rates come at the cost of lower statistical power (see Supplementary Figures \ref{fig:plotPowerGeo01} - \ref{fig:plotPowerGeo05}).
-In additional simulations where the true relationship between the predictor and outcome variable was known, power analyses showed that both continent fixed effects and the 2000km radius control had less than 80\% power to detect moderate true correlations (\emph{r} = 0.3) under strong spatial autocorrelation.
+In additional simulations where the true relationship between the predictor and outcome variable was known, power analyses showed that both continent fixed effects and the 2000km radius control had less than 80\% power to detect moderate true correlations (r = 0.3) under strong spatial autocorrelation.
 
@@
 False positives are not completely eliminated with these random effects models.
-Nevertheless, these methods have at least 80\% power to detect moderate (\emph{r} = 0.3) and large (\emph{r} = 0.5) true correlations between variables under all levels of spatial autocorrelation (see Supplementary Figures \ref{fig:plotPowerGeo01} - \ref{fig:plotPowerGeo05}).
+Nevertheless, these methods have at least 80\% power to detect moderate (r = 0.3) and large (r = 0.5) true correlations between variables under all levels of spatial autocorrelation (see Supplementary Figures \ref{fig:plotPowerGeo01} - \ref{fig:plotPowerGeo05}).
 
@@
 By contrast, models with random effects covarying according to linguistic proximity completely eliminate false positives across all degrees of cultural phylogenetic autocorrelation.
-This approach is also the only method that is able to detect large true correlations (\emph{r} = 0.5) with at least 80\% power (see Supplementary Figures \ref{fig:plotPowerLin01} - \ref{fig:plotPowerLin05}).
+This approach is also the only method that is able to detect large true correlations (r = 0.5) with at least 80\% power (see Supplementary Figures \ref{fig:plotPowerLin01} - \ref{fig:plotPowerLin05}).
 Random effects models that additionally account for geographic proximity between nations perform equally well, though models with only a spatial Gaussian process continue to produce false positives.
@@
 
-Our systematic review and simulation study have shown that controls for non-independence are rare in cross-national studies of economic development and cultural values, and when studies do attempt to control for non-independence, the methods typically used are unable to sufficiently reduce false positive rates.
+Our literature review and simulation study have shown that controls for non-independence are rare in cross-national studies of economic development and cultural values, and when studies do attempt to control for non-independence, the methods typically used are unable to sufficiently reduce false positive rates.
 This raises the worrying possibility that the cross-national literature in economics and psychology is populated with spurious relationships.
 
-To determine how widespread this issue of spurious cross-national relationships might be, we reanalysed a subset of twelve previous cross-national analyses from our systematic review, controlling for spatial and cultural phylogenetic non-independence using global geographic and linguistic proximity matrices.
-Out of the 100 papers included in our systematic review, primary or secondary data were publicly available for 47 papers.
+To determine how widespread this issue of spurious cross-national relationships might be, we reanalysed a subset of twelve previous cross-national analyses from our literature review, controlling for spatial and cultural phylogenetic non-independence using global geographic and linguistic proximity matrices.
+Out of the 100 papers included in our literature review, primary or secondary data were publicly available for 47 papers.
 We attempted to replicate key statistically significant cross-national correlations from these papers --- mostly initial bivariate regression specifications without covariates (see Methods) --- and stopped when we had sampled a set of twelve analyses for which we were able to replicate the original result.
-The final set included six analyses from our economic development review\textsuperscript{53--58} and six from our cultural values review\textsuperscript{13,14,16,59--61} that had available data and were able to be replicated.
+The final set included six analyses from our economic development review\textsuperscript{55--60} and six from our cultural values review\textsuperscript{13,14,16,61--63} that had available data and were able to be replicated.
 We pre-registered this set before running any control models (\url{https://osf.io/uywx8/}).
@@
 \includegraphics{manuscript_files/figure-latex/plotReplications1-1.pdf}
-\caption{\label{fig:plotReplications1}\emph{Posterior correlations from our reanalysis of twelve previous cross-national analyses.} For each previous cross-national relationship, we plot the posterior slopes from a naive regression (red), a regression controlling for spatial non-independence (green), a regression controlling for cultural phylogenetic non-independence (blue), and a regression controlling for both spatial and cultural phylogenetic non-independence simultaneously (purple).
+\caption{\label{fig:plotReplications1}Posterior correlations from our reanalysis of twelve previous cross-national analyses.
+For each previous cross-national relationship, we plot the posterior slopes from a naive regression (red), a regression controlling for spatial non-independence (green), a regression controlling for cultural phylogenetic non-independence (blue), and a regression controlling for both spatial and cultural phylogenetic non-independence simultaneously (purple).
 All outcome and predictor variables are standardised.
 Most analyses are simple bivariate cross-national correlations, but Gelfand et al.~(2011) is a partial correlation controlling for log gross national income and Adamczyk and Pitt (2009) is a multilevel model including several covariates.
-Coloured ridges are full posterior distributions, and points and black lines represent posterior medians and 95\% credible intervals.
+Coloured ridges are full posterior distributions, and points and black lines represent posterior medians and equal-tailed 95\% credible intervals.
+Numbers of observations from the models are as follows, from top to bottom: n = 69, n = 45, n = 103, n = 98, n = 63, n = 89, n = 33, n = 75, n = 67, n = 28, n = 38, and n = 28.
 GDP = gross domestic product.
@@
 
+To understand why some cross-national correlations were attenuated by controls for non-independence while others were robust, we further explored our fitted models for evidence of spatial and cultural autocorrelation.
+For each outcome variable, our Gaussian process models provided varying estimates of how quickly spatial autocorrelation declined with distance (Supplementary Figure \ref{fig:plotReplications3}).
+For example, in Skidmore and Toya\textsuperscript{60} gross domestic product growth was only moderately spatially autocorrelated at 1,000 km distance (posterior median spatial autocorrelation at 1,000 km = 0.42, equal-tailed 95\% CI {[}0.07 0.90{]}), whereas in Inglehart and Baker\textsuperscript{16} traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1,000 km = 0.96, 95\% CI {[}0.81 0.99{]}).
+We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \ref{fig:plotReplications4}), with some outcome variables expressing low signal (e.g.~confidence in institutions\textsuperscript{63}; posterior median = 0.07, 95\% CI {[}0.00 0.53{]}) and others expressing high signal (e.g.~female labour force participation\textsuperscript{62}; posterior median = 0.89, 95\% CI {[}0.63 0.98{]}).
+Exploratory regressions provided suggestive evidence that stronger estimates of spatial autocorrelation or cultural phylogenetic signal resulted in a more pronounced reduction in the effect size when controlling for non-independence between nations (Supplementary Figure \ref{fig:plotReplications5}).
+However, these negative slopes for spatial autocorrelation (b = -0.19, 95\% CI {[}-1.33 1.04{]}) and cultural phylogenetic signal (b = -0.28, 95\% CI {[}-1.26 0.70{]}) were both very uncertain, due to the small number of analyses and the posterior uncertainty in effect sizes and estimates of non-independence.
+
 \hypertarget{discussion}{%
@@
 
-In a systematic literature review and simulation, we found that cross-national studies in economics and psychology rarely account for non-independence between nations, and, when they do, the methods they use are insufficient to reduce false positives in non-independent data.
+In a literature review and simulation, we found that cross-national studies in economics and psychology rarely account for non-independence between nations, and, when they do, the methods they use are insufficient to reduce false positives in non-independent data.
 In a reanalysis of twelve cross-national correlations, we further showed that neglecting to account for non-independence has resulted in potentially spurious relationships in the published literature, with half of the correlations failing to replicate when controlling for spatial or cultural non-independence with more rigorous methods.
@@
 Our initial analyses add to and clarify existing evidence regarding the degree of non-independence for national-level economic and cultural variables.
-One previous study suggested that geographic proximity is more important than deep cultural ancestry in explaining the distribution of human development across Eurasian nations, though the authors noted that their small sample of 44 nations and regional focus limited their statistical power\textsuperscript{62}.
+One previous study suggested that geographic proximity is more important than deep cultural ancestry in explaining the distribution of human development across Eurasian nations, though the authors noted that their small sample of 44 nations and regional focus limited their statistical power\textsuperscript{64}.
 By contrast, our global samples of over 160 nations revealed strong cultural phylogenetic signal, as well as geographic signal, for the Human Development Index, GDP per capita, and the Gini index of inequality.
@@
 
-Crucially, our systematic literature review and simulation study revealed that the most commonly used controls for non-independence do not sufficiently deal with the issue.
+Crucially, our literature review and simulation study revealed that the most commonly used controls for non-independence do not sufficiently deal with the issue.
 In our simulations, controlling for either latitude or longitude did not reduce false positive rates.
 This result calls into question the use of controls like distance to the equator to account for non-independence in cross-national regression models, though these controls may still be suitable to account for regional or latitudinal variation in climate, ecology, and natural threats (e.g.~pathogens), which we did not simulate.
-High false positive rates persisted with Conley standard errors, despite recent claims that these standard error corrections are sufficient to deal with spatial non-independence\textsuperscript{63}.
-The simulation also confirmed the assertion that fixed effects for spatial or cultural groupings (e.g.~continent or language family fixed effects) are insufficient because non-independence still remains within groupings\textsuperscript{41}.
-This logic further applies to analyses that control for non-independence by separately analysing different regions (e.g.\textsuperscript{64}).
-Controlling for the mean of the predictor variable within a 2000km radius\textsuperscript{46} eliminated false positive rates in spatially autocorrelated data, but had reduced statistical power to detect true associations.
-Across all model types in our simulation, the only methods that markedly reduced false positive rates while retaining high statistical power were the random effects models with covariance matrices.
-The important advantage of these models is not that they are Bayesian \emph{per se} (any of the approaches used here could be implemented in a Bayesian framework), but rather that they explicitly model covariance as a function of spatial or cultural distance.
+High false positive rates persisted with Conley standard errors, despite recent claims that these standard error corrections are sufficient to deal with spatial non-independence\textsuperscript{65}.
+The simulation also confirmed the assertion that fixed effects for spatial or cultural groupings (e.g.~continent or language family fixed effects) are insufficient because non-independence still remains within groupings\textsuperscript{43}.
+This logic further applies to analyses that control for non-independence by separately analysing different regions (e.g.\textsuperscript{66}).
+Controlling for the mean of the predictor variable within a 2000km radius\textsuperscript{48} eliminated false positive rates in spatially autocorrelated data, but had reduced statistical power to detect true associations.
+Across all model types in our simulation, the only methods that reduced false positive rates while retaining high statistical power were the random effects models with covariance matrices.
+The important advantage of these models is not that they are Bayesian per se (any of the approaches used here could be implemented in a Bayesian framework), but rather that they explicitly model covariance as a function of spatial or cultural distance.
 
 There are other approaches to controlling for non-independence that we did not include in our simulation.
-For example, conditional autoregressive models\textsuperscript{31} and generalised additive models\textsuperscript{65} are approaches that can be applied in both frequentist and Bayesian frameworks.
-There are also alternative ways to operationalise cultural distances between nations beyond linguistic distances, including metrics like cultural fixation indices (e.g.~cultural \(F_{ST}\)\textsuperscript{6,66}), covariance based on genetic distances\textsuperscript{67}, and phylogenetic distances between religious traditions.
+For example, conditional autoregressive models\textsuperscript{34} and generalised additive models\textsuperscript{67} are approaches that can be applied in both frequentist and Bayesian frameworks.
+There are also alternative ways to operationalise cultural distances between nations beyond linguistic distances, including metrics like cultural fixation indices (e.g.~cultural \(F_{ST}\)\textsuperscript{6,68}), covariance based on genetic distances\textsuperscript{69}, and phylogenetic distances between religious traditions.
 We see merit in each of these approaches and the use of one over another will depend on the specific question at hand.
-We decided to focus on linguistic distances in this study since language is a tangible socially-learned trait that has previously been used to successfully track the effects of deep cultural ancestry on modern national outcomes\textsuperscript{62,68,69}.
+We decided to focus on linguistic distances in this study since language is a tangible socially-learned trait that has previously been used to successfully track the effects of deep cultural ancestry on modern national outcomes\textsuperscript{64,70,71}.
 Future work should explore whether other approaches are sufficient to reduce false positive rates in spatially and culturally non-independent data.
-In addition to spatial proximity and shared cultural ancestry, we did not simulate other sources of non-independence that potentially exist in real cross-national datasets, such as modern connections between nations due to flows of people and information (e.g.~flight networks, social media networks) and shared histories of colonialism and capitalist modes of production that have shaped today's global landscape\textsuperscript{70,71}.
+In addition to spatial proximity and shared cultural ancestry, we did not simulate other sources of non-independence that potentially exist in real cross-national datasets, such as modern connections between nations due to flows of people and information (e.g.~flight networks, social media networks) and shared histories of colonialism and capitalist modes of production that have shaped today's global landscape\textsuperscript{72,73}.
 Additional controls will be required to ensure that these sources of non-independence do not confound cross-national inferences.
@@
 Ours is not the first review to show that studies are misapplying statistical methods in ways that inflate false positive rates.
-For example, other literature reviews have shown that studies in the social sciences tend to use small samples of participants\textsuperscript{72}, treat ordinal data as metric\textsuperscript{73}, incorrectly handle missing values\textsuperscript{74}, and ignore best practices in meta-analyses\textsuperscript{75}.
+For example, other literature reviews have shown that studies in the social sciences tend to use small samples of participants\textsuperscript{74}, treat ordinal data as metric\textsuperscript{75}, incorrectly handle missing values\textsuperscript{76}, and ignore best practices in meta-analyses\textsuperscript{77}.
 Why do cross-national studies also rarely account for non-independence?
-At the institutional level, one possibility is that such practices are incentivised \emph{because} they generate statistically significant relationships, which increase the probability that a study is published\textsuperscript{72}.
-Indeed, we found that controls for non-independence were \emph{less} common among articles published in high impact journals, suggesting that researchers are rewarded for such practices.
+At the institutional level, one possibility is that such practices are incentivised because they generate statistically significant relationships, which increase the probability that a study is published\textsuperscript{74}.
+Indeed, we found that controls for non-independence were less common among articles published in high impact journals, suggesting that researchers are rewarded for such practices.
 At the individual level, another possibility is that researchers outside of anthropology and ecology are less aware of the problem, or believe that the problem does not apply to analyses of nations.
@@
 These institutional- and individual-level barriers can be combatted.
-First, cross-national replication studies like ours and others\textsuperscript{40--44}, combined with the methodological reviews included in Registered Reports\textsuperscript{76}, might change incentive structures and encourage researchers to analyse the world's nations with more rigorous methods.
-Second, more explicit descriptions of causal models could promote controls for non-independence by clearly outlining the nature of confounding and the sources of autocorrelation in cross-national data\textsuperscript{77}.
+First, cross-national replication studies like ours and others\textsuperscript{42--46}, combined with the methodological reviews included in Registered Reports\textsuperscript{78}, might change incentive structures and encourage researchers to analyse the world's nations with more rigorous methods.
+Second, more explicit descriptions of causal models could promote controls for non-independence by clearly outlining the nature of confounding and the sources of autocorrelation in cross-national data\textsuperscript{79}.
 The causal model outlined in Supplementary Figure \ref{fig:dag} is a useful example, but individual studies must outline their own particular causal assumptions, which may include further sources of non-independence and confounding variables to control for (e.g.~post-communist status, colony status).
@@
 Indeed, in our review, economists studying economic development dealt with national-level non-independence more than psychologists studying cultural values, likely because economics studies tend to be lengthy statistical exercises that systematically incorporate or exclude numerous variables in an attempt to infer causation.
-Third, the recent widespread accessibility of open source statistical software, such as the programming language Stan\textsuperscript{78} and the R package \emph{brms}\textsuperscript{79}, should promote the use of more rigorous methods to control for non-independence.
-Using \emph{brms}, for example, Bayesian Gaussian process regression is straightforward to conduct, requiring only longitude and latitude values for nations.
+Third, the recent widespread accessibility of open source statistical software, such as the programming language Stan\textsuperscript{80} and the R package brms\textsuperscript{81}, should promote the use of more rigorous methods to control for non-independence.
+Using brms, for example, Bayesian Gaussian process regression is straightforward to conduct, requiring only longitude and latitude values for nations.
 We have provided an online tutorial to help researchers apply these methods to their own cross-cultural datasets (\url{https://scottclaessens.github.io/blog/2022/crossnational/}).
@@
 Until such changes are implemented and sufficient controls for non-independence are the norm, existing cross-national correlations should be interpreted with caution.
-In our reanalyses, we found that half of the cross-national correlations had 95\% credible intervals that included zero when controlling for spatial and/or cultural phylogenetic non-independence.
-While these results are in line with previous reanalyses\textsuperscript{40}, we note that we are unable to outright reject the claims from these studies, since we only reanalysed the first bivariate regression specifications presented in the papers\textsuperscript{63}.
+In our reanalyses, we found that half of the cross-national correlations had equal-tailed 95\% credible intervals that included zero when controlling for spatial and/or cultural phylogenetic non-independence.
+While these results are in line with previous reanalyses\textsuperscript{42}, we note that we are unable to outright reject the claims from these studies, since we only reanalysed the first bivariate regression specifications presented in the papers\textsuperscript{65}.
 More detailed sets of reanalyses would be required to comprehensively challenge the claims from these specific papers.
-Nevertheless, these reanalyses do show, more broadly, that the problem of statistical non-independence applies to a wider range of national-level variables than those identified by previous work, such as parasite stress and democratic outcomes\textsuperscript{41}.
+Nevertheless, these reanalyses do show, more broadly, that the problem of statistical non-independence applies to a wider range of national-level variables than those identified by previous work, such as parasite stress and democratic outcomes\textsuperscript{43}.
 Moreover, given our finding that most studies in the current cross-national literature do not deal with non-independence at all, our reanalyses raise the worrying possibility that this literature is populated with spurious relationships.
@@
 On the contrary, such work promises to deepen understanding of our world, including the causes and consequences of economic development and cultural values.
-Moreover, cross-national studies allow social scientists to broaden their scope of study beyond Western populations\textsuperscript{7}, providing the representative samples necessary to test evolutionary and socio-ecological theories of human behaviour\textsuperscript{8,80}.
+Moreover, cross-national studies allow social scientists to broaden their scope of study beyond Western populations\textsuperscript{7}, providing the representative samples necessary to test evolutionary and socio-ecological theories of human behaviour\textsuperscript{8,82}.
 But in order to minimise spurious relationships in global datasets, we urge researchers to control for spatial and cultural phylogenetic non-independence when reporting cross-national correlations.
@@
 
-\hypertarget{geographic-and-cultural-phylogenetic-signal}{%
-\subsection{Geographic and cultural phylogenetic signal}\label{geographic-and-cultural-phylogenetic-signal}}
+\hypertarget{geographic-and-cultural-phylogenetic-signal-for-measures-of-economic-development-and-cultural-values}{%
+\subsection{Geographic and cultural phylogenetic signal for measures of economic development and cultural values}\label{geographic-and-cultural-phylogenetic-signal-for-measures-of-economic-development-and-cultural-values}}
 
 To estimate the degree of spatial and cultural phylogenetic non-independence in economic development and cultural values, we calculated geographic and cultural phylogenetic signal for global measures of development and values.
-For economic development variables, we retrieved longitudinal data on the Human Development Index\textsuperscript{45} (1990 - 2019; n = 189 nations), gross domestic product per capita (1960 - 2021; n = 209 nations), annual percentage growth in gross domestic product per capita (1961 - 2021; n = 208 nations), and the Gini coefficient of income inequality (1967 - 2021; n = 167 nations).
+For economic development variables, we retrieved longitudinal data on the Human Development Index\textsuperscript{47} (1990 - 2019; n = 189 nations), gross domestic product per capita (1960 - 2021; n = 209 nations), annual percentage growth in gross domestic product per capita (1961 - 2021; n = 208 nations), and the Gini coefficient of income inequality (1967 - 2021; n = 167 nations).
 Human development data were retrieved from the United Nations Development Programme (\url{https://hdr.undp.org/en/content/download-data}) and data for all other economic development variables were retrieved from the World Bank (\url{https://data.worldbank.org/}).
@@
 We downloaded the full Integrated Values Survey, which included all waves from the World Values Survey and the European Values Survey, and computed the two dimensions of cultural values following procedures from previous research\textsuperscript{16}.
-Additionally, we retrieved cross-sectional data on cultural tightness (n = 57 nations) and individualism (n = 97 nations) from previous work\textsuperscript{13,81}.
+Additionally, we retrieved cross-sectional data on cultural tightness (n = 57 nations) and individualism (n = 97 nations) from previous work\textsuperscript{13,83}.
 
 To calculate geographic and cultural phylogenetic signal, we created two proximity matrices for 269 of the world's nations: a geographic proximity matrix and a linguistic proximity matrix.
-Geographic proximity was converted from logged geodesic distances between nation capital cities.
-Linguistic proximity was calculated as the cultural proximity between all languages spoken within nations, weighted by speaker percentages (see Supplementary Methods).
+Geographic distance between two nations was calculated as the logged geodesic distance between nation capital cities (data from the maps R package\textsuperscript{84}) using the geosphere R package\textsuperscript{85}.
+The geographic proximity matrix was computed as one minus the log geographic distance matrix scaled between 0 and 1.
+Linguistic proximity between two nations was calculated as the cultural proximity between all languages spoken within those nations, weighted by speaker percentages.
+We acquired cultural proximity data by combining the language family trees provided by Glottolog v3.0\textsuperscript{86} into one global language tree (undated and unresolved).
+We calculated cultural proximity \(s\) between two languages \(j\) and \(k\) as the distance (in number of nodes traversed) of their most recent common ancestor \(i\) to the root of the tree, through the formula:
+
+\begin{gather}
+s_{jk} = \frac{n_{r}-n_{i}}{n_{r}}
+\end{gather}
+
+where \(n_{r}\) is the maximum path length (in number of nodes traversed) leading to the pan-human root \(r\), and \(n_{i}\) is the maximum path length leading to node \(i\).
+We then combined these proximities with speaker data from Ethnologue 21\textsuperscript{87} and compared every language spoken within those nations by at least 1 permille of the population, weighted by speaker percentages, through the formula:
+
+\begin{gather}
+w_{lm} = {\Sigma}{\Sigma}p_{lj}p_{mk}s_{jk}
+\end{gather}
+
+where \(p_{lj}\) is the percentage of the population in nation \(l\) speaking language \(j\), \(p_{mk}\) is the percentage of the population in nation \(m\) speaking language \(k\), and \(s_{jk}\) is the proximity measure between languages \(j\) and \(k\)\textsuperscript{88}.
+This calculation resulted in a linguistic proximity matrix with values between 0 and 1.
+
 We included these matrices in Bayesian multilevel models, allowing nation random intercepts to covary according to both geographic and linguistic proximity simultaneously.
-These models were fitted with the R package \emph{brms}\textsuperscript{79} and converged normally (\(\hat{R}\) \textless{} 1.1).
+These models were fitted with the R package brms\textsuperscript{81} and converged normally (\(\hat{R}\) \textless{} 1.1).
+The assumptions of these models were met: residuals were approximately normally distributed, though this was not formally tested.
 Estimates of geographic and cultural phylogenetic signal were computed as the proportion of national-level variance in these models explained by geographic and linguistic proximity matrices.
 
-\hypertarget{systematic-literature-review}{%
-\subsection{Systematic literature review}\label{systematic-literature-review}}
+\hypertarget{literature-review}{%
+\subsection{Literature review}\label{literature-review}}
 
@@
 We coded common attempts to control for non-independence between nations.
-These included: (1) any higher-level control variables for spatial regional groupings (e.g.~continent fixed effects); (2) any geographic distance control variables (e.g.~distance between capital cities, distance from equator, latitude); (3) any control variables capturing shared cultural history (e.g.~former colony, legal origin fixed effects, linguistic history, cultural influence); and (4) any other control variables, tests, or approaches that were deemed as attempts to control for non-independence (e.g.~eigenvector filtering\textsuperscript{82}, controls for trade-weightings between nations, cross-sectional dependence tests\textsuperscript{83}, separate analyses for subsets of nations).
+These included: (1) any higher-level control variables for spatial regional groupings (e.g.~continent fixed effects); (2) any geographic distance control variables (e.g.~distance between capital cities, distance from equator, latitude); (3) any control variables capturing shared cultural history (e.g.~former colony, legal origin fixed effects, linguistic history, cultural influence); and (4) any other control variables, tests, or approaches that were deemed as attempts to control for non-independence (e.g.~eigenvector filtering\textsuperscript{89}, controls for trade-weightings between nations, cross-sectional dependence tests\textsuperscript{90}, separate analyses for subsets of nations).
 These were coded by the first author.
@@
 We also calculated the proportion of articles employing the different types of control listed above at least once: regional fixed effects, distance, shared cultural history, or other.
-For these proportions, we calculated 95\% bootstrap confidence intervals with 1,000 bootstrap iterations.
+For these proportions, we calculated two-tailed 95\% bootstrap confidence intervals with 1,000 bootstrap iterations.
 Additionally, we predicted the probability of an article attempting to control for non-independence at least once using Bayesian logistic regression, including in separate models log journal impact factor and year of publication as linear and spline predictors, respectively.
@@
 We fitted these models separately for overall attempts to control for non-independence and split by method type.
-We report the adjusted proportions with 95\% credible intervals (see Results and Supplementary Figure \ref{fig:plotReview2}).
+We report the adjusted proportions with equal-tailed 95\% credible intervals (see Results and Supplementary Figure \ref{fig:plotReview2}).
 Additionally, we predicted the probability of an analysis attempting to control for non-independence using Bayesian multilevel logistic regression with random intercepts for articles.
 In separate models, we included log journal impact factor and year of publication as linear and spline predictors, respectively.
-All Bayesian models were fitted with the \emph{brms} R package\textsuperscript{79}.
+All Bayesian models were fitted with the brms R package\textsuperscript{81}.
 Our priors were informed by prior predictive checks, and all models converged normally (\(\hat{R}\) \textless{} 1.1).
+The assumptions of these models (i.e., binary nested data) were met.
 
@@
 Models employing Conley standard errors either required latitude and longitude values or coordinates for genetic distances.
-To determine distance cutoffs, we employed an approach recommended in previous work\textsuperscript{84}: we fitted models with a range of feasible distance cutoffs and retained the model with the largest standard error for the slope parameter.
-These models were fitted using the \emph{conleyreg} R package\textsuperscript{85}.
-Bayesian models were fitted using the \emph{brms} R package\textsuperscript{79}.
+To determine distance cutoffs, we employed an approach recommended in previous work\textsuperscript{91}: we fitted models with a range of feasible distance cutoffs and retained the model with the largest standard error for the slope parameter.
+These models were fitted using the conleyreg R package\textsuperscript{92}.
+Bayesian models were fitted using the brms R package\textsuperscript{81}.
 Our choice of priors was based on prior predictive simulation.
 All models converged normally (\(\hat{R}\) \textless{} 1.1).
-Across all model types and parameter combinations, we calculated the false positive rate as the proportion of models that estimated slopes with a 95\% confidence / credible interval excluding zero when \(r = 0\).
+Across all model types and parameter combinations, we calculated the false positive rate as the proportion of models that estimated slopes with a two-tailed 95\% confidence / credible interval excluding zero when \(r = 0\).
 We calculated statistical power as the proportion of models that estimated slopes with a 95\% confidence / credible interval excluding zero when \(r > 0\).
-We calculated 95\% bootstrap confidence intervals for these false positive rates and statistical power estimates with 1,000 bootstrap iterations.
+We calculated two-tailed 95\% bootstrap confidence intervals for these false positive rates and statistical power estimates with 1,000 bootstrap iterations.
 
@@
 
-We searched the individual analyses from our systematic review for statistically significant cross-national correlations with available primary or secondary data.
+We searched the individual analyses from our literature review for statistically significant cross-national correlations with available primary or secondary data.
 We restricted our search to one analysis per paper, and searched until we had a set of twelve analyses, six from economic development papers and six from cultural values papers, for which we were able to replicate the original result (i.e.~find a cross-national correlation with the same sign and roughly the same effect size).
@@
 
-The twelve analyses that we settled on\textsuperscript{13,14,16,53--61} were mostly bivariate cross-national correlations, except for two.
-One analysis\textsuperscript{14} additionally controlled for log gross national income, and another analysis\textsuperscript{58} is a multilevel model including random intercepts for nations and several individual-level and national-level covariates (see Model 5 in original paper).
-Before running any additional models, we pre-registered these twelve analyses on the Open Science Framework on 25\textsuperscript{th} January 2022 (\url{https://osf.io/uywx8/}).
+The twelve analyses that we settled on\textsuperscript{13,14,16,55--63} were mostly bivariate cross-national correlations, except for two.
+One analysis\textsuperscript{14} additionally controlled for log gross national income, and another analysis\textsuperscript{60} is a multilevel model including random intercepts for nations and several individual-level and national-level covariates (see Model 5 in original paper).
+Before running any additional models, we pre-registered these twelve analyses on the Open Science Framework on 25\textsuperscript{th} January 2022 (\url{https://osf.io/u8tbf}).
 We endeavoured to keep the sample sizes of our reanalyses as close to the original analyses as possible, though there were some deviations (see Supplementary Table \ref{tab:tableDeviations}).
@@
 
-We fitted these models using the \emph{brms} R package\textsuperscript{79}.
+We fitted these models using the brms R package\textsuperscript{81}.
 Our choice of priors was based on prior predictive simulation.
-All models converged normally (\(\hat{R}\) \textless{} 1.1), though for some models we resorted to using approximate Gaussian processes\textsuperscript{86} to reach convergence.
+All models converged normally (\(\hat{R}\) \textless{} 1.1), though for some models we resorted to using approximate Gaussian processes\textsuperscript{93} to reach convergence.
+The assumptions of these models were met: residuals were approximately normally distributed, though this was not formally tested.
 
@@
 
-All data and code are accessible at our Open Science Framework repository (\url{https://osf.io/uywx8/}).
-We used the \emph{targets} R package\textsuperscript{87} to create a reproducible data analysis pipeline and the \emph{papaja} R package\textsuperscript{88} to reproducibly generate the manuscript.
+All data and code are accessible on GitHub (\url{https://github.com/ScottClaessens/crossNationalCorrelations}).
+We used the targets R package\textsuperscript{94} to create a reproducible data analysis pipeline and the papaja R package\textsuperscript{95} to reproducibly generate the manuscript.
+
+\hypertarget{ethics}{%
+\subsection{Ethics}\label{ethics}}
+
+Ethical approval was not required for the study as we analysed only secondary and simulated data. We did not apply for ethical approval through an ethics board.
 
@@
 \nolinenumbers
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-This work was supported by a Royal Society of New Zealand Marsden grant (20-UOA123) to QDA.
-
-\hypertarget{author-contributions}{%
-\section{Author Contributions}\label{author-contributions}}
-
-SC and QDA conceived of and designed the study.
-SC curated the data, produced all code for analysis and visualisation, and wrote the original draft of the manuscript.
-TK and QDA developed and compiled the geographic and linguistic distance matrices.
-QDA provided funding and input on manuscript preparation and revision.
-All authors reviewed and edited the final draft of the manuscript.
-
-\hypertarget{competing-interests}{%
-\section{Competing Interests}\label{competing-interests}}
-
-The authors declare no competing interests.
 
@@
 
-All data can be found on the Open Science Framework: \url{https://osf.io/uywx8/}
+All data can be found on GitHub (\url{https://github.com/ScottClaessens/crossNationalCorrelations}).
+Data on human development were retrieved from the United Nations Development Programme (\url{https://hdr.undp.org/en/content/download-data}).
+Data on GDP per capita, annual GDP per capita growth, and the Gini coefficient were retrieved from the World Bank (\url{https://data.worldbank.org/}).
+Data on traditional vs.~secular values and survival vs.~self-expression values were retrieved from the World Values Survey (\url{https://www.worldvaluessurvey.org/wvs.jsp}).
+Data on cultural tightness were retrieved from the OSF repository for Gelfand et al.~2021 (\url{https://osf.io/47pe8/}).
+The review data generated in this study are provided in the Supplementary Data file.
+All other datasets (e.g., for replications) were retrieved from tables and supplementary tables directly from papers cited in the main text\textsuperscript{13,14,16,55--63} - these datasets can be found in our GitHub repository.
 
@@
 
-All code to reproduce the statistical analyses in this manuscript can be found on the Open Science Framework: \url{https://osf.io/uywx8/}
-
-\hypertarget{ethical-approval}{%
-\section{Ethical Approval}\label{ethical-approval}}
-
-Ethical approval was not required for the study as we analysed only secondary and simulated data.
+All code to reproduce the statistical analyses in this manuscript can be found on GitHub (\url{https://github.com/ScottClaessens/crossNationalCorrelations}).
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Murdock1969}{}}%
+\CSLLeftMargin{28. }%
+\CSLRightInline{Murdock, G. P. \& White, D. R. \href{https://doi.org/10.2307/3772907}{Standard cross-cultural sample}. \emph{Ethnology} \textbf{8}, 329--369 (1969).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Gray2009}{}}%
+\CSLLeftMargin{29. }%
+\CSLRightInline{Gray, R.
+D., Drummond, A.
+J.
+\& Greenhill, S.
+J.
+\href{https://doi.org/10.1126/science.1166858}{Language phylogenies reveal expansion pulses and pauses in {Pacific} settlement}.
+\emph{Science} \textbf{323}, 479--483 (2009).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Green1999}{}}%
+\CSLLeftMargin{30. }%
+\CSLRightInline{Green, R.
+\& Pawley, A.
+Early {Oceanic} architectural forms and settlement patterns: Linguistic, archaeological and ethnological perspectives.
+in \emph{Archaeology and language {III}: Artefacts, languages and texts} (eds.
+Blench, R.
+\& Spriggs, M.) 31--89 (Routledge, 1999).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Jarque1987}{}}%
-\CSLLeftMargin{28. }%
+\CSLLeftMargin{31. }%
 \CSLRightInline{Jarque, C.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Legendre1993}{}}%
-\CSLLeftMargin{29. }%
+\CSLLeftMargin{32. }%
 \CSLRightInline{Legendre, P. \href{https://doi.org/10.2307/1939924}{Spatial autocorrelation: Trouble or new paradigm?} \emph{Ecology} \textbf{74}, 1659--1673 (1993).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Roberts2013}{}}%
-\CSLLeftMargin{30. }%
+\CSLLeftMargin{33. }%
 \CSLRightInline{Roberts, J., Seán AND Winters.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Lichstein2002}{}}%
-\CSLLeftMargin{31. }%
+\CSLLeftMargin{34. }%
 \CSLRightInline{Lichstein, J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Loftin1983}{}}%
-\CSLLeftMargin{32. }%
+\CSLLeftMargin{35. }%
 \CSLRightInline{Loftin, C.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Ferguson1997}{}}%
-\CSLLeftMargin{33. }%
+\CSLLeftMargin{36. }%
 \CSLRightInline{Ferguson, J. Anthropology and its evil twin. in (eds. Cooper, F. \& Packard, R.) 150--175 (University of California Press Berkeley, 1997).}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Murdock1969}{}}%
-\CSLLeftMargin{34. }%
-\CSLRightInline{Murdock, G. P. \& White, D. R. \href{https://doi.org/10.2307/3772907}{Standard cross-cultural sample}. \emph{Ethnology} \textbf{8}, 329--369 (1969).}
-
 \leavevmode\vadjust pre{\hypertarget{ref-Dow2008}{}}%
-\CSLLeftMargin{35. }%
+\CSLLeftMargin{37. }%
 \CSLRightInline{Dow, M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Eff2004}{}}%
-\CSLLeftMargin{36. }%
+\CSLLeftMargin{38. }%
 \CSLRightInline{Eff, E. A. Does {M}r. {G}alton still have a problem? Autocorrelation in the standard cross-cultural sample. \emph{World Cultures} \textbf{15}, 153--170 (2004).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Symonds2014}{}}%
-\CSLLeftMargin{37. }%
+\CSLLeftMargin{39. }%
 \CSLRightInline{Symonds, M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Watts2018}{}}%
-\CSLLeftMargin{38. }%
+\CSLLeftMargin{40. }%
 \CSLRightInline{Watts, J., Sheehan, O., Bulbulia, J., Gray, R.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Atkinson2016}{}}%
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{41. }%
 \CSLRightInline{Atkinson, Q.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Kelly2020}{}}%
-\CSLLeftMargin{40. }%
+\CSLLeftMargin{42. }%
 \CSLRightInline{Kelly, M. \href{http://ssrn.com/abstract=3688200}{Understanding persistence}. \emph{CEPR Discussion Paper No. DP15246} (2020).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Bromham2018}{}}%
-\CSLLeftMargin{41. }%
+\CSLLeftMargin{43. }%
 \CSLRightInline{Bromham, L., Hua, X., Cardillo, M., Schneemann, H.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Bromham2021}{}}%
-\CSLLeftMargin{42. }%
+\CSLLeftMargin{44. }%
 \CSLRightInline{Bromham, L., Skeels, A., Schneemann, H., Dinnage, R.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Currie2012}{}}%
-\CSLLeftMargin{43. }%
+\CSLLeftMargin{45. }%
 \CSLRightInline{Currie, T.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Passmore2022}{}}%
-\CSLLeftMargin{44. }%
+\CSLLeftMargin{46. }%
 \CSLRightInline{Passmore, S.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hdi}{}}%
-\CSLLeftMargin{45. }%
+\CSLLeftMargin{47. }%
 \CSLRightInline{\emph{Human development report}. \url{http://hdr.undp.org/en/composite/HDI}.}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Schulz2022}{}}%
-\CSLLeftMargin{46. }%
+\CSLLeftMargin{48. }%
 \CSLRightInline{Schulz, J. F. \href{https://doi.org/10.1093/ej/ueac027}{Kin networks and institutional development}. \emph{The Economic Journal} \textbf{132}, 2578--2613 (2022).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Conley1999}{}}%
-\CSLLeftMargin{47. }%
+\CSLLeftMargin{49. }%
 \CSLRightInline{Conley, T. G. \href{https://doi.org/10.1016/S0304-4076(98)00084-0}{{GMM} estimation with cross sectional dependence}. \emph{Journal of Econometrics} \textbf{92}, 1--45 (1999).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Conley2010}{}}%
-\CSLLeftMargin{48. }%
+\CSLLeftMargin{50. }%
 \CSLRightInline{Conley, T. G. Spatial econometrics. in \emph{Microeconometrics} (eds. Durlauf, S. N. \& Blume, L. E.) 303--313 (Palgrave Macmillan, 2010).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Currie2009}{}}%
-\CSLLeftMargin{49. }%
+\CSLLeftMargin{51. }%
 \CSLRightInline{Currie, T.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-McElreath2020}{}}%
-\CSLLeftMargin{50. }%
+\CSLLeftMargin{52. }%
 \CSLRightInline{McElreath, R. \emph{Statistical rethinking: A {B}ayesian course with examples in {R} and {Stan}}. (CRC Press, 2020).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Neal1998}{}}%
-\CSLLeftMargin{51. }%
+\CSLLeftMargin{53. }%
 \CSLRightInline{Neal, R.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-deVillemereuil2014}{}}%
-\CSLLeftMargin{52. }%
+\CSLLeftMargin{54. }%
 \CSLRightInline{Villemereuil, P.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Beck2003}{}}%
-\CSLLeftMargin{53. }%
+\CSLLeftMargin{55. }%
 \CSLRightInline{Beck, T., Demirgäç-Kunt, A.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Beck2005}{}}%
-\CSLLeftMargin{54. }%
+\CSLLeftMargin{56. }%
 \CSLRightInline{Beck, T., Demirgäç-Kunt, A. \& Levine, R. {SME}s, growth, and poverty: Cross-country evidence. \emph{Journal of Economic Growth} \textbf{10}, 199--229 (2005).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Bockstette2002}{}}%
-\CSLLeftMargin{55. }%
+\CSLLeftMargin{57. }%
 \CSLRightInline{Bockstette, V., Chanda, A. \& Putterman, L. States and markets: The advantage of an early start. \emph{Journal of Economic growth} \textbf{7}, 347--369 (2002).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Easterly2003}{}}%
-\CSLLeftMargin{56. }%
+\CSLLeftMargin{58. }%
 \CSLRightInline{Easterly, W.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Easterly2007}{}}%
-\CSLLeftMargin{57. }%
+\CSLLeftMargin{59. }%
 \CSLRightInline{Easterly, W.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Skidmore2002}{}}%
-\CSLLeftMargin{58. }%
+\CSLLeftMargin{60. }%
 \CSLRightInline{Skidmore, M. \& Toya, H. \href{https://doi.org/10.1093/ei/40.4.664}{Do natural disasters promote long-run growth?} \emph{Economic Inquiry} \textbf{40}, 664--687 (2002).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Adamczyk2009}{}}%
-\CSLLeftMargin{59. }%
+\CSLLeftMargin{61. }%
 \CSLRightInline{Adamczyk, A.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Alesina2013}{}}%
-\CSLLeftMargin{60. }%
+\CSLLeftMargin{62. }%
 \CSLRightInline{Alesina, A., Giuliano, P. \& Nunn, N. On the origins of gender roles: Women and the plough. \emph{Quarterly Journal of Economics} \textbf{128}, 469--530 (2013).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Knack1997}{}}%
-\CSLLeftMargin{61. }%
+\CSLLeftMargin{63. }%
 \CSLRightInline{Knack, S.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Sookias2018}{}}%
-\CSLLeftMargin{62. }%
+\CSLLeftMargin{64. }%
 \CSLRightInline{Sookias, R.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Voth2021}{}}%
-\CSLLeftMargin{63. }%
+\CSLLeftMargin{65. }%
 \CSLRightInline{Voth, H.-J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Smith2017}{}}%
-\CSLLeftMargin{64. }%
+\CSLLeftMargin{66. }%
 \CSLRightInline{Smith, M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Hastie2017}{}}%
-\CSLLeftMargin{65. }%
+\CSLLeftMargin{67. }%
 \CSLRightInline{Hastie, T. J. \& Tibshirani, R. J. \emph{Generalized additive models}. (Routledge, 2017).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Muthukrishna2020}{}}%
-\CSLLeftMargin{66. }%
+\CSLLeftMargin{68. }%
 \CSLRightInline{Muthukrishna, M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Spolaore2009}{}}%
-\CSLLeftMargin{67. }%
+\CSLLeftMargin{69. }%
 \CSLRightInline{Spolaore, E. \& Wacziarg, R. The diffusion of development. \emph{The Quarterly Journal of Economics} \textbf{124}, 469--529 (2009).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Kyritsis2022}{}}%
-\CSLLeftMargin{68. }%
+\CSLLeftMargin{70. }%
 \CSLRightInline{Kyritsis, T., Matthews, L.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Matthews2016}{}}%
-\CSLLeftMargin{69. }%
+\CSLLeftMargin{71. }%
 \CSLRightInline{Matthews, L.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Trouillot2016}{}}%
-\CSLLeftMargin{70. }%
+\CSLLeftMargin{72. }%
 \CSLRightInline{Trouillot, M. \emph{Global transformations: Anthropology and the modern world}. (Springer, 2016).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Wolf2010}{}}%
-\CSLLeftMargin{71. }%
+\CSLLeftMargin{73. }%
 \CSLRightInline{Wolf, E. R. Europe and the people without history. \emph{University of California Press} (2010).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Smaldino2016}{}}%
-\CSLLeftMargin{72. }%
+\CSLLeftMargin{74. }%
 \CSLRightInline{Smaldino, P. E. \& McElreath, R. \href{https://doi.org/10.1098/rsos.160384}{The natural selection of bad science}. \emph{Royal Society Open Science} \textbf{3}, 160384 (2016).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Liddell2018}{}}%
-\CSLLeftMargin{73. }%
+\CSLLeftMargin{75. }%
 \CSLRightInline{Liddell, T.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Nicholson2017}{}}%
-\CSLLeftMargin{74. }%
+\CSLLeftMargin{76. }%
 \CSLRightInline{Nicholson, J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Tipton2019}{}}%
-\CSLLeftMargin{75. }%
+\CSLLeftMargin{77. }%
 \CSLRightInline{Tipton, E., Pustejovsky, J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Chambers2021}{}}%
-\CSLLeftMargin{76. }%
+\CSLLeftMargin{78. }%
 \CSLRightInline{Chambers, C. D. \& Tzavella, L. The past, present and future of {Registered Reports}. \emph{Nature Human Behaviour} \textbf{6}, 29--42 (2021).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Deffner2022}{}}%
-\CSLLeftMargin{77. }%
+\CSLLeftMargin{79. }%
 \CSLRightInline{Deffner, D., Rohrer, J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-rstan}{}}%
-\CSLLeftMargin{78. }%
+\CSLLeftMargin{80. }%
 \CSLRightInline{Stan Development Team. \href{http://mc-stan.org/}{{RStan}: The {R} interface to {Stan}}. (2020).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Burkner2017}{}}%
-\CSLLeftMargin{79. }%
+\CSLLeftMargin{81. }%
 \CSLRightInline{Bürkner, P.-C.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Oishi2013}{}}%
-\CSLLeftMargin{80. }%
+\CSLLeftMargin{82. }%
 \CSLRightInline{Oishi, S. \href{https://doi.org/10.1146/annurev-psych-030413-152156}{Socioecological psychology}. \emph{Annual Review of Psychology} \textbf{65}, 581--609 (2014).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Gelfand2021}{}}%
-\CSLLeftMargin{81. }%
+\CSLLeftMargin{83. }%
 \CSLRightInline{Gelfand, M.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Brownrigg2018}{}}%
+\CSLLeftMargin{84. }%
+\CSLRightInline{Brownrigg, R. \emph{\href{https://CRAN.R-project.org/package=maps}{{maps}: Draw geographical maps}}. (2018).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Hijmans2019}{}}%
+\CSLLeftMargin{85. }%
+\CSLRightInline{Hijmans, R. J. \emph{\href{https://CRAN.R-project.org/package=geosphere}{{geosphere}: Spherical trigonometry}}. (2019).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Glottolog}{}}%
+\CSLLeftMargin{86. }%
+\CSLRightInline{Hammarström, H., Forkel, R., Haspelmath, M.
+\& Bank, S.
+\emph{Glottolog 3.0}.
+(Max Planck Institute for the Science of Human History, 2017).
+doi:\href{https://doi.org/10.5281/zenodo.4061162}{10.5281/zenodo.4061162}.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Ethnologue}{}}%
+\CSLLeftMargin{87. }%
+\CSLRightInline{\emph{Ethnologue: Languages of the world}. (SIL International, 2018).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Eff2008}{}}%
+\CSLLeftMargin{88. }%
+\CSLRightInline{Eff, E. A. Weight matrices for cultural proximity: Deriving weights from a language phylogeny. \emph{Structure and Dynamics} \textbf{3}, (2008).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Tiefelsdorf2007}{}}%
-\CSLLeftMargin{82. }%
+\CSLLeftMargin{89. }%
 \CSLRightInline{Tiefelsdorf, M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-DeHoyos2006}{}}%
-\CSLLeftMargin{83. }%
+\CSLLeftMargin{90. }%
 \CSLRightInline{De Hoyos, R.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Colella2019}{}}%
-\CSLLeftMargin{84. }%
+\CSLLeftMargin{91. }%
 \CSLRightInline{Colella, F., Lalive, R., Sakalli, S.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Duben2021}{}}%
-\CSLLeftMargin{85. }%
+\CSLLeftMargin{92. }%
 \CSLRightInline{Düben, C. \emph{\href{https://CRAN.R-project.org/package=conleyreg}{{conleyreg}: Estimations using {C}onley standard errors}}. (2021).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Solin2020}{}}%
-\CSLLeftMargin{86. }%
+\CSLLeftMargin{93. }%
 \CSLRightInline{Solin, A. \& Särkkä, S. Hilbert space methods for reduced-rank {G}aussian process regression. \emph{Statistics and Computing} \textbf{30}, 419--446 (2020).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
-\CSLLeftMargin{87. }%
+\CSLLeftMargin{94. }%
 \CSLRightInline{Landau, W.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Aust2020}{}}%
-\CSLLeftMargin{88. }%
+\CSLLeftMargin{95. }%
 \CSLRightInline{Aust, F. \& Barth, M. \emph{\href{https://github.com/crsh/papaja}{{papaja}: {Create} {APA} manuscripts with {R Markdown}}}. (2020).}
@@
 
-\hypertarget{appendix-appendix}{%
-\appendix}
-
-\renewcommand{\appendixname}{\bf{Supplementary Material}}
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+This work was supported by a Royal Society of New Zealand Marsden grant (20-UOA123) to QDA.
+
+\hypertarget{author-contributions-statement}{%
+\section{Author Contributions Statement}\label{author-contributions-statement}}
+
+SC and QDA conceived of and designed the study.
+SC curated the data, produced all code for analysis and visualisation, and wrote the original draft of the manuscript.
+TK and QDA developed and compiled the geographic and linguistic distance matrices.
+QDA provided funding and input on manuscript preparation and revision.
+All authors reviewed and edited the final draft of the manuscript.
+
+\hypertarget{competing-interests-statement}{%
+\section{Competing Interests Statement}\label{competing-interests-statement}}
+
+The authors declare no competing interests.
+
+\newpage
+\vspace*{60mm}
+
 \renewcommand{\figurename}{Supplementary Figure}
 \renewcommand{\tablename}{Supplementary Table}
-\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
-\renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
-\renewcommand{\theequation}{S\arabic{table}} \setcounter{equation}{0}
-\pagenumbering{gobble}
-
-\hypertarget{section}{%
-\section{}\label{section}}
+\renewcommand{\thefigure}{\arabic{figure}} \setcounter{figure}{0}
+\renewcommand{\thetable}{\arabic{table}} \setcounter{table}{0}
+\renewcommand{\theequation}{S\arabic{equation}} \setcounter{equation}{0}
+
+\hypertarget{supplementary-information}{%
+\section{\texorpdfstring{\textbf{Supplementary Information}}{Supplementary Information}}\label{supplementary-information}}
+
+\setcounter{page}{1}
+\centering
+
+\noindent \small Cross-national analyses require additional controls to account for the non-independence of nations \newline
+\hspace*{1cm} \small Scott Claessens\textsuperscript{1}, Thanos Kyritsis\textsuperscript{1}, \& Quentin D. Atkinson\textsuperscript{1,2} \newline
+
+\raggedright
+
+\noindent \footnotesize \textsuperscript{1} School of Psychology, University of Auckland, New Zealand \newline
+\noindent \footnotesize \textsuperscript{2} School of Anthropology and Museum Ethnography, University of Oxford, United Kingdom
+\normalsize
+\newpage
 
@@
 
-\hypertarget{calculating-global-geographic-and-linguistic-proximity-matrices}{%
-\subsubsection{Calculating global geographic and linguistic proximity matrices}\label{calculating-global-geographic-and-linguistic-proximity-matrices}}
-
-Geographic distance between two nations was calculated as the logged geodesic distance between nation capital cities (data from the R package \emph{maps}; Brownrigg, 2018) using the R package \emph{geosphere} (Hijmans, 2019).
-The geographic proximity matrix was computed as one minus the log geographic distance matrix scaled between 0 and 1.
-
-Linguistic proximity between two nations was calculated as the cultural proximity between all languages spoken within those nations, weighted by speaker percentages.
-We acquired cultural proximity data by combining the language family trees provided by Glottolog v3.0 (Hammarström et al., 2017) into one global language tree (undated and unresolved).
-We calculated cultural proximity \(s\) between two languages \(j\) and \(k\) as the distance (in number of nodes traversed) of their most recent common ancestor \(i\) to the root of the tree, through the formula:
-
-\[
-s_{jk} = \frac{n_{r}-n_{i}}{n_{r}}
-\]
-
-where \(n_{r}\) is the maximum path length (in number of nodes traversed) leading to the pan-human root \(r\), and \(n_{i}\) is the maximum path length leading to node \(i\).
-We then combined these proximities with speaker data from Ethnologue 21 (Ethnologue, 2018) and compared every language spoken within those nations by at least 1 permille of the population, weighted by speaker percentages, through the formula:
-
-\[
-w_{lm} = {\Sigma}{\Sigma}p_{lj}p_{mk}s_{jk}
-\]
-
-where \(p_{lj}\) is the percentage of the population in nation \(l\) speaking language \(j\), \(p_{mk}\) is the percentage of the population in nation \(m\) speaking language \(k\), and \(s_{jk}\) is the proximity measure between languages \(j\) and \(k\) (Eff, 2008).
-The resulting linguistic proximity matrix was also scaled between 0 and 1 before analysis.
-
 \hypertarget{bayesian-models-for-reanalysis}{%
@@
 
-\[
-\begin{aligned}
+\begin{align}
 Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
-\mu_{i} &= \alpha + \beta X_{i} \\
-\alpha &\sim \text{Normal}(0, 0.4) \\
-\beta &\sim \text{Normal}(0, 0.4) \\
-\sigma &\sim \text{Exponential}(5)
-\end{aligned}
-\]
+\mu_{i} &= \alpha + \beta X_{i} \nonumber \\
+\alpha &\sim \text{Normal}(0, 0.4) \nonumber \\
+\beta &\sim \text{Normal}(0, 0.4) \nonumber \\
+\sigma &\sim \text{Exponential}(5) \nonumber
+\end{align}
+
 The priors in this model were arrived at by prior predictive checks, with wider priors making predictions beyond the scale of standardised outcome variables and narrower priors being too informative.
 
-\newpage
-
 To control for spatial non-independence, we add a Gaussian process to this model and feed it a scaled geographic distance matrix \(D\) based on Euclidean distances between latitude and longitude coordinates.
-This distance matrix is computed internally by the R package \emph{brms} (Bürkner, 2017).
+This distance matrix is computed internally by the \emph{brms} R package\textsuperscript{1}.
 The Gaussian process uses an exponentiated quadratic covariance kernel, the only covariance kernel currently supported by \emph{brms}.
@@
 
-\[
-\begin{aligned}
+\begin{align}
 Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
-\mu_{i} &= \alpha + \kappa_{\text{NATION}[i]} + \beta X_{i} \\
+\mu_{i} &= \alpha + \kappa_{\text{NATION}[i]} + \beta X_{i} \nonumber \\
 \begin{pmatrix}
-\kappa_{1} \\ \kappa_{2} \\ \kappa_{3} \\ ... \\ \kappa_{n}
+\kappa_{1} \\ \kappa_{2} \\  ... \\ \kappa_{n}
 \end{pmatrix} &\sim \text{MVNormal}
@@
 \begin{pmatrix}
-0 \\ 0 \\ 0 \\ ... \\ 0
+0 \\ 0 \\ ... \\ 0
 \end{pmatrix},\textbf{K}
-\end{pmatrix}\\
-\textbf{K}_{ij} &= sdgp^2 \text{exp} \big (-D_{ij}^2 / (2 lscale^2) \big )\\
-\alpha &\sim \text{Normal}(0, 0.4) \\
-\beta &\sim \text{Normal}(0, 0.4) \\
-\sigma &\sim \text{Exponential}(5) \\
-sdgp &\sim \text{Exponential}(5) \\
-lscale &\sim \text{InverseGamma}(?,?)
-\end{aligned}
-\]
+\end{pmatrix} \nonumber \\
+\textbf{K}_{ij} &= sdgp^2 \text{exp} \big (-D_{ij}^2 / (2 lscale^2) \big ) \nonumber \\
+\alpha &\sim \text{Normal}(0, 0.4) \nonumber \\
+\beta &\sim \text{Normal}(0, 0.4) \nonumber \\
+\sigma &\sim \text{Exponential}(5) \nonumber \\
+sdgp &\sim \text{Exponential}(5) \nonumber \\
+lscale &\sim \text{InverseGamma}(?,?) \nonumber 
+\end{align}
+
 where \(n\) is the number of nations, and \(D^2_{ij}\) reflects the squared Euclidean distances between latitude and longitude coordinates for the \(i\)-th and \(j\)-th nations.
@@
 This is because the \emph{brms} package intelligently tunes the prior for this parameter based on the covariates of the Gaussian process (see \url{https://betanalpha.github.io/assets/case_studies/gp_part3/part3.html}).
-
-\newpage
 
@@
 
-\[
-\begin{aligned}
+\begin{align}
 Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
-\mu_{i} &= \alpha + z_{\text{NATION}[i]}\sigma_{\alpha}L + \beta X_{i} \\
-\alpha &\sim \text{Normal}(0, 0.4) \\
-\beta &\sim \text{Normal}(0, 0.4) \\
-z_{j} &\sim \text{Normal}(0, 1)\\
-\sigma_{\alpha} &\sim \text{Exponential}(5) \\
-\sigma &\sim \text{Exponential}(5)
-\end{aligned}
-\]
-\newpage
+\mu_{i} &= \alpha + z_{\text{NATION}[i]}\sigma_{\alpha}L + \beta X_{i} \nonumber \\
+\alpha &\sim \text{Normal}(0, 0.4) \nonumber \\
+\beta &\sim \text{Normal}(0, 0.4) \nonumber \\
+z_{j} &\sim \text{Normal}(0, 1) \nonumber \\
+\sigma_{\alpha} &\sim \text{Exponential}(5) \nonumber \\
+\sigma &\sim \text{Exponential}(5) \nonumber 
+\end{align}
 
@@
 
-\[
-\begin{aligned}
+\begin{align}
 Y_{i} &\sim \text{Normal}(\mu_{i},\sigma) \\
-\mu_{i} &= \alpha + \kappa_{\text{NATION}[i]} + z_{\text{NATION}[i]}\sigma_{\alpha}L + \beta X_{i} \\
+\mu_{i} &= \alpha + \kappa_{\text{NATION}[i]} + z_{\text{NATION}[i]}\sigma_{\alpha}L + \beta X_{i} \nonumber \\
 \begin{pmatrix}
-\kappa_{1} \\ \kappa_{2} \\ \kappa_{3} \\ ... \\ \kappa_{n}
+\kappa_{1} \\ \kappa_{2} \\ ... \\ \kappa_{n}
 \end{pmatrix} &\sim \text{MVNormal}
@@
 \begin{pmatrix}
-0 \\ 0 \\ 0 \\ ... \\ 0
+0 \\ 0 \\ ... \\ 0
 \end{pmatrix},\textbf{K}
-\end{pmatrix}\\
-\textbf{K}_{ij} &= sdgp^2 \text{exp} \big (-D_{ij}^2 / (2 lscale^2) \big )\\
-\alpha &\sim \text{Normal}(0, 0.4) \\
-\beta &\sim \text{Normal}(0, 0.4) \\
-z_{j} &\sim \text{Normal}(0, 1)\\
-\sigma_{\alpha} &\sim \text{Exponential}(5) \\
-\sigma &\sim \text{Exponential}(5) \\
-sdgp &\sim \text{Exponential}(5) \\
-lscale &\sim \text{InverseGamma}(?,?)
-\end{aligned}
-\]
-
-\newpage
-
-\hypertarget{supplementary-results}{%
-\subsection{Supplementary Results}\label{supplementary-results}}
-
-\hypertarget{further-exploration-of-twelve-reanalyses}{%
-\subsubsection{Further exploration of twelve reanalyses}\label{further-exploration-of-twelve-reanalyses}}
-
-To understand why some cross-national correlations were attenuated by controls for non-independence while others were robust, we further explored our fitted models for evidence of spatial and cultural autocorrelation.
-For each outcome variable, our Gaussian process models provided varying estimates of how quickly spatial autocorrelation declined with distance (Supplementary Figure \ref{fig:plotReplications3}).
-For example, in Skidmore and Toya\textsuperscript{58} gross domestic product growth was only moderately spatially autocorrelated at 1,000 km distance (posterior median spatial autocorrelation at 1,000 km = 0.42, 95\% CI {[}0.07 0.90{]}), whereas in Inglehart and Baker\textsuperscript{16} traditional values were strongly spatially autocorrelated at the same distance (posterior median spatial autocorrelation at 1,000 km = 0.96, 95\% CI {[}0.81 0.99{]}).
-We also found varying estimates of cultural phylogenetic signal (Supplementary Figure \ref{fig:plotReplications4}), with some outcome variables expressing low signal (e.g.~confidence in institutions\textsuperscript{61}; posterior median = 0.07, 95\% CI {[}0.00 0.53{]}) and others expressing high signal (e.g.~female labour force participation\textsuperscript{60}; posterior median = 0.89, 95\% CI {[}0.63 0.98{]}).
-Exploratory regressions provided suggestive evidence that stronger estimates of spatial autocorrelation or cultural phylogenetic signal resulted in a more pronounced reduction in the effect size when controlling for non-independence between nations (Supplementary Figure \ref{fig:plotReplications5}).
-However, these negative slopes for spatial autocorrelation (b = -0.19, 95\% CI {[}-1.33 1.04{]}) and cultural phylogenetic signal (b = -0.28, 95\% CI {[}-1.26 0.70{]}) were both very uncertain, due to the small number of analyses and the posterior uncertainty in effect sizes and estimates of non-independence.
+\end{pmatrix} \nonumber \\
+\textbf{K}_{ij} &= sdgp^2 \text{exp} \big (-D_{ij}^2 / (2 lscale^2) \big ) \nonumber \\
+\alpha &\sim \text{Normal}(0, 0.4) \nonumber \\
+\beta &\sim \text{Normal}(0, 0.4) \nonumber \\
+z_{j} &\sim \text{Normal}(0, 1) \nonumber \\
+\sigma_{\alpha} &\sim \text{Exponential}(5) \nonumber \\
+\sigma &\sim \text{Exponential}(5) \nonumber \\
+sdgp &\sim \text{Exponential}(5) \nonumber \\
+lscale &\sim \text{InverseGamma}(?,?) \nonumber 
+\end{align}
 
@@
 \begin{figure}[H]
-\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotReview2-1} \caption{\emph{Analysis-level results from systematic review of 100 highly-cited cross-national studies of economic development (red) and cultural values (blue).} (a) Proportion of analyses accounting for non-independence, overall and split by common methods of controlling for non-independence.
+\includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotReview2-1} \caption{\emph{Analysis-level results from literature review of 100 highly-cited cross-national studies of economic development (red) and cultural values (blue).} (a) Proportion of analyses accounting for non-independence, overall and split by common methods of controlling for non-independence.
+Points represent posterior median proportions and ranges represent equal-tailed 95\% credible intervals (n = 4308 observations).
 (b) The association between journal impact factor and the probability that an analysis accounts for non-independence.
 (c) Estimated trend over time for the probability that an analysis accounts for non-independence.
-Point ranges represent estimated proportions and 95\% credible intervals.
-Lines and shaded areas are posterior median regression lines and 50\% credible intervals from Bayesian multilevel models.
+Lines and shaded areas are posterior median regression lines and equal-tailed 50\% credible intervals from Bayesian multilevel models (n = 4308 observations).
 Histograms represent relative counts for individual analyses that did (top) or did not (bottom) account for non-independence.
@@
 \includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotSim1-1} \caption{\emph{Distribution of cross-national correlations from simulation study under strong spatial autocorrelation.} In these simulations, the strength of spatial autocorrelation is set to 0.8 for both outcome and predictor variables, and the true correlation is set to 0.
-For frequentist regression models, point ranges represent correlation estimates and 95\% confidence intervals.
-For Bayesian regression models, point ranges represent posterior means and 95\% credible intervals.
+For frequentist regression models, points represent correlation estimates and ranges represent two-tailed 95\% confidence intervals (n = 236 observations).
+For Bayesian regression models, points represent posterior means and ranges represent equal-tailed 95\% credible intervals (n = 236 observations).
 Correlations are ordered by effect size independently in each panel.
 Red point ranges indicate that the 95\% confidence / credible interval excludes zero.
-Black point ranges indicate that the 95\% confidence / credible interval includes zero.}\label{fig:plotSim1}
+Black point ranges indicate that the 95\% confidence / credible interval includes zero.
+SEs = standard errors.}\label{fig:plotSim1}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotSim2-1} \caption{\emph{Distribution of cross-national correlations from simulation study under strong cultural phylogenetic autocorrelation.} In these simulations, the strength of cultural phylogenetic autocorrelation is set to 0.8 for both outcome and predictor variables, and the true correlation is set to 0.
-For frequentist regression models, point ranges represent correlation estimates and 95\% confidence intervals.
-For Bayesian regression models, point ranges represent posterior means and 95\% credible intervals.
+For frequentist regression models, points represent correlation estimates and ranges represent two-tailed 95\% confidence intervals (n = 236 observations).
+For Bayesian regression models, points represent posterior means and ranges represent equal-tailed 95\% credible intervals (n = 236 observations).
 Correlations are ordered by effect size independently in each panel.
 Red point ranges indicate that the 95\% confidence / credible interval excludes zero.
-Black point ranges indicate that the 95\% confidence / credible interval includes zero.}\label{fig:plotSim2}
+Black point ranges indicate that the 95\% confidence / credible interval includes zero.
+SEs = standard errors.}\label{fig:plotSim2}
 \end{figure}
@@
 We simulated 100 datasets per parameter combination assuming a small true correlation between variables (\emph{r} = 0.1) and fitted different models to each dataset.
-Statistical power was operationalised as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero.
-Point ranges represent proportions and 95\% bootstrap confidence intervals, and dashed lines indicate 80\% power.
+Statistical power was operationalised as the proportion of models that estimated a slope with a two-tailed 95\% confidence / credible interval excluding zero.
+Points represent raw proportions of models with slope 95\% confidence / credible intervals excluding zero, ranges represent two-tailed 95\% bootstrap confidence intervals (n = 1000 bootstrap samples), and dashed lines indicate 80\% power.
+Colours indicate whether the strength of autocorrelation for the predictor variable is 0.2 (red), 0.5, (green) or 0.8 (blue).
 SEs = standard errors.}\label{fig:plotPowerGeo01}
@@
 We simulated 100 datasets per parameter combination assuming a medium true correlation between variables (\emph{r} = 0.3) and fitted different models to each dataset.
-Statistical power was operationalised as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero.
-Point ranges represent proportions and 95\% bootstrap confidence intervals, and dashed lines indicate 80\% power.
+Statistical power was operationalised as the proportion of models that estimated a slope with a two-tailed 95\% confidence / credible interval excluding zero.
+Points represent raw proportions of models with slope 95\% confidence / credible intervals excluding zero, ranges represent two-tailed 95\% bootstrap confidence intervals (n = 1000 bootstrap samples), and dashed lines indicate 80\% power.
+Colours indicate whether the strength of autocorrelation for the predictor variable is 0.2 (red), 0.5, (green) or 0.8 (blue).
 SEs = standard errors.}\label{fig:plotPowerGeo03}
@@
 We simulated 100 datasets per parameter combination assuming a large true correlation between variables (\emph{r} = 0.5) and fitted different models to each dataset.
-Statistical power was operationalised as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero.
-Point ranges represent proportions and 95\% bootstrap confidence intervals, and dashed lines indicate 80\% power.
+Statistical power was operationalised as the proportion of models that estimated a slope with a two-tailed 95\% confidence / credible interval excluding zero.
+Points represent raw proportions of models with slope 95\% confidence / credible intervals excluding zero, ranges represent two-tailed 95\% bootstrap confidence intervals (n = 1000 bootstrap samples), and dashed lines indicate 80\% power.
+Colours indicate whether the strength of autocorrelation for the predictor variable is 0.2 (red), 0.5, (green) or 0.8 (blue).
 SEs = standard errors.}\label{fig:plotPowerGeo05}
@@
 We simulated 100 datasets per parameter combination assuming a small true correlation between variables (\emph{r} = 0.1) and fitted different models to each dataset.
-Statistical power was operationalised as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero.
-Point ranges represent proportions and 95\% bootstrap confidence intervals, and dashed lines indicate 80\% power.
+Statistical power was operationalised as the proportion of models that estimated a slope with a two-tailed 95\% confidence / credible interval excluding zero.
+Points represent raw proportions of models with slope 95\% confidence / credible intervals excluding zero, ranges represent two-tailed 95\% bootstrap confidence intervals (n = 1000 bootstrap samples), and dashed lines indicate 80\% power.
+Colours indicate whether the strength of autocorrelation for the predictor variable is 0.2 (red), 0.5, (green) or 0.8 (blue).
 SEs = standard errors.}\label{fig:plotPowerLin01}
@@
 We simulated 100 datasets per parameter combination assuming a medium true correlation between variables (\emph{r} = 0.3) and fitted different models to each dataset.
-Statistical power was operationalised as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero.
-Point ranges represent proportions and 95\% bootstrap confidence intervals, and dashed lines indicate 80\% power.
+Statistical power was operationalised as the proportion of models that estimated a slope with a two-tailed 95\% confidence / credible interval excluding zero.
+Points represent raw proportions of models with slope 95\% confidence / credible intervals excluding zero, ranges represent two-tailed 95\% bootstrap confidence intervals (n = 1000 bootstrap samples), and dashed lines indicate 80\% power.
+Colours indicate whether the strength of autocorrelation for the predictor variable is 0.2 (red), 0.5, (green) or 0.8 (blue).
 SEs = standard errors.}\label{fig:plotPowerLin03}
@@
 We simulated 100 datasets per parameter combination assuming a large true correlation between variables (\emph{r} = 0.5) and fitted different models to each dataset.
-Statistical power was operationalised as the proportion of models that estimated a slope with a 95\% confidence / credible interval excluding zero.
-Point ranges represent proportions and 95\% bootstrap confidence intervals, and dashed lines indicate 80\% power.
+Statistical power was operationalised as the proportion of models that estimated a slope with a two-tailed 95\% confidence / credible interval excluding zero.
+Points represent raw proportions of models with slope 95\% confidence / credible intervals excluding zero, ranges represent two-tailed 95\% bootstrap confidence intervals (n = 1000 bootstrap samples), and dashed lines indicate 80\% power.
+Colours indicate whether the strength of autocorrelation for the predictor variable is 0.2 (red), 0.5, (green) or 0.8 (blue).
 SEs = standard errors.}\label{fig:plotPowerLin05}
@@
 In all reanalyses, outcome and predictor variables are standardised, making regression slopes comparable to Pearson's correlation coefficients.
-Lines and shaded areas represent posterior median regression lines and 95\% credible intervals.
-For models with covariates (Adamzyck and Pitt 2009; Gelfand et al.~2011), marginal effects are presented holding all covariates at zero or their reference categories.}\label{fig:plotReplications2}
+Lines and shaded areas represent posterior median regression lines and equal-tailed 95\% credible intervals.
+Colours indicate predictions from models with no control (red), spatial control only (green), cultural control only (blue), or both spatial and cultural control (purple).
+For models with covariates (Adamzyck and Pitt 2009; Gelfand et al.~2011), marginal effects are presented holding all covariates at zero or their reference categories.
+GDP = gross domestic product; SME = small and medium-sized enterprise.}\label{fig:plotReplications2}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotReplications3-1} \caption{\emph{Posterior estimates of Gaussian process functions mapping spatial autocorrelation onto geographic distance from our reanalyses of economic development (a) and cultural values (b) studies.} The y-axis represents the amount of spatial autocorrelation between data points with increasing distance between those points on the x-axis (logged distance in kilometres).
-Lines and shaded areas represent median posterior spatial autocorrelation functions and 50\% and 95\% credible intervals.}\label{fig:plotReplications3}
+Lines and shaded areas represent median posterior spatial autocorrelation functions and equal-tailed 50\% and 95\% credible intervals.}\label{fig:plotReplications3}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{manuscript_files/figure-latex/plotReplications4-1} \caption{\emph{Posterior estimates of cultural phylogenetic signal from our reanalyses.} Cultural phylogenetic signal is operationalised as the proportion of national-level variance explained by linguistic proximity between nations.
-Ridges are full posterior distributions, and points are posterior medians, and lines represent 50\% and 95\% credible intervals.}\label{fig:plotReplications4}
+Ridges are full posterior distributions, points are posterior medians, and lines represent equal-tailed 50\% and 95\% credible intervals.
+Numbers of observations from the models are as follows, from top to bottom: n = 69, n = 45, n = 103, n = 98, n = 63, n = 89, n = 33, n = 75, n = 67, n = 28, n = 38, and n = 28.}\label{fig:plotReplications4}
 \end{figure}
@@
 In both panels, the y-axis represents the ratio of the effect size when controlling for spatial and cultural non-independence to the original effect size (from naive regression model), and the x-axis represents posterior median model estimates.
-Regression lines are plotted with 95\% credible intervals.}
+Regression lines are plotted with equal-tailed 95\% credible intervals.}
 \end{figure}
@@
 
-\begin{center}
-\begin{ThreePartTable}
-
-\footnotesize{
-
-\begin{longtable}{m{2cm}m{11cm}m{2cm}}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
-\caption{\label{tab:tableArticlesReview}List of 100 papers included in literature review, sorted by annual rate of citations since publication.}\\
-\toprule
-Review & \multicolumn{1}{c}{Reference} & \multicolumn{1}{c}{Citations per year}\\
-\midrule
-\endfirsthead
-\caption*{\normalfont{Table \ref{tab:tableArticlesReview} continued}}\\
-\toprule
-Review & \multicolumn{1}{c}{Reference} & \multicolumn{1}{c}{Citations per year}\\
-\midrule
-\endhead
-Cultural values & Knack, S., \& Keefer, P. (1997). Does social capital have an economic payoff? A cross-country investigation. The Quarterly journal of economics, 112(4), 1251-1288. & 121.50\\
- & Inglehart, R., \& Baker, W. E. (2000). Modernization, cultural change, and the persistence of traditional values. American sociological review, 19-51. & 119.33\\
- & Gelfand, M.
-J., Raver, J.
-L., Nishii, L., Leslie, L.
-M., Lun, J., Lim, B.
-C., ...
-\& Yamaguchi, S.
-(2011).
-Differences between tight and loose cultures: A 33-nation study. science, 332(6033), 1100-1104.
-& 91.33\\
- & Alesina, A., Giuliano, P., \& Nunn, N. (2013). On the origins of gender roles: Women and the plough. The quarterly journal of economics, 128(2), 469-530. & 57.12\\
- & Schmitt, D.
-P., Realo, A., Voracek, M., \& Allik, J.
-(2008).
-Why can't a man be more like a woman?
-Sex differences in Big Five personality traits across 55 cultures. Journal of personality and social psychology, 94(1), 168.
-& 49.15\\
- & Hofstede, G., \& McCrae, R. R. (2004). Personality and culture revisited: Linking traits and dimensions of culture. Cross-cultural research, 38(1), 52-88. & 35.53\\
- & Delhey, J., Newton, K., \& Welzel, C. (2011). How general is trust in “most people”? Solving the radius of trust problem. American Sociological Review, 76(5), 786-807. & 35.10\\
- & Fincher, C.
-L., Thornhill, R., Murray, D.
-R., \& Schaller, M.
-(2008).
-Pathogen prevalence predicts human cross-cultural variability in individualism/collectivism. Proceedings of the Royal Society B: Biological Sciences, 275(1640), 1279-1285.
-& 33.38\\
- & Schneider, S. L. (2008). Anti-immigrant attitudes in Europe: Outgroup size and perceived ethnic threat. European Sociological Review, 24(1), 53-67. & 30.54\\
- & Franzen, A., \& Meyer, R. (2010). Environmental attitudes in cross-national perspective: A multilevel analysis of the ISSP 1993 and 2000. European sociological review, 26(2), 219-234. & 30.27\\
- & Steenkamp, J.
-B.
-E., Ter Hofstede, F., \& Wedel, M.
-(1999).
-A cross-national investigation into the individual and national cultural antecedents of consumer innovativeness. Journal of marketing, 63(2), 55-69.
-& 29.32\\
- & Paxton, P. (2002). Social capital and democracy: An interdependent relationship. American sociological review, 254-277. & 29.21\\
- & Santos, H. C., Varnum, M. E., \& Grossmann, I. (2017). Global increases in individualism. Psychological science, 28(9), 1228-1239. & 28.25\\
- & Van Zanten, B.
-T., Van Berkel, D.
-B., Meentemeyer, R.
-K., Smith, J.
-W., Tieskens, K.
-F., \& Verburg, P.
-H.
-(2016).
-Continental-scale quantification of landscape values using social media data. Proceedings of the National Academy of Sciences, 113(46), 12974-12979.
-& 27.00\\
- & Beugelsdijk, S., \& Welzel, C. (2018). Dimensions and dynamics of national culture: Synthesizing Hofstede with Inglehart. Journal of cross-cultural psychology, 49(10), 1469-1505. & 24.00\\
- & Stephan, U., \& Uhlaner, L.
-M.
-(2010).
-Performance-based vs socially supportive culture: A cross-national study of descriptive norms and entrepreneurship. Journal of International Business Studies, 41(8), 1347-1364.
-& 23.91\\
- & Dakhli, M., \& De Clercq, D. (2004). Human capital, social capital, and innovation: a multi-country study. Entrepreneurship \& regional development, 16(2), 107-128. & 23.53\\
- & Adamczyk, A., \& Pitt, C. (2009). Shaping attitudes about homosexuality: The role of religion and cultural context. Social Science Research, 38(2), 338-351. & 21.75\\
- & Mayda, A. M., \& Rodrik, D. (2005). Why are some people (and countries) more protectionist than others?. European Economic Review, 49(6), 1393-1430. & 21.69\\
- & Johnson, T., Kulesa, P., Cho, Y.
-I., \& Shavitt, S.
-(2005).
-The relation between culture and response styles: Evidence from 19 countries. Journal of Cross-cultural psychology, 36(2), 264-277.
-& 20.62\\
- & Tam, K.
-P., \& Chan, H.
-W.
-(2018).
-Generalized trust narrows the gap between environmental concern and pro-environmental behavior: Multilevel evidence. Global Environmental Change, 48, 182-194.
-& 20.00\\
- & Kuppens, P., Realo, A., \& Diener, E.
-(2008).
-The role of positive and negative emotions in life satisfaction judgment across nations. Journal of personality and social psychology, 95(1), 66.
-& 19.62\\
- & Tam, K.
-P., \& Chan, H.
-W.
-(2017).
-Environmental concern has a weaker association with pro-environmental behavior in some societies than others: A cross-cultural psychology perspective. Journal of Environmental Psychology, 53, 213-223.
-& 18.75\\
- & Miska, C., Szőcs, I., \& Schiffinger, M. (2018). Culture’s effects on corporate sustainability practices: A multi-domain and multi-level view. Journal of World Business, 53(2), 263-279. & 17.67\\
- & Oishi, S., Diener, E., Lucas, R.
-E., \& Suh, E.
-M.
-(2009).
-Cross-cultural variations in predictors of life satisfaction: Perspectives from needs and values.
-In Culture and well-being (pp.
-109-127).
-Springer, Dordrecht.
-& 16.00\\
- & Malka, A., Soto, C.
-J., Inzlicht, M., \& Lelkes, Y.
-(2014).
-Do needs for security and certainty predict cultural and economic conservatism?
-A cross-national analysis. Journal of personality and social psychology, 106(6), 1031.
-& 16.00\\
- & Eom, K., Kim, H.
-S., Sherman, D.
-K., \& Ishii, K.
-(2016).
-Cultural variability in the link between environmental concern and support for environmental action. Psychological Science, 27(10), 1331-1339.
-& 15.80\\
- & Larsen, C. A. (2008). The institutional logic of welfare attitudes: How welfare regimes influence public support. Comparative political studies, 41(2), 145-168. & 15.62\\
- & Sortheix, F. M., \& Schwartz, S. H. (2017). Values that underlie and undermine well–being: Variability across countries. European Journal of Personality, 31(2), 187-201. & 15.50\\
- & Van der Meer, T., \& Hakhverdian, A.
-(2017).
-Political trust as the evaluation of process and performance: A cross-national study of 42 European countries. Political Studies, 65(1), 81-102.
-& 15.25\\
- & Resick, C.
-J., Hanges, P.
-J., Dickson, M.
-W., \& Mitchelson, J.
-K.
-(2006).
-A cross-cultural examination of the endorsement of ethical leadership. Journal of Business Ethics, 63(4), 345-359.
-& 15.20\\
- & Murray, D.
-R., \& Schaller, M.
-(2010).
-Historical prevalence of infectious diseases within 230 geopolitical regions: A tool for investigating origins of culture. Journal of Cross-Cultural Psychology, 41(1), 99-108.
-& 15.00\\
- & Welzel, C., Inglehart, R., \& Kligemann, H. D. (2003). The theory of human development: A cross‐cultural analysis. European Journal of Political Research, 42(3), 341-379. & 14.94\\
- & Löckenhoff, C.
-E., De Fruyt, F., Terracciano, A., McCrae, R.
-R., De Bolle, M., Costa, P.
-T., ...
-\& Yik, M.
-(2009).
-Perceptions of aging across 26 cultures and their culture-level associates. Psychology and aging, 24(4), 941.
-& 14.75\\
- & Fulmer, C.
-A., Gelfand, M.
-J., Kruglanski, A.
-W., Kim-Prieto, C., Diener, E., Pierro, A., \& Higgins, E.
-T.
-(2010).
-On “feeling right” in cultural contexts: How person-culture match affects self-esteem and subjective well-being. Psychological Science, 21(11), 1563-1569.
-& 14.64\\
- & Gifford, R., Scannell, L., Kormos, C., Smolova, L., Biel, A., Boncu, S., ...
-\& Uzzell, D.
-(2009).
-Temporal pessimism and spatial optimism in environmental assessments: An 18-nation study. Journal of environmental psychology, 29(1), 1-12.
-& 14.25\\
- & Tranter, B., \& Booth, K. (2015). Scepticism in a changing climate: A cross-national study. Global Environmental Change, 33, 154-164. & 14.17\\
- & Pisano, I., \& Lubell, M. (2017). Environmental behavior in cross-national perspective: A multilevel analysis of 30 countries. Environment and Behavior, 49(1), 31-58. & 13.75\\
- & Kesler, C., \& Bloemraad, I.
-(2010).
-Does immigration erode social capital?
-The conditional effects of immigration-generated diversity on trust, membership, and participation across 19 countries, 1981–2000. Canadian Journal of Political Science/Revue canadienne de science politique, 43(2), 319-347.
-& 13.73\\
- & Gelissen, J. (2007). Explaining popular support for environmental protection: A multilevel analysis of 50 nations. Environment and behavior, 39(3), 392-415. & 13.71\\
- & Thorisdottir, H., Jost, J.
-T., Liviatan, I., \& Shrout, P.
-E.
-(2007).
-Psychological needs and values underlying left-right political orientation: Cross-national evidence from Eastern and Western Europe. Public Opinion Quarterly, 71(2), 175-203.
-& 13.50\\
- & Hörisch, J., Kollat, J., \& Brieger, S.
-A.
-(2017).
-What influences environmental entrepreneurship?
-A multilevel analysis of the determinants of entrepreneurs’ environmental orientation. Small Business Economics, 48(1), 47-69.
-& 13.50\\
- & Fungáčová, Z., Hasan, I., \& Weill, L. (2019). Trust in banks. Journal of Economic Behavior \& Organization, 157, 452-476. & 13.50\\
- & Cuddy, A.
-J., Wolf, E.
-B., Glick, P., Crotty, S., Chong, J., \& Norton, M.
-I.
-(2015).
-Men as cultural ideals: Cultural values moderate gender stereotype content. Journal of personality and social psychology, 109(4), 622.
-& 13.17\\
- & Mata, R., Josef, A. K., \& Hertwig, R. (2016). Propensity for risk taking across the life span and around the globe. Psychological science, 27(2), 231-243. & 13.00\\
- & Venaik, S., \& Brewer, P. (2010). Avoiding uncertainty in Hofstede and GLOBE. Journal of international business studies, 41(8), 1294-1315. & 12.55\\
- & Schimmack, U., Oishi, S., \& Diener, E.
-(2005).
-Individualism: A valid and important dimension of cultural differences between nations. Personality and Social Psychology Review, 9(1), 17-31.
-& 12.31\\
- & Kvaløy, B., Finseraas, H., \& Listhaug, O. (2012). The publics’ concern for global warming: A cross-national study of 47 countries. Journal of Peace Research, 49(1), 11-22. & 12.00\\
- & Mikucka, M., Sarracino, F., \& Dubrow, J.
-K.
-(2017).
-When does economic growth improve life satisfaction?
-Multilevel analysis of the roles of social trust and income inequality in 46 countries, 1981–2012. World Development, 93, 447-459.
-& 12.00\\
- & Miyamoto, Y., Yoo, J., Levine, C.
-S., Park, J., Boylan, J.
-M., Sims, T., ...
-\& Ryff, C.
-D.
-(2018).
-Culture and social hierarchy: Self-and other-oriented correlates of socioeconomic status across cultures. Journal of Personality and Social Psychology, 115(3), 427.
-& 12.00\\
-Economic development & Inglehart, R., \& Baker, W. E. (2000). Modernization, cultural change, and the persistence of traditional values. American sociological review, 19-51. & 119.33\\
- & King, R. G., \& Levine, R. (1993). Finance and growth: Schumpeter might be right. The quarterly journal of economics, 108(3), 717-737. & 90.00\\
- & Treisman, D. (2000). The causes of corruption: a cross-national study. Journal of public economics, 76(3), 399-457. & 73.67\\
- & Selden, T.
-M., \& Song, D.
-(1994).
-Environmental quality and development: is there a Kuznets curve for air pollution emissions?. Journal of Environmental Economics and management, 27(2), 147-162.
-& 49.07\\
- & Benhabib, J., \& Spiegel, M. M. (1994). The role of human capital in economic development evidence from aggregate cross-country data. Journal of Monetary economics, 34(2), 143-173. & 47.89\\
- & Acemoglu, D., Johnson, S., Robinson, J. A., \& Yared, P. (2008). Income and democracy. American Economic Review, 98(3), 808-42. & 44.62\\
- & Easterly, W., \& Levine, R. (2003). Tropics, germs, and crops: how endowments influence economic development. Journal of monetary economics, 50(1), 3-39. & 36.10\\
- & Wimmer, A., Cederman, L. E., \& Min, B. (2009). Ethnic politics and armed conflict: A configurational analysis of a new global data set. American sociological review, 74(2), 316-337. & 34.42\\
- & Redding, S., \& Venables, A. J. (2004). Economic geography and international inequality. Journal of international Economics, 62(1), 53-82. & 32.65\\
- & Liobikienė, G., Mandravickaitė, J., \& Bernatonienė, J.
-(2016).
-Theory of planned behavior approach to understand the green purchasing behavior in the EU: A cross-cultural study. Ecological Economics, 125, 38-46.
-& 31.00\\
- & McCrae, R. R., \& Terracciano, A. (2005). Personality profiles of cultures: aggregate personality traits. Journal of personality and social psychology, 89(3), 407. & 26.12\\
- & Faccio, M. (2010). Differences between politically connected and nonconnected firms: A cross‐country analysis. Financial management, 39(3), 905-928. & 25.73\\
- & Ginarte, J. C., \& Park, W. G. (1997). Determinants of patent rights: A cross-national study. Research policy, 26(3), 283-301. & 25.08\\
- & Beugelsdijk, S., \& Welzel, C. (2018). Dimensions and dynamics of national culture: Synthesizing Hofstede with Inglehart. Journal of cross-cultural psychology, 49(10), 1469-1505. & 24.00\\
- & Jorgenson, A. K., \& Clark, B. (2012). Are the economy and the environment decoupling? A comparative international study, 1960–2005. American Journal of Sociology, 118(1), 1-44. & 23.00\\
- & Durlauf, S. N., \& Johnson, P. A. (1995). Multiple regimes and cross‐country growth behaviour. Journal of applied econometrics, 10(4), 365-384. & 22.88\\
- & Easterly, W. (2007). Inequality does cause underdevelopment: Insights from a new instrument. Journal of development economics, 84(2), 755-776. & 21.93\\
- & Fuwa, M. (2004). Macro-level gender inequality and the division of household labor in 22 countries. American sociological review, 69(6), 751-767. & 21.24\\
- & Smith, M.
-D., Rabbitt, M.
-P., \& Coleman-Jensen, A.
-(2017).
-Who are the world’s food insecure?
-New evidence from the Food and Agriculture Organization’s food insecurity experience scale. World Development, 93, 402-412.
-& 20.75\\
- & Beck, T., Demirgüç-Kunt, A., \& Levine, R. (2003). Law, endowments, and finance. Journal of financial Economics, 70(2), 137-181. & 20.28\\
- & Schreinemachers, P., \& Tipraqsa, P. (2012). Agricultural pesticides and land use intensification in high, middle and low income countries. Food policy, 37(6), 616-626. & 20.11\\
- & Poe, S.
-C., Tate, C.
-N., \& Keith, L.
-C.
-(1999).
-Repression of the human right to personal integrity revisited: A global cross-national study covering the years 1976–1993. International studies quarterly, 43(2), 291-313.
-& 19.36\\
- & Bloom, D. E., Sachs, J. D., Collier, P., \& Udry, C. (1998). Geography, demography, and economic growth in Africa. Brookings papers on economic activity, 1998(2), 207-295. & 18.61\\
- & Skidmore, M., \& Toya, H. (2002). Do natural disasters promote long‐run growth?. Economic inquiry, 40(4), 664-687. & 18.47\\
- & Kurtz, M. J., \& Schrank, A. (2007). Growth and governance: Models, measures, and mechanisms. The Journal of Politics, 69(2), 538-554. & 17.53\\
- & Klasen, S., \& Lamanna, F.
-(2009).
-The impact of gender inequality in education and employment on economic growth: new evidence for a panel of countries. Feminist economics, 15(3), 91-132.
-& 17.25\\
- & Welzel, C., Inglehart, R., \& Kligemann, H. D. (2003). The theory of human development: A cross‐cultural analysis. European Journal of Political Research, 42(3), 341-379. & 14.94\\
- & Brülhart, M., \& Sbergami, F. (2009). Agglomeration and growth: Cross-country evidence. Journal of Urban Economics, 65(1), 48-63. & 14.33\\
- & Beck, T., Demirguc-Kunt, A., \& Levine, R. (2005). SMEs, growth, and poverty: Cross-country evidence. Journal of economic growth, 10(3), 199-229. & 14.12\\
- & Ehrhardt‐Martinez, K., Crenshaw, E.
-M., \& Jenkins, J.
-C.
-(2002).
-Deforestation and the environmental Kuznets curve: A cross‐national investigation of intervening mechanisms. Social Science Quarterly, 83(1), 226-243.
-& 13.05\\
- & Rai, D., Zitko, P., Jones, K., Lynch, J., \& Araya, R.
-(2013).
-Country-and individual-level socioeconomic determinants of depression: multilevel cross-national comparison. The British Journal of Psychiatry, 202(3), 195-203.
-& 12.75\\
- & Holzner, M. (2011). Tourism and economic development: The beach disease?. Tourism Management, 32(4), 922-933. & 12.70\\
- & Calderón, C., Moral‐Benito, E., \& Servén, L. (2015). Is infrastructure capital productive? A dynamic heterogeneous approach. Journal of Applied Econometrics, 30(2), 177-198. & 12.67\\
- & Roskam, A.
-J.
-R., Kunst, A.
-E., Van Oyen, H., Demarest, S., Klumbiene, J., Regidor, E., ...
-\& additional participants to the study.
-(2010).
-Comparative appraisal of educational inequalities in overweight and obesity among adults in 19 European countries. International journal of epidemiology, 39(2), 392-404.
-& 12.64\\
- & Curtis, J. E., Baer, D. E., \& Grabb, E. G. (2001). Nations of joiners: Explaining voluntary association membership in democratic societies. American Sociological Review, 783-805. & 12.40\\
- & Adams, S., \& Klobodu, E. K. M. (2018). Financial development and environmental degradation: does political regime matter?. Journal of Cleaner Production, 197, 1472-1479. & 12.33\\
- & Qian, Y.
-(2007).
-Do national patent laws stimulate domestic innovation in a global patenting environment?
-A cross-country analysis of pharmaceutical patent protection, 1978–2002. The Review of Economics and Statistics, 89(3), 436-453.
-& 12.29\\
- & Gaddis, I., \& Klasen, S. (2014). Economic development, structural change, and women’s labor force participation. Journal of Population Economics, 27(3), 639-681. & 11.86\\
- & Bockstette, V., Chanda, A., \& Putterman, L. (2002). States and markets: The advantage of an early start. Journal of Economic growth, 7(4), 347-369. & 11.84\\
- & Levels, M., Dronkers, J., \& Kraaykamp, G.
-(2008).
-Immigrant children's educational achievement in western countries: origin, destination, and community effects on mathematical performance. American Sociological Review, 73(5), 835-853.
-& 11.62\\
- & Klasen, S.
-(2002).
-Low schooling for girls, slower growth for all?
-Cross‐country evidence on the effect of gender inequality in education on economic development. The World Bank Economic Review, 16(3), 345-373.
-& 11.42\\
- & Knudsen, K., \& Wærness, K. (2008). National context and spouses’ housework in 34 countries. European Sociological Review, 24(1), 97-113. & 11.15\\
- & Rigobon, R., \& Rodrik, D. (2005). Rule of law, democracy, openness, and income: Estimating the interrelationships. Economics of transition, 13(3), 533-564. & 11.00\\
- & Beladi, H., Chao, C. C., Ee, M. S., \& Hollas, D. (2019). Does medical tourism promote economic growth? A cross-country analysis. Journal of Travel Research, 58(1), 121-135. & 11.00\\
- & Roe, M.
-J., \& Siegel, J.
-I.
-(2011).
-Political instability: Effects on financial development, roots in the severity of economic inequality. Journal of Comparative Economics, 39(3), 279-309.
-& 10.90\\
- & Roberts, J. T., \& Grimes, P. E. (1997). Carbon intensity and economic development 1962–1991: A brief exploration of the environmental Kuznets curve. World development, 25(2), 191-198. & 10.71\\
- & Reinsberg, B., Kentikelenis, A., Stubbs, T., \& King, L.
-(2019).
-The world system and the hollowing out of state capacity: How structural adjustment programs affect bureaucratic quality in developing countries. American Journal of Sociology, 124(4), 1222-1257.
-& 10.50\\
- & Jones-Smith, J.
-C., Gordon-Larsen, P., Siddiqi, A., \& Popkin, B.
-M.
-(2011).
-Cross-national comparisons of time trends in overweight inequality by socioeconomic status among women using repeated cross-sectional surveys from 37 developing countries, 1989–2007. American journal of epidemiology, 173(6), 667-675.
-& 10.40\\
- & Demirgüç-Kunt, A., \& Levine, R. (1996). Stock market development and financial intermediaries: stylized facts. The World Bank Economic Review, 10(2), 291-321. & 10.36\\
- & Sethi, D., \& Acharya, D. (2018). Financial inclusion and economic growth linkage: Some cross country evidence. Journal of Financial Economic Policy. & 10.33\\
-\bottomrule
-\end{longtable}
-
-}
-
-\end{ThreePartTable}
-\end{center}
-
-\newpage
-
 \begin{lltable}
@@
 \caption{\label{tab:tableSim1}False positive rates in simulation with varying levels of spatial autocorrelation for the predictor variable (\(\rho\)) and for the outcome variable (\(\lambda\)), assuming that the true correlation is 0.
-Numbers represent the total number of analyses, out of 100, that estimated a slope with a 95\% confidence / credible interval excluding zero, falsely inferring a relationship when none is present.}\\
+Numbers represent the total number of analyses, out of 100, that estimated a slope with a two-tailed 95\% confidence / credible interval excluding zero, falsely inferring a relationship when none is present.}\\
 \toprule
@@
 \caption{\label{tab:tableSim2}False positive rates in simulation with varying levels of cultural phylogenetic autocorrelation for the predictor variable (\(\rho\)) and for the outcome variable (\(\lambda\)), assuming that the true correlation is 0.
-Numbers represent the total number of analyses, out of 100, that estimated a slope with a 95\% confidence / credible interval excluding zero, falsely inferring a relationship when none is present.}\\
+Numbers represent the total number of analyses, out of 100, that estimated a slope with a two-tailed 95\% confidence / credible interval excluding zero, falsely inferring a relationship when none is present.}\\
 \toprule
@@
 \caption{\label{tab:tableReplications}Estimated cross-national correlations from our reanalysis of twelve previous cross-national analyses.
-Numbers are posterior median slopes (equivalent to Pearson's r correlation coefficients) with 95\% credible intervals.}\\
+Numbers are posterior median slopes (equivalent to Pearson's r correlation coefficients) with equal-tailed 95\% credible intervals.}\\
 \toprule
@@
 
-\footnotesize{
-
-\begin{longtable}{m{4cm}m{2cm}m{2cm}m{7cm}}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\scriptsize{
+
+\begin{longtable}{m{2cm}m{2cm}m{2cm}m{9cm}}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
 \caption{\label{tab:tableDeviations}List of sample size deviations between original analyses and reanalyses.}\\
@@
 
-Brownrigg, R. \emph{maps: Draw geographical maps.} (2018).
-
-Bürkner, P.-C. brms: An R package for Bayesian multilevel models using Stan. \emph{Journal of Statistical Software} \textbf{80}, 1---28 (2017).
-
-Eff, E. A. Weight matrices for cultural proximity: Deriving weights from a language phylogeny. \emph{Structure and Dynamics} \textbf{3}, (2008).
-
-\emph{Ethnologue: Languages of the world.} (SIL International, 2018).
-
-Hammarström, H., Forkel, R., Haspelmath, M.
-\& Bank, S.
-\emph{Glottolog 3.0.} (Max Planck Institute for the Science of Human History, 2017).
-doi: \href{https://doi.org/10.5281/zenodo.4061162}{10.5281/zenodo.4061162}
-
-Hijmans, R. J. \emph{geosphere: Spherical trigonometry.} (2019).
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Bürkner, P.-C. brms: An R package for Bayesian multilevel models using Stan. \emph{Journal of Statistical Software} \textbf{80}, 1--28 (2017).
+\end{enumerate}
 

### 2023-07-10 R/functionsSimulation.R

@@ -334,69 +334,57 @@ plotSimInd <- function(olsModel1, olsModel2, olsModel3, olsModel4, olsModel5,
 plotSimAll <- function(olsModel1, olsModel2, olsModel3, olsModel4, olsModel5,
                        olsModel6, conleyModel1, conleyModel2, brmsModel1,
                        brmsModel2, brmsModel3, type, file) {
-  # plotting function
-  plotFun <- function(data, title, ylab, xlab) {
+  # data wrangling function
+  dataWrangle <- function(data, label) {
     data %>%
       filter(r == 0) %>%
       group_by(lambda, rho) %>%
-      summarise(prop = sum(sig) / n(), n = n()) %>%
+      summarise(prop = sum(sig) / n(), n = n(), .groups = "drop") %>%
       mutate(lower = map2(prop, n, function(x, y) getBootCIProp(x, y, 1000)[1]), 
              upper = map2(prop, n, function(x, y) getBootCIProp(x, y, 1000)[2])) %>%
       unnest(c(lower, upper)) %>%
-      ggplot(aes(x = lambda, y = prop, group = factor(rho), colour = factor(rho))) +
-      geom_hline(yintercept = 0.05, linetype = "dashed") +
-      geom_pointrange(aes(x = lambda, y = prop, ymin = lower, ymax = upper), 
-                      position = position_dodge(0.05), size = 0.15) +
-      geom_line(position = position_dodge(0.05)) +
-      scale_x_continuous(name = xlab, limits = c(0.1, 0.9), breaks = c(0.2, 0.5, 0.8)) +
-      scale_y_continuous(name = ylab, limits = c(0, 1), breaks = c(0, 0.25, 0.5, 0.75, 1)) +
-      guides(colour = guide_legend(title = "Strength of\nautocorrelation\nfor predictor\nvariable")) +
-      ggtitle(title) +
-      theme_classic()
+      mutate(label = label)
   }
-  # individual plots
-  pA <- plotFun(olsModel1, "No control", "", "")
-  pB <- plotFun(olsModel2, "Latitude", "", "")
-  pC <- plotFun(olsModel3, "Longitude", "", "")
-  pD <- plotFun(olsModel4, "Continent", "", "")
-  pE <- plotFun(olsModel5, "Language family", "False positive rate", "")
-  pF <- plotFun(olsModel6, "Mean 2000km radius", "", "")
[…]

### 2023-07-10 _targets.R

@@ -147,8 +147,6 @@ list(
   tar_target(plotReview1, plotReviewArticle(review, yearArticle, ifArticle)), 
   tar_target(plotReview2, plotReviewAnalysis(review, yearAnalysis, ifAnalysis, 
                                              postRM1, postRM2, postRM3, postRM4, postRM5)),
-  # table of articles in review
-  tar_target(tableArticlesReview, makeTableArticlesReview(review)),
   
   #### Simulation ####
   

### 2023-07-10 manuscript.tex
add zenodo doi citation

@@
 
-All data and code are accessible on GitHub (\url{https://github.com/ScottClaessens/crossNationalCorrelations}).
-We used the targets R package\textsuperscript{94} to create a reproducible data analysis pipeline and the papaja R package\textsuperscript{95} to reproducibly generate the manuscript.
+All data and code are accessible on GitHub (\url{https://github.com/ScottClaessens/crossNationalCorrelations})\textsuperscript{94}.
+We used the targets R package\textsuperscript{95} to create a reproducible data analysis pipeline and the papaja R package\textsuperscript{96} to reproducibly generate the manuscript.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-github}{}}%
+\CSLLeftMargin{94. }%
+\CSLRightInline{Claessens, S., Kyritsis, T.
+\& Atkinson, Q.
+D.
+Cross-national analyses require additional controls to account for the non-independence of nations - ScottClaessens/crossNationalCorrelations: v1.0.0.
+(2023) doi:\href{https://doi.org/10.5281/zenodo.8128702}{10.5281/zenodo.8128702}.}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
-\CSLLeftMargin{94. }%
+\CSLLeftMargin{95. }%
 \CSLRightInline{Landau, W.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Aust2020}{}}%
-\CSLLeftMargin{95. }%
+\CSLLeftMargin{96. }%
 \CSLRightInline{Aust, F. \& Barth, M. \emph{\href{https://github.com/crsh/papaja}{{papaja}: {Create} {APA} manuscripts with {R Markdown}}}. (2020).}

### 2023-08-08 manuscript.Rmd
add zenodo doi citation
final author changes

@@
   - name          : "Scott Claessens"
-    affiliation   : "1"
+    affiliation   : "1*"
   - name          : "Thanos Kyritsis"
@@
   - name          : "Quentin D. Atkinson"
-    affiliation   : "1,2"
+    affiliation   : "1,2*"
 
@@
   - id            : "2"
-    institution   : "\\footnotesize School of Anthropology and Museum Ethnography, University of Oxford, Oxford, United Kingdom"
-
-authornote: "Correspondence concerning this article should be addressed to Scott Claessens (scott.claessens@gmail.com) and Quentin D.
-Atkinson (q.atkinson@auckland.ac.nz), Floor 2, Building 302, 23 Symonds Street, Auckland, 1010, New Zealand.
-\\newline \\indent The authors made the following contributions.
-Scott Claessens: Conceptualization, Data curation, Formal analysis, Investigation, Software, Visualization, Writing - original draft, Writing - review & editing; Thanos Kyritsis: Data curation, Writing - review & editing; Quentin D.
-Atkinson: Conceptualization, Data curation, Funding acquisition, Supervision, Writing - review & editing."
+    institution   : "\\footnotesize School of Anthropology and Museum Ethnography, University of Oxford, Oxford, United Kingdom \\newline *email: scott.claessens@gmail.com; q.atkinson@auckland.ac.nz"
 
@@
 ```
+
+# Introduction
 
@@
 We removed articles that did not report original research, were not relevant to economic development or cultural values, or did not report at least one cross-national analysis.
-We then retained the 100 articles (50 for economic development, 50 for cultural values) with the highest annual rate of citations (see Supplementary Data).
+We then retained the 100 articles (50 for economic development, 50 for cultural values) with the highest annual rate of citations (see Supplementary Data 1).
 For each of these highly-cited articles, we exhaustively recorded every cross-national analysis reported in the main text (n = `r printnum(as.numeric(nrow(review)), digits = 0)`), identifying in each case whether or not the analysis attempted to control for spatial, cultural, or any other form of non-independence between nations (see Methods for detailed search criteria and coding decisions).
@@
 The second search was for the terms "values" AND ("cross-national" OR "cross-cultural" OR "cross-country"), which returned 6806 articles.
+As this was not a formal systematic literature review, we did not follow PRISMA[@Page2021] guidelines for systematic literature reviews.
+
 Once exported, we ordered the articles by descending number of citations per year since initial publication, using citation counts reported by Web of Science.
-
 We then coded each article, in order, for inclusion in our review.
@@
 
-All data and code are accessible on GitHub (https://github.com/ScottClaessens/crossNationalCorrelations).
+All data and code are accessible on GitHub[@github].
 We used the targets R package[@Landau2021] to create a reproducible data analysis pipeline and the papaja R package[@Aust2020] to reproducibly generate the manuscript.
@@
 
-Ethical approval was not required for the study as we analysed only secondary and simulated data. We did not apply for ethical approval through an ethics board.
+We did not apply for ethical approval through an ethics board as we analysed only publicly available and simulated data in this study.
 
@@
 
-All data can be found on GitHub (https://github.com/ScottClaessens/crossNationalCorrelations).
+All data to reproduce the statistical analyses in this manuscript can be found on GitHub[@github].
 Data on human development were retrieved from the United Nations Development Programme (https://hdr.undp.org/en/content/download-data).
@@
 self-expression values were retrieved from the World Values Survey (https://www.worldvaluessurvey.org/wvs.jsp).
-Data on cultural tightness were retrieved from the OSF repository for Gelfand et al.
-2021 (https://osf.io/47pe8/).
-The review data generated in this study are provided in the Supplementary Data file.
-All other datasets (e.g., for replications) were retrieved from tables and supplementary tables directly from papers cited in the main text[@Adamczyk2009;@Alesina2013;@Beck2003;@Beck2005;@Bockstette2002;@Easterly2003;@Easterly2007;@Fincher2008;@Gelfand2011;@Inglehart2000;@Knack1997;@Skidmore2002] - these datasets can be found in our GitHub repository.
+Data on cultural tightness were retrieved from the OSF repository for ref[@Gelfand2021] (https://osf.io/47pe8/).
+The review data generated in this study are provided in Supplementary Data 1.
+All other datasets (e.g., for replications) were retrieved from tables and supplementary tables directly from papers cited in the main text[@Adamczyk2009;@Alesina2013;@Beck2003;@Beck2005;@Bockstette2002;@Easterly2003;@Easterly2007;@Fincher2008;@Gelfand2011;@Inglehart2000;@Knack1997;@Skidmore2002] - these datasets can be found in our GitHub repository[@github].
 
@@
 
-All code to reproduce the statistical analyses in this manuscript can be found on GitHub (https://github.com/ScottClaessens/crossNationalCorrelations).
+All code to reproduce the statistical analyses in this manuscript can be found on GitHub[@github].
 

### 2023-08-08 manuscript.tex
final author changes

@@
   pdftitle={Cross-national analyses require additional controls to account for the non-independence of nations},
-  pdfauthor={Scott Claessens1, Thanos Kyritsis1, \& Quentin D. Atkinson1,2},
+  pdfauthor={Scott Claessens1*, Thanos Kyritsis1, \& Quentin D. Atkinson1,2*},
   pdflang={en-EN},
@@
 \title{Cross-national analyses require additional controls to account for the non-independence of nations}
-\author{Scott Claessens\textsuperscript{1}, Thanos Kyritsis\textsuperscript{1}, \& Quentin D. Atkinson\textsuperscript{1,2}}
+\author{Scott Claessens\textsuperscript{1*}, Thanos Kyritsis\textsuperscript{1}, \& Quentin D. Atkinson\textsuperscript{1,2*}}
 \date{}
@@
 
-\authornote{Correspondence concerning this article should be addressed to Scott Claessens (\href{mailto:scott.claessens@gmail.com}{\nolinkurl{scott.claessens@gmail.com}}) and Quentin D.
-Atkinson (\href{mailto:q.atkinson@auckland.ac.nz}{\nolinkurl{q.atkinson@auckland.ac.nz}}), Floor 2, Building 302, 23 Symonds Street, Auckland, 1010, New Zealand.
-\newline \indent The authors made the following contributions.
-Scott Claessens: Conceptualization, Data curation, Formal analysis, Investigation, Software, Visualization, Writing - original draft, Writing - review \& editing; Thanos Kyritsis: Data curation, Writing - review \& editing; Quentin D.
-Atkinson: Conceptualization, Data curation, Funding acquisition, Supervision, Writing - review \& editing.}
-
-\affiliation{\vspace{0.5cm}\textsuperscript{1} \footnotesize School of Psychology, University of Auckland, Auckland, New Zealand\\\textsuperscript{2} \footnotesize School of Anthropology and Museum Ethnography, University of Oxford, Oxford, United Kingdom}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} \footnotesize School of Psychology, University of Auckland, Auckland, New Zealand\\\textsuperscript{2} \footnotesize School of Anthropology and Museum Ethnography, University of Oxford, Oxford, United Kingdom \newline *email: \href{mailto:scott.claessens@gmail.com}{\nolinkurl{scott.claessens@gmail.com}}; \href{mailto:q.atkinson@auckland.ac.nz}{\nolinkurl{q.atkinson@auckland.ac.nz}}}
 
@@
 
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
 Nations are an important population unit structuring modern human life.
@@
 We removed articles that did not report original research, were not relevant to economic development or cultural values, or did not report at least one cross-national analysis.
-We then retained the 100 articles (50 for economic development, 50 for cultural values) with the highest annual rate of citations (see Supplementary Data).
+We then retained the 100 articles (50 for economic development, 50 for cultural values) with the highest annual rate of citations (see Supplementary Data 1).
 For each of these highly-cited articles, we exhaustively recorded every cross-national analysis reported in the main text (n = 4,308), identifying in each case whether or not the analysis attempted to control for spatial, cultural, or any other form of non-independence between nations (see Methods for detailed search criteria and coding decisions).
@@
 The second search was for the terms ``values'' AND (``cross-national'' OR ``cross-cultural'' OR ``cross-country''), which returned 6806 articles.
+As this was not a formal systematic literature review, we did not follow PRISMA\textsuperscript{89} guidelines for systematic literature reviews.
+
 Once exported, we ordered the articles by descending number of citations per year since initial publication, using citation counts reported by Web of Science.
-
 We then coded each article, in order, for inclusion in our review.
@@
 We coded common attempts to control for non-independence between nations.
-These included: (1) any higher-level control variables for spatial regional groupings (e.g.~continent fixed effects); (2) any geographic distance control variables (e.g.~distance between capital cities, distance from equator, latitude); (3) any control variables capturing shared cultural history (e.g.~former colony, legal origin fixed effects, linguistic history, cultural influence); and (4) any other control variables, tests, or approaches that were deemed as attempts to control for non-independence (e.g.~eigenvector filtering\textsuperscript{89}, controls for trade-weightings between nations, cross-sectional dependence tests\textsuperscript{90}, separate analyses for subsets of nations).
+These included: (1) any higher-level control variables for spatial regional groupings (e.g.~continent fixed effects); (2) any geographic distance control variables (e.g.~distance between capital cities, distance from equator, latitude); (3) any control variables capturing shared cultural history (e.g.~former colony, legal origin fixed effects, linguistic history, cultural influence); and (4) any other control variables, tests, or approaches that were deemed as attempts to control for non-independence (e.g.~eigenvector filtering\textsuperscript{90}, controls for trade-weightings between nations, cross-sectional dependence tests\textsuperscript{91}, separate analyses for subsets of nations).
 These were coded by the first author.
@@
 Models employing Conley standard errors either required latitude and longitude values or coordinates for genetic distances.
-To determine distance cutoffs, we employed an approach recommended in previous work\textsuperscript{91}: we fitted models with a range of feasible distance cutoffs and retained the model with the largest standard error for the slope parameter.
-These models were fitted using the conleyreg R package\textsuperscript{92}.
+To determine distance cutoffs, we employed an approach recommended in previous work\textsuperscript{92}: we fitted models with a range of feasible distance cutoffs and retained the model with the largest standard error for the slope parameter.
+These models were fitted using the conleyreg R package\textsuperscript{93}.
 Bayesian models were fitted using the brms R package\textsuperscript{81}.
@@
 Our choice of priors was based on prior predictive simulation.
-All models converged normally (\(\hat{R}\) \textless{} 1.1), though for some models we resorted to using approximate Gaussian processes\textsuperscript{93} to reach convergence.
+All models converged normally (\(\hat{R}\) \textless{} 1.1), though for some models we resorted to using approximate Gaussian processes\textsuperscript{94} to reach convergence.
 The assumptions of these models were met: residuals were approximately normally distributed, though this was not formally tested.
@@
 
-All data and code are accessible on GitHub (\url{https://github.com/ScottClaessens/crossNationalCorrelations})\textsuperscript{94}.
-We used the targets R package\textsuperscript{95} to create a reproducible data analysis pipeline and the papaja R package\textsuperscript{96} to reproducibly generate the manuscript.
+All data and code are accessible on GitHub\textsuperscript{95}.
+We used the targets R package\textsuperscript{96} to create a reproducible data analysis pipeline and the papaja R package\textsuperscript{97} to reproducibly generate the manuscript.
 
@@
 
-All data can be found on GitHub (\url{https://github.com/ScottClaessens/crossNationalCorrelations}).
+All data to reproduce the statistical analyses in this manuscript can be found on GitHub\textsuperscript{95}.
 Data on human development were retrieved from the United Nations Development Programme (\url{https://hdr.undp.org/en/content/download-data}).
@@
 Data on traditional vs.~secular values and survival vs.~self-expression values were retrieved from the World Values Survey (\url{https://www.worldvaluessurvey.org/wvs.jsp}).
-Data on cultural tightness were retrieved from the OSF repository for Gelfand et al.~2021 (\url{https://osf.io/47pe8/}).
-The review data generated in this study are provided in the Supplementary Data file.
-All other datasets (e.g., for replications) were retrieved from tables and supplementary tables directly from papers cited in the main text\textsuperscript{13,14,16,55--63} - these datasets can be found in our GitHub repository.
+Data on cultural tightness were retrieved from the OSF repository for ref\textsuperscript{83} (\url{https://osf.io/47pe8/}).
+The review data generated in this study are provided in Supplementary Data 1.
+All other datasets (e.g., for replications) were retrieved from tables and supplementary tables directly from papers cited in the main text\textsuperscript{13,14,16,55--63} - these datasets can be found in our GitHub repository\textsuperscript{95}.
 
@@
 
-All code to reproduce the statistical analyses in this manuscript can be found on GitHub (\url{https://github.com/ScottClaessens/crossNationalCorrelations}).
+All code to reproduce the statistical analyses in this manuscript can be found on GitHub\textsuperscript{95}.
 
@@
 \CSLLeftMargin{47. }%
-\CSLRightInline{\emph{Human development report}. \url{http://hdr.undp.org/en/composite/HDI}.}
+\CSLRightInline{United Nations Development Programme. \emph{Human development report}. \url{http://hdr.undp.org/en/composite/HDI} (2021).}
 
@@
 \CSLLeftMargin{87. }%
-\CSLRightInline{\emph{Ethnologue: Languages of the world}. (SIL International, 2018).}
+\CSLRightInline{Eberhard, D. M., Simons, G. F. \& Fennig, C. D. \emph{\href{https://www.ethnologue.com/}{Ethnologue: Languages of the world}}. (SIL International, 2018).}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Page2021}{}}%
+\CSLLeftMargin{89. }%
+\CSLRightInline{Page, M. J. \emph{et al.} \href{https://doi.org/10.1136/bmj.n71}{The {PRISMA} 2020 statement: An updated guideline for reporting systematic reviews}. \emph{BMJ} \textbf{372}, (2021).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Tiefelsdorf2007}{}}%
-\CSLLeftMargin{89. }%
+\CSLLeftMargin{90. }%
 \CSLRightInline{Tiefelsdorf, M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-DeHoyos2006}{}}%
-\CSLLeftMargin{90. }%
+\CSLLeftMargin{91. }%
 \CSLRightInline{De Hoyos, R.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Colella2019}{}}%
-\CSLLeftMargin{91. }%
+\CSLLeftMargin{92. }%
 \CSLRightInline{Colella, F., Lalive, R., Sakalli, S.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Duben2021}{}}%
-\CSLLeftMargin{92. }%
+\CSLLeftMargin{93. }%
 \CSLRightInline{Düben, C. \emph{\href{https://CRAN.R-project.org/package=conleyreg}{{conleyreg}: Estimations using {C}onley standard errors}}. (2021).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Solin2020}{}}%
-\CSLLeftMargin{93. }%
+\CSLLeftMargin{94. }%
 \CSLRightInline{Solin, A. \& Särkkä, S. Hilbert space methods for reduced-rank {G}aussian process regression. \emph{Statistics and Computing} \textbf{30}, 419--446 (2020).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-github}{}}%
-\CSLLeftMargin{94. }%
+\CSLLeftMargin{95. }%
 \CSLRightInline{Claessens, S., Kyritsis, T.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
-\CSLLeftMargin{95. }%
+\CSLLeftMargin{96. }%
 \CSLRightInline{Landau, W.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Aust2020}{}}%
-\CSLLeftMargin{96. }%
+\CSLLeftMargin{97. }%
 \CSLRightInline{Aust, F. \& Barth, M. \emph{\href{https://github.com/crsh/papaja}{{papaja}: {Create} {APA} manuscripts with {R Markdown}}}. (2020).}
